### 2011-04-15 paper/walnuts/paper.tex
Paper abstract

% $Id: thesis.tex,v 1.60 2010/04/13 11:12:30 dvd Exp $

\documentclass{article}
\usepackage{algpseudocode}
\usepackage[ruled]{algorithm}
\usepackage{url}
\usepackage{framed}
\usepackage{amsfonts,amsmath,amsthm}
\usepackage{graphicx}
\usepackage{url}
\usepackage{color}

\newcommand {\mean} {\ensuremath {\mathop{\mathrm{mean}}}}
\newcommand {\median} {\ensuremath {\mathop{\mathrm{median}}}}
\newcommand {\N} {\ensuremath {\mathcal{N}}}
\newcommand {\IE} {\ensuremath {\mathbb{E}}}
\newcommand {\cov} {\ensuremath {\mathop{\mathrm{cov}}}}
\newcommand {\BEL} {\ensuremath {\mathop{\mathrm{BEL}}}}

\newtheorem{lemma}{Lemma}

\title{Efficient Sampling Policies for the Walnut Merchant Problem}
\author {David Tolpin \and Solomon Eyal Shimony \\
Department of Computer Science, \\
Ben-Gurion University of the Negev, Beer Sheva, Israel \\
\{tolpin,shimony\}@cs.bgu.ac.il}

\begin{document}

\maketitle

\begin{abstract}

Policies for the Multiarmed Bandit (MB) proved successful for Monte Carlo
tree sampling (MCTS). However, MCTS differs from MB in that only the
final choice, and not all hand pulls, brings a reward. We define the
Walnut Merchant problem (WM) as a model for decision-making in MCTS,
propose efficient sampling policies, theoretirally analyze and
empirically evaluate their performance. The results suggest that
WM policies are better suited for MCTS than MB policies.

\end{abstract}

\section{Introduction}

\section{Background}

\section{Walnut Merchant Problem}

\section{Best Possible Reward in WM}

\section{Efficient Sampling Policies}

\section{Empirical Evaluation}

\section*{Acknowledgments}

The research is partially supported by Israel
Science Foundation grant 305/09, by the Lynne and William Frankel
Center for Computer Sciences, and by the Paul Ivanier Center for
Robotics Research and Production Management.

\bibliographystyle{named}
\bibliography{refs}

\end{document}

### 2011-05-28 code/figures.R
twolevel results

@@ -1,6 +1,6 @@
 ALGORITHMS <- c('rnd', 'ucb', 'sve')
 
-draw.regrets <- function(regrets_file) {
+draw.regrets <- function(regrets_file, legend.position='topright') {
 	regrets <- read.table(regrets_file, header=T)
 	xlim <- range(regrets$nsamples)
 	ylim <- range(0, regrets[-1])
@@ -11,7 +11,7 @@ draw.regrets <- function(regrets_file) {
 			  type="o", pch=i)
 		i <- i+1
 	}
-	legend(x='topright', legend=sapply(names(regrets)[-1], function(n) substring(n, 3)), pch=1:length(names(regrets)[-1]))
+	legend(x=legend.position, legend=sapply(names(regrets)[-1], function(n) substring(n, 3)), pch=1:length(names(regrets)[-1]))
 }
 
 draw.experiment <- function(prefix="") {

### 2011-05-29 code/figures.R
added illustration

@@ -36,10 +36,10 @@ compare.random <- function(upper_file='random-upper.txt', greedy_file='random-gr
   legend(x='topright', legend=c('upper', 'greedy', 'semigreedy'), pch=1:3)
 }
 
-draw.random <- function() {
+draw.random <- function(prefix='random', legend.position='topright') {
   layout(cbind(c(1,2),c(3,4)))
-  draw.regrets('random-8.txt')
-  draw.regrets('random-16.txt')
-  draw.regrets('random-32.txt')
-  draw.regrets('random-64.txt')
+  draw.regrets(paste(prefix,'8.txt',sep='-'), legend.position=legend.position)
+  draw.regrets(paste(prefix,'16.txt',sep='-'), legend.position=legend.position)
+  draw.regrets(paste(prefix,'32.txt',sep='-'), legend.position=legend.position)
+  draw.regrets(paste(prefix,'64.txt',sep='-'), legend.position=legend.position)
 }

### 2011-05-29 code/figures.R
better illustration

@@ -37,6 +37,7 @@ compare.random <- function(upper_file='random-upper.txt', greedy_file='random-gr
 }
 
 draw.random <- function(prefix='random', legend.position='topright') {
+  par(mar=c(4,4,2,2))
   layout(cbind(c(1,2),c(3,4)))
   draw.regrets(paste(prefix,'8.txt',sep='-'), legend.position=legend.position)
   draw.regrets(paste(prefix,'16.txt',sep='-'), legend.position=legend.position)

### 2011-05-29 code/figures.R
vararm experiment

@@ -1,10 +1,10 @@
 ALGORITHMS <- c('rnd', 'ucb', 'sve')
 
-draw.regrets <- function(regrets_file, legend.position='topright') {
+draw.regrets <- function(regrets_file, legend.position='topright', xlab=expresion(N[samples])) {
 	regrets <- read.table(regrets_file, header=T)
 	xlim <- range(regrets$nsamples)
 	ylim <- range(0, regrets[-1])
-	plot(x=regrets$nsamples, type='n', log="x", xlim=xlim, ylim=ylim, yaxs='i', xlab=expression(N[samples]), ylab='Regret', main=unlist(strsplit(regrets_file, "\\."))[[1]])
+	plot(x=regrets$nsamples, type='n', log="x", xlim=xlim, ylim=ylim, yaxs='i', xlab=xlab, ylab='Regret', main=unlist(strsplit(regrets_file, "\\."))[[1]])
 	i <- 1
 	for(alg in names(regrets)[-1]) {
 		lines(x=regrets$nsamples, y=regrets[[alg]],

### 2011-05-29 paper/baskets.tex
more figures

\documentclass{article}
\usepackage{tikz}
\usetikzlibrary{positioning}
\usetikzlibrary{shapes.geometric}
\usetikzlibrary{shapes.symbols}
\usetikzlibrary{shadows}
\usetikzlibrary{arrows}

\pagestyle{empty}

\begin{document}

\begin{tikzpicture}
	[basket/.style={cylinder, shape border rotate=90,aspect=0.25, minimum height=1cm,minimum width=1.5cm},
	poor/.style={basket, draw=gray, fill=gray!20!white},
	good/.style={basket, thick, draw=green!50!black, fill=green!20!white}]

\draw [->,thick,draw=,>=stealth] (-2,-0.75) node [above right,gray] {0} -- (6.5, -0.75) node [above left,green!80!black] {1} node [below left,black] {Quality};
\node [poor] at (0.6,0) {Poor};
\node [poor] at (0.2,-0.5) {Poor};
\node [poor] at (-0.2,-1) {Poor};
\node [poor] at (-0.6,-1.5) {Poor};
\node [good] at (4,-0.75) {Good};

\end{tikzpicture}
\end{document}

### 2011-05-30 paper/baskets-p-m-g.tex
added p-m-g

\documentclass{article}
\usepackage{dingbat}
\usepackage{tikz}
\usetikzlibrary{positioning}
\usetikzlibrary{shapes.geometric}
\usetikzlibrary{shapes.symbols}
\usetikzlibrary{shadows}
\usetikzlibrary{arrows}

\pagestyle{empty}

\begin{document}

\begin{tikzpicture}
	[basket/.style={cylinder, shape border rotate=90,aspect=0.25, minimum height=1.1cm,minimum width=1.75cm},
	poor/.style={basket, draw=gray, fill=gray!20!white},
	mediocre/.style={basket, thick, green!30!gray!20!black, draw=green!50!gray!50!black, fill=green!40!gray!20!white},
	good/.style={basket, very thick, green!30!black, draw=green!50!black, fill=green!20!white}]

\draw [->,thick,draw=,>=stealth] (-2,0) node [above right,gray] {0} -- (6.5, 0) node [above left,green!80!black] {1} node [below left,black] {Quality};
\node [poor] at (-0.5,0.0) {Poor};
\node [mediocre] at (2.0,-0.05) {Mediocre};
\node [good] at (4.0,0.0) {Good};

\end{tikzpicture}
\end{document}

### 2011-05-31 slides/slides.tex
some slides

\documentclass{beamer}
\usepackage[latin1]{inputenc}

\usepackage[noend]{algpseudocode}
\usepackage[ruled]{algorithm}
\usepackage{url}
\usepackage{framed}
\usepackage{amsfonts,amsmath,amsthm}
\usepackage{graphicx}
\usepackage{url}
\usepackage{color}

\title{Rational Monte Carlo Tree Search}
\author{David Tolpin, Solomon Eyal Shimony}
\institute{Ben-Gurion University of the Negev\\Beer Sheva, Israel}
\date{June 1, 2011}

\AtBeginSection[]
{
  \begin{frame}<beamer>
    \frametitle{Outline}
    \tableofcontents[currentsection]
  \end{frame}
}

\begin{document}

\begin{frame}
\titlepage
\end{frame}

\begin{frame}{UCT}
UCT (Upper Confidence Bounds on Trees) is popular for Monte
  Carlo search in large trees:
\begin{itemize}
\item Selects an action $a_i$ that maximizes $\overline r_i + C\sqrt
  {\frac {\log n} {n_i}}$
\item Chooses a non-optimal action $n_i\ge\rho\log n$ times ($\rho$ is
  some constant).
\item Based on UCB1, that achieves logarithmic regret for multi-armed
  bandits.
\item But {\bf{\Large no bandits in Monte Carlo:}} {\bf no reward} is given for {\bf
  sampling} a good action.
\end{itemize}
\end{frame}

\begin{frame}{Doing better than UCT on sets}
\begin{figure}[h]
\centering
\includegraphics[scale=0.8]{onelevel-tree.pdf}
\end{figure}
When an arm is selected based on the {\bf sample mean}:
\begin{itemize}
\item Regret of UCB decreases {\it polynomially} with $n$.
\item Regret of $\epsilon$-greedy decreases {\it exponentially} with
  $n$.
\item Regret of UVB: $\max V_i$, $V_{i_{best}}=\frac {1-1/k}
  {n_{i_{best}}},\quad V_{i_{other}}=\frac {1/k} {n_{i_{other}}}$
  decreases exponentially with $n$, faster than $\epsilon$-greedy.
\end{itemize}
\end{frame}

\begin{frame}{UCB vs. $\epsilon$-greedy vs UVB}
\begin{figure}[h]
\centering
\includegraphics[scale=0.45]{onelevel-64.pdf}

64 Bernoulli arms, randomly generated
\end{figure}
\end{frame}

\begin{frame}{Doing Better Than UCT on Trees}
Uniform sampling is useless in this tree:
\begin{figure}[h]
\centering
\includegraphics[scale=0.8]{twolevel-tree.pdf}
\end{figure}
Rational sampling:
\begin{itemize}
\item first, choose an action that maximizes VOI (UVB);
\item then, choose actions that maximize average reward (UCB).
\end{itemize}
\end{frame}{}

\begin{frame}{UVT vs. VCT (UVB+UCT) vs. UCT}
\begin{figure}[h]
\centering
\includegraphics[scale=0.45]{twolevel-64.pdf}

64 Bernoulli arms, randomly generated
\end{figure}
\end{frame}

\end{document}

### 2011-05-31 code/figures.R

@@ -1,6 +1,6 @@
 ALGORITHMS <- c('rnd', 'ucb', 'sve')
 
-draw.regrets <- function(regrets_file, legend.position='topright', xlab=expresion(N[samples])) {
+draw.regrets <- function(regrets_file, legend.position='topright', xlab=expression(N[samples])) {
 	regrets <- read.table(regrets_file, header=T)
 	xlim <- range(regrets$nsamples)
 	ylim <- range(0, regrets[-1])

### 2011-06-12 code/sailing/splitlog.py
sailing:model:fixed bad-leg-p, unit test added

@@ -4,7 +4,7 @@ def do(logf,template='costs-size=%s-nsamples=%s.txt'):
     resf = None
     size='X'
     for line in logf:
-        m = re.match('.*\(exp1 .* :size (\d+)\)\s*', line)
+        m = re.match('.*\(exp1 .* :size (\d+)\s*', line)
         if m:
             size = int(m.group(1))
             continue

### 2011-06-12 paper/rmcts/paper.tex

--- 2011-04-15 paper/walnuts/paper.tex
@@
-% $Id: thesis.tex,v 1.60 2010/04/13 11:12:30 dvd Exp $
-
 \documentclass{article}
@@
 
-\title{Efficient Sampling Policies for the Walnut Merchant Problem}
+\title{Doing Better Than UCT: Rational Monte Carlo Sampling in Trees}
 \author {David Tolpin \and Solomon Eyal Shimony \\
@@
 tree sampling (MCTS). However, MCTS differs from MB in that only the
-final choice, and not all hand pulls, brings a reward. We define the
-Walnut Merchant problem (WM) as a model for decision-making in MCTS,
-propose efficient sampling policies, theoretirally analyze and
-empirically evaluate their performance. The results suggest that
-WM policies are better suited for MCTS than MB policies.
+final choice, and not all hand pulls, brings a reward.
 
@@
 
-\section{Background}
+\section{Main Results}
 
-\section{Walnut Merchant Problem}
+\section{Doing Better Than UCB}
 
-\section{Best Possible Reward in WM}
+\section{Doing Better Than UCT}
 
-\section{Efficient Sampling Policies}
-
-\section{Empirical Evaluation}
+\section{Summary and Future Work}
 
@@
 
-\bibliographystyle{named}
+\bibliographystyle{plain}
 \bibliography{refs}

### 2011-06-21 code/flatexp/uvexp.py
scratcha dded

@@ -20,7 +20,7 @@ GRD = 'GRD'
 UVB = 'UVB'
 UVB2 = 'UVB2'
 
-ALGORITHMS = [UCB, GRD, UVB]
+ALGORITHMS = [RND, UCB, GRD, UVB]
 # ALGORITHMS = [RND, UCB, UCBt, GRD, UVB, UVB2]
 
 class Exp:

### 2011-06-21 code/flatexp/uvrexp.py

@@ -44,7 +44,7 @@ def randomexp(nhandles=10, nruns=1000, samples=range(1, 16)):
     def mkhandles():
         return [RandomHandle(transform(random.random())) for i in range(nhandles)]
     handles = mkhandles()
-    for nsamples in [nhandles*i for i in samples]:
+    for nsamples in [10*nhandles*i for i in samples]:
         print nsamples,
         for alg in ALGORITHMS:
             print repeat_alg(alg=alg, mkhandles=mkhandles, nsamples=nsamples, nruns=nruns)[1],

### 2011-06-21 paper/rmcts/paper.tex
paper preparations
eps-greedy regret
added forgotten makefile
paper in progress

@@
 \usepackage{framed}
-\usepackage{amsfonts,amsmath,amsthm}
+\usepackage{amsfonts,amsmath,amsthm,amssymb}
 \usepackage{graphicx}
@@
 tree sampling (MCTS). However, MCTS differs from MB in that only the
-final choice, and not all hand pulls, brings a reward.
+final choice, rather than all arm pulls, brings a reward.
 
@@
 \section{Introduction}
+
+\cite{Auer.ucb}
 
@@
 
+\pagebreak
+
+\appendix
+
+\section{Facts}
+
+{\bf Chernoff bound:} for $m$ independent random variables $X_1, X_2, ..., X_m$
+taking values 0 or 1, $X=\sum_{i=1}^m X_i$, and $0\le\delta\le 1$:
+
+\begin{equation}
+\label{eq:chernoff-bound}
+\Pr[X < (1-\delta)\IE[X]] < e^{-\delta^2\IE[X]/2}
+\end{equation}
+
+\section{Finite-time regret bounds}
+
+\subsection{$\varepsilon$-greedy}
+
+Regret $R_\varepsilon$:
+
+\begin{eqnarray}
+\IE[R_\varepsilon]&=&\sum_{j=1}^K\Delta_j\Pr(\overline X_j=\max_i\overline X_i)
+\end{eqnarray}
+
+For $0<\varepsilon\le1-\frac 1 K$ and $x_0>0$ (see Section 3, Proofs, in \cite{Auer.ucb}):
+
+\begin{eqnarray}
+\Pr(\overline X_j=\max_i\overline X_i)&\le&2\left(x_0\cdot \Pr\left\{T_j^R(n)\le x_0\right\} + \frac 2{\Delta_j^2}e^{-\Delta_j^2\lfloor x_0 \rfloor/2}\right)\nonumber\\
+&\le&2\left(x_0\cdot \Pr\left\{T_j^R(n)\le x_0\right\} + \frac {2 \cdot
+  e^{\Delta_j^2/2}}{\Delta_j^2}e^{-\Delta_j^2 x_0 /2}\right)
+\end{eqnarray}
+
+Number of times $\mu$ the $j$th arm is selected {\it randomly}:
+
+\begin{equation}
+\mu\triangleq\IE\left[T_j^R(n)\right]=\frac {n\varepsilon} K
+\end{equation}
+
+By choosing $x_0=\frac \mu 2$ and applying the Chernoff bound (\ref{eq:chernoff-bound}), obtain:
+
+\begin{eqnarray}
+\Pr(\overline X_j=\max_i\overline X_i)&\le& 2\left(\frac {\mu}{2} e^{-\mu/8} + \frac {2e^{\Delta_j^2/2}}{\Delta_j^2}e^{-\Delta_j^2 \mu/4}\right)\nonumber\\
+&\le&\left(\mu + \frac {4\sqrt e}{\Delta_j^2}\right)e^{-\Delta_j^2\mu/8}
+\end{eqnarray}
+
+\begin{eqnarray}
+\IE[R_\varepsilon]&\le&\sum_{j=1}^K\Delta_j\left(\mu + \frac {4\sqrt e}
+{\Delta_j^2}\right)e^{-\Delta_j^2\mu/8}\nonumber\\
+&\le&K\left(\mu + \frac {4\sqrt e}{\Delta_{\min}}\right)e^{-\Delta_{\min}^2\mu/8}
+\end{eqnarray}
+
+\subsection{UCB}
+
+Regret $R_{UCB}$:
+
+\subsection{UVB}
+
+Regret $R_{UVB}$:
+
+\section{Asymptotic bounds for $\varepsilon$-greedy policies}
+
+\section{Case study: $K-1$ and One Arms}
+
 \end{document}

### 2011-06-21 code/flatexp/uvrexp.py
merge

@@ -44,7 +44,7 @@ def randomexp(nhandles=10, nruns=1000, samples=range(1, 16)):
     def mkhandles():
         return [RandomHandle(transform(random.random())) for i in range(nhandles)]
     handles = mkhandles()
-    for nsamples in [10*nhandles*i for i in samples]:
+    for nsamples in [nhandles*i for i in samples]:
         print nsamples,
         for alg in ALGORITHMS:
             print repeat_alg(alg=alg, mkhandles=mkhandles, nsamples=nsamples, nruns=nruns)[1],

### 2011-06-21 paper/rmcts/paper.tex
paper in progress

@@
 
+$\hat{C}_f$
+
 \bibliographystyle{plain}
@@
 
-{\bf Chernoff bound:} for $m$ independent random variables $X_1, X_2, ..., X_m$
+{\bf Chernoff bound (see \cite{Hagerup.chernoff}):} for $m$ independent random variables $X_1, X_2, ..., X_m$
 taking values 0 or 1, $X=\sum_{i=1}^m X_i$, and $0\le\delta\le 1$:
@@
 \section{Finite-time regret bounds}
+
+For an analysis of pure exploration of uniform and UCB algorithms, see also
+\cite{Bubeck.pure}.
 

### 2011-06-21 paper/rmcts/paper.tex
paper in progress: figures added

@@
 \usepackage{color}
+\usepackage{geometry}
+
+\geometry{margin=1.2in}
 
@@
 
-\title{Doing Better Than UCT: Rational Monte Carlo Sampling in Trees}
-\author {David Tolpin \and Solomon Eyal Shimony \\
+\title{Doing Better Than UCT: \\ Rational Monte Carlo Sampling in Trees}
+\author {David Tolpin, Solomon Eyal Shimony \\
 Department of Computer Science, \\
@@
 \begin{abstract}
-
-Policies for the Multiarmed Bandit (MB) proved successful for Monte Carlo
-tree sampling (MCTS). However, MCTS differs from MB in that only the
-final choice, rather than all arm pulls, brings a reward.
-
+UCT, a state-of-the art algorithm for Monte Carlo tree sampling
+(MCTS), is based on UCB, a sampling policy for the Multi-armed Bandit
+Problem (MAB) that minimizes the accumulated regret. However, MCTS
+differs from MAB in that only the final choice, rather than all arm
+pulls, brings a reward, that is, the simple regret, as opposite to the
+accumulated regret, must be minimized. This work introduces policies for
+multi-armed bandits with lower simple regret than UCB, and an
+algorithm for MCTS which combines accumulated and simple regret
+minization and outperforms UCT. Finite-time and asymptotic analysis of
+the policies is provided, and the algorithms are empirically compared.
 \end{abstract}
@@
 
-\cite{Auer.ucb}
+The simple regret of a sampling policy for the Multi-armed Bandit
+Problem is the expected difference between the best expected reward
+$\mu_*$ and the expected reward of the arm with the best sample mean
+$\mu_j,\;\overline X_j=\max_i\overline X_i$:
+\begin{equation}
+\label{eq:simple-regret}
+\IE[R]=\sum_{j=1}^K\Delta_j\Pr(\overline X_j=\max_i\overline X_i)
+\end{equation}
+where $\Delta_j=\mu_*-\mu_j$.
 
@@
 
-\section{Doing Better Than UCB}
+\subsection{Doing better than UCB}
 
-\section{Doing Better Than UCT}
+\begin{figure}[t]
+\centering
+\includegraphics[scale=1.2]{onelevel-tree.pdf}
+\caption{Simple regret in MAB: the best arm}
+\label{fig:onelevel-tree}
+\end{figure}
+
+\begin{figure}[t]
+\centering
+\includegraphics[scale=0.64]{onelevel-64.pdf}
+\caption{Simple regret in MAB: UCB vs. $\varepsilon$-greedy vs. UVB}
+\label{fig:onelevel-64}
+\end{figure}
+
+\subsection{Doing better than UCT}
+
+\begin{figure}[t]
+\centering
+\includegraphics[scale=1.2]{twolevel-tree.pdf}
+\caption{MCTS: a path to the best arm}
+\label{fig:twolevel-tree}
+\end{figure}
+
+\begin{figure}[t]
+\centering
+\includegraphics[scale=0.64,trim=0pt 0pt 0pt 24pt,clip]{twolevel-64.pdf}
+\caption{MCTS: UCT vs. UVT vs. VCT vs. Random}
+\label{fig:twolevel-64}
+\end{figure}
+
+\section{Empirical Evaluation}
+
+\subsection{Simple regret in MAB}
+
+\begin{figure}[t]
+\centering
+\includegraphics[scale=0.64,trim=0pt 0pt 0pt 0pt,clip]{onelevel-8-16-32-64.pdf}
+\caption{Simple regret in MAB, varying number of arms}
+\label{fig:onelevel-8-16-32-64}
+\end{figure}
+
+\subsection{Monte Carlo tree search}
+
+\begin{figure}[t]
+\centering
+\includegraphics[scale=0.64,trim=0pt 0pt 0pt 0pt,clip]{twolevel-8-16-32-64.pdf}
+\caption{Monte carlo tree search, varying number of arms}
+\label{fig:twolevel-8-16-32-64}
+\end{figure}
+
+\subsection{The sailing domain}
+
+\begin{figure}[t]
+  \begin{minipage}[b]{0.5\linewidth} \centering
+    \includegraphics[scale=0.45,trim=0pt 0pt 0pt
+    0pt,clip]{sailing-size=8-factor=_8438.pdf}
+  \end{minipage}
+  \begin{minipage}[b]{0.5\linewidth} \centering
+    \includegraphics[scale=0.45,trim=0pt 0pt 0pt
+    0pt,clip]{sailing-size=8-nsamples=185.pdf}
+  \end{minipage}
+  \caption{The sailing domain, a smaller lake}
+  \label{fig:sailing-8}
+\end{figure}
+
+\begin{figure}[t]
+  \begin{minipage}[b]{0.5\linewidth} \centering
+    \includegraphics[scale=0.45,trim=0pt 0pt 0pt
+    0pt,clip]{sailing-size=13-factor=_5625.pdf}
+  \end{minipage}
+  \begin{minipage}[b]{0.5\linewidth} \centering
+    \includegraphics[scale=0.45,trim=0pt 0pt 0pt
+    0pt,clip]{sailing-size=13-nsamples=240.pdf}
+  \end{minipage}
+  \caption{The sailing domain, a larger lake}
+  \label{fig:sailing-13}
+\end{figure}
 
@@
 Robotics Research and Production Management.
-
-$\hat{C}_f$
 
@@
 
-\section{Finite-time regret bounds}
+\section{Derivations}
+
+\subsection{Finite-time regret bounds}
 
@@
 
-\subsection{$\varepsilon$-greedy}
-
-Regret $R_\varepsilon$:
-
-\begin{eqnarray}
-\IE[R_\varepsilon]&=&\sum_{j=1}^K\Delta_j\Pr(\overline X_j=\max_i\overline X_i)
-\end{eqnarray}
+\subsubsection{$\varepsilon$-greedy}
 
@@
 \begin{eqnarray}
+\label{eq:pr-epsilon-greedy}
 \Pr(\overline X_j=\max_i\overline X_i)&\le& 2\left(\frac {\mu}{2} e^{-\mu/8} + \frac {2e^{\Delta_j^2/2}}{\Delta_j^2}e^{-\Delta_j^2 \mu/4}\right)\nonumber\\
@@
 \end{eqnarray}
+
+A bound on the simple regret $R_\varepsilon$ of an
+$\varepsilon$-greedy policy follows from substituting
+(\ref{eq:pr-epsilon-greedy}) into (\ref{eq:simple-regret}):
 
@@
 
-\subsection{UCB}
+\subsubsection{UCB}
 
@@
 
-\subsection{UVB}
+\subsubsection{UVB}
 
@@
 
-\section{Asymptotic bounds for $\varepsilon$-greedy policies}
-
-\section{Case study: $K-1$ and One Arms}
+\subsection{Asymptotic bounds for $\varepsilon$-greedy policies}
 

### 2011-06-22 paper/rmcts/paper.tex
added grt algorithms

@@
   \includegraphics[scale=0.64]{onelevel-64.pdf}
-\caption{Simple regret in MAB: UCB vs. $\varepsilon$-greedy vs. UVB}
+  \caption{Simple regret in MAB: UCB vs. $\varepsilon$-greedy vs. Random}
   \label{fig:onelevel-64}
@@
   \includegraphics[scale=0.64,trim=0pt 0pt 0pt 24pt,clip]{twolevel-64.pdf}
-\caption{MCTS: UCT vs. UVT vs. VCT vs. Random}
+  \caption{MCTS: UCT vs. GRT vs. GCT vs. Random}
   \label{fig:twolevel-64}
@@
 \section{Empirical Evaluation}
-
-\subsection{Simple regret in MAB}
-
-\begin{figure}[t]
-\centering
-\includegraphics[scale=0.64,trim=0pt 0pt 0pt 0pt,clip]{onelevel-8-16-32-64.pdf}
-\caption{Simple regret in MAB, varying number of arms}
-\label{fig:onelevel-8-16-32-64}
-\end{figure}
 
@@
     0pt,clip]{sailing-size=8-factor=_8438.pdf}
+    {a. varying number of samples} 
  \end{minipage}
@@
     0pt,clip]{sailing-size=8-nsamples=185.pdf}
+    {b. varying exploration factor}
   \end{minipage}
@@
     0pt,clip]{sailing-size=13-factor=_5625.pdf}
+    {a. varying number of samples}
   \end{minipage}
@@
     0pt,clip]{sailing-size=13-nsamples=240.pdf}
+    {b. varying exploration factor}
   \end{minipage}

### 2011-06-22 code/figures.R

@@ -3,8 +3,8 @@ ALGORITHMS <- c('rnd', 'ucb', 'sve')
 draw.regrets <- function(regrets_file, legend.position='topright', xlab=expression(N[samples])) {
 	regrets <- read.table(regrets_file, header=T)
 	xlim <- range(regrets$nsamples)
-	ylim <- range(0, regrets[-1])
-	plot(x=regrets$nsamples, type='n', log="x", xlim=xlim, ylim=ylim, yaxs='i', xlab=xlab, ylab='Regret', main=unlist(strsplit(regrets_file, "\\."))[[1]])
+	ylim <- range(regrets[-1])
+	plot(x=regrets$nsamples, type='n', log="xy", xlim=xlim, ylim=ylim, yaxs='i', xlab=xlab, ylab='Regret', main=unlist(strsplit(regrets_file, "\\."))[[1]])
 	i <- 1
 	for(alg in names(regrets)[-1]) {
 		lines(x=regrets$nsamples, y=regrets[[alg]],
@@ -44,3 +44,16 @@ draw.random <- function(prefix='random', legend.position='topright') {
   draw.regrets(paste(prefix,'32.txt',sep='-'), legend.position=legend.position)
   draw.regrets(paste(prefix,'64.txt',sep='-'), legend.position=legend.position)
 }
+
+draw.curves <- function (imin=5,imax=10,max=100, alpha=1, gamma=0.5) {
+	x=0.01*(imin:imax)*max
+	rnd <- sapply(x, function (x) exp(-gamma*x/2))
+	grd <- sapply(x, function (x) exp(-gamma*x))
+	ucb <- sapply(x, function (x) x^(-alpha))
+	ylim=range(rnd, grd, ucb)
+	plot(x=x, y=rnd, log="xy", type='o', pch=1, ylim=ylim)
+	lines(x=x, y=grd, type='o', pch=2)
+	lines(x=x, y=ucb, type='o', pch=3)
+	legend(x='topright', legend=c('RND','GRD','UCB'),pch=c(1,2,3))
+}
+

### 2011-06-22 code/flatexp/uvexp.py

@@ -18,10 +18,9 @@ UCB = 'UCB'
 UCBt = 'UCBt'
 GRD = 'GRD'
 UVB = 'UVB'
-UVB2 = 'UVB2'
 
 ALGORITHMS = [RND, UCB, GRD, UVB]
-# ALGORITHMS = [RND, UCB, UCBt, GRD, UVB, UVB2]
+# ALGORITHMS = [RND, UCB, UCBt, GRD, UVB]
 
 class Exp:
 
@@ -97,16 +96,6 @@ class Exp:
         voi = (avg==a and 1-kappa or kappa)/ni
         return voi
 
-    def voi2(self, o, a):
-        ni = len(o)
-        si = sum(o)
-        avg = si/ni
-        kappa = 1.0/len(self.outcomes)
-        kappa = kappa*kappa
-        voi = (avg==a and 1-kappa or kappa)/ni
-        return voi
-
-
     def UVB(self):
         # find best average
         abest = -1.0
@@ -124,25 +113,6 @@ class Exp:
             if vois[i] >= vbest:
                 ibest, vbest = i, vois[i]
         self.draw(ibest)
-
-    def UVB2(self):
-        # find best average
-        abest = -1.0
-        for o in self.outcomes:
-            avg = sum(o)/len(o)
-            if avg > abest:
-                abest = avg
-    
-        # find best handle
-        vois = [self.voi2(o, abest) for o in self.outcomes]
-        ibest, vbest = -1, -1
-        indices = range(len(vois))
-        random.shuffle(indices) # break ties randomly
-        for i in indices:
-            if vois[i] >= vbest:
-                ibest, vbest = i, vois[i]
-        self.draw(ibest)
-
 
     def regret(self):
         avgs = [sum(o)/len(o) for o in self.outcomes]
@@ -202,4 +172,3 @@ if __name__=="__main__":
     else:
         handles = sys.argv[1]
     experiment(locals()['handles_'+handles])
-

### 2011-06-22 code/flatexp/uvrexp.py

@@ -37,9 +37,9 @@ def highkey(x):
 def lowkey(x):
     return 0.5*(1+x**4)
 
-transform = lowkey # identity
+transform = identity
     
-def randomexp(nhandles=10, nruns=1000, samples=range(1, 16)):
+def randomexp(nhandles=10, nruns=1000, samples=[4,8,16,32,64,128]):
     print "nsamples "+" ".join("r_"+alg for alg in ALGORITHMS)
     def mkhandles():
         return [RandomHandle(transform(random.random())) for i in range(nhandles)]

### 2011-06-23 code/flatexp/uvexp.py
UQB in flat experiment

@@ -18,8 +18,9 @@ UCB = 'UCB'
 UCBt = 'UCBt'
 GRD = 'GRD'
 UVB = 'UVB'
+UQB = 'UQB'
 
-ALGORITHMS = [RND, UCB, GRD, UVB]
+ALGORITHMS = [RND, UCB, GRD, UVB, UQB]
 # ALGORITHMS = [RND, UCB, UCBt, GRD, UVB]
 
 class Exp:
@@ -69,6 +70,21 @@ class Exp:
             avg = sum(self.outcomes[i])/ni
             var = sum(o*o for o in self.outcomes[i])/ni-avg*avg
             v = avg+sqrt(min(0.25,var+sqrt(2*log(n)/ni))*log(n)/ni)
+            if v > vbest:
+                ibest, vbest = i, v
+        self.draw(ibest)
+
+    # Upper Qonfidence (sQuare root instead of logarithm) Bounds
+    def UQB(self):
+        n = sum(len(o) for o in self.outcomes)
+        k = len(self.outcomes)
+        ibest, vbest = -1, -1
+        indices = range(len(self.outcomes))
+        random.shuffle(indices) # break ties randomly
+        for i in indices:
+            ni = len(self.outcomes[i])
+            avg = sum(self.outcomes[i])/ni
+            v = avg+sqrt(sqrt(n)/ni/k)
             if v > vbest:
                 ibest, vbest = i, v
         self.draw(ibest)

### 2011-06-23 code/flatexp/uvrexp.py

@@ -39,18 +39,23 @@ def lowkey(x):
 
 transform = identity
     
-def randomexp(nhandles=10, nruns=1000, samples=[4,8,16,32,64,128]):
+MINSF=8
+def randomexp(nhandles=10, nruns=1000, minsf=MINSF, sfstep=2, nsf=8):
     print "nsamples "+" ".join("r_"+alg for alg in ALGORITHMS)
     def mkhandles():
         return [RandomHandle(transform(random.random())) for i in range(nhandles)]
     handles = mkhandles()
-    for nsamples in [nhandles*i for i in samples]:
+    nsamples = minsf*nhandles
+    for i in range(nsf):
         print nsamples,
-        for alg in ALGORITHMS:
-            print repeat_alg(alg=alg, mkhandles=mkhandles, nsamples=nsamples, nruns=nruns)[1],
-        print
         sys.stdout.flush()
+        for alg in ALGORITHMS:
+            print " %#8f" % repeat_alg(alg=alg, mkhandles=mkhandles, nsamples=nsamples, nruns=nruns)[1],
+            sys.stdout.flush()
+        print
+        nsamples = int(nsamples*sfstep)
 
 if __name__=="__main__":
     transform = locals()[sys.argv[3]]
-    randomexp(nhandles=int(sys.argv[1]), nruns=int(sys.argv[2]))
+    minsf = len(sys.argv)==4 and MINSF or int(sys.argv[4])
+    randomexp(nhandles=int(sys.argv[1]), nruns=int(sys.argv[2]), minsf=minsf)

### 2011-06-23 code/flatexp/uvexp.py
more experiments

@@ -74,7 +74,7 @@ class Exp:
                 ibest, vbest = i, v
         self.draw(ibest)
 
-    # Upper Qonfidence (sQuare root instead of logarithm) Bounds
+    # Upper Qonfidence (root instead of log) Bounds
     def UQB(self):
         n = sum(len(o) for o in self.outcomes)
         k = len(self.outcomes)
@@ -84,7 +84,7 @@ class Exp:
         for i in indices:
             ni = len(self.outcomes[i])
             avg = sum(self.outcomes[i])/ni
-            v = avg+sqrt(sqrt(n)/ni/k)
+            v = avg+sqrt(sqrt(n)/ni)
             if v > vbest:
                 ibest, vbest = i, v
         self.draw(ibest)

### 2011-06-23 code/flatexp/uvexp.py
case results

@@ -172,12 +172,12 @@ def repeat_alg(alg=UCB, handles=handles_symmetric, nsamples=10, nruns=1000):
     regret = sum(r[1] for r in results)/nruns
     return (drawcounts, regret)
 
-def experiment(handles, nruns=10000, samples=range(4, 16)):
+def experiment(handles, nruns=10000, samples=[2**i for i in range(2,8)]):
     print "nsamples "+" ".join("r_"+alg for alg in ALGORITHMS)
     for nsamples in [len(handles)*i for i in samples]:
         print nsamples,
         for alg in ALGORITHMS:
-            print repeat_alg(alg=alg, handles=handles, nsamples=nsamples, nruns=nruns)[1],
+            print "%6f" % repeat_alg(alg=alg, handles=handles, nsamples=nsamples, nruns=nruns)[1],
         print
         sys.stdout.flush()
 

### 2011-06-23 code/flatexp/uvexp.py
better range for cases

@@ -172,7 +172,7 @@ def repeat_alg(alg=UCB, handles=handles_symmetric, nsamples=10, nruns=1000):
     regret = sum(r[1] for r in results)/nruns
     return (drawcounts, regret)
 
-def experiment(handles, nruns=10000, samples=[2**i for i in range(2,8)]):
+def experiment(handles, nruns=10000, samples=[2**i for i in range(7)]):
     print "nsamples "+" ".join("r_"+alg for alg in ALGORITHMS)
     for nsamples in [len(handles)*i for i in samples]:
         print nsamples,

### 2011-07-26 code/sailing/figures.R
rcq reporting

@@ -35,3 +35,10 @@ draw.costs.of.nsamples <- function(costs_file, legend.position='right', log="x")
   legend(x=legend.position, legend=algorithms, pch=1:length(algorithms))
 }
   
+draw.rcq <- function() {
+	layout(cbind(c(1,2), c(3,4))); 
+	for(x in c(199, 397, 793, 1585)) 
+		draw.costs.of.factor(paste('../../exp/sailing/rcq-size=6-nsamples=', x, '.txt', sep=''),
+							 legend.position='right')
+}
+							 

### 2011-07-27 code/flatexp/uvexp.py
removed legacy python code

@@ -1,190 +1,0 @@
-import random
-from math import sqrt, log, exp
-
-import sys
-
-class Handle:
-    def __init__(self, mean, draw):
-        self.mean = mean
-        self.draw = draw
-
-def perm(seq):
-    seq = seq[:]
-    random.shuffle(seq)
-    return seq
-
-RND = 'RND'
-UCB = 'UCB'
-UCBt = 'UCBt'
-GRD = 'GRD'
-UVB = 'UVB'
-UQB = 'UQB'
-
-ALGORITHMS = [RND, UCB, GRD, UQB]
-# ALGORITHMS = [RND, UCB, UCBt, GRD, UVB, UQB]
-
-class Exp:
-
-    def __init__(self, alg, handles):
-        self.alg = getattr(self, alg)
-        self.handles = handles
-
-    def run(self, nsamples):
-        self.predraw()
-        for i in range(nsamples):
-            self.alg()
-        return self.regret()
-
-    def predraw(self):
-        self.outcomes = [[handle.draw()] for handle in self.handles]
-        
-    def draw(self, i):
-        self.outcomes[i].append(self.handles[i].draw())
-
-    # Random
-    def RND(self):
-        self.draw(random.randrange(len(self.handles)))
-    
-    # Upper Confidence Bounds
-    def UCB(self):
-        n = sum(len(o) for o in self.outcomes)
-        ibest, vbest = -1, -1
-        indices = range(len(self.outcomes))
-        random.shuffle(indices) # break ties randomly
-        for i in indices:
-            ni = len(self.outcomes[i])
-            avg = sum(self.outcomes[i])/ni
-            v = avg+sqrt(2*log(n)/ni)
-            if v > vbest:
-                ibest, vbest = i, v
-        self.draw(ibest)
-
-    # UCB Tuned 
-    def UCBt(self):
-        n = sum(len(o) for o in self.outcomes)
-        ibest, vbest = -1, -1
-        indices = range(len(self.outcomes))
-        random.shuffle(indices) # break ties randomly
-        for i in indices:
-            ni = len(self.outcomes[i])
-            avg = sum(self.outcomes[i])/ni
-            var = sum(o*o for o in self.outcomes[i])/ni-avg*avg
-            v = avg+sqrt(min(0.25,var+sqrt(2*log(n)/ni))*log(n)/ni)
-            if v > vbest:
-                ibest, vbest = i, v
-        self.draw(ibest)
-
[…]

### 2011-07-27 code/flatexp/uvrexp.py

@@ -1,66 +1,0 @@
-from uvexp import *
-import math
-import sys
-
-MINSF=8        # initial sampling factor in the random experiment
-MAXRWRD=1.0    # reward of a random arm is [1-MAXRWRD, MAXRWRD)
-
-class RandomHandle(Handle):
-    def __init__(self, mean):
-        flip = random.randrange(3)
-        if flip==0: # bi-value
-            draw = lambda: random.random() < mean and 1.0 or 0.0
-        elif flip==1: # constant
-            draw = lambda: mean
-        else: # triangular
-            mode = mean
-            mean = (mode+1)/3
-            draw = lambda: random.triangular(0, 1, mode)
-
-        Handle.__init__(self, mean, draw)
-
-def repeat_alg(alg=UCB, mkhandles=lambda: [RandomHandle(random.random()) for i in range(10)], nsamples=10, nruns=1000):
-    
-    results = [ try_alg(alg=alg, handles=mkhandles(), nsamples=nsamples)
-                for i in range(nruns) ]
-    drawcounts = [round(sum(len(r[0][i]) for r in results)/float(nruns))
-                  for i in range(len(results[0][0]))]
-    regret = sum(r[1] for r in results)/nruns
-    return (drawcounts, regret)
-
-def identity(x): 
-    return x
-
-def sigmoid(x):
-    return math.tanh(5*(x-0.5))
-    
-def highkey(x):
-    return 0.5*x**0.25
-
-def lowkey(x):
-    return 0.5*(1+x**4)
-
-transform = identity
-    
-def randomexp(nhandles=10, nruns=1000, minsf=MINSF, sfstep=2, nsf=8):
-    assert (MAXRWRD<=1.0 and MAXRWRD>0.5)
-    print "nsamples "+" ".join("r_"+alg for alg in ALGORITHMS)
-    def mkhandles():
-        return [RandomHandle(transform(1-MAXRWRD
-                                       +(2*MAXRWRD-1)*random.random()))
-                for i in range(nhandles)]
-    handles = mkhandles()
-    nsamples = minsf*nhandles
-    for i in range(nsf):
-        print nsamples,
-        sys.stdout.flush()
-        for alg in ALGORITHMS:
-            print " %#8f" % repeat_alg(alg=alg, mkhandles=mkhandles, nsamples=nsamples, nruns=nruns)[1],
-            sys.stdout.flush()
-        print
[…]

### 2011-07-28 paper/rmcts/paper.tex
pdfs automatically

@@
 
-\begin{figure}[t]
-  \begin{minipage}[b]{0.5\linewidth} \centering
-    \includegraphics[scale=0.45,trim=0pt 0pt 0pt
-    0pt,clip]{sailing-size=8-factor=_8438.pdf}
-    {a. varying number of samples} 
- \end{minipage}
-  \begin{minipage}[b]{0.5\linewidth} \centering
-    \includegraphics[scale=0.45,trim=0pt 0pt 0pt
-    0pt,clip]{sailing-size=8-nsamples=185.pdf}
-    {b. varying exploration factor}
-  \end{minipage}
-  \caption{The sailing domain, a smaller lake}
-  \label{fig:sailing-8}
-\end{figure}
-
-\begin{figure}[t]
-  \begin{minipage}[b]{0.5\linewidth} \centering
-    \includegraphics[scale=0.45,trim=0pt 0pt 0pt
-    0pt,clip]{sailing-size=13-factor=_5625.pdf}
-    {a. varying number of samples}
-  \end{minipage}
-  \begin{minipage}[b]{0.5\linewidth} \centering
-    \includegraphics[scale=0.45,trim=0pt 0pt 0pt
-    0pt,clip]{sailing-size=13-nsamples=240.pdf}
-    {b. varying exploration factor}
-  \end{minipage}
-  \caption{The sailing domain, a larger lake}
-  \label{fig:sailing-13}
-\end{figure}
+%\begin{figure}[t]
+%  \begin{minipage}[b]{0.5\linewidth} \centering
+%    \includegraphics[scale=0.45,trim=0pt 0pt 0pt
+%    0pt,clip]{sailing-size=8-factor=_8438.pdf}
+%    {a. varying number of samples} 
+% \end{minipage}
+%  \begin{minipage}[b]{0.5\linewidth} \centering
+%    \includegraphics[scale=0.45,trim=0pt 0pt 0pt
+%    0pt,clip]{sailing-size=8-nsamples=185.pdf}
+%    {b. varying exploration factor}
+%  \end{minipage}
+%  \caption{The sailing domain, a smaller lake}
+%  \label{fig:sailing-8}
+%\end{figure}
+%
+%\begin{figure}[t]
+%  \begin{minipage}[b]{0.5\linewidth} \centering
+%    \includegraphics[scale=0.45,trim=0pt 0pt 0pt
+%    0pt,clip]{sailing-size=13-factor=_5625.pdf}
+%    {a. varying number of samples}
+%  \end{minipage}
+%  \begin{minipage}[b]{0.5\linewidth} \centering
+%    \includegraphics[scale=0.45,trim=0pt 0pt 0pt
+%    0pt,clip]{sailing-size=13-nsamples=240.pdf}
+%    {b. varying exploration factor}
+%  \end{minipage}
+%  \caption{The sailing domain, a larger lake}
+%  \label{fig:sailing-13}
+%\end{figure}
 

### 2011-07-28 code/figures.R

@@ -1,12 +1,12 @@
 ALGORITHMS <- c('rnd', 'ucb', 'sve')
 
 
-draw.regrets <- function(regrets_file, log="xy", legend.position='topright', xlab=expression(N[samples])) {
+draw.regrets <- function(regrets_file, log="xy", legend.position='topright', xlab=expression(N[samples]), main=NA) {
 	regrets <- read.table(regrets_file, header=T)
 	xlim <- range(regrets$nsamples)
 	values <- unlist(regrets[-1])
 	ylim <- range(values[values>0])
-	plot(x=regrets$nsamples, type='n', log=log, xlim=xlim, ylim=ylim, yaxs='i', xlab=xlab, ylab='Regret', main=unlist(strsplit(regrets_file, "\\."))[[1]])
+	plot(x=regrets$nsamples, type='n', log=log, xlim=xlim, ylim=ylim, yaxs='i', xlab=xlab, ylab='Regret', main=if(is.na(main)) unlist(strsplit(regrets_file, "\\."))[[1]] else main)
 	i <- 1
 	for(alg in names(regrets)[-1]) {
 		y <- regrets[[alg]]
@@ -61,3 +61,13 @@ draw.curves <- function (imin=5,imax=10,max=100, alpha=1, gamma=0.5,log="xy") {
 	legend(x='topright', legend=c('RND','GRD','UCB'),pch=c(1,2,3))
 }
 
+make.pdfs <- function() {
+  for(prefix in c('flat-low-key', 'flat-trilevel', 'tree-trilevel', 'tree-identity')) {
+    for(k in c('16', '32', '64')) {
+      basename <- paste(prefix, '-k=', k, '-uqb=8', sep='')
+      pdf(file=paste(basename, 'pdf', sep='.'), width=6, height=4.5)
+      draw.regrets(paste(basename,'txt', sep='.'), legend.position='right', main='')
+      dev.off()
+    }
+  }
+}

### 2011-07-30 code/sailing/figures.R
paper: limit derivation draft

@@ -1,22 +1,23 @@
-draw.costs.of.factor <- function(costs_file, legend.position='bottomright', log="x") {
+draw.costs.of.factor <- function(costs_file, legend.position='bottomright', log="x",main=NA) {
   costs <- read.table(costs_file, header=T)
   algorithms <- names(costs)[-1]
   xlim <- range(costs$FACTOR)
   ylim <- range(costs[algorithms])
   plot(x=costs$nsamples, type='n', log=log, xlim=xlim, ylim=ylim, xaxs='i',
        xlab='exploration factor', ylab='cost',
-       main=do.call(paste,as.list(unlist(strsplit(costs_file,
-         "-|\\.txt",fixed=FALSE)))[-1]))
+       main=if(is.na(main)) do.call(paste,as.list(unlist(strsplit(costs_file,
+                            "-|\\.txt",fixed=FALSE)))[-1])
+            else main)
   i <- 1
   for(alg in algorithms) {
     lines(x=costs$FACTOR, y=costs[[alg]],
           type="o", pch=i)
     i <- i+1
   }
-  legend(x=legend.position, legend=algorithms, pch=1:length(algorithms))
+  legend(x=legend.position, legend=algorithms, pch=1:length(algorithms), bg='white')
 }
 
-draw.costs.of.nsamples <- function(costs_file, legend.position='right', log="x") {
+draw.costs.of.nsamples <- function(costs_file, legend.position='right', log="x", main=NA) {
   costs <- read.table(costs_file, header=T)
   algorithms <- names(costs)[-1]
   xlim <- range(costs$SAMPLES)
@@ -24,15 +25,16 @@ draw.costs.of.nsamples <- function(costs_file, legend.position='right', log="x")
   print(costs)
   plot(x=costs$nsamples, type='n', log=log, xlim=xlim, ylim=ylim, xaxs='i',
        xlab='samples per node', ylab='cost',
-       main=do.call(paste,as.list(unlist(strsplit(costs_file,
-         "-|\\.txt",fixed=FALSE)))[-1]))
+       main=if(is.na(main)) do.call(paste,as.list(unlist(strsplit(costs_file,
+                            "-|\\.txt",fixed=FALSE)))[-1])
+            else main)
   i <- 1
   for(alg in algorithms) {
     lines(x=costs$SAMPLES, y=costs[[alg]],
           type="o", pch=i)
     i <- i+1
   }
-  legend(x=legend.position, legend=algorithms, pch=1:length(algorithms))
[…]

### 2011-08-02 paper/rmcts/paper.tex
paper: limit derivation draft
derivations

@@
 
-\begin{figure}[t]
-  \centering
-  \includegraphics[scale=1.2]{onelevel-tree.pdf}
-  \caption{Simple regret in MAB: the best arm}
-  \label{fig:onelevel-tree}
-\end{figure}
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[scale=0.64]{onelevel-64.pdf}
-  \caption{Simple regret in MAB: UCB vs. $\varepsilon$-greedy vs. Random}
-  \label{fig:onelevel-64}
-\end{figure}
-
 \subsection{Doing better than UCT}
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[scale=1.2]{twolevel-tree.pdf}
-  \caption{MCTS: a path to the best arm}
-  \label{fig:twolevel-tree}
-\end{figure}
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[scale=0.64,trim=0pt 0pt 0pt 24pt,clip]{twolevel-64.pdf}
-  \caption{MCTS: UCT vs. GRT vs. GCT vs. Random}
-  \label{fig:twolevel-64}
-\end{figure}
 
@@
 
+\subsection{Simple regret in multi-armed bandits}
+
+\begin{figure}[t]
+  \begin{minipage}[c]{0.5\linewidth}
+    \centering
+    \includegraphics[scale=1.0]{onelevel-tree.pdf}\\
+    \vspace{4em}
+    a. search tree
+  \end{minipage}
+  \begin{minipage}[c]{0.5\linewidth}
+    \centering
+    \includegraphics[scale=0.5]{flat-trilevel-k=64-uqb=8.pdf}\\
+    b. regret vs. number of samples
+  \end{minipage}
+  \label{fig:mab-simple-regret}
+  \caption{Simple regret in MAB}
+\end{figure}
+
 \subsection{Monte Carlo tree search}
 
-\begin{figure}[t]
+\begin{figure}
+  \begin{minipage}[c]{0.5\linewidth}
     \centering
-  \includegraphics[scale=0.64,trim=0pt 0pt 0pt 0pt,clip]{twolevel-8-16-32-64.pdf}
-  \caption{Monte carlo tree search, varying number of arms}
-  \label{fig:twolevel-8-16-32-64}
+    \includegraphics[scale=0.8]{twolevel-tree.pdf}\\
+    a. search tree
+  \end{minipage}
+  \begin{minipage}[c]{0.5\linewidth}
+    \centering
+    \includegraphics[scale=0.4]{tree-identity-k=16-uqb=8.pdf}\\ 
+    b. 16 arms \\
+    \vspace{1em}
+    \includegraphics[scale=0.4]{tree-identity-k=64-uqb=8.pdf} \\
+    c. 64 arms
+ \end{minipage}
+  \label{fig:mcts-regret}
+  \caption{MCTS: a path to the best arm}
 \end{figure}
@@
 
-%\begin{figure}[t]
-%  \begin{minipage}[b]{0.5\linewidth} \centering
-%    \includegraphics[scale=0.45,trim=0pt 0pt 0pt
-%    0pt,clip]{sailing-size=8-factor=_8438.pdf}
-%    {a. varying number of samples} 
-% \end{minipage}
-%  \begin{minipage}[b]{0.5\linewidth} \centering
-%    \includegraphics[scale=0.45,trim=0pt 0pt 0pt
-%    0pt,clip]{sailing-size=8-nsamples=185.pdf}
-%    {b. varying exploration factor}
-%  \end{minipage}
-%  \caption{The sailing domain, a smaller lake}
-%  \label{fig:sailing-8}
-%\end{figure}
+\begin{figure}[t]
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[scale=0.45]{costs-size=6-group=median.pdf}\\
+    a. median cost
+  \end{minipage}
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[scale=0.45]{costs-size=6-group=minimum.pdf}\\
+    b. minimum cost
+  \end{minipage}
+  \caption{The sailing domain, $6\times 6$ lake, cost vs. number of playouts}
+  \label{fig:sailing-cost-vs-nsamples}
+\end{figure}
+
+\begin{figure}[t]
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[scale=0.45]{costs-size=6-nsamples=199.pdf}\\
+    a. 199 playouts\\
+    \vspace{1em}
+    \includegraphics[scale=0.45]{costs-size=6-nsamples=397.pdf}\\
+    b. 397 playouts\\
+  \end{minipage}
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[scale=0.45]{costs-size=6-nsamples=793.pdf}\\
+    c. 793 playouts\\
+    \vspace{1em}
+    \includegraphics[scale=0.45]{costs-size=6-nsamples=1585.pdf}\\
+    d. 1585 playouts\\
+  \end{minipage}
+  \caption{The sailing domain, $6\times 6$ lake, cost vs. factor}
+  \label{fig:sailing-cost-vs-factor}
+\end{figure}
+
+\begin{figure}[t]
+  \begin{minipage}[b]{0.333\linewidth}
+    \centering
+    \includegraphics[scale=0.35]{costs-size=3-nsamples=397.pdf}\\
+    a. $3\times 3$ lake
+  \end{minipage}
+  \begin{minipage}[b]{0.333\linewidth}
+    \centering
+    \includegraphics[scale=0.35]{costs-size=6-nsamples=397.pdf}\\
+    b. $6\times 6$ lake
+  \end{minipage}
+  \begin{minipage}[b]{0.333\linewidth}
+    \centering
+    \includegraphics[scale=0.35]{costs-size=10-nsamples=397.pdf}\\
+    b. $10\times 10$ lake
+  \end{minipage}
+  \caption{The sailing domain, 397 samples, cost vs. factor}
+  \label{fig:sailing-lake-size}
+\end{figure}
+
+\begin{figure}[t]
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[scale=0.45]{rcq-size=6-nsamples=199.pdf}\\
+    a. 199 playouts\\
+    \vspace{1em}
+    \includegraphics[scale=0.45]{rcq-size=6-nsamples=397.pdf}\\
+    b. 397 playouts\\
+  \end{minipage}
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[scale=0.45]{rcq-size=6-nsamples=793.pdf}\\
+    c. 793 playouts\\
+    \vspace{1em}
+    \includegraphics[scale=0.45]{rcq-size=6-nsamples=1585.pdf}\\
+    d. 1585 playouts\\
+  \end{minipage}
+  \caption{The sailing domain, log vs. sqrt, $6\times 6$ lake}
+  \label{fig:sailing-cost-vs-factor}
+\end{figure}
 %
@@
 
-A bound on the simple regret $R_\varepsilon$ of an
+A bound on the simple regret $r_\varepsilon$ of an
 $\varepsilon$-greedy policy follows from substituting
@@
 \begin{eqnarray}
-\IE[R_\varepsilon]&\le&\sum_{j=1}^K\Delta_j\left(\mu + \frac {4\sqrt e}
+  \IE r_\varepsilon&\le&\sum_{j=1}^K\Delta_j\left(\mu + \frac {4\sqrt e}
 {\Delta_j^2}\right)e^{-\Delta_j^2\mu/8}\nonumber\\
-&\le&K\left(\mu + \frac {4\sqrt e}{\Delta_{\min}}\right)e^{-\Delta_{\min}^2\mu/8}
+&\le&K\left(\mu + \frac {4\sqrt
+    e}{\Delta_{\min}}\right)e^{-\Delta_{\min}^2\mu/8}
+\label{eq:deriv-epsilon-greedy-b}
 \end{eqnarray}
 
-\subsubsection{UCB}
+\subsubsection{UCB($\alpha$)}
 
-Regret $R_{UCB}$:
+$n_b$ pulls of the best arm and $n_j$ pulls of the $j$th arm are
+achieved simultaneously when either
 
-\subsubsection{UVB}
+\begin{equation}
+\sqrt {\frac {\alpha \log n} {n_j-1}} \ge 1 + \sqrt { \frac {\alpha \log n} {n_b}}
+\end{equation}
 
-Regret $R_{UVB}$:
+or
 
-\subsection{Asymptotic bounds for $\varepsilon$-greedy policies}
+\begin{equation}
+\sqrt {\frac {\alpha \log n} {n_j}} \le 1 + \sqrt { \frac {\alpha \log n} {n_b - 1}}
+\label{eq:deriv-ucb-alpha-a}
+\end{equation}
+
+Inequality (\ref{eqn:deriv-ucb-alpha-a}) can be used to derive a lower
+bound on the number of pulls of a non-optimal arm,  and, consequently, the upper
+bound on the probability of selecting a non-optimal arm:
+
+\begin{eqnarray}
+\frac {\alpha \log n} {n_j}&\le& 1 + 2  \sqrt {\frac {\alpha \log n}
+  {n_b-1}} + \frac {\alpha \log n} {n_b-1}\nonumber\\
+n_j &\ge& \frac {\alpha \log n} {1+2 \sqrt {\frac {\alpha \log n}
+    {n_b-1}} + \frac {\alpha \log n} {n_b-1}}\nonumber\\
+    &\ge& \frac {\alpha \log n} {4} \mbox{ for } \alpha \log n \le n_b-1
+\end{eqnarray}
+
+The probability that the $j$th arm will be chosen can be bounded based
+on the Chernoff inequality and the union bound:
+\begin{equation}
+P_j \le 2 \exp\left(-\frac {\alpha \Delta_j^2 \log n} 8 \right) = 2n^{\frac {-\alpha \Delta_j^2} 8}
+\end{equation}
+
+The expected simple regret of an UCB($\alpha$) algorithm is thus
+bounded by the following inequality:
+\begin{equation}
+\IE r_{ucb} \le 2\sum_{j=1}^K \Delta_jn^{\frac {-\alpha \Delta_j^2} 8}
+\label{eq:deriv-ucb-alpha-b}
+\end{equation}
+
+\cite{Bubeck.pure} provides for UCB($\alpha$) a similar bound,
+polynomial in the number of pulls $n$.
+
+\subsection{Asymptotic regret analysis}
+
+Bounds (\ref{eq:deriv-epsilon-greedy-b}, \ref{eq:deriv-ucb-alpha-b}) where
+derived ignoring dependency of the allocation on expected values
+of non-optimal arms. Derivation of better bounds seems complicated; instead, insights
+can be obtained by analysing  asymptotic behavior of each of the
+algorithms in assumption that the sample mean approaches the expected
+value.
+
+\subsubsection{$\varepsilon$-greedy}
+
+\subsubsection{UCB($\alpha$)}
 

### 2011-08-03 paper/rmcts/paper.tex
in progress

@@
 
-$n_b$ pulls of the best arm and $n_j$ pulls of the $j$th arm are
+$n_*$ pulls of the best arm and $n_j$ pulls of the $j$th arm are
 achieved simultaneously when either
@@
 \begin{equation}
-\sqrt {\frac {\alpha \log n} {n_j-1}} \ge 1 + \sqrt { \frac {\alpha \log n} {n_b}}
+\sqrt {\frac {\alpha \log n} {n_j-1}} \ge 1 + \sqrt { \frac {\alpha \log n} {n_*}}
 \end{equation}
@@
 \begin{equation}
-\sqrt {\frac {\alpha \log n} {n_j}} \le 1 + \sqrt { \frac {\alpha \log n} {n_b - 1}}
+\sqrt {\frac {\alpha \log n} {n_j}} \le 1 + \sqrt { \frac {\alpha \log n} {n_* - 1}}
 \label{eq:deriv-ucb-alpha-a}
@@
 \frac {\alpha \log n} {n_j}&\le& 1 + 2  \sqrt {\frac {\alpha \log n}
-  {n_b-1}} + \frac {\alpha \log n} {n_b-1}\nonumber\\
+  {n_*-1}} + \frac {\alpha \log n} {n_*-1}\nonumber\\
 n_j &\ge& \frac {\alpha \log n} {1+2 \sqrt {\frac {\alpha \log n}
-    {n_b-1}} + \frac {\alpha \log n} {n_b-1}}\nonumber\\
-    &\ge& \frac {\alpha \log n} {4} \mbox{ for } \alpha \log n \le n_b-1
+    {n_*-1}} + \frac {\alpha \log n} {n_*-1}}\nonumber\\
+    &\ge& \frac {\alpha \log n} {4} \mbox{ for } \alpha \log n \le n_*-1
 \end{eqnarray}
@@
 The probability that the $j$th arm will be chosen can be bounded based
-on the Chernoff inequality and the union bound:
+on the Chernoff-Hoeffding bound:
 \begin{equation}
@@
 
-Bounds (\ref{eq:deriv-epsilon-greedy-b}, \ref{eq:deriv-ucb-alpha-b}) where
+Bounds (\ref{eq:deriv-epsilon-greedy-b}, \ref{eq:deriv-ucb-alpha-b}) were
 derived ignoring dependency of the allocation on expected values
-of non-optimal arms. Derivation of better bounds seems complicated; instead, insights
-can be obtained by analysing  asymptotic behavior of each of the
-algorithms in assumption that the sample mean approaches the expected
-value.
+of non-optimal arms. Derivation of better bounds seems
+complicated. Instead, one can analyse symptotic behavior of each of the
+algorithms in assumption that the sample mean is close to the expectation.
 
 \subsubsection{$\varepsilon$-greedy}
+
+A common technique \cite{Auer.ucb} to bound the probability of
+selecting a non-optimal arm $j$ is to split the interval between
+$\mu_j$ and $\mu_*$ at the middle and bound the probability of
+selecting the $j$th arm by the sum of the probabilities that the samle
+mean $\overline x_j$ of the $j$th arm is greater than $\mu_j+\frac
+{\Delta_j} 2$ and the sample mean $\overline x_*$ of the optimal arm is
+less than $\mu_*- \frac {\Delta_j} 2$. Each of the probabilities can
+be bound using the Chernoff-Hoefding bound.
+
+Splitting the interval between $\mu_j$ and $\mu_*$ half-way is an
+arbirtrary decision. By selecting a different split point
+depending on the allocation of samples, one can obtain a tighter
+bound. On the other hand, there is a value of $\varepsilon$ that
+minimizes the bound for the $\varepsilon$-greedy allocation scheme.
+Thus, the  $\varepsilon$-greedy allocation scheme can be tuned by
+finding $\varepsilon$ that minimizes the bound obtained by
+selecting the best split point.
+
+Split the interval $[\mu_j, \mu_*]$ at $\mu_j+\delta_j$. The
+probability $P_j$ for selecting the $j$th arm is bounded as:
+
+\begin{equation}
+P_j\le exp\left(-2\delta_j^2n_j\right)+exp\left(-2(\Delta_j-\delta_j)^2n_*\right)
+\end{equation}
+
+For the $\varepsilon$-greedy scheme, $n_i=\frac {\varepsilon n} K$,
+$n_*=(1-\varepsilon)n$, thus
+
+\begin{equation}
+P_j \le \exp\left(-\frac {2\delta_j^2\varepsilon n} (K-1)\right) +\exp\left(-2(\Delta_j-\delta_j)^2(1-\varepsilon)n\right)
+\end{equation}
+
+In general, $\varepsilon$ and $\delta$ which minimize $P_j$ will
+depend of $n$. For a constant $\varepsilon$, require that the
+exponents in both terms are equal
+\begin{eqnarray}
+\exp\left(-\frac {2\delta_j^2\varepsilon n} (K-1)\right)
+   &=& \exp\left(-2(\Delta_j-\delta_j)^2(1-\varepsilon)n\right)\\
+\frac {\delta_j} {\Delta_j - \delta_j}
+   &=& \sqrt { \frac {(K-1)(1-\varepsilon)} \varepsilon }
+\label{eq:deriv-exps-equal}
+\end{eqnarray}
+and find conditional minima of $P_j(\delta_j, \epsilon)$ by requiring
+either $\frac {\partial P_j} {\partial \delta_j}=0$ or ${\partial P_j}
+{\partial \varepsilon}=0$.
+
+$\frac {\partial P_j} {\partial \delta_j}=0$:
+
+\begin{eqnarray*}
+\frac {4\delta_j\varepsilon n} (K-1)
+  &-& 4(\Delta_j-\delta_j)(1-\varepsilon)n=0\\
+\frac {\delta_j} {\Delta_j-\delta_j}&=&{ \frac {(K-1)(1-\varepsilon)} \varepsilon }
+\end{eqnarray*}
+and together with (\ref{eq:deriv-exps-equal}) this gives:
+\begin{eqnarray}
+\delta_j&=&\frac {Delta_j} 2\nonumber\\
+\varepsilon&=&1-\frac 1 K
+\label{eq:deriv-uniform-sampling}
+\end{eqnarray}
+
+That is, the case $\frac {\partial P_j} {\partial \delta_j}=0$
+corresponds to uniform random sampling, the tightest bound is obtained
+by splitting the interval at the middle and the convergence rate
+is $exp\left(-\frac {\Delta_j^2n} {2K}\right)$.
 

### 2011-08-03 paper/rmcts/paper.tex
derivations

@@
 
-\begin{figure}[t]
+\begin{figure}[ht]
   \begin{minipage}[c]{0.5\linewidth}
@@
 
-\begin{figure}[t]
+\begin{figure}[ht]
   \begin{minipage}[b]{0.5\linewidth}
@@
 
-\begin{figure}[t]
+\begin{figure}[ht]
   \begin{minipage}[b]{0.333\linewidth}
@@
   \caption{The sailing domain, log vs. sqrt, $6\times 6$ lake}
-  \label{fig:sailing-cost-vs-factor}
+  \label{fig:sailing-rcq-vs-factor}
 \end{figure}
 %
-%\begin{figure}[t]
+%\begin{figure}[ht]
 %  \begin{minipage}[b]{0.5\linewidth} \centering
@@
 
-Inequality (\ref{eqn:deriv-ucb-alpha-a}) can be used to derive a lower
+Inequality (\ref{eq:deriv-ucb-alpha-a}) can be used to derive a lower
 bound on the number of pulls of a non-optimal arm,  and, consequently, the upper
@@
 
-The probability that the $j$th arm will be chosen can be bounded based
-on the Chernoff-Hoeffding bound:
+The probability that the $j$th arm will be chosen can be bounded using
+the Chernoff-Hoeffding bound:
 \begin{equation}
@@
 
-\cite{Bubeck.pure} provides for UCB($\alpha$) a similar bound,
+\cite{Bubeck.pure} establishes for UCB($\alpha$) a similar bound,
 polynomial in the number of pulls $n$.
@@
 \begin{equation}
-P_j \le \exp\left(-\frac {2\delta_j^2\varepsilon n} (K-1)\right) +\exp\left(-2(\Delta_j-\delta_j)^2(1-\varepsilon)n\right)
+P_j \le \exp\left(-\frac {2\delta_j^2\varepsilon n} {K-1}\right) +\exp\left(-2(\Delta_j-\delta_j)^2(1-\varepsilon)n\right)
 \end{equation}
@@
 \begin{eqnarray}
-\exp\left(-\frac {2\delta_j^2\varepsilon n} (K-1)\right)
+\exp\left(-\frac {2\delta_j^2\varepsilon n} {K-1}\right)
    &=& \exp\left(-2(\Delta_j-\delta_j)^2(1-\varepsilon)n\right)\\
@@
 \end{eqnarray}
-and find conditional minima of $P_j(\delta_j, \epsilon)$ by requiring
-either $\frac {\partial P_j} {\partial \delta_j}=0$ or ${\partial P_j}
+and find conditional minima of $P_j(\delta_j, \epsilon)$ satisfying
+either $\frac {\partial P_j} {\partial \delta_j}=0$ or $\frac {\partial P_j}
 {\partial \varepsilon}=0$.
 
-$\frac {\partial P_j} {\partial \delta_j}=0$:
+\vspace{1em}
+\paragraph{a. {\large $\frac {\partial P_j} {\partial \delta_j}=0$}}
 
 \begin{eqnarray*}
-\frac {4\delta_j\varepsilon n} (K-1)
+\frac {4\delta_j\varepsilon n} {K-1}
   &-& 4(\Delta_j-\delta_j)(1-\varepsilon)n=0\\
@@
 \begin{eqnarray}
-\delta_j&=&\frac {Delta_j} 2\nonumber\\
-\varepsilon&=&1-\frac 1 K
+\delta_j&=&\frac {\Delta_j} 2\nonumber\\
+\varepsilon&=&1-\frac 1 K\nonumber\\
+\IE r_{\varepsilon=1-1/K}&\le&B_{uniform}=2\exp\left(-\frac {\Delta_j^2n} {2K}\right)
 \label{eq:deriv-uniform-sampling}
@@
 That is, the case $\frac {\partial P_j} {\partial \delta_j}=0$
-corresponds to uniform random sampling, the tightest bound is obtained
-by splitting the interval at the middle and the convergence rate
-is $exp\left(-\frac {\Delta_j^2n} {2K}\right)$.
+corresponds to \emph{uniform random sampling}, and the tightest bound is obtained
+by splitting the interval at the middle.
+
+\paragraph{b. {\large $\frac {\partial P_j} {\partial \varepsilon}=0$}}
+
+\begin{eqnarray*}
+\frac {2\delta_j^2 n} {K-1}
+  &-& 2(\Delta_j-\delta_j)^2n=0\\
+\frac {\delta_j} {\Delta_j-\delta_j}&=&\sqrt {K-1}
+\end{eqnarray*}
+and together with (\ref{eq:deriv-exps-equal}) this gives:
+\begin{eqnarray}
+\frac {\delta_j} {\Delta_j}&=&\frac 1 {\sqrt {K-1}+1}\nonumber\\
+\varepsilon&=&\frac 1 2\nonumber\\
+\IE
+r_{\varepsilon=1/2}&\le&B_{1/2\mbox{-}greedy}=2\exp\left(-\frac
+  {\Delta_j^2n} {(\sqrt{K-1}+1)^2}\right)\nonumber\\
+\lim_{K\to\infty}
+\mathbb{B}r_{1/2\mbox{-}greedy}^K&=&2\exp\left(-{\Delta_j^2n}\right)\nonumber\\
+\mbox{for large $K$: }\mathbb{B}r_{1/2\mbox{-}greedy}&\approx&2\exp\left(-\frac {\Delta_j^2n} K\right)
+\label{eq:deriv-half-greedy-sampling}
+\end{eqnarray}
+
+That is, the case $\frac {\partial P_j} {\partial \varepsilon}=0$
+corresponds to \emph{0.5-greedy sampling}, and for a large number of
+arms $K$ 
+\begin{equation}
+\mathbb{B}r_{1/2\mbox{-}greedy}\approx \mathbb{B}r_{uniform}^2
+\label{eq:derive-uniform-vs-half-greedy}
+\end{equation}
+Equation (\ref{eq:derive-uniform-vs-half-greedy}) suggests that
+0.5-greedy sampling has a much higher convergence rate than uniform
+sampling for a large number of arms.
 
@@
 
+The $UCB(\alpha)$ allocation scheme distributes arm pulls such that
+the upper bound on the probability of selecting a non-optimal arm is
+asymptotically the same for all arms. Indeed, for large n
+\begin{eqnarray}
+n_i&\ll&n\\
+n_*&\approx&n\\
+\sqrt {\frac {2 \log n} {n_j}}&\approx&\Delta_j+\sqrt {\frac {2\log n}
+  {n_*}} \\
+&\Rightarrow& n_j \approx \frac {2\log n} {\Delta_j^2}
+\label{eq:deriv-ucb-asymptotic-nj}
+\end{eqnarray}
+and by Hoeffding-Chernoff bound the upper bound $\mathbb{B}P_j$ on the probability
+$P_j^{ucb}$ to choose the $j$th arm is
+\begin{equation}
+P_j^{ucb}\le\mathbb{B}P_j^{ucb}\approx2\exp\left(-2\Delta_j^2\frac {2\log n}{\Delta_j^2}\right)
+             =2n^{-4}
+\end{equation}
+
+That is, UCB($\alpha$) supposedly distributes pulls of non-optimal
+arms better than $\varepsilon$-greedy
+scheme which may undersample arms with high expectation. The
+polynomial, rather than exponential, convergence rate (Equation
+\ref{eq:deriv-ucb-alpha-b}) is due to oversampling the best arm.
+
+Replacing the logarithm with a faster growing subpolynomial
+function, like square root (UQB), results in a superpolynomial convergence
+rate of the upper bound:
+\begin{equation}
+P_j^{uqb}=\mathbb{B}P_j^{uqb}\approx2\exp\left(-\beta\Delta_j^2\frac {\sqrt n}{\Delta_j^2}\right)
+            =2\exp(-2\beta\sqrt n)
+ \end{equation}
+where $\beta$ is a constant depending on reward bounds.
+Finite time convergence can still be shown similarly to UCB.
+
 \end{document}

### 2011-08-03 paper/rmcts/paper.tex
main results

@@
 
-\section{Introduction}
+\section{Introduction and definitions}
+
+The Multi-armed Bandit is a set of $K$ arms. Each arm can be pulled
+multiple times. When the $i$th arm is pulled, a random reward $X_i$
+from an unknown stationary distribution is returned.  The reward is 
+bounded between 0 and 1.
 
@@
 Problem is the expected difference between the best expected reward
-$\mu_*$ and the expected reward of the arm with the best sample mean
-$\mu_j,\;\overline X_j=\max_i\overline X_i$:
+$\mu_*$ and the expected reward $\mu_j$ of the arm with the best sample mean
+$\overline X_j=\max_i\overline X_i$:
 \begin{equation}
@@
 where $\Delta_j=\mu_*-\mu_j$.
+Strategies that minimize the simple reward are called pure exploration
+strategies \cite{Bubeck.pure}.
+
+Such strategies are used to select the best arm, and, by
+extension, the best action in search, when every action in the
+current world state is viewed as a different arm, and the best arm
+must be chosen. Algorithm UCB that maximizes the accumulated reward
+\cite{Auer.ucb} had been extended to tree search as UCT \cite{Kocsis.uct}.
+
+UCT-driven search achieves good results in many search problems. At each
+search step, the algorithm allocates Monte Carlo playouts to choose
+the best action, thereby minimizing the simple regret. However, the
+simple and the accumulated regret cannot be minimized simultaneously;
+moreover, \cite{Bubeck.pure} shows that in many cases the smaller the
+accumulated regret, the greater is the simple regret.
+
+UCT performance can be improved by combining UCB with a sampling
+scheme that minimizes the simple regret of selecting an action at the
+current root node. Indeed, the algorithm must select an action with
+the minimum regret \emph{given assumption that after performing the selected action
+the algorithm performs optimally}, which corresponds to maximizing the
+value of partial information \cite{Russell.aima}. Therefore, an
+improved allocation scheme would
+\begin{itemize}
+\item maximize the value of partial
+information by sampling actions to minimize the \textbf{simple} regret at the
+current root node, and then 
+\item minimize the \textbf{accumulated} regret of the playouts from the second
+  step on to approach the optimum behavior. 
+\end{itemize}
 
@@
 
+The $\varepsilon$-greedy allocation scheme has exponential convergence
+rate of the simple regret $\IE r$ (see Appendix~\ref{app:derivations-finite-eps}):
+\begin{equation}
+  \IE r_\varepsilon\le\sum_{j=1}^K\Delta_j\left(\frac {n\varepsilon} K + \frac {4\sqrt e}
+{\Delta_j^2}\right)e^{-\Delta_j^2n\varepsilon/8K}
+\end{equation}
+while UCB exhibits polynomial convergence
+rate (see Appendix~\ref{app:derivations-finite-ucb}):
+\begin{equation}
+\IE r_{ucb} \le 2\sum_{j=1}^K \Delta_jn^{\frac {-\alpha \Delta_j^2} 8}
+\end{equation}
+The best $\varepsilon$ is $\frac 1 2$, and for a large number of arms
+the upper regret bound for the 1/2-greedy scheme approaches the square of the
+upper regret bound for uniform sampling (see
+Appendix~\ref{app:derivations-asym-eps}).
+
+Another improved allocation scheme (UQB) for minimizing the simple
+regret is obtained by substituting $\sqrt{\cdot}$ instead of $\log(\cdot)$ into
+the formula for UCB and has a superpolynomial convergence rate
+(see Appendix~\ref{app:derivations-asym-ucb}).
+
 \subsection{Doing better than UCT}
+
+By extension, a Monte Carlo tree search sampling scheme that selects
+an action at the current root node according to an algorithm that
+minimizes the simple regret, such as $1/2-greedy$ or $UQB$, and then
+selects actions according to $UCB$, outperforms the UCT sampling
+scheme that selects actions based on UCB only.
+
+Such sampling scheme is significantly less sensitive to the tuning of
+the empirical upper bound ($C_p$ in \cite{Kocsis.uct}) of UCT, since
+the conflict \cite{Bubeck.pure} between the need for a larger value of $C_p$ on the first
+step (simple regret) and a smaller value for the rest of the playout
+(accumulated regret) is resolved. In fact, a sampling scheme that uses
+UCB at all steps but a larger value of $C_p$ for the first step than
+for the rest of the steps, outperforms UCT.
 
@@
 
+The results were empirically verified on Multi-armed Bandit instances
+(Section~\ref{seq:emp-mab}), on search trees
+(Section~\ref{seq:emp-mcts}), and on the sailing domain
+(Section~\ref{seq:emp-sailing}), as defined in \cite{Kocsis.uct}. The
+experiments confirmed the hypotheses.
+
 \subsection{Simple regret in multi-armed bandits}
+\label{seq:emp-mab}
 
@@
 \subsection{Monte Carlo tree search}
+\label{seq:emp-mcts}
 
@@
 \subsection{The sailing domain}
+\label{seq:emp-sailing}
 
@@
 \end{figure}
-%
-%\begin{figure}[ht]
-%  \begin{minipage}[b]{0.5\linewidth} \centering
-%    \includegraphics[scale=0.45,trim=0pt 0pt 0pt
-%    0pt,clip]{sailing-size=13-factor=_5625.pdf}
-%    {a. varying number of samples}
-%  \end{minipage}
-%  \begin{minipage}[b]{0.5\linewidth} \centering
-%    \includegraphics[scale=0.45,trim=0pt 0pt 0pt
-%    0pt,clip]{sailing-size=13-nsamples=240.pdf}
-%    {b. varying exploration factor}
-%  \end{minipage}
-%  \caption{The sailing domain, a larger lake}
-%  \label{fig:sailing-13}
-%\end{figure}
 
-\section{Summary and Future Work}
+%\section{Summary and Future Work}
 
@@
 \section{Derivations}
+\label{app:derivations}
 
 \subsection{Finite-time regret bounds}
+\label{app:derivations-finite}
 
@@
 \subsubsection{$\varepsilon$-greedy}
+\label{app:derivations-finite-eps}
 
@@
 
-Number of times $\mu$ the $j$th arm is selected {\it randomly}:
+Number of times $\xi$ the $j$th arm is selected {\it randomly}:
 
 \begin{equation}
-\mu\triangleq\IE\left[T_j^R(n)\right]=\frac {n\varepsilon} K
+\xi\triangleq\IE\left[T_j^R(n)\right]=\frac {n\varepsilon} K
 \end{equation}
 
-By choosing $x_0=\frac \mu 2$ and applying the Chernoff bound (\ref{eq:chernoff-bound}), obtain:
+By choosing $x_0=\frac \xi 2$ and applying the Chernoff bound (\ref{eq:chernoff-bound}), obtain:
 
@@
 \label{eq:pr-epsilon-greedy}
-\Pr(\overline X_j=\max_i\overline X_i)&\le& 2\left(\frac {\mu}{2} e^{-\mu/8} + \frac {2e^{\Delta_j^2/2}}{\Delta_j^2}e^{-\Delta_j^2 \mu/4}\right)\nonumber\\
-&\le&\left(\mu + \frac {4\sqrt e}{\Delta_j^2}\right)e^{-\Delta_j^2\mu/8}
+\Pr(\overline X_j=\max_i\overline X_i)&\le& 2\left(\frac {\xi}{2} e^{-\xi/8} + \frac {2e^{\Delta_j^2/2}}{\Delta_j^2}e^{-\Delta_j^2 \xi/4}\right)\nonumber\\
+&\le&\left(\xi + \frac {4\sqrt e}{\Delta_j^2}\right)e^{-\Delta_j^2\xi/8}
 \end{eqnarray}
@@
 \begin{eqnarray}
-  \IE r_\varepsilon&\le&\sum_{j=1}^K\Delta_j\left(\mu + \frac {4\sqrt e}
-{\Delta_j^2}\right)e^{-\Delta_j^2\mu/8}\nonumber\\
-&\le&K\left(\mu + \frac {4\sqrt
-    e}{\Delta_{\min}}\right)e^{-\Delta_{\min}^2\mu/8}
+  \IE r_\varepsilon&\le&\sum_{j=1}^K\Delta_j\left(\xi + \frac {4\sqrt e}
+{\Delta_j^2}\right)e^{-\Delta_j^2\xi/8}\nonumber\\
+&\le&K\left(\xi + \frac {4\sqrt
+    e}{\Delta_{\min}}\right)e^{-\Delta_{\min}^2\xi/8}
 \label{eq:deriv-epsilon-greedy-b}
@@
 \subsubsection{UCB($\alpha$)}
+\label{app:derivations-finite-ucb}
 
@@
 \subsection{Asymptotic regret analysis}
+\label{app:derivations-asym}
 
@@
 \subsubsection{$\varepsilon$-greedy}
+\label{app:derivations-asym-eps}
 
@@
 selecting the $j$th arm by the sum of the probabilities that the samle
-mean $\overline x_j$ of the $j$th arm is greater than $\mu_j+\frac
-{\Delta_j} 2$ and the sample mean $\overline x_*$ of the optimal arm is
+mean $\overline X_j$ of the $j$th arm is greater than $\mu_j+\frac
+{\Delta_j} 2$ and the sample mean $\overline X_*$ of the optimal arm is
 less than $\mu_*- \frac {\Delta_j} 2$. Each of the probabilities can
@@
 \subsubsection{UCB($\alpha$)}
+\label{app:derivations-asym-ucb}
 
@@
  \end{equation}
-where $\beta$ is a constant depending on reward bounds.
+where $\beta$ is a constant that depends on reward bounds only.
 Finite time convergence can still be shown similarly to UCB.

### 2011-08-03 paper/rmcts/paper.tex
initial draft version

@@
 Such sampling scheme is significantly less sensitive to the tuning of
-the empirical upper bound ($C_p$ in \cite{Kocsis.uct}) of UCT, since
+the exploration factor ($C_p$ in \cite{Kocsis.uct}) of UCT, since
 the conflict \cite{Bubeck.pure} between the need for a larger value of $C_p$ on the first
@@
 \label{seq:emp-mab}
+
+Figure~\ref{fig:mab-simple-regret} presents a comparison of simple
+regret minimization algorithms for Multi-armed
+bandits. Figure~\ref{fig:mab-simple-regret}.a shows the search tree
+corresponding to a problem instance. Each arm returns a random reward
+drawn from a Bernoulli distribution. The search selects an arm
+and compares the expected reward, unknown to the algorithm during the
+sampling, to the expected reward of the best arm.
+
+Figure~\ref{fig:mab-simple-regret}.b shows the regret
+vs. the number of samples, averages over $10^4$ experiments, for uniform
+random sampling (RND), UCT (same as UCB in this case),
+1/2-greedy (GCT), and QCT (UCT with $\log(\cdot)$ replaced by
+$\sqrt{\cdot}$), for randomly generated instances of 32 arms.
+
+For smaller numbers of samples, 1/2-greedy achieves the best
+performance; for large number of samples, QCT outperforms GCT. QCT is
+better than UCT everywhere except for very small numbers of samples. A
+combination of GCT and QCT dominates UCT over the whole range.
 
@@
 \label{seq:emp-mcts}
+
+The second set of experiments was performed on randomly generated
+trees crafted in such a way that uniform random sampling selects a
+direction at the root randomly. The degree of the root is a parameter of
+the tree generator. The degree of all nodes at distance 1 from the
+root is 2, and all nodes at distance 2 from the roots are leaves. The
+average reward of two children of each node at distance 1 is
+0.5. Thus, a uniform sampling scheme results in the same average reward for
+all edges at the root, and an adaptive sampling scheme, such as UCT,
+has to be used.
+
+Figure~\ref{fig:mcts-regret} shows a sketch of the search tree
+(Figure~\ref{fig:mcts-regret}.a) and the dependency of the regret vs. the
+number of samples for trees with root degree 16
+(Figure~\ref{fig:mcts-regret}.b) and 64 (Figure~\ref{fig:mcts-regret}.c). The
+dependencies look differently from Multi-armed bandit instances, but
+the algorithms exhibit a similar relative performance: GCT and QCT
+simultaneously are better than UCT, QCT dominates UCT everywhere
+except for small numbers of instances. The advantage of QCT grows with
+the number of arms.
 
@@
 
+Figures~\ref{fig:sailing-cost-vs-nsamples}--\ref{fig:sailing-rcq-vs-factor}
+show results of experiments on the sailing
+domain. Figure~\ref{fig:sailing-cost-vs-nsamples} shows the regret
+vs. the number of samples, computed for a range of values of
+$C_p$. Figure~\ref{fig:sailing-cost-vs-nsamples}.a shows the median
+cost, and Figure~\ref{fig:sailing-cost-vs-nsamples}.b --- the minimum
+costs. UCT is always worse than either 1/2-greedy (GCT) or QCT, and is sensitive to
+the value of $C_p$: the median cost is much higher than the minimum
+cost for UCT. For both 1/2-greedy and QCT, the difference is
+significantly less prominent.
+
 \begin{figure}[ht]
@@
 \end{figure}
+
+Figure~\ref{fig:sailing-cost-vs-factor} shows the rerget vs. the
+factor for different numbers of samples. QCT is always better than
+UCT, and 1/2-greedy is better than UCT expect for a small range of
+values of the exploration factor $C_p$.
 
@@
 
+Figure~\ref{fig:sailing-lake-size} shows the cost vs. the exploration
+factor for lakes of different sizes. The relative difference between
+the sampling schemes becomes more prominent when the lake size
+increases.
+
 \begin{figure}[ht]
@@
 \end{figure}
+
+Figure~\ref{fig:sailing-rcq-vs-factor} compares QCT with UCT with a
+different exploration factor at the root (CCT). $C_p$ for the rest of
+the steps was chosen to ensure the best performance from earlier
+experiments on the domain. Both algorithms exhibit similar dependency,
+but as the number of samples grows, QCT achieves smaller average
+regrets and is less sensitive to the choice of the value for $C_p$ at
+the first step.
 

### 2011-08-10 paper/rmcts/paper.tex
fixed formula 26
uvb
rmcts paper: mdp definition,algorithm

@@
 
-Such strategies are used to select the best arm, and, by
-extension, the best action in search, when every action in the
-current world state is viewed as a different arm, and the best arm
-must be chosen. Algorithm UCB that maximizes the accumulated reward
-\cite{Auer.ucb} had been extended to tree search as UCT \cite{Kocsis.uct}.
+Such strategies are used to select the best arm, and, by extension,
+the best action in Monte Carlo tree search. Monte Carlo tree search is
+used to solve Markov Decision Processes (MDP) approximately. An MDP is
+defined by the set of states $S$, the set of actions $A$, the
+transition table $T(s, a, s')$, the reward table $R(s, a, s')$, the
+initial state $s$ and the goal state $t$: $(S, A, T, R, s, t)$.  Monte
+Carlo tree search explores an MDP by \emph{rollouts}---trajectories
+from the current state to a state in which a termination condition is
+satisfied (either the goal state, or a cutoff state for which the
+reward is evaluated approximately). Algorithm UCB that maximizes the
+accumulated reward \cite{Auer.ucb} had been extended to tree search as
+UCT \cite{Kocsis.uct}.
 
 UCT-driven search achieves good results in many search problems. At each
-search step, the algorithm allocates Monte Carlo playouts to choose
+search step, the algorithm allocates Monte Carlo rollouts to choose
 the best action, thereby minimizing the simple regret. However, the
@@
 current root node, and then 
-\item minimize the \textbf{accumulated} regret of the playouts from the second
+\item minimize the \textbf{accumulated} regret of the rollouts from the second
   step on to approach the optimum behavior. 
@@
 The best $\varepsilon$ is $\frac 1 2$, and for a large number of arms
-the upper regret bound for the 1/2-greedy scheme approaches the square of the
+the upper regret bound for the $\frac 1 2$-greedy scheme approaches the square of the
 upper regret bound for uniform sampling (see
@@
 
-By extension, a Monte Carlo tree search sampling scheme that selects
-an action at the current root node according to an algorithm that
-minimizes the simple regret, such as $1/2-greedy$ or $UQB$, and then
-selects actions according to $UCB$, outperforms the UCT sampling
-scheme that selects actions based on UCB only.
+The {\bf two-stage} Monte Carlo tree search sampling
+scheme selects an action at the current root node according to an
+algorithm that minimizes the simple regret, such as $\frac 1 2$-greedy or
+UQB, and then selects actions according to UCB. Such two-stage
+outperforms the UCT sampling scheme that selects actions based on
+UCB only.
 
 Such sampling scheme is significantly less sensitive to the tuning of
-the exploration factor ($C_p$ in \cite{Kocsis.uct}) of UCT, since
-the conflict \cite{Bubeck.pure} between the need for a larger value of $C_p$ on the first
-step (simple regret) and a smaller value for the rest of the playout
-(accumulated regret) is resolved. In fact, a sampling scheme that uses
-UCB at all steps but a larger value of $C_p$ for the first step than
-for the rest of the steps, outperforms UCT.
+the exploration factor ($C_p$ in \cite{Kocsis.uct}) of UCT, since the
+conflict \cite{Bubeck.pure} between the need for a larger value of
+$C_p$ on the first step (simple regret) and a smaller value for the
+rest of the rollout (accumulated regret) is resolved. In fact, a
+sampling scheme that uses UCB at all steps but a larger value of $C_p$
+for the first step than for the rest of the steps, outperforms
+UCT. The pseudocode of the two-stage rollout is in
+Algorithm~\ref{alg:two-stage-mcts}.
+
+\begin{algorithm}[t]
+\caption{Two-stage Monte-Carlo tree search sampling}
+\label{alg:two-stage-mcts}
+\begin{algorithmic}[1]
+\Procedure{Rollout}{node}
+  \State $depth \gets 0$
+  \State $reward \gets 0$
+  \While {$\lnot leaf(node)$}
+    \If {$depth=1$}
+      \State $action \gets$\Call{FirstAction}{node} \Comment minimizes simple regret, e.g. $\frac 1 2$-greedy
+    \Else
+      \State $action \gets$\Call{NextAction}{node} \Comment  UCB --- minimizes accumulated regret
+    \EndIf
+    \State $node' \gets$\Call{NextState}{node, action}
+    \State $reward \gets reward+$\Call{TransitionReward}{node, action, node'}
+    \State $node \gets node'$
+    \State $depth \gets depth+1$
+  \EndWhile
+  \State \textbf{return} $reward$
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
 
@@
 random sampling (RND), UCT (same as UCB in this case),
-1/2-greedy (GCT), and QCT (UCT with $\log(\cdot)$ replaced by
+$\frac 1 2$-greedy (GCT), and QCT (UCT with $\log(\cdot)$ replaced by
 $\sqrt{\cdot}$), for randomly generated instances of 32 arms.
 
-For smaller numbers of samples, 1/2-greedy achieves the best
+For smaller numbers of samples, $\frac 1 2$-greedy achieves the best
 performance; for large number of samples, QCT outperforms GCT. QCT is
@@
 
-\begin{figure}[ht]
+\begin{figure}[t]
   \begin{minipage}[c]{0.5\linewidth}
@@
 cost, and Figure~\ref{fig:sailing-cost-vs-nsamples}.b --- the minimum
-costs. UCT is always worse than either 1/2-greedy (GCT) or QCT, and is sensitive to
+costs. UCT is always worse than either $\frac 1 2$-greedy (GCT) or QCT, and is sensitive to
 the value of $C_p$: the median cost is much higher than the minimum
-cost for UCT. For both 1/2-greedy and QCT, the difference is
+cost for UCT. For both $\frac 1 2$-greedy and QCT, the difference is
 significantly less prominent.
 
-\begin{figure}[ht]
+\begin{figure}[t]
   \begin{minipage}[b]{0.5\linewidth}
@@
   \end{minipage}
-  \caption{The sailing domain, $6\times 6$ lake, cost vs. number of playouts}
+  \caption{The sailing domain, $6\times 6$ lake, cost vs. number of rollouts}
   \label{fig:sailing-cost-vs-nsamples}
@@
 factor for different numbers of samples. QCT is always better than
-UCT, and 1/2-greedy is better than UCT expect for a small range of
+UCT, and $\frac 1 2$-greedy is better than UCT expect for a small range of
 values of the exploration factor $C_p$.
 
-\begin{figure}[ht]
+\begin{figure}[t]
   \begin{minipage}[b]{0.5\linewidth}
@@
     \includegraphics[scale=0.45]{costs-size=6-nsamples=199.pdf}\\
-    a. 199 playouts\\
+    a. 199 rollouts\\
     \vspace{1em}
     \includegraphics[scale=0.45]{costs-size=6-nsamples=397.pdf}\\
-    b. 397 playouts\\
+    b. 397 rollouts\\
   \end{minipage}
@@
     \includegraphics[scale=0.45]{costs-size=6-nsamples=793.pdf}\\
-    c. 793 playouts\\
+    c. 793 rollouts\\
     \vspace{1em}
     \includegraphics[scale=0.45]{costs-size=6-nsamples=1585.pdf}\\
-    d. 1585 playouts\\
+    d. 1585 rollouts\\
   \end{minipage}
@@
 
-\begin{figure}[ht]
+\begin{figure}[t]
   \begin{minipage}[b]{0.333\linewidth}
@@
 
-\begin{figure}[ht]
+\begin{figure}[t]
   \begin{minipage}[b]{0.5\linewidth}
@@
     \includegraphics[scale=0.45]{rcq-size=6-nsamples=199.pdf}\\
-    a. 199 playouts\\
+    a. 199 rollouts\\
     \vspace{1em}
     \includegraphics[scale=0.45]{rcq-size=6-nsamples=397.pdf}\\
-    b. 397 playouts\\
+    b. 397 rollouts\\
   \end{minipage}
@@
     \includegraphics[scale=0.45]{rcq-size=6-nsamples=793.pdf}\\
-    c. 793 playouts\\
+    c. 793 rollouts\\
     \vspace{1em}
     \includegraphics[scale=0.45]{rcq-size=6-nsamples=1585.pdf}\\
-    d. 1585 playouts\\
+    d. 1585 rollouts\\
   \end{minipage}
@@
 
-%\section{Summary and Future Work}
+\section{Summary and Future Work}
+
+The Monte Carlo tree search algorithms presented in the paper differ
+from UCT at the first step of the rollout, when the `simple' selection
+regret is minimized instead of the accumulated regret. Both the
+theoretical analysis and the empirical evaluation provide evidence for
+better general performance of the proposed algorithms.
+
+The improvement is based on the notion of value of information (VOI),
+but VOI is used implicitly in the analysis of the algorithm, rather
+than computed or learned explicitly in order to plan the rollouts. A
+further improvement can be achieved by computing or estimating the VOI
+of the rollouts and chosing a rollout that maximizes the VOI. Using VOI
+to guide Monte Carlo sampling instead of relying on the sample means
+of different actions  is particularly beneficial when different
+actions have qualitatively different distributions: for example, when the
+distribution variance varies signficantly between the actions.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[scale=0.45,trim=0 0 0 16pt,clip]{flat-trilevel-vct-k=64.pdf}\\
+  \caption{VOI-based action selection (VCT) vs. other algorithms}
+  \label{fig:flat-trilevel-vct}
+\end{figure}
+
+VOI of a rollout can be computed when the sample distribution of an
+action is known up to the parameters, such as the normal distribution
+with an unknown mean and/or variance. Alternatively, the value of
+information can be estimated from the set of samples, and the need to
+assume a particular shape of the distribution is be lifted. In one
+realization of the latter approach the VOI of performing an action is
+estimated as the value of perfect information learned from the outcome
+of earlier samples of the action divided by the number of
+samples. Early experiments with this approach showed the best
+performance for sets of Bernoulli arms, and even a more prominent
+improvement for mixed arms (e.g. Bernoulli, triangularly distributed,
+and fixed arms) (Figure~\ref{fig:flat-trilevel-vct}).  Sample-based
+VOI estimation requires keeping more information than an algorithm
+based only on the sample mean, but the overhead is small and justifies
+the improvement.
 
@@
 
-\bibliographystyle{plain}
-\bibliography{refs}
-
-\pagebreak
-
+\clearpage
 \appendix
@@
 
-{\bf Chernoff bound (see \cite{Hagerup.chernoff}):} for $m$ independent random variables $X_1, X_2, ..., X_m$
-taking values 0 or 1, $X=\sum_{i=1}^m X_i$, and $0\le\delta\le 1$:
-
+{\bf Chernoff bounds (see \cite{Hagerup.chernoff}):} for $m$ independent random variables $X_1, X_2, ..., X_m$
+taking on values 0 or 1, $X=\sum_{i=1}^m X_i$, and $0\le\delta\le 1$:
 \begin{equation}
+\Pr[X < (1-\delta)\IE[X]] < e^{-\delta^2\IE[X]/2} \mbox{ (Chernoff bound)}
 \label{eq:chernoff-bound}
-\Pr[X < (1-\delta)\IE[X]] < e^{-\delta^2\IE[X]/2}
+\end{equation}
+Further on, if $\forall i. \IE[X_i]=\mu$:
+\begin{equation}
+\Pr[X < n\mu-a] <  e^{-2a^2/n}
+\label{eq:chernoff-hoeffding-bound} \mbox{ (Hoeffding-Chernoff bound)}
 \end{equation}
@@
 \begin{equation}
-P_j^{uqb}=\mathbb{B}P_j^{uqb}\approx2\exp\left(-\beta\Delta_j^2\frac {\sqrt n}{\Delta_j^2}\right)
-            =2\exp(-2\beta\sqrt n)
+P_j^{uqb}\le\mathbb{B}P_j^{uqb}\approx2\exp\left(-\beta\Delta_j^2\frac {\sqrt n}{\Delta_j^2}\right)
+            =2\exp(-\beta\sqrt n)
  \end{equation}
@@
 
+\bibliographystyle{plain}
+\bibliography{refs}
+
 \end{document}

### 2011-08-13 paper/rmcts/paper.tex
algorithm fixed

@@
 transition table $T(s, a, s')$, the reward table $R(s, a, s')$, the
-initial state $s$ and the goal state $t$: $(S, A, T, R, s, t)$.  Monte
-Carlo tree search explores an MDP by \emph{rollouts}---trajectories
-from the current state to a state in which a termination condition is
-satisfied (either the goal state, or a cutoff state for which the
-reward is evaluated approximately). Algorithm UCB that maximizes the
-accumulated reward \cite{Auer.ucb} had been extended to tree search as
-UCT \cite{Kocsis.uct}.
+initial state $s$ and the goal state $t$: $(S, A, T, R, s, t)$
+\cite{Russell.aima}.  Monte Carlo tree search explores an MDP by
+\emph{rollouts}---trajectories from the current state to a state in
+which a termination condition is satisfied (either the goal state, or
+a cutoff state for which the reward is evaluated
+approximately). Algorithm UCB that maximizes the accumulated reward
+\cite{Auer.ucb} had been extended to tree search as UCT
+\cite{Kocsis.uct}.
 
@@
 \begin{algorithmic}[1]
-\Procedure{Rollout}{node}
-  \State $depth \gets 0$
-  \State $reward \gets 0$
-  \While {$\lnot leaf(node)$}
-    \If {$depth=1$}
-      \State $action \gets$\Call{FirstAction}{node} \Comment minimizes simple regret, e.g. $\frac 1 2$-greedy
+\Procedure{Rollout}{node, depth=1}
+  \If {\Call{IsLeaf}{node, depth}}
+    \State \textbf{return} 0
   \Else
-      \State $action \gets$\Call{NextAction}{node} \Comment  UCB --- minimizes accumulated regret
+    \If {depth=1}
+      \State action $\gets$ \Call{FirstAction}{node} \Comment minimizes simple regret, e.g. $\frac 1 2$-greedy
+    \Else
+      \State action $\gets$ \Call{NextAction}{node} \Comment  UCB --- minimizes accumulated regret
     \EndIf
-    \State $node' \gets$\Call{NextState}{node, action}
-    \State $reward \gets reward+$\Call{TransitionReward}{node, action, node'}
-    \State $node \gets node'$
-    \State $depth \gets depth+1$
-  \EndWhile
-  \State \textbf{return} $reward$
+    \State next-node $\gets$ \Call{NextState}{node, action}
+    \State reward $\gets$ \Call{TransitionReward}{node, action, next-node}
+    \State \hspace{4em} + \Call{Rollout}{next-node, depth+1} \Comment call {\sc Rollout} recursively
+    \State \Call{UpdateStats}{node, action, reward}
+  \EndIf
 \EndProcedure
@@
 
-The improvement is based on the notion of value of information (VOI),
+The improvement is inspired by the notion of value of information (VOI),
 but VOI is used implicitly in the analysis of the algorithm, rather
@@
 That is, the case $\frac {\partial P_j} {\partial \varepsilon}=0$
-corresponds to \emph{0.5-greedy sampling}, and for a large number of
+corresponds to \emph{$\frac 1 2$-greedy sampling}, and for a large number of
 arms $K$ 
@@
 Equation (\ref{eq:derive-uniform-vs-half-greedy}) suggests that
-0.5-greedy sampling has a much higher convergence rate than uniform
+$\frac 1 2$-greedy sampling has a much higher convergence rate than uniform
 sampling for a large number of arms.

### 2011-08-17 paper/rmcts/paper.tex
some fixes
some fixes after eyals notes

@@
 pulls, brings a reward, that is, the simple regret, as opposite to the
-accumulated regret, must be minimized. This work introduces policies for
+cumulative regret, must be minimized. This work introduces policies for
 multi-armed bandits with lower simple regret than UCB, and an
-algorithm for MCTS which combines accumulated and simple regret
+algorithm for MCTS which combines cumulative and simple regret
 minization and outperforms UCT. Finite-time and asymptotic analysis of
@@
 a cutoff state for which the reward is evaluated
-approximately). Algorithm UCB that maximizes the accumulated reward
+approximately). Algorithm UCB that minimizes the cumulative regret
 \cite{Auer.ucb} had been extended to tree search as UCT
@@
 search step, the algorithm allocates Monte Carlo rollouts to choose
-the best action, thereby minimizing the simple regret. However, the
-simple and the accumulated regret cannot be minimized simultaneously;
+the best action, thereby minimizing the cumulative regret. However, the
+simple and the cumulative regret cannot be minimized simultaneously;
 moreover, \cite{Bubeck.pure} shows that in many cases the smaller the
-accumulated regret, the greater is the simple regret.
+cumulative regret, the greater the simple regret.
 
@@
 current root node. Indeed, the algorithm must select an action with
-the minimum regret \emph{given assumption that after performing the selected action
+the minimum regret \emph{given the assumption that after performing the selected action
 the algorithm performs optimally}, which corresponds to maximizing the
@@
 current root node, and then 
-\item minimize the \textbf{accumulated} regret of the rollouts from the second
+\item minimize the \textbf{cumulative} regret of the rollouts from the second
   step on to approach the optimum behavior. 
@@
 $C_p$ on the first step (simple regret) and a smaller value for the
-rest of the rollout (accumulated regret) is resolved. In fact, a
+rest of the rollout (cumulative regret) is resolved. In fact, a
 sampling scheme that uses UCB at all steps but a larger value of $C_p$
@@
     \Else
-      \State action $\gets$ \Call{NextAction}{node} \Comment  UCB --- minimizes accumulated regret
+      \State action $\gets$ \Call{NextAction}{node} \Comment  UCB --- minimizes cumulative regret
     \EndIf
@@
 Figure~\ref{fig:mab-simple-regret}.b shows the regret
-vs. the number of samples, averages over $10^4$ experiments, for uniform
+vs. the number of samples, averaged over $10^4$ experiments, for uniform
 random sampling (RND), UCT (same as UCB in this case),
@@
 dependencies look differently from Multi-armed bandit instances, but
-the algorithms exhibit a similar relative performance: GCT and QCT
-simultaneously are better than UCT, QCT dominates UCT everywhere
+the algorithms exhibit a similar relative performance: either GCT or QCT
+is always better than UCT, QCT dominates UCT everywhere
 except for small numbers of instances. The advantage of QCT grows with
@@
 Figure~\ref{fig:sailing-cost-vs-factor} shows the rerget vs. the
-factor for different numbers of samples. QCT is always better than
+exploration factor for different numbers of samples. QCT is always better than
 UCT, and $\frac 1 2$-greedy is better than UCT expect for a small range of
@@
 from UCT at the first step of the rollout, when the `simple' selection
-regret is minimized instead of the accumulated regret. Both the
+regret is minimized instead of the cumulative regret. Both the
 theoretical analysis and the empirical evaluation provide evidence for
@@
 \label{app:derivations-finite}
+
+%TODO: bounds on regret based on information gathered, independent of
+%the way the choice is made.
 
@@
 
+%TODO: Oracle instead of asymptotic
+
 Bounds (\ref{eq:deriv-epsilon-greedy-b}, \ref{eq:deriv-ucb-alpha-b}) were
@@
 of non-optimal arms. Derivation of better bounds seems
-complicated. Instead, one can analyse symptotic behavior of each of the
-algorithms in assumption that the sample mean is close to the expectation.
+complicated. Instead, one can analyse the asymptotic behavior of each of the
+algorithms under the assumption that the sample mean is close to the expectation.
 
@@
  \end{equation}
-where $\beta$ is a constant that depends on reward bounds only.
+where $\beta$ is a constant that depends only on reward bounds.
 Finite time convergence can still be shown similarly to UCB.

### 2011-08-17 paper/rmcts/paper.tex
vct with uqb

@@
 
-Another improved allocation scheme (UQB) for minimizing the simple
-regret is obtained by substituting $\sqrt{\cdot}$ instead of $\log(\cdot)$ into
-the formula for UCB and has a superpolynomial convergence rate
-(see Appendix~\ref{app:derivations-asym-ucb}).
+Another improved allocation scheme for minimizing the simple regret,
+called \textbf{UQB} here, is obtained by substituting $\sqrt{\cdot}$
+instead of $\log(\cdot)$ into the formula for UCB and has a
+superpolynomial convergence rate (see
+Appendix~\ref{app:derivations-asym-ucb}).
 
@@
 improvement for mixed arms (e.g. Bernoulli, triangularly distributed,
-and fixed arms) (Figure~\ref{fig:flat-trilevel-vct}).  Sample-based
-VOI estimation requires keeping more information than an algorithm
-based only on the sample mean, but the overhead is small and justifies
+and fixed arms) (Figure~\ref{fig:flat-trilevel-vct}):
+\textbf{VCT}, the algorithm which selects an action with the maximum VOI estimate on
+the first step exhibits lower average regret than other algorithms.
+Sample-based VOI estimation requires keeping more information than an
+algorithm based only on the sample mean, but the overhead is small and justifies
 the improvement.

### 2011-08-17 paper/rmcts/paper.tex
a more formal statement of the results

@@
 
-The $\varepsilon$-greedy allocation scheme has exponential convergence
+\begin{enumerate}
+\item The $\varepsilon$-greedy allocation scheme has exponential convergence
 rate of the simple regret $\IE r$ (see Appendix~\ref{app:derivations-finite-eps}):
@@
 \end{equation}
-while UCB exhibits polynomial convergence
+\item The UCB allocation scheme exhibits polynomial convergence
 rate (see Appendix~\ref{app:derivations-finite-ucb}):
@@
 \end{equation}
-The best $\varepsilon$ is $\frac 1 2$, and for a large number of arms
-the upper regret bound for the $\frac 1 2$-greedy scheme approaches the square of the
-upper regret bound for uniform sampling (see
+\item A tighter upper regret bound for the $\varepsilon$-greedy allocation scheme depends on
+$\varepsilon$. The lowest bound is
+acheived for $\varepsilon=\frac 1
+2$, and for a large number of arms the bound for the $\frac 1 2$-greedy scheme approaches the square of the
+bound for uniform sampling (see
 Appendix~\ref{app:derivations-asym-eps}).
+\end{enumerate}
 
@@
 (Section~\ref{seq:emp-sailing}), as defined in \cite{Kocsis.uct}. The
-experiments confirmed the hypotheses.
+experiments confirmed the hypotheses. The algorithms used in the
+experiments are:
+\begin{description}
+\item[RND:] uniform \underline{r}a\underline{nd}om samping;
+\item[UCT:] \underline{U}pper \underline{C}onfidence Bounds applied to
+  \underline{T}rees \cite{Kocsis.uct};
+\item[GCT:] U\underline{CT} with the first step
+  replaced with $\frac 1 2$-\underline{g}reedy sampling;
+\item[QCT:] U\underline{CT} with the first step
+  replaced with U\underline{Q}B.
+\end{description}
 
@@
 Figure~\ref{fig:mab-simple-regret}.b shows the regret
-vs. the number of samples, averaged over $10^4$ experiments, for uniform
-random sampling (RND), UCT (same as UCB in this case),
-$\frac 1 2$-greedy (GCT), and QCT (UCT with $\log(\cdot)$ replaced by
-$\sqrt{\cdot}$), for randomly generated instances of 32 arms.
+vs. the number of samples, averaged
+over $10^4$ experiments for randomly generated instances of 32 arms.
 
@@
 realization of the latter approach the VOI of performing an action is
-estimated as the value of perfect information learned from the outcome
+estimated as \emph{the value of perfect information learned from the outcome
 of earlier samples of the action divided by the number of
-samples. Early experiments with this approach showed the best
+samples}. Early experiments with this approach showed the best
 performance for sets of Bernoulli arms, and even a more prominent
@@
 Sample-based VOI estimation requires keeping more information than an
-algorithm based only on the sample mean, but the overhead is small and justifies
-the improvement.
+algorithm based only on the sample
+mean, but the overhead is small and
+is justified by the improvement.
 
@@
 $\mu_j$ and $\mu_*$ at the middle and bound the probability of
-selecting the $j$th arm by the sum of the probabilities that the samle
+selecting the $j$th arm by the sum of the probabilities that the sample
 mean $\overline X_j$ of the $j$th arm is greater than $\mu_j+\frac
@@
 Split the interval $[\mu_j, \mu_*]$ at $\mu_j+\delta_j$. The
-probability $P_j$ for selecting the $j$th arm is bounded as:
+probability $P_j$ for selecting the
+$j$th arm is bounded, using the union bound and the Chernoff-Hoeffding bound, as:
 
-\begin{equation}
-P_j\le exp\left(-2\delta_j^2n_j\right)+exp\left(-2(\Delta_j-\delta_j)^2n_*\right)
-\end{equation}
+\begin{eqnarray}
+P_j&\le&\Pr[\overline X_j>\mu_j+\delta_j]+\Pr[\overline X_*<\mu_*-(\Delta_j-\delta_j)]\nonumber\\
+   &\le&\exp\left(-2\delta_j^2n_j\right)+\exp\left(-2(\Delta_j-\delta_j)^2n_*\right)
+\end{eqnarray}
 

### 2011-08-17 paper/rmcts/paper.tex
formal statement of results

@@
 Appendix~\ref{app:derivations-asym-eps}).
-\end{enumerate}
 
-Another improved allocation scheme for minimizing the simple regret,
+\item Another improved allocation scheme for minimizing the simple regret,
 called \textbf{UQB} here, is obtained by substituting $\sqrt{\cdot}$
@@
 Appendix~\ref{app:derivations-asym-ucb}).
+\end{enumerate}
 
@@
 \end{equation}
-Equation (\ref{eq:derive-uniform-vs-half-greedy}) suggests that
-$\frac 1 2$-greedy sampling has a much higher convergence rate than uniform
-sampling for a large number of arms.
+Equation (\ref{eq:derive-uniform-vs-half-greedy}) suggests
+that \emph{for a large number of arms} $\frac 1 2$-greedy
+sampling has a much higher convergence rate than uniform
+sampling.
 

### 2011-09-13 paper/voisam/paper.tex
added bibliography references

--- 2011-08-17 paper/rmcts/paper.tex
@@
 
-\title{Doing Better Than UCT: \\ Rational Monte Carlo Sampling in Trees}
+\title{VOI-based Monte Carlo Sampling}
 \author {David Tolpin, Solomon Eyal Shimony \\
@@
 \begin{abstract}
-UCT, a state-of-the art algorithm for Monte Carlo tree sampling
-(MCTS), is based on UCB, a sampling policy for the Multi-armed Bandit
-Problem (MAB) that minimizes the accumulated regret. However, MCTS
-differs from MAB in that only the final choice, rather than all arm
-pulls, brings a reward, that is, the simple regret, as opposite to the
-cumulative regret, must be minimized. This work introduces policies for
-multi-armed bandits with lower simple regret than UCB, and an
-algorithm for MCTS which combines cumulative and simple regret
-minization and outperforms UCT. Finite-time and asymptotic analysis of
-the policies is provided, and the algorithms are empirically compared.
 \end{abstract}
 
-\section{Introduction and definitions}
-
-The Multi-armed Bandit is a set of $K$ arms. Each arm can be pulled
-multiple times. When the $i$th arm is pulled, a random reward $X_i$
-from an unknown stationary distribution is returned.  The reward is 
-bounded between 0 and 1.
-
-The simple regret of a sampling policy for the Multi-armed Bandit
-Problem is the expected difference between the best expected reward
-$\mu_*$ and the expected reward $\mu_j$ of the arm with the best sample mean
-$\overline X_j=\max_i\overline X_i$:
-\begin{equation}
-\label{eq:simple-regret}
-\IE[R]=\sum_{j=1}^K\Delta_j\Pr(\overline X_j=\max_i\overline X_i)
-\end{equation}
-where $\Delta_j=\mu_*-\mu_j$.
-Strategies that minimize the simple reward are called pure exploration
-strategies \cite{Bubeck.pure}.
-
-Such strategies are used to select the best arm, and, by extension,
-the best action in Monte Carlo tree search. Monte Carlo tree search is
-used to solve Markov Decision Processes (MDP) approximately. An MDP is
-defined by the set of states $S$, the set of actions $A$, the
-transition table $T(s, a, s')$, the reward table $R(s, a, s')$, the
-initial state $s$ and the goal state $t$: $(S, A, T, R, s, t)$
-\cite{Russell.aima}.  Monte Carlo tree search explores an MDP by
-\emph{rollouts}---trajectories from the current state to a state in
-which a termination condition is satisfied (either the goal state, or
-a cutoff state for which the reward is evaluated
-approximately). Algorithm UCB that minimizes the cumulative regret
-\cite{Auer.ucb} had been extended to tree search as UCT
-\cite{Kocsis.uct}.
-
-UCT-driven search achieves good results in many search problems. At each
-search step, the algorithm allocates Monte Carlo rollouts to choose
-the best action, thereby minimizing the cumulative regret. However, the
-simple and the cumulative regret cannot be minimized simultaneously;
-moreover, \cite{Bubeck.pure} shows that in many cases the smaller the
-cumulative regret, the greater the simple regret.
-
-UCT performance can be improved by combining UCB with a sampling
-scheme that minimizes the simple regret of selecting an action at the
-current root node. Indeed, the algorithm must select an action with
-the minimum regret \emph{given the assumption that after performing the selected action
-the algorithm performs optimally}, which corresponds to maximizing the
-value of partial information \cite{Russell.aima}. Therefore, an
-improved allocation scheme would
-\begin{itemize}
-\item maximize the value of partial
-information by sampling actions to minimize the \textbf{simple} regret at the
-current root node, and then 
-\item minimize the \textbf{cumulative} regret of the rollouts from the second
-  step on to approach the optimum behavior. 
-\end{itemize}
-
-\section{Main Results}
-
-\subsection{Doing better than UCB}
-
-The $\varepsilon$-greedy allocation scheme has exponential convergence
-rate of the simple regret $\IE r$ (see Appendix~\ref{app:derivations-finite-eps}):
-\begin{equation}
-  \IE r_\varepsilon\le\sum_{j=1}^K\Delta_j\left(\frac {n\varepsilon} K + \frac {4\sqrt e}
-{\Delta_j^2}\right)e^{-\Delta_j^2n\varepsilon/8K}
-\end{equation}
-while UCB exhibits polynomial convergence
-rate (see Appendix~\ref{app:derivations-finite-ucb}):
-\begin{equation}
-\IE r_{ucb} \le 2\sum_{j=1}^K \Delta_jn^{\frac {-\alpha \Delta_j^2} 8}
-\end{equation}
-The best $\varepsilon$ is $\frac 1 2$, and for a large number of arms
-the upper regret bound for the $\frac 1 2$-greedy scheme approaches the square of the
-upper regret bound for uniform sampling (see
-Appendix~\ref{app:derivations-asym-eps}).
-
-Another improved allocation scheme for minimizing the simple regret,
-called \textbf{UQB} here, is obtained by substituting $\sqrt{\cdot}$
-instead of $\log(\cdot)$ into the formula for UCB and has a
-superpolynomial convergence rate (see
-Appendix~\ref{app:derivations-asym-ucb}).
-
-\subsection{Doing better than UCT}
-
-The {\bf two-stage} Monte Carlo tree search sampling
-scheme selects an action at the current root node according to an
-algorithm that minimizes the simple regret, such as $\frac 1 2$-greedy or
-UQB, and then selects actions according to UCB. Such two-stage
-outperforms the UCT sampling scheme that selects actions based on
-UCB only.
-
-Such sampling scheme is significantly less sensitive to the tuning of
-the exploration factor ($C_p$ in \cite{Kocsis.uct}) of UCT, since the
-conflict \cite{Bubeck.pure} between the need for a larger value of
-$C_p$ on the first step (simple regret) and a smaller value for the
-rest of the rollout (cumulative regret) is resolved. In fact, a
-sampling scheme that uses UCB at all steps but a larger value of $C_p$
-for the first step than for the rest of the steps, outperforms
-UCT. The pseudocode of the two-stage rollout is in
-Algorithm~\ref{alg:two-stage-mcts}.
-
-\begin{algorithm}[t]
-\caption{Two-stage Monte-Carlo tree search sampling}
-\label{alg:two-stage-mcts}
-\begin{algorithmic}[1]
-\Procedure{Rollout}{node, depth=1}
-  \If {\Call{IsLeaf}{node, depth}}
-    \State \textbf{return} 0
-  \Else
-    \If {depth=1}
-      \State action $\gets$ \Call{FirstAction}{node} \Comment minimizes simple regret, e.g. $\frac 1 2$-greedy
-    \Else
-      \State action $\gets$ \Call{NextAction}{node} \Comment  UCB --- minimizes cumulative regret
-    \EndIf
-    \State next-node $\gets$ \Call{NextState}{node, action}
-    \State reward $\gets$ \Call{TransitionReward}{node, action, next-node}
-    \State \hspace{4em} + \Call{Rollout}{next-node, depth+1} \Comment call {\sc Rollout} recursively
-    \State \Call{UpdateStats}{node, action, reward}
-  \EndIf
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
-
-\section{Empirical Evaluation}
-
-The results were empirically verified on Multi-armed Bandit instances
-(Section~\ref{seq:emp-mab}), on search trees
-(Section~\ref{seq:emp-mcts}), and on the sailing domain
-(Section~\ref{seq:emp-sailing}), as defined in \cite{Kocsis.uct}. The
-experiments confirmed the hypotheses.
-
-\subsection{Simple regret in multi-armed bandits}
-\label{seq:emp-mab}
-
-Figure~\ref{fig:mab-simple-regret} presents a comparison of simple
-regret minimization algorithms for Multi-armed
-bandits. Figure~\ref{fig:mab-simple-regret}.a shows the search tree
-corresponding to a problem instance. Each arm returns a random reward
-drawn from a Bernoulli distribution. The search selects an arm
-and compares the expected reward, unknown to the algorithm during the
-sampling, to the expected reward of the best arm.
-
-Figure~\ref{fig:mab-simple-regret}.b shows the regret
-vs. the number of samples, averaged over $10^4$ experiments, for uniform
-random sampling (RND), UCT (same as UCB in this case),
-$\frac 1 2$-greedy (GCT), and QCT (UCT with $\log(\cdot)$ replaced by
-$\sqrt{\cdot}$), for randomly generated instances of 32 arms.
-
-For smaller numbers of samples, $\frac 1 2$-greedy achieves the best
-performance; for large number of samples, QCT outperforms GCT. QCT is
-better than UCT everywhere except for very small numbers of samples. A
-combination of GCT and QCT dominates UCT over the whole range.
-
-\begin{figure}[t]
-  \begin{minipage}[c]{0.5\linewidth}
-    \centering
-    \includegraphics[scale=1.0]{onelevel-tree.pdf}\\
-    \vspace{4em}
-    a. search tree
-  \end{minipage}
-  \begin{minipage}[c]{0.5\linewidth}
-    \centering
-    \includegraphics[scale=0.5]{flat-trilevel-k=64-uqb=8.pdf}\\
-    b. regret vs. number of samples
-  \end{minipage}
-  \label{fig:mab-simple-regret}
-  \caption{Simple regret in MAB}
-\end{figure}
-
-\subsection{Monte Carlo tree search}
-\label{seq:emp-mcts}
-
-The second set of experiments was performed on randomly generated
-trees crafted in such a way that uniform random sampling selects a
-direction at the root randomly. The degree of the root is a parameter of
-the tree generator. The degree of all nodes at distance 1 from the
-root is 2, and all nodes at distance 2 from the roots are leaves. The
-average reward of two children of each node at distance 1 is
-0.5. Thus, a uniform sampling scheme results in the same average reward for
-all edges at the root, and an adaptive sampling scheme, such as UCT,
-has to be used.
-
-Figure~\ref{fig:mcts-regret} shows a sketch of the search tree
-(Figure~\ref{fig:mcts-regret}.a) and the dependency of the regret vs. the
-number of samples for trees with root degree 16
-(Figure~\ref{fig:mcts-regret}.b) and 64 (Figure~\ref{fig:mcts-regret}.c). The
-dependencies look differently from Multi-armed bandit instances, but
-the algorithms exhibit a similar relative performance: either GCT or QCT
-is always better than UCT, QCT dominates UCT everywhere
-except for small numbers of instances. The advantage of QCT grows with
-the number of arms.
-
-\begin{figure}
-  \begin{minipage}[c]{0.5\linewidth}
-    \centering
-    \includegraphics[scale=0.8]{twolevel-tree.pdf}\\
-    a. search tree
-  \end{minipage}
-  \begin{minipage}[c]{0.5\linewidth}
-    \centering
-    \includegraphics[scale=0.4]{tree-identity-k=16-uqb=8.pdf}\\ 
-    b. 16 arms \\
-    \vspace{1em}
-    \includegraphics[scale=0.4]{tree-identity-k=64-uqb=8.pdf} \\
-    c. 64 arms
- \end{minipage}
-  \label{fig:mcts-regret}
-  \caption{MCTS: a path to the best arm}
-\end{figure}
-
-\subsection{The sailing domain}
-\label{seq:emp-sailing}
-
-Figures~\ref{fig:sailing-cost-vs-nsamples}--\ref{fig:sailing-rcq-vs-factor}
-show results of experiments on the sailing
-domain. Figure~\ref{fig:sailing-cost-vs-nsamples} shows the regret
-vs. the number of samples, computed for a range of values of
-$C_p$. Figure~\ref{fig:sailing-cost-vs-nsamples}.a shows the median
-cost, and Figure~\ref{fig:sailing-cost-vs-nsamples}.b --- the minimum
-costs. UCT is always worse than either $\frac 1 2$-greedy (GCT) or QCT, and is sensitive to
-the value of $C_p$: the median cost is much higher than the minimum
-cost for UCT. For both $\frac 1 2$-greedy and QCT, the difference is
-significantly less prominent.
-
-\begin{figure}[t]
-  \begin{minipage}[b]{0.5\linewidth}
-    \centering
-    \includegraphics[scale=0.45]{costs-size=6-group=median.pdf}\\
-    a. median cost
-  \end{minipage}
-  \begin{minipage}[b]{0.5\linewidth}
-    \centering
-    \includegraphics[scale=0.45]{costs-size=6-group=minimum.pdf}\\
-    b. minimum cost
-  \end{minipage}
-  \caption{The sailing domain, $6\times 6$ lake, cost vs. number of rollouts}
-  \label{fig:sailing-cost-vs-nsamples}
-\end{figure}
-
-Figure~\ref{fig:sailing-cost-vs-factor} shows the rerget vs. the
-exploration factor for different numbers of samples. QCT is always better than
-UCT, and $\frac 1 2$-greedy is better than UCT expect for a small range of
-values of the exploration factor $C_p$.
-
-\begin{figure}[t]
-  \begin{minipage}[b]{0.5\linewidth}
-    \centering
-    \includegraphics[scale=0.45]{costs-size=6-nsamples=199.pdf}\\
-    a. 199 rollouts\\
-    \vspace{1em}
-    \includegraphics[scale=0.45]{costs-size=6-nsamples=397.pdf}\\
-    b. 397 rollouts\\
-  \end{minipage}
-  \begin{minipage}[b]{0.5\linewidth}
-    \centering
-    \includegraphics[scale=0.45]{costs-size=6-nsamples=793.pdf}\\
-    c. 793 rollouts\\
-    \vspace{1em}
-    \includegraphics[scale=0.45]{costs-size=6-nsamples=1585.pdf}\\
-    d. 1585 rollouts\\
-  \end{minipage}
-  \caption{The sailing domain, $6\times 6$ lake, cost vs. factor}
-  \label{fig:sailing-cost-vs-factor}
-\end{figure}
-
-Figure~\ref{fig:sailing-lake-size} shows the cost vs. the exploration
-factor for lakes of different sizes. The relative difference between
-the sampling schemes becomes more prominent when the lake size
-increases.
-
-\begin{figure}[t]
-  \begin{minipage}[b]{0.333\linewidth}
-    \centering
-    \includegraphics[scale=0.35]{costs-size=3-nsamples=397.pdf}\\
-    a. $3\times 3$ lake
-  \end{minipage}
-  \begin{minipage}[b]{0.333\linewidth}
-    \centering
-    \includegraphics[scale=0.35]{costs-size=6-nsamples=397.pdf}\\
-    b. $6\times 6$ lake
-  \end{minipage}
-  \begin{minipage}[b]{0.333\linewidth}
-    \centering
-    \includegraphics[scale=0.35]{costs-size=10-nsamples=397.pdf}\\
-    b. $10\times 10$ lake
-  \end{minipage}
-  \caption{The sailing domain, 397 samples, cost vs. factor}
-  \label{fig:sailing-lake-size}
-\end{figure}
-
-Figure~\ref{fig:sailing-rcq-vs-factor} compares QCT with UCT with a
-different exploration factor at the root (CCT). $C_p$ for the rest of
-the steps was chosen to ensure the best performance from earlier
-experiments on the domain. Both algorithms exhibit similar dependency,
-but as the number of samples grows, QCT achieves smaller average
-regrets and is less sensitive to the choice of the value for $C_p$ at
-the first step.
-
-\begin{figure}[t]
-  \begin{minipage}[b]{0.5\linewidth}
-    \centering
-    \includegraphics[scale=0.45]{rcq-size=6-nsamples=199.pdf}\\
-    a. 199 rollouts\\
-    \vspace{1em}
-    \includegraphics[scale=0.45]{rcq-size=6-nsamples=397.pdf}\\
-    b. 397 rollouts\\
-  \end{minipage}
-  \begin{minipage}[b]{0.5\linewidth}
-    \centering
-    \includegraphics[scale=0.45]{rcq-size=6-nsamples=793.pdf}\\
-    c. 793 rollouts\\
-    \vspace{1em}
-    \includegraphics[scale=0.45]{rcq-size=6-nsamples=1585.pdf}\\
-    d. 1585 rollouts\\
-  \end{minipage}
-  \caption{The sailing domain, log vs. sqrt, $6\times 6$ lake}
-  \label{fig:sailing-rcq-vs-factor}
-\end{figure}
-
-\section{Summary and Future Work}
-
-The Monte Carlo tree search algorithms presented in the paper differ
-from UCT at the first step of the rollout, when the `simple' selection
-regret is minimized instead of the cumulative regret. Both the
-theoretical analysis and the empirical evaluation provide evidence for
-better general performance of the proposed algorithms.
-
-The improvement is inspired by the notion of value of information (VOI),
-but VOI is used implicitly in the analysis of the algorithm, rather
-than computed or learned explicitly in order to plan the rollouts. A
-further improvement can be achieved by computing or estimating the VOI
-of the rollouts and chosing a rollout that maximizes the VOI. Using VOI
-to guide Monte Carlo sampling instead of relying on the sample means
-of different actions  is particularly beneficial when different
-actions have qualitatively different distributions: for example, when the
-distribution variance varies signficantly between the actions.
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[scale=0.45,trim=0 0 0 16pt,clip]{flat-trilevel-vct-k=64.pdf}\\
-  \caption{VOI-based action selection (VCT) vs. other algorithms}
-  \label{fig:flat-trilevel-vct}
-\end{figure}
-
-VOI of a rollout can be computed when the sample distribution of an
-action is known up to the parameters, such as the normal distribution
-with an unknown mean and/or variance. Alternatively, the value of
-information can be estimated from the set of samples, and the need to
-assume a particular shape of the distribution is be lifted. In one
-realization of the latter approach the VOI of performing an action is
-estimated as the value of perfect information learned from the outcome
-of earlier samples of the action divided by the number of
-samples. Early experiments with this approach showed the best
-performance for sets of Bernoulli arms, and even a more prominent
-improvement for mixed arms (e.g. Bernoulli, triangularly distributed,
-and fixed arms) (Figure~\ref{fig:flat-trilevel-vct}):
-\textbf{VCT}, the algorithm which selects an action with the maximum VOI estimate on
-the first step exhibits lower average regret than other algorithms.
-Sample-based VOI estimation requires keeping more information than an
-algorithm based only on the sample mean, but the overhead is small and justifies
-the improvement.
-
-\section*{Acknowledgments}
-
-The research is partially supported by Israel
-Science Foundation grant 305/09, by the Lynne and William Frankel
-Center for Computer Sciences, and by the Paul Ivanier Center for
-Robotics Research and Production Management.
-
-\clearpage
-\appendix
-
-\section{Facts}
-
-{\bf Chernoff bounds (see \cite{Hagerup.chernoff}):} for $m$ independent random variables $X_1, X_2, ..., X_m$
-taking on values 0 or 1, $X=\sum_{i=1}^m X_i$, and $0\le\delta\le 1$:
-\begin{equation}
-\Pr[X < (1-\delta)\IE[X]] < e^{-\delta^2\IE[X]/2} \mbox{ (Chernoff bound)}
-\label{eq:chernoff-bound}
-\end{equation}
-Further on, if $\forall i. \IE[X_i]=\mu$:
-\begin{equation}
-\Pr[X < n\mu-a] <  e^{-2a^2/n}
-\label{eq:chernoff-hoeffding-bound} \mbox{ (Hoeffding-Chernoff bound)}
-\end{equation}
-
-\section{Derivations}
-\label{app:derivations}
-
-\subsection{Finite-time regret bounds}
-\label{app:derivations-finite}
-
-%TODO: bounds on regret based on information gathered, independent of
-%the way the choice is made.
-
-For an analysis of pure exploration of uniform and UCB algorithms, see also
-\cite{Bubeck.pure}.
-
-\subsubsection{$\varepsilon$-greedy}
-\label{app:derivations-finite-eps}
-
-For $0<\varepsilon\le1-\frac 1 K$ and $x_0>0$ (see Section 3, Proofs, in \cite{Auer.ucb}):
-
-\begin{eqnarray}
-\Pr(\overline X_j=\max_i\overline X_i)&\le&2\left(x_0\cdot \Pr\left\{T_j^R(n)\le x_0\right\} + \frac 2{\Delta_j^2}e^{-\Delta_j^2\lfloor x_0 \rfloor/2}\right)\nonumber\\
-&\le&2\left(x_0\cdot \Pr\left\{T_j^R(n)\le x_0\right\} + \frac {2 \cdot
-  e^{\Delta_j^2/2}}{\Delta_j^2}e^{-\Delta_j^2 x_0 /2}\right)
-\end{eqnarray}
-
-Number of times $\xi$ the $j$th arm is selected {\it randomly}:
-
-\begin{equation}
-\xi\triangleq\IE\left[T_j^R(n)\right]=\frac {n\varepsilon} K
-\end{equation}
-
-By choosing $x_0=\frac \xi 2$ and applying the Chernoff bound (\ref{eq:chernoff-bound}), obtain:
-
-\begin{eqnarray}
-\label{eq:pr-epsilon-greedy}
-\Pr(\overline X_j=\max_i\overline X_i)&\le& 2\left(\frac {\xi}{2} e^{-\xi/8} + \frac {2e^{\Delta_j^2/2}}{\Delta_j^2}e^{-\Delta_j^2 \xi/4}\right)\nonumber\\
-&\le&\left(\xi + \frac {4\sqrt e}{\Delta_j^2}\right)e^{-\Delta_j^2\xi/8}
-\end{eqnarray}
-
-A bound on the simple regret $r_\varepsilon$ of an
-$\varepsilon$-greedy policy follows from substituting
-(\ref{eq:pr-epsilon-greedy}) into (\ref{eq:simple-regret}):
-
-\begin{eqnarray}
-  \IE r_\varepsilon&\le&\sum_{j=1}^K\Delta_j\left(\xi + \frac {4\sqrt e}
-{\Delta_j^2}\right)e^{-\Delta_j^2\xi/8}\nonumber\\
-&\le&K\left(\xi + \frac {4\sqrt
-    e}{\Delta_{\min}}\right)e^{-\Delta_{\min}^2\xi/8}
-\label{eq:deriv-epsilon-greedy-b}
-\end{eqnarray}
-
-\subsubsection{UCB($\alpha$)}
-\label{app:derivations-finite-ucb}
-
-$n_*$ pulls of the best arm and $n_j$ pulls of the $j$th arm are
-achieved simultaneously when either
-
-\begin{equation}
-\sqrt {\frac {\alpha \log n} {n_j-1}} \ge 1 + \sqrt { \frac {\alpha \log n} {n_*}}
-\end{equation}
-
-or
-
-\begin{equation}
-\sqrt {\frac {\alpha \log n} {n_j}} \le 1 + \sqrt { \frac {\alpha \log n} {n_* - 1}}
-\label{eq:deriv-ucb-alpha-a}
-\end{equation}
-
-Inequality (\ref{eq:deriv-ucb-alpha-a}) can be used to derive a lower
-bound on the number of pulls of a non-optimal arm,  and, consequently, the upper
-bound on the probability of selecting a non-optimal arm:
-
-\begin{eqnarray}
-\frac {\alpha \log n} {n_j}&\le& 1 + 2  \sqrt {\frac {\alpha \log n}
-  {n_*-1}} + \frac {\alpha \log n} {n_*-1}\nonumber\\
-n_j &\ge& \frac {\alpha \log n} {1+2 \sqrt {\frac {\alpha \log n}
-    {n_*-1}} + \frac {\alpha \log n} {n_*-1}}\nonumber\\
-    &\ge& \frac {\alpha \log n} {4} \mbox{ for } \alpha \log n \le n_*-1
-\end{eqnarray}
-
-The probability that the $j$th arm will be chosen can be bounded using
-the Chernoff-Hoeffding bound:
-\begin{equation}
-P_j \le 2 \exp\left(-\frac {\alpha \Delta_j^2 \log n} 8 \right) = 2n^{\frac {-\alpha \Delta_j^2} 8}
-\end{equation}
-
-The expected simple regret of an UCB($\alpha$) algorithm is thus
-bounded by the following inequality:
-\begin{equation}
-\IE r_{ucb} \le 2\sum_{j=1}^K \Delta_jn^{\frac {-\alpha \Delta_j^2} 8}
-\label{eq:deriv-ucb-alpha-b}
-\end{equation}
-
-\cite{Bubeck.pure} establishes for UCB($\alpha$) a similar bound,
-polynomial in the number of pulls $n$.
-
-\subsection{Asymptotic regret analysis}
-\label{app:derivations-asym}
-
-%TODO: Oracle instead of asymptotic
-
-Bounds (\ref{eq:deriv-epsilon-greedy-b}, \ref{eq:deriv-ucb-alpha-b}) were
-derived ignoring dependency of the allocation on expected values
-of non-optimal arms. Derivation of better bounds seems
-complicated. Instead, one can analyse the asymptotic behavior of each of the
-algorithms under the assumption that the sample mean is close to the expectation.
-
-\subsubsection{$\varepsilon$-greedy}
-\label{app:derivations-asym-eps}
-
-A common technique \cite{Auer.ucb} to bound the probability of
-selecting a non-optimal arm $j$ is to split the interval between
-$\mu_j$ and $\mu_*$ at the middle and bound the probability of
-selecting the $j$th arm by the sum of the probabilities that the samle
-mean $\overline X_j$ of the $j$th arm is greater than $\mu_j+\frac
-{\Delta_j} 2$ and the sample mean $\overline X_*$ of the optimal arm is
-less than $\mu_*- \frac {\Delta_j} 2$. Each of the probabilities can
-be bound using the Chernoff-Hoefding bound.
-
-Splitting the interval between $\mu_j$ and $\mu_*$ half-way is an
-arbirtrary decision. By selecting a different split point
-depending on the allocation of samples, one can obtain a tighter
-bound. On the other hand, there is a value of $\varepsilon$ that
-minimizes the bound for the $\varepsilon$-greedy allocation scheme.
-Thus, the  $\varepsilon$-greedy allocation scheme can be tuned by
-finding $\varepsilon$ that minimizes the bound obtained by
-selecting the best split point.
-
-Split the interval $[\mu_j, \mu_*]$ at $\mu_j+\delta_j$. The
-probability $P_j$ for selecting the $j$th arm is bounded as:
-
-\begin{equation}
-P_j\le exp\left(-2\delta_j^2n_j\right)+exp\left(-2(\Delta_j-\delta_j)^2n_*\right)
-\end{equation}
-
-For the $\varepsilon$-greedy scheme, $n_i=\frac {\varepsilon n} K$,
-$n_*=(1-\varepsilon)n$, thus
-
-\begin{equation}
-P_j \le \exp\left(-\frac {2\delta_j^2\varepsilon n} {K-1}\right) +\exp\left(-2(\Delta_j-\delta_j)^2(1-\varepsilon)n\right)
-\end{equation}
-
-In general, $\varepsilon$ and $\delta$ which minimize $P_j$ will
-depend of $n$. For a constant $\varepsilon$, require that the
-exponents in both terms are equal
-\begin{eqnarray}
-\exp\left(-\frac {2\delta_j^2\varepsilon n} {K-1}\right)
-   &=& \exp\left(-2(\Delta_j-\delta_j)^2(1-\varepsilon)n\right)\\
-\frac {\delta_j} {\Delta_j - \delta_j}
-   &=& \sqrt { \frac {(K-1)(1-\varepsilon)} \varepsilon }
-\label{eq:deriv-exps-equal}
-\end{eqnarray}
-and find conditional minima of $P_j(\delta_j, \epsilon)$ satisfying
-either $\frac {\partial P_j} {\partial \delta_j}=0$ or $\frac {\partial P_j}
-{\partial \varepsilon}=0$.
-
-\vspace{1em}
-\paragraph{a. {\large $\frac {\partial P_j} {\partial \delta_j}=0$}}
-
-\begin{eqnarray*}
-\frac {4\delta_j\varepsilon n} {K-1}
-  &-& 4(\Delta_j-\delta_j)(1-\varepsilon)n=0\\
-\frac {\delta_j} {\Delta_j-\delta_j}&=&{ \frac {(K-1)(1-\varepsilon)} \varepsilon }
-\end{eqnarray*}
-and together with (\ref{eq:deriv-exps-equal}) this gives:
-\begin{eqnarray}
-\delta_j&=&\frac {\Delta_j} 2\nonumber\\
-\varepsilon&=&1-\frac 1 K\nonumber\\
-\IE r_{\varepsilon=1-1/K}&\le&B_{uniform}=2\exp\left(-\frac {\Delta_j^2n} {2K}\right)
-\label{eq:deriv-uniform-sampling}
-\end{eqnarray}
-
-That is, the case $\frac {\partial P_j} {\partial \delta_j}=0$
-corresponds to \emph{uniform random sampling}, and the tightest bound is obtained
-by splitting the interval at the middle.
-
-\paragraph{b. {\large $\frac {\partial P_j} {\partial \varepsilon}=0$}}
-
-\begin{eqnarray*}
-\frac {2\delta_j^2 n} {K-1}
-  &-& 2(\Delta_j-\delta_j)^2n=0\\
-\frac {\delta_j} {\Delta_j-\delta_j}&=&\sqrt {K-1}
-\end{eqnarray*}
-and together with (\ref{eq:deriv-exps-equal}) this gives:
-\begin{eqnarray}
-\frac {\delta_j} {\Delta_j}&=&\frac 1 {\sqrt {K-1}+1}\nonumber\\
-\varepsilon&=&\frac 1 2\nonumber\\
-\IE
-r_{\varepsilon=1/2}&\le&B_{1/2\mbox{-}greedy}=2\exp\left(-\frac
-  {\Delta_j^2n} {(\sqrt{K-1}+1)^2}\right)\nonumber\\
-\lim_{K\to\infty}
-\mathbb{B}r_{1/2\mbox{-}greedy}^K&=&2\exp\left(-{\Delta_j^2n}\right)\nonumber\\
-\mbox{for large $K$: }\mathbb{B}r_{1/2\mbox{-}greedy}&\approx&2\exp\left(-\frac {\Delta_j^2n} K\right)
-\label{eq:deriv-half-greedy-sampling}
-\end{eqnarray}
-
-That is, the case $\frac {\partial P_j} {\partial \varepsilon}=0$
-corresponds to \emph{$\frac 1 2$-greedy sampling}, and for a large number of
-arms $K$ 
-\begin{equation}
-\mathbb{B}r_{1/2\mbox{-}greedy}\approx \mathbb{B}r_{uniform}^2
-\label{eq:derive-uniform-vs-half-greedy}
-\end{equation}
-Equation (\ref{eq:derive-uniform-vs-half-greedy}) suggests that
-$\frac 1 2$-greedy sampling has a much higher convergence rate than uniform
-sampling for a large number of arms.
-
-\subsubsection{UCB($\alpha$)}
-\label{app:derivations-asym-ucb}
-
-The $UCB(\alpha)$ allocation scheme distributes arm pulls such that
-the upper bound on the probability of selecting a non-optimal arm is
-asymptotically the same for all arms. Indeed, for large n
-\begin{eqnarray}
-n_i&\ll&n\\
-n_*&\approx&n\\
-\sqrt {\frac {2 \log n} {n_j}}&\approx&\Delta_j+\sqrt {\frac {2\log n}
-  {n_*}} \\
-&\Rightarrow& n_j \approx \frac {2\log n} {\Delta_j^2}
-\label{eq:deriv-ucb-asymptotic-nj}
-\end{eqnarray}
-and by Hoeffding-Chernoff bound the upper bound $\mathbb{B}P_j$ on the probability
-$P_j^{ucb}$ to choose the $j$th arm is
-\begin{equation}
-P_j^{ucb}\le\mathbb{B}P_j^{ucb}\approx2\exp\left(-2\Delta_j^2\frac {2\log n}{\Delta_j^2}\right)
-             =2n^{-4}
-\end{equation}
-
-That is, UCB($\alpha$) supposedly distributes pulls of non-optimal
-arms better than $\varepsilon$-greedy
-scheme which may undersample arms with high expectation. The
-polynomial, rather than exponential, convergence rate (Equation
-\ref{eq:deriv-ucb-alpha-b}) is due to oversampling the best arm.
-
-Replacing the logarithm with a faster growing subpolynomial
-function, like square root (UQB), results in a superpolynomial convergence
-rate of the upper bound:
-\begin{equation}
-P_j^{uqb}\le\mathbb{B}P_j^{uqb}\approx2\exp\left(-\beta\Delta_j^2\frac {\sqrt n}{\Delta_j^2}\right)
-            =2\exp(-\beta\sqrt n)
- \end{equation}
-where $\beta$ is a constant that depends only on reward bounds.
-Finite time convergence can still be shown similarly to UCB.
+\section{Introduction}
 

### 2011-09-13 paper/rmcts/paper.tex
small fix
more papers

@@
 pulls, brings a reward, that is, the simple regret, as opposite to the
-cumulative regret, must be minimized. This work introduces policies for
+cumulative regret, must be minimized. This ongoing work aims at applying
+meta-reasoning techniques to MCTS, which is non-trivial.
+We begin by introducing policies for
 multi-armed bandits with lower simple regret than UCB, and an
 algorithm for MCTS which combines cumulative and simple regret
-minization and outperforms UCT. Finite-time and asymptotic analysis of
-the policies is provided, and the algorithms are empirically compared.
+minimization and outperforms UCT. We also develop a sampling scheme loosely based
+on a myopic version of perfect value of information.
+Finite-time and asymptotic analysis of
+the policies is provided, and the algorithms are compared empirically.
 \end{abstract}
@@
 
-The Multi-armed Bandit is a set of $K$ arms. Each arm can be pulled
+Monte-Carlo tree sampling, and especially a version based on the
+UCT formula \cite{Kocsis.uct} appears in numerous search applications,
+such as \cite{Eyerich.ctp}. Although these methods are shown to be successful empirically,
+most authors appear to be using the UCT formula ``because it has been shown
+to be successful in the past'', and ``because it does a good job of
+trading off exploration and exploitation''. While the latter statement may be
+correct for the multi-armed bandit and for the UCB method \cite{Auer.ucb},
+we argue that it is inappropriate for search. The problem is not that
+UCT does not work; rather, we argue that a simple reconsideration from basic
+principles can result in schemes that outperform UCT. This is demonstrated both
+theoretically and empirically.
+
+In the Multi-armed Bandit problem we have a set of $K$ arms. Each arm can be pulled
 multiple times. When the $i$th arm is pulled, a random reward $X_i$
@@
 where $\Delta_j=\mu_*-\mu_j$.
-Strategies that minimize the simple reward are called pure exploration
+Strategies that minimize the simple regret are called pure exploration
 strategies \cite{Bubeck.pure}.
@@
 a cutoff state for which the reward is evaluated
-approximately). Algorithm UCB that minimizes the cumulative regret
-\cite{Auer.ucb} had been extended to tree search as UCT
-\cite{Kocsis.uct}.
+approximately). The UCB algorithm (that attempts to minimize the cumulative regret)
+\cite{Auer.ucb} had been extended into the tree search sampling scheme
+known as UCT \cite{Kocsis.uct}.
 
@@
 search step, the algorithm allocates Monte Carlo rollouts to choose
-the best action, thereby minimizing the cumulative regret. However, the
-simple and the cumulative regret cannot be minimized simultaneously;
+the best action, thereby (approximately) minimizing the cumulative regret. 
+The core issue is that in search (especially for adversarial search
+and for ``games against nature'' - optimizing behavior under uncertainty) the goal is typically
+to either find a good (or optimal) strategy, or even to find the best first
+action of such a policy. Once such an action is discovered, it is not beneficial
+to further sample that action, ``exploitation'' is thus meaningless for
+search problems. Finding a good first action is closer to the pure exploration
+variant, as seen in the selection problem \cite{Bubeck.pure,TolpinShimony.blinkered}. In the selection
+problem, it is much better to minimize the {\em simple} regret.
+However, the simple and the cumulative regret cannot be minimized simultaneously;
 moreover, \cite{Bubeck.pure} shows that in many cases the smaller the
@@
 
-UCT performance can be improved by combining UCB with a sampling
+UCT performance can thus be improved by combining UCB with a sampling
 scheme that minimizes the simple regret of selecting an action at the
@@
 \item maximize the value of partial
-information by sampling actions to minimize the \textbf{simple} regret at the
-current root node, and then 
-\item minimize the \textbf{cumulative} regret of the rollouts from the second
-  step on to approach the optimum behavior. 
+information by sampling actions to
+minimize the \textbf{simple} regret
+of the selection at the current root node, and
+\item as the goal of sampling in deeper tree nodes is estimating the value of
+a node, rather than selection, it makes sense to
+minimize the \textbf{cumulative} regret of the rollouts from the second
+step onwards. 
 \end{itemize}
+
+The ultimate goal of this ongoing work is ``optimal'' sampling
+in the meta-reasoning sense. Nevertheless, this task is daunting for
+the following reasons:
+\begin{itemize}
+\item Defining the cost of a sample is not easy, and even if we simply use time-cost
+as an approximation, we get an intractable meta-reasoning problem.
+\item Applying the standard myopic and subtree independence assumptions, we run into
+serious problems. Even in the standard selection problem \cite{TolpinShimony.blinkered}, we get a non-concave
+utility function and premature stopping of the algorithm. This is due to the fact
+that the value of information of a single measurement (analogous
+to a sample in MCTS) is frequently less than its
+time-cost, even though this is not true for multiple measurements.
+When applying the selection problem to MCTS, the situation is exacerbated. 
+The utility of an action is usually bounded, and thus in many cases a single sample
+may be insufficient to change the current best action, {\em regardless} of its
+outcome. As a result, we frequently
+get a {\em zero} ``myopic'' value of information for a single sample.
+\end{itemize}
+
+As the ultimate task is extremely difficult
+to achieve, and even harder to analyze, we begin with
+simple schemes more amenable to analysis. We then
+develop a crude value of information (VOI)
+based scheme. Although the latter is a somewhat ad-hoc attempt to
+estimate value of information of a sample, it already appears (empirically) to do
+better than UCT as well as better than all the simpler schemes we propose.
 
@@
 $\varepsilon$. The lowest bound is
-acheived for $\varepsilon=\frac 1
+achieved for $\varepsilon=\frac 1
 2$, and for a large number of arms the bound for the $\frac 1 2$-greedy scheme approaches the square of the
@@
 
-Figure~\ref{fig:sailing-cost-vs-factor} shows the rerget vs. the
+Figure~\ref{fig:sailing-cost-vs-factor} shows the regret vs. the
 exploration factor for different numbers of samples. QCT is always better than
@@
 further improvement can be achieved by computing or estimating the VOI
-of the rollouts and chosing a rollout that maximizes the VOI. Using VOI
+of the rollouts and choosing a rollout that maximizes the VOI. Using VOI
 to guide Monte Carlo sampling instead of relying on the sample means
@@
 actions have qualitatively different distributions: for example, when the
-distribution variance varies signficantly between the actions.
+distribution variance varies significantly between the actions.
 
@@
 information can be estimated from the set of samples, and the need to
-assume a particular shape of the distribution is be lifted. In one
+assume a particular shape of the distribution can be lifted. In one
 realization of the latter approach the VOI of performing an action is
@@
 algorithm based only on the sample
-mean, but the overhead is small and
-is justified by the improvement.
+mean, but the overhead is small and justified by the improvement.
 
@@
 of non-optimal arms. Derivation of better bounds seems
-complicated. Instead, one can analyse the asymptotic behavior of each of the
+complicated. Instead, one can analyze the asymptotic behavior of each of the
 algorithms under the assumption that the sample mean is close to the expectation.
@@
 Splitting the interval between $\mu_j$ and $\mu_*$ half-way is an
-arbirtrary decision. By selecting a different split point
+arbitrary decision. By selecting a different split point
 depending on the allocation of samples, one can obtain a tighter

### 2011-09-13 paper/voisam/paper.tex
more papers

@@
 
-\section{Introduction}
+\section{Introduction and Definitions}
+
+Taking a sequence of samples in order to minimize the
+regret of a decision based on the samples is abstracted by the
+{\em Multi-armed Bandit Problem.} In the Multi-armed Bandit problem
+we have a set of $K$ arms. Each arm can be pulled multiple
+times. When the $i$th arm is pulled, a random reward $X_i$ from an
+unknown stationary distribution is returned.  The reward is bounded
+between 0 and 1.
+
+The simple regret of a sampling policy for the Multi-armed Bandit
+Problem is the expected difference between the best expected reward
+$\mu_*$ and the expected reward $\mu_j$ of the arm with the best sample mean
+$\overline X_j=\max_i\overline X_i$:
+\begin{equation}
+\label{eq:simple-regret}
+\IE[R]=\sum_{j=1}^K\Delta_j\Pr(\overline X_j=\max_i\overline X_i)
+\end{equation}
+where $\Delta_j=\mu_*-\mu_j$.
+Strategies that minimize the simple regret are called pure exploration
+strategies \cite{Bubeck.pure}. Principles of rational metareasoning
+\cite{Russel.right} suggest that at each step the arm with the great
+value of information (VOI) must be pulled, and the sampling must be
+stopped and a decision must be made when no arm has positive VOI.
+
+To estimate the VOI of pulling an arm, either a particular 
+distribution of the rewards should be assumed (and updated based on
+pull outcomes), or a distribution-independent bound on the VOI can be
+used as the estimate. In this paper, we use {\em concentration inequalities}
+to derive distribution-independent bounds on the VOI.
+
+\section{Some Concentration Inequalities}
+
+\begin{description}
+\item[Markov' inequality \cite{}:] Let $X \ge 0$ be a random non-negative variable
+  with finite mean $\IE[X]$. Then for all $a > 0$
+\begin{equation}
+\Pr(X\ge a)\le \frac {\IE[X]} a
+\label{eq:conc-markov-upper}
+\end{equation}
+If additionally $X\le 1$, then by susbstituting $Y=1-X$
+\begin{equation}
+\Pr(X\le a)\le \frac {1-\IE[X]} {1-a}
+\label{eq:conc-markov-lower}
+\end{equation}
+\item[Hoeffding's inequality \cite{}:] Let $X_1, \ldots, X_n$ be
+independent random variables with values from $[0,1]$,
+$X=\frac 1 n \sum_{i=1}^n X_i$. Then 
+\begin{eqnarray}
+\Pr(X-\IE[X] \ge a) &\le& \exp ( -2na^2)\nonumber\\
+\Pr(\IE[X]-X \ge a) &\le& \exp ( -2na^2)
+\label{eq:conc-hoeffding}
+\end{eqnarray}
+\item[Maurer \& Pontil's inequality \cite{MaurerPontil.benrstein}:] \footnote{A
+  modification of Bennet's inequality \cite{} based on the sample,
+  rather than the true, variance.}
+\end{description}
+
+\section{Upper Bounds on Value of Information}
+
+\section{VOI-based Sampling Control}
+
+\subsection{Selection Criterion}
+
+\subsection{Termination Condition}
 

### 2011-09-13 paper/voisam/paper.tex
bounds

@@
 
+Let $X_1, \ldots, X_n$ be i.i.d. random variables with values from $[0,1]$,
+$X=\frac 1 n \sum_{i=1}^n X_i$. Then 
 \begin{description}
-\item[Markov' inequality \cite{}:] Let $X \ge 0$ be a random non-negative variable
-  with finite mean $\IE[X]$. Then for all $a > 0$
+\item[Hoeffding's inequality \rm{\cite{Hoeffding.ineq}}:] 
 \begin{equation}
-\Pr(X\ge a)\le \frac {\IE[X]} a
-\label{eq:conc-markov-upper}
+\Pr(X-\IE[X] \ge a) \le \exp ( -2na^2)
+\label{eq:conc-hoeffding}
 \end{equation}
-If additionally $X\le 1$, then by susbstituting $Y=1-X$
+\item[Empirical Bernstein's inequality \rm{\cite{MaurerPontil.benrstein}}:]
 \begin{equation}
-\Pr(X\le a)\le \frac {1-\IE[X]} {1-a}
-\label{eq:conc-markov-lower}
+\Pr(X-\IE[X] \ge a) \le 2\exp \left( \frac {-na^2} {10a+2\overline\sigma_n^2}\right)
+\label{eq:conc-empbernstein}
 \end{equation}
-\item[Hoeffding's inequality \cite{}:] Let $X_1, \ldots, X_n$ be
-independent random variables with values from $[0,1]$,
-$X=\frac 1 n \sum_{i=1}^n X_i$. Then 
-\begin{eqnarray}
-\Pr(X-\IE[X] \ge a) &\le& \exp ( -2na^2)\nonumber\\
-\Pr(\IE[X]-X \ge a) &\le& \exp ( -2na^2)
-\label{eq:conc-hoeffding}
-\end{eqnarray}
-\item[Maurer \& Pontil's inequality \cite{MaurerPontil.benrstein}:] \footnote{A
-  modification of Bennet's inequality \cite{} based on the sample,
-  rather than the true, variance.}
+where sample variance $\overline\sigma_n^2$ is
+\begin{equation}
+\overline\sigma_n^2=\frac 1 {n(n-1)} \sum_{1\le i < j\le n}(X_i-X_j)^2
+\label{eq:sample-variance}
+\end{equation}
 \end{description}
+Bounds (\ref{eq:conc-hoeffding}, \ref{eq:conc-empbernstein}) are symmetrical
+around the mean. Bound (\ref{eq:conc-empbernstein}) is tighter than
+(\ref{eq:conc-hoeffding}) for small $a$ and $\overline\sigma_n^2$.
 

### 2011-09-14 paper/voisam/paper.tex
added derivation for empirical bernstein inequality

@@
 
-To estimate the VOI of pulling an arm, either a particular 
+To estimate the VOI of pulling an arm, either a certain 
 distribution of the rewards should be assumed (and updated based on
-pull outcomes), or a distribution-independent bound on the VOI can be
-used as the estimate. In this paper, we use {\em concentration inequalities}
+observed rewards), or a distribution-independent bound on the VOI can be
+used as the VOI estimate. In this paper, we use {\em concentration inequalities}
 to derive distribution-independent bounds on the VOI.
@@
 \end{equation}
-\item[Empirical Bernstein's inequality \rm{\cite{MaurerPontil.benrstein}}:]
-\begin{equation}
-\Pr(X-\IE[X] \ge a) \le 2\exp \left( \frac {-na^2} {10a+2\overline\sigma_n^2}\right)
+\item[Empirical Bernstein's inequality
+  \rm{\cite{MaurerPontil.benrstein}}:]\footnote{see
+    Appendix~\ref{app:deriv-conc-empbernstein} for derivation}
+\begin{eqnarray}
+\Pr(X-\IE[X] \ge a) &\le& 2\exp \left( - \frac {na^2} {\frac {14} {3}
+                          \frac {n} {n-1}a+2\overline\sigma_n^2}\right)\nonumber\\
+                    &<& 2\exp \left( - \frac {na^2} {10a+2\overline\sigma_n^2}\right)
 \label{eq:conc-empbernstein}
-\end{equation}
+\end{eqnarray}
 where sample variance $\overline\sigma_n^2$ is
@@
 Bounds (\ref{eq:conc-hoeffding}, \ref{eq:conc-empbernstein}) are symmetrical
-around the mean. Bound (\ref{eq:conc-empbernstein}) is tighter than
+around the mean. Bound~(\ref{eq:conc-empbernstein}) is tighter than
 (\ref{eq:conc-hoeffding}) for small $a$ and $\overline\sigma_n^2$.
@@
 
+\appendix
+
+\section{Empirical Bernstein Inequality}
+\label{app:deriv-conc-empbernstein}
+
+Theorem~4 in~\cite{MaurerPontil.benrstein} states that
+\[\Pr\left(\IE[X]-\overline X_n \ge \sqrt { \frac {2\overline\sigma_n^2 \ln 2/\delta} n } + \frac {7 \ln 2/\delta} {3(n-1)}\right)\le \delta,\]
+Therefore
+\[\Pr\left(\IE[X]-\overline X_n \ge \sqrt { \left(\frac {7 \ln 2/\delta} {3(n-1)}\right)^2+\frac {2\overline\sigma_n^2 \ln 2/\delta} n } + \frac {7 \ln 2/\delta} {3(n-1)}\right)\le
+\delta.\]
+$t=\sqrt { \left(\frac {7 \ln 2/\delta} {3(n-1)}\right)^2+\frac {2\overline\sigma_n^2 \ln 2/\delta} n } + \frac {7 \ln 2/\delta} {3(n-1)}$ is a root of
+square equation
+\[t^2-t\frac {14 \ln 2/\delta} {3(n-1)} -\frac {2\overline\sigma_n^2 \ln 2/\delta} n=0\]
+which, solved for $\delta\triangleq\Pr(\IE[X]-\overline X_n\ge t)$,
+gives
+\[\Pr(\IE[X]-\overline X_n\ge t)\le 2\exp \left( - \frac {na^2} {\frac {14} {3} \frac {n} {n-1}a+2\overline\sigma_n^2}\right)\]
+Other derivations, giving slightly different results, are possible.
+
 \bibliographystyle{plain}

### 2011-09-14 paper/voisam/paper.tex
first draft for the bounds

@@
 \begin{abstract}
+Upper bounds for the VOI are provided for pure exploitaiton in the
+Mulit-armed Bandit Problem. Sampling policies based on the upper
+bounds are suggested.
 \end{abstract}
@@
 
+The intrinsic VOI of pulling an arm is the expected decrease
+in the regret compared to selecting an arm without pulling any arm at
+all. The intrinsic VOI can be estimated as the intrinsic 
+value of perfect information $\Lambda_i$ about the mean reward of the $i$th arm. Two
+cases are possible:
+\begin{itemize}
+\item the arm ($\alpha$) with the highest sample mean is pulled, and the 
+mean of the arm is lower than the sample mean the second-best arm ($\beta$);
+\item another arm is pulled, and the mean of the arm is higher
+than the current highest sample mean ($\alpha$).
+\end{itemize}
+\begin{eqnarray}
+\Lambda_\alpha&=&\int_0^{\overline X_\beta}p(\IE[X_\alpha]=x)(\overline X_\beta-x)dx\nonumber\\
+\Lambda_{i|i\ne\alpha}&=&\int_{\overline X_\alpha}^1p(\IE[X_i]=x)(x-\overline X_\alpha)dx
+\end{eqnarray}
+Using the concentration inequalities, $\Lambda_i$ can be bounded from
+above as the probability that a different arm is selected times the
+maximum possible increase in the reward:
+\begin{eqnarray}
+\Lambda_\alpha&\le&\Pr(\IE[X_\alpha]\le\overline X_\beta)\overline X_\beta\nonumber\\
+\Lambda_{i|i\ne \alpha}&\le&\Pr(\IE[X_i]\ge\overline
+X_\alpha)(1-\overline X_\alpha)
+\label{eq:lambda-general-bounds}
+\end{eqnarray}
+
+The probabilities in bounds~(\ref{eq:lambda-general-bounds}) can be
+bounded from above by the concentration
+inequalities~(\ref{eq:conc-hoeffding}, \ref{eq:conc-empbernstein});
+thus upper bounds on $\Lambda$ corresponding to each of the
+concentration inequalities are obtained. For the Hoeffding's
+inequality~(\ref{eq:conc-hoeffding}):
+\begin{eqnarray}
+\Lambda_\alpha&\le&\hat\Lambda_\alpha^h=\overline X_\beta \exp\left(-2n(\overline  X_\alpha - \overline X_\beta)^2\right)\nonumber\\
+\Lambda_{i|i\ne \alpha}&\le&\hat\Lambda_{i|i\ne \alpha}^h=(1-\overline X_\alpha) \exp\left(-2n(\overline X_\alpha - \overline X_i)^2\right)
+\label{eq:lambda-hoeffding-bounds}
+\end{eqnarray}
+For the empirical Bernstein's inequality~(\ref{eq:conc-empbernstein}):
+\begin{eqnarray}
+\Lambda_\alpha&\le&\hat\Lambda_\alpha^h=\overline X_\beta \exp\left(-\frac {n(\overline X_\alpha - \overline X_\beta)^2} {10(\overline X_\alpha - \overline X_\beta)+2\overline\sigma_{\alpha,n}^2}\right)\nonumber\\
+\Lambda_\alpha&\le&\hat\Lambda_{i|i\ne \alpha}^h=(1-\overline X_\alpha) \exp\left(-\frac {2n(\overline X_\alpha - \overline X_i)^2} {10(\overline X_\alpha - \overline X_i)+2\overline\sigma_{i,n}^2}\right)
+\label{eq:lambda-bernstein-bounds}
+\end{eqnarray}
+and a tighter bound can be constructed as $\hat\Lambda^{h,b}=\min(\hat\Lambda^h,\hat\Lambda^b)$.
+
 \section{VOI-based Sampling Control}
@@
 
+Following the principles of rational metareasoning, an arm with the highest upper bound $\hat
+Lambda$ on the perfect value of information should be pulled at each
+step. This way, arms known to have a low VOI  would be pulled less frequently.
+
 \subsection{Termination Condition}
+
+The upper bounds~(\ref{eq:lambda-hoeffding-bounds}, \ref{eq:lambda-bernstein-bounds}) decrease exponentially with the
+number of pulls $n$. When the upper bound of the VOI for all arms
+becomes lower than a threshold $\lambda$, which can be chosen based on
+resource constraints, the sampling should be stopped and an arm should
+be chosen.
 

### 2011-09-14 paper/voisam/paper.tex
spell-checked
typos fixed

@@
 \begin{abstract}
-Upper bounds for the VOI are provided for pure exploitaiton in the
-Mulit-armed Bandit Problem. Sampling policies based on the upper
+Upper bounds for the VOI are provided for pure exploration in the
+Multi-armed Bandit Problem. Sampling policies based on the upper
 bounds are suggested.
@@
                           \frac {n} {n-1}a+2\overline\sigma_n^2}\right)\nonumber\\
-                    &<& 2\exp \left( - \frac {na^2} {10a+2\overline\sigma_n^2}\right)
+                    &\le& 2\exp \left( - \frac {na^2} {10a+2\overline\sigma_n^2}\right)
 \label{eq:conc-empbernstein}
@@
 \begin{eqnarray}
-\Lambda_\alpha&\le&\hat\Lambda_\alpha^h=\overline X_\beta \exp\left(-\frac {n(\overline X_\alpha - \overline X_\beta)^2} {10(\overline X_\alpha - \overline X_\beta)+2\overline\sigma_{\alpha,n}^2}\right)\nonumber\\
-\Lambda_\alpha&\le&\hat\Lambda_{i|i\ne \alpha}^h=(1-\overline X_\alpha) \exp\left(-\frac {2n(\overline X_\alpha - \overline X_i)^2} {10(\overline X_\alpha - \overline X_i)+2\overline\sigma_{i,n}^2}\right)
+\Lambda_\alpha&\le&\hat\Lambda_\alpha^h=2\overline X_\beta \exp\left(-\frac {n(\overline X_\alpha - \overline X_\beta)^2} {10(\overline X_\alpha - \overline X_\beta)+2\overline\sigma_{\alpha,n}^2}\right)\nonumber\\
+\Lambda_{i|i\ne \alpha}&\le&\hat\Lambda_{i|i\ne \alpha}^h=2(1-\overline X_\alpha) \exp\left(-\frac {n(\overline X_\alpha - \overline X_i)^2} {10(\overline X_\alpha - \overline X_i)+2\overline\sigma_{i,n}^2}\right)
 \label{eq:lambda-bernstein-bounds}
@@
 Following the principles of rational metareasoning, an arm with the highest upper bound $\hat
-Lambda$ on the perfect value of information should be pulled at each
+\Lambda$ on the perfect value of information should be pulled at each
 step. This way, arms known to have a low VOI  would be pulled less frequently.
@@
 becomes lower than a threshold $\lambda$, which can be chosen based on
-resource constraints, the sampling should be stopped and an arm should
+resource constraints, the sampling should be stopped, and an arm should
 be chosen.

### 2011-09-21 paper/voisam/onbetterbound.tex
on better bound

\documentclass{article}
\usepackage{algpseudocode}
\usepackage[ruled]{algorithm}
\usepackage{url}
\usepackage{framed}
\usepackage{amsfonts,amsmath,amsthm,amssymb}
\usepackage{graphicx}
\usepackage{url}
\usepackage{color}
\usepackage{geometry}

\geometry{margin=1in}

\newcommand {\mean} {\ensuremath {\mathop{\mathrm{mean}}}}
\newcommand {\median} {\ensuremath {\mathop{\mathrm{median}}}}
\newcommand {\N} {\ensuremath {\mathcal{N}}}
\newcommand {\IE} {\ensuremath {\mathbb{E}}}
\newcommand {\cov} {\ensuremath {\mathop{\mathrm{cov}}}}
\newcommand {\BEL} {\ensuremath {\mathop{\mathrm{BEL}}}}

\newtheorem{lemma}{Lemma}

\title{On Better Bound on VOI of Monte Carlo Sampling}
\author {David Tolpin, Solomon Eyal Shimony \\
\{tolpin,shimony\}@cs.bgu.ac.il}

\begin{document}

\maketitle

Consider the case of VOI of measurement the current best item (sample mean
$\alpha$). The probability that the true mean of the item $\mu_\alpha$ is at
most $x$ is bounded by the Hoeffding's inequality as:
\[ \Pr(\mu_\alpha \le x) \le \exp\left(-2n(\alpha-x)^2\right) \]
An upper bound on the VOI is thus the product of the probability that
the true mean of the current best item is less than the sample mean
$\beta$ of the current second-best item times the maximum reward,
$\beta$:
\[ V \le \beta \exp\left(-2n(\alpha-\beta)^2\right) \]
The bound can be supposedly be improved by selecting a midpoint
$0 < \gamma < \beta$ and computing the bound as the sum of two parts:
\begin{itemize}
\item $\beta-\gamma$ multiplied by the probability that
  $\mu_\alpha \le \beta$;
\item $\beta$ multiplied by the probability that $\mu_\alpha\le
  \gamma$.
\end{itemize}.
\[ V^* \le (\beta-\gamma)\exp\left(-2n(\alpha-\beta)^2\right)+\beta\exp\left(-2n(\alpha-\gamma)^2\right) \]

\vspace{\baselineskip}

The minimum of $V^*$ is achieved when $\frac {dV^*} {d\gamma}=0$, that
is, when $\gamma$ is the root of the following equation:
\[ 4\beta n(\alpha-\gamma)=\exp\left(-2n\left(\frac {\alpha-\beta}
    {\alpha-\gamma}\right)^2\right) \]
If a root in the interval  $0\le\gamma\le\beta$ exists, then the
number of samples is bounded as
\[n\le\frac 1 {4\beta(\alpha-\beta)}\]
by observing that the righthand side 
is at most $1$ (a negative power), and the left-hand side is at least
$4\beta n(\alpha-\beta)$.
So, the bound can supposedly be improved for smaller values of
$n$. The improvement is more significant when the current best and
second-best sample means are close.

The derivation for the other case
(sampling an item that can be better than the current best) is
obtained by substitution $1-\alpha, 1-\beta, 1-\gamma$ instead of
$\alpha, \beta, \gamma$. The anticipated influence of the improved
VOI estimate would be that the selected item will be a less discovered
one and further from the current best or second-best.

A closed-form solution for $\gamma$ cannot be achieved, but the value
can be efficiently computed. It should be determined empirically
whether the improved estimate has justified influence on the
performance of the algorithm.

\end{document}

### 2011-09-21 paper/voisam/onbetterbound.tex
spell checked

@@
 \[n\le\frac 1 {4\beta(\alpha-\beta)}\]
-by observing that the righthand side 
+by observing that the right-hand side 
 is at most $1$ (a negative power), and the left-hand side is at least
@@
 
-A closed-form solution for $\gamma$ cannot be achieved, but the value
-can be efficiently computed. It should be determined empirically
-whether the improved estimate has justified influence on the
-performance of the algorithm.
+A closed-form solution for $\gamma$ cannot be obtained, but given
+$\alpha, \beta, n$, the value of $\gamma$ can be efficiently
+computed. It should be determined empirically whether the improved
+estimate has justified influence on the performance of the algorithm.
 

### 2011-11-26 code/figures.R
fixed swapped results

@@ -71,3 +71,36 @@ make.pdfs <- function() {
     }
   }
 }
+
+
+go.times <- c('5000', '7000', '10000', '15000')
+go.pchs <- list('5000'=1, '7000'=2, '10000'=3, '15000'=4)
+draw.go.curves <- function(agent="vct") {
+  res <- list()
+  lo <- 1.0
+  hi <- 0.0
+  for(time in go.times) {
+    fname <- paste("uct", agent, "-time=", time, ".shomersu.dat", sep="")
+    res[[time]] <- read.table(fname, header=T)
+    res[[time]]$MAXVOI <- res[[time]]$MAXVOI+1E-8
+    if(lo > min(res[[time]]$WIN_W)) lo <- min(res[[time]]$WIN_W)
+    if(hi < max(res[[time]]$WIN_W)) hi <- max(res[[time]]$WIN_W)
+  }
+  lo <- min(lo, 0.3)
+  hi <- min(hi, 0.7)
+  
+
+  plot(res[['5000']]$MAXVOI, res[['5000']]$WIN_W, ylim=c(lo,hi), log="x", type='n',
+       main=NA, xlab=expression(VOI[min]), ylab='UCT wins, %')
+  abline(h=0.5, lty='dashed')
+  lines
+  for(time in go.times) {
+    lines(res[[time]]$MAXVOI, res[[time]]$WIN_W, type='o', pch=go.pchs[[time]])
+  }
+
+  legend(x='topleft', legend=go.times, pch=sapply(go.times, function(time) go.pchs[[time]]))
+}
+
+  
+  
+

### 2011-11-28 code/almostbest.py
all results, drawing best bars

@@ -1,0 +1,17 @@
+import sys
+
+bestweight = len(sys.argv)==1 and 1.0 or float(sys.argv[1])
+
+header = sys.stdin.readline().split()
+alpha = 1.0
+beta = 1.0
+for line in sys.stdin:
+	data = dict(zip(header, line.split()))
+	value = float(data['WIN_W'])
+	if value < alpha:
+		beta = alpha
+		alpha = value
+	elif value < beta:
+		beta = value
+print (bestweight*alpha+(1-bestweight)*beta)
+

### 2011-11-28 code/figures.R

@@ -77,22 +77,23 @@ go.times <- c('5000', '7000', '10000', '15000')
 go.pchs <- list('5000'=1, '7000'=2, '10000'=3, '15000'=4)
 draw.go.curves <- function(agent="vct") {
   res <- list()
-  lo <- 1.0
+  lo <- 100.0
   hi <- 0.0
   for(time in go.times) {
     fname <- paste("uct", agent, "-time=", time, ".shomersu.dat", sep="")
     res[[time]] <- read.table(fname, header=T)
     res[[time]]$MAXVOI <- res[[time]]$MAXVOI+1E-8
+    res[[time]]$WIN_W <- res[[time]]$WIN_W*100 
     if(lo > min(res[[time]]$WIN_W)) lo <- min(res[[time]]$WIN_W)
     if(hi < max(res[[time]]$WIN_W)) hi <- max(res[[time]]$WIN_W)
   }
-  lo <- min(lo, 0.3)
-  hi <- min(hi, 0.7)
+  lo <- min(lo, 30)
+  hi <- min(hi, 70)
   
 
   plot(res[['5000']]$MAXVOI, res[['5000']]$WIN_W, ylim=c(lo,hi), log="x", type='n',
-       main=NA, xlab=expression(VOI[min]), ylab='UCT wins, %')
-  abline(h=0.5, lty='dashed')
+       main=NA, xlab=expression(VOI[min]+10^-8), ylab='UCT wins, %')
+  abline(h=50, lty='dashed')
   lines
   for(time in go.times) {
     lines(res[[time]]$MAXVOI, res[[time]]$WIN_W, type='o', pch=go.pchs[[time]])
@@ -101,6 +102,12 @@ draw.go.curves <- function(agent="vct") {
   legend(x='topleft', legend=go.times, pch=sapply(go.times, function(time) go.pchs[[time]]))
 }
 
-  
-  
-
+go.draw.bests <- function(bestsfile, colors=rainbow(3)) {
+  angles <- c(70, -45, 30)
+  densities <- c(20, 20, 30)
+  bests <- read.table(bestsfile, header=T)
+  barplot(t(as.matrix(bests[,-1])), names.arg = bests$NSMPLS,
+          density=densities, angle=angles, col=colors, beside=T)
+  legend(x="topright", names(bests)[-1], angle=angles,
+         density=densities, fill=colors, bty="n", ncol=3, bg=rainbow(3))
+}

### 2011-11-28 code/gobests.sh

@@ -1,0 +1,14 @@
+#!/bin/sh
+
+BESTWEIGHT=${1:-1.0}
+almostbest="python `dirname $0`/almostbest.py"
+
+echo NSMPLS BCT ECT VCT
+for time in 5000 7000 10000 15000; do
+ echo -n $time
+ for agent in bct ect vct; do
+  echo -n "" `$almostbest $BESTWEIGHT < uct$agent-time=$time.shomersu.dat`
+ done
+ echo ""
+done
+  

### 2011-11-28 paper/vago/paper.tex
vago paper stub

--- 2011-09-14 paper/voisam/paper.tex
@@
 
-\title{VOI-based Monte Carlo Sampling}
+\title{VOI-aware Monte Carlo Sampling in Trees}
 \author {David Tolpin, Solomon Eyal Shimony \\
@@
 \begin{abstract}
+Upper bounds for the VOI are provided for pure exploration in the
+Multi-armed Bandit Problem. Sampling policies based on the upper
+bounds are suggested. Empirical evaluation of the policies is provided
+on random problems as well on the Go game.
 \end{abstract}
 
-\section{Introduction and Definitions}
+\section{Empirical Evaluation}
 
-Taking a sequence of samples in order to minimize the
-regret of a decision based on the samples is abstracted by the
-{\em Multi-armed Bandit Problem.} In the Multi-armed Bandit problem
-we have a set of $K$ arms. Each arm can be pulled multiple
-times. When the $i$th arm is pulled, a random reward $X_i$ from an
-unknown stationary distribution is returned.  The reward is bounded
-between 0 and 1.
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.8]{vct-wins.pdf}
+\caption{Winning rate: UCT against VCT}
+\label{fig:uct-against-vct}
+\end{figure}
 
-The simple regret of a sampling policy for the Multi-armed Bandit
-Problem is the expected difference between the best expected reward
-$\mu_*$ and the expected reward $\mu_j$ of the arm with the best sample mean
-$\overline X_j=\max_i\overline X_i$:
-\begin{equation}
-\label{eq:simple-regret}
-\IE[R]=\sum_{j=1}^K\Delta_j\Pr(\overline X_j=\max_i\overline X_i)
-\end{equation}
-where $\Delta_j=\mu_*-\mu_j$.
-Strategies that minimize the simple regret are called pure exploration
-strategies \cite{Bubeck.pure}. Principles of rational metareasoning
-\cite{Russel.right} suggest that at each step the arm with the great
-value of information (VOI) must be pulled, and the sampling must be
-stopped and a decision must be made when no arm has positive VOI.
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.8]{ect-wins.pdf}
+\caption{Winning rate: UCT against ECT}
+\label{fig:uct-against-ect}
+\end{figure}
 
-To estimate the VOI of pulling an arm, either a certain 
-distribution of the rewards should be assumed (and updated based on
-observed rewards), or a distribution-independent bound on the VOI can be
-used as the VOI estimate. In this paper, we use {\em concentration inequalities}
-to derive distribution-independent bounds on the VOI.
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.8]{bct-wins.pdf}
+\caption{Winning rate: UCT against BCT}
+\label{fig:uct-against-ect}
+\end{figure}
 
-\section{Some Concentration Inequalities}
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.8]{bests-colorful.pdf}
+\caption{Best winning rate comparison}
+\label{fig:best-winning-rate}
+\end{figure}
 
-Let $X_1, \ldots, X_n$ be i.i.d. random variables with values from $[0,1]$,
-$X=\frac 1 n \sum_{i=1}^n X_i$. Then 
-\begin{description}
-\item[Hoeffding's inequality \rm{\cite{Hoeffding.ineq}}:] 
-\begin{equation}
-\Pr(X-\IE[X] \ge a) \le \exp ( -2na^2)
-\label{eq:conc-hoeffding}
-\end{equation}
-\item[Empirical Bernstein's inequality
-  \rm{\cite{MaurerPontil.benrstein}}:]\footnote{see
-    Appendix~\ref{app:deriv-conc-empbernstein} for derivation}
-\begin{eqnarray}
-\Pr(X-\IE[X] \ge a) &\le& 2\exp \left( - \frac {na^2} {\frac {14} {3}
-                          \frac {n} {n-1}a+2\overline\sigma_n^2}\right)\nonumber\\
-                    &<& 2\exp \left( - \frac {na^2} {10a+2\overline\sigma_n^2}\right)
-\label{eq:conc-empbernstein}
-\end{eqnarray}
-where sample variance $\overline\sigma_n^2$ is
-\begin{equation}
-\overline\sigma_n^2=\frac 1 {n(n-1)} \sum_{1\le i < j\le n}(X_i-X_j)^2
-\label{eq:sample-variance}
-\end{equation}
-\end{description}
-Bounds (\ref{eq:conc-hoeffding}, \ref{eq:conc-empbernstein}) are symmetrical
-around the mean. Bound~(\ref{eq:conc-empbernstein}) is tighter than
-(\ref{eq:conc-hoeffding}) for small $a$ and $\overline\sigma_n^2$.
-
-\section{Upper Bounds on Value of Information}
-
-\section{VOI-based Sampling Control}
-
-\subsection{Selection Criterion}
-
-\subsection{Termination Condition}
-
-\appendix
-
-\section{Empirical Bernstein Inequality}
-\label{app:deriv-conc-empbernstein}
-
-Theorem~4 in~\cite{MaurerPontil.benrstein} states that
-\[\Pr\left(\IE[X]-\overline X_n \ge \sqrt { \frac {2\overline\sigma_n^2 \ln 2/\delta} n } + \frac {7 \ln 2/\delta} {3(n-1)}\right)\le \delta,\]
-Therefore
-\[\Pr\left(\IE[X]-\overline X_n \ge \sqrt { \left(\frac {7 \ln 2/\delta} {3(n-1)}\right)^2+\frac {2\overline\sigma_n^2 \ln 2/\delta} n } + \frac {7 \ln 2/\delta} {3(n-1)}\right)\le
-\delta.\]
-$t=\sqrt { \left(\frac {7 \ln 2/\delta} {3(n-1)}\right)^2+\frac {2\overline\sigma_n^2 \ln 2/\delta} n } + \frac {7 \ln 2/\delta} {3(n-1)}$ is a root of
-square equation
-\[t^2-t\frac {14 \ln 2/\delta} {3(n-1)} -\frac {2\overline\sigma_n^2 \ln 2/\delta} n=0\]
-which, solved for $\delta\triangleq\Pr(\IE[X]-\overline X_n\ge t)$,
-gives
-\[\Pr(\IE[X]-\overline X_n\ge t)\le 2\exp \left( - \frac {na^2} {\frac {14} {3} \frac {n} {n-1}a+2\overline\sigma_n^2}\right)\]
-Other derivations, giving slightly different results, are possible.
+\clearpage
 

### 2011-11-28 code/figures.R

@@ -75,7 +75,7 @@ make.pdfs <- function() {
 
 go.times <- c('5000', '7000', '10000', '15000')
 go.pchs <- list('5000'=1, '7000'=2, '10000'=3, '15000'=4)
-draw.go.curves <- function(agent="vct") {
+go.draw.curves <- function(agent="vct") {
   res <- list()
   lo <- 100.0
   hi <- 0.0
@@ -102,9 +102,8 @@ draw.go.curves <- function(agent="vct") {
   legend(x='topleft', legend=go.times, pch=sapply(go.times, function(time) go.pchs[[time]]))
 }
 
-go.draw.bests <- function(bestsfile, colors=rainbow(3)) {
+go.draw.bests <- function(bestsfile, colors=rainbow(3), densities=c(20, 20, 30)) {
   angles <- c(70, -45, 30)
-  densities <- c(20, 20, 30)
   bests <- read.table(bestsfile, header=T)
   barplot(t(as.matrix(bests[,-1])), names.arg = bests$NSMPLS,
           density=densities, angle=angles, col=colors, beside=T)

### 2011-12-04 paper/vago/paper.tex
vago in progress

@@
 
-\newtheorem{lemma}{Lemma}
+\newtheorem{dfn}{Definition}
+\newtheorem{thm}{Theorem}
+\newtheorem{lmm}{Lemma}
 
@@
 \end{abstract}
+
+\section{Introduction and Definitions}
+
+Taking a sequence of samples in order to minimize the
+regret of a decision based on the samples is abstracted by the
+{\em Multi-armed Bandit Problem.} In the Multi-armed Bandit problem
+we have a set of $K$ arms. Each arm can be pulled multiple
+times. When the $i$th arm is pulled, a random reward $X_i$ from an
+unknown stationary distribution is returned.  The reward is bounded
+between 0 and 1.
+
+The simple regret of a sampling policy for the Multi-armed Bandit
+Problem is the expected difference between the best expected reward
+$\mu_*$ and the expected reward $\mu_j$ of the arm with the best sample mean
+$\overline X_j=\max_i\overline X_i$:
+\begin{equation}
+\label{eqn:simple-regret}
+\IE[R]=\sum_{j=1}^K\Delta_j\Pr(\overline X_j=\max_i\overline X_i)
+\end{equation}
+where $\Delta_j=\mu_*-\mu_j$.
+Strategies that minimize the simple regret are called pure exploration
+strategies \cite{Bubeck.pure}. Principles of rational metareasoning
+\cite{Russell.right} suggest that at each step the arm with the great
+value of information (VOI) must be pulled, and the sampling must be
+stopped and a decision must be made when no arm has positive VOI.
+
+To estimate the VOI of pulling an arm, either a certain 
+distribution of the rewards should be assumed (and updated based on
+observed rewards), or a distribution-independent bound on the VOI can be
+used as the VOI estimate. In this paper, we use {\em concentration inequalities}
+to derive distribution-independent bounds on the VOI.
+
+\section{Some Concentration Inequalities}
+
+Let $X_1, \ldots, X_n$ be i.i.d. random variables with values from $[0,1]$,
+$X=\frac 1 n \sum_{i=1}^n X_i$. Then 
+\begin{description}
+\item[Hoeffding's inequality \rm{\cite{Hoeffding.ineq}}:] 
+\begin{equation}
+\Pr(X-\IE[X] \ge a) \le \exp ( -2na^2)
+\label{eqn:conc-hoeffding}
+\end{equation}
+\item[Empirical Bernstein's inequality
+  \rm{\cite{MaurerPontil.benrstein}}:]\footnote{see
+    Appendix~\ref{app:deriv-conc-empbernstein} for derivation}
+\begin{eqnarray}
+\Pr(X-\IE[X] \ge a) &\le& 2\exp \left( - \frac {na^2} {\frac {14} {3}
+                          \frac {n} {n-1}a+2\overline\sigma_n^2}\right)\nonumber\\
+                    &\le& 2\exp \left( - \frac {na^2} {10a+2\overline\sigma_n^2}\right)
+\label{eqn:conc-empbernstein}
+\end{eqnarray}
+where sample variance $\overline\sigma_n^2$ is
+\begin{equation}
+\overline\sigma_n^2=\frac 1 {n(n-1)} \sum_{1\le i < j\le n}(X_i-X_j)^2
+\label{eqn:sample-variance}
+\end{equation}
+\end{description}
+Bounds (\ref{eqn:conc-hoeffding}, \ref{eqn:conc-empbernstein}) are symmetrical
+around the mean. Bound~(\ref{eqn:conc-empbernstein}) is tighter than
+(\ref{eqn:conc-hoeffding}) for small $a$ and $\overline\sigma_n^2$.
+
+\section{Upper Bounds on Value of Information}
+
+\begin{thm} The intrinsic value of perfect information $\Lambda_i^p$ about the $i$th arm is
+  bounded from above as
+\begin{equation}
+  \Lambda_i^p \le \left\{
+  \begin{array}{l l}
+    \Pr(\IE[X_i] \le \overline X_\beta)\overline X_\beta & \mbox{if $i=\alpha$} \\
+    \Pr(\IE[X_i] \ge \overline X_\alpha)(1-\overline X_\alpha) & \mbox{otherwise}
+  \end{array} \right.
+\label{eqn:thm-vopi}
+\end{equation}
+\end{thm}
+
+\begin{thm} The blinkered estimate $\Lambda_i^b$ of intrinsic value of information of sampling
+  the $i$th arm for the remaining budget of $N$ samples is bounded from above as
+\begin{equation}
+  \Lambda_i^b \le \left\{
+  \begin{array}{l l}
+    \Pr(\overline X_i'\le\overline X_\beta)\overline X_\beta\frac N
+       {N+n_i} < \Pr(\overline X_i'\le\overline X_\beta)\overline X_\beta\frac N
+       {n_i} & \mbox{if $i=\alpha$} \\
+    \Pr(\overline X_i'\ge\overline X_\alpha)(1-\overline X_\alpha)\frac N {N+n_i}&  \mbox{otherwise}
+  \end{array} \right.
+\label{eqn:thm-be}
+\end{equation}
+where $\overline X_i'$ is the sample mean of the $i$th arm after $N$ more
+samples.
+\end{thm}
+
+ The probabilities in equations (\ref{eqn:thm-vopi}, \ref{eqn:thm-be}) can be bounded from above using concentration
+inequalities. In particular, Lemma~\ref{lemma:hoeffding-prob-bounds} is
+based on the Hoeffding inequality (\ref{eqn:conc-hoeffding}):
+\begin{lmm} The probabilities in equations (\ref{eqn:thm-vopi}, \ref{eqn:thm-be}) are bounded from above as
+\begin{eqnarray}
+\Pr(\IE[X_i] \le \overline X_\beta|i=\alpha)& \le & \exp(-2 (\overline X_i - \overline X_\beta)^2 n_i)\nonumber\\
+\Pr(\IE[X_i] \ge \overline X_\alpha|i\ne\alpha)& \le & \exp(-2 (\overline X_\alpha - \overline X_i)^2 n_i)\nonumber\\
+\Pr(X_i' \le \overline X_\beta|i=\alpha)& \le & 2\exp(-0.5(\overline X_i - \overline X_\beta)^2 n_i)\nonumber\\
+\Pr(X_i' \ge \overline X_\alpha|i\ne\alpha)& \le & 2\exp(-0.5(\overline X_\alpha - \overline X_i)^2 n_i)
+\label{eqn:probound-perf-hoeffding}
+\end{eqnarray}
+\label{lemma:hoeffding-prob-bounds}
+\end{lmm}
+
+Better bounds can be obtained through tighter estimates on
+the probabilities, for example, based on the empirical Bernstein
+inequality (\ref{eqn:conc-empbernstein}).
 
@@
 \caption{Winning rate: UCT against BCT}
-\label{fig:uct-against-ect}
+\label{fig:uct-against-bct}
 \end{figure}
@@
 
+\appendix
+
+\section{Empirical Bernstein Inequality}
+\label{app:deriv-conc-empbernstein}
+
+Theorem~4 in~\cite{MaurerPontil.benrstein} states that
+\[\Pr\left(\IE[X]-\overline X_n \ge \sqrt { \frac {2\overline\sigma_n^2 \ln 2/\delta} n } + \frac {7 \ln 2/\delta} {3(n-1)}\right)\le \delta,\]
+Therefore
+\[\Pr\left(\IE[X]-\overline X_n \ge \sqrt { \left(\frac {7 \ln 2/\delta} {3(n-1)}\right)^2+\frac {2\overline\sigma_n^2 \ln 2/\delta} n } + \frac {7 \ln 2/\delta} {3(n-1)}\right)\le
+\delta.\]
+$a=\sqrt { \left(\frac {7 \ln 2/\delta} {3(n-1)}\right)^2+\frac {2\overline\sigma_n^2 \ln 2/\delta} n } + \frac {7 \ln 2/\delta} {3(n-1)}$ is a root of
+square equation
+\[a^2-a\frac {14 \ln 2/\delta} {3(n-1)} -\frac {2\overline\sigma_n^2 \ln 2/\delta} n=0\]
+which, solved for $\delta\triangleq\Pr(\IE[X]-\overline X_n\ge a)$,
+gives
+\[\Pr(\IE[X]-\overline X_n\ge a)\le 2\exp \left( - \frac {na^2} {\frac {14} {3} \frac {n} {n-1}a+2\overline\sigma_n^2}\right)\]
+Other derivations, giving slightly different results, are possible.
+
 \bibliographystyle{plain}

### 2011-12-04 paper/vago/paper.tex
paper in progress

@@
 bounds are suggested. Empirical evaluation of the policies is provided
-on random problems as well on the Go game.
+on random problems as well as on the Go game.
 \end{abstract}
@@
   \begin{array}{l l}
-    \Pr(\overline X_i'\le\overline X_\beta)\overline X_\beta\frac N
-       {N+n_i} < \Pr(\overline X_i'\le\overline X_\beta)\overline X_\beta\frac N
-       {n_i} & \mbox{if $i=\alpha$} \\
-    \Pr(\overline X_i'\ge\overline X_\alpha)(1-\overline X_\alpha)\frac N {N+n_i}&  \mbox{otherwise}
+    \Pr(\overline X_i'\le\overline X_\beta)\overline X_\beta\frac N {N+n_i} < \Pr(\overline X_i'\le\overline X_\beta)\overline X_\beta\frac N   {n_i} & \mbox{if $i=\alpha$} \\
+    \Pr(\overline X_i'\ge\overline X_\alpha)(1-\overline X_\alpha)\frac N {N+n_i} < \Pr(\overline X_i'\ge\overline X_\alpha)(1-\overline X_\alpha)\frac N {n_i}&  \mbox{otherwise}
   \end{array} \right.
@@
 \Pr(\IE[X_i] \ge \overline X_\alpha|i\ne\alpha)& \le & \exp(-2 (\overline X_\alpha - \overline X_i)^2 n_i)\nonumber\\
-\Pr(X_i' \le \overline X_\beta|i=\alpha)& \le & 2\exp(-0.5(\overline X_i - \overline X_\beta)^2 n_i)\nonumber\\
-\Pr(X_i' \ge \overline X_\alpha|i\ne\alpha)& \le & 2\exp(-0.5(\overline X_\alpha - \overline X_i)^2 n_i)
+\Pr(X_i' \le \overline X_\beta|i=\alpha)& \le & 2\exp\left(-\frac {(\overline X_i - \overline X_\beta)^2 n_i} 2 \right)\nonumber\\
+\Pr(X_i' \ge \overline X_\alpha|i\ne\alpha)& \le & 2\exp\left(-\frac {(\overline X_\alpha - \overline X_i)^2 n_i} 2 \right)
 \label{eqn:probound-perf-hoeffding}
@@
 the probabilities, for example, based on the empirical Bernstein
-inequality (\ref{eqn:conc-empbernstein}).
+inequality (\ref{eqn:conc-empbernstein}) or through a more careful
+application of the Hoeffding inequality (Appendix~\ref{app:better-hoeffding-bound}).
 
@@
 
+\section{VOI-based Sampling Control}
+
+\subsection{Selection Criterion}
+
+Following the principles of rational metareasoning, an arm with the highest upper bound $\hat
+\Lambda$ on the perfect value of information should be pulled at each
+step. This way, arms known to have a low VOI  would be pulled less frequently.
+
+\subsection{Termination Condition}
+
+The upper bounds~(\ref{eq:lambda-hoeffding-bounds}, \ref{eq:lambda-bernstein-bounds}) decrease exponentially with the
+number of pulls $n$. When the upper bound of the VOI for all arms
+becomes lower than a threshold $\lambda$, which can be chosen based on
+resource constraints, the sampling should be stopped, and an arm should
+be chosen.
+
+\subsection{Sample Redistribution in Trees}
+
+UCT forwards samples to the next search stage when the winner at the
+current state is known with high confidence. VOI-based termination
+condition can be used to stop the sampling in the current state early
+and save the remaining samples for re-use in a later search state.
+
 \clearpage
@@
 
+\section{Better Hoeffding-Based Bound on Value of Perfect Information}
+\label{app:better-hoeffding-bound}
+
+The bound can be supposedly be improved by selecting a midpoint
+$0 < \gamma < \overline X_\beta$ and computing the bound as the sum of two parts:
+\begin{itemize}
+\item $\overline X_\beta-\gamma$ multiplied by the probability that
+  $\mu_\alpha \le \overline X_\beta$;
+\item $\overline X_\beta$ multiplied by the probability that $\mu_\alpha\le
+  \gamma$.
+\end{itemize}.
+\[ V_{\overline X=\overline X_\alpha} \le (\overline X_\beta-\gamma)\exp\left(-2n(\overline X_\alpha-\overline X_\beta)^2\right)+\overline X_\beta\exp\left(-2n(\overline X_\alpha-\gamma)^2\right) \]
+
+\vspace{\baselineskip}
+
+The minimum of $V^*$ is achieved when $\frac {dV^*} {d\gamma}=0$, that
+is, when $\gamma$ is the root of the following equation:
+\[ 4\overline X_\beta n(\overline X_\alpha-\gamma)=\exp\left(-2n\left(\frac {\overline X_\alpha-\overline X_\beta}
+    {\overline X_\alpha-\gamma}\right)^2\right) \]
+If a root in the interval  $0\le\gamma\le\beta$ exists, then the
+number of samples is bounded as
+\[n\le\frac 1 {4\overline X_\beta(\overline X_\alpha-\overline X_\beta)}\]
+by observing that the right-hand side 
+is at most $1$ (a negative power), and the left-hand side is at least
+$4\overline X_\beta n(\overline X_\alpha-\overline X_\beta)$.
+So, the bound can supposedly be improved for smaller values of
+$n$. The improvement is more significant when the current best and
+second-best sample means are close.
+
+The derivation for the other case
+(sampling an item that can be better than the current best) is
+obtained by substitution $1-\overline X, 1-\overline X_\alpha, 1-\gamma$ instead of
+$\overline X_\alpha, \overline X_\beta, \gamma$:
+
+\[ V_{\overline X\ne\overline X_\alpha} \le (\gamma-\overline X_\alpha)\exp\left(-2n(\overline X_\alpha-\overline
+  X)^2\right)+(1-\overline X_\alpha)\exp\left(-2n(\gamma-\overline X)^2\right) \]
+
+The anticipated influence of the improved
+VOI estimate would be that the selected item will be a less discovered
+one and further from the current best or second-best.
+
+A closed-form solution for $\gamma$ cannot be obtained, but given
+$\overline X_\alpha, \overline X_\beta, n$, the value of $\gamma$ can be efficiently
+computed. It should be determined empirically whether the improved
+estimate has justified influence on the performance of the algorithm.
+
 \bibliographystyle{plain}

### 2011-12-04 paper/vago/paper.tex
paper in progress

@@
 
+The intrinsic VOI of pulling an arm is the expected decrease
+in the regret compared to selecting an arm without pulling any arm at
+all. The intrinsic VOI can be estimated as the intrinsic 
+value of perfect information $\Lambda_i$ about the mean reward of the $i$th arm. Two
+cases are possible:
+\begin{itemize}
+\item the arm ($\alpha$) with the highest sample mean is pulled, and the 
+mean of the arm is lower than the sample mean the second-best arm ($\beta$);
+\item another arm is pulled, and the mean of the arm is higher
+than the current highest sample mean ($\alpha$).
+\end{itemize}
+$\Lambda_i$ can be bounded from above as the probability that a different arm is selected times the
+maximum possible increase in the reward:
+
 \begin{thm} The intrinsic value of perfect information $\Lambda_i^p$ about the $i$th arm is
@@
 \Pr(\IE[X_i] \ge \overline X_\alpha|i\ne\alpha)& \le & \exp(-2 (\overline X_\alpha - \overline X_i)^2 n_i)\nonumber\\
-\Pr(X_i' \le \overline X_\beta|i=\alpha)& \le & 2\exp\left(-\frac {(\overline X_i - \overline X_\beta)^2 n_i} 2 \right)\nonumber\\
-\Pr(X_i' \ge \overline X_\alpha|i\ne\alpha)& \le & 2\exp\left(-\frac {(\overline X_\alpha - \overline X_i)^2 n_i} 2 \right)
+\Pr(\overline X_i' \le \overline X_\beta|i=\alpha)& \le &  2\exp\left(-\frac {2(\overline X_i - \overline X_\beta)^2  (n_i+N) n_i} {(\sqrt n_i + \sqrt {n_i + N})^2}  \right) \le 2\exp\left(-\frac {(\overline X_i - \overline X_\beta)^2 n_i} 2 \right)\nonumber\\
+\Pr(\overline X_i' \ge \overline X_\alpha|i\ne\alpha)& \le & 2\exp\left(-\frac {(\overline X_\alpha - \overline X_i)^2 n_i} 2 \right)
 \label{eqn:probound-perf-hoeffding}
@@
 
+\subsection{Selecting best arm}
+
+Figure~\ref{fig:random-instances}
+
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.8]{flat.pdf}
+\caption{Random instances: regret vs. number of samples}
+\label{fig:random-instances}
+\end{figure}
+
+\subsection{Playing Go against UCT}
+
+Figure~\ref{fig:uct-against-vct}, Figure~\ref{fig:uct-against-ect}, Figure~\ref{fig:uct-against-bct},
+Figure~\ref{fig:best-winning-rate}.
+
 \begin{figure}[h]
@@
 \includegraphics[scale=0.8]{vct-wins.pdf}
-\caption{Winning rate: UCT against VCT}
+\caption{Go: winning rate --- UCT against VCT}
 \label{fig:uct-against-vct}
@@
 \includegraphics[scale=0.8]{ect-wins.pdf}
-\caption{Winning rate: UCT against ECT}
+\caption{Go: winning rate --- UCT against ECT}
 \label{fig:uct-against-ect}
@@
 \includegraphics[scale=0.8]{bct-wins.pdf}
-\caption{Winning rate: UCT against BCT}
+\caption{Go: winning rate --- UCT against BCT}
 \label{fig:uct-against-bct}
@@
 \includegraphics[scale=0.8]{bests-colorful.pdf}
-\caption{Best winning rate comparison}
+\caption{Go: best winning rate comparison}
 \label{fig:best-winning-rate}

### 2011-12-04 code/figures.R

@@ -8,7 +8,7 @@ draw.regrets <- function(regrets_file, log="xy", legend.position='topright', xla
 	ylim <- range(values[values>0])
 	plot(x=regrets$nsamples, type='n', log=log, xlim=xlim, ylim=ylim, yaxs='i', xlab=xlab, ylab='Regret', main=if(is.na(main)) unlist(strsplit(regrets_file, "\\."))[[1]] else main)
 	i <- 1
-	for(alg in names(regrets)[-1]) {
+	for(alg in algs) {
 		y <- regrets[[alg]]
 		lines(x=regrets$nsamples, y=regrets[[alg]],
 			  type="o", pch=i)

### 2011-12-05 paper/vago/paper.tex
paper in even more progress
paper in progress

@@
 Multi-armed Bandit Problem. Sampling policies based on the upper
-bounds are suggested. Empirical evaluation of the policies is provided
-on random problems as well as on the Go game.
+bounds are suggested. Empirical evaluation of the policies and
+comparison to the UCB1 and UCT policies is provided
+on random problem instances as well as on the Go game.
 \end{abstract}
@@
 
-The intrinsic VOI of pulling an arm is the expected decrease
+The intrinsic VOI $\Lambda_i$ of pulling an arm is the expected decrease
 in the regret compared to selecting an arm without pulling any arm at
-all. The intrinsic VOI can be estimated as the intrinsic 
-value of perfect information $\Lambda_i$ about the mean reward of the $i$th arm. Two
+all. The \textit{myopic} VOI estimate is of limited applicability to
+Monte Carlo sampling, since the effect of a single sample is small,
+and the myopic VOI estimate will often be non-positive, resulting in premature
+termination of the search. However, $\Lambda_i$ can be estimated as the intrinsic 
+value of perfect information $\Lambda_i^p$ about the mean reward of the $i$th arm. Two
 cases are possible:
@@
 \end{itemize}
-$\Lambda_i$ can be bounded from above as the probability that a different arm is selected times the
+$\Lambda_i^p$ can be bounded from above as the probability that a
+different arm is selected, multiplied by the
 maximum possible increase in the reward:
@@
 
-\begin{thm} The blinkered estimate $\Lambda_i^b$ of intrinsic value of information of sampling
-  the $i$th arm for the remaining budget of $N$ samples is bounded from above as
+The search time is finite, and in a simple case the \textit{search
+  budget} specified as the maximum number of samples.
+An estimate based on the perfect intrinsic VOI does not take in
+consideration the remaining number of samples. Given two arms
+with the same intrinsic perfect VOI, the VOI
+estimate of the arm pulled fewer times so far should be higher.
+
+\begin{dfn} The \textbf{blinkered estimate} of intrinsic VOI information of the
+  $i$th arm is the intrinsic VOI of pulling the $i$th arm for the
+  remaining budget.
+\end{dfn}
+
+\begin{thm} Denote the current number of samples of the $i$th arm as
+  $n_i$. The blinkered estimate $\Lambda_i^b$ of intrinsic value of
+  information of pulling the $i$th arm for the remaining budget of $N$
+  samples is bounded from above as
 \begin{equation}
@@
   \begin{array}{l l}
-    \Pr(\overline X_i'\le\overline X_\beta)\overline X_\beta\frac N {N+n_i} < \Pr(\overline X_i'\le\overline X_\beta)\overline X_\beta\frac N   {n_i} & \mbox{if $i=\alpha$} \\
-    \Pr(\overline X_i'\ge\overline X_\alpha)(1-\overline X_\alpha)\frac N {N+n_i} < \Pr(\overline X_i'\ge\overline X_\alpha)(1-\overline X_\alpha)\frac N {n_i}&  \mbox{otherwise}
+    \Pr(\overline X_i'\le\overline X_\beta)\overline X_\beta\frac N {N+n_i} & \mbox{if $i=\alpha$} \\
+    \Pr(\overline X_i'\ge\overline X_\alpha)(1-\overline  X_\alpha)\frac N {N+n_i} &  \mbox{otherwise}
   \end{array} \right.
@@
 \end{equation}
-where $\overline X_i'$ is the sample mean of the $i$th arm after $N$ more
+where $\overline X_i'$ is the sample mean of the $i$th arm after $n_i+N$ 
 samples.
@@
 \Pr(\IE[X_i] \ge \overline X_\alpha|i\ne\alpha)& \le & \exp(-2 (\overline X_\alpha - \overline X_i)^2 n_i)\nonumber\\
-\Pr(\overline X_i' \le \overline X_\beta|i=\alpha)& \le &  2\exp\left(-\frac {2(\overline X_i - \overline X_\beta)^2  (n_i+N) n_i} {(\sqrt n_i + \sqrt {n_i + N})^2}  \right) \le 2\exp\left(-\frac {(\overline X_i - \overline X_\beta)^2 n_i} 2 \right)\nonumber\\
-\Pr(\overline X_i' \ge \overline X_\alpha|i\ne\alpha)& \le & 2\exp\left(-\frac {(\overline X_\alpha - \overline X_i)^2 n_i} 2 \right)
+\Pr(\overline X_i' \le \overline X_\beta|i=\alpha)& \le &  2\exp\left(-\frac {2(\overline X_i - \overline X_\beta)^2  (n_i+N) n_i} {(\sqrt n_i + \sqrt {n_i + N})^2}  \right) < 2\exp\left(-\frac {(\overline X_i - \overline X_\beta)^2 n_i} 2 \right)\nonumber\\
+\Pr(\overline X_i' \ge \overline X_\alpha|i\ne\alpha)& \le & 2\exp\left(-\frac {(\overline X_\alpha - \overline X_i)^2 (n_i+N) n_i} {(\sqrt n_i + \sqrt {n_i + N})^2} \right) <  2\exp\left(-\frac {(\overline X_\alpha - \overline X_i)^2 n_i} 2 \right)
 \label{eqn:probound-perf-hoeffding}
@@
 
+\section{VOI-based Sampling Control}
+
+\subsection{Selection Criterion}
+
+Following the principles of rational metareasoning, an arm with the highest upper bound $\hat
+\Lambda$ on the perfect value of information should be pulled at each
+step. This way, arms known to have a low VOI  would be pulled less frequently.
+
+\subsection{Termination Condition}
+
+The upper bounds~(\ref{eq:lambda-hoeffding-bounds}, \ref{eq:lambda-bernstein-bounds}) decrease exponentially with the
+number of pulls $n$. When the upper bound of the VOI for all arms
+becomes lower than a threshold $\lambda$, which can be chosen based on
+resource constraints, the sampling should be stopped, and an arm should
+be chosen.
+
+\subsection{Sample Redistribution in Trees}
+
+UCT forwards samples to the next search stage when the winner at the
+current state is known with high confidence. VOI-based termination
+condition can be used to stop the sampling in the current state early
+and save the remaining samples for re-use in a later search state.
+
 \section{Empirical Evaluation}
 
-\subsection{Selecting best arm}
+\subsection{Selecting The Best Arm}
 
@@
 
-\subsection{Playing Go against UCT}
+\subsection{Playing Go Against UCT}
 
@@
 
-\section{VOI-based Sampling Control}
-
-\subsection{Selection Criterion}
-
-Following the principles of rational metareasoning, an arm with the highest upper bound $\hat
-\Lambda$ on the perfect value of information should be pulled at each
-step. This way, arms known to have a low VOI  would be pulled less frequently.
-
-\subsection{Termination Condition}
-
-The upper bounds~(\ref{eq:lambda-hoeffding-bounds}, \ref{eq:lambda-bernstein-bounds}) decrease exponentially with the
-number of pulls $n$. When the upper bound of the VOI for all arms
-becomes lower than a threshold $\lambda$, which can be chosen based on
-resource constraints, the sampling should be stopped, and an arm should
-be chosen.
-
-\subsection{Sample Redistribution in Trees}
-
-UCT forwards samples to the next search stage when the winner at the
-current state is known with high confidence. VOI-based termination
-condition can be used to stop the sampling in the current state early
-and save the remaining samples for re-use in a later search state.
-
 \clearpage
@@
 The bound can be supposedly be improved by selecting a midpoint
-$0 < \gamma < \overline X_\beta$ and computing the bound as the sum of two parts:
+$0 < y < \overline X_\beta$ and computing the bound as the sum of two parts:
 \begin{itemize}
-\item $\overline X_\beta-\gamma$ multiplied by the probability that
+\item $\overline X_\beta-y$ multiplied by the probability that
   $\mu_\alpha \le \overline X_\beta$;
 \item $\overline X_\beta$ multiplied by the probability that $\mu_\alpha\le
-  \gamma$.
+  y$.
 \end{itemize}.
-\[ V_{\overline X=\overline X_\alpha} \le (\overline X_\beta-\gamma)\exp\left(-2n(\overline X_\alpha-\overline X_\beta)^2\right)+\overline X_\beta\exp\left(-2n(\overline X_\alpha-\gamma)^2\right) \]
+\[ V_{\overline X=\overline X_\alpha} \le (\overline X_\beta-y)\exp\left(-2n(\overline X_\alpha-\overline X_\beta)^2\right)+\overline X_\beta\exp\left(-2n(\overline X_\alpha-y)^2\right) \]
 
@@
 
-The minimum of $V^*$ is achieved when $\frac {dV^*} {d\gamma}=0$, that
-is, when $\gamma$ is the root of the following equation:
-\[ 4\overline X_\beta n(\overline X_\alpha-\gamma)=\exp\left(-2n\left(\frac {\overline X_\alpha-\overline X_\beta}
-    {\overline X_\alpha-\gamma}\right)^2\right) \]
-If a root in the interval  $0\le\gamma\le\beta$ exists, then the
+The minimum of $V^*$ is achieved when $\frac {dV^*} {dy}=0$, that
+is, when $y$ is the root of the following equation:
+\[ 4\overline X_\beta n(\overline X_\alpha-y)=\exp\left(-2n\left(\frac {\overline X_\alpha-\overline X_\beta}
+    {\overline X_\alpha-y}\right)^2\right) \]
+If a root in the interval  $0\le y\le\beta$ exists, then the
 number of samples is bounded as
@@
 (sampling an item that can be better than the current best) is
-obtained by substitution $1-\overline X, 1-\overline X_\alpha, 1-\gamma$ instead of
-$\overline X_\alpha, \overline X_\beta, \gamma$:
+obtained by substitution $1-\overline X, 1-\overline X_\alpha, 1-y$ instead of
+$\overline X_\alpha, \overline X_\beta, y$:
 
-\[ V_{\overline X\ne\overline X_\alpha} \le (\gamma-\overline X_\alpha)\exp\left(-2n(\overline X_\alpha-\overline
-  X)^2\right)+(1-\overline X_\alpha)\exp\left(-2n(\gamma-\overline X)^2\right) \]
+\[ V_{\overline X\ne\overline X_\alpha} \le (y-\overline X_\alpha)\exp\left(-2n(\overline X_\alpha-\overline
+  X)^2\right)+(1-\overline X_\alpha)\exp\left(-2n(y-\overline X)^2\right) \]
 
@@
 
-A closed-form solution for $\gamma$ cannot be obtained, but given
-$\overline X_\alpha, \overline X_\beta, n$, the value of $\gamma$ can be efficiently
+A closed-form solution for $y$ cannot be obtained, but given
+$\overline X_\alpha, \overline X_\beta, n$, the value of $y$ can be efficiently
 computed. It should be determined empirically whether the improved

### 2011-12-07 paper/rmcts.arxiv/paper.tex
initial commit

--- 2011-09-13 paper/rmcts/paper.tex
@@
 where $\Delta_j=\mu_*-\mu_j$.
-Strategies that minimize the simple regret are called pure exploration
+Strategies that minimize the simple reward are called pure exploration
 strategies \cite{Bubeck.pure}.
@@
 \item The $\varepsilon$-greedy allocation scheme has exponential convergence
-rate of the simple regret $\IE r$ (see Appendix~\ref{app:derivations-finite-eps}):
+rate of the simple regret $\IE r$:
 \begin{equation}
@@
 \item The UCB allocation scheme exhibits polynomial convergence
-rate (see Appendix~\ref{app:derivations-finite-ucb}):
+rate:
 \begin{equation}
@@
 2$, and for a large number of arms the bound for the $\frac 1 2$-greedy scheme approaches the square of the
-bound for uniform sampling (see
-Appendix~\ref{app:derivations-asym-eps}).
+bound for uniform sampling.
 
@@
 instead of $\log(\cdot)$ into the formula for UCB and has a
-superpolynomial convergence rate (see
-Appendix~\ref{app:derivations-asym-ucb}).
+superpolynomial convergence rate.
 \end{enumerate}
+
+Proofs and derivations for the above statements are in the full paper,
+removed for brevity.
 
@@
 
-\begin{figure}[t]
-  \centering
-  \includegraphics[scale=0.45,trim=0 0 0 16pt,clip]{flat-trilevel-vct-k=64.pdf}\\
-  \caption{VOI-based action selection (VCT) vs. other algorithms}
-  \label{fig:flat-trilevel-vct}
-\end{figure}
-
 VOI of a rollout can be computed when the sample distribution of an
@@
 improvement for mixed arms (e.g. Bernoulli, triangularly distributed,
-and fixed arms) (Figure~\ref{fig:flat-trilevel-vct}):
-\textbf{VCT}, the algorithm which selects an action with the maximum VOI estimate on
-the first step exhibits lower average regret than other algorithms.
-Sample-based VOI estimation requires keeping more information than an
-algorithm based only on the sample
+and fixed arms. Sample-based VOI estimation requires keeping more
+information than an algorithm based only on the sample
 mean, but the overhead is small and justified by the improvement.
@@
 
-\clearpage
-\appendix
-
-\section{Facts}
-
-{\bf Chernoff bounds (see \cite{Hagerup.chernoff}):} for $m$ independent random variables $X_1, X_2, ..., X_m$
-taking on values 0 or 1, $X=\sum_{i=1}^m X_i$, and $0\le\delta\le 1$:
-\begin{equation}
-\Pr[X < (1-\delta)\IE[X]] < e^{-\delta^2\IE[X]/2} \mbox{ (Chernoff bound)}
-\label{eq:chernoff-bound}
-\end{equation}
-Further on, if $\forall i. \IE[X_i]=\mu$:
-\begin{equation}
-\Pr[X < n\mu-a] <  e^{-2a^2/n}
-\label{eq:chernoff-hoeffding-bound} \mbox{ (Hoeffding-Chernoff bound)}
-\end{equation}
-
-\section{Derivations}
-\label{app:derivations}
-
-\subsection{Finite-time regret bounds}
-\label{app:derivations-finite}
-
-%TODO: bounds on regret based on information gathered, independent of
-%the way the choice is made.
-
-For an analysis of pure exploration of uniform and UCB algorithms, see also
-\cite{Bubeck.pure}.
-
-\subsubsection{$\varepsilon$-greedy}
-\label{app:derivations-finite-eps}
-
-For $0<\varepsilon\le1-\frac 1 K$ and $x_0>0$ (see Section 3, Proofs, in \cite{Auer.ucb}):
-
-\begin{eqnarray}
-\Pr(\overline X_j=\max_i\overline X_i)&\le&2\left(x_0\cdot \Pr\left\{T_j^R(n)\le x_0\right\} + \frac 2{\Delta_j^2}e^{-\Delta_j^2\lfloor x_0 \rfloor/2}\right)\nonumber\\
-&\le&2\left(x_0\cdot \Pr\left\{T_j^R(n)\le x_0\right\} + \frac {2 \cdot
-  e^{\Delta_j^2/2}}{\Delta_j^2}e^{-\Delta_j^2 x_0 /2}\right)
-\end{eqnarray}
-
-Number of times $\xi$ the $j$th arm is selected {\it randomly}:
-
-\begin{equation}
-\xi\triangleq\IE\left[T_j^R(n)\right]=\frac {n\varepsilon} K
-\end{equation}
-
-By choosing $x_0=\frac \xi 2$ and applying the Chernoff bound (\ref{eq:chernoff-bound}), obtain:
-
-\begin{eqnarray}
-\label{eq:pr-epsilon-greedy}
-\Pr(\overline X_j=\max_i\overline X_i)&\le& 2\left(\frac {\xi}{2} e^{-\xi/8} + \frac {2e^{\Delta_j^2/2}}{\Delta_j^2}e^{-\Delta_j^2 \xi/4}\right)\nonumber\\
-&\le&\left(\xi + \frac {4\sqrt e}{\Delta_j^2}\right)e^{-\Delta_j^2\xi/8}
-\end{eqnarray}
-
-A bound on the simple regret $r_\varepsilon$ of an
-$\varepsilon$-greedy policy follows from substituting
-(\ref{eq:pr-epsilon-greedy}) into (\ref{eq:simple-regret}):
-
-\begin{eqnarray}
-  \IE r_\varepsilon&\le&\sum_{j=1}^K\Delta_j\left(\xi + \frac {4\sqrt e}
-{\Delta_j^2}\right)e^{-\Delta_j^2\xi/8}\nonumber\\
-&\le&K\left(\xi + \frac {4\sqrt
-    e}{\Delta_{\min}}\right)e^{-\Delta_{\min}^2\xi/8}
-\label{eq:deriv-epsilon-greedy-b}
-\end{eqnarray}
-
-\subsubsection{UCB($\alpha$)}
-\label{app:derivations-finite-ucb}
-
-$n_*$ pulls of the best arm and $n_j$ pulls of the $j$th arm are
-achieved simultaneously when either
-
-\begin{equation}
-\sqrt {\frac {\alpha \log n} {n_j-1}} \ge 1 + \sqrt { \frac {\alpha \log n} {n_*}}
-\end{equation}
-
-or
-
-\begin{equation}
-\sqrt {\frac {\alpha \log n} {n_j}} \le 1 + \sqrt { \frac {\alpha \log n} {n_* - 1}}
-\label{eq:deriv-ucb-alpha-a}
-\end{equation}
-
-Inequality (\ref{eq:deriv-ucb-alpha-a}) can be used to derive a lower
-bound on the number of pulls of a non-optimal arm,  and, consequently, the upper
-bound on the probability of selecting a non-optimal arm:
-
-\begin{eqnarray}
-\frac {\alpha \log n} {n_j}&\le& 1 + 2  \sqrt {\frac {\alpha \log n}
-  {n_*-1}} + \frac {\alpha \log n} {n_*-1}\nonumber\\
-n_j &\ge& \frac {\alpha \log n} {1+2 \sqrt {\frac {\alpha \log n}
-    {n_*-1}} + \frac {\alpha \log n} {n_*-1}}\nonumber\\
-    &\ge& \frac {\alpha \log n} {4} \mbox{ for } \alpha \log n \le n_*-1
-\end{eqnarray}
-
-The probability that the $j$th arm will be chosen can be bounded using
-the Chernoff-Hoeffding bound:
-\begin{equation}
-P_j \le 2 \exp\left(-\frac {\alpha \Delta_j^2 \log n} 8 \right) = 2n^{\frac {-\alpha \Delta_j^2} 8}
-\end{equation}
-
-The expected simple regret of an UCB($\alpha$) algorithm is thus
-bounded by the following inequality:
-\begin{equation}
-\IE r_{ucb} \le 2\sum_{j=1}^K \Delta_jn^{\frac {-\alpha \Delta_j^2} 8}
-\label{eq:deriv-ucb-alpha-b}
-\end{equation}
-
-\cite{Bubeck.pure} establishes for UCB($\alpha$) a similar bound,
-polynomial in the number of pulls $n$.
-
-\subsection{Asymptotic regret analysis}
-\label{app:derivations-asym}
-
-%TODO: Oracle instead of asymptotic
-
-Bounds (\ref{eq:deriv-epsilon-greedy-b}, \ref{eq:deriv-ucb-alpha-b}) were
-derived ignoring dependency of the allocation on expected values
-of non-optimal arms. Derivation of better bounds seems
-complicated. Instead, one can analyze the asymptotic behavior of each of the
-algorithms under the assumption that the sample mean is close to the expectation.
-
-\subsubsection{$\varepsilon$-greedy}
-\label{app:derivations-asym-eps}
-
-A common technique \cite{Auer.ucb} to bound the probability of
-selecting a non-optimal arm $j$ is to split the interval between
-$\mu_j$ and $\mu_*$ at the middle and bound the probability of
-selecting the $j$th arm by the sum of the probabilities that the sample
-mean $\overline X_j$ of the $j$th arm is greater than $\mu_j+\frac
-{\Delta_j} 2$ and the sample mean $\overline X_*$ of the optimal arm is
-less than $\mu_*- \frac {\Delta_j} 2$. Each of the probabilities can
-be bound using the Chernoff-Hoefding bound.
-
-Splitting the interval between $\mu_j$ and $\mu_*$ half-way is an
-arbitrary decision. By selecting a different split point
-depending on the allocation of samples, one can obtain a tighter
-bound. On the other hand, there is a value of $\varepsilon$ that
-minimizes the bound for the $\varepsilon$-greedy allocation scheme.
-Thus, the  $\varepsilon$-greedy allocation scheme can be tuned by
-finding $\varepsilon$ that minimizes the bound obtained by
-selecting the best split point.
-
-Split the interval $[\mu_j, \mu_*]$ at $\mu_j+\delta_j$. The
-probability $P_j$ for selecting the
-$j$th arm is bounded, using the union bound and the Chernoff-Hoeffding bound, as:
-
-\begin{eqnarray}
-P_j&\le&\Pr[\overline X_j>\mu_j+\delta_j]+\Pr[\overline X_*<\mu_*-(\Delta_j-\delta_j)]\nonumber\\
-   &\le&\exp\left(-2\delta_j^2n_j\right)+\exp\left(-2(\Delta_j-\delta_j)^2n_*\right)
-\end{eqnarray}
-
-For the $\varepsilon$-greedy scheme, $n_i=\frac {\varepsilon n} K$,
-$n_*=(1-\varepsilon)n$, thus
-
-\begin{equation}
-P_j \le \exp\left(-\frac {2\delta_j^2\varepsilon n} {K-1}\right) +\exp\left(-2(\Delta_j-\delta_j)^2(1-\varepsilon)n\right)
-\end{equation}
-
-In general, $\varepsilon$ and $\delta$ which minimize $P_j$ will
-depend of $n$. For a constant $\varepsilon$, require that the
-exponents in both terms are equal
-\begin{eqnarray}
-\exp\left(-\frac {2\delta_j^2\varepsilon n} {K-1}\right)
-   &=& \exp\left(-2(\Delta_j-\delta_j)^2(1-\varepsilon)n\right)\\
-\frac {\delta_j} {\Delta_j - \delta_j}
-   &=& \sqrt { \frac {(K-1)(1-\varepsilon)} \varepsilon }
-\label{eq:deriv-exps-equal}
-\end{eqnarray}
-and find conditional minima of $P_j(\delta_j, \epsilon)$ satisfying
-either $\frac {\partial P_j} {\partial \delta_j}=0$ or $\frac {\partial P_j}
-{\partial \varepsilon}=0$.
-
-\vspace{1em}
-\paragraph{a. {\large $\frac {\partial P_j} {\partial \delta_j}=0$}}
-
-\begin{eqnarray*}
-\frac {4\delta_j\varepsilon n} {K-1}
-  &-& 4(\Delta_j-\delta_j)(1-\varepsilon)n=0\\
-\frac {\delta_j} {\Delta_j-\delta_j}&=&{ \frac {(K-1)(1-\varepsilon)} \varepsilon }
-\end{eqnarray*}
-and together with (\ref{eq:deriv-exps-equal}) this gives:
-\begin{eqnarray}
-\delta_j&=&\frac {\Delta_j} 2\nonumber\\
-\varepsilon&=&1-\frac 1 K\nonumber\\
-\IE r_{\varepsilon=1-1/K}&\le&B_{uniform}=2\exp\left(-\frac {\Delta_j^2n} {2K}\right)
-\label{eq:deriv-uniform-sampling}
-\end{eqnarray}
-
-That is, the case $\frac {\partial P_j} {\partial \delta_j}=0$
-corresponds to \emph{uniform random sampling}, and the tightest bound is obtained
-by splitting the interval at the middle.
-
-\paragraph{b. {\large $\frac {\partial P_j} {\partial \varepsilon}=0$}}
-
-\begin{eqnarray*}
-\frac {2\delta_j^2 n} {K-1}
-  &-& 2(\Delta_j-\delta_j)^2n=0\\
-\frac {\delta_j} {\Delta_j-\delta_j}&=&\sqrt {K-1}
-\end{eqnarray*}
-and together with (\ref{eq:deriv-exps-equal}) this gives:
-\begin{eqnarray}
-\frac {\delta_j} {\Delta_j}&=&\frac 1 {\sqrt {K-1}+1}\nonumber\\
-\varepsilon&=&\frac 1 2\nonumber\\
-\IE
-r_{\varepsilon=1/2}&\le&B_{1/2\mbox{-}greedy}=2\exp\left(-\frac
-  {\Delta_j^2n} {(\sqrt{K-1}+1)^2}\right)\nonumber\\
-\lim_{K\to\infty}
-\mathbb{B}r_{1/2\mbox{-}greedy}^K&=&2\exp\left(-{\Delta_j^2n}\right)\nonumber\\
-\mbox{for large $K$: }\mathbb{B}r_{1/2\mbox{-}greedy}&\approx&2\exp\left(-\frac {\Delta_j^2n} K\right)
-\label{eq:deriv-half-greedy-sampling}
-\end{eqnarray}
-
-That is, the case $\frac {\partial P_j} {\partial \varepsilon}=0$
-corresponds to \emph{$\frac 1 2$-greedy sampling}, and for a large number of
-arms $K$ 
-\begin{equation}
-\mathbb{B}r_{1/2\mbox{-}greedy}\approx \mathbb{B}r_{uniform}^2
-\label{eq:derive-uniform-vs-half-greedy}
-\end{equation}
-Equation (\ref{eq:derive-uniform-vs-half-greedy}) suggests
-that \emph{for a large number of arms} $\frac 1 2$-greedy
-sampling has a much higher convergence rate than uniform
-sampling.
-
-\subsubsection{UCB($\alpha$)}
-\label{app:derivations-asym-ucb}
-
-The $UCB(\alpha)$ allocation scheme distributes arm pulls such that
-the upper bound on the probability of selecting a non-optimal arm is
-asymptotically the same for all arms. Indeed, for large n
-\begin{eqnarray}
-n_i&\ll&n\\
-n_*&\approx&n\\
-\sqrt {\frac {2 \log n} {n_j}}&\approx&\Delta_j+\sqrt {\frac {2\log n}
-  {n_*}} \\
-&\Rightarrow& n_j \approx \frac {2\log n} {\Delta_j^2}
-\label{eq:deriv-ucb-asymptotic-nj}
-\end{eqnarray}
-and by Hoeffding-Chernoff bound the upper bound $\mathbb{B}P_j$ on the probability
-$P_j^{ucb}$ to choose the $j$th arm is
-\begin{equation}
-P_j^{ucb}\le\mathbb{B}P_j^{ucb}\approx2\exp\left(-2\Delta_j^2\frac {2\log n}{\Delta_j^2}\right)
-             =2n^{-4}
-\end{equation}
-
-That is, UCB($\alpha$) supposedly distributes pulls of non-optimal
-arms better than $\varepsilon$-greedy
-scheme which may undersample arms with high expectation. The
-polynomial, rather than exponential, convergence rate (Equation
-\ref{eq:deriv-ucb-alpha-b}) is due to oversampling the best arm.
-
-Replacing the logarithm with a faster growing subpolynomial
-function, like square root (UQB), results in a superpolynomial convergence
-rate of the upper bound:
-\begin{equation}
-P_j^{uqb}\le\mathbb{B}P_j^{uqb}\approx2\exp\left(-\beta\Delta_j^2\frac {\sqrt n}{\Delta_j^2}\right)
-            =2\exp(-\beta\sqrt n)
- \end{equation}
-where $\beta$ is a constant that depends only on reward bounds.
-Finite time convergence can still be shown similarly to UCB.
-
 \bibliographystyle{plain}

### 2011-12-07 paper/vago/paper.tex
better bounds on blinkered voi
initial commit

@@
 
+\section{Related Work}
+
+\textit{TODO}
+
 \section{Some Concentration Inequalities}
@@
 \item[Empirical Bernstein's inequality
-  \rm{\cite{MaurerPontil.benrstein}}:]\footnote{see
-    Appendix~\ref{app:deriv-conc-empbernstein} for derivation}
+  \rm{\cite{MaurerPontil.benrstein}} \textrm{(derived in Appendix~\ref{app:deriv-conc-empbernstein})}:]
 \begin{eqnarray}
@@
 \end{equation}
+\label{thm:vopi}
 \end{thm}
+
+\begin{proof} For the case $i\ne \alpha$, the probability that the perfect
+  information about the $i$th arm changes the final choice is
+  $\Pr(\IE[X_i] \ge \overline X_\alpha)$. $\IE[X_i] \le 1$ by definition,
+  therefore the maximum increase in the expected reward is
+  $(1-\overline X_\alpha)$. Thus the intrinsic value of perfect
+  information is at most $\Pr(\IE[X_i] \ge \overline
+  X_\alpha)(1-\overline X_\alpha)$.
+  Proof for the case $i=\alpha$ is similar.
+\end{proof}
 
@@
   \begin{array}{l l}
-    \Pr(\overline X_i'\le\overline X_\beta)\overline X_\beta\frac N {N+n_i} & \mbox{if $i=\alpha$} \\
-    \Pr(\overline X_i'\ge\overline X_\alpha)(1-\overline  X_\alpha)\frac N {N+n_i} &  \mbox{otherwise}
+    \Pr(\overline X_i'\le\overline X_\beta)\overline X_\beta\frac N {N+n_i}
+       \le N\Pr(\overline X_i'\le\overline X_\beta)\frac {\overline X_\beta} {n_i} & \mbox{if $i=\alpha$} \\
+    \Pr(\overline X_i'\ge\overline X_\alpha)(1-\overline  X_\alpha)\frac N {N+n_i}
+       \le N\Pr(\overline X_i'\ge\overline X_\alpha)\frac {(1-\overline  X_\alpha)} {n_i} & \mbox{otherwise}
   \end{array} \right.
@@
 samples.
+\label{thm:be}
 \end{thm}
 
- The probabilities in equations (\ref{eqn:thm-vopi}, \ref{eqn:thm-be}) can be bounded from above using concentration
-inequalities. In particular, Lemma~\ref{lemma:hoeffding-prob-bounds} is
-based on the Hoeffding inequality (\ref{eqn:conc-hoeffding}):
-\begin{lmm} The probabilities in equations (\ref{eqn:thm-vopi}, \ref{eqn:thm-be}) are bounded from above as
+\begin{proof}
+The proof is similar to the proof of Theorem~\ref{thm:vopi}, with
+$\overline X'_i$ substituted instead of $\IE[X_i]$, and, if
+$i\ne\alpha$, the upper bound on $X_i'$ is $X_\alpha+(1-X_\alpha)
+\frac N {N+n_i}$; if $i=\alpha$, the lower bound on $X_i'$ is
+$X_\beta - X_\beta \frac N {N+n_i}$.
+\end{proof}
+
+The probabilities in equations (\ref{eqn:thm-vopi}, \ref{eqn:thm-be})
+can be bounded from above using concentration inequalities. In
+particular, Lemma~\ref{lemma:hoeffding-prob-bounds} (proved in 
+Appendix~\ref{app:hoeffding-prob-bounds-proof}) is based on the
+Hoeffding inequality (\ref{eqn:conc-hoeffding}):
+\begin{lmm} The probabilities in equations (\ref{eqn:thm-vopi},
+\ref{eqn:thm-be}) are bounded from above as
 \begin{eqnarray}
-\Pr(\IE[X_i] \le \overline X_\beta|i=\alpha)& \le & \exp(-2 (\overline X_i - \overline X_\beta)^2 n_i)\nonumber\\
-\Pr(\IE[X_i] \ge \overline X_\alpha|i\ne\alpha)& \le & \exp(-2 (\overline X_\alpha - \overline X_i)^2 n_i)\nonumber\\
-\Pr(\overline X_i' \le \overline X_\beta|i=\alpha)& \le &  2\exp\left(-\frac {2(\overline X_i - \overline X_\beta)^2  (n_i+N) n_i} {(\sqrt n_i + \sqrt {n_i + N})^2}  \right) < 2\exp\left(-\frac {(\overline X_i - \overline X_\beta)^2 n_i} 2 \right)\nonumber\\
-\Pr(\overline X_i' \ge \overline X_\alpha|i\ne\alpha)& \le & 2\exp\left(-\frac {(\overline X_\alpha - \overline X_i)^2 (n_i+N) n_i} {(\sqrt n_i + \sqrt {n_i + N})^2} \right) <  2\exp\left(-\frac {(\overline X_\alpha - \overline X_i)^2 n_i} 2 \right)
+\Pr(\IE[X_i] \le \overline X_\beta|i=\alpha)&
+   \le & \exp(-2 (\overline X_i - \overline X_\beta)^2 n_i)\nonumber\\
+\Pr(\IE[X_i] \ge \overline X_\alpha|i\ne\alpha)&
+   \le & \exp(-2 (\overline X_\alpha - \overline X_i)^2 n_i)
 \label{eqn:probound-perf-hoeffding}
+\end{eqnarray}
+\vspace{-1.5\baselineskip}
+\begin{eqnarray}
+\Pr(\overline X_i' \le \overline X_\beta|i=\alpha) &
+   \le & 2\exp\left(- 2\left(\frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}}(\overline X_i - \overline X_\beta)\right)^2 n_i \right)
+   \le 2\exp\left(- 1.37(\overline X_i - \overline X_\beta)^2 n_i \right) \nonumber\\
+\Pr(\overline X_i' \ge \overline X_\alpha|i\ne\alpha) &
+   \le & 2\exp\left(- 2\left(\frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}} (\overline X_\alpha - \overline X_i)\right)^2 n_i \right)
+    \le 2\exp\left(- 1.37(\overline X_\alpha - \overline X_i)^2 n_i \right)
+\label{eqn:probound-blnk-hoeffding}
 \end{eqnarray}
@@
 
-Better bounds can be obtained through tighter estimates on
-the probabilities, for example, based on the empirical Bernstein
-inequality (\ref{eqn:conc-empbernstein}) or through a more careful
-application of the Hoeffding inequality (Appendix~\ref{app:better-hoeffding-bound}).
+An upper bound on the intrinsic value of perfect information is obtained by substituting
+(\ref{eqn:probound-perf-hoeffding}) into (\ref{eqn:thm-vopi}):
+\begin{equation}
+  \Lambda_i^p \le \hat\Lambda_i^p=\left\{
+  \begin{array}{l l}
+    \exp\left(- 2(\overline X_i - \overline X_\beta)^2 n_i \right)\overline X_\beta & \mbox{if $i=\alpha$} \\
+    \exp\left(- 2(\overline X_\alpha - \overline X_i)^2 n_i \right)(1-\overline  X_\alpha)  &  \mbox{otherwise}
+  \end{array} \right.
+\label{eqn:bound-perf-hoeffding}
+\end{equation}
+An upper bound on the blinkered VOI estimate with ordering independent of $N$ is obtained
+by substituting (\ref{eqn:probound-blnk-hoeffding}) into (\ref{eqn:thm-be}).
+\begin{equation}
+  \Lambda_i^b \le \hat\Lambda_i^b=\left\{
+  \begin{array}{l l}
+    C \exp\left(- 1.37(\overline X_i - \overline X_\beta)^2 n_i\right)\frac {\overline X_\beta} {n_i}
+      & \mbox{if $i=\alpha$} \\
+    C \exp\left(- 1.37(\overline X_\alpha - \overline X_i)^2 n_i\right)\frac {(1-\overline  X_\alpha)} {n_i}
+      &  \mbox{otherwise}
+  \end{array} \right.
+\label{eqn:bound-blnk-hoeffding}
+\end{equation}
+where $C=2N$. The bound is generally tighter when $N$ is small relative to $n_i$,
+and the estimate $\hat\Lambda_i^b$ can be viewed as a relaxation of the myopic VOI
+estimate to longer sequences of actions, such
+that the VOI estimate is positive for some of the actions.
+
+Better bounds can be obtained through tighter estimates on the
+probabilities, for example, based on the empirical Bernstein inequality
+(\ref{eqn:conc-empbernstein}) or through a more careful application of
+the Hoeffding inequality (Appendix~\ref{app:better-hoeffding-bound}).
 
@@
 
-Following the principles of rational metareasoning, an arm with the highest upper bound $\hat
-\Lambda$ on the perfect value of information should be pulled at each
-step. This way, arms known to have a low VOI  would be pulled less frequently.
+Following the principles of rational metareasoning, an arm with
+the highest VOI should be pulled at each step. The upper bounds
+established in Theorems~\ref{thm:vopi},~\ref{thm:be} can be used
+as VOI estimates. Blinkered VOI estimates~(\ref{eqn:thm-be},
+ \ref{eqn:bound-blnk-hoeffding}) can be viewed
+as approximations of VOI estimates for myopic or
+receding horison policies. As illustrated by the empirical evaluation
+(Section~\ref{sec:empirical-evaluation}), estimates based on upper
+bounds on the VOI result in rational sampling policies, and exhibit
+performance comparable to the performance of some state-of-the-art
+heuristic algorithms.
 
@@
 
-The upper bounds~(\ref{eq:lambda-hoeffding-bounds}, \ref{eq:lambda-bernstein-bounds}) decrease exponentially with the
-number of pulls $n$. When the upper bound of the VOI for all arms
-becomes lower than a threshold $\lambda$, which can be chosen based on
-resource constraints, the sampling should be stopped, and an arm should
-be chosen.
+The simplest termination condition for a sampling policy is the
+budget---a fixed number of samples performed before a decision is
+made. When a sample has a cost commensurable with the value of
+information of a measurement, an upper bound on the intrinsic VOI can
+be used to stop the sampling if the intrinsic VOI of any action
+is less than the cost of sampling $C$:
+\begin{equation}
+\mbox{stop if } \max_i \Lambda_i \le C
+\end{equation}
+Blinkered VOI estimates~(\ref{eqn:thm-be},
+\ref{eqn:bound-blnk-hoeffding}) include the remaining budget $N$ as a
+factor, but given the cost of a single sample $c$, the cost of the
+remaining samples accounted for in estimating the intrinsic VOI is
+$C=cN$. $N$ can be dropped on both sides of the inequality,
+and a viable stopping condition is
+\begin{equation}
+\begin{array}{c r r l l}
+&\frac 1 N \Lambda_\alpha^b
+\le\Pr(\overline X_\alpha'\le\overline X_\beta)\frac {\overline  X_\beta} {n_\alpha}&\le& c
+&\nonumber\\
+\mbox{and}&&&&\nonumber\\
+&\frac 1 N \max_i\Lambda_i^b
+\le \max_i\Pr(\overline X_i'\ge\overline X_\alpha)\frac {(1-\overline  X_\alpha)} {n_i}&\le& c
+ & \forall i: i\ne\alpha
+\end{array}
+\label{eqn:stopping-blnk}
+\end{equation}
+The empirical evaluation (Section~\ref{sec:empirical-evaluation})
+confirms viability of this stopping condition and illustrates the
+influence of the sample cost $c$ on the performance of
+the sampling policy.
 
@@
 \section{Empirical Evaluation}
+\label{sec:empirical-evaluation}
 
@@
 
-Figure~\ref{fig:uct-against-vct}, Figure~\ref{fig:uct-against-ect}, Figure~\ref{fig:uct-against-bct},
-Figure~\ref{fig:best-winning-rate}.
+Figure~\ref{fig:uct-against-vct}, Figure~\ref{fig:uct-against-ect},
+Figure~\ref{fig:uct-against-bct}, Figure~\ref{fig:best-winning-rate}.
 
@@
 \centering
-\includegraphics[scale=0.8]{bests-colorful.pdf}
+\includegraphics[scale=0.8]{bests-bw.pdf}
 \caption{Go: best winning rate comparison}
@@
 \end{figure}
+
+\section{Summary and Future Work}
+
+\textit{TODO}
 
@@
 
+\section{Proof of Lemma~\ref{lemma:hoeffding-prob-bounds}}
+\label{app:hoeffding-prob-bounds-proof}
+
+\begin{proof}[]
+Equations (\ref{eqn:probound-perf-hoeffding}) is a direct
+application of the Hoeffding inequality (\ref{eqn:conc-hoeffding}).
+
+Equations (\ref{eqn:probound-blnk-hoeffding}) follow from the
+observation that if $i\ne\alpha$, $\overline X_i'>\overline X_\alpha$
+if and only if the mean $\overline X_i^+$ of $N$ samples from $n_i+1$
+to $N$ is at least $\overline X_\alpha+(\overline X_\alpha-\overline
+X_i)\frac {n_i} N$.
+
+For any $\delta$, the probability that $\overline X_i'$ is \textit{not
+greater} than
+$X_\alpha$ is more than the probability that $\IE[X_i]\le\overline
+X_i+\delta$ \textit{and} $\overline X_i'\ge \IE[X_i]+\overline X_\alpha - \overline
+  X_i - \delta +(\overline X_\alpha - \overline X_i)\frac {n_i} N$;
+therefore, the probability that $\overline X_i'$ is \textit{greater}
+than than $X_\alpha$ is less than the probability that $\IE[X_i]\ge\overline
+X_i+\delta$ \textit{or} $\overline X_i'\le \IE[X_i]+\overline X_\alpha - \overline
+  X_i - \delta +(\overline X_\alpha - \overline X_i)\frac {n_i} N$.
+Using the union bound,
+\begin{equation}
+\Pr(\overline X_i'\ge \overline X_\alpha)\le
+  \Pr(\IE[X_i]-\overline X_i \ge \delta) +
+  \Pr(\overline X_i^+ - \IE[X_i] \ge \overline X_\alpha - \overline
+  X_i - \delta +(\overline X_\alpha - \overline X_i)\frac {n_i} N)
+\end{equation}
+Bounding the probabilities on the right side using the Hoeffding
+inequality, obtain:
+\begin{equation}
+\Pr(\overline X_i'\ge \overline X_\alpha)\le
+  \exp(-2\delta^2n_i)+\exp\left(-2\left((\overline X_\alpha - \overline
+  X_i)\left(1+\frac {n_i} N\right) - \delta\right)^2N\right)
+\end{equation}
+$\delta^2n = -2\left((\overline X_\alpha - \overline X_i)
+(1+\frac {n_i} N) - \delta\right)^2N$ when
+$\delta=\frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}} (\overline X_\alpha
+- \overline X_i) \ge (\sqrt 2 - 1)(\overline X_\alpha-\overline X_i)$,
+therefore
+\begin{eqnarray}
+\Pr(\overline X_i'\ge \overline X_\alpha) 
+& \le & 2\exp\left(-2\left( \frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}}(\overline X_\alpha - \overline X_i)\right)^2 n_i\right)\nonumber \\
+& \le & 2\exp(-2(\sqrt 2 - 1)^2(\overline X_\alpha - \overline X_i)^2n_i)
+ \le 2\exp(-1.37(\overline X_\alpha - \overline X_i)^2n_i)
+\end{eqnarray}
+Derivation for the case $i=\alpha$ is similar.
+\end{proof}
+
 \section{Empirical Bernstein Inequality}
@@
 Theorem~4 in~\cite{MaurerPontil.benrstein} states that
-\[\Pr\left(\IE[X]-\overline X_n \ge \sqrt { \frac {2\overline\sigma_n^2 \ln 2/\delta} n } + \frac {7 \ln 2/\delta} {3(n-1)}\right)\le \delta,\]
+\begin{equation}
+\Pr\left(\IE[X]-\overline X_n
+    \ge \sqrt { \frac {2\overline\sigma_n^2 \ln 2/\delta} n } + \frac {7 \ln 2/\delta} {3(n-1)}\right)\le \delta
+\end{equation}
 Therefore
-\[\Pr\left(\IE[X]-\overline X_n \ge \sqrt { \left(\frac {7 \ln 2/\delta} {3(n-1)}\right)^2+\frac {2\overline\sigma_n^2 \ln 2/\delta} n } + \frac {7 \ln 2/\delta} {3(n-1)}\right)\le
-\delta.\]
-$a=\sqrt { \left(\frac {7 \ln 2/\delta} {3(n-1)}\right)^2+\frac {2\overline\sigma_n^2 \ln 2/\delta} n } + \frac {7 \ln 2/\delta} {3(n-1)}$ is a root of
-square equation
-\[a^2-a\frac {14 \ln 2/\delta} {3(n-1)} -\frac {2\overline\sigma_n^2 \ln 2/\delta} n=0\]
+\begin{equation}
+\Pr\left(\IE[X]-\overline X_n \ge
+              \sqrt { \left(\frac {7 \ln 2/\delta} {3(n-1)}\right)^2+\frac {2\overline\sigma_n^2 \ln 2/\delta} n }
+                      + \frac {7 \ln 2/\delta} {3(n-1)}\right)
+   \le \delta.
+\end{equation}
+$a=\sqrt { \left(\frac {7 \ln 2/\delta} {3(n-1)}\right)^2+\frac {2\overline\sigma_n^2 \ln 2/\delta} n } + \frac {7 \ln 2/\delta} {3(n-1)}$
+ is a root of square equation $a^2-a\frac {14 \ln 2/\delta} {3(n-1)} -\frac {2\overline\sigma_n^2 \ln 2/\delta} n=0$
 which, solved for $\delta\triangleq\Pr(\IE[X]-\overline X_n\ge a)$,
 gives
-\[\Pr(\IE[X]-\overline X_n\ge a)\le 2\exp \left( - \frac {na^2} {\frac {14} {3} \frac {n} {n-1}a+2\overline\sigma_n^2}\right)\]
+\begin{equation*}
+\Pr(\IE[X]-\overline X_n\ge a)\le 2\exp \left( - \frac {na^2} {\frac {14} {3} \frac {n} {n-1}a+2\overline\sigma_n^2}\right)
+\end{equation*}
 Other derivations, giving slightly different results, are possible.
@@
 
-The bound can be supposedly be improved by selecting a midpoint
-$0 < y < \overline X_\beta$ and computing the bound as the sum of two parts:
+Bounds (\ref{eqn:bound-perf-hoeffding}, \ref{eqn:bound-blnk-hoeffding}) have the form
+\begin{equation}
+  \Lambda_i \le \hat\Lambda_i=\left\{
+  \begin{array}{l l}
+    \exp\left(- \gamma(\overline X_i - \overline X_\beta)^2 n_i \right)(\overline X_\beta-B) & \mbox{if $i=\alpha$} \\
+    \exp\left(- \gamma(\overline X_\alpha - \overline X_i)^2 n_i \right)(A-\overline  X_\alpha)  &  \mbox{otherwise}
+  \end{array} \right.
+\label{eqn:bound-generic-hoeffding}
+\end{equation}
+where $A$ is the upper bound on the posterior sample mean $\overline X_i'$ if $i\ne \alpha$, $B$ is the lower bound on $\overline X_i'$ if $i=\alpha$.
+For $i=\alpha$, the bound can be improved by selecting a midpoint
+$B < y < \overline X_\beta$ and computing the bound as the sum of two parts:
 \begin{itemize}
 \item $\overline X_\beta-y$ multiplied by the probability that
-  $\mu_\alpha \le \overline X_\beta$;
-\item $\overline X_\beta$ multiplied by the probability that $\mu_\alpha\le
+  $\overline X_i' \le \overline X_\beta$;
+\item $\overline X_\beta-B$ multiplied by the probability that $\overline X_i'\le
   y$.
 \end{itemize}.
-\[ V_{\overline X=\overline X_\alpha} \le (\overline X_\beta-y)\exp\left(-2n(\overline X_\alpha-\overline X_\beta)^2\right)+\overline X_\beta\exp\left(-2n(\overline X_\alpha-y)^2\right) \]
-
-\vspace{\baselineskip}
-
-The minimum of $V^*$ is achieved when $\frac {dV^*} {dy}=0$, that
+\begin{equation}
+\Lambda_{i|i=\alpha} \le \hat \Lambda_{i|i=\alpha} =
+ (\overline X_\beta-y)\exp\left(-\gamma n_i(\overline X_i-\overline X_\beta)^2\right)
+ +(\overline X_\beta-B)\exp\left(-\gamma n_i(\overline X_i-y)^2\right)
+\label{eqn:bound-twopart-hoeffding-alpha}
+\end{equation}
+If $y$ that minimizes $\hat\Lambda_{i|i=\alpha}$ exists, the minimum is achieved when
+ $\frac {d\hat\Lambda_{i|i=\alpha}} {dy}=0$, that
 is, when $y$ is the root of the following equation:
-\[ 4\overline X_\beta n(\overline X_\alpha-y)=\exp\left(-2n\left(\frac {\overline X_\alpha-\overline X_\beta}
-    {\overline X_\alpha-y}\right)^2\right) \]
-If a root in the interval  $0\le y\le\beta$ exists, then the
-number of samples is bounded as
-\[n\le\frac 1 {4\overline X_\beta(\overline X_\alpha-\overline X_\beta)}\]
-by observing that the right-hand side 
-is at most $1$ (a negative power), and the left-hand side is at least
-$4\overline X_\beta n(\overline X_\alpha-\overline X_\beta)$.
-So, the bound can supposedly be improved for smaller values of
-$n$. The improvement is more significant when the current best and
-second-best sample means are close.
-
-The derivation for the other case
-(sampling an item that can be better than the current best) is
-obtained by substitution $1-\overline X, 1-\overline X_\alpha, 1-y$ instead of
-$\overline X_\alpha, \overline X_\beta, y$:
-
-\[ V_{\overline X\ne\overline X_\alpha} \le (y-\overline X_\alpha)\exp\left(-2n(\overline X_\alpha-\overline
-  X)^2\right)+(1-\overline X_\alpha)\exp\left(-2n(y-\overline X)^2\right) \]
-
-The anticipated influence of the improved
-VOI estimate would be that the selected item will be a less discovered
-one and further from the current best or second-best.
-
+\begin{equation}
+2\gamma n_i(\overline X_\beta-B)(\overline X_i-y)=\exp\left(-\gamma n_i\left(\frac {\overline X_i-\overline X_\beta}
+    {\overline X_i-y}\right)^2\right)
+\end{equation}
+The derivation for the case $i\ne \alpha$ is obtained by substitution $1-A, 1-\overline X_i, 1-\overline X_\alpha, 1-y$ instead of
+$B, \overline X_i, \overline X_\beta, y$ into (\ref{eqn:bound-twopart-hoeffding-alpha}):
+\begin{equation}
+\Lambda_{i|i\ne\alpha} \le \hat\Lambda_{i|i\ne\alpha} = (y-\overline X_\alpha)\exp\left(-\gamma n_i(\overline X_\alpha-\overline
+  X_i)^2\right)+(A-\overline X_\alpha)\exp\left(-\gamma n_i(y-\overline X_i)^2\right)
+\label{eqn:bound-twopart-hoeffding-rest}
+\end{equation}
 A closed-form solution for $y$ cannot be obtained, but given
-$\overline X_\alpha, \overline X_\beta, n$, the value of $y$ can be efficiently
-computed. It should be determined empirically whether the improved
-estimate has justified influence on the performance of the algorithm.
+$A, B, \overline X_\alpha, \overline X_\beta, n$, the value of $y$ can be efficiently
+computed.
 

### 2011-12-07 paper/voisam/onbetterbound.tex
initial commit

@@
 $\beta$:
-\[ V \le \beta \exp\left(-2n(\alpha-\beta)^2\right) \]
+\[ V_{\overline x=\alpha} \le \beta \exp\left(-2n(\alpha-\beta)^2\right) \]
 The bound can be supposedly be improved by selecting a midpoint
@@
 \end{itemize}.
-\[ V^* \le (\beta-\gamma)\exp\left(-2n(\alpha-\beta)^2\right)+\beta\exp\left(-2n(\alpha-\gamma)^2\right) \]
+\[ V_{\overline x=\alpha} \le (\beta-\gamma)\exp\left(-2n(\alpha-\beta)^2\right)+\beta\exp\left(-2n(\alpha-\gamma)^2\right) \]
 
@@
 (sampling an item that can be better than the current best) is
-obtained by substitution $1-\alpha, 1-\beta, 1-\gamma$ instead of
-$\alpha, \beta, \gamma$. The anticipated influence of the improved
+obtained by substitution $1-\overline x, 1-\alpha, 1-\gamma$ instead of
+$\alpha, \beta, \gamma$:
+
+\[ V_{\overline x\ne\alpha} \le (\gamma-\alpha)\exp\left(-2n(\alpha-\overline
+  x)^2\right)+(1-\alpha)\exp\left(-2n(\gamma-\overline x)^2\right) \]
+
+The anticipated influence of the improved
 VOI estimate would be that the selected item will be a less discovered

### 2011-12-07 paper/voisam/paper.tex
initial commit

@@
 \delta.\]
-$t=\sqrt { \left(\frac {7 \ln 2/\delta} {3(n-1)}\right)^2+\frac {2\overline\sigma_n^2 \ln 2/\delta} n } + \frac {7 \ln 2/\delta} {3(n-1)}$ is a root of
+$a=\sqrt { \left(\frac {7 \ln 2/\delta} {3(n-1)}\right)^2+\frac {2\overline\sigma_n^2 \ln 2/\delta} n } + \frac {7 \ln 2/\delta} {3(n-1)}$ is a root of
 square equation
-\[t^2-t\frac {14 \ln 2/\delta} {3(n-1)} -\frac {2\overline\sigma_n^2 \ln 2/\delta} n=0\]
-which, solved for $\delta\triangleq\Pr(\IE[X]-\overline X_n\ge t)$,
+\[a^2-a\frac {14 \ln 2/\delta} {3(n-1)} -\frac {2\overline\sigma_n^2 \ln 2/\delta} n=0\]
+which, solved for $\delta\triangleq\Pr(\IE[X]-\overline X_n\ge a)$,
 gives
-\[\Pr(\IE[X]-\overline X_n\ge t)\le 2\exp \left( - \frac {na^2} {\frac {14} {3} \frac {n} {n-1}a+2\overline\sigma_n^2}\right)\]
+\[\Pr(\IE[X]-\overline X_n\ge a)\le 2\exp \left( - \frac {na^2} {\frac {14} {3} \frac {n} {n-1}a+2\overline\sigma_n^2}\right)\]
 Other derivations, giving slightly different results, are possible.

### 2011-12-07 code/figures.R

@@ -1,0 +1,112 @@
+ALGORITHMS <- c('rnd', 'ucb', 'sve')
+
+
+draw.regrets <- function(regrets_file, log="xy", legend.position='topright', xlab=expression(N[samples]), main=NA) {
+	regrets <- read.table(regrets_file, header=T)
+	xlim <- range(regrets$nsamples)
+	values <- unlist(regrets[-1])
+	ylim <- range(values[values>0])
+	plot(x=regrets$nsamples, type='n', log=log, xlim=xlim, ylim=ylim, yaxs='i', xlab=xlab, ylab='Regret', main=if(is.na(main)) unlist(strsplit(regrets_file, "\\."))[[1]] else main)
+	i <- 1
+	for(alg in algs) {
+		y <- regrets[[alg]]
+		lines(x=regrets$nsamples, y=regrets[[alg]],
+			  type="o", pch=i)
+		i <- i+1
+	}
+	legend(x=legend.position, legend=sapply(names(regrets)[-1], function(n) substring(n, 3)), pch=1:length(names(regrets)[-1]))
+}
+
+draw.experiment <- function(prefix="") {
+  	old.par <- par(no.readonly=TRUE)
+	par(mar=c(4,4,2,2))
+	layout(cbind(c(1,2,3),c(4,5,6)))
+	for(result in paste(c('symmetric', 'twothirds', 'threequarters', 'quarter', 'third', 'many'), "txt", sep=".")) draw.regrets(paste(prefix,result,sep=""))
+        par(old.par)
+}
+
+compare.random <- function(upper_file='random-upper.txt', greedy_file='random-greedy.txt', semigreedy_file='random-semigreedy.txt') {
+  upper <- read.table(upper_file, header=T)
+  greedy <- read.table(greedy_file, header=T)
+  semigreedy <- read.table(semigreedy_file, header=T)
+
+  xlim <- range(upper$nsamples)
+  ylim <- range(0, upper$r_sve, greedy$r_sve, semigreedy$r_sve)
+  plot(x=upper$nsamples, type='n', xlim=xlim, ylim=ylim, xlab=expression(N[samples]), ylab='Regret', main='upper vs. greedy vs. semigreedy')
+  lines(x=upper$nsamples, upper$r_sve, type="o", pch=1)
+  lines(x=greedy$nsamples, greedy$r_sve, type="o", pch=2)
[…]

### 2011-12-07 code/sailing/figures.R

@@ -1,0 +1,72 @@
+draw.costs.of.factor <- function(costs_file, legend.position='bottomright', log="x",main=NA) {
+  costs <- read.table(costs_file, header=T)
+  algorithms <- names(costs)[-1]
+  xlim <- range(costs$FACTOR)
+  ylim <- range(costs[algorithms])
+  plot(x=costs$nsamples, type='n', log=log, xlim=xlim, ylim=ylim, xaxs='i',
+       xlab='exploration factor', ylab='cost',
+       main=if(is.na(main)) do.call(paste,as.list(unlist(strsplit(costs_file,
+                            "-|\\.txt",fixed=FALSE)))[-1])
+            else main)
+  i <- 1
+  for(alg in algorithms) {
+    lines(x=costs$FACTOR, y=costs[[alg]],
+          type="o", pch=i)
+    i <- i+1
+  }
+  legend(x=legend.position, legend=algorithms, pch=1:length(algorithms), bg='white')
+}
+
+draw.costs.of.nsamples <- function(costs_file, legend.position='right', log="x", main=NA) {
+  costs <- read.table(costs_file, header=T)
+  algorithms <- names(costs)[-1]
+  xlim <- range(costs$SAMPLES)
+  ylim <- range(costs[algorithms])
+  print(costs)
+  plot(x=costs$nsamples, type='n', log=log, xlim=xlim, ylim=ylim, xaxs='i',
+       xlab='samples per node', ylab='cost',
+       main=if(is.na(main)) do.call(paste,as.list(unlist(strsplit(costs_file,
+                            "-|\\.txt",fixed=FALSE)))[-1])
+            else main)
+  i <- 1
+  for(alg in algorithms) {
+    lines(x=costs$SAMPLES, y=costs[[alg]],
+          type="o", pch=i)
+    i <- i+1
+  }
+  legend(x=legend.position, legend=algorithms, pch=1:length(algorithms), bg='white')
+}
+  
+draw.rcq <- function() {
+	layout(cbind(c(1,2), c(3,4))); 
+	for(x in c(199, 397, 793, 1585)) 
+		draw.costs.of.factor(paste('../../exp/sailing/rcq-size=6-nsamples=', x, '.txt', sep=''),
+							 legend.position='right')
+}
+							 
+
+make.pdfs <- function () {
+  for(size in c('3', '6', '10')) {
+    for(group in c('minimum', 'median')) {
+      basename <- paste('costs-size=', size, '-group=', group, sep='')
[…]

### 2011-12-07 code/sailing/group.py

@@ -1,0 +1,51 @@
+import re 
+import sys
+
+def minimum(vals):
+  return min(vals)
+
+def median(vals):
+  vals.sort()
+  if len(vals) % 2 == 1:
+    return vals[(len(vals)+1)/2-1]
+  else:
+    lower = vals[len(vals)/2-1]
+    upper = vals[len(vals)/2]
+    return (float(lower + upper))/2.0
+
+def mean(vals):
+    return float(sum(vals))/len(vals)
+
+def do(filenames,group):
+    regrets = {}
+    names = None
+    size = None
+    for filename in filenames:
+        m = re.match('costs-size=(\d+)-nsamples=(\d+).txt', filename)
+        size = m.group(1)
+        nsamples = int(m.group(2))
+        f = file(filename)
+        names = f.readline().split()
+        values = [[] for name in names[1:]]
+        for line in f:
+            lvs = [float(v) for v in line.split()[1:]]
+            for i in range(len(values)):
+                values[i].append(lvs[i])
+            regrets[nsamples] = [globals()[group](vs) for vs in values]
+        f.close()
+    names = ['SAMPLES']+names[1:]
+    f = file("costs-size=%s-group=%s.txt" % (size,group), "w")
+    print >> f, ' '.join(names)
+    for nsamples in sorted(regrets.keys()):
+        print >> f, nsamples, ' '.join([str(x) for x in regrets[nsamples]])
+    f.close()
+
+if __name__=='__main__':
+    import sys
+    for group in ['minimum', 'median']:
+      do(sys.argv[1:], group)
+    
+            
+            
+            
+    

### 2011-12-07 code/sailing/splitlog.py

@@ -1,0 +1,35 @@
+import sys, re
+
+def do(logf,template='costs-size=%s-nsamples=%s.txt'):
+    resf = None
+    size='X'
+    for line in logf:
+        m = re.match('.*\(exp1 .*:size (\d+)\s*', line)
+        if m:
+            size = int(m.group(1))
+            continue
+        m = re.match('\[(\d+) samples\].*', line)
+        if m:
+            nsamples = int(m.group(1))
+            if resf:
+                resf.close()
+            resf = file(template % (size, nsamples), "w")
+            continue
+        if resf:
+            m = re.match('\s*[A-Z]+(\s+[A-Z]+)* *', line) \
+                or re.match('\s*[-\d.]+(\s+[-\d.]+)* *', line)
+            if m:
+                print >>resf, line,
+    if resf:
+        resf.close()
+
+if __name__=="__main__":
+    if len(sys.argv)>1:
+        logf = file(sys.argv[1])
+        if len(sys.argv)==3:
+            do(logf, sys.argv[2])
+        else:
+            do(logf)
+    else:
+        do(sys.stdin)
+        

### 2011-12-07 paper/vago/paper.tex
redistribution

@@
 \subsection{Termination Condition}
+\label{sec:control-termination-condition}
 
@@
 
-UCT forwards samples to the next search stage when the winner at the
-current state is known with high confidence. VOI-based termination
-condition can be used to stop the sampling in the current state early
-and save the remaining samples for re-use in a later search state.
+Monte-Carlo tree search \cite{Chaslot.montecarlo} selects at each search step an action
+that appears to be the best based on outcomes of \textit{search rollouts}. Two different
+criteria are employed in action selection during a rollout:
+\begin{itemize}
+\item at the first step, the simple regret of selecting an action must be minimized;
+\item starting with the second step of a rollout, the expected reward of the rollout
+ must be made as close as possible to the optimum reward, so that the value of the
+ action at the first step of the rollout is evaluated correctly. Thus, starting with
+ the second step, the cumulative regret must be minimized.
+\end{itemize}
+
+A natural approach would be to apply an algorithm that minimizes the
+simple regret in Multi-armed bandits at the first step, and the
+cumulative regret (solves the exploration-exploitation trade-off)
+during the rest of the rollout.  However, this approach assumes that
+the information obtained from rollouts in the current state is
+discarded after an action is selected. In practice, most successful
+Monte-Carlo tree search algorithms re-use rollouts originating from an
+earlier search state and passing through the current search state;
+thus, the value of information of a rollout is determined not just by
+the influence on the choice of the action at the current state, but
+also by the influence on the choice at future search states, provided
+the rollout passes through the search states which will be visited.
+
+One way to account for this re-use would be to incroprorate the
+`future' value of information into a VOI estimate. However, this 
+approach appears to be complicated. Alternatively, one can behave
+myopically in search tree depth:
+\begin{itemize}
+\item estimate VOI as though the information is discarded after each step;
+\item stop early if the VOI is below a certain threshold
+   (see Section~\ref{sec:control-termination-condition});
+\item save the unused sample budget for search in future state, such that
+   if the nominal budget is $N$, and the unused budget in the last state
+   is $N_u$, the search budget in the next state will be $N+N_u$.
+\end{itemize}
+In this approach, the cost of a sample in the current state is the VOI
+of increasing the budget of a future state by one sample. It is unclear
+whether this cost can be accurately estimated, but supposedly a fixed value
+for a given problem type and algorithm implementation would work. Indeed,
+the empirical evaluation (Section~\ref{sec:emp-go}) confirms that a learned
+fixed cost and sample redistribution substantially improve performance of
+the VOI-based sampling policy in game tree search.
 
@@
 \subsection{Playing Go Against UCT}
+\label{sec:emp-go}
 

### 2011-12-07 paper/vago/paper.tex
emp intro

@@
   \begin{array}{l l}
-    \Pr(\IE[X_i] \le \overline X_\beta)\overline X_\beta & \mbox{if $i=\alpha$} \\
-    \Pr(\IE[X_i] \ge \overline X_\alpha)(1-\overline X_\alpha) & \mbox{otherwise}
+    \overline X_\beta \Pr(\IE[X_i] \le \overline X_\beta) & \mbox{if $i=\alpha$} \\
+    (1-\overline X_\alpha)\Pr(\IE[X_i] \ge \overline X_\alpha) & \mbox{otherwise}
   \end{array} \right.
@@
   \begin{array}{l l}
-    \Pr(\overline X_i'\le\overline X_\beta)\overline X_\beta\frac N {N+n_i}
-       \le N\Pr(\overline X_i'\le\overline X_\beta)\frac {\overline X_\beta} {n_i} & \mbox{if $i=\alpha$} \\
-    \Pr(\overline X_i'\ge\overline X_\alpha)(1-\overline  X_\alpha)\frac N {N+n_i}
-       \le N\Pr(\overline X_i'\ge\overline X_\alpha)\frac {(1-\overline  X_\alpha)} {n_i} & \mbox{otherwise}
+    \overline X_\beta\frac N {N+n_i}\Pr(\overline X_i'\le\overline X_\beta)
+    \le N\frac {\overline X_\beta} {n_i}\Pr(\overline X_i'\le\overline X_\beta) & \mbox{if $i=\alpha$} \\
+    (1-\overline  X_\alpha)\frac N {N+n_i}\Pr(\overline X_i'\ge\overline X_\alpha)
+       \le N\frac {(1-\overline  X_\alpha)} {n_i}\Pr(\overline X_i'\ge\overline X_\alpha) & \mbox{otherwise}
   \end{array} \right.
@@
   \begin{array}{l l}
-    \exp\left(- 2(\overline X_i - \overline X_\beta)^2 n_i \right)\overline X_\beta & \mbox{if $i=\alpha$} \\
-    \exp\left(- 2(\overline X_\alpha - \overline X_i)^2 n_i \right)(1-\overline  X_\alpha)  &  \mbox{otherwise}
+    \overline X_\beta\exp\left(- 2(\overline X_i - \overline X_\beta)^2 n_i \right) & \mbox{if $i=\alpha$} \\
+    (1-\overline  X_\alpha)\exp\left(- 2(\overline X_\alpha - \overline X_i)^2 n_i \right)(1-\overline  X_\alpha)  &  \mbox{otherwise}
   \end{array} \right.
@@
   \begin{array}{l l}
-    C \exp\left(- 1.37(\overline X_i - \overline X_\beta)^2 n_i\right)\frac {\overline X_\beta} {n_i}
+    C \frac {\overline X_\beta} {n_i}\exp\left(- 1.37(\overline X_i - \overline X_\beta)^2 n_i\right)
       & \mbox{if $i=\alpha$} \\
-    C \exp\left(- 1.37(\overline X_\alpha - \overline X_i)^2 n_i\right)\frac {(1-\overline  X_\alpha)} {n_i}
+    C \frac {(1-\overline  X_\alpha)} {n_i} \exp\left(- 1.37(\overline X_\alpha - \overline X_i)^2 n_i\right)
       &  \mbox{otherwise}
@@
 &\frac 1 N \Lambda_\alpha^b
-\le\Pr(\overline X_\alpha'\le\overline X_\beta)\frac {\overline  X_\beta} {n_\alpha}&\le& c
+\le\frac {\overline  X_\beta} {n_\alpha}\Pr(\overline X_\alpha'\le\overline X_\beta)&\le& c
 &\nonumber\\
@@
 &\frac 1 N \max_i\Lambda_i^b
-\le \max_i\Pr(\overline X_i'\ge\overline X_\alpha)\frac {(1-\overline  X_\alpha)} {n_i}&\le& c
+\le \max_i\frac {(1-\overline  X_\alpha)} {n_i}\Pr(\overline X_i'\ge\overline X_\alpha)&\le& c
  & \forall i: i\ne\alpha
@@
 
-Monte-Carlo tree search \cite{Chaslot.montecarlo} selects at each search step an action
-that appears to be the best based on outcomes of \textit{search rollouts}. Two different
-criteria are employed in action selection during a rollout:
+Monte-Carlo tree search \cite{Chaslot.montecarlo} selects at each
+search step an action that appears to be the best based on outcomes
+of \textit{search rollouts}. Two different criteria are employed in
+action selection during a rollout:
 \begin{itemize}
-\item at the first step, the simple regret of selecting an action must be minimized;
-\item starting with the second step of a rollout, the expected reward of the rollout
- must be made as close as possible to the optimum reward, so that the value of the
- action at the first step of the rollout is evaluated correctly. Thus, starting with
- the second step, the cumulative regret must be minimized.
+\item at the first step, the simple regret of selecting an action
+must be minimized;
+\item starting with the second step of a rollout, the expected reward
+of the rollout
+ must be made as close as possible to the optimum reward, so that the
+ value of the action at the first step of the rollout is evaluated
+ correctly. Thus, starting with the second step, the cumulative regret
+ must be minimized.
 \end{itemize}
@@
 myopically in search tree depth:
-\begin{itemize}
+\begin{enumerate}
 \item estimate VOI as though the information is discarded after each step;
@@
    is $N_u$, the search budget in the next state will be $N+N_u$.
-\end{itemize}
-In this approach, the cost of a sample in the current state is the VOI
-of increasing the budget of a future state by one sample. It is unclear
-whether this cost can be accurately estimated, but supposedly a fixed value
-for a given problem type and algorithm implementation would work. Indeed,
-the empirical evaluation (Section~\ref{sec:emp-go}) confirms that a learned
-fixed cost and sample redistribution substantially improve performance of
-the VOI-based sampling policy in game tree search.
+\end{enumerate}
+In this approach, the cost of a sample in the current state is the
+VOI of increasing the budget of a future state by one sample.  It is
+unclear whether this cost can be accurately estimated, but supposedly a
+fixed value for a given problem type and algorithm implementation would
+work. Indeed, the empirical evaluation (Section~\ref{sec:emp-go})
+confirms that a learned fixed cost and sample redistribution
+substantially improve the performance of the VOI-based sampling policy
+in game tree search.
 
@@
 
+The experiments compare the UCT algorithm \cite{Kocsis.uct} with modified
+versions of UCT in which the samples at the first step are selected according
+to their VOI estimates. Three blinkered VOI estimates were used for the comparison:
+\begin{description}
+\item[VCT:] based on Hoeffding inequality (\ref{eqn:conc-hoeffding});
+\item[ECT:] based on Hoeffding inequality with midpoint (see Appendix~\ref{app:better-hoeffding-bound});
+\item[BCT:] based on empirical Bernstein inequality (\ref{eqn:conc-empbernstein}).
+\end{description}
+For simplicity, the bounds on probability for unlimited number of samples were used in the estimate formulas. The estimates where computed as follows:
+\begin{eqnarray}\
+  \Lambda_i^{VCT}& =&\left\{
+  \begin{array}{l l}
+    \frac {\overline X_\beta} {n_i}\exp\left(- 2(\overline X_i - \overline X_\beta)^2 n_i\right)
+      & \mbox{if $i=\alpha$} \\
+    \frac {(1-\overline  X_\alpha)} {n_i}\exp\left(- 2(\overline X_\alpha - \overline X_i)^2 n_i\right)
+      &  \mbox{otherwise}
+  \end{array}  \right.\nonumber\\
+  \Lambda_i^{ECT}& =&\min\left(\Lambda_i^{VCT}, \min_y\left\{
+  \begin{array}{l l}
+    \frac {\overline X_\beta-y} {n_i}\exp\left(-2 n_i(\overline X_i-\overline X_\beta)^2\right)
+    +\frac {\overline X_\beta} {n_i}\exp\left(-2 n_i(\overline X_i-y)^2\right)
+      & \mbox{if $i=\alpha$} \\
+\frac {y-\overline X_\alpha} {n_i}\exp\left(-2 n_i(\overline X_\alpha-\overline
+  X_i)^2\right)+\frac {1-\overline X_\alpha} {n_i}\exp\left(-2 n_i(y-\overline X_i)^2\right)
+      &  \mbox{otherwise}
+  \end{array} \right.\right) \nonumber\\
+  \Lambda_i^{BCT}& =&\min\left(\Lambda_i^{VCT},\left\{
+  \begin{array}{l l}
+    2\frac {\overline X_\beta} {n_i}\exp\left(- \frac {(\overline X_i - \overline X_\beta)^2 n_i} {\frac {14} 3 \frac {n_i} {n_i-1}(\overline X_i - \overline X_\beta) + 2 \overline X_i(1-\overline X_i)} \right)
+      & \mbox{if $i=\alpha$} \\
+    2\frac {(1-\overline  X_\alpha)} {n_i}\exp\left(- \frac {(\overline X_\alpha - \overline X_i)^2 n_i} {\frac {14} 3 \frac {n_i} {n_i-1}(\overline X_\alpha - \overline X_i) + 2 \overline X_i(1-\overline X_i)} \right)
+      &  \mbox{otherwise}
+  \end{array} \right.\right)
+\label{eqn:emp-estimates}
+\end{eqnarray}
+
 \subsection{Selecting The Best Arm}
+\label{sec:emp-arm}
 
@@
 therefore, the probability that $\overline X_i'$ is \textit{greater}
-than than $X_\alpha$ is less than the probability that $\IE[X_i]\ge\overline
+than $X_\alpha$ is less than the probability that $\IE[X_i]\ge\overline
 X_i+\delta$ \textit{or} $\overline X_i'\le \IE[X_i]+\overline X_\alpha - \overline

### 2011-12-08 paper/vago/paper.tex
experiments on bandits

@@
 VOI of increasing the budget of a future state by one sample.  It is
-unclear whether this cost can be accurately estimated, but supposedly a
-fixed value for a given problem type and algorithm implementation would
-work. Indeed, the empirical evaluation (Section~\ref{sec:emp-go})
-confirms that a learned fixed cost and sample redistribution
-substantially improve the performance of the VOI-based sampling policy
-in game tree search.
+unclear whether this cost can be accurately estimated, but supposedly
+a fixed value for a given problem type and algorithm implementation
+would work. Indeed, the empirical evaluation (Section~\ref{sec:emp-go})
+confirms that stopping and sample redistribution based on a learned
+fixed cost  substantially improve the performance of the VOI-based
+sampling policy in game tree search.
 
@@
 
-Figure~\ref{fig:random-instances}
-
 \begin{figure}[h]
@@
 \end{figure}
+
+The sampling policies are first compared on random multi-armed bandit 
+problem instances, where the Monte Carlo Tree Search algorithm
+is reduced to the sample selection policy at the first step: UCT is 
+actually UCB1. In the problem of pure exploration in multi-armed bandits
+\cite{Bubeck.pure}, the smaller is the exploration factor, the higher
+is the expected simple regret; indeed, UCT exhibits the lowest simple
+regret in the experiments for the default exploration factor $Cp=\frac
+{\sqrt 2} 2$ \cite{Kocsis.uct}.
+
+Figure~\ref{fig:random-instances} shows experiment results for
+randomly-generated multi-armed bandits with 32 Bernoulli arms, with
+the mean rewards of the arms distributed uniformly in the range $[0,
+  1]$, for a range of sample budgets $32..2048$, with multiplicative
+step of $2$. The experiment for each number of samples was repeated
+20000 times. UCT is worse than any of the VOI-aware sampling policies,
+for example, for 384 samples the simple regret of UCT is $\approx 40$
+times greater than of VCT, and $\approx 100$ times greater than of
+ECT. On average, it takes $2.5$ times as many samples for UCT to reach
+the regret of VCT.
 

### 2011-12-08 paper/vago/paper.tex
Go: introduction

@@
 ECT. On average, it takes $2.5$ times as many samples for UCT to reach
-the regret of VCT.
+the regret of VCT. Different numbers of arms and distributions of
+means give similar results.
 
@@
 \label{sec:emp-go}
+
+One search domain in which Monte-Carlo tree search based on the UCT
+sampling policy has been particularly successful is playing the Go
+game \cite{Gelly.mogo}. This series of experiments compares
+the winning rate of VOI-aware policies (VCT, ECT, BCT) against UCT 
+in Go and investigates the effect of early stopping and sample
+redistribution.
+
+For the experiments, a modified version of Pachi \cite{Braudis.pachi},
+a state of the art Go program, was used:
+\begin{itemize}
+\item The UCT engine of Pachi was extended with VOI-aware sampling
+  policies at the first step. 
+\item The stopping condition for the VOI-aware policies was
+  modified and based solely on the sample cost, specified as
+  a constant parameter. 
+\item The time-allocation mode based on the fixed number of samples
+  was modified such that 
+  \begin{enumerate}
+    \item the same number of samples is available to
+      the agent at each step, independently of the number of pre-simulated
+      games;  
+    \item if samples were unused at the current step,
+      they become available at the next step.
+  \end{enumerate}
+\end{itemize}
 

### 2011-12-08 paper/vago/paper.tex
Go: introduction
if, otherwise in rm
V,E,BCT vs UCT

@@
   \begin{array}{l l}
-    \overline X_\beta \Pr(\IE[X_i] \le \overline X_\beta) & \mbox{if $i=\alpha$} \\
-    (1-\overline X_\alpha)\Pr(\IE[X_i] \ge \overline X_\alpha) & \mbox{otherwise}
+    \overline X_\beta \Pr(\IE[X_i] \le \overline X_\beta) & \mbox{{\rm if} $i=\alpha$} \\
+    (1-\overline X_\alpha)\Pr(\IE[X_i] \ge \overline X_\alpha) & \mbox{\rm otherwise}
   \end{array} \right.
@@
     \overline X_\beta\frac N {N+n_i}\Pr(\overline X_i'\le\overline X_\beta)
-    \le N\frac {\overline X_\beta} {n_i}\Pr(\overline X_i'\le\overline X_\beta) & \mbox{if $i=\alpha$} \\
+    \le N\frac {\overline X_\beta} {n_i}\Pr(\overline X_i'\le\overline X_\beta) & \mbox{{\rm if} $i=\alpha$} \\
     (1-\overline  X_\alpha)\frac N {N+n_i}\Pr(\overline X_i'\ge\overline X_\alpha)
-       \le N\frac {(1-\overline  X_\alpha)} {n_i}\Pr(\overline X_i'\ge\overline X_\alpha) & \mbox{otherwise}
+       \le N\frac {(1-\overline  X_\alpha)} {n_i}\Pr(\overline X_i'\ge\overline X_\alpha) & \mbox{\rm otherwise}
   \end{array} \right.
@@
   \begin{array}{l l}
-    \overline X_\beta\exp\left(- 2(\overline X_i - \overline X_\beta)^2 n_i \right) & \mbox{if $i=\alpha$} \\
-    (1-\overline  X_\alpha)\exp\left(- 2(\overline X_\alpha - \overline X_i)^2 n_i \right)(1-\overline  X_\alpha)  &  \mbox{otherwise}
+    \overline X_\beta\exp\left(- 2(\overline X_i - \overline X_\beta)^2 n_i \right) & \mbox{{\rm if} $i=\alpha$} \\
+    (1-\overline  X_\alpha)\exp\left(- 2(\overline X_\alpha - \overline X_i)^2 n_i \right)(1-\overline  X_\alpha)  &  \mbox{\rm otherwise}
   \end{array} \right.
@@
     C \frac {\overline X_\beta} {n_i}\exp\left(- 1.37(\overline X_i - \overline X_\beta)^2 n_i\right)
-      & \mbox{if $i=\alpha$} \\
+      & \mbox{{\rm if} $i=\alpha$} \\
     C \frac {(1-\overline  X_\alpha)} {n_i} \exp\left(- 1.37(\overline X_\alpha - \overline X_i)^2 n_i\right)
-      &  \mbox{otherwise}
+      &  \mbox{\rm otherwise}
   \end{array} \right.
@@
 \subsection{Sample Redistribution in Trees}
+\label{sec:control-redistribution}
 
@@
     \frac {\overline X_\beta} {n_i}\exp\left(- 2(\overline X_i - \overline X_\beta)^2 n_i\right)
-      & \mbox{if $i=\alpha$} \\
+      & \mbox{{\rm if} $i=\alpha$} \\
     \frac {(1-\overline  X_\alpha)} {n_i}\exp\left(- 2(\overline X_\alpha - \overline X_i)^2 n_i\right)
-      &  \mbox{otherwise}
+      &  \mbox{\rm otherwise}
   \end{array}  \right.\nonumber\\
@@
     +\frac {\overline X_\beta} {n_i}\exp\left(-2 n_i(\overline X_i-y)^2\right)
-      & \mbox{if $i=\alpha$} \\
+      & \mbox{{\rm if} $i=\alpha$} \\
 \frac {y-\overline X_\alpha} {n_i}\exp\left(-2 n_i(\overline X_\alpha-\overline
   X_i)^2\right)+\frac {1-\overline X_\alpha} {n_i}\exp\left(-2 n_i(y-\overline X_i)^2\right)
-      &  \mbox{otherwise}
+      &  \mbox{\rm otherwise}
   \end{array} \right.\right) \nonumber\\
@@
     2\frac {\overline X_\beta} {n_i}\exp\left(- \frac {(\overline X_i - \overline X_\beta)^2 n_i} {\frac {14} 3 \frac {n_i} {n_i-1}(\overline X_i - \overline X_\beta) + 2 \overline X_i(1-\overline X_i)} \right)
-      & \mbox{if $i=\alpha$} \\
+      & \mbox{{\rm if} $i=\alpha$} \\
     2\frac {(1-\overline  X_\alpha)} {n_i}\exp\left(- \frac {(\overline X_\alpha - \overline X_i)^2 n_i} {\frac {14} 3 \frac {n_i} {n_i-1}(\overline X_\alpha - \overline X_i) + 2 \overline X_i(1-\overline X_i)} \right)
-      &  \mbox{otherwise}
+      &  \mbox{\rm otherwise}
   \end{array} \right.\right)
@@
 \end{itemize}
-
-Figure~\ref{fig:uct-against-vct}, Figure~\ref{fig:uct-against-ect},
-Figure~\ref{fig:uct-against-bct}, Figure~\ref{fig:best-winning-rate}.
+While the UCT engine is not the most powerful engine of Pachi, it is still
+a strong player. On the other hand, additional
+features of more advanced engines would obstruct the Monte Carlo
+sampling phenomena which are the subject of the experiment.
 
@@
 \end{figure}
-
 \begin{figure}[h]
@@
 \end{figure}
+The engines were compared on the 9x9 board, for 5000, 7000, 1000, and
+15000 samples (game simulations) per ply, each experiment was repeated
+1000 times. Figures~\ref{fig:uct-against-vct}--\ref{fig:uct-against-bct} show the
+winning rate of UCT against VCT, ECT, BCT, correspondingly, vs. the
+stopping threshold (if the maximum VOI of a sample is below the
+threshold, the simulation is stopped, and a move is chosen). Each curve
+in the figures corresponds to a certain number of samples per ply. The
+lower the curve, the better is the corresponding policy compared to UCT.
+
+As the results show, without sample redistribution UCT is as good as
+better as the VOI-aware policies for lower numbers of samples per ply
+(5000, 7000), and the advantage of VOI-aware policies for larger
+number of samples is rather modest (e.g., ECT wins in 55\% of games for
+10000 samples per ply). For the stopping threshold of $10^-6$,
+however, the VOI-aware policies are almost always better than UCT
+(only VCT for 5000 samples wins in less than 50\% of games), with
+stronger policies (ECT, VCT) reaching the winning rate of 67\%.
+
+In agreement with intuition
+(Section~\ref{sec:control-redistribution}), VOI-based stopping and sample
+redistribution is most influential for medium numbers of samples per
+ply. When the maximum number of samples is too low, early stopping would
+result in poorly selected moves. On the other hand, when the maximum
+number of samples is sufficiently high, the VOI of increasing the
+maximum number of samples in a future state is low.
 
@@
 \end{figure}
+Figure~\ref{fig:best-winning-rate}
 
@@
   \begin{array}{l l}
-    \exp\left(- \gamma(\overline X_i - \overline X_\beta)^2 n_i \right)(\overline X_\beta-B) & \mbox{if $i=\alpha$} \\
-    \exp\left(- \gamma(\overline X_\alpha - \overline X_i)^2 n_i \right)(A-\overline  X_\alpha)  &  \mbox{otherwise}
+    \exp\left(- \gamma(\overline X_i - \overline X_\beta)^2 n_i \right)(\overline X_\beta-B) & \mbox{{\rm if} $i=\alpha$} \\
+    \exp\left(- \gamma(\overline X_\alpha - \overline X_i)^2 n_i \right)(A-\overline  X_\alpha)  &  \mbox{\rm otherwise}
   \end{array} \right.

### 2011-12-08 paper/vago/paper.tex
empirical evaluation done

@@
 (only VCT for 5000 samples wins in less than 50\% of games), with
-stronger policies (ECT, VCT) reaching the winning rate of 67\%.
+stronger policies (ECT, VCT) reaching the winning rate of 66\%.
 
@@
 \end{figure}
-Figure~\ref{fig:best-winning-rate}
+Figure~\ref{fig:best-winning-rate} presents relative performance of
+different VOI-aware policies depending on the number of samples per
+ply by comparing the best achieved winning rates against UCT (again,
+the lower the bar, the better is the policy). Both ECT (based on the
+Hoeffding inequality with midpoint) and BCT (based on the empirical
+Benrstein inequality) are better than VCT. The difference between ECT
+and VCT is most prominent for smaller number of samples, and the
+winning rate of ECT increases relatively slowly with the number of
+samples. In contrast, BCT is worse than ECT and only slightly better
+than VCT for smaller number of samples, but the winning rate rapidly
+increases with the number of samples, and for 10000 and 15000 samples
+per ply BCT is the best of the policies with the winning rate reaching
+67\%.
 

### 2011-12-08 paper/vago/paper.tex
draft done

@@
 number of samples is rather modest (e.g., ECT wins in 55\% of games for
-10000 samples per ply). For the stopping threshold of $10^-6$,
+10000 samples per ply). For the stopping threshold of $10^{-6}$,
 however, the VOI-aware policies are almost always better than UCT

### 2011-12-08 paper/vago/paper.tex
spelling edits

@@
 \begin{description}
-\item[Hoeffding's inequality \rm{\cite{Hoeffding.ineq}}:] 
+\item[Hoeffding inequality \rm{\cite{Hoeffding.ineq}}:] 
 \begin{equation}
@@
 \end{equation}
-\item[Empirical Bernstein's inequality
+\item[Empirical Bernstein inequality
   \rm{\cite{MaurerPontil.benrstein}} \textrm{(derived in Appendix~\ref{app:deriv-conc-empbernstein})}:]
@@
 \Pr(\overline X_i' \le \overline X_\beta|i=\alpha) &
-   \le & 2\exp\left(- 2\left(\frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}}(\overline X_i - \overline X_\beta)\right)^2 n_i \right)
-   \le 2\exp\left(- 1.37(\overline X_i - \overline X_\beta)^2 n_i \right) \nonumber\\
+   \le & \exp\left(- 2\left(\frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}}(\overline X_i - \overline X_\beta)\right)^2 n_i \right)
+   \le \exp\left(- 1.37(\overline X_i - \overline X_\beta)^2 n_i \right) \nonumber\\
 \Pr(\overline X_i' \ge \overline X_\alpha|i\ne\alpha) &
-   \le & 2\exp\left(- 2\left(\frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}} (\overline X_\alpha - \overline X_i)\right)^2 n_i \right)
-    \le 2\exp\left(- 1.37(\overline X_\alpha - \overline X_i)^2 n_i \right)
+   \le & \exp\left(- 2\left(\frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}} (\overline X_\alpha - \overline X_i)\right)^2 n_i \right)
+    \le \exp\left(- 1.37(\overline X_\alpha - \overline X_i)^2 n_i \right)
 \label{eqn:probound-blnk-hoeffding}
@@
   \begin{array}{l l}
-    C \frac {\overline X_\beta} {n_i}\exp\left(- 1.37(\overline X_i - \overline X_\beta)^2 n_i\right)
+    N \frac {\overline X_\beta} {n_i}\exp\left(- 1.37(\overline X_i - \overline X_\beta)^2 n_i\right)
       & \mbox{{\rm if} $i=\alpha$} \\
-    C \frac {(1-\overline  X_\alpha)} {n_i} \exp\left(- 1.37(\overline X_\alpha - \overline X_i)^2 n_i\right)
+    N \frac {(1-\overline  X_\alpha)} {n_i} \exp\left(- 1.37(\overline X_\alpha - \overline X_i)^2 n_i\right)
       &  \mbox{\rm otherwise}
@@
 \end{equation}
-where $C=2N$. The bound is generally tighter when $N$ is small relative to $n_i$,
+The bound is generally tighter when $N$ is small relative to $n_i$,
 and the estimate $\hat\Lambda_i^b$ can be viewed as a relaxation of the myopic VOI
@@
 
-As the results show, without sample redistribution UCT is as good as
-better as the VOI-aware policies for lower numbers of samples per ply
+As the results show, without sample redistribution UCT is as good as or
+better than the VOI-aware policies for lower numbers of samples per ply
 (5000, 7000), and the advantage of VOI-aware policies for larger
@@
 
-For any $\delta$, the probability that $\overline X_i'$ is \textit{not
-greater} than
-$X_\alpha$ is more than the probability that $\IE[X_i]\le\overline
-X_i+\delta$ \textit{and} $\overline X_i'\ge \IE[X_i]+\overline X_\alpha - \overline
-  X_i - \delta +(\overline X_\alpha - \overline X_i)\frac {n_i} N$;
-therefore, the probability that $\overline X_i'$ is \textit{greater}
+For any $\delta$, the probability that $\overline X_i'$ is greater
 than $X_\alpha$ is less than the probability that $\IE[X_i]\ge\overline
-X_i+\delta$ \textit{or} $\overline X_i'\le \IE[X_i]+\overline X_\alpha - \overline
-  X_i - \delta +(\overline X_\alpha - \overline X_i)\frac {n_i} N$.
-Using the union bound,
+X_i+\delta$ \emph{or} $\overline X_i'\le \IE[X_i]+\overline X_\alpha - \overline
+  X_i - \delta +(\overline X_\alpha - \overline X_i)\frac {n_i} N$, thus
+less than the maximum of probabilities of the two events:
 \begin{equation}
 \Pr(\overline X_i'\ge \overline X_\alpha)\le
-  \Pr(\IE[X_i]-\overline X_i \ge \delta) +
-  \Pr(\overline X_i^+ - \IE[X_i] \ge \overline X_\alpha - \overline
-  X_i - \delta +(\overline X_\alpha - \overline X_i)\frac {n_i} N)
+  \max\left(\Pr(\IE[X_i]-\overline X_i \ge \delta),\,
+  \Pr\left(\overline X_i^+ - \IE[X_i] \ge \overline X_\alpha - \overline
+  X_i - \delta +(\overline X_\alpha - \overline X_i)\frac {n_i} N\right)\right)
 \end{equation}
@@
 \Pr(\overline X_i'\ge \overline X_\alpha)\le 
-  \exp(-2\delta^2n_i)+\exp\left(-2\left((\overline X_\alpha - \overline
-  X_i)\left(1+\frac {n_i} N\right) - \delta\right)^2N\right)
+\max\left(\exp(-2\delta^2n_i),\,
+          \exp\left(-2\left((\overline X_\alpha 
+          - \overline X_i)\left(1+\frac {n_i} N\right)
+          - \delta\right)^2N\right)\right)
 \end{equation}
@@
 \Pr(\overline X_i'\ge \overline X_\alpha) 
-& \le & 2\exp\left(-2\left( \frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}}(\overline X_\alpha - \overline X_i)\right)^2 n_i\right)\nonumber \\
-& \le & 2\exp(-2(\sqrt 2 - 1)^2(\overline X_\alpha - \overline X_i)^2n_i)
- \le 2\exp(-1.37(\overline X_\alpha - \overline X_i)^2n_i)
+& \le & \exp\left(-2\left( \frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}}
+                          (\overline X_\alpha - \overline X_i)\right)^2 n_i\right)\nonumber \\
+& \le & \exp(-2(\sqrt 2 - 1)^2(\overline X_\alpha - \overline X_i)^2n_i)
+ \le \exp(-1.37(\overline X_\alpha - \overline X_i)^2n_i)
 \end{eqnarray}

### 2011-12-08 paper/vago/paper.tex
spelling edits

@@
 as approximations of VOI estimates for myopic or
-receding horison policies. As illustrated by the empirical evaluation
+receding horizon policies. As illustrated by the empirical evaluation
 (Section~\ref{sec:empirical-evaluation}), estimates based on upper
@@
 
-One way to account for this re-use would be to incroprorate the
+One way to account for this re-use would be to incorporate the
 `future' value of information into a VOI estimate. However, this 
@@
 Hoeffding inequality with midpoint) and BCT (based on the empirical
-Benrstein inequality) are better than VCT. The difference between ECT
+Bernstein inequality) are better than VCT. The difference between ECT
 and VCT is most prominent for smaller number of samples, and the

### 2011-12-08 paper/vago/paper.tex
eliminated 2 in bounds

@@
 For any $\delta$, the probability that $\overline X_i'$ is greater
-than $X_\alpha$ is less than the probability that $\IE[X_i]\ge\overline
+than $\overline X_\alpha$ is less than the probability that $\IE[X_i]\ge\overline
 X_i+\delta$ \emph{or} $\overline X_i'\le \IE[X_i]+\overline X_\alpha - \overline
@@
           - \delta\right)^2N\right)\right)
+\label{eqn:app-hoeffding-le-maxexp}
 \end{equation}
-$\delta^2n = -2\left((\overline X_\alpha - \overline X_i)
-(1+\frac {n_i} N) - \delta\right)^2N$ when
-$\delta=\frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}} (\overline X_\alpha
-- \overline X_i) \ge (\sqrt 2 - 1)(\overline X_\alpha-\overline X_i)$,
-therefore
+The maximum is equal to either of the arguments when the arguments are equal. Let
+\begin{equation}
+\exp(-\delta^2n) = \exp(-2\left((\overline X_\alpha - \overline X_i)(1+\frac {n_i} N) - \delta\right)^2N)\label{eqn:app-hoeffding-eq-exp}
+\end{equation}
+Solve (\ref{eqn:app-hoeffding-eq-exp}) for $\delta$: $\delta=\frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}} (\overline X_\alpha
+- \overline X_i) \ge (\sqrt 2 - 1)(\overline X_\alpha-\overline X_i)$. Substitute $\delta$ into 
+(\ref{eqn:app-hoeffding-le-maxexp}) and obtain
 \begin{eqnarray}

### 2011-12-09 paper/vago/paper.tex
in progress relwork
related work

@@
 \section{Related Work}
+
+Efficient algorithms for Multi-Armed Bandits based on
+distribution-independent bounds, in particular UCB1, are introduced in
+\cite{Auer.ucb}. The UCT algorithm, an extension of UCB1 to Monte
+Carlo Tree Search is described in \cite{Kocsis.uct}, and a successful
+application of UCT to playing the Go game is discussed in
+\cite{Gelly.mogo}.
+
+Pure exploration in multi-armed bandit problems is explored in
+\cite{Bubeck.pure}. On the one hand, the paper proves certain upper
+and lower bounds for UCB1 and uniform sampling, showing that an upper
+bound on the simple regret is exponential in the number of samples for
+uniform sampling, while only polynomial for UCB1. On the other hand,
+empirical performance of UCB1 appears to be better than of uniform
+sampling. \cite{Mnih.bernstop} investigate stopping criteria for
+sampling based on the empirical Bernstein inequality; however, the
+stopping criteria are based on error probabilities rather than on
+value of information measures, and do not directly address the
+objective of regret minimization.
+
+The principles of bounded rationality appear in
+\cite{Horvitz.reasoningabout}. \cite{Russell.right} provided a formal
+description of rational metareasoning and case studies of applications
+in several problem domains. One obstacle to straightforward
+application of the principles of rational metareasoning to Monte Carlo
+sampling is the metagreedy assumption, according to which samples must
+be selected as though at most one sample can be taken before an action
+is chosen. In Monte Carlo sampling, the value of information of any
+single sample in a given search state is often zero, so a different
+approximating assumption must be used instead.
 
@@
 game \cite{Gelly.mogo}. This series of experiments compares
-the winning rate of VOI-aware policies (VCT, ECT, BCT) against UCT 
+the winning rate of UCT against VOI-aware policies (VCT, ECT, BCT)
 in Go and investigates the effect of early stopping and sample

### 2011-12-09 paper/vago/paper.tex
related work
editorial changes

@@
 
-\textit{TODO}
-
 \section{Some Concentration Inequalities}
@@
 \item the arm ($\alpha$) with the highest sample mean is pulled, and the 
-mean of the arm is lower than the sample mean the second-best arm ($\beta$);
-\item another arm is pulled, and the mean of the arm is higher
+mean of the arm becomes lower than the sample mean of the second-best arm ($\beta$);
+\item another arm is pulled, and the mean of the arm becomes higher
 than the current highest sample mean ($\alpha$).
@@
 $\overline X'_i$ substituted instead of $\IE[X_i]$, and, if
-$i\ne\alpha$, the upper bound on $X_i'$ is $X_\alpha+(1-X_\alpha)
-\frac N {N+n_i}$; if $i=\alpha$, the lower bound on $X_i'$ is
+$i\ne\alpha$, an upper bound on $X_i'$ is $X_\alpha+(1-X_\alpha)
+\frac N {N+n_i}$; if $i=\alpha$, a lower bound on $X_i'$ is
 $X_\beta - X_\beta \frac N {N+n_i}$.
@@
 
-\begin{figure}[h]
+\begin{figure}[t]
 \centering
@@
 \end{figure}
-\begin{figure}[h]
+\begin{figure}[t]
 \centering
@@
 \textit{TODO}
-
-\clearpage
 

### 2011-12-09 paper/vago/paper.tex
consistent formulas

@@
 \begin{itemize}
-\item the arm ($\alpha$) with the highest sample mean is pulled, and the 
-mean of the arm becomes lower than the sample mean of the second-best arm ($\beta$);
-\item another arm is pulled, and the mean of the arm becomes higher
-than the current highest sample mean ($\alpha$).
+\item the arm $\alpha$ with the highest sample mean is pulled, and the 
+true mean of the arm is lower than the sample mean of the second-best arm $\beta$;
+\item another arm is pulled, and the true mean of the arm is higher
+than the sample mean of the best arm $\alpha$.
 \end{itemize}
@@
 maximum possible increase in the reward:
-
 \begin{thm} The intrinsic value of perfect information $\Lambda_i^p$ about the $i$th arm is
@@
   \begin{array}{l l}
-    \frac {\overline X_\beta-y} {n_i}\exp\left(-2 n_i(\overline X_i-\overline X_\beta)^2\right)
-    +\frac {\overline X_\beta} {n_i}\exp\left(-2 n_i(\overline X_i-y)^2\right)
+    \frac {\overline X_\beta-y} {n_i}\exp\left(-2 (\overline X_i-\overline X_\beta)^2n_i\right)
+    +\frac {\overline X_\beta} {n_i}\exp\left(-2 (\overline X_i-y)^2n_i\right)
       & \mbox{{\rm if} $i=\alpha$} \\
-\frac {y-\overline X_\alpha} {n_i}\exp\left(-2 n_i(\overline X_\alpha-\overline
-  X_i)^2\right)+\frac {1-\overline X_\alpha} {n_i}\exp\left(-2 n_i(y-\overline X_i)^2\right)
+\frac {y-\overline X_\alpha} {n_i}\exp\left(-2(\overline X_\alpha-\overline X_i)^2n_i\right)
+    +\frac {1-\overline X_\alpha} {n_i}\exp\left(-2 (y-\overline X_i)^2n_i\right)
       &  \mbox{\rm otherwise}
@@
   \begin{array}{l l}
-    \exp\left(- \gamma(\overline X_i - \overline X_\beta)^2 n_i \right)(\overline X_\beta-B) & \mbox{{\rm if} $i=\alpha$} \\
-    \exp\left(- \gamma(\overline X_\alpha - \overline X_i)^2 n_i \right)(A-\overline  X_\alpha)  &  \mbox{\rm otherwise}
+    (\overline X_\beta-B)\exp\left(- \gamma(\overline X_i - \overline X_\beta)^2 n_i \right) & \mbox{{\rm if} $i=\alpha$} \\
+    (A-\overline  X_\alpha)\exp\left(- \gamma(\overline X_\alpha - \overline X_i)^2 n_i \right)  &  \mbox{\rm otherwise}
   \end{array} \right.
@@
 \Lambda_{i|i=\alpha} \le \hat \Lambda_{i|i=\alpha} =
- (\overline X_\beta-y)\exp\left(-\gamma n_i(\overline X_i-\overline X_\beta)^2\right)
- +(\overline X_\beta-B)\exp\left(-\gamma n_i(\overline X_i-y)^2\right)
+ (\overline X_\beta-y)\exp\left(-\gamma (\overline X_i-\overline X_\beta)^2n_i\right)
+ +(\overline X_\beta-B)\exp\left(-\gamma (\overline X_i-y)^2n_i\right)
 \label{eqn:bound-twopart-hoeffding-alpha}
@@
 \begin{equation}
-2\gamma n_i(\overline X_\beta-B)(\overline X_i-y)=\exp\left(-\gamma n_i\left(\frac {\overline X_i-\overline X_\beta}
-    {\overline X_i-y}\right)^2\right)
+  2\gamma (\overline X_\beta-B)(\overline X_i-y)n_i = \exp\left(-\gamma \left(\frac {\overline X_i-\overline X_\beta}
+    {\overline X_i-y}\right)^2n_i\right)
 \end{equation}
@@
 \begin{equation}
-\Lambda_{i|i\ne\alpha} \le \hat\Lambda_{i|i\ne\alpha} = (y-\overline X_\alpha)\exp\left(-\gamma n_i(\overline X_\alpha-\overline
-  X_i)^2\right)+(A-\overline X_\alpha)\exp\left(-\gamma n_i(y-\overline X_i)^2\right)
+\Lambda_{i|i\ne\alpha} \le \hat\Lambda_{i|i\ne\alpha} = (y-\overline X_\alpha)\exp\left(-\gamma(\overline X_\alpha-\overline
+  X_i)^2 n_i\right)+(A-\overline X_\alpha)\exp\left(-\gamma (y-\overline X_i)^2n_i\right)
 \label{eqn:bound-twopart-hoeffding-rest}

### 2011-12-09 paper/vago/paper.tex
paper draft done, all parts in place

@@
 \end{equation}
-\item[Empirical Bernstein inequality
-  \rm{\cite{MaurerPontil.benrstein}} \textrm{(derived in Appendix~\ref{app:deriv-conc-empbernstein})}:]
+\item[Empirical Bernstein inequality \rm{(derived in Appendix~\ref{app:deriv-conc-empbernstein})}:]
 \begin{eqnarray}
@@
 
-\textit{TODO}
+This work suggested Monte Carlo sampling policies in which sample
+selection and termination are based on upper bounds on the value of
+information. A simplified model of accounting for future value of
+information of a sample, based on early termination and sample
+redistribution, was proposed for Monte Carlo sampling in
+trees. Empirical evaluation showed that the policies outperform
+heuristic algorithms for simple regret minimization in multi-armed
+bandits, as well as for tree search.
 
-\appendix
+Monte Carlo tree search still remains a largely unexplored field of
+application of VOI-aware algorithms. More elaborated VOI estimates,
+taking into consideration re-use of samples in future search states
+should be considered. The policies introduced in the paper differ from
+the UCT algorithm only at the first step, where the VOI-aware
+decisions are made. Consistent application of principles of rational
+metareasoning at all steps of a rollout may further improve the
+sampling policies.
 

### 2011-12-09 paper/vago/paper.tex
Monte Carlo => Monte-Carlo

@@
 
-\title{VOI-aware Monte Carlo Sampling in Trees}
+\title{VOI-aware Monte-Carlo Sampling in Trees}
 \author {David Tolpin, Solomon Eyal Shimony \\
@@
 distribution-independent bounds, in particular UCB1, are introduced in
-\cite{Auer.ucb}. The UCT algorithm, an extension of UCB1 to Monte
-Carlo Tree Search is described in \cite{Kocsis.uct}, and a successful
+\cite{Auer.ucb}. The UCT algorithm, an extension of UCB1 to
+Monte-Carlo Tree Search is described in \cite{Kocsis.uct}, and a successful
 application of UCT to playing the Go game is discussed in
@@
 in several problem domains. One obstacle to straightforward
-application of the principles of rational metareasoning to Monte Carlo
+application of the principles of rational metareasoning to Monte-Carlo
 sampling is the metagreedy assumption, according to which samples must
 be selected as though at most one sample can be taken before an action
-is chosen. In Monte Carlo sampling, the value of information of any
+is chosen. In Monte-Carlo sampling, the value of information of any
 single sample in a given search state is often zero, so a different
@@
 all. The \textit{myopic} VOI estimate is of limited applicability to
-Monte Carlo sampling, since the effect of a single sample is small,
+Monte-Carlo sampling, since the effect of a single sample is small,
 and the myopic VOI estimate will often be non-positive, resulting in premature
@@
 The sampling policies are first compared on random multi-armed bandit 
-problem instances, where the Monte Carlo Tree Search algorithm
+problem instances, where the Monte-Carlo Tree Search algorithm
 is reduced to the sample selection policy at the first step: UCT is 
@@
 a strong player. On the other hand, additional
-features of more advanced engines would obstruct the Monte Carlo
+features of more advanced engines would obstruct the Monte-Carlo
 sampling phenomena which are the subject of the experiment.
@@
 
-This work suggested Monte Carlo sampling policies in which sample
+This work suggested Monte-Carlo sampling policies in which sample
 selection and termination are based on upper bounds on the value of
@@
 information of a sample, based on early termination and sample
-redistribution, was proposed for Monte Carlo sampling in
+redistribution, was proposed for Monte-Carlo sampling in
 trees. Empirical evaluation showed that the policies outperform
@@
 
-Monte Carlo tree search still remains a largely unexplored field of
+Monte-Carlo tree search still remains a largely unexplored field of
 application of VOI-aware algorithms. More elaborated VOI estimates,

### 2011-12-10 paper/vago/paper.tex
typo fixed

@@
 Problem is the expected difference between the best expected reward
-$\mu_*$ and the expected reward $\mu_j$ of the arm with the best sample mean
-$\overline X_j=\max_i\overline X_i$:
+$\max\limits_i\IE[X_i]$ and the expected reward $\IE[X_j]$ of the arm
+with the best sample mean $\overline X_j=\max\limits_i\overline X_i$:
 \begin{equation}
@@
 \end{equation}
-where $\Delta_j=\mu_*-\mu_j$.
-Strategies that minimize the simple regret are called pure exploration
-strategies \cite{Bubeck.pure}. Principles of rational metareasoning
+where $\Delta_j=\max\limits_i\IE[X_i]-\IE[X_j]$.  Strategies that
+minimize the simple regret are called pure exploration strategies
+\cite{Bubeck.pure}. Principles of rational metareasoning
 \cite{Russell.right} suggest that at each step the arm with the great
@@
 
-Pure exploration in multi-armed bandit problems is explored in
+Pure exploration in Multi-armed bandits is explored in
 \cite{Bubeck.pure}. On the one hand, the paper proves certain upper
@@
 \begin{equation}
-\Pr(X-\IE[X] \ge a) \le \exp ( -2na^2)
+\Pr(X-\IE[X] \ge a) \le \exp ( -2a^2n)
 \label{eqn:conc-hoeffding}
@@
 \begin{eqnarray}
-\Pr(X-\IE[X] \ge a) &\le& 2\exp \left( - \frac {na^2} {\frac {14} {3}
+\Pr(X-\IE[X] \ge a) &\le& 2\exp \left( - \frac {a^2n} {\frac {14} {3}
                           \frac {n} {n-1}a+2\overline\sigma_n^2}\right)\nonumber\\
-                    &\le& 2\exp \left( - \frac {na^2} {10a+2\overline\sigma_n^2}\right)
+                    &\le& 2\exp \left( - \frac {a^2n} {10a+2\overline\sigma_n^2}\right)
 \label{eqn:conc-empbernstein}
@@
 The search time is finite, and in a simple case the \textit{search
-  budget} specified as the maximum number of samples.
+  budget} is specified as the maximum number of samples.
 An estimate based on the perfect intrinsic VOI does not take in
@@
     \overline X_\beta\exp\left(- 2(\overline X_i - \overline X_\beta)^2 n_i \right) & \mbox{{\rm if} $i=\alpha$} \\
-    (1-\overline  X_\alpha)\exp\left(- 2(\overline X_\alpha - \overline X_i)^2 n_i \right)(1-\overline  X_\alpha)  &  \mbox{\rm otherwise}
+    (1-\overline  X_\alpha)\exp\left(- 2(\overline X_\alpha - \overline X_i)^2 n_i \right)  &  \mbox{\rm otherwise}
   \end{array} \right.
@@
 simple regret in Multi-armed bandits at the first step, and the
-cumulative regret (solves the exploration-exploitation trade-off)
-during the rest of the rollout.  However, this approach assumes that
-the information obtained from rollouts in the current state is
-discarded after an action is selected. In practice, most successful
-Monte-Carlo tree search algorithms re-use rollouts originating from an
-earlier search state and passing through the current search state;
-thus, the value of information of a rollout is determined not just by
-the influence on the choice of the action at the current state, but
-also by the influence on the choice at future search states, provided
-the rollout passes through the search states which will be visited.
+cumulative regret during the rest of the rollout.  However, this
+approach assumes that the information obtained from rollouts in the
+current state is discarded after an action is selected. In practice,
+most successful Monte-Carlo tree search algorithms re-use rollouts
+originating from an earlier search state and passing through the
+current search state; thus, the value of information of a rollout is
+determined not just by the influence on the choice of the action at
+the current state, but also by the influence on the choice at future
+search states, provided the rollout passes through the search states
+which will be visited.
 
@@
 
-The sampling policies are first compared on random multi-armed bandit 
+The sampling policies are first compared on random Multi-armed bandit 
 problem instances, where the Monte-Carlo Tree Search algorithm
 is reduced to the sample selection policy at the first step: UCT is 
-actually UCB1. In the problem of pure exploration in multi-armed bandits
+actually UCB1. In the problem of pure exploration in Multi-armed bandits
 \cite{Bubeck.pure}, the smaller is the exploration factor, the higher
@@
 Figure~\ref{fig:random-instances} shows experiment results for
-randomly-generated multi-armed bandits with 32 Bernoulli arms, with
+randomly-generated Multi-armed bandits with 32 Bernoulli arms, with
 the mean rewards of the arms distributed uniformly in the range $[0,
@@
 redistribution, was proposed for Monte-Carlo sampling in
-trees. Empirical evaluation showed that the policies outperform
-heuristic algorithms for simple regret minimization in multi-armed
+trees. Empirical evaluation showed that these policies outperform
+heuristic algorithms for simple regret minimization in Multi-armed
 bandits, as well as for tree search.

### 2011-12-10 paper/vago/paper.tex
fixed typos in proof of lemma 1

@@
 than $\overline X_\alpha$ is less than the probability that $\IE[X_i]\ge\overline
-X_i+\delta$ \emph{or} $\overline X_i'\le \IE[X_i]+\overline X_\alpha - \overline
+X_i+\delta$ \emph{or} $\overline X_i^+\ge \IE[X_i]+\overline X_\alpha - \overline
   X_i - \delta +(\overline X_\alpha - \overline X_i)\frac {n_i} N$, thus

### 2011-12-10 paper/vago/paper.tex
2 in blinkered is back, that was a temporary blackout, sorry

--- 2011-12-10 paper/vago/paper.tex
@@
 \Pr(\overline X_i' \le \overline X_\beta|i=\alpha) &
-   \le & \exp\left(- 2\left(\frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}}(\overline X_i - \overline X_\beta)\right)^2 n_i \right)
-   \le \exp\left(- 1.37(\overline X_i - \overline X_\beta)^2 n_i \right) \nonumber\\
+   \le & 2\exp\left(- 2\left(\frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}}(\overline X_i - \overline X_\beta)\right)^2 n_i \right)
+   \le 2\exp\left(- 1.37(\overline X_i - \overline X_\beta)^2 n_i \right) \nonumber\\
 \Pr(\overline X_i' \ge \overline X_\alpha|i\ne\alpha) &
-   \le & \exp\left(- 2\left(\frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}} (\overline X_\alpha - \overline X_i)\right)^2 n_i \right)
-    \le \exp\left(- 1.37(\overline X_\alpha - \overline X_i)^2 n_i \right)
+   \le & 2\exp\left(- 2\left(\frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}} (\overline X_\alpha - \overline X_i)\right)^2 n_i \right)
+    \le 2\exp\left(- 1.37(\overline X_\alpha - \overline X_i)^2 n_i \right)
 \label{eqn:probound-blnk-hoeffding}
@@
   \begin{array}{l l}
-    \overline X_\beta\exp\left(- 2(\overline X_i - \overline X_\beta)^2 n_i \right) & \mbox{{\rm if} $i=\alpha$} \\
-    (1-\overline  X_\alpha)\exp\left(- 2(\overline X_\alpha - \overline X_i)^2 n_i \right)  &  \mbox{\rm otherwise}
+    \overline X_\beta\cdot \exp\left(- 2(\overline X_i - \overline X_\beta)^2 n_i \right) & \mbox{{\rm if} $i=\alpha$} \\
+    (1-\overline  X_\alpha)\cdot \exp\left(- 2(\overline X_\alpha - \overline X_i)^2 n_i \right)  &  \mbox{\rm otherwise}
   \end{array} \right.
@@
   \begin{array}{l l}
-    N \frac {\overline X_\beta} {n_i}\exp\left(- 1.37(\overline X_i - \overline X_\beta)^2 n_i\right)
+    N\frac {\overline X_\beta} {n_i}\cdot 2\exp\left(- 1.37(\overline X_i - \overline X_\beta)^2 n_i\right)
       & \mbox{{\rm if} $i=\alpha$} \\
-    N \frac {(1-\overline  X_\alpha)} {n_i} \exp\left(- 1.37(\overline X_\alpha - \overline X_i)^2 n_i\right)
+    N\frac {(1-\overline  X_\alpha)} {n_i}\cdot 2\exp\left(- 1.37(\overline X_\alpha - \overline X_i)^2 n_i\right)
       &  \mbox{\rm otherwise}
@@
   \begin{array}{l l}
-    \frac {\overline X_\beta} {n_i}\exp\left(- 2(\overline X_i - \overline X_\beta)^2 n_i\right)
+    \frac {\overline X_\beta} {n_i} \cdot \exp\left(- 2(\overline X_i - \overline X_\beta)^2 n_i\right)
       & \mbox{{\rm if} $i=\alpha$} \\
-    \frac {(1-\overline  X_\alpha)} {n_i}\exp\left(- 2(\overline X_\alpha - \overline X_i)^2 n_i\right)
+    \frac {(1-\overline  X_\alpha)} {n_i}\cdot \exp\left(- 2(\overline X_\alpha - \overline X_i)^2 n_i\right)
       &  \mbox{\rm otherwise}
@@
   \begin{array}{l l}
-    \frac {\overline X_\beta-y} {n_i}\exp\left(-2 (\overline X_i-\overline X_\beta)^2n_i\right)
-    +\frac {\overline X_\beta} {n_i}\exp\left(-2 (\overline X_i-y)^2n_i\right)
+    \frac {\overline X_\beta-y} {n_i}\cdot \exp\left(-2 (\overline X_i-\overline X_\beta)^2n_i\right)
+    +\frac {\overline X_\beta} {n_i}\cdot \exp\left(-2 (\overline X_i-y)^2n_i\right)
       & \mbox{{\rm if} $i=\alpha$} \\
-\frac {y-\overline X_\alpha} {n_i}\exp\left(-2(\overline X_\alpha-\overline X_i)^2n_i\right)
-    +\frac {1-\overline X_\alpha} {n_i}\exp\left(-2 (y-\overline X_i)^2n_i\right)
+\frac {y-\overline X_\alpha} {n_i}\cdot \exp\left(-2(\overline X_\alpha-\overline X_i)^2n_i\right)
+    +\frac {1-\overline X_\alpha} {n_i}\cdot \exp\left(-2 (y-\overline X_i)^2n_i\right)
       &  \mbox{\rm otherwise}
@@
   \begin{array}{l l}
-    2\frac {\overline X_\beta} {n_i}\exp\left(- \frac {(\overline X_i - \overline X_\beta)^2 n_i} {\frac {14} 3 \frac {n_i} {n_i-1}(\overline X_i - \overline X_\beta) + 2 \overline X_i(1-\overline X_i)} \right)
+    \frac {\overline X_\beta} {n_i}\cdot 2\exp\left(- \frac {(\overline X_i - \overline X_\beta)^2 n_i} {\frac {14} 3 \frac {n_i} {n_i-1}(\overline X_i - \overline X_\beta) + 2 \overline X_i(1-\overline X_i)} \right)
       & \mbox{{\rm if} $i=\alpha$} \\
-    2\frac {(1-\overline  X_\alpha)} {n_i}\exp\left(- \frac {(\overline X_\alpha - \overline X_i)^2 n_i} {\frac {14} 3 \frac {n_i} {n_i-1}(\overline X_\alpha - \overline X_i) + 2 \overline X_i(1-\overline X_i)} \right)
+    \frac {(1-\overline  X_\alpha)} {n_i}\cdot 2\exp\left(- \frac {(\overline X_\alpha - \overline X_i)^2 n_i} {\frac {14} 3 \frac {n_i} {n_i-1}(\overline X_\alpha - \overline X_i) + 2 \overline X_i(1-\overline X_i)} \right)
       &  \mbox{\rm otherwise}
@@
 For any $\delta$, the probability that $\overline X_i'$ is greater
-than $\overline X_\alpha$ is less than the probability that $\IE[X_i]\ge\overline
-X_i+\delta$ \emph{or} $\overline X_i'\le \IE[X_i]+\overline X_\alpha - \overline
-  X_i - \delta +(\overline X_\alpha - \overline X_i)\frac {n_i} N$, thus
-less than the maximum of probabilities of the two events:
+than $\overline X_\alpha$ is less than the probability that
+$\IE[X_i]\ge\overline X_i+\delta$
+\emph{or} $\overline X_i^+\ge \IE[X_i]+\overline X_\alpha
+- \overline X_i - \delta +(\overline X_\alpha - \overline X_i)\frac {n_i} N$,
+thus, by the union bound, less than the sum of the probabilities:
 \begin{equation}
 \Pr(\overline X_i'\ge \overline X_\alpha)\le
-  \max\left(\Pr(\IE[X_i]-\overline X_i \ge \delta),\,
-  \Pr\left(\overline X_i^+ - \IE[X_i] \ge \overline X_\alpha - \overline
-  X_i - \delta +(\overline X_\alpha - \overline X_i)\frac {n_i} N\right)\right)
+  \Pr(\IE[X_i]-\overline X_i \ge \delta)+
+  \Pr\left(\overline X_i^+ - \IE[X_i] \ge \overline X_\alpha
+           - \overline X_i - \delta +(\overline X_\alpha - \overline X_i)\frac {n_i} N\right)
 \end{equation}
-Bounding the probabilities on the right side using the Hoeffding
+Bounding the probabilities on the right-hand side using the Hoeffding
 inequality, obtain:
@@
 \Pr(\overline X_i'\ge \overline X_\alpha)\le 
-\max\left(\exp(-2\delta^2n_i),\,
+    \exp(-2\delta^2n_i)+
          \exp\left(-2\left((\overline X_\alpha 
          - \overline X_i)\left(1+\frac {n_i} N\right)
-          - \delta\right)^2N\right)\right)
+         - \delta\right)^2N\right)
 \label{eqn:app-hoeffding-le-maxexp}
 \end{equation}
-The maximum is equal to either of the arguments when the arguments are equal. Let
+Find $\delta$ for which the two terms on the right-hand side of
+(\ref{eqn:app-hoeffding-le-maxexp}) are equal:
 \begin{equation}
@@
 \Pr(\overline X_i'\ge \overline X_\alpha) 
-& \le & \exp\left(-2\left( \frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}}
+& \le & 2\exp\left(-2\left( \frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}}
                           (\overline X_\alpha - \overline X_i)\right)^2 n_i\right)\nonumber \\
-& \le & \exp(-2(\sqrt 2 - 1)^2(\overline X_\alpha - \overline X_i)^2n_i)
- \le \exp(-1.37(\overline X_\alpha - \overline X_i)^2n_i)
+& \le & 2\exp(-2(\sqrt 2 - 1)^2(\overline X_\alpha - \overline X_i)^2n_i)
+ \le 2\exp(-1.37(\overline X_\alpha - \overline X_i)^2n_i)
 \end{eqnarray}

### 2011-12-12 paper/vago/paper.tex
fixes
corrections per eyal requests

@@
 \newtheorem{lmm}{Lemma}
+\newtheorem{crl}{Corollary}
 
@@
 \cite{Bubeck.pure}. Principles of rational metareasoning
-\cite{Russell.right} suggest that at each step the arm with the great
-value of information (VOI) must be pulled, and the sampling must be
-stopped and a decision must be made when no arm has positive VOI.
+\cite{Russell.right} suggest that at each step the arm with the
+greatest value of information (VOI) estimate must be pulled, and the
+sampling must be stopped and a decision must be made when no arm has
+positive VOI estimate.
 
-To estimate the VOI of pulling an arm, either a certain 
-distribution of the rewards should be assumed (and updated based on
-observed rewards), or a distribution-independent bound on the VOI can be
-used as the VOI estimate. In this paper, we use {\em concentration inequalities}
-to derive distribution-independent bounds on the VOI.
+To estimate the VOI of pulling an arm, either a certain distribution
+of the rewards should be assumed (and updated based on observed
+rewards), or a distribution-independent bound on the VOI can be used
+as the VOI estimate. In this paper, we use {\em concentration
+  inequalities} to derive distribution-independent bounds on the VOI.
 
@@
 can be bounded from above using concentration inequalities. In
-particular, Lemma~\ref{lemma:hoeffding-prob-bounds} (proved in 
+particular, Theorem~\ref{thm:hoeffding-prob-bounds} (proved in 
 Appendix~\ref{app:hoeffding-prob-bounds-proof}) is based on the
 Hoeffding inequality (\ref{eqn:conc-hoeffding}):
-\begin{lmm} The probabilities in equations (\ref{eqn:thm-vopi},
+\begin{thm} The probabilities in equations (\ref{eqn:thm-vopi},
 \ref{eqn:thm-be}) are bounded from above as
@@
 \end{eqnarray}
-\label{lemma:hoeffding-prob-bounds}
-\end{lmm}
+\label{thm:hoeffding-prob-bounds}
+\end{thm}
 
-An upper bound on the intrinsic value of perfect information is obtained by substituting
+\begin{crl} An upper bound on the intrinsic value of perfect information is obtained by substituting
 (\ref{eqn:probound-perf-hoeffding}) into (\ref{eqn:thm-vopi}):
@@
 \end{equation}
+\label{crl:bound-perf-hoeffding}
+\end{crl}
+\begin{crl}
 An upper bound on the blinkered VOI estimate with ordering independent of $N$ is obtained
@@
 \end{equation}
-The bound is generally tighter when $N$ is small relative to $n_i$,
-and the estimate $\hat\Lambda_i^b$ can be viewed as a relaxation of the myopic VOI
-estimate to longer sequences of actions, such
-that the VOI estimate is positive for some of the actions.
-
-Better bounds can be obtained through tighter estimates on the
+\label{crl:bound-blnk-hoeffding}
+\end{crl}
+Estimate $\hat\Lambda_i^b$ can be viewed as a relaxation of the myopic VOI
+estimate to longer sequences of actions, such that the VOI estimate is
+positive for some of the actions. Better bounds can be obtained through tighter estimates on the
 probabilities, for example, based on the empirical Bernstein inequality
@@
 \begin{itemize}
-\item at the first step, the simple regret of selecting an action
-must be minimized;
-\item starting with the second step of a rollout, the expected reward
-of the rollout
- must be made as close as possible to the optimum reward, so that the
- value of the action at the first step of the rollout is evaluated
- correctly. Thus, starting with the second step, the cumulative regret
- must be minimized.
+\item at the first step, the value of partial
+information should be maximized to minimize the \textit{simple} regret
+of the selection at the current root node;
+\item as the goal of sampling in deeper tree nodes is estimating the
+value of a node, rather than selection, it makes sense to minimize
+the \textit{cumulative} regret of the rollouts from the second step
+onwards.
 \end{itemize}
@@
 
-The experiments compare the UCT algorithm \cite{Kocsis.uct} with modified
-versions of UCT in which the samples at the first step are selected according
-to their VOI estimates. Three blinkered VOI estimates were used for the comparison:
+The experiments compare the UCT algorithm \cite{Kocsis.uct} with
+modified versions of UCT in which the samples at the first step are
+selected according to their VOI estimates. Three blinkered VOI
+estimates were used for the comparison:
 \begin{description}
 \item[VCT:] based on Hoeffding inequality (\ref{eqn:conc-hoeffding});
-\item[ECT:] based on Hoeffding inequality with midpoint (see Appendix~\ref{app:better-hoeffding-bound});
-\item[BCT:] based on empirical Bernstein inequality (\ref{eqn:conc-empbernstein}).
+\item[ECT:] based on Hoeffding inequality with an intermediate point (see
+Appendix~\ref{app:better-hoeffding-bound});
+\item[BCT:] based on empirical Bernstein inequality
+(\ref{eqn:conc-empbernstein}).
 \end{description}
-For simplicity, the bounds on probability for unlimited number of samples were used in the estimate formulas. The estimates where computed as follows:
+For simplicity, the bounds on probability for an unlimited number
+of samples were used in the estimate formulas. The estimates where
+computed as follows:
 \begin{eqnarray}\
@@
 One search domain in which Monte-Carlo tree search based on the UCT
-sampling policy has been particularly successful is playing the Go
-game \cite{Gelly.mogo}. This series of experiments compares
+sampling policy has been particularly successful is playing the game Go
+\cite{Gelly.mogo}. This series of experiments compares
 the winning rate of UCT against VOI-aware policies (VCT, ECT, BCT)
@@
 the lower the bar, the better is the policy). Both ECT (based on the
-Hoeffding inequality with midpoint) and BCT (based on the empirical
+Hoeffding inequality with an intermediate point) and BCT (based on the empirical
 Bernstein inequality) are better than VCT. The difference between ECT
@@
 
-\section{Proof of Lemma~\ref{lemma:hoeffding-prob-bounds}}
+\section{Proof of Theorem~\ref{thm:hoeffding-prob-bounds}}
 \label{app:hoeffding-prob-bounds-proof}
@@
 \end{equation}
-where $A$ is the upper bound on the posterior sample mean $\overline X_i'$ if $i\ne \alpha$, $B$ is the lower bound on $\overline X_i'$ if $i=\alpha$.
-For $i=\alpha$, the bound can be improved by selecting a midpoint
-$B < y < \overline X_\beta$ and computing the bound as the sum of two parts:
+where $A$ is the upper bound on the posterior sample mean $\overline
+X_i'$ if $i\ne \alpha$, $B$ is the lower bound on $\overline X_i'$ if
+$i=\alpha$. For $i=\alpha$, the bound can be improved by selecting an
+intermediate point $B < y < \overline X_\beta$ and computing the bound
+as the sum of two parts: 
 \begin{itemize}

### 2011-12-12 paper/vago/paper.tex
a few more fixes

@@
 Monte-Carlo sampling, since the effect of a single sample is small,
-and the myopic VOI estimate will often be non-positive, resulting in premature
+and the myopic VOI estimate will often be zero, resulting in premature
 termination of the search. However, $\Lambda_i$ can be estimated as the intrinsic 
@@
 \item the arm $\alpha$ with the highest sample mean is pulled, and the 
-true mean of the arm is lower than the sample mean of the second-best arm $\beta$;
-\item another arm is pulled, and the true mean of the arm is higher
+actual arm value is lower than the sample mean of the second-best arm $\beta$;
+\item another arm is pulled, and the actual arm value is higher
 than the sample mean of the best arm $\alpha$.
@@
 
-\begin{thm} Denote the current number of samples of the $i$th arm as
+\begin{thm} Denote the current number of samples of the $i$th arm by
   $n_i$. The blinkered estimate $\Lambda_i^b$ of intrinsic value of
@@
 \end{crl}
-Estimate $\hat\Lambda_i^b$ can be viewed as a relaxation of the myopic VOI
-estimate to longer sequences of actions, such that the VOI estimate is
-positive for some of the actions. Better bounds can be obtained through tighter estimates on the
-probabilities, for example, based on the empirical Bernstein inequality
+Bound~\ref{eqn:bound-blnk-hoeffding} is generally tighter when $N$ is
+small relative to $n_i$. Estimate $\hat\Lambda_i^b$ can be viewed as a
+relaxation of the myopic assumption such that sequences consisting
+of more than one action are taken into considertion. Better bounds can
+be obtained through tighter estimates on the probabilities, for
+example, based on the empirical Bernstein inequality
 (\ref{eqn:conc-empbernstein}) or through a more careful application of
@@
 current state is discarded after an action is selected. In practice,
-most successful Monte-Carlo tree search algorithms re-use rollouts
+many successful Monte-Carlo tree search algorithms re-use rollouts
 originating from an earlier search state and passing through the
@@
 actually UCB1. In the problem of pure exploration in Multi-armed bandits
-\cite{Bubeck.pure}, the smaller is the exploration factor, the higher
-is the expected simple regret; indeed, UCT exhibits the lowest simple
+\cite{Bubeck.pure}, the smaller the exploration factor, the higher
+the expected simple regret is; indeed, UCT exhibits the lowest simple
 regret in the experiments for the default exploration factor $Cp=\frac

### 2011-12-12 paper/vago/paper.tex
in progress

@@
 $i=\alpha$. For $i=\alpha$, the bound can be improved by selecting an
-intermediate point $B < y < \overline X_\beta$ and computing the bound
+intermediate point $y$, $B < y < \overline X_\beta$, and computing the bound
 as the sum of two parts: 

### 2011-12-18 paper/vago/paper.tex
changes of eyal, but safek

@@
 \documentclass{article}
+\usepackage{enumerate}
 \usepackage{algpseudocode}
@@
 
-To estimate the VOI of pulling an arm, either a certain distribution
-of the rewards should be assumed (and updated based on observed
-rewards), or a distribution-independent bound on the VOI can be used
-as the VOI estimate. In this paper, we use {\em concentration
-  inequalities} to derive distribution-independent bounds on the VOI.
+In order to estimate the VOI of pulling an arm, a sampling model is needed.
+The generic sampling model used is: the actual value of an arm has a certain distribution.
+Pulling an arm is an i.i.d. noisy measurement of the actual value of the arm.
+In \cite{HayRussell.MCTS}, Section 4, two instances of this model are
+examined:
+\begin{enumerate}[a)]
+\item uniformly distributed prior of the value, with Bernouli
+  sampling, and 
+\item normally distributed prior of the value, with normally
+  distributed sampling.
+\end{enumerate}
+
+Here we examine the same general setting (i.i.d. samples), but instead
+of using a specific distribution model, use {\em concentration
+  inequalities} to derive distribution-independent bounds on the
+VOI. Our bounds are applicable to any distribution that conforms to
+our assumptions, including the case of normally distributed prior of
+the value, with normally distributed sampling.
 

### 2011-12-18 paper/vago/paper.tex
softer formulation of relation to hay and russell

@@
 
-In order to estimate the VOI of pulling an arm, a sampling model is needed.
-The generic sampling model used is: the actual value of an arm has a certain distribution.
-Pulling an arm is an i.i.d. noisy measurement of the actual value of the arm.
+In order to estimate the VOI of pulling an arm, a sampling model is
+needed.  The generic sampling model used is: 
+\begin{itemize}
+\item The actual value of an arm has a certain distribution.
+\item Pulling an arm is an i.i.d. noisy measurement of the actual
+  value of the arm.
+\end{itemize}
 In \cite{HayRussell.MCTS}, Section 4, two instances of this model are
@@
 \end{enumerate}
-
 Here we examine the same general setting (i.i.d. samples), but instead
@@
 inequalities} to derive distribution-independent bounds on the
-VOI. Our bounds are applicable to any distribution that conforms to
-our assumptions, including the case of normally distributed prior of
-the value, with normally distributed sampling.
+VOI. Our bounds are applicable to any bounded distribution. Similar
+bounds can be derived for certain unbounded distributions, such as the
+normally distributed prior of the value with normally disributed
+sampling, a subject of future work.
 

### 2011-12-28 paper/vago/paper.tex
some refinements

@@
 \end{itemize}
-While the UCT engine is not the most powerful engine of Pachi, it is still
-a strong player. On the other hand, additional
-features of more advanced engines would obstruct the Monte-Carlo
-sampling phenomena which are the subject of the experiment.
+While the UCT engine is not the most powerful engine of Pachi, it is
+still a strong player. On the other hand, additional features of more
+advanced engines would obstruct the Monte-Carlo sampling phenomena
+which are the subject of the experiment. It is worth noting that both
+the original UCT policy and the VOI-aware sampling policies may stop
+the sampling prior to reaching the maximum number of samples in the
+given node based on a policy-specific stopping condition, and the
+unused samples become abailable at the next step of the algorithm.
 
@@
 15000 samples (game simulations) per ply, each experiment was repeated
-1000 times. Figures~\ref{fig:uct-against-vct}--\ref{fig:uct-against-bct} show the
-winning rate of UCT against VCT, ECT, BCT, correspondingly, vs. the
-stopping threshold (if the maximum VOI of a sample is below the
-threshold, the simulation is stopped, and a move is chosen). Each curve
-in the figures corresponds to a certain number of samples per ply. The
-lower the curve, the better is the corresponding policy compared to UCT.
+1000 times. Figures~\ref{fig:uct-against-vct}--\ref{fig:uct-against-bct}
+show the winning rate of UCT against VCT, ECT, BCT, correspondingly,
+vs. the stopping threshold (if the maximum VOI of a sample is below
+the threshold, the simulation is stopped, and a move is chosen). Each
+curve in the figures corresponds to a certain number of samples per
+ply. The lower the curve, the better is the corresponding policy
+compared to UCT.
 
-As the results show, without sample redistribution UCT is as good as or
-better than the VOI-aware policies for lower numbers of samples per ply
-(5000, 7000), and the advantage of VOI-aware policies for larger
-number of samples is rather modest (e.g., ECT wins in 55\% of games for
-10000 samples per ply). For the stopping threshold of $10^{-6}$,
-however, the VOI-aware policies are almost always better than UCT
-(only VCT for 5000 samples wins in less than 50\% of games), with
-stronger policies (ECT, VCT) reaching the winning rate of 66\%.
+As the results show, without VOI-aware sample redistribution the
+VOI-aware policies are equal or inferior to UCT for lower numbers of
+samples per ply (5000, 7000), and the advantage of the VOI-aware policies
+for larger number of samples is rather modest (e.g., ECT wins in 55\%
+of games for 10000 samples per ply). For the stopping threshold of
+$10^{-6}$, however, the VOI-aware policies are almost always better
+than UCT (only VCT for 5000 samples wins in less than 50\% of games),
+with stronger policies (ECT, VCT) reaching the winning rate of 66\%.
 
 In agreement with intuition
-(Section~\ref{sec:control-redistribution}), VOI-based stopping and sample
-redistribution is most influential for medium numbers of samples per
-ply. When the maximum number of samples is too low, early stopping would
-result in poorly selected moves. On the other hand, when the maximum
-number of samples is sufficiently high, the VOI of increasing the
-maximum number of samples in a future state is low.
+(Section~\ref{sec:control-redistribution}), VOI-based stopping and
+sample redistribution is most influential for medium numbers of
+samples per ply. When the maximum number of samples is too low, early
+stopping would result in poorly selected moves. On the other hand,
+when the maximum number of samples is sufficiently high, the VOI of
+increasing the maximum number of samples in a future state is low.
 

### 2011-12-28 paper/vago/paper.tex
refinement to the paper

--- 2011-12-18 paper/vago/paper.tex
@@
 \item The time-allocation mode based on the fixed number of samples
-  was modified such that 
+  was modified for both the original UCT policy and the VOI-aware
+  policies such that 
   \begin{enumerate}
@@
 \end{itemize}
-While the UCT engine is not the most powerful engine of Pachi, it is still
-a strong player. On the other hand, additional
-features of more advanced engines would obstruct the Monte-Carlo
-sampling phenomena which are the subject of the experiment.
+While the UCT engine is not the most powerful engine of Pachi, it is
+still a strong player. On the other hand, additional features of more
+advanced engines would obstruct the Monte-Carlo sampling phenomena
+which are the subject of the experiment.
 
@@
 15000 samples (game simulations) per ply, each experiment was repeated
-1000 times. Figures~\ref{fig:uct-against-vct}--\ref{fig:uct-against-bct} show the
-winning rate of UCT against VCT, ECT, BCT, correspondingly, vs. the
-stopping threshold (if the maximum VOI of a sample is below the
-threshold, the simulation is stopped, and a move is chosen). Each curve
-in the figures corresponds to a certain number of samples per ply. The
-lower the curve, the better is the corresponding policy compared to UCT.
+1000 times. Figures~\ref{fig:uct-against-vct}--\ref{fig:uct-against-bct}
+show the winning rate of UCT against VCT, ECT, BCT, correspondingly,
+vs. the stopping threshold (if the maximum VOI of a sample is below
+the threshold, the simulation is stopped, and a move is chosen). Each
+curve in the figures corresponds to a certain number of samples per
+ply. The lower the curve, the better is the corresponding policy
+compared to UCT.
 
-As the results show, without sample redistribution UCT is as good as or
-better than the VOI-aware policies for lower numbers of samples per ply
-(5000, 7000), and the advantage of VOI-aware policies for larger
-number of samples is rather modest (e.g., ECT wins in 55\% of games for
-10000 samples per ply). For the stopping threshold of $10^{-6}$,
-however, the VOI-aware policies are almost always better than UCT
-(only VCT for 5000 samples wins in less than 50\% of games), with
-stronger policies (ECT, VCT) reaching the winning rate of 66\%.
+As the results show, without VOI-aware sample redistribution the
+VOI-aware policies are at best equal to UCT for lower numbers of
+samples per ply (5000, 7000), and the advantage of the VOI-aware policies
+for larger number of samples is rather modest (e.g., ECT wins in 55\%
+of games for 10000 samples per ply). For the stopping threshold of
+$10^{-6}$, however, the VOI-aware policies are almost always better
+than UCT (only VCT for 5000 samples wins in less than 50\% of games),
+with stronger policies (ECT, VCT) reaching the winning rate of 66\%.
 
 In agreement with intuition
-(Section~\ref{sec:control-redistribution}), VOI-based stopping and sample
-redistribution is most influential for medium numbers of samples per
-ply. When the maximum number of samples is too low, early stopping would
-result in poorly selected moves. On the other hand, when the maximum
-number of samples is sufficiently high, the VOI of increasing the
-maximum number of samples in a future state is low.
+(Section~\ref{sec:control-redistribution}), VOI-based stopping and
+sample redistribution is most influential for medium numbers of
+samples per ply. When the maximum number of samples is too low, early
+stopping would result in poorly selected moves. On the other hand,
+when the maximum number of samples is sufficiently high, the VOI of
+increasing the maximum number of samples in a future state is low.
 

### 2012-01-11 paper/rmcts/paper.tex
paper in progress

@@
 
-\title{Doing Better Than UCT: \\ Rational Monte Carlo Sampling in Trees}
+\title{Rational Monte Carlo Sampling in Trees}
 \author {David Tolpin, Solomon Eyal Shimony \\
@@
 
-The ultimate goal of this ongoing work is ``optimal'' sampling
-in the meta-reasoning sense. Nevertheless, this task is daunting for
-the following reasons:
+We begin with introduction of several sampling schemes with
+better bounds for the simple regret on sets (Section~\ref{sec:sampling-on-sets}). We then
+extend the resutls to sampling in trees by combining the proposed sampling
+schemes on the first step of a rollout with UCT during the rest of the
+rollout (Section~\ref{sec:sampling-in-trees}). Finally, we empirically
+evaluate the performance of the proposed sampling schemes on sets of
+Bernoulli arms (Section~\ref{sec:emp-mab}), in randomly generated
+trees (Section~\ref{sec:emp-mcts}), and on the sailing domain
+(Section~\ref{sec:emp-sailing}). Finally, we describe future research
+directions and provide preliminary results for a sampling scheme
+explicitly based on the value of information of a rollout (Section~\ref{sec:summary}).
+
+\section{Related Work}
+\label{sec:related-work}
+
+Efficient algorithms for Multi-Armed Bandits based on
+distribution-independent bounds, in particular UCB1, are introduced in
+\cite{Auer.ucb}. The UCT algorithm, an extension of UCB1 to
+Monte-Carlo Tree Search is described in \cite{Kocsis.uct}. Pure
+exploration in Multi-armed bandits is explored in
+\cite{Bubeck.pure}. On the one hand, the paper proves certain upper
+and lower bounds for UCB1 and uniform sampling, showing that an upper
+bound on the simple regret is exponential in the number of samples for
+uniform sampling, while only polynomial for UCB1. On the other hand,
+empirical performance of UCB1 appears to be better than of uniform
+sampling.
+
+The principles of bounded rationality appear in
+\cite{Horvitz.reasoningabout}. \cite{Russell.right} provided a formal
+description of rational metareasoning and case studies of applications
+in several problem domains. One obstacle to straightforward
+application of the principles of rational metareasoning to Monte-Carlo
+sampling is the metagreedy assumption, according to which samples must
+be selected as though at most one sample can be taken before an action
+is chosen. In Monte-Carlo sampling, the value of information of any
+single sample in a given search state is often zero, so a different
+approximating assumption must be used instead.
+
+\section{Main Results}
+\label{sec:results}
+
+The ultimate goal is ``optimal'' sampling in the meta-reasoning
+sense. Nevertheless, this task is daunting for the following reasons:
 \begin{itemize}
-\item Defining the cost of a sample is not easy, and even if we simply use time-cost
-as an approximation, we get an intractable meta-reasoning problem.
-\item Applying the standard myopic and subtree independence assumptions, we run into
-serious problems. Even in the standard selection problem \cite{TolpinShimony.blinkered}, we get a non-concave
-utility function and premature stopping of the algorithm. This is due to the fact
-that the value of information of a single measurement (analogous
-to a sample in MCTS) is frequently less than its
-time-cost, even though this is not true for multiple measurements.
-When applying the selection problem to MCTS, the situation is exacerbated. 
-The utility of an action is usually bounded, and thus in many cases a single sample
-may be insufficient to change the current best action, {\em regardless} of its
-outcome. As a result, we frequently
-get a {\em zero} ``myopic'' value of information for a single sample.
+\item Defining the cost of a sample is not easy, and even if we simply
+  use time-cost as an approximation, we get an intractable
+  meta-reasoning problem.
+\item Applying the standard myopic and subtree independence
+  assumptions, we run into serious problems. Even in the standard
+  selection problem \cite{TolpinShimony.blinkered}, we get a
+  non-concave utility function and premature stopping of the
+  algorithm. This is due to the fact that the value of information of
+  a single measurement (analogous to a sample in MCTS) is frequently
+  less than its time-cost, even though this is not true for multiple
+  measurements.  When applying the selection problem to MCTS, the
+  situation is exacerbated.  The utility of an action is usually
+  bounded, and thus in many cases a single sample may be insufficient
+  to change the current best action, {\em regardless} of its
+  outcome. As a result, we frequently get a {\em zero} ``myopic''
+  value of information for a single sample.
 \end{itemize}
@@
 As the ultimate task is extremely difficult
-to achieve, and even harder to analyze, we begin with
-simple schemes more amenable to analysis. We then
-develop a crude value of information (VOI)
-based scheme. Although the latter is a somewhat ad-hoc attempt to
-estimate value of information of a sample, it already appears (empirically) to do
-better than UCT as well as better than all the simpler schemes we propose.
+to achieve, and even harder to analyze, we introduce
+simple schemes more amenable to analysis, and compare
+them to UCB (on sets) and UCT (in trees).
 
-\section{Main Results}
-
-\subsection{Doing better than UCB}
+\subsection{Sampling on Sets}
+\label{sec:sampling-on-sets}
 
 \begin{enumerate}
-\item The $\varepsilon$-greedy allocation scheme has exponential convergence
-rate of the simple regret $\IE r$ (see Appendix~\ref{app:derivations-finite-eps}):
+\item The $\varepsilon$-greedy allocation scheme has exponential convergencerate of the simple regret $\IE r$ (see Appendix~\ref{app:derivations-finite-eps}):
 \begin{equation}
@@
 \end{equation}
-\item A tighter upper regret bound for the $\varepsilon$-greedy allocation scheme depends on
-$\varepsilon$. The lowest bound is
-achieved for $\varepsilon=\frac 1
-2$, and for a large number of arms the bound for the $\frac 1 2$-greedy scheme approaches the square of the
-bound for uniform sampling (see
-Appendix~\ref{app:derivations-asym-eps}).
+\item An asymptotic upper bound on the simple regret for the
+$\varepsilon$-greedy allocation scheme non-monotonically depends on
+$\varepsilon$. The lowest bound is achieved for $\varepsilon=\frac 1
+2$, and for a large number of arms the bound for the $\frac 1
+2$-greedy scheme approaches the square of the bound for uniform
+sampling (see Appendix~\ref{app:derivations-asym-eps}).
 
@@
 
-\subsection{Doing better than UCT}
+\subsection{Sampling in Trees}
+\label{sec:sampling-in-trees}
 
@@
 \section{Empirical Evaluation}
+\label{sec:emp}
 
 The results were empirically verified on Multi-armed Bandit instances
-(Section~\ref{seq:emp-mab}), on search trees
-(Section~\ref{seq:emp-mcts}), and on the sailing domain
-(Section~\ref{seq:emp-sailing}), as defined in \cite{Kocsis.uct}. The
+(Section~\ref{sec:emp-mab}), on search trees
+(Section~\ref{sec:emp-mcts}), and on the sailing domain
+(Section~\ref{sec:emp-sailing}), as defined in \cite{Kocsis.uct}. The
 experiments confirmed the hypotheses. The algorithms used in the
@@
 \subsection{Simple regret in multi-armed bandits}
-\label{seq:emp-mab}
+\label{sec:emp-mab}
 
-Figure~\ref{fig:mab-simple-regret} presents a comparison of simple
-regret minimization algorithms for Multi-armed
-bandits. Figure~\ref{fig:mab-simple-regret}.a shows the search tree
+Figure~\ref{fig:mab-simple-regret} presents a comparison of MCTS sampling
+schemes on Multi-armed bandits. Figure~\ref{fig:mab-simple-regret}.a shows the search tree
 corresponding to a problem instance. Each arm returns a random reward
@@
 and compares the expected reward, unknown to the algorithm during the
-sampling, to the expected reward of the best arm.
+sampling, to the expected reward of the best arm. For such problem
+instances, the MCTS sampling schemes coincide with the selection
+algorithms employed at the first step of a rollout.
 
@@
 
-For smaller numbers of samples, $\frac 1 2$-greedy achieves the best
+For smaller numbers of samples, GCT achieves the best
 performance; for large number of samples, QCT outperforms GCT. QCT is
@@
 \subsection{Monte Carlo tree search}
-\label{seq:emp-mcts}
+\label{sec:emp-mcts}
 
@@
 \subsection{The sailing domain}
-\label{seq:emp-sailing}
+\label{sec:emp-sailing}
 
@@
 \section{Summary and Future Work}
+\label{sec:summary}
 
@@
 realization of the latter approach the VOI of performing an action is
-estimated as \emph{the value of perfect information learned from the outcome
-of earlier samples of the action divided by the number of
+estimated as \emph{the value of perfect information learned from the
+  outcome of earlier samples of the action divided by the number of
   samples}. Early experiments with this approach showed the best
@@
 improvement for mixed arms (e.g. Bernoulli, triangularly distributed,
-and fixed arms) (Figure~\ref{fig:flat-trilevel-vct}):
-\textbf{VCT}, the algorithm which selects an action with the maximum VOI estimate on
-the first step exhibits lower average regret than other algorithms.
-Sample-based VOI estimation requires keeping more information than an
-algorithm based only on the sample
-mean, but the overhead is small and justified by the improvement.
+and fixed arms) (Figure~\ref{fig:flat-trilevel-vct} shows results for
+different algorithms for 32 randomly generated arms): \textbf{VCT},
+the algorithm which selects a rollout with the maximum VOI estimate
+exhibits lower average regret than other algorithms. Sample-based VOI
+estimation requires keeping more information than an algorithm based
+only on the sample mean, but the overhead is small and justified by
+the improvement.
 

### 2012-01-11 paper/rmcts/paper.tex
paper in progress

@@
 algorithm that minimizes the simple regret, such as $\frac 1 2$-greedy or
-UQB, and then selects actions according to UCB. Such two-stage
+UQB, and then selects actions according to UCB. This scheme
 outperforms the UCT sampling scheme that selects actions based on
@@
 
-Such sampling scheme is significantly less sensitive to the tuning of
-the exploration factor ($C_p$ in \cite{Kocsis.uct}) of UCT, since the
-conflict \cite{Bubeck.pure} between the need for a larger value of
-$C_p$ on the first step (simple regret) and a smaller value for the
-rest of the rollout (cumulative regret) is resolved. In fact, a
-sampling scheme that uses UCB at all steps but a larger value of $C_p$
-for the first step than for the rest of the steps, outperforms
-UCT. The pseudocode of the two-stage rollout is in
-Algorithm~\ref{alg:two-stage-mcts}.
+Additionaly, such two-stage sampling scheme is significantly less
+sensitive to the tuning of the exploration factor ($C_p$ in
+\cite{Kocsis.uct}) of UCT, since the conflict \cite{Bubeck.pure}
+between the need for a larger value of $C_p$ on the first step (simple
+regret) and a smaller value for the rest of the rollout (cumulative
+regret) is resolved. In fact, a sampling scheme that uses UCB at all
+steps but a larger value of $C_p$ for the first step than for the rest
+of the steps, outperforms UCT. The pseudocode of the two-stage rollout
+is in Algorithm~\ref{alg:two-stage-mcts}.
 

### 2012-01-11 paper/rmcts/paper.tex
one more small change

@@
 \item An asymptotic upper bound on the simple regret for the
-$\varepsilon$-greedy allocation scheme non-monotonically depends on
+$\varepsilon$-greedy allocation scheme depends \emph{non-monotonically} on
 $\varepsilon$. The lowest bound is achieved for $\varepsilon=\frac 1
@@
   \underline{T}rees \cite{Kocsis.uct};
+\item[RCT:] U\underline{CT} with the first step replaced with uniform
+  \underline{r}andom sampling;
 \item[GCT:] U\underline{CT} with the first step

### 2012-01-12 paper/rmcts/aaai/paper.tex
aaai, first draft

--- 2012-01-11 paper/rmcts/paper.tex
@@
-\documentclass{article}
+%File: formatting-instruction.tex
+\documentclass[letterpaper]{article}
+\usepackage{aaai}
+\usepackage{times}
+\usepackage{helvet}
+\usepackage{courier}
 \usepackage{algpseudocode}
@@
 \usepackage{color}
-\usepackage{geometry}
 
-\geometry{margin=1.2in}
+\newtheorem{lemma}{Lemma}
+\frenchspacing
+\pdfinfo{
+/Title (Rational Monte Carlo Sampling in Trees)
+/Subject (AAAI Publications)
+/Author (AAAI Press)}
 
+\setcounter{secnumdepth}{0}  
 \newcommand {\mean} {\ensuremath {\mathop{\mathrm{mean}}}}
@@
 \newcommand {\BEL} {\ensuremath {\mathop{\mathrm{BEL}}}}
-
-\newtheorem{lemma}{Lemma}
 
@@
 We begin with introduction of several sampling schemes with
-better bounds for the simple regret on sets (Section~\ref{sec:sampling-on-sets}). We then
+better bounds for the simple regret on sets. We then
 extend the resutls to sampling in trees by combining the proposed sampling
 schemes on the first step of a rollout with UCT during the rest of the
-rollout (Section~\ref{sec:sampling-in-trees}). Finally, we empirically
+rollout. Finally, we empirically
 evaluate the performance of the proposed sampling schemes on sets of
-Bernoulli arms (Section~\ref{sec:emp-mab}), in randomly generated
-trees (Section~\ref{sec:emp-mcts}), and on the sailing domain
-(Section~\ref{sec:emp-sailing}). Finally, we describe future research
+Bernoulli arms, in randomly generated
+trees, and on the sailing domain
+. Finally, we describe future research
 directions and provide preliminary results for a sampling scheme
-explicitly based on the value of information of a rollout (Section~\ref{sec:summary}).
+explicitly based on the value of information of a rollout.
 
@@
     \If {depth=1}
-      \State action $\gets$ \Call{FirstAction}{node} \Comment minimizes simple regret, e.g. $\frac 1 2$-greedy
+      \State action $\gets$ \Call{FirstAction}{node}
     \Else
-      \State action $\gets$ \Call{NextAction}{node} \Comment  UCB --- minimizes cumulative regret
+      \State action $\gets$ \Call{NextAction}{node}
     \EndIf
     \State next-node $\gets$ \Call{NextState}{node, action}
-    \State reward $\gets$ \Call{TransitionReward}{node, action, next-node}
-    \State \hspace{4em} + \Call{Rollout}{next-node, depth+1} \Comment call {\sc Rollout} recursively
+    \State reward $\gets$ \Call{Reward}{node, action, next-node}
+     \State \hspace{4em} + \Call{Rollout}{next-node, depth+1}
     \State \Call{UpdateStats}{node, action, reward}
@@
 
-The results were empirically verified on Multi-armed Bandit instances
-(Section~\ref{sec:emp-mab}), on search trees
-(Section~\ref{sec:emp-mcts}), and on the sailing domain
-(Section~\ref{sec:emp-sailing}), as defined in \cite{Kocsis.uct}. The
+The results were empirically verified on Multi-armed Bandit instances,
+on search trees, and on the sailing domain, as defined in \cite{Kocsis.uct}. The
 experiments confirmed the hypotheses. The algorithms used in the
@@
 \begin{figure}[t]
-  \begin{minipage}[c]{0.5\linewidth}
+  \begin{minipage}[c]{1.0\linewidth}
     \centering
     \includegraphics[scale=1.0]{onelevel-tree.pdf}\\
-    \vspace{4em}
     a. search tree
+    \vspace{1em}
   \end{minipage}
-  \begin{minipage}[c]{0.5\linewidth}
+  \begin{minipage}[c]{1.0\linewidth}
     \centering
@@
 
-\begin{figure}
-  \begin{minipage}[c]{0.5\linewidth}
+\begin{figure}[h!]
+  \begin{minipage}[c]{1.0\linewidth}
     \centering
@@
     a. search tree
+    \vspace{1em}
   \end{minipage}
-  \begin{minipage}[c]{0.5\linewidth}
+  \begin{minipage}[c]{1.0\linewidth}
     \centering
     \includegraphics[scale=0.4]{tree-identity-k=16-uqb=8.pdf}\\ 
-    b. 16 arms \\
+    b. 16 arms
     \vspace{1em}
+  \end{minipage}
+  \begin{minipage}[c]{1.0\linewidth}
+    \centering
     \includegraphics[scale=0.4]{tree-identity-k=64-uqb=8.pdf} \\
@@
 \begin{figure}[t]
-  \begin{minipage}[b]{0.5\linewidth}
+  \begin{minipage}[b]{1.0\linewidth}
     \centering
@@
     a. median cost
+    \vspace{1em}
   \end{minipage}
-  \begin{minipage}[b]{0.5\linewidth}
+  \begin{minipage}[b]{1.0\linewidth}
     \centering
@@
 \begin{figure}[t]
-  \begin{minipage}[b]{0.5\linewidth}
   \centering
-    \includegraphics[scale=0.45]{costs-size=6-nsamples=199.pdf}\\
-    a. 199 rollouts\\
+  \includegraphics[scale=0.45]{costs-size=6-nsamples=397.pdf}\\
+  a. 397 rollouts\\
   \vspace{1em}
-    \includegraphics[scale=0.45]{costs-size=6-nsamples=397.pdf}\\
-    b. 397 rollouts\\
-  \end{minipage}
-  \begin{minipage}[b]{0.5\linewidth}
-    \centering
   \includegraphics[scale=0.45]{costs-size=6-nsamples=793.pdf}\\
-    c. 793 rollouts\\
+  b. 793 rollouts\\
   \vspace{1em}
   \includegraphics[scale=0.45]{costs-size=6-nsamples=1585.pdf}\\
-    d. 1585 rollouts\\
-  \end{minipage}
+  c. 1585 rollouts
   \caption{The sailing domain, $6\times 6$ lake, cost vs. factor}
@@
 \begin{figure}[t]
-  \begin{minipage}[b]{0.333\linewidth}
+  \begin{minipage}[b]{1.0\linewidth}
     \centering
-    \includegraphics[scale=0.35]{costs-size=3-nsamples=397.pdf}\\
+    \includegraphics[scale=0.45]{costs-size=3-nsamples=397.pdf}\\
     a. $3\times 3$ lake
+    \vspace{1em}
   \end{minipage}
-  \begin{minipage}[b]{0.333\linewidth}
+  \begin{minipage}[b]{1.0\linewidth}
     \centering
-    \includegraphics[scale=0.35]{costs-size=6-nsamples=397.pdf}\\
+    \includegraphics[scale=0.45]{costs-size=6-nsamples=397.pdf}\\
     b. $6\times 6$ lake
+    \vspace{1em}
   \end{minipage}
-  \begin{minipage}[b]{0.333\linewidth}
+  \begin{minipage}[b]{1.0\linewidth}
     \centering
-    \includegraphics[scale=0.35]{costs-size=10-nsamples=397.pdf}\\
-    b. $10\times 10$ lake
+    \includegraphics[scale=0.45]{costs-size=10-nsamples=397.pdf}\\
+    c. $10\times 10$ lake
   \end{minipage}
@@
 \begin{figure}[t]
-  \begin{minipage}[b]{0.5\linewidth}
   \centering
-    \includegraphics[scale=0.45]{rcq-size=6-nsamples=199.pdf}\\
-    a. 199 rollouts\\
-    \vspace{1em}
   \includegraphics[scale=0.45]{rcq-size=6-nsamples=397.pdf}\\
   b. 397 rollouts\\
-  \end{minipage}
-  \begin{minipage}[b]{0.5\linewidth}
-    \centering
+  \vspace{1em}
   \includegraphics[scale=0.45]{rcq-size=6-nsamples=793.pdf}\\
@@
   d. 1585 rollouts\\
-  \end{minipage}
   \caption{The sailing domain, log vs. sqrt, $6\times 6$ lake}
@@
   \centering
-  \includegraphics[scale=0.45,trim=0 0 0 16pt,clip]{flat-trilevel-vct-k=64.pdf}\\
+  \includegraphics[scale=0.35,trim=0 0 0 16pt,clip]{flat-trilevel-vct-k=64.pdf}\\
   \caption{VOI-based action selection (VCT) vs. other algorithms}
@@
 
-\clearpage
-\appendix
+\section{Appendixes}
 
-\section{Facts}
+\subsection{Facts}
 
@@
 
-\section{Derivations}
+\subsection{Derivations}
 \label{app:derivations}
 
-\subsection{Finite-time regret bounds}
+\subsubsection{Finite-time regret bounds}
 \label{app:derivations-finite}
@@
 
-\subsubsection{$\varepsilon$-greedy}
+\paragraph{$\varepsilon$-greedy}
 \label{app:derivations-finite-eps}
@@
 \begin{eqnarray}
-\Pr(\overline X_j=\max_i\overline X_i)&\le&2\left(x_0\cdot \Pr\left\{T_j^R(n)\le x_0\right\} + \frac 2{\Delta_j^2}e^{-\Delta_j^2\lfloor x_0 \rfloor/2}\right)\nonumber\\
-&\le&2\left(x_0\cdot \Pr\left\{T_j^R(n)\le x_0\right\} + \frac {2 \cdot
-  e^{\Delta_j^2/2}}{\Delta_j^2}e^{-\Delta_j^2 x_0 /2}\right)
+\Pr(\overline X_j&=&\max_i\overline X_i)\nonumber\\
+   &\le&2\left(x_0\cdot \Pr\left\{T_j^R(n)\le x_0\right\} + \frac 2{\Delta_j^2}e^{-\Delta_j^2\lfloor x_0 \rfloor/2}\right)\nonumber\\
+   &\le&2\left(x_0\cdot \Pr\left\{T_j^R(n)\le x_0\right\} + \frac {2
+  \cdot  e^{\Delta_j^2/2}}{\Delta_j^2}e^{-\Delta_j^2 x_0
+  /2}\right)\nonumber\\
+   &&
 \end{eqnarray}
@@
 
-\subsubsection{UCB($\alpha$)}
+\paragraph{UCB($\alpha$)}
 \label{app:derivations-finite-ucb}
@@
 
-\subsection{Asymptotic regret analysis}
+\subsubsection{Asymptotic regret analysis}
 \label{app:derivations-asym}
@@
 
-\subsubsection{$\varepsilon$-greedy}
+\paragraph{$\varepsilon$-greedy}
 \label{app:derivations-asym-eps}
@@
 \varepsilon&=&\frac 1 2\nonumber\\
-\IE
-r_{\varepsilon=1/2}&\le&B_{1/2\mbox{-}greedy}=2\exp\left(-\frac
-  {\Delta_j^2n} {(\sqrt{K-1}+1)^2}\right)\nonumber\\
+\IE r_{\varepsilon=1/2}&\le&B_{1/2\mbox{-}greedy}\nonumber\\
+                     &=&2\exp\left(-\frac{\Delta_j^2n} {(\sqrt{K-1}+1)^2}\right)\nonumber\\
 \lim_{K\to\infty}
@@
 
-\subsubsection{UCB($\alpha$)}
+\paragraph{UCB($\alpha$)}
 \label{app:derivations-asym-ucb}
@@
 
-\bibliographystyle{plain}
+\bibliographystyle{aaai}
 \bibliography{refs}

### 2012-01-19 paper/rmcts/aaai/formatting-instructions-latex.tex

%File: formatting-instruction.tex
\documentclass[letterpaper]{article}
\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\frenchspacing
\pdfinfo{
/Title (Formatting Instructions for Authors Using LaTeX)
/Subject (AAAI Publications)
/Author (AAAI Press)}
\setcounter{secnumdepth}{0}  
 \begin{document}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\title{Formatting Instructions for Authors \\Using \LaTeX{}}
\author{AAAI Press\\
Association for the Advancement of Artificial Intelligence\\
445 Burgess Drive\\
Menlo Park, California 94025\\
}
\maketitle
\begin{abstract}
\begin{quote}
AAAI creates proceedings, working notes, and technical reports directly from electronic source furnished by the authors. To ensure that all papers in the publication have a uniform appearance, authors must adhere to the following 
instructions. 
\end{quote}
\end{abstract}

\noindent Congratulations on having a paper selected for inclusion in an AAAI Press proceedings or technical report! This document details the requirements necessary to get your accepted paper published using \LaTeX{}. If you are using Microsoft Word, instructions are provided in a different document. If you want to use some other formatting software, you must obtain permission from AAAI Press first.

The instructions herein are provided as a general guide for experienced \LaTeX{} users who would like to use that software to format their paper for an AAAI Press publication or report. If you are not an experienced \LaTeX{} user, do not use it to format your paper. AAAI cannot provide you with support and the accompanying style files are \textbf{not} guaranteed to work. If the results you obtain are not in accordance with the specifications you received, you must correct your source file to achieve the correct result.

These instructions are generic. Consequently, they do not include specific dates, page charges, and so forth. Please consult your specific written conference instructions for details regarding your submission. Please review the entire document for specific instructions that might apply to your particular situation. All authors must comply with the following:

\begin{itemize}
\item You must use the latest AAAI Press \LaTeX{} macro .
\item Download the author kit.
\item Complete, sign, and return by the deadline the AAAI copyright form (proceedings authors) or distribution license (technical report authors).
\item Read and format your paper source and PDF according to the formatting instructions for authors.
\item Submit your electronic files and abstract using our electronic submission form \textbf{on time.}
\item Submit your copyright form, and any required page or formatting charges to AAAI Press so that they are received by the deadline.
\item Check every page of your paper before submitting it.
\end{itemize}

\section{Copyright}
All papers submitted for publication by AAAI Press must be accompanied by a valid signed copyright form or, in the case of technical reports, by a valid signed permission to distribute form. There are no exceptions to this requirement. You must send us the original version of this form. However, to meet the deadline, you may fax (1-650-321-4457) or scan and e-mail the form (publications11@aaai.org) to AAAI by the submission deadline, and then mail the original via postal mail to the AAAI office. If you fail to send in a signed copyright or permission form, your paper will not be published. You will find PDF versions of the AAAI copyright and permission to distribute forms in the author kit.

\section{Formatting Requirements in Brief}
We need source and PDF files that can be used in a variety of ways and can be output on a variety of devices. AAAI imposes some requirements on your source and PDF files that must be followed. Most of these requirements are based on our efforts to standardize conference manuscript properties and layout. These requirements are as follows, and all papers submitted to AAAI for publication must comply:

\begin{itemize}
\item All fonts must be embedded in the PDF file.
\item Modifications to the style sheet (or your document) in an effort to avoid extra page charges are NOT allowed.
\item No type 3 fonts may be used (even in illustrations).
\item Your title must follow US capitalization rules.
\item \LaTeX{} documents must use the Times or Nimbus font package (do not use Computer Modern for the text of your paper).
\item No \LaTeX{} 209 documents may be used or submitted.
\item Fonts that require non-English language support (CID and Identity-H) must be converted to outlines or removed from the document (even if they are in a graphics file embedded in the document). 
\item Two-column format in AAAI style is required for all papers.
\item The paper size for final submission must be US letter. No exceptions.
\item The source file must exactly match the PDF.
\item The document margins must be as specified in the formatting instructions.
\item The number of pages and the file size must be as specified for your event.
\item No document may be password protected.
\item Neither the PDFs nor the source may contain any embedded links or bookmarks.
\item Your source and PDF must not have any page numbers, footers, or headers.
\item Your PDF must be compatible with Acrobat 5 or higher.
\end{itemize}

If you do not follow the above requirements, it is likely that we will be unable to publish your paper.

\section{What Files to Submit}
You must submit the following items to ensure that your paper is published:
\begin{itemize}
\item A fully-compliant PDF file.
\item Your  \LaTeX{}  source file submitted as a \textit{single} .tex file that \textit{includes} the bibliography formatted using aaai.bib, and any custom macros your paper uses (include them in the preamble). Your source must compile on our system.
\item All your graphics files.
\item The .aux and .log file for your compiled source.
\item All the style files used in your document.
\end{itemize}

Your \LaTeX{} source will be reviewed and recompiled on our system (if it does not compile, you may incur late fees).   \textbf{Do not submit your source in multiple text files.} Your single \LaTeX{} source file must include all your text, your bibliography (formatted using aaai.bst), and any custom macros. Accompanying this source file, you must also supply all your referenced style files and graphics files.

Your files should work without any supporting files (other than the program itself) on any computer. Place your PDF and source files in a single tar, zipped, gzipped, stuffed, or compressed archive. Name your source file with your last (family) name.

\textbf{Do not send files that are not actually used in the paper.} We don't want you to send us any files not needed for compiling your paper, including, for example, this instructions file, unused graphics files, and so forth.  A shell script that might help you create the \LaTeX{} source package is included in the Author Kit.

\section{Using \LaTeX{} to Format Your Paper}

The latest version of the AAAI style file is available on AAAI's website. Download this file and place it in a file named ``aaai.sty" in the \TeX\ search path. Placing it in the same directory as the paper should also work. You must download the latest version of the complete author kit so that you will have the latest instruction set.

\subsection{Document Preamble}

In the \LaTeX{} source for your paper, you \textbf{must} place the following lines as shown in the example in this subsection. This command set-up is for three authors. Add or subtract author and address lines as necessary, and uncomment the portions that apply to you. In most instances, this is all you need to do to format your paper in the Times font. The helvet package will cause Helvetica to be used for sans serif, and the courier package will cause Courier to be used for the typewriter font. These files are part of the PSNFSS2e package, which is freely available from many Internet sites (and is often part of a standard installation).

Leave the setcounter for section number depth commented out and set at 0 unless you want to add section numbers to your paper. If you do add section numbers, you must uncomment this line and change the number to 1 (for section numbers), or 2 (for section and subsection numbers). The style file will not work properly with numbering of subsubsections, so do not use a number higher than 2.

To add the required metadata for your paper, uncomment one of the two metadata sections. Metadata will be explained in greater detail in the next subsection.

\begin{quote}
\begin{small}
\textbackslash documentclass[letterpaper]{article}\\
\% \textit{Required Packages}\\
\textbackslash usepackage\{aaai\}\\
\textbackslash usepackage\{times\}\\
\textbackslash usepackage\{helvet\}\\
\textbackslash usepackage\{courier\}\\
\%\%\%\%\%\%\%\%\%\%\\
\% \textit{PDFMARK for TeX and GhostScript}\\
\% Uncomment and complete the following for \\
\% metadata if your paper is typeset using TeX and \\
\% GhostScript (e.g  if you use .ps or .eps files in your paper):\\
\% \textbackslash special\{! /pdfmark where\\
\% \{pop\} \{userdict /pdfmark /cleartomark load put\} ifelse\\
\% [ /Author (John Doe, Jane Doe)\\
\% /Title (Input Your Paper Title Here)\\
\% /Subject (Input the Proceedings Title Here)\\
\% /Keywords (Input your paper's keywords here)\\
\% /DOCINFO pdfmark\\
\% \}\\
\%\%\%\%\%\%\%\%\%\%\\
\% \textit{PDFINFO for PDF\TeX{}}\\
\% Uncomment and complete the following for metadata if\\
\% your paper is typeset using PDF\TeX{}\\
\% \textbackslash pdfinfo\{\\
\% /Title (Input Your Paper Title Here)\\
\% /Author (John Doe, Jane Doe)\\
\% /Subject (Input the Proceedings Title Here)\\
\% /Keywords (Input your paper's keywords here)\\
\% \}\\
\%\%\%\%\%\%\%\%\%\%\\
\% \textit{Section Numbers}\\
\% Uncomment if you want to use section numbers\\
\% and change the 0 to a 1 or 2\\
\% \textbackslash setcounter\{secnumdepth\}\{0\}\\
\%\%\%\%\%\%\%\%\%\%\\
\% \textit{Title, Author, and Address Information}\\
\textbackslash title\{Title\}\\
\textbackslash author\{Author 1 \textbackslash and Author 2\textbackslash\textbackslash \\ 
Address line\textbackslash\textbackslash\\ Address line\textbackslash\textbackslash \\
\textbackslash And\\
Author 3\textbackslash\textbackslash\\ Address line\textbackslash\textbackslash\\ Address line\}\\
\%\%\%\%\%\%\%\%\%\%\\
\% \textit{Body of Paper Begins}\\
\textbackslash begin\{document\}\\
\textbackslash maketitle\\
...\\
\%\%\%\%\%\%\%\%\%\%\\
\% \textit{References and End of Paper}\\
\textbackslash bibliography\{Bibliography-File\}\\
\textbackslash bibliographystyle\{aaai\}\\
\textbackslash end\{document\}
\end{small}
\end{quote}

\subsection{Inserting Document Metadata with \LaTeX{}}
PDF files contain document summary information that enables us to create an Acrobat index (pdx) file, and also allows search engines to locate and present your paper more accurately. \textit{Inserting metadata is a requirement of submission.}

If your paper includes illustrations that are not compatible with PDF\TeX{} (such as .eps or .ps documents), you will be using GhostScript to create your PDF. To add metadata to your paper, you should uncomment the  \textit{PDFMARK for \TeX{} and GhostScript} section in the preamble shown above.

If you have no illustrations, or all of your illustrations are compatible with PDF\TeX{} (that is, they are PDF or another compatible format), you should use PDF\TeX{}. To add metadata to your paper, you should uncomment the \textit{PDFINFO for PDF\TeX{}} section in the preamble instead.

\textit{Important:} Do not include \textit{any} \LaTeX{} code or nonascii characters (including accented characters) in the metadata. The data in the metadata must be completely plain ascii. It may not include slashes, accents, linebreaks, unicode, or any \LaTeX{} commands. Type the title exactly as it appears on the paper (minus all formatting). Input the author names in the order in which they appear on the paper (minus all accents), separating each author by a comma. You may also include keywords in the Keywords field. If you know the full title of the proceedings, include it in the subject line. Leave any additional metadata fields blank.

\subsection{Preparing Your Paper}

After the preamble above, you should prepare your paper as follows:

\begin{quote}
\begin{small}
\textbackslash begin\{document\}\\
\textbackslash maketitle\\
...\\
\textbackslash bibliography\{Bibliography-File\}\\
\textbackslash bibliographystyle\{aaai\}\\
\textbackslash end\{document\}\\
\end{small}
\end{quote}
\subsection{Incompatible Packages}
The following packages are incompatible with aaai.sty and/or aaai.bst and must not be used (this list is not exhaustive):
\begin{itemize}
\item hyperref
\item natbib
\item geometry
\item titlesec
\item layout
\item T1 fontenc package (install the CM super fonts package instead)
\end{itemize}

\subsection{Paper Size, Margins, and Column Width}
Papers must be formatted to print in two-column format on 8.5 x 11 inch US letter-sized paper. The margins must be exactly as follows: 
\begin{itemize}
\item Top margin: .75 inches
\item Left margin: .75 inches
\item Right margin: .75 inches
\item Bottom margin: 1.25 inches
\end{itemize}

The default paper size in most installations of \LaTeX{} is A4. However, because we require that your electronic paper be formatted in US letter size, you will need to alter the default for this paper to US letter size. Assuming you are using the 2e version of \LaTeX{}, you can do this by including the [letterpaper] option at the beginning of your file: 
\textbackslash documentclass[letterpaper]{article}.

This command is usually sufficient to change the format. Sometimes, however, it may not work, especially if you use PDF\LaTeX{}, and you may need to make some alterations to your source. \textbf{Do not use the Geometry package to alter the page size.} Use of this style file alters aaai.sty and will result in your paper being rejected.

It is also usually necessary to configure dvips. Try passing the -tletter option to dvips. Those using RedHat Linux 8.0 and \LaTeX{} should also check the paper size setting in ``/usr/share/texmf/dvips/config/config.ps" --- it may be that ``A4" is the default, rather than ``letter." This can result in incorrect top and bottom margins in documents you prepare with \LaTeX{}. You may need to edit the config file to correct the problem. (Once you've edited to config file for US letter, however, don't forget to change it back --- otherwise it may not be possible for you to print your papers locally.)

\subsubsection{Column Width and Margins.}
To ensure maximum readability, your paper must include two columns. Each column should be 3.3 inches wide (slightly more than 3.25 inches), with a .375 inch (.952 cm) gutter of white space between the two columns. The aaai.sty file will automatically create these columns for you.

\subsection{Overlength Papers}
If your paper is too long, turn on \textbackslash frenchspacing, which will reduce the space after periods. Next, alter the linespacing to 11 point and/or shrink the size of your graphics. Use \textbackslash centering instead of \textbackslash begin\{center\} in your figure environment. If these two methods don't work, you may minimally use the following. For floats (tables and figures), you may minimally reduce \textbackslash floatsep, \textbackslash textfloatsep, \textbackslash abocaptionskip, and \textbackslash belowcaptionskip. For mathematical environments, you may minimally reduce \textbackslash abovedisplayskip, \textbackslash belowdisplayskip, and \textbackslash arraycolsep. You may also alter the size of your bibliography to \textbackslash small.

Commands that alter page layout are forbidden. These include \textbackslash columnsep, \textbackslash topmargin, \textbackslash topskip, \textbackslash textheight, \textbackslash textwidth, \textbackslash oddsidemargin, and \textbackslash evensizemargin. \textbf{If you alter page layout, your paper will be classified as late, and it will either be excluded from the proceedings or you will be required to pay a significant late fee.} Other commands that are questionable and may cause your paper to be rejected include  \textbackslash parindent, and \textbackslash parskip. Commands that alter the space between sections are also questionable. Regardless of the above, if your paper is obviously ``squeezed" it is not going to to be accepted. Before using every trick you know to make your paper a certain length, try cutting text instead or (if allowed) paying the extra page charge. It will be cheaper in the long run.

\subsection{Type Font and Size}
Your paper must be formatted in Times Roman or Nimbus. We will not accept papers formatted using Computer Modern as the text or heading typeface. Sans serif, when used, should be Courier. Use Symbol or Lucida or Computer Modern for \textit{mathematics only. }

Do not use type 3 fonts for any portion of your paper, including graphics. Type 3 bitmapped fonts are designed for fixed resolution printers. Most print at 300 dpi even if the printer resolution is 1200 dpi or higher. They also often cause high resolution imagesetter devices and our PDF indexing software to crash. Consequently, AAAI will not accept electronic files containing obsolete type 3 fonts. Files containing those fonts (even in graphics) will be rejected.

Fortunately, there are effective workarounds that will prevent your file from embedding type 3 bitmapped fonts. The easiest workaround is to use the required times, helvet, and courier packages with \LaTeX{}2e. (Note that papers formatted in this way will still use Computer Modern for the mathematics. To make the math look good, you'll either have to use Symbol or Lucida, or you will need to install type 1 Computer Modern fonts --- for more on these fonts, see the section ``Obtaining Type 1 Computer Modern.")

If you are unsure if your paper contains type 3 fonts, view the PDF in Acrobat Reader. The Properties/Fonts window will display the font name, font type, and encoding properties of all the fonts in the document. If you are unsure if your graphics contain type 3 fonts (and they are PostScript or encapsulated PostScript documents), create PDF versions of them, and consult the properties window in Acrobat Reader.

The default size for your type should be ten-point with eleven- or twelve-point leading (line spacing). Twelve point leading is  easier to read, however if your paper is running long, you may change the leading to eleven point. Start all pages (except the first) directly under the top margin. (See the next section for instructions on formatting the title page.) Indent ten points when beginning a new paragraph, unless the paragraph begins directly below a heading or subheading.

\subsubsection{Obtaining Type 1 Computer Modern for \LaTeX{}.}

If you use Computer Modern for the mathematics in your paper (you cannot use it for the text) you may need to download type 1 Computer fonts. They are available without charge from the American Mathematical Society:
http://www.ams.org/tex/type1-fonts.html.

\subsection{Title and Authors}
Your title must appear in mixed case (nouns, pronouns, and verbs are capitalized) near the top of the first page, centered over both columns in sixteen-point bold type (twenty-four point leading). This style is called ``mixed case." Author's names should appear below the title of the paper, centered in twelve-point type (with fifteen point leading), along with affiliation(s) and complete address(es) (including electronic mail address if available) in nine-point roman type (the twelve point leading). (If the title is long, or you have many authors, you may reduce the specified point sizes by up to two points.) You should begin the two-column format when you come to the abstract.

\subsubsection{Formatting Author Information}
Author information can be set in a number of different styles, depending on the number of authors and the number of affiliations you need to display. For several authors from the same institution, use \textbackslash and:

\begin{quote}
\begin{small}
\textbackslash author\{Author 1 \textbackslash and  ...  \textbackslash and Author \textit{n}\textbackslash \textbackslash  \\
Address line \textbackslash \textbackslash ~... \textbackslash \textbackslash ~Address line\}
\end{small}
\end{quote}

\noindent If the names do not fit well on one line use:

\begin{quote}
\begin{small}
\textbackslash author\{Author 1\}\textbackslash \textbackslash \\ \{\textbackslash bf Author 2\}\textbackslash \textbackslash ~ 
... \textbackslash \textbackslash ~\{\textbackslash bf Author \textit{n}\}\textbackslash \textbackslash \\
Address line \textbackslash \textbackslash ~ ... \textbackslash \textbackslash ~ Address line\}
\end{small}
\end{quote}

\noindent For authors from different institutions, use \textbackslash And:

\begin{quote}
\begin{small}
\textbackslash author\{Author 1\textbackslash \textbackslash ~ Address line \textbackslash \textbackslash ~...  \textbackslash \textbackslash ~ Address line
\textbackslash And  ...  
\textbackslash And
Author \textit{n}\textbackslash \textbackslash \\ Address line\textbackslash \textbackslash ~
... \textbackslash \textbackslash ~
Address line\}
\end{small}
\end{quote}

\noindent To start a separate ``row" of authors, use \textbackslash AND:
\begin{quote}
\begin{small}
\textbackslash author\{Author 1\textbackslash\textbackslash ~ Address line \textbackslash\textbackslash ~
...  \textbackslash \textbackslash ~ Address line\textbackslash\textbackslash \\
\textbackslash AND\\
Author 2 \textbackslash\textbackslash ~ Address line \textbackslash\textbackslash ~
...  \textbackslash \textbackslash ~ Address line\textbackslash\textbackslash \\
\textbackslash And\\
Author 3 \textbackslash\textbackslash ~ Address line \textbackslash\textbackslash ~
...  \textbackslash \textbackslash ~ Address line\textbackslash\textbackslash \\\}
\end{small}
\end{quote}

\noindent If the title and author information does not fit in the area allocated, place
\textbackslash setlength\textbackslash titlebox\{\textit{height}\}
after the \textbackslash documentclass line where \{\textit{height}\} is something like 2.5in.

\subsection{\LaTeX{} Copyright Notice}
The copyright notice automatically appears if you use aaai.sty. If you are creating a technical report, it is not necessary to include this notice. You may disable the copyright line using the \verb+\+nocopyrightcommand. To change the entire text of the copyright slug, use:
\textbackslash copyrighttext \{\emph{text}\}.
Either of these must appear before \textbackslash maketitle. Please be advised, however, that \textit{if you disable or change the copyright line and transfer of copyright is required, your paper will not be published.}

\subsection{Credits}
Any credits to a sponsoring agency should appear in the acknowledgments section, unless the agency requires different placement. If it is necessary to include this information on the front page, use
\textbackslash thanks in either the \textbackslash author or \textbackslash title commands.
For example:
\begin{quote}
\begin{small}
\textbackslash title\{Very Important Results in AI\textbackslash thanks\{This work is
 supported by everybody.\}\}
\end{small}
\end{quote}
Multiple \textbackslash thanks commands can be given. Each will result in a separate footnote indication in the author or title with the corresponding text at the botton of the first column of the document. Note that the \textbackslash thanks command is fragile. You will need to use \textbackslash protect.

Please do not include \textbackslash pubnote commands in your document.

\subsection{Abstract}
The abstract must be placed at the beginning of the first column, indented ten points from the left and right margins. The title “Abstract” should appear in ten-point bold type, centered above the body of the abstract. The abstract should be set in nine-point type with ten-point leading. This concise, one-paragraph summary should describe the general thesis and conclusion of your paper. A reader should be able to learn the purpose of the paper and the reason for its importance from the abstract. The abstract should be no more than two hundred words in length. (Authors who are submitting short one- or two-page extended extracts should provide a short abstract of only a sentence or so.) \textbf{Do not include references in your abstract!}

\subsection{Page Numbers}

Do not \textbf{ever} print any page numbers on your paper.

\subsection{Text }
The main body of the paper must be formatted in ten-point with eleven-point or twelve-point leading (line spacing).

\subsection{Citations}
Citations within the text should include the author's last name and year, for example (Newell 1980). Append lower-case letters to the year in cases of ambiguity. Multiple authors should be treated as follows: (Feigenbaum and Engelmore 1988) or (Ford, Hayes, and Glymour 1992). In the case of four or more authors, list only the first author, followed by et al. (Ford et al. 1997).

\subsection{Extracts}
Long quotations and extracts should be indented ten points from the left and right margins.

\begin{quote}
This is an example of an extract or quotation. Note the indent on both sides. Quotation marks are not necessary if you offset the text in a block like this, and properly identify and cite the quotation in the text.

\end{quote}

\subsection{Footnotes}
Avoid footnotes as much as possible; they interrupt the reading of the text. When essential, they should be consecutively numbered throughout with superscript Arabic numbers. Footnotes should appear at the bottom of the page, separated from the text by a blank line space and a thin, half-point rule.

\subsection{Headings and Sections}
When necessary, headings should be used to separate major sections of your paper. Remember, you are writing a short paper, not a lengthy book! An overabundance of headings will tend to make your paper look more like an outline than a paper.

First-level heads should be twelve-point Times Roman bold type, mixed case (initial capitals followed by lower case on all words except articles, conjunctions, and prepositions, which should appear entirely in lower case), with fifteen-point leading, centered, with one blank line preceding them and three additional points of leading following them. Second-level headings should be eleven-point Times Roman bold type, mixed case, with thirteen-point leading, flush left, with one blank line preceding them and three additional points of leading following them. Do not skip a line between paragraphs. Third-level headings should be run in with the text, ten-point Times Roman bold type, mixed case, with twelve-point leading, flush left, with six points of additional space preceding them and no additional points of leading following them.

\subsubsection{Section Numbers}
The use of section numbers in AAAI Press papers is optional. To use section numbers in \LaTeX{}, uncomment the setcounter line in your document preamble and change the 0 to a 1 or 2. Section numbers should not be used in short poster papers.

\subsubsection{Section Headings.}
Sections should be arranged and headed as follows:

\subsubsection{Acknowledgments.}
The acknowledgments section, if included, appears after the main body of text and is headed ``Acknowledgments." This section includes acknowledgments of help from associates and colleagues, credits to sponsoring agencies, financial support, and permission to publish. Please acknowledge other contributors, grant support, and so forth, in this section. Do not put acknowledgments in a footnote on the first page. If your grant agency requires acknowledgment of the grant on page 1, limit the footnote to the required statement, and put the remaining acknowledgments at the back. Please try to limit acknowledgments to no more than three sentences.

\subsubsection{Appendices.}
Any appendices follow the acknowledgments, if included, or after the main body of text if no acknowledgments appear.

\subsubsection{References}
The references section should be labeled ``References" and should appear at the very end of the paper (don't end the paper with references, and then put a figure by itself on the last page). A sample list of references is given later on in these instructions. Please use a consistent format for references. Poorly prepared or sloppy references reflect badly on the quality of your paper and your research. Please prepare complete and accurate citations.

\subsection{Illustrations and Figures}
Figures, drawings, tables, and photographs should be placed throughout the paper near the place where they are first discussed. Do not group them together at the end of the paper. If placed at the top or bottom of the paper, illustrations may run across both columns. Figures must not invade the top, bottom, or side margin areas. Figures must be inserted using your page-formatting software. Number figures sequentially, for example, figure 1, and so on.

The illustration number and caption should appear under the illustration. Leave some space between the figure and the caption and surrounding type; .25 inches should suffice. Captions should be presented in nine-point Times Roman italic. Labels, and other text in illustrations must be at least nine-point type.

\subsubsection{Low-Resolution Bitmaps.}
You may not use low-resolution (such as 72 dpi) screen-dumps and GIF files---these files contain so few pixels that they are always blurry, and illegible when printed. If they are color, they will become an indecipherable mess when converted to black and white. This is always the case with gif files, which should never be used. The resolution of screen dumps can be increased by reducing the print size of the original file while retaining the same number of pixels. You can also enlarge files by manipulating them in software such as PhotoShop. Your figures should be a minimum of 266 dpi when incorporated into your document.

\subsubsection{\LaTeX{} Overflow.}
\LaTeX{} users please beware: \LaTeX{} will sometimes put portions of the figure or table or an equation in the margin. If this happens, you need to scale the figure or table down, or reformat the equation. Check your log file! You must fix any overflow into the margin (that means no overfull boxes in \LaTeX{}). If you don't, the overflow text will simply be eliminated. \textbf{Nothing is permitted to intrude into the margins.}

\subsubsection{Using Color.}
Your paper will be printed in black and white and grayscale. Consequently, because conversion to grayscale can cause undesirable effects (red changes to black, yellow can disappear, and so forth), we strongly suggest you avoid placing color figures in your document. Of course, any reference to color will be indecipherable to your reader.

\subsubsection{Drawings.}
We suggest you use computer drawing software (such as Adobe Illustrator or, (if unavoidable), the drawing tools in Microsoft Word) to create your illustrations. Do not use Microsoft Publisher. These illustrations will look best if all line widths are uniform (half- to two-point in size), and you do not create labels over shaded areas. Shading should be 133 lines per inch if possible. Use Times Roman or Helvetica for all figure call-outs. \textbf{Do not use hairline width lines} --- be sure that the stroke width of all lines is at least .5 pt. Zero point lines will print on a laser printer, but will completely disappear on the high-resolution devices used by our printers.

\subsubsection{Photographs and Images.}
Photographs and other images should be in grayscale (color photographs will not reproduce well; for example, red tones will reproduce as black, yellow may turn to white, and so forth) and set to a minimum of 266 dpi. Do not prescreen images.

\subsection{References} 
The aaai.sty file includes a set of definitions for use in formatting references with BibTeX. These definitions make the bibliography style fairly close to the one specified below. To use these definitions, you also need the BibTeX style file ``aaai.bst," available in the author kit on the AAAI web site. Then, at the end of your paper but before \textbackslash end{document}, you need to put the following lines:

\begin{quote}
\begin{small}
\textbackslash bibliographystyle\{aaai\}
\textbackslash bibliography\{bibfile1,bibfile2,...\}
\end{small}
\end{quote}

The list of files in the \textbackslash  bibliography command should be the names of your BibTeX source files (that is, the .bib files referenced in your paper).

The following commands are available for your use in citing references:
\begin{quote}
\begin{small}
\textbackslash cite: Cites the given reference(s) with a full citation. This appears as ``(Author Year)'' for one reference, or ``(Author Year; Author Year)'' for multiple references.\\
\textbackslash shortcite: Cites the given reference(s) with just the year. This appears as ``(Year)'' for one reference, or ``(Year; Year)'' for multiple references.\\
\textbackslash citeauthor: Cites the given reference(s) with just the author name(s) and no parentheses.\\
\textbackslash citeyear: Cites the given reference(s) with just the date(s) and no parentheses.
\end{small}
\end{quote}

\textbf{Warning:} The aaai.sty file is incompatible with the hyperref and natbib packages. If you use either, your references will be garbled.

Formatted bibliographies should look like the following examples.

\smallskip \noindent \textit{Book with Multiple Authors}\\
Engelmore, R., and Morgan, A. eds. 1986. \textit{Blackboard Systems.} Reading, Mass.: Addison-Wesley.

\smallskip \noindent \textit{Journal Article}\\
Robinson, A. L. 1980a. New Ways to Make Microcircuits Smaller. \textit{Science} 208: 1019--1026.

\smallskip \noindent \textit{Magazine Article}\\
Hasling, D. W.; Clancey, W. J.; and Rennels, G. R. 1983. Strategic Explanations in Consultation. \textit{The International Journal of Man-Machine Studies} 20(1): 3--19.

\smallskip \noindent \textit{Proceedings Paper Published by a Society}\\
Clancey, W. J. 1983b. Communication, Simulation, and Intelligent Agents: Implications of Personal Intelligent Machines for Medical Education. In Proceedings of the Eighth International Joint Conference on Artificial Intelligence, 556--560. Menlo Park, Calif.: International Joint Conferences on Artificial Intelligence, Inc.

\smallskip \noindent \textit{Proceedings Paper Published by a Press or Publisher}\\
Clancey, W. J. 1984. Classification Problem Solving. In \textit{Proceedings of the Fourth National Conference on Artificial Intelligence,} 49--54. Menlo Park, Calif.: AAAI Press.

\smallskip \noindent \textit{University Technical Report}\\
Rice, J. 1986. Poligon: A System for Parallel Problem Solving, Technical Report, KSL-86-19, Dept. of Computer Science, Stanford Univ.

\smallskip \noindent \textit{Dissertation or Thesis}\\
Clancey, W. J. 1979b. Transfer of Rule-Based Expertise through a Tutorial Dialogue. Ph.D. diss., Dept. of Computer Science, Stanford Univ., Stanford, Calif.

\smallskip \noindent \textit{Forthcoming Publication}\\
Clancey, W. J. 1986a. The Engineering of Qualitative Models. Forthcoming.

\section{Producing Reliable PDF\\Documents with \LaTeX{}}
Generally speaking, PDF files are platform independent and accessible to everyone. When creating a paper for a proceedings or publication in which many PDF documents must be merged and then printed on high-resolution PostScript RIPs, several requirements must be met that are not normally of concern. Thus to ensure that your paper will look like it does when printed on your own machine, you must take several precautions:
\begin{itemize}
\item Use type 1 fonts (not type 3 fonts)
\item Use only standard Times, Nimbus, and CMR font packages (not fonts like F3 or fonts with tildes in the names or fonts---other than Computer Modern---that are created for specific point sizes, like Times\~{}19) or fonts with strange combinations of numbers and letters
\item Embed all fonts when producing the PDF
\item Do not use the [T1]{fontenc} package (install the CM super fonts package instead)
\end{itemize}

\subsection{Traditional \LaTeX{} Output}
Most authors using traditional \LaTeX{} output methods will have success by taking the following three steps in creating their paper (called, in the example, proceedingspaper.tex).

\begin{quote}
\begin{small}
latex proceedingspaper\\
dvips -Ppdf -G0 -tletter proceedingspaper\\
ps2pdf -dPDFSETTINGS=/printer\\
-dCompatibilityLevel=1.4 -dMaxSubsetPct=0\\
-dSubsetFonts=false -dEmbedAllFonts=true\\
-sPAPERSIZE=letter proceedingspaper.ps
\end{small}
\end{quote}

Note that the ps2pdf command should be typed all on one line. You can then proceed to distill your PostScript file into a PDF file using GhostScript or Acrobat Distiller.

If your PostScript output still includes type 3 fonts, you should run dvips with option ``dvips -Ppdf -G0 -o papername.ps papername.dvi" (If your machine or site has type 1 fonts, they will probably be loaded.) Note that it is a zero following the ``-G." This tells dvips to use the config.pdf file (and this file refers to a better font mapping). If that doesn't work, you'll have to download the fonts and create a font substitution list.

\subsection{Creating Output Using PDF\LaTeX{}}
PDF\LaTeX{} is a good alternative solution to the \LaTeX{} font problem. By using the PDF\TeX{} program instead of straight \LaTeX{} or \TeX{}, you will probably avoid the type 3 font problem altogether. PDF\LaTeX{} enables you to create a PDF document directly from \LaTeX{} source. The one requirement of this software is that all your graphics and images are available in a format that PDF\LaTeX{} understands (normally PDF).

PDF\LaTeX{}'s default is to create documents with type 1 fonts. If you find that it is not doing so in your case, it is likely that one or more fonts are missing from your system or are not in a path that is known to PDF\LaTeX{}.

One problem with PDF\LaTeX{}, however, is that, by default, it will not embed base 14 fonts. AAAI will embed these fonts for you, but we will not be able to proofread the results. If you are concerned that font substitution may alter your paper in an adverse way (and it can), we recommend that you alter the PDF\TeX{}.cfg configuration file so that the following lines are present and uncommented:

\begin{quote}
\begin{small}
map +bsr.map \% CM/AMS fonts\\
map +bsr-interpolated.map \% additional sizes\\
map +hoekwater.map \% additional fonts\\
\end{small}
\end{quote}

The base 14 Nimbus fonts can be embedded by replacing the line
\textit{map acrobat-std-adobe-suildin.map}
with
\textit{map acrobat-std-urw-kb.map}. 
If this doesn't work, you should look at the PDF\TeX{} mailing list for hints on how to configure PDF\TeX{} or PDF\LaTeX{} to properly embed the typefaces: http://tug.org/pipermail/PDF\TeX{}/2002-July/002803.html

\subsubsection{dvipdf Script}
Scripts such as dvipdf which ostensibly bypass the Postscript intermediary should not be used since they generally do not instruct dvips to use the config.pdf file.

\subsubsection{dvipdfm}
Do not use this dvi-PDF conversion package if your document contains graphics (and we recommend you avoid it even if your document does not contain graphics).

\subsection{Ghostscript}
\LaTeX{} users using GhostScript should make sure that they are using v7.04 or newer. The older versions do not create acceptable PDF files on most platforms.

\subsection{Graphics}
If you are still finding type 3 fonts in your PDF file, look at your graphics! \LaTeX{} users should check all their imported graphics files as well for font problems.

\section{Proofreading Your PDF}
Please check all the pages of your PDF file. Is the page size A4? Are there any type 3, Identity-H, or CID fonts? Are all the fonts embedded? Are there any areas where equations or figures run into the margins? Did you include all your figures? Did you follow mixed case capitalization rules for your title? Did you include a copyright notice? Do any of the pages scroll slowly (because the graphics draw slowly on the page)? Are URLs underlined and in color? You will need to fix these common errors before submitting your file.

\section{Improperly Formatted Files }
In the past, AAAI has corrected improperly formatted files submitted by the authors. Unfortunately, this has become an increasingly burdensome expense that we can no longer absorb. Consequently, if your file is improperly formatted, it may not be possible to include your paper in the publication. If time allows, however, you will be notified via e-mail (with a copy to the program chair) of the problems with your file and given the option of correcting the file yourself (and paying a late fee) or asking that AAAI have the file corrected for you, for an additional fee. If you opt to correct the file yourself, please note that we cannot provide you with any additional advice beyond that given in your packet. Files that are not corrected after a second attempt will be withdrawn.

\subsection{\LaTeX{} 209 Warning}
If you use \LaTeX{} 209 it is extremely unlikely that we will be able to publish your paper. Convert your paper to \LaTeX{}2e.

\section{Naming Your Electronic File}
We request that you name your \LaTeX{} source file with your last name (family name) so that it can easily be differentiated from other submissions. If you name your files with the name of the event or ``aaai" or ``paper" or ``camera-ready" or some other generic or indecipherable name, you bear all risks of loss --- it is extremely likely that your file may be overwritten.

\section{Submitting Your Electronic Files to AAAI}
Submitting your files to AAAI is a two-step process. It is explained fully in the author registration and submission instructions. Please consult this document for details on how to submit your paper.

\section{Inquiries} 
If you have any questions about the preparation or submission of your paper as instructed in this document, please contact AAAI Press at the address given below. If you have technical questions about implementation of the aaai style file, please contact an expert at your site. We do not provide technical support for \LaTeX{} or any other software package. To avoid problems, please keep your paper simple, and do not incorporate complicated macros and style files.

\begin{quote}
\noindent AAAI Press\\
445 Burgess Drive\\
Menlo Park, California 94025\\ 
\textit{Telephone:} (650) 328-3123\\ 
\textit{E-mail:} See the submission instructions for your particular conference or event.
\end{quote}

\section{Possible Bugs in the AAAI Style File}
Some users have found that the aaai.sty does not work properly at their site. They have submitted suggestions for improvement of the macro. You will find those suggestions in the buglist file that is part of author kit, and also as a separate file on the AAAI website. Some of these suggestions have already been implemented, while others seem to be dependent on individual site conditions. If you're having problems with aaai.sty, we suggest you look at the ``bug list" first. The style file is \textbf{not} guaranteed to work in all situations and on all platforms. If you make bug fixes or improvements, please let us know so that we might share them with others.

\section{Additional Resources}
\LaTeX{} is a difficult program to master. If you've used that software, and this document didn't help or some items were not explained clearly, we recommend you read Michael Shell's excellent document (testflow doc.txt V1.0a 2002/08/13) about obtaining correct PS/PDF output on \LaTeX{} systems. (It was written for another purpose, but it has general application as well). It is available at www.ctan.org in the tex-archive.

\section{ Acknowledgments}
AAAI is especially grateful to Peter Patel Schneider for his work in implementing the aaai.sty file, liberally using the ideas of other style hackers, including Barbara Beeton. We also acknowledge with thanks the work of George Ferguson for his guide to using the style and BibTeX files --- which has been incorporated into this document  --- and Hans Guesgen, who provided several timely modifications, as well as the many others who have, from time to time, sent in suggestions on improvements to the AAAI style.

The preparation of the \LaTeX{} and Bib\TeX{} files that implement these instructions was supported by Schlumberger Palo Alto Research, AT\&T Bell Laboratories, Morgan Kaufmann Publishers, The Live Oak Press, LLC, and AAAI Press. Bibliography style changes were added by Sunil Issar. \verb+\+pubnote was added by J. Scott Penberthy. George Ferguson added support for printing the AAAI copyright slug. Additional changes to aaai.sty and aaai.bst have been made by the AAAI staff.

\bigskip
\noindent Thank you for reading these instructions carefully. We look forward to receiving your electronic files!

\end{document}

### 2012-01-19 paper/rmcts/aaai/paper.tex
rearrangment in progress

@@
 
-\title{Rational Monte Carlo Sampling in Trees}
+\title{Simple-Regret Based MCTS}
 \author {David Tolpin, Solomon Eyal Shimony \\
@@
 \{tolpin,shimony\}@cs.bgu.ac.il}
-
 \begin{document}
@@
 \begin{abstract}
-UCT, a state-of-the art algorithm for Monte Carlo tree sampling
-(MCTS), is based on UCB, a sampling policy for the Multi-armed Bandit
-Problem (MAB) that minimizes the accumulated regret. However, MCTS
-differs from MAB in that only the final choice, rather than all arm
-pulls, brings a reward, that is, the simple regret, as opposite to the
-cumulative regret, must be minimized. This ongoing work aims at applying
-meta-reasoning techniques to MCTS, which is non-trivial.
-We begin by introducing policies for
-multi-armed bandits with lower simple regret than UCB, and an
-algorithm for MCTS which combines cumulative and simple regret
-minimization and outperforms UCT. We also develop a sampling scheme loosely based
-on a myopic version of perfect value of information.
-Finite-time and asymptotic analysis of
-the policies is provided, and the algorithms are compared empirically.
+UCT, a state-of-the art algorithm for Monte Carlo tree search (MCTS)
+in games and Markov decision processes, is based on UCB, a sampling
+policy for the Multi-armed Bandit problem (MAB) that asymptotically
+minimizes the cumulative regret.  However, search differs from MAB in
+that in MCTS it is usually only the final ``arm pull'' (the actual
+move selection) that collects a reward, rather than all ``arm pulls''.
+Therefore, it makes more sense to minimize the simple regret, as
+opposed to the cumulative regret. We begin by introducing policies for
+multi-armed bandits with lower finite-time and asymptotic simple
+regret than UCB, using it to develop an algorithm (USRT) for MCTS
+which outperforms UCT empirically.
+
+We then observe that optimizing the sampling process is itself a
+meta-reasoning problem, a solution of which can use value of
+information (VOI) techniques.  Although the theory of VOI for search
+exists, applying it to MCTS is non-trivial, as typical myopic
+assumptions fail. Lacking a working VOI theory for MCTS, we
+nevertheless propose a sampling scheme that is ``aware'' of VOI,
+achieving an algorithm that in empirical evaluation outperforms USRT,
+as well as UCT.
 \end{abstract}
 
-\section{Introduction and definitions}
+\section{Introduction}
 
@@
 UCT does not work; rather, we argue that a simple reconsideration from basic
-principles can result in schemes that outperform UCT. This is demonstrated both
-theoretically and empirically.
+principles can result in schemes that outperform UCT.
+
+The core issue is that in search (especially for adversarial search
+and for ``games against nature'' - optimizing behavior under
+uncertainty) the goal is typically to either find a good (or optimal)
+strategy, or even to find the best first action of such a policy. Once
+such an action is discovered, it is not beneficial to further sample
+that action, ``exploitation'' is thus meaningless for search
+problems. Finding a good first action is closer to the pure
+exploration variant, as seen in the selection problem
+\cite{Bubeck.pure,TolpinShimony.blinkered}. In the selection problem,
+it is much better to minimize the {\em simple} regret.  However, the
+simple and the cumulative regret cannot be minimized simultaneously;
+moreover, \cite{Bubeck.pure} shows that in many cases the smaller the
+cumulative regret, the greater the simple regret.
+
+We begin with introduction of several sampling schemes with better
+bounds for the simple regret on sets. We then extend the resutls to
+sampling in trees by combining the proposed sampling schemes on the
+first step of a rollout with UCT during the rest of the
+rollout. Finally, we empirically evaluate the performance of the
+proposed sampling schemes on sets of Bernoulli arms, in randomly
+generated trees, and on the sailing domain.
+
+\section{Related Work}
+\label{sec:related-work}
+
+Efficient algorithms for Multi-Armed Bandits based on
+distribution-independent bounds, in particular UCB1, are introduced in
+\cite{Auer.ucb}. The UCT algorithm, an extension of UCB1 to
+Monte-Carlo Tree Search is described in \cite{Kocsis.uct}. Pure
+exploration in Multi-armed bandits is explored in
+\cite{Bubeck.pure}. On the one hand, the paper proves certain upper
+and lower bounds for UCB1 and uniform sampling, showing that an upper
+bound on the simple regret is exponential in the number of samples for
+uniform sampling, while only polynomial for UCB1. On the other hand,
+empirical performance of UCB1 appears to be better than of uniform
+sampling.
+
+The principles of bounded rationality appear in
+\cite{Horvitz.reasoningabout}. \cite{Russell.right} provided a formal
+description of rational metareasoning and case studies of applications
+in several problem domains. One obstacle to straightforward
+application of the principles of rational metareasoning to Monte-Carlo
+sampling is the metagreedy assumption, according to which samples must
+be selected as though at most one sample can be taken before an action
+is chosen. In Monte-Carlo sampling, the value of information of any
+single sample in a given search state is often zero, so a different
+approximating assumption must be used instead.
+
+\section{Monte Carlo Search for Simple Regret Minimization}
+\label{sec:results}
 
@@
 
-UCT-driven search achieves good results in many search problems. At each
-search step, the algorithm allocates Monte Carlo rollouts to choose
-the best action, thereby (approximately) minimizing the cumulative regret. 
-The core issue is that in search (especially for adversarial search
-and for ``games against nature'' - optimizing behavior under uncertainty) the goal is typically
-to either find a good (or optimal) strategy, or even to find the best first
-action of such a policy. Once such an action is discovered, it is not beneficial
-to further sample that action, ``exploitation'' is thus meaningless for
-search problems. Finding a good first action is closer to the pure exploration
-variant, as seen in the selection problem \cite{Bubeck.pure,TolpinShimony.blinkered}. In the selection
-problem, it is much better to minimize the {\em simple} regret.
-However, the simple and the cumulative regret cannot be minimized simultaneously;
-moreover, \cite{Bubeck.pure} shows that in many cases the smaller the
-cumulative regret, the greater the simple regret.
-
-UCT performance can thus be improved by combining UCB with a sampling
+UCT performance can be improved by combining UCB with a sampling
 scheme that minimizes the simple regret of selecting an action at the
@@
 \end{itemize}
-
-We begin with introduction of several sampling schemes with
-better bounds for the simple regret on sets. We then
-extend the resutls to sampling in trees by combining the proposed sampling
-schemes on the first step of a rollout with UCT during the rest of the
-rollout. Finally, we empirically
-evaluate the performance of the proposed sampling schemes on sets of
-Bernoulli arms, in randomly generated
-trees, and on the sailing domain
-. Finally, we describe future research
-directions and provide preliminary results for a sampling scheme
-explicitly based on the value of information of a rollout.
-
-\section{Related Work}
-\label{sec:related-work}
-
-Efficient algorithms for Multi-Armed Bandits based on
-distribution-independent bounds, in particular UCB1, are introduced in
-\cite{Auer.ucb}. The UCT algorithm, an extension of UCB1 to
-Monte-Carlo Tree Search is described in \cite{Kocsis.uct}. Pure
-exploration in Multi-armed bandits is explored in
-\cite{Bubeck.pure}. On the one hand, the paper proves certain upper
-and lower bounds for UCB1 and uniform sampling, showing that an upper
-bound on the simple regret is exponential in the number of samples for
-uniform sampling, while only polynomial for UCB1. On the other hand,
-empirical performance of UCB1 appears to be better than of uniform
-sampling.
-
-The principles of bounded rationality appear in
-\cite{Horvitz.reasoningabout}. \cite{Russell.right} provided a formal
-description of rational metareasoning and case studies of applications
-in several problem domains. One obstacle to straightforward
-application of the principles of rational metareasoning to Monte-Carlo
-sampling is the metagreedy assumption, according to which samples must
-be selected as though at most one sample can be taken before an action
-is chosen. In Monte-Carlo sampling, the value of information of any
-single sample in a given search state is often zero, so a different
-approximating assumption must be used instead.
-
-\section{Main Results}
-\label{sec:results}
-
-The ultimate goal is ``optimal'' sampling in the meta-reasoning
-sense. Nevertheless, this task is daunting for the following reasons:
-\begin{itemize}
-\item Defining the cost of a sample is not easy, and even if we simply
-  use time-cost as an approximation, we get an intractable
-  meta-reasoning problem.
-\item Applying the standard myopic and subtree independence
-  assumptions, we run into serious problems. Even in the standard
-  selection problem \cite{TolpinShimony.blinkered}, we get a
-  non-concave utility function and premature stopping of the
-  algorithm. This is due to the fact that the value of information of
-  a single measurement (analogous to a sample in MCTS) is frequently
-  less than its time-cost, even though this is not true for multiple
-  measurements.  When applying the selection problem to MCTS, the
-  situation is exacerbated.  The utility of an action is usually
-  bounded, and thus in many cases a single sample may be insufficient
-  to change the current best action, {\em regardless} of its
-  outcome. As a result, we frequently get a {\em zero} ``myopic''
-  value of information for a single sample.
-\end{itemize}
-
-As the ultimate task is extremely difficult
-to achieve, and even harder to analyze, we introduce
-simple schemes more amenable to analysis, and compare
+Here, we introduce simple schemes more amenable to analysis, and compare
 them to UCB (on sets) and UCT (in trees).
@@
   \vspace{1em}
-  \includegraphics[scale=0.45]{costs-size=6-nsamples=793.pdf}\\
-  b. 793 rollouts\\
-  \vspace{1em}
   \includegraphics[scale=0.45]{costs-size=6-nsamples=1585.pdf}\\
@@
 \begin{figure}[t]
-  \begin{minipage}[b]{1.0\linewidth}
-    \centering
-    \includegraphics[scale=0.45]{costs-size=3-nsamples=397.pdf}\\
-    a. $3\times 3$ lake
-    \vspace{1em}
-  \end{minipage}
-  \begin{minipage}[b]{1.0\linewidth}
    \centering
    \includegraphics[scale=0.45]{costs-size=6-nsamples=397.pdf}\\
-    b. $6\times 6$ lake
+   b. $6\times 6$ lake \\
    \vspace{1em}
-  \end{minipage}
-  \begin{minipage}[b]{1.0\linewidth}
-    \centering
    \includegraphics[scale=0.45]{costs-size=10-nsamples=397.pdf}\\
    c. $10\times 10$ lake
-  \end{minipage}
   \caption{The sailing domain, 397 samples, cost vs. factor}
@@
   b. 397 rollouts\\
-  \vspace{1em}
-  \includegraphics[scale=0.45]{rcq-size=6-nsamples=793.pdf}\\
-  c. 793 rollouts\\
   \vspace{1em}

### 2012-01-20 paper/rmcts/aaai/paper.tex
aaai paper in progress

@@
 
-\title{Simple-Regret Based MCTS}
+\newtheorem{dfn}{Definition}
+\newtheorem{thm}{Theorem}
+\newtheorem{lmm}{Lemma}
+\newtheorem{crl}{Corollary}
+
+\title{MCTS Based on Simple Regret}
 \author {David Tolpin, Solomon Eyal Shimony \\
@@
 
-\section{Monte Carlo Search for Simple Regret Minimization}
+\section{Sampling Based on Simple Regret}
 \label{sec:results}
 
-In the Multi-armed Bandit problem we have a set of $K$ arms. Each arm can be pulled
-multiple times. When the $i$th arm is pulled, a random reward $X_i$
-from an unknown stationary distribution is returned.  The reward is 
-bounded between 0 and 1.
+In the Multi-armed Bandit problem \cite{Vermorel.bandits} we have a set
+of $K$ arms. Each arm can be pulled multiple times. When the $i$th arm
+is pulled, a random reward $X_i$ from an unknown stationary
+distribution is returned.  The reward is bounded between 0 and 1.
 
-The simple regret of a sampling policy for the Multi-armed Bandit
+\begin{dfn}
+The \textbf{simple regret} of a sampling policy for the Multi-armed Bandit
 Problem is the expected difference between the best expected reward
-$\mu_*$ and the expected reward $\mu_j$ of the arm with the best sample mean
+$\mu_*$ and the expected reward $\mu_j$ of the empirically best arm
 $\overline X_j=\max_i\overline X_i$:
@@
 where $\Delta_j=\mu_*-\mu_j$.
+\end{dfn}
+
 Strategies that minimize the simple regret are called pure exploration
-strategies \cite{Bubeck.pure}.
-
-Such strategies are used to select the best arm, and, by extension,
-the best action in Monte Carlo tree search. Monte Carlo tree search is
-used to solve Markov Decision Processes (MDP) approximately. An MDP is
-defined by the set of states $S$, the set of actions $A$, the
-transition table $T(s, a, s')$, the reward table $R(s, a, s')$, the
-initial state $s$ and the goal state $t$: $(S, A, T, R, s, t)$
-\cite{Russell.aima}.  Monte Carlo tree search explores an MDP by
-\emph{rollouts}---trajectories from the current state to a state in
-which a termination condition is satisfied (either the goal state, or
-a cutoff state for which the reward is evaluated
-approximately). The UCB algorithm (that attempts to minimize the cumulative regret)
-\cite{Auer.ucb} had been extended into the tree search sampling scheme
-known as UCT \cite{Kocsis.uct}.
+strategies \cite{Bubeck.pure}. Such strategies are used to select the
+best arm, and, by extension, the best action in Monte Carlo tree
+search. Monte Carlo tree search is used to solve Markov Decision
+Processes (MDP) approximately. An MDP is defined by the set of states
+$S$, the set of actions $A$, the transition table $T(s, a, s')$, the
+reward table $R(s, a, s')$, the initial state $s$ and the goal state
+$t$: $(S, A, T, R, s, t)$ \cite{Russell.aima}.  Monte Carlo tree
+search explores an MDP by \emph{rollouts}---trajectories from the
+current state to a state in which a termination condition is satisfied
+(either the goal state, or a cutoff state for which the reward is
+evaluated approximately). The UCB algorithm (that attempts to minimize
+the cumulative regret) \cite{Auer.ucb} had been extended into the tree
+search sampling scheme known as UCT \cite{Kocsis.uct}.
 
@@
 \end{itemize}
-Here, we introduce simple schemes more amenable to analysis, and compare
+Ultimately, an ``optimal'' sampling in the meta-reasoning sense should
+be used at the first step in the above scheme. Nevertheless, this task
+is daunting for the following reasons:
+\begin{itemize}
+\item Defining the cost of a sample is not easy, and even if we simply
+  use time-cost as an approximation, we get an intractable
+  meta-reasoning problem.
+\item Applying the standard myopic and subtree independence
+  assumptions, we run into serious problems. Even in the standard
+  selection problem \cite{TolpinShimony.blinkered}, we get a
+  non-concave utility function and premature stopping of the
+  algorithm. This is due to the fact that the value of information of
+  a single measurement (analogous to a sample in MCTS) is frequently
+  less than its time-cost, even though this is not true for multiple
+  measurements.  When applying the selection problem to MCTS, the
+  situation is exacerbated.  The utility of an action is usually
+  bounded, and thus in many cases a single sample may be insufficient
+  to change the current best action, {\em regardless} of its
+  outcome. As a result, we frequently get a {\em zero} ``myopic''
+  value of information for a single sample.
+\end{itemize}
+As the ultimate goal is extremely difficult
+to achieve, and even harder to analyze, we introduce
+simple schemes more amenable to analysis, and compare
 them to UCB (on sets) and UCT (in trees).
@@
 
-\begin{enumerate}
-\item The $\varepsilon$-greedy allocation scheme has exponential convergencerate of the simple regret $\IE r$ (see Appendix~\ref{app:derivations-finite-eps}):
+\begin{dfn} Scheme $\mathbf{UCB(\alpha)}$ repeatedly pulls arm $i$ that maximizes 
+upper confidence bound $b_i$ on the reward:
 \begin{equation}
-  \IE r_\varepsilon\le\sum_{j=1}^K\Delta_j\left(\frac {n\varepsilon} K + \frac {4\sqrt e}
-{\Delta_j^2}\right)e^{-\Delta_j^2n\varepsilon/8K}
+b_i=\overline X_i+\sqrt {\frac {\alpha \ln n} {n_i}}
 \end{equation}
-\item The UCB allocation scheme exhibits polynomial convergence
-rate (see Appendix~\ref{app:derivations-finite-ucb}):
+where $\overline X_i$ is the average reward obtained from arm $i$,
+$n_i$ is the number of times arm $i$ was pulled, and $n$ is the total
+number of pulls so far. \end{dfn} The best known upper bound on the simple
+regret of $UCB(\alpha)$ is polynomially decreasing in the number of samples
+(see \cite{Bubeck.pure}, Theorems~2,3).
+
+An upper bound on the simple regret of uniform sampling is
+exponentially decreasing in the number of samples (see
+\cite{Bubeck.pure}, Proposition~1). However, empirically
+$UCB(\alpha)$ yields a lower simple regret than uniform
+sampling.
+
+We introduce here two sampling schemes with exponentially
+decreasing upper bounds on the simple regret. The bounds
+suggest that these schemes achieve a lower simple regret
+than uniform sampling; indeed, this is confirmed
+by experiments.
+
+We first consider $\varepsilon$-greedy sampling as a straightforward
+generalization of uniform sampling:
+\begin{dfn} The \textbf{$\mathbf{\varepsilon}$-greedy} sampling scheme
+pulls the empirically best arm with probability $\varepsilon$ and any other
+arm with equal probability $\frac {1-\varepsilon} {K-1}$. 
+\end{dfn}
+
+\begin{thm} Simple regret of the  $\varepsilon$-greedy
+sampling scheme is asymptotically bounded from above as
 \begin{equation}
-\IE r_{ucb} \le 2\sum_{j=1}^K \Delta_jn^{\frac {-\alpha \Delta_j^2} 8}
+\IE r_{\varepsilon\mbox{-}greedy}\le...
 \end{equation}
-\item An asymptotic upper bound on the simple regret for the
-$\varepsilon$-greedy allocation scheme depends \emph{non-monotonically} on
-$\varepsilon$. The lowest bound is achieved for $\varepsilon=\frac 1
-2$, and for a large number of arms the bound for the $\frac 1
-2$-greedy scheme approaches the square of the bound for uniform
-sampling (see Appendix~\ref{app:derivations-asym-eps}).
+the bound is minimized for $\varepsilon=\frac 1 2$.
+\end{thm}
 
-\item Another improved allocation scheme for minimizing the simple regret,
-called \textbf{UQB} here, is obtained by substituting $\sqrt{\cdot}$
-instead of $\log(\cdot)$ into the formula for UCB and has a
-superpolynomial convergence rate (see
-Appendix~\ref{app:derivations-asym-ucb}).
-\end{enumerate}
+\begin{proof}[Proof sketch:] TODO
+\end{proof}
 
+\begin{dfn} Scheme $\mathbf{UCB_{\sqrt{\cdot}}(\alpha)}$ repeatedly pulls arm $i$ that
+maximizes $b_i$:
+\begin{equation}
+b_i=\overline X_i+\sqrt {\frac {\alpha \sqrt n} {n_i}}
+\end{equation}
+where, as before, $\overline X_i$ is the average reward obtained from arm $i$,
+$n_i$ is the number of times arm $i$ was pulled, and $n$ is the total
+number of pulls so far. \end{dfn}
+
+\begin{thm} Simple regret of the $UCB_{\sqrt{}}(\alpha)$ sampling
+  scheme is asymptotically bounded from above as.
+\begin{equation}
+\IE r_{ucb\sqrt{\cdot}} \le ...
+\end{equation}
+\end{thm}
+\begin{proof}[Proof sketch:] TODO
+\end{proof}
 \subsection{Sampling in Trees}
@@
 
-The {\bf two-stage} Monte Carlo tree search sampling
-scheme selects an action at the current root node according to an
-algorithm that minimizes the simple regret, such as $\frac 1 2$-greedy or
-UQB, and then selects actions according to UCB. This scheme
-outperforms the UCT sampling scheme that selects actions based on
-UCB only.
+We suggest an improvement on the UCT algorithm, which combines
+different sampling schemes on the first step and during the rest of a rollout:
+\begin{dfn}
+The {\bf two-stage MCTS sampling scheme} selects an action at the
+current root node according to an algorithm that minimizes the simple
+regret, such as $\frac 1 2$-greedy or $UCB_{\sqrt{\cdot}}$, and then
+selects actions according to $UCB$.
+\end{dfn}
 
-Additionaly, such two-stage sampling scheme is significantly less
-sensitive to the tuning of the exploration factor ($C_p$ in
-\cite{Kocsis.uct}) of UCT, since the conflict \cite{Bubeck.pure}
-between the need for a larger value of $C_p$ on the first step (simple
+This scheme outperforms the UCT sampling scheme that selects actions based on
+UCB only. Additionaly, such two-stage sampling scheme is significantly less
+sensitive to the tuning of the exploration factor $\alpha$ of UCT,
+since the conflict \cite{Bubeck.pure}
+between the need for a larger value of $\alpha$ on the first step (simple
 regret) and a smaller value for the rest of the rollout (cumulative
 regret) is resolved. In fact, a sampling scheme that uses UCB at all
-steps but a larger value of $C_p$ for the first step than for the rest
+steps but a larger value of $\alpha$ for the first step than for the rest
 of the steps, outperforms UCT. The pseudocode of the two-stage rollout
@@
 \end{algorithm}
+
+\subsection{VOI-aware Sampling}
+
+TODO
 
@@
 Figure~\ref{fig:mab-simple-regret}.b shows the regret
-vs. the number of samples, averaged
-over $10^4$ experiments for randomly generated instances of 32 arms.
+vs. the number of samples, averaged over $10^4$ experiments for
+randomly generated instances of 32 arms.
 
 For smaller numbers of samples, GCT achieves the best
-performance; for large number of samples, QCT outperforms GCT. QCT is
+performance; for larger numbers of samples, QCT outperforms GCT. QCT is
 better than UCT everywhere except for very small numbers of samples. A
@@
 vs. the number of samples, computed for a range of values of
-$C_p$. Figure~\ref{fig:sailing-cost-vs-nsamples}.a shows the median
+$\alpha$. Figure~\ref{fig:sailing-cost-vs-nsamples}.a shows the median
 cost, and Figure~\ref{fig:sailing-cost-vs-nsamples}.b --- the minimum
-costs. UCT is always worse than either $\frac 1 2$-greedy (GCT) or QCT, and is sensitive to
-the value of $C_p$: the median cost is much higher than the minimum
-cost for UCT. For both $\frac 1 2$-greedy and QCT, the difference is
-significantly less prominent.
+costs. UCT is always worse than either $\frac 1 2$-greedy (GCT) or
+QCT, and is sensitive to the value of $\alpha$: the median cost is
+much higher than the minimum cost for UCT. For both $\frac 1 2$-greedy
+and QCT, the difference is significantly less prominent.
 
@@
 UCT, and $\frac 1 2$-greedy is better than UCT expect for a small range of
-values of the exploration factor $C_p$.
+values of the exploration factor $\alpha$.
 
@@
 increases.
-
 \begin{figure}[t]
@@
 Figure~\ref{fig:sailing-rcq-vs-factor} compares QCT with UCT with a
-different exploration factor at the root (CCT). $C_p$ for the rest of
+different exploration factor at the root (CCT). $\alpha$ for the rest of
 the steps was chosen to ensure the best performance from earlier
@@
 but as the number of samples grows, QCT achieves smaller average
-regrets and is less sensitive to the choice of the value for $C_p$ at
+regrets and is less sensitive to the choice of the value for $\alpha$ at
 the first step.
-
 \begin{figure}[t]
@@
 \end{figure}
-
 VOI of a rollout can be computed when the sample distribution of an
@@
 
-\section{Appendixes}
-
-\subsection{Facts}
-
-{\bf Chernoff bounds (see \cite{Hagerup.chernoff}):} for $m$ independent random variables $X_1, X_2, ..., X_m$
-taking on values 0 or 1, $X=\sum_{i=1}^m X_i$, and $0\le\delta\le 1$:
-\begin{equation}
-\Pr[X < (1-\delta)\IE[X]] < e^{-\delta^2\IE[X]/2} \mbox{ (Chernoff bound)}
-\label{eq:chernoff-bound}
-\end{equation}
-Further on, if $\forall i. \IE[X_i]=\mu$:
-\begin{equation}
-\Pr[X < n\mu-a] <  e^{-2a^2/n}
-\label{eq:chernoff-hoeffding-bound} \mbox{ (Hoeffding-Chernoff bound)}
-\end{equation}
-
-\subsection{Derivations}
-\label{app:derivations}
-
-\subsubsection{Finite-time regret bounds}
-\label{app:derivations-finite}
-
-%TODO: bounds on regret based on information gathered, independent of
-%the way the choice is made.
-
-For an analysis of pure exploration of uniform and UCB algorithms, see also
-\cite{Bubeck.pure}.
-
-\paragraph{$\varepsilon$-greedy}
-\label{app:derivations-finite-eps}
-
-For $0<\varepsilon\le1-\frac 1 K$ and $x_0>0$ (see Section 3, Proofs, in \cite{Auer.ucb}):
-
-\begin{eqnarray}
-\Pr(\overline X_j&=&\max_i\overline X_i)\nonumber\\
-   &\le&2\left(x_0\cdot \Pr\left\{T_j^R(n)\le x_0\right\} + \frac 2{\Delta_j^2}e^{-\Delta_j^2\lfloor x_0 \rfloor/2}\right)\nonumber\\
-   &\le&2\left(x_0\cdot \Pr\left\{T_j^R(n)\le x_0\right\} + \frac {2
-  \cdot  e^{\Delta_j^2/2}}{\Delta_j^2}e^{-\Delta_j^2 x_0
-  /2}\right)\nonumber\\
-   &&
-\end{eqnarray}
-
-Number of times $\xi$ the $j$th arm is selected {\it randomly}:
-
-\begin{equation}
-\xi\triangleq\IE\left[T_j^R(n)\right]=\frac {n\varepsilon} K
-\end{equation}
-
-By choosing $x_0=\frac \xi 2$ and applying the Chernoff bound (\ref{eq:chernoff-bound}), obtain:
-
-\begin{eqnarray}
-\label{eq:pr-epsilon-greedy}
-\Pr(\overline X_j=\max_i\overline X_i)&\le& 2\left(\frac {\xi}{2} e^{-\xi/8} + \frac {2e^{\Delta_j^2/2}}{\Delta_j^2}e^{-\Delta_j^2 \xi/4}\right)\nonumber\\
-&\le&\left(\xi + \frac {4\sqrt e}{\Delta_j^2}\right)e^{-\Delta_j^2\xi/8}
-\end{eqnarray}
-
-A bound on the simple regret $r_\varepsilon$ of an
-$\varepsilon$-greedy policy follows from substituting
-(\ref{eq:pr-epsilon-greedy}) into (\ref{eq:simple-regret}):
-
-\begin{eqnarray}
-  \IE r_\varepsilon&\le&\sum_{j=1}^K\Delta_j\left(\xi + \frac {4\sqrt e}
-{\Delta_j^2}\right)e^{-\Delta_j^2\xi/8}\nonumber\\
-&\le&K\left(\xi + \frac {4\sqrt
-    e}{\Delta_{\min}}\right)e^{-\Delta_{\min}^2\xi/8}
-\label{eq:deriv-epsilon-greedy-b}
-\end{eqnarray}
-
-\paragraph{UCB($\alpha$)}
-\label{app:derivations-finite-ucb}
-
-$n_*$ pulls of the best arm and $n_j$ pulls of the $j$th arm are
-achieved simultaneously when either
-
-\begin{equation}
-\sqrt {\frac {\alpha \log n} {n_j-1}} \ge 1 + \sqrt { \frac {\alpha \log n} {n_*}}
-\end{equation}
-
-or
-
-\begin{equation}
-\sqrt {\frac {\alpha \log n} {n_j}} \le 1 + \sqrt { \frac {\alpha \log n} {n_* - 1}}
-\label{eq:deriv-ucb-alpha-a}
-\end{equation}
-
-Inequality (\ref{eq:deriv-ucb-alpha-a}) can be used to derive a lower
-bound on the number of pulls of a non-optimal arm,  and, consequently, the upper
-bound on the probability of selecting a non-optimal arm:
-
-\begin{eqnarray}
-\frac {\alpha \log n} {n_j}&\le& 1 + 2  \sqrt {\frac {\alpha \log n}
-  {n_*-1}} + \frac {\alpha \log n} {n_*-1}\nonumber\\
-n_j &\ge& \frac {\alpha \log n} {1+2 \sqrt {\frac {\alpha \log n}
-    {n_*-1}} + \frac {\alpha \log n} {n_*-1}}\nonumber\\
-    &\ge& \frac {\alpha \log n} {4} \mbox{ for } \alpha \log n \le n_*-1
-\end{eqnarray}
-
-The probability that the $j$th arm will be chosen can be bounded using
-the Chernoff-Hoeffding bound:
-\begin{equation}
-P_j \le 2 \exp\left(-\frac {\alpha \Delta_j^2 \log n} 8 \right) = 2n^{\frac {-\alpha \Delta_j^2} 8}
-\end{equation}
-
-The expected simple regret of an UCB($\alpha$) algorithm is thus
-bounded by the following inequality:
-\begin{equation}
-\IE r_{ucb} \le 2\sum_{j=1}^K \Delta_jn^{\frac {-\alpha \Delta_j^2} 8}
-\label{eq:deriv-ucb-alpha-b}
-\end{equation}
-
-\cite{Bubeck.pure} establishes for UCB($\alpha$) a similar bound,
-polynomial in the number of pulls $n$.
-
-\subsubsection{Asymptotic regret analysis}
-\label{app:derivations-asym}
-
-%TODO: Oracle instead of asymptotic
-
-Bounds (\ref{eq:deriv-epsilon-greedy-b}, \ref{eq:deriv-ucb-alpha-b}) were
-derived ignoring dependency of the allocation on expected values
-of non-optimal arms. Derivation of better bounds seems
-complicated. Instead, one can analyze the asymptotic behavior of each of the
-algorithms under the assumption that the sample mean is close to the expectation.
-
-\paragraph{$\varepsilon$-greedy}
-\label{app:derivations-asym-eps}
-
-A common technique \cite{Auer.ucb} to bound the probability of
-selecting a non-optimal arm $j$ is to split the interval between
-$\mu_j$ and $\mu_*$ at the middle and bound the probability of
-selecting the $j$th arm by the sum of the probabilities that the sample
-mean $\overline X_j$ of the $j$th arm is greater than $\mu_j+\frac
-{\Delta_j} 2$ and the sample mean $\overline X_*$ of the optimal arm is
-less than $\mu_*- \frac {\Delta_j} 2$. Each of the probabilities can
-be bound using the Chernoff-Hoefding bound.
-
-Splitting the interval between $\mu_j$ and $\mu_*$ half-way is an
-arbitrary decision. By selecting a different split point
-depending on the allocation of samples, one can obtain a tighter
-bound. On the other hand, there is a value of $\varepsilon$ that
-minimizes the bound for the $\varepsilon$-greedy allocation scheme.
-Thus, the  $\varepsilon$-greedy allocation scheme can be tuned by
-finding $\varepsilon$ that minimizes the bound obtained by
-selecting the best split point.
-
-Split the interval $[\mu_j, \mu_*]$ at $\mu_j+\delta_j$. The
-probability $P_j$ for selecting the
-$j$th arm is bounded, using the union bound and the Chernoff-Hoeffding bound, as:
-
-\begin{eqnarray}
-P_j&\le&\Pr[\overline X_j>\mu_j+\delta_j]+\Pr[\overline X_*<\mu_*-(\Delta_j-\delta_j)]\nonumber\\
-   &\le&\exp\left(-2\delta_j^2n_j\right)+\exp\left(-2(\Delta_j-\delta_j)^2n_*\right)
-\end{eqnarray}
-
-For the $\varepsilon$-greedy scheme, $n_i=\frac {\varepsilon n} K$,
-$n_*=(1-\varepsilon)n$, thus
-
-\begin{equation}
-P_j \le \exp\left(-\frac {2\delta_j^2\varepsilon n} {K-1}\right) +\exp\left(-2(\Delta_j-\delta_j)^2(1-\varepsilon)n\right)
-\end{equation}
-
-In general, $\varepsilon$ and $\delta$ which minimize $P_j$ will
-depend of $n$. For a constant $\varepsilon$, require that the
-exponents in both terms are equal
-\begin{eqnarray}
-\exp\left(-\frac {2\delta_j^2\varepsilon n} {K-1}\right)
-   &=& \exp\left(-2(\Delta_j-\delta_j)^2(1-\varepsilon)n\right)\\
-\frac {\delta_j} {\Delta_j - \delta_j}
-   &=& \sqrt { \frac {(K-1)(1-\varepsilon)} \varepsilon }
-\label{eq:deriv-exps-equal}
-\end{eqnarray}
-and find conditional minima of $P_j(\delta_j, \epsilon)$ satisfying
-either $\frac {\partial P_j} {\partial \delta_j}=0$ or $\frac {\partial P_j}
-{\partial \varepsilon}=0$.
-
-\vspace{1em}
-\paragraph{a. {\large $\frac {\partial P_j} {\partial \delta_j}=0$}}
-
-\begin{eqnarray*}
-\frac {4\delta_j\varepsilon n} {K-1}
-  &-& 4(\Delta_j-\delta_j)(1-\varepsilon)n=0\\
-\frac {\delta_j} {\Delta_j-\delta_j}&=&{ \frac {(K-1)(1-\varepsilon)} \varepsilon }
-\end{eqnarray*}
-and together with (\ref{eq:deriv-exps-equal}) this gives:
-\begin{eqnarray}
-\delta_j&=&\frac {\Delta_j} 2\nonumber\\
-\varepsilon&=&1-\frac 1 K\nonumber\\
-\IE r_{\varepsilon=1-1/K}&\le&B_{uniform}=2\exp\left(-\frac {\Delta_j^2n} {2K}\right)
-\label{eq:deriv-uniform-sampling}
-\end{eqnarray}
-
-That is, the case $\frac {\partial P_j} {\partial \delta_j}=0$
-corresponds to \emph{uniform random sampling}, and the tightest bound is obtained
-by splitting the interval at the middle.
-
-\paragraph{b. {\large $\frac {\partial P_j} {\partial \varepsilon}=0$}}
-
-\begin{eqnarray*}
-\frac {2\delta_j^2 n} {K-1}
-  &-& 2(\Delta_j-\delta_j)^2n=0\\
-\frac {\delta_j} {\Delta_j-\delta_j}&=&\sqrt {K-1}
-\end{eqnarray*}
-and together with (\ref{eq:deriv-exps-equal}) this gives:
-\begin{eqnarray}
-\frac {\delta_j} {\Delta_j}&=&\frac 1 {\sqrt {K-1}+1}\nonumber\\
-\varepsilon&=&\frac 1 2\nonumber\\
-\IE r_{\varepsilon=1/2}&\le&B_{1/2\mbox{-}greedy}\nonumber\\
-                     &=&2\exp\left(-\frac{\Delta_j^2n} {(\sqrt{K-1}+1)^2}\right)\nonumber\\
-\lim_{K\to\infty}
-\mathbb{B}r_{1/2\mbox{-}greedy}^K&=&2\exp\left(-{\Delta_j^2n}\right)\nonumber\\
-\mbox{for large $K$: }\mathbb{B}r_{1/2\mbox{-}greedy}&\approx&2\exp\left(-\frac {\Delta_j^2n} K\right)
-\label{eq:deriv-half-greedy-sampling}
-\end{eqnarray}
-
-That is, the case $\frac {\partial P_j} {\partial \varepsilon}=0$
-corresponds to \emph{$\frac 1 2$-greedy sampling}, and for a large number of
-arms $K$ 
-\begin{equation}
-\mathbb{B}r_{1/2\mbox{-}greedy}\approx \mathbb{B}r_{uniform}^2
-\label{eq:derive-uniform-vs-half-greedy}
-\end{equation}
-Equation (\ref{eq:derive-uniform-vs-half-greedy}) suggests
-that \emph{for a large number of arms} $\frac 1 2$-greedy
-sampling has a much higher convergence rate than uniform
-sampling.
-
-\paragraph{UCB($\alpha$)}
-\label{app:derivations-asym-ucb}
-
-The $UCB(\alpha)$ allocation scheme distributes arm pulls such that
-the upper bound on the probability of selecting a non-optimal arm is
-asymptotically the same for all arms. Indeed, for large n
-\begin{eqnarray}
-n_i&\ll&n\\
-n_*&\approx&n\\
-\sqrt {\frac {2 \log n} {n_j}}&\approx&\Delta_j+\sqrt {\frac {2\log n}
-  {n_*}} \\
-&\Rightarrow& n_j \approx \frac {2\log n} {\Delta_j^2}
-\label{eq:deriv-ucb-asymptotic-nj}
-\end{eqnarray}
-and by Hoeffding-Chernoff bound the upper bound $\mathbb{B}P_j$ on the probability
-$P_j^{ucb}$ to choose the $j$th arm is
-\begin{equation}
-P_j^{ucb}\le\mathbb{B}P_j^{ucb}\approx2\exp\left(-2\Delta_j^2\frac {2\log n}{\Delta_j^2}\right)
-             =2n^{-4}
-\end{equation}
-
-That is, UCB($\alpha$) supposedly distributes pulls of non-optimal
-arms better than $\varepsilon$-greedy
-scheme which may undersample arms with high expectation. The
-polynomial, rather than exponential, convergence rate (Equation
-\ref{eq:deriv-ucb-alpha-b}) is due to oversampling the best arm.
-
-Replacing the logarithm with a faster growing subpolynomial
-function, like square root (UQB), results in a superpolynomial convergence
-rate of the upper bound:
-\begin{equation}
-P_j^{uqb}\le\mathbb{B}P_j^{uqb}\approx2\exp\left(-\beta\Delta_j^2\frac {\sqrt n}{\Delta_j^2}\right)
-            =2\exp(-\beta\sqrt n)
- \end{equation}
-where $\beta$ is a constant that depends only on reward bounds.
-Finite time convergence can still be shown similarly to UCB.
+\clearpage
 

### 2012-01-20 paper/rmcts/aaai/paper.tex
in progress

@@
 \pdfinfo{
-/Title (Rational Monte Carlo Sampling in Trees)
+/Title (MCTS Based on Simple Regret)
 /Subject (AAAI Publications)
@@
 \cite{Bubeck.pure,TolpinShimony.blinkered}. In the selection problem,
-it is much better to minimize the {\em simple} regret.  However, the
+it is much better to minimize the \emph{simple} regret.  However, the
 simple and the cumulative regret cannot be minimized simultaneously;
@@
 strategies \cite{Bubeck.pure}. Such strategies are used to select the
-best arm, and, by extension, the best action in Monte Carlo tree
-search. Monte Carlo tree search is used to solve Markov Decision
-Processes (MDP) approximately. An MDP is defined by the set of states
-$S$, the set of actions $A$, the transition table $T(s, a, s')$, the
-reward table $R(s, a, s')$, the initial state $s$ and the goal state
-$t$: $(S, A, T, R, s, t)$ \cite{Russell.aima}.  Monte Carlo tree
-search explores an MDP by \emph{rollouts}---trajectories from the
+best arm, and, by extension, the best action in MCTS. MCTS is used to
+solve Markov Decision Processes (MDP) approximately. An MDP is defined
+by the set of states $S$, the set of actions $A$, the transition table
+$T(s, a, s')$, the reward table $R(s, a, s')$, the initial state $s$
+and the goal state $t$: $(S, A, T, R, s, t)$ \cite{Russell.aima}.
+MCTS explores an MDP by \emph{rollouts}---trajectories from the
 current state to a state in which a termination condition is satisfied
@@
   bounded, and thus in many cases a single sample may be insufficient
-  to change the current best action, {\em regardless} of its
-  outcome. As a result, we frequently get a {\em zero} ``myopic''
+  to change the current best action, \emph{regardless} of its
+  outcome. As a result, we frequently get a \emph{zero} ``myopic''
   value of information for a single sample.
@@
 pulls the empirically best arm with probability $\varepsilon$ and any other
-arm with equal probability $\frac {1-\varepsilon} {K-1}$. 
+arm with probability $\frac {1-\varepsilon} {K-1}$. 
 \end{dfn}
@@
 \end{proof}
+
 \subsection{Sampling in Trees}
@@
 
-We suggest an improvement on the UCT algorithm, which combines
-different sampling schemes on the first step and during the rest of a rollout:
+The UCT sampling scheme \cite{Kocsis.uct} is a widely used generalization
+of UCB for MCTS.  UCT applies the UCB sampling
+scheme at each step of a rollout.  We suggest an improvement on UCT,
+which combines different sampling schemes on the first step and during
+the rest of a rollout:
 \begin{dfn}
@@
 
-This scheme outperforms the UCT sampling scheme that selects actions based on
-UCB only. Additionaly, such two-stage sampling scheme is significantly less
-sensitive to the tuning of the exploration factor $\alpha$ of UCT,
-since the conflict \cite{Bubeck.pure}
-between the need for a larger value of $\alpha$ on the first step (simple
-regret) and a smaller value for the rest of the rollout (cumulative
-regret) is resolved. In fact, a sampling scheme that uses UCB at all
-steps but a larger value of $\alpha$ for the first step than for the rest
-of the steps, outperforms UCT. The pseudocode of the two-stage rollout
-is in Algorithm~\ref{alg:two-stage-mcts}.
+Such two-stage sampling scheme outperforms UCT, and, additionally, is
+significantly less sensitive to the tuning of the exploration factor
+$\alpha$ of UCT, since the conflict between the need for a larger
+value of $\alpha$ on the first step (simple regret) and a smaller
+value for the rest of the rollout (cumulative regret)
+\cite{Bubeck.pure} is resolved. In fact, a sampling scheme that uses
+UCB at all steps but a larger value of $\alpha$ for the first step
+than for the rest of the steps, outperforms UCT. The pseudocode of the
+two-stage rollout is in Algorithm~\ref{alg:two-stage-mcts}.
 
@@
 
-TODO
+A further improvement can be achieved by computing or estimating the VOI
+of the rollouts and choosing a rollout that maximizes the VOI. VOI of
+a rollout can be computed when the sample distribution of an
+action is known up to the parameters, such as the normal distribution
+with an unknown mean and/or variance. Alternatively, the value of
+information can be estimated from the set of samples, and the need to
+assume a particular shape of the distribution can be lifted. In one
+realization of the latter approach the VOI of performing an action is
+estimated as an \emph{upper bound on the value of perfect information about
+an arm, divided by the number of pulls of the arm}. 
+\begin{equation}
+\end{equation}
+Early experiments with this approach demonstrated significantly lower
+simple regret on a wide range of problems.
 
@@
   \end{minipage}
+  \caption{Simple regret in MAB}
   \label{fig:mab-simple-regret}
-  \caption{Simple regret in MAB}
 \end{figure}
@@
 
-The improvement is inspired by the notion of value of information (VOI),
-but VOI is used implicitly in the analysis of the algorithm, rather
-than computed or learned explicitly in order to plan the rollouts. A
-further improvement can be achieved by computing or estimating the VOI
-of the rollouts and choosing a rollout that maximizes the VOI. Using VOI
-to guide Monte Carlo sampling instead of relying on the sample means
-of different actions  is particularly beneficial when different
-actions have qualitatively different distributions: for example, when the
-distribution variance varies significantly between the actions.
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[scale=0.35,trim=0 0 0 16pt,clip]{flat-trilevel-vct-k=64.pdf}\\
-  \caption{VOI-based action selection (VCT) vs. other algorithms}
-  \label{fig:flat-trilevel-vct}
-\end{figure}
-VOI of a rollout can be computed when the sample distribution of an
-action is known up to the parameters, such as the normal distribution
-with an unknown mean and/or variance. Alternatively, the value of
-information can be estimated from the set of samples, and the need to
-assume a particular shape of the distribution can be lifted. In one
-realization of the latter approach the VOI of performing an action is
-estimated as \emph{the value of perfect information learned from the
-  outcome of earlier samples of the action divided by the number of
-  samples}. Early experiments with this approach showed the best
-performance for sets of Bernoulli arms, and even a more prominent
-improvement for mixed arms (e.g. Bernoulli, triangularly distributed,
-and fixed arms) (Figure~\ref{fig:flat-trilevel-vct} shows results for
-different algorithms for 32 randomly generated arms): \textbf{VCT},
-the algorithm which selects a rollout with the maximum VOI estimate
-exhibits lower average regret than other algorithms. Sample-based VOI
-estimation requires keeping more information than an algorithm based
-only on the sample mean, but the overhead is small and justified by
-the improvement.
-
 \section*{Acknowledgments}

### 2012-01-21 paper/rmcts/aaai/paper.tex
in progress
added formulas for VOI estimates

@@
 \label{eq:simple-regret}
-\IE[R]=\sum_{j=1}^K\Delta_j\Pr(\overline X_j=\max_i\overline X_i)
+\IE r=\sum_{j=1}^K\Delta_j\Pr(\overline X_j=\max_i\overline X_i)
 \end{equation}
@@
 
-The UCT sampling scheme \cite{Kocsis.uct} is a widely used generalization
-of UCB for MCTS.  UCT applies the UCB sampling
-scheme at each step of a rollout.  We suggest an improvement on UCT,
-which combines different sampling schemes on the first step and during
-the rest of a rollout:
+UCT \cite{Kocsis.uct} is a generalization of UCB for MCTS.  UCT
+applies UCB at each step of a rollout.  We suggest an improvement on
+UCT, which combines different sampling schemes on the first step and
+during the rest of a rollout:
 \begin{dfn}
-The {\bf two-stage MCTS sampling scheme} selects an action at the
-current root node according to an algorithm that minimizes the simple
-regret, such as $\frac 1 2$-greedy or $UCB_{\sqrt{\cdot}}$, and then
-selects actions according to $UCB$.
+The \textbf{SR+CR MCTS sampling scheme} selects an action at the
+current root node according to a scheme suitable for minimizing 
+simple regret (\textbf{SR}), such as $\frac 1 2$-greedy or $UCB_{\sqrt{\cdot}}$, and
+then selects actions according to $UCB$, minimizing cumulative regret (\textbf{CR}).
 \end{dfn}
@@
 
-A further improvement can be achieved by computing or estimating the VOI
-of the rollouts and choosing a rollout that maximizes the VOI. VOI of
-a rollout can be computed when the sample distribution of an
-action is known up to the parameters, such as the normal distribution
-with an unknown mean and/or variance. Alternatively, the value of
-information can be estimated from the set of samples, and the need to
-assume a particular shape of the distribution can be lifted. In one
-realization of the latter approach the VOI of performing an action is
-estimated as an \emph{upper bound on the value of perfect information about
-an arm, divided by the number of pulls of the arm}. 
-\begin{equation}
-\end{equation}
-Early experiments with this approach demonstrated significantly lower
-simple regret on a wide range of problems.
+A further improvement can be achieved by computing or estimating the
+value of information (VOI) of the rollouts and choosing a rollout that
+maximizes the VOI. VOI of a rollout can be computed when the sample
+distribution of an action is known up to the parameters, such as the
+normal distribution with an unknown mean and/or
+variance. Alternatively, the value of information can be estimated
+from the set of samples, and the need to assume a particular shape of
+the distribution can be lifted. In one realization of the latter
+approach the VOI of performing an action is estimated as an
+\emph{upper bound on the value of perfect information about an arm,
+divided by the number of pulls of the arm}.
+\begin{eqnarray}
+VOI_\alpha&\approx&\frac {\overline X_\beta} {n_\alpha+1}
+\exp\left(-2(\overline X_\alpha - \overline X_\beta)^2 n_\alpha\right)\\
+VOI_i&\approx&\frac {1-\overline X_\alpha} {n_i+1}
+\exp\left(-2(\overline X_\alpha - \overline X_i)^2 n_i\right),\; i\ne\alpha\nonumber\\
+\mbox{where }\alpha&=&\arg\max_i \overline X_i\nonumber\\
+             \beta&=&\arg\max_{i,\,i\ne\alpha} \overline X_i\nonumber
+\end{eqnarray}
+Early experiments with this approach demonstrated
+significantly lower simple regret on a wide range of problems.
 
@@
 
-\begin{figure}[t]
+\begin{figure}[h!]
   \begin{minipage}[c]{1.0\linewidth}
@@
 
-\begin{figure}[t]
+\begin{figure}[h!]
   \begin{minipage}[b]{1.0\linewidth}
@@
 
-\begin{figure}[t]
+\begin{figure}[h!]
   \centering
@@
 increases.
-\begin{figure}[t]
+\begin{figure}[h!]
    \centering
@@
 the first step.
-\begin{figure}[t]
+\begin{figure}[h!]
   \centering
@@
 
-The Monte Carlo tree search algorithms presented in the paper differ
+The MTCS SR+CR scheme presented in the paper differs
 from UCT at the first step of the rollout, when the `simple' selection
@@
 theoretical analysis and the empirical evaluation provide evidence for
-better general performance of the proposed algorithms.
+better general performance of the proposed scheme.
 
@@
 
-\clearpage
-
  \bibliographystyle{aaai}

### 2012-01-21 paper/rmcts/aaai/paper.tex
lots of changes to graphics

@@
 number of pulls so far. \end{dfn} The best known upper bound on the simple
-regret of $UCB(\alpha)$ is polynomially decreasing in the number of samples
+regret of UCB($\alpha$) is polynomially decreasing in the number of samples
 (see \cite{Bubeck.pure}, Theorems~2,3).
@@
 \cite{Bubeck.pure}, Proposition~1). However, empirically
-$UCB(\alpha)$ yields a lower simple regret than uniform
+UCB($\alpha$) yields a lower simple regret than uniform
 sampling.
@@
 
-\begin{thm} Simple regret of the $UCB_{\sqrt{}}(\alpha)$ sampling
+\begin{thm} Simple regret of the UCB$_{\sqrt{}}$($\alpha$) sampling
   scheme is asymptotically bounded from above as.
@@
 current root node according to a scheme suitable for minimizing 
-simple regret (\textbf{SR}), such as $\frac 1 2$-greedy or $UCB_{\sqrt{\cdot}}$, and
-then selects actions according to $UCB$, minimizing cumulative regret (\textbf{CR}).
+simple regret (\textbf{SR}), such as $\frac 1 2$-greedy or UCB$_{\sqrt{\cdot}}$, and
+then selects actions according to UCB, minimizing cumulative regret (\textbf{CR}).
 \end{dfn}
@@
 The results were empirically verified on Multi-armed Bandit instances,
-on search trees, and on the sailing domain, as defined in \cite{Kocsis.uct}. The
-experiments confirmed the hypotheses. The algorithms used in the
-experiments are:
-\begin{description}
-\item[RND:] uniform \underline{r}a\underline{nd}om samping;
-\item[UCT:] \underline{U}pper \underline{C}onfidence Bounds applied to
-  \underline{T}rees \cite{Kocsis.uct};
-\item[RCT:] U\underline{CT} with the first step replaced with uniform
-  \underline{r}andom sampling;
-\item[GCT:] U\underline{CT} with the first step
-  replaced with $\frac 1 2$-\underline{g}reedy sampling;
-\item[QCT:] U\underline{CT} with the first step
-  replaced with U\underline{Q}B.
-\end{description}
+on search trees, and on the sailing domain, as defined in
+\cite{Kocsis.uct}. In most cases, the experiments showed lower average
+simple regret for $\frac 1 2$-greedy an UCB$_{\sqrt{\cdot}}$ than for
+UCB on sets, and for the SR+CR scheme than for UCT in trees.
 
@@
     \centering
-    \includegraphics[scale=0.4]{tree-identity-k=16-uqb=8.pdf}\\ 
+    \includegraphics[scale=0.5]{tree-identity-k=16-uqb=8.pdf}\\ 
     b. 16 arms
@@
     \centering
-    \includegraphics[scale=0.4]{tree-identity-k=64-uqb=8.pdf} \\
+    \includegraphics[scale=0.5]{tree-identity-k=64-uqb=8.pdf} \\
     c. 64 arms

### 2012-01-21 code/figures.R

@@ -2,19 +2,22 @@ ALGORITHMS <- c('rnd', 'ucb', 'sve')
 
 
 draw.regrets <- function(regrets_file, log="xy", legend.position='topright', xlab=expression(N[samples]), main=NA) {
-	regrets <- read.table(regrets_file, header=T)
-	xlim <- range(regrets$nsamples)
-	values <- unlist(regrets[-1])
+	regrets <- read.table(regrets_file, header=F)
+
+        names <- as.matrix(regrets[1,])
+        regrets <- as.matrix(regrets[-1,])
+        regrets <- matrix(as.numeric(regrets), nrow=nrow(regrets), ncol=ncol(regrets))
+
+        xlim <- range(regrets[,1])
+	values <- unlist(regrets[,-1])
 	ylim <- range(values[values>0])
-	plot(x=regrets$nsamples, type='n', log=log, xlim=xlim, ylim=ylim, yaxs='i', xlab=xlab, ylab='Regret', main=if(is.na(main)) unlist(strsplit(regrets_file, "\\."))[[1]] else main)
-	i <- 1
-	for(alg in algs) {
-		y <- regrets[[alg]]
-		lines(x=regrets$nsamples, y=regrets[[alg]],
-			  type="o", pch=i)
-		i <- i+1
+
+	plot(x=regrets[1], type='n', log=log, xlim=xlim, ylim=ylim, yaxs='i', xlab=xlab, ylab='Regret', main=if(is.na(main)) unlist(strsplit(regrets_file, "\\."))[[1]] else main)
+        for(i in (2:ncol(regrets))) {
+          lines(x=regrets[,1], y=regrets[,i],
+                type="o", pch=i-1)
 	}
-	legend(x=legend.position, legend=sapply(names(regrets)[-1], function(n) substring(n, 3)), pch=1:length(names(regrets)[-1]))
+	legend(x=legend.position, legend=sapply(names[-1], function(n) substring(n, 3)), pch=1:length(names[-1]))
 }
 
 draw.experiment <- function(prefix="") {

### 2012-01-21 code/figures.R
better figures

@@ -12,7 +12,7 @@ draw.regrets <- function(regrets_file, log="xy", legend.position='topright', xla
 	values <- unlist(regrets[,-1])
 	ylim <- range(values[values>0])
 
-	plot(x=regrets[1], type='n', log=log, xlim=xlim, ylim=ylim, yaxs='i', xlab=xlab, ylab='Regret', main=if(is.na(main)) unlist(strsplit(regrets_file, "\\."))[[1]] else main)
+	plot(x=regrets[,1], type='n', log=log, xlim=xlim, ylim=ylim, yaxs='i', xlab=xlab, ylab='Regret', main=if(is.na(main)) unlist(strsplit(regrets_file, "\\."))[[1]] else main)
         for(i in (2:ncol(regrets))) {
           lines(x=regrets[,1], y=regrets[,i],
                 type="o", pch=i-1)

### 2012-01-21 code/sailing/figures.R

@@ -1,40 +1,45 @@
 draw.costs.of.factor <- function(costs_file, legend.position='bottomright', log="x",main=NA) {
-  costs <- read.table(costs_file, header=T)
-  algorithms <- names(costs)[-1]
-  xlim <- range(costs$FACTOR)
-  ylim <- range(costs[algorithms])
-  plot(x=costs$nsamples, type='n', log=log, xlim=xlim, ylim=ylim, xaxs='i',
+  costs <- read.table(costs_file, header=F)
+  names <- as.matrix(costs[1,])[-1]
+  costs <- as.matrix(costs[-1,])
+  costs <- matrix(as.numeric(costs), nrow=nrow(costs), ncol=ncol(costs))
+
+  fctr <- costs[,1]
+  fctr <- fctr*fctr*4
+  xlim <- range(fctr)
+  ylim <- range(unlist(costs[,-1]))
+  plot(x=fctr, type='n', log=log, xlim=xlim, ylim=ylim, xaxs='i',
        xlab='exploration factor', ylab='cost',
        main=if(is.na(main)) do.call(paste,as.list(unlist(strsplit(costs_file,
                             "-|\\.txt",fixed=FALSE)))[-1])
             else main)
-  i <- 1
-  for(alg in algorithms) {
-    lines(x=costs$FACTOR, y=costs[[alg]],
-          type="o", pch=i)
-    i <- i+1
+
+  for(i in (2:ncol(costs))) {
+    lines(x=fctr, y=costs[,i],
+          type="o", pch=i-1)
   }
-  legend(x=legend.position, legend=algorithms, pch=1:length(algorithms), bg='white')
+  legend(x=legend.position, legend=names, pch=1:length(names), bg='white')
 }
 
 draw.costs.of.nsamples <- function(costs_file, legend.position='right', log="x", main=NA) {
-  costs <- read.table(costs_file, header=T)
-  algorithms <- names(costs)[-1]
-  xlim <- range(costs$SAMPLES)
-  ylim <- range(costs[algorithms])
-  print(costs)
-  plot(x=costs$nsamples, type='n', log=log, xlim=xlim, ylim=ylim, xaxs='i',
+  costs <- read.table(costs_file, header=F)
+  names <- as.matrix(costs[1,])[-1]
+  costs <- as.matrix(costs[-1,])
+  costs <- matrix(as.numeric(costs), nrow=nrow(costs), ncol=ncol(costs))
+
+  smpls <- costs[,1]
[…]

### 2012-01-21 paper/rmcts/aaai/paper.tex
better figures
figures done

@@
 multi-armed bandits with lower finite-time and asymptotic simple
-regret than UCB, using it to develop an algorithm (USRT) for MCTS
+regret than UCB, using it to develop a two-stage scheme (SR+CR) for MCTS
 which outperforms UCT empirically.
@@
 nevertheless propose a sampling scheme that is ``aware'' of VOI,
-achieving an algorithm that in empirical evaluation outperforms USRT,
-as well as UCT.
+achieving an algorithm that in empirical evaluation outperforms 
+both UCT and the other proposed algorithms.
 \end{abstract}
@@
 
-For smaller numbers of samples, GCT achieves the best
-performance; for larger numbers of samples, QCT outperforms GCT. QCT is
-better than UCT everywhere except for very small numbers of samples. A
-combination of GCT and QCT dominates UCT over the whole range.
+For smaller numbers of samples, $\frac 1 2$-greedy achieves the best
+performance; for larger numbers of samples, UCB$_{\sqrt{\cdot}}$
+outperforms $\frac 1 2$-greedy. A combination of $\frac 1
+2$-greedy  and UCB$_{\sqrt{\cdot}}$ dominates UCB over the
+whole range.
 
@@
 dependencies look differently from Multi-armed bandit instances, but
-the algorithms exhibit a similar relative performance: either GCT or QCT
-is always better than UCT, QCT dominates UCT everywhere
-except for small numbers of instances. The advantage of QCT grows with
-the number of arms.
+the algorithms exhibit a similar relative performance: either $\frac 1
+2$-greedy+UCT or UCB$_{\sqrt{\cdot}}$+UCT
+gives the lowest regret, UCB$_{\sqrt{\cdot}}$+UCT dominates UCT everywhere
+except for small numbers of instances. The advantage of both $\frac 1
+2$-greedy+UCT and UCB$_{\sqrt{\cdot}}$+UCT grows with the number of arms.
 
@@
 cost, and Figure~\ref{fig:sailing-cost-vs-nsamples}.b --- the minimum
-costs. UCT is always worse than either $\frac 1 2$-greedy (GCT) or
-QCT, and is sensitive to the value of $\alpha$: the median cost is
-much higher than the minimum cost for UCT. For both $\frac 1 2$-greedy
-and QCT, the difference is significantly less prominent.
+costs. UCT is always worse than either $\frac 1 2$-greedy+UCT or
+UCB$_{\sqrt{\cdot}}$+UCT, and is sensitive to the value of $\alpha$: the median cost is
+much higher than the minimum cost for UCT. For both $\frac 1 2$-greedy+UCT
+and UCB$_{\sqrt{\cdot}}$+UCT, the difference is significantly less prominent.
 
@@
   \end{minipage}
-  \caption{The sailing domain, $6\times 6$ lake, cost vs. number of rollouts}
+  \caption{The sailing domain, $6\times 6$ lake, cost vs. samples}
   \label{fig:sailing-cost-vs-nsamples}
@@
 Figure~\ref{fig:sailing-cost-vs-factor} shows the regret vs. the
-exploration factor for different numbers of samples. QCT is always better than
-UCT, and $\frac 1 2$-greedy is better than UCT expect for a small range of
+exploration factor for different numbers of samples. UCB$_{\sqrt{\cdot}}$+UCT is always better than
+UCT, and $\frac 1 2$-greedy+UCT is better than UCT expect for a small range of
 values of the exploration factor $\alpha$.
@@
 
-Figure~\ref{fig:sailing-rcq-vs-factor} compares QCT with UCT with a
-different exploration factor at the root (CCT). $\alpha$ for the rest of
+Figure~\ref{fig:sailing-rcq-vs-factor} compares UCB$_{\sqrt{\cdot}}$+UCT with UCT with a
+different exploration factor at the root (UCB+UCT). $\alpha$ for the rest of
 the steps was chosen to ensure the best performance from earlier
 experiments on the domain. Both algorithms exhibit similar dependency,
-but as the number of samples grows, QCT achieves smaller average
+but as the number of samples grows, UCB$_{\sqrt{\cdot}}$+UCT achieves smaller average
 regrets and is less sensitive to the choice of the value for $\alpha$ at

### 2012-01-21 paper/rmcts/aaai/paper.tex
future work

@@
 We begin with introduction of several sampling schemes with better
-bounds for the simple regret on sets. We then extend the resutls to
+bounds for the simple regret on sets. We then extend the results to
 sampling in trees by combining the proposed sampling schemes on the
@@
 b_i=\overline X_i+\sqrt {\frac {\alpha \ln n} {n_i}}
+\label{eqn:ucb}
 \end{equation}
@@
 
-We introduce here two sampling schemes with exponentially
+We introduce here two sampling schemes with superpolynomially
 decreasing upper bounds on the simple regret. The bounds
@@
 \end{dfn}
-
-\begin{thm} Simple regret of the  $\varepsilon$-greedy
-sampling scheme is asymptotically bounded from above as
+This sampling scheme exhibits an exponentially decreasing simple regret:
+\begin{thm} For any $\pi>0$  and $\gamma>0$ there exists $N$ such that for
+  any number of samples $n>N$ the simple regret of the  $\varepsilon$-greedy
+sampling scheme is  bounded from above as
 \begin{equation}
-\IE r_{\varepsilon\mbox{-}greedy}\le...
+\IE r_{\varepsilon\mbox{-}greedy}\le (1+\gamma) \sum_{i=1}^K\Delta_i\exp(\cdot)
 \end{equation}
+with probability at least $1-\pi$.
 the bound is minimized for $\varepsilon=\frac 1 2$.
@@
 
-\begin{proof}[Proof sketch:] TODO
+\begin{proof}[Proof outline:] TODO
 \end{proof}
 
+$\varepsilon$-greedy is based solely on sampling the empirically best
+arm with a higher probability then the rest of the arms, and ignores
+information about empirical means of other arms. On the other hand,
+UCB distributes samples in accordance with empirical means, but, in order to
+minimize cumulative regret, chooses the empirically best arm too often.
+Intuitively, a better scheme for simple regret minimization would
+distribute samples in a way similar to UCB, but would sample the best arm
+less. This can be achieved by replacing $\log(\cdot)$ in
+Equation~\ref{eqn:ucb} with a faster growing sublinear function, for
+example, $\sqrt\cdot$.
 \begin{dfn} Scheme $\mathbf{UCB_{\sqrt{\cdot}}(\alpha)}$ repeatedly pulls arm $i$ that
@@
 number of pulls so far. \end{dfn}
+This scheme also exhibits a superpolynomially decreasing simple regret:
+\begin{thm}  For any $\pi>0$  and $\gamma>0$ there exists $N$ such that for
+  any number of samples $n>N$ the simple regret of the  UCB$_{\sqrt{}}$($\alpha$)
+sampling scheme is  bounded from above as
+\begin{equation}
+\IE r_{ucb\sqrt{\cdot}} \le (1+\gamma)\sum_{i=1}^K\Delta_i\exp\left(-2\alpha\sqrt{n}\right)
+\end{equation}
+with probability at least $1-\pi$.
+\end{thm}
 
-\begin{thm} Simple regret of the UCB$_{\sqrt{}}$($\alpha$) sampling
-  scheme is asymptotically bounded from above as.
-\begin{equation}
-\IE r_{ucb\sqrt{\cdot}} \le ...
-\end{equation}
-\end{thm}
-\begin{proof}[Proof sketch:] TODO
+\begin{proof}[Proof outline:] TODO
 \end{proof}
@@
 current root node according to a scheme suitable for minimizing 
-simple regret (\textbf{SR}), such as $\frac 1 2$-greedy or UCB$_{\sqrt{\cdot}}$, and
-then selects actions according to UCB, minimizing cumulative regret (\textbf{CR}).
+the simple regret (\textbf{SR}), such as $\frac 1 2$-greedy or UCB$_{\sqrt{\cdot}}$, and
+then selects actions according to UCB, minimizing the cumulative regret (\textbf{CR}).
 \end{dfn}
@@
 \subsection{VOI-aware Sampling}
+\label{sec:voi-sampling}
 
@@
 Early experiments with this approach demonstrated
-significantly lower simple regret on a wide range of problems.
+a significantly lower simple regret on a wide range of problems.
 
@@
 on search trees, and on the sailing domain, as defined in
-\cite{Kocsis.uct}. In most cases, the experiments showed lower average
+\cite{Kocsis.uct}. In most cases, the experiments showed a lower average
 simple regret for $\frac 1 2$-greedy an UCB$_{\sqrt{\cdot}}$ than for
@@
   d. 1585 rollouts\\
-  \caption{The sailing domain, log vs. sqrt, $6\times 6$ lake}
+  \caption{The sailing domain, UCB vs. UCB$_{\sqrt{\cdot}}$, $6\times 6$ lake}
   \label{fig:sailing-rcq-vs-factor}
@@
 
-The MTCS SR+CR scheme presented in the paper differs
+The MCTS SR+CR scheme presented in the paper differs
 from UCT at the first step of the rollout, when the `simple' selection
@@
 better general performance of the proposed scheme.
+
+The improvement is inspired by the notion of value of information
+(VOI), but VOI is used implicitly in the analysis of the algorithm,
+rather than computed or learned explicitly in order to plan the
+rollouts. Preliminary results on VOI-aware sampling suggest that a
+further improvement in the sampling performance can be
+achieved. However, application of the theory of rational metareasoning
+to Monte Carlo Tree Search is an open problem \cite{HayRussell.MCTS},
+and both a solid theoretical model and empirically efficient VOI
+estimates still need to be developed.
 

### 2012-01-21 paper/rmcts/aaai/paper.tex
kind of proof for epsilon-greedy

@@
 \begin{equation}
-\label{eq:simple-regret}
 \IE r=\sum_{j=1}^K\Delta_j\Pr(\overline X_j=\max_i\overline X_i)
+\label{eqn:simple-regret}
 \end{equation}
@@
 \begin{dfn} The \textbf{$\mathbf{\varepsilon}$-greedy} sampling scheme
-pulls the empirically best arm with probability $\varepsilon$ and any other
+pulls the empirically best arm with probability
+$0<\varepsilon<1$ and any other
 arm with probability $\frac {1-\varepsilon} {K-1}$. 
@@
 This sampling scheme exhibits an exponentially decreasing simple regret:
-\begin{thm} For any $\pi>0$  and $\gamma>0$ there exists $N$ such that for
+\begin{thm} For any $0<\pi<10$  and $\gamma>1$ there exists $N$ such that for
   any number of samples $n>N$ the simple regret of the  $\varepsilon$-greedy
@@
 \begin{equation}
-\IE r_{\varepsilon\mbox{-}greedy}\le (1+\gamma) \sum_{i=1}^K\Delta_i\exp(\cdot)
+\IE r_{\varepsilon\mbox{-}greedy}\le 2\gamma \sum_{i=1}^K\Delta_i\exp\left(-\frac {2\Delta_j^2n\varepsilon}
+  {\left(1+\sqrt{\frac {(K-1)\varepsilon}
+        {1-\varepsilon}}\right)^2}\right)
 \end{equation}
 with probability at least $1-\pi$.
-the bound is minimized for $\varepsilon=\frac 1 2$.
 \end{thm}
 
-\begin{proof}[Proof outline:] TODO
+\begin{proof}[Proof outline:]
+
+Bound the probability $P_j$ that a non-optimal arm $j$ is selected. Split the interval
+ $[\mu_j, \mu_*]$ at $\mu_j+\delta_j$. By Chernoff-Hoeffding bound:
+\begin{eqnarray}
+P_j&\le&\Pr[\overline X_j>\mu_j+\delta_j]+\Pr[\overline X_*<\mu_*-(\Delta_j-\delta_j)]\nonumber\\
+   &\le&\exp\left(-2\delta_j^2n_j\right)+\exp\left(-2(\Delta_j-\delta_j)^2n_*\right)
+\end{eqnarray}
+Observe that, in probability, $\overline X_j \rightarrow \mu_j$ as $n\rightarrow\infty$, 
+therefore $n_*\rightarrow n\varepsilon$, $n_j\rightarrow\frac
+{n(1-\varepsilon)} {K-1}$ as $n\rightarrow \infty$. Conclude that for
+any $0<\pi<1$, $\gamma>1$ there exists $N$ such that for any $n>N$ and
+all non-optimal arms $j$:
+\begin{equation}
+P_j \le \gamma \left(\exp\left(-\frac {2\delta_j^2n(1-\varepsilon)}{K-1}\right)
++\exp\left(-2(\Delta_j-\delta_j)^2n\varepsilon\right)\right)
+\label{eqn:epsgreedy-probbound}
+\end{equation}
+Require
+\begin{eqnarray}
+\exp\left(-\frac {2\delta_j^2n(1-\varepsilon)}{K-1}\right)
+&=&\exp\left(-2(\Delta_j-\delta_j)^2n\varepsilon\right)\nonumber\\
+\frac {\delta_j} {\Delta_j-\delta_j}&=&\sqrt{\frac {(K-1)\varepsilon} {1-\varepsilon}}
+\label{eqn:epsgreedy-constant-varepsilon}
+\end{eqnarray}
+Substitute (\ref{eqn:epsgreedy-probbound}) together with
+(\ref{eqn:epsgreedy-constant-varepsilon}) into
+(\ref{eqn:simple-regret}) and obtain
+\begin{equation}
+\IE r_{\varepsilon\mbox{-}greedy}\le 2\gamma \sum_{i=1}^K\Delta_i\exp\left(-\frac {2\Delta_j^2n\varepsilon}
+  {\left(1+\sqrt{\frac {(K-1)\varepsilon}
+        {1-\varepsilon}}\right)^2}\right)
+\end{equation}
 \end{proof}
+
+In particular, as the number of arms $K$ grows, the bound for $\frac 1
+2$-greedy sampling ($\varepsilon=\frac 1 2$) becomes considerably tighter than for uniform
+random sampling ($\varepsilon=\frac 1 K$):
+\begin{crl}
+For uniform random sampling, 
+\begin{equation}
+\IE r_{uniform}\le 2\gamma \sum_{i=1}^K\Delta_i\exp\left(-\frac {\Delta_j^2n} {K}\right)
+\end{equation}
+For $\frac 1 2$-greedy sampling,
+\begin{eqnarray}
+\IE r_{\frac 1 2\mbox{-}greedy}&\le& 2\gamma \sum_{i=1}^K\Delta_i\exp\left(-\frac {2\Delta_j^2n}
+  {\left(1+\sqrt{(K-1)}\right)^2}\right)\nonumber\\
+  &\approx& 2\gamma \sum_{i=1}^K\Delta_i\exp\left(-\frac
+    {2\Delta_j^2n} {K}\right)\mbox{ for large K}\nonumber\\
+\end{eqnarray}
+\end{crl}
 
@@
 This scheme also exhibits a superpolynomially decreasing simple regret:
-\begin{thm}  For any $\pi>0$  and $\gamma>0$ there exists $N$ such that for
-  any number of samples $n>N$ the simple regret of the  UCB$_{\sqrt{}}$($\alpha$)
+\begin{thm}  For any $0<\pi<1$  and $\gamma>1$ there exists $N$ such that for
+  any number of samples $n>N$ the simple regret of the  UCB$_{\sqrt{\cdot}}$($\alpha$)
 sampling scheme is  bounded from above as
 \begin{equation}
-\IE r_{ucb\sqrt{\cdot}} \le (1+\gamma)\sum_{i=1}^K\Delta_i\exp\left(-2\alpha\sqrt{n}\right)
+\IE r_{ucb\sqrt{\cdot}} \le \gamma\sum_{i=1}^K\Delta_i\exp\left(-2\alpha\sqrt{n}\right)
 \end{equation}
@@
 and both a solid theoretical model and empirically efficient VOI
-estimates still need to be developed.
+estimates need to be developed.
 

### 2012-01-21 paper/srcr/paper.tex
moved to a new place

--- 2012-01-21 paper/rmcts/aaai/paper.tex
@@
 
-We introduce here two sampling schemes with superpolynomially
+We introduce here two sampling schemes with super-polynomially
 decreasing upper bounds on the simple regret. The bounds
@@
 
-Bound the probability $P_j$ that a non-optimal arm $j$ is selected. Split the interval
- $[\mu_j, \mu_*]$ at $\mu_j+\delta_j$. By Chernoff-Hoeffding bound:
+Bound the probability $P_i$ that a non-optimal arm $i$ is selected. Split the interval
+ $[\mu_i, \mu_*]$ at $\mu_i+\delta_i$. By Chernoff-Hoeffding bound:
 \begin{eqnarray}
-P_j&\le&\Pr[\overline X_j>\mu_j+\delta_j]+\Pr[\overline X_*<\mu_*-(\Delta_j-\delta_j)]\nonumber\\
-   &\le&\exp\left(-2\delta_j^2n_j\right)+\exp\left(-2(\Delta_j-\delta_j)^2n_*\right)
+P_i&\le&\Pr[\overline X_i>\mu_i+\delta_i]+\Pr[\overline X_*<\mu_*-(\Delta_i-\delta_i)]\nonumber\\
+   &\le&\exp\left(-2\delta_i^2n_i\right)+\exp\left(-2(\Delta_i-\delta_i)^2n_*\right)
 \end{eqnarray}
-Observe that, in probability, $\overline X_j \rightarrow \mu_j$ as $n\rightarrow\infty$, 
-therefore $n_*\rightarrow n\varepsilon$, $n_j\rightarrow\frac
+Observe that, in probability, $\overline X_i \rightarrow \mu_i$ as $n\rightarrow\infty$, 
+therefore $n_*\rightarrow n\varepsilon$, $n_i\rightarrow\frac
 {n(1-\varepsilon)} {K-1}$ as $n\rightarrow \infty$. Conclude that for
 any $0<\pi<1$, $\gamma>1$ there exists $N$ such that for any $n>N$ and
-all non-optimal arms $j$:
+all non-optimal arms $i$:
 \begin{equation}
-P_j \le \gamma \left(\exp\left(-\frac {2\delta_j^2n(1-\varepsilon)}{K-1}\right)
-+\exp\left(-2(\Delta_j-\delta_j)^2n\varepsilon\right)\right)
+P_i \le \gamma \left(\exp\left(-\frac {2\delta_i^2n(1-\varepsilon)}{K-1}\right)
++\exp\left(-2(\Delta_i-\delta_i)^2n\varepsilon\right)\right)
 \label{eqn:epsgreedy-probbound}
@@
 \begin{eqnarray}
-\exp\left(-\frac {2\delta_j^2n(1-\varepsilon)}{K-1}\right)
-&=&\exp\left(-2(\Delta_j-\delta_j)^2n\varepsilon\right)\nonumber\\
-\frac {\delta_j} {\Delta_j-\delta_j}&=&\sqrt{\frac {(K-1)\varepsilon} {1-\varepsilon}}
+\exp\left(-\frac {2\delta_i^2n(1-\varepsilon)}{K-1}\right)
+&=&\exp\left(-2(\Delta_i-\delta_i)^2n\varepsilon\right)\nonumber\\
+\frac {\delta_i} {\Delta_i-\delta_i}&=&\sqrt{\frac {(K-1)\varepsilon} {1-\varepsilon}}
 \label{eqn:epsgreedy-constant-varepsilon}
@@
 \begin{equation}
-\IE r_{\varepsilon\mbox{-}greedy}\le 2\gamma \sum_{i=1}^K\Delta_i\exp\left(-\frac {2\Delta_j^2n\varepsilon}
+\IE r_{\varepsilon\mbox{-}greedy}\le 2\gamma \sum_{i=1}^K\Delta_i\exp\left(-\frac {2\Delta_i^2n\varepsilon}
   {\left(1+\sqrt{\frac {(K-1)\varepsilon}
@@
 \begin{equation}
-\IE r_{uniform}\le 2\gamma \sum_{i=1}^K\Delta_i\exp\left(-\frac {\Delta_j^2n} {K}\right)
+\IE r_{uniform}\le 2\gamma \sum_{i=1}^K\Delta_i\exp\left(-\frac {\Delta_i^2n} {K}\right)
 \end{equation}
@@
 \begin{eqnarray}
-\IE r_{\frac 1 2\mbox{-}greedy}&\le& 2\gamma \sum_{i=1}^K\Delta_i\exp\left(-\frac {2\Delta_j^2n}
+\IE r_{\frac 1 2\mbox{-}greedy}&\le& 2\gamma \sum_{i=1}^K\Delta_i\exp\left(-\frac {2\Delta_i^2n}
   {\left(1+\sqrt{(K-1)}\right)^2}\right)\nonumber\\
   &\approx& 2\gamma \sum_{i=1}^K\Delta_i\exp\left(-\frac
-    {2\Delta_j^2n} {K}\right)\mbox{ for large K}\nonumber\\
+    {2\Delta_i^2n} {K}\right)\mbox{ for large K}\nonumber\\
 \end{eqnarray}
@@
 number of pulls so far. \end{dfn}
-This scheme also exhibits a superpolynomially decreasing simple regret:
+This scheme also exhibits a super-polynomially decreasing simple regret:
 \begin{thm}  For any $0<\pi<1$  and $\gamma>1$ there exists $N$ such that for
@@
 \begin{equation}
-\IE r_{ucb\sqrt{\cdot}} \le \gamma\sum_{i=1}^K\Delta_i\exp\left(-2\alpha\sqrt{n}\right)
+\IE r_{ucb\sqrt{\cdot}} \le
+2\gamma\sum_{i=1}^K\Delta_i\exp\left(-\frac {\alpha\sqrt{n}} 2\right)
 \end{equation}
@@
 
-\begin{proof}[Proof outline:] TODO
+\begin{proof}[Proof outline:] Bound the probability $P_i$ that a
+  non-optimal arm $i$ is chosen. Split the interval $[\mu_i, \mu_*]$
+  at $\mu_i+\frac {\Delta_i} 2$. By Chernoff-Hoeffding bound:
+\begin{eqnarray}
+P_i&\le&Pr\left[\overline X_i>\mu_i+\frac {\Delta_i} 2\right]+\Pr\left[\overline X_*<\mu_*-\frac {\Delta_i}
+  2\right]\nonumber \\
+   &\le&\exp\left(-\frac {\Delta_i^2n_i} 2\right)+\exp\left(-\frac {\Delta_i^2n_*} 2\right)
+\end{eqnarray}
+Observe that, in probability, $n_i\to \frac {\alpha \sqrt n}
+{\Delta_i^2}$, $n_i\le n_*$ as $n\to\infty$. Conclude that for
+any $0<\pi<1$, $\gamma>1$ there exists $N$ such that for any $n>N$ and
+all non-optimal arms $i$:
+\begin{equation}
+P_i \le 2\gamma \exp\left(-\frac {\alpha \sqrt n} 2\right)
+\label{eqn:usbsqrt-probbound}
+\end{equation}
+Substitute (\ref{eqn:usbsqrt-probbound}) into
+(\ref{eqn:simple-regret}) and obtain
+\begin{equation}
+\IE r_{ucb\sqrt{\cdot}} \le 2\gamma\sum_{i=1}^K\Delta_i\exp\left(-\frac {\alpha\sqrt{n}} 2\right)
+\end{equation}
 \end{proof}

### 2012-01-22 paper/srcr/paper.tex
small changes

@@
 
-\begin{dfn} Scheme $\mathbf{UCB(\alpha)}$ repeatedly pulls arm $i$ that maximizes 
+\begin{dfn} Scheme $\mathbf{UCB(\alpha)}$ pulls arm $i$ that maximizes 
 upper confidence bound $b_i$ on the reward:
@@
 This sampling scheme exhibits an exponentially decreasing simple regret:
-\begin{thm} For any $0<\pi<10$  and $\gamma>1$ there exists $N$ such that for
+\begin{thm} For any $0<\eta<1$  and $\gamma>1$ there exists $N$ such that for
   any number of samples $n>N$ the simple regret of the  $\varepsilon$-greedy
@@
 \end{equation}
-with probability at least $1-\pi$.
+with probability at least $1-\eta$.
 \end{thm}
@@
 {n(1-\varepsilon)} {K-1}$ as $n\rightarrow \infty$. Conclude that for
-any $0<\pi<1$, $\gamma>1$ there exists $N$ such that for any $n>N$ and
+any $0<\eta<1$, $\gamma>1$ there exists $N$ such that for any $n>N$ and
 all non-optimal arms $i$:
@@
 This scheme also exhibits a super-polynomially decreasing simple regret:
-\begin{thm}  For any $0<\pi<1$  and $\gamma>1$ there exists $N$ such that for
+\begin{thm}  For any $0<\eta<1$  and $\gamma>1$ there exists $N$ such that for
   any number of samples $n>N$ the simple regret of the  UCB$_{\sqrt{\cdot}}$($\alpha$)
@@
 {\Delta_i^2}$, $n_i\le n_*$ as $n\to\infty$. Conclude that for
-any $0<\pi<1$, $\gamma>1$ there exists $N$ such that for any $n>N$ and
+any $0<\eta<1$, $\gamma>1$ there exists $N$ such that for any $n>N$ and
 all non-optimal arms $i$:
@@
 and compares the expected reward, unknown to the algorithm during the
-sampling, to the expected reward of the best arm. For such problem
-instances, the MCTS sampling schemes coincide with the selection
-algorithms employed at the first step of a rollout.
+sampling, to the expected reward of the best arm.
 
@@
 vs. the number of samples, averaged over $10^4$ experiments for
-randomly generated instances of 32 arms.
-
-For smaller numbers of samples, $\frac 1 2$-greedy achieves the best
+randomly generated instances of 32 arms. For smaller numbers of
+samples, $\frac 1 2$-greedy achieves the best
 performance; for larger numbers of samples, UCB$_{\sqrt{\cdot}}$
@@
   \end{minipage}
+  \caption{MCTS in random trees}
   \label{fig:mcts-regret}
-  \caption{MCTS: a path to the best arm}
 \end{figure}
@@
 
-Figures~\ref{fig:sailing-cost-vs-nsamples}--\ref{fig:sailing-rcq-vs-factor}
+Figures~\ref{fig:sailing-cost-vs-nsamples}--\ref{fig:sailing-lake-size}
 show results of experiments on the sailing
@@
 
-Figure~\ref{fig:sailing-rcq-vs-factor} compares UCB$_{\sqrt{\cdot}}$+UCT with UCT with a
-different exploration factor at the root (UCB+UCT). $\alpha$ for the rest of
-the steps was chosen to ensure the best performance from earlier
-experiments on the domain. Both algorithms exhibit similar dependency,
-but as the number of samples grows, UCB$_{\sqrt{\cdot}}$+UCT achieves smaller average
-regrets and is less sensitive to the choice of the value for $\alpha$ at
-the first step.
+\subsection{VOI-aware MCTS}
+
+Finally, the VOI-aware sampling scheme was empirically compared to
+other sampling schemes (UCT, $\frac 1 2$-greedy+UCT,
+UCT$_{\sqrt{\cdot}}$+UCT). Again, the experiments were performed on
+randomly generated trees with structure shown in
+Figure~\ref{fig:mcts-regret}.a. Figure~\ref{fig:mcts-regret-voi} shows
+the results for 32 arms. VOI+UCT, the scheme based on a VOI estimate,
+outperforms all other sampling schemes in this example. 
 \begin{figure}[h!]
   \centering
-  \includegraphics[scale=0.45]{rcq-size=6-nsamples=397.pdf}\\
-  b. 397 rollouts\\
-  \vspace{1em}
-  \includegraphics[scale=0.45]{rcq-size=6-nsamples=1585.pdf}\\
-  d. 1585 rollouts\\
-  \caption{The sailing domain, UCB vs. UCB$_{\sqrt{\cdot}}$, $6\times 6$ lake}
-  \label{fig:sailing-rcq-vs-factor}
+  \includegraphics[scale=0.45]{tree-identity-k=32-uqb=8+voi.pdf}\\
+  \caption{MCTS in random trees, VOI+UCT vs. other schemes.}
+  \label{fig:mcts-regret-voi}
 \end{figure}

### 2012-01-23 paper/srcr/paper.tex
eyals input, my corrections

@@
 
+%\title{MCTS Based on Simple Regret}
+%\author {David Tolpin, Solomon Eyal Shimony \\
+%Department of Computer Science, \\
+%Ben-Gurion University of the Negev, Beer Sheva, Israel \\
+%\{tolpin,shimony\}@cs.bgu.ac.il}
+
 \title{MCTS Based on Simple Regret}
-\author {David Tolpin, Solomon Eyal Shimony \\
-Department of Computer Science, \\
-Ben-Gurion University of the Negev, Beer Sheva, Israel \\
-\{tolpin,shimony\}@cs.bgu.ac.il}
+\author {blinded}
+
 \begin{document}
@@
 exists, applying it to MCTS is non-trivial, as typical myopic
-assumptions fail. Lacking a working VOI theory for MCTS, we
+assumptions fail. Lacking a complete working VOI theory for MCTS, we
 nevertheless propose a sampling scheme that is ``aware'' of VOI,
@@
 
-The core issue is that in search (especially for adversarial search
-and for ``games against nature'' - optimizing behavior under
-uncertainty) the goal is typically to either find a good (or optimal)
+The core issue is that in adversarial search
+and search in ``games against nature'' --- optimizing behavior under
+uncertainty, the goal is typically to either find a good (or optimal)
 strategy, or even to find the best first action of such a policy. Once
-such an action is discovered, it is not beneficial to further sample
+such an action is discovered, it is usually not beneficial to further sample
 that action, ``exploitation'' is thus meaningless for search
@@
 
-\section{Related Work}
+\section{Background and Related Work}
 \label{sec:related-work}
 
+Monte-Carlo tree search was initially suggested as a scheme for
+finding approximately optimal policies for Markov Decision Processes (MDP).
+An MDP is defined by the set of states $S$, the set of actions $A$, the transition table
+$T(s, a, s')$, the reward table $R(s, a, s')$, the initial state $s$
+and the goal state $t$: $(S, A, T, R, s, t)$ \cite{Russell.aima}.
+Several MCTS schemes explore an MDP by performing \emph{rollouts}---trajectories from the
+current state to a state in which a termination condition is satisfied
+(either the goal state, or a cutoff state for which the reward is
+evaluated approximately). The UCB algorithm (that attempts to minimize
+the cumulative regret) \cite{Auer.ucb} had been extended into the tree
+search sampling scheme known as UCT \cite{Kocsis.uct}, and shown to outperform
+many state of the art search algorithms in both MDP and adversarial search \cite{Eyerich.ctp}. %% more refs
+
 Efficient algorithms for Multi-Armed Bandits based on
-distribution-independent bounds, in particular UCB1, are introduced in
-\cite{Auer.ucb}. The UCT algorithm, an extension of UCB1 to
+distribution-independent bounds, in particular UCB1, or simply UCB, (which is
+asymptotically optimal in minimizing the cumulative regret)  were introduced in
+\cite{Auer.ucb}. The UCT algorithm, an extension of UCB to
 Monte-Carlo Tree Search is described in \cite{Kocsis.uct}. Pure
-exploration in Multi-armed bandits is explored in
-\cite{Bubeck.pure}. On the one hand, the paper proves certain upper
-and lower bounds for UCB1 and uniform sampling, showing that an upper
+exploration in Multi-armed bandits is examined in
+\cite{Bubeck.pure}. On the one hand, the \cite{Bubeck.pure} prove certain upper
+and lower bounds for UCB and uniform sampling, showing that an upper
 bound on the simple regret is exponential in the number of samples for
-uniform sampling, while only polynomial for UCB1. On the other hand,
-empirical performance of UCB1 appears to be better than of uniform
+uniform sampling, while only polynomial for UCB. On the other hand,
+empirical performance of UCB appears to be better than that of uniform
 sampling.
 
-The principles of bounded rationality appear in
-\cite{Horvitz.reasoningabout}. \cite{Russell.right} provided a formal
+A completely different direction of controlling sampling can use the
+principles of bounded rationality \cite{Horvitz.reasoningabout}
+and meta-reasoning -  \cite{Russell.right} provided a formal
 description of rational metareasoning and case studies of applications
-in several problem domains. One obstacle to straightforward
-application of the principles of rational metareasoning to Monte-Carlo
-sampling is the metagreedy assumption, according to which samples must
-be selected as though at most one sample can be taken before an action
-is chosen. In Monte-Carlo sampling, the value of information of any
-single sample in a given search state is often zero, so a different
-approximating assumption must be used instead.
+in several problem domains. Using the principles of meta-reasoning,
+one can ideally find an ``optimal'' sampling scheme, to be used
+for selecting what to sample, both at the root node \cite{HayRussell.MCTS} and elsewhere.
+However, this task is daunting for the following reasons:
+\begin{itemize}
+\item Defining the cost of a sample is not easy, and even just
+  using time-cost as an approximation, we get an intractable
+  meta-reasoning problem.
+\item To handle the complexity, one must use simplifying assumptions.
+      Applying the standard meta-reasoning myopic assumption, according to 
+    which samples would be selected as though at most one sample
+   can be taken before an action is chosen, we run into serious problems. Even the basic
+  selection problem \cite{TolpinShimony.blinkered} exhibits a
+  non-concave utility function and results in premature stopping of the
+  standard myopic algorithms. This is due to the fact that the value of information of
+  a single measurement (analogous to a sample in MCTS) is frequently
+  less than its time-cost, even though this is not true for multiple
+  measurements.  When applying the selection problem to MCTS, the
+  situation is exacerbated.  The utility of an action is usually
+  bounded, and thus in many cases a single sample may be insufficient
+  to change the current best action, \emph{regardless} of its
+  outcome. As a result, we frequently get a \emph{zero} ``myopic''
+  value of information for a single sample.
+\item Rational meta-reasoning requires a known distribution model, which may be
+  difficult to obtain.
+\end{itemize}
+
+As the above ultimate goal is extremely difficult
+to achieve, and even harder to analyze, we introduce
+simple schemes more amenable to analysis, loosely based on the meta-reasoning
+concept of value of information, and compare
+them to UCB (on sets) and UCT (in trees).
 
@@
 In the Multi-armed Bandit problem \cite{Vermorel.bandits} we have a set
-of $K$ arms. Each arm can be pulled multiple times. When the $i$th arm
+of $K$ arms. Each arm can be pulled multiple times (sometimes a cost is
+associated with each pulling action). When the $i$th arm
 is pulled, a random reward $X_i$ from an unknown stationary
-distribution is returned.  The reward is bounded between 0 and 1.
+distribution is encountered.  The reward is bounded between 0 and 1.
+In the cumulative setting, all encountered rewards are collected
+by the agent. In the simple (selection) setting, the agent gets to
+collect only the reward of the last pull.
 
@@
 Strategies that minimize the simple regret are called pure exploration
-strategies \cite{Bubeck.pure}. Such strategies are used to select the
-best arm, and, by extension, the best action in MCTS. MCTS is used to
-solve Markov Decision Processes (MDP) approximately. An MDP is defined
-by the set of states $S$, the set of actions $A$, the transition table
-$T(s, a, s')$, the reward table $R(s, a, s')$, the initial state $s$
-and the goal state $t$: $(S, A, T, R, s, t)$ \cite{Russell.aima}.
-MCTS explores an MDP by \emph{rollouts}---trajectories from the
-current state to a state in which a termination condition is satisfied
-(either the goal state, or a cutoff state for which the reward is
-evaluated approximately). The UCB algorithm (that attempts to minimize
-the cumulative regret) \cite{Auer.ucb} had been extended into the tree
-search sampling scheme known as UCT \cite{Kocsis.uct}.
-
-UCT performance can be improved by combining UCB with a sampling
-scheme that minimizes the simple regret of selecting an action at the
-current root node. Indeed, the algorithm must select an action with
-the minimum regret \emph{given the assumption that after performing the selected action
-the algorithm performs optimally}, which corresponds to maximizing the
-value of partial information \cite{Russell.aima}. Therefore, an
-improved allocation scheme would
-\begin{itemize}
-\item maximize the value of partial
-information by sampling actions to
-minimize the \textbf{simple} regret
-of the selection at the current root node, and
-\item as the goal of sampling in deeper tree nodes is estimating the value of
-a node, rather than selection, it makes sense to
-minimize the \textbf{cumulative} regret of the rollouts from the second
-step onwards. 
-\end{itemize}
-Ultimately, an ``optimal'' sampling in the meta-reasoning sense should
-be used at the first step in the above scheme. Nevertheless, this task
-is daunting for the following reasons:
-\begin{itemize}
-\item Defining the cost of a sample is not easy, and even if we simply
-  use time-cost as an approximation, we get an intractable
-  meta-reasoning problem.
-\item Applying the standard myopic and subtree independence
-  assumptions, we run into serious problems. Even in the standard
-  selection problem \cite{TolpinShimony.blinkered}, we get a
-  non-concave utility function and premature stopping of the
-  algorithm. This is due to the fact that the value of information of
-  a single measurement (analogous to a sample in MCTS) is frequently
-  less than its time-cost, even though this is not true for multiple
-  measurements.  When applying the selection problem to MCTS, the
-  situation is exacerbated.  The utility of an action is usually
-  bounded, and thus in many cases a single sample may be insufficient
-  to change the current best action, \emph{regardless} of its
-  outcome. As a result, we frequently get a \emph{zero} ``myopic''
-  value of information for a single sample.
-\end{itemize}
-As the ultimate goal is extremely difficult
-to achieve, and even harder to analyze, we introduce
-simple schemes more amenable to analysis, and compare
-them to UCB (on sets) and UCT (in trees).
+strategies \cite{Bubeck.pure}.
 
@@
 \begin{equation}
-\IE r_{\varepsilon\mbox{-}greedy}\le 2\gamma \sum_{i=1}^K\Delta_i\exp\left(-\frac {2\Delta_j^2n\varepsilon}
+\IE r_{\varepsilon\mbox{-}greedy}\le 2\gamma \sum_{i=1}^K\Delta_i\exp\left(\frac {-2\Delta_j^2n\varepsilon}
   {\left(1+\sqrt{\frac {(K-1)\varepsilon}
@@
 \begin{equation}
-P_i \le \gamma \left(\exp\left(-\frac {2\delta_i^2n(1-\varepsilon)}{K-1}\right)
+P_i \le \gamma \left(\exp\left(\frac {-2\delta_i^2n(1-\varepsilon)}{K-1}\right)
 +\exp\left(-2(\Delta_i-\delta_i)^2n\varepsilon\right)\right)
@@
 \begin{equation}
-\IE r_{\varepsilon\mbox{-}greedy}\le 2\gamma \sum_{i=1}^K\Delta_i\exp\left(-\frac {2\Delta_i^2n\varepsilon}
+\IE r_{\varepsilon\mbox{-}greedy}\le 2\gamma \sum_{i=1}^K\Delta_i\exp\left(\frac {-2\Delta_i^2n\varepsilon}
   {\left(1+\sqrt{\frac {(K-1)\varepsilon}
@@
 \begin{eqnarray}
-\IE r_{\frac 1 2\mbox{-}greedy}&\le& 2\gamma \sum_{i=1}^K\Delta_i\exp\left(-\frac {2\Delta_i^2n}
-  {\left(1+\sqrt{(K-1)}\right)^2}\right)\nonumber\\
-  &\approx& 2\gamma \sum_{i=1}^K\Delta_i\exp\left(-\frac
-    {2\Delta_i^2n} {K}\right)\mbox{ for large K}\nonumber\\
+\IE r_{\frac 1 2\mbox{-}greedy}&\le& 2\gamma \sum_{i=1}^K\Delta_i\exp\left(\frac {-2\Delta_i^2n}
+  {\left(1+\sqrt{K-1}\right)^2}\right)\\
+  &\approx& 2\gamma \sum_{i=1}^K\Delta_i\exp\left(\frac
+    {-2\Delta_i^2n} {K}\right)\mbox{ for }K\gg 1\nonumber
 \end{eqnarray}
@@
 UCT \cite{Kocsis.uct} is a generalization of UCB for MCTS.  UCT
-applies UCB at each step of a rollout.  We suggest an improvement on
-UCT, which combines different sampling schemes on the first step and
-during the rest of a rollout:
+applies UCB at each step of a rollout.  
+At the root node, the goal of sampling in MCTS is usually finding the
+first action to perform. Search is re-started, either from scratch or using some
+previously collected information, after observing the actual outcome (in MDPs) or
+the opponent's move (in adversarial games). As in many domains neither can be
+predicted with any degree of certainty, there is usually little utility in performing additional
+sampling once one action is shown to be the best choice with high confidence.
+As the choice of move is unlikely to change, the value of information of additional samples
+is low. Therefore, one should be able to do better than UCT by optimizing {\em simple regret},
+rather than {\em cumulative regret},
+at the root node.
+
+For nodes deeper in the search tree the situation is not as clear, but here the task appears different.
+In order to support an optimal move choice at the root, a precise {\em value} of the search
+tree nodes is beneficial. For these internal nodes optimizing simple regret is not the answer, and 
+cumulative regret optimization seems more reasonable. Lacking a complete meta-reasoning for sampling,
+which would indicate the optimal way to sample both root nodes and internal nodes,
+our suggested improvement to UCT thus combines different sampling schemes on the first step and
+during the rest of each rollout:
 \begin{dfn}
@@
 
-\begin{algorithm}[t]
+\begin{algorithm}[h!]
 \caption{Two-stage Monte-Carlo tree search sampling}
@@
 
-A further improvement can be achieved by computing or estimating the
-value of information (VOI) of the rollouts and choosing a rollout that
-maximizes the VOI. VOI of a rollout can be computed when the sample
-distribution of an action is known up to the parameters, such as the
-normal distribution with an unknown mean and/or
-variance. Alternatively, the value of information can be estimated
-from the set of samples, and the need to assume a particular shape of
-the distribution can be lifted. In one realization of the latter
-approach the VOI of performing an action is estimated as an
-\emph{upper bound on the value of perfect information about an arm,
-divided by the number of pulls of the arm}.
+Further improvement can be achieved by computing or estimating the
+value of information (VOI) of the rollouts and choosing rollouts that
+maximize the VOI. This is done by maintaining a current best move at the root,
+and finding the expected gain from finding another move to be better than the
+current best \cite{Russell.aima}.
+However, as indicated above, actually computing the VOI is infeasible.
+Instead we suggest the following scheme based on the following
+features of value of information:
+\begin{enumerate}
+\item An estimate of the probability that one or more rollouts will make another action
+appear better than the current best.
+\item An estimate of the gain that may be incurred if such a change occurs.
+\end{enumerate}
+
+If the distribution of results generated by the rollouts were known, the
+above features could be easily computed. However, this is not the case for
+most MCTS applications. Therefore, we estimate bounds on the feature values from
+the current set of samples, based on the myopic assumption that the algorithm will only
+sample one of the actions, and use these bounds as the feature values, to get:
 \begin{eqnarray}
@@
 \exp\left(-2(\overline X_\alpha - \overline X_i)^2 n_i\right),\; i\ne\alpha\nonumber\\
-\mbox{where }\alpha&=&\arg\max_i \overline X_i\nonumber\\
-             \beta&=&\arg\max_{i,\,i\ne\alpha} \overline X_i\nonumber
+\mbox{where }&&\alpha=\arg\max_i \overline X_i,\quad
+             \beta=\arg\max_{i,\,i\ne\alpha} \overline X_i\nonumber
 \end{eqnarray}
-Early experiments with this approach demonstrated
-a significantly lower simple regret on a wide range of problems.
+with $VOI_{\alpha}$ being (approximate) value for sampling the current best action,
+and $VOI_i$ is the (approximate) value for sampling some other action $i$.
+
+These equations were derived as follows. The gain from switching from the current best action $\alpha$ to another
+action can be bounded:  by  the current expectation of the value the current second-best action
+for the case where we sample only $\alpha$, and by 1 (the maximum reward) minus the current expectation
+of $\alpha $ when sampling any other action. 
+The probability that another action be found
+best can be bounded by an exponential function of the difference in expectations when the true value
+of the actions becomes known. But the effect of each individual sample on the sample mean
+is inversely proportional to the current number of samples, hence the current number of samples (plus one
+in order to handle the initial case of no previous samples) in the denominator.
+Early experiments with this approach demonstrate a significantly lower simple regret.
 
@@
     \centering
-    \includegraphics[scale=1.0]{onelevel-tree.pdf}\\
+    \includegraphics[scale=0.9]{onelevel-tree.pdf}\\
     a. search tree
@@
     \centering
-    \includegraphics[scale=0.5]{flat-trilevel-k=64-uqb=8.pdf}\\
+    \includegraphics[scale=0.45]{flat-trilevel-k=64-uqb=8.pdf}\\
     b. regret vs. number of samples
@@
     \centering
-    \includegraphics[scale=0.8]{twolevel-tree.pdf}\\
+    \includegraphics[scale=0.7]{twolevel-tree.pdf}\\
     a. search tree
-    \vspace{1em}
+    \vspace{0.5em}
   \end{minipage}
@@
     \centering
-    \includegraphics[scale=0.5]{tree-identity-k=16-uqb=8.pdf}\\ 
+    \includegraphics[scale=0.45]{tree-identity-k=16-uqb=8.pdf}\\ 
     b. 16 arms
-    \vspace{1em}
+    \vspace{0.5em}
   \end{minipage}
@@
     \centering
-    \includegraphics[scale=0.5]{tree-identity-k=64-uqb=8.pdf} \\
+    \includegraphics[scale=0.45]{tree-identity-k=64-uqb=8.pdf} \\
     c. 64 arms
@@
     \centering
-    \includegraphics[scale=0.45]{costs-size=6-group=median.pdf}\\
+    \includegraphics[scale=0.425]{costs-size=6-group=median.pdf}\\
     a. median cost
-    \vspace{1em}
+    \vspace{0.5em}
   \end{minipage}
@@
     \centering
-    \includegraphics[scale=0.45]{costs-size=6-group=minimum.pdf}\\
+    \includegraphics[scale=0.425]{costs-size=6-group=minimum.pdf}\\
     b. minimum cost
@@
   \centering
-  \includegraphics[scale=0.45]{costs-size=6-nsamples=397.pdf}\\
+  \includegraphics[scale=0.425]{costs-size=6-nsamples=397.pdf}\\
   a. 397 rollouts\\
-  \vspace{1em}
-  \includegraphics[scale=0.45]{costs-size=6-nsamples=1585.pdf}\\
+  \vspace{0.5em}
+  \includegraphics[scale=0.425]{costs-size=6-nsamples=1585.pdf}\\
   c. 1585 rollouts
@@
    \centering
-   \includegraphics[scale=0.45]{costs-size=6-nsamples=397.pdf}\\
+   \includegraphics[scale=0.425]{costs-size=6-nsamples=397.pdf}\\
    b. $6\times 6$ lake \\
-   \vspace{1em}
-   \includegraphics[scale=0.45]{costs-size=10-nsamples=397.pdf}\\
+   \vspace{0.5em}
+   \includegraphics[scale=0.425]{costs-size=10-nsamples=397.pdf}\\
    c. $10\times 10$ lake
@@
 
+UCT-based Monte-Carlo tree search has been shown to be very effective
+for finding good actions in both MDPs and adversarial games.
+Further improvement of the sampling scheme is thus of interest in
+numerous search applications. We argue that although UCT is already very efficient,
+one can do better if the sampling scheme is considered from a meta-reasoning
+perspective of value of information (VOI).
+
 The MCTS SR+CR scheme presented in the paper differs
-from UCT at the first step of the rollout, when the `simple' selection
+from UCT mainly in the first step of the rollout, when the `simple' selection
 regret is minimized instead of the cumulative regret. Both the
@@
 
-The improvement is inspired by the notion of value of information
-(VOI), but VOI is used implicitly in the analysis of the algorithm,
+Although SR+CR is inspired  by the notion of VOI,
+the VOI is used there implicitly in the analysis of the algorithm,
 rather than computed or learned explicitly in order to plan the
-rollouts. Preliminary results on VOI-aware sampling suggest that a
-further improvement in the sampling performance can be
-achieved. However, application of the theory of rational metareasoning
+rollouts. Ideally, using VOI to control sampling ab-initio should do even better,
+but the theory for doing that is still not up to speed. Instead we suggest
+a  ``VOI-aware'' sampling scheme based on crude probability and value estimates,
+which despite its simplicity already shows a marked improvement in minimizing regret.
+However, application of the theory of rational metareasoning
 to Monte Carlo Tree Search is an open problem \cite{HayRussell.MCTS},
@@
 
-The research is partially supported by Israel
-Science Foundation grant 305/09, by the Lynne and William Frankel
-Center for Computer Sciences, and by the Paul Ivanier Center for
-Robotics Research and Production Management.
+Blinded.
+
+%The research is partially supported by Israel
+%Science Foundation grant 305/09, by the Lynne and William Frankel
+%Center for Computer Sciences, and by the Paul Ivanier Center for
+%Robotics Research and Production Management.
 

### 2012-01-24 paper/srcr/paper.tex
some corrections

@@
 \title{MCTS Based on Simple Regret}
-\author {blinded}
+\author {Submission 198}
 
@@
 in games and Markov decision processes, is based on UCB, a sampling
-policy for the Multi-armed Bandit problem (MAB) that asymptotically
+policy for the Multi-armed Bandit problem (MAB) that 
 minimizes the cumulative regret.  However, search differs from MAB in
@@
 Monte-Carlo tree search was initially suggested as a scheme for
-finding approximately optimal policies for Markov Decision Processes (MDP).
-An MDP is defined by the set of states $S$, the set of actions $A$, the transition table
-$T(s, a, s')$, the reward table $R(s, a, s')$, the initial state $s$
-and the goal state $t$: $(S, A, T, R, s, t)$ \cite{Russell.aima}.
-Several MCTS schemes explore an MDP by performing \emph{rollouts}---trajectories from the
-current state to a state in which a termination condition is satisfied
-(either the goal state, or a cutoff state for which the reward is
-evaluated approximately). The UCB algorithm (that attempts to minimize
-the cumulative regret) \cite{Auer.ucb} had been extended into the tree
-search sampling scheme known as UCT \cite{Kocsis.uct}, and shown to outperform
-many state of the art search algorithms in both MDP and adversarial search \cite{Eyerich.ctp}. %% more refs
+finding approximately optimal policies for Markov Decision Processes
+(MDP).  An MDP is defined by the set of states $S$, the set of actions
+$A$, the transition table $T(s, a, s')$, the reward table $R(s, a,
+s')$, the initial state $s$ and the goal state $t$: $(S, A, T, R, s,
+t)$ \cite{Russell.aima}.  Several MCTS schemes explore an MDP by
+performing \emph{rollouts}---trajectories from the current state to a
+state in which a termination condition is satisfied (either the goal
+state, or a cutoff state for which the reward is evaluated
+approximately).
 
 Efficient algorithms for Multi-Armed Bandits based on
-distribution-independent bounds, in particular UCB1, or simply UCB, (which is
-asymptotically optimal in minimizing the cumulative regret)  were introduced in
-\cite{Auer.ucb}. The UCT algorithm, an extension of UCB to
-Monte-Carlo Tree Search is described in \cite{Kocsis.uct}. Pure
-exploration in Multi-armed bandits is examined in
-\cite{Bubeck.pure}. On the one hand, the \cite{Bubeck.pure} prove certain upper
-and lower bounds for UCB and uniform sampling, showing that an upper
-bound on the simple regret is exponential in the number of samples for
-uniform sampling, while only polynomial for UCB. On the other hand,
-empirical performance of UCB appears to be better than that of uniform
-sampling.
+distribution-independent bounds, and, in particular, UCB (which is
+close to being optimal in minimizing the cumulative regret), were
+introduced in \cite{Auer.ucb}. The UCT algorithm, an extension of UCB
+to Monte-Carlo Tree Search is described in \cite{Kocsis.uct}, and
+shown to outperform many state of the art search algorithms in both
+MDP and adversarial search \cite{Eyerich.ctp}. Pure exploration in
+Multi-armed bandits is examined in \cite{Bubeck.pure}. On the one
+hand, the \cite{Bubeck.pure} prove certain upper and lower bounds for
+UCB and uniform sampling, showing that an upper bound on the simple
+regret is exponential in the number of samples for uniform sampling,
+while only polynomial for UCB. On the other hand, empirical
+performance of UCB appears to be better than that of uniform sampling.
 
@@
 
-\begin{dfn} Scheme $\mathbf{UCB(\alpha)}$ pulls arm $i$ that maximizes 
+\begin{dfn} Scheme $\mathbf{UCB(c)}$ pulls arm $i$ that maximizes 
 upper confidence bound $b_i$ on the reward:
 \begin{equation}
-b_i=\overline X_i+\sqrt {\frac {\alpha \ln n} {n_i}}
+b_i=\overline X_i+\sqrt {\frac {c \ln n} {n_i}}
 \label{eqn:ucb}
@@
 number of pulls so far. \end{dfn} The best known upper bound on the simple
-regret of UCB($\alpha$) is polynomially decreasing in the number of samples
+regret of UCB($c$) is polynomially decreasing in the number of samples
 (see \cite{Bubeck.pure}, Theorems~2,3).
@@
 \cite{Bubeck.pure}, Proposition~1). However, empirically
-UCB($\alpha$) yields a lower simple regret than uniform
+UCB($c$) yields a lower simple regret than uniform
 sampling.
@@
 example, $\sqrt\cdot$.
-\begin{dfn} Scheme $\mathbf{UCB_{\sqrt{\cdot}}(\alpha)}$ repeatedly pulls arm $i$ that
+\begin{dfn} Scheme $\mathbf{UCB_{\sqrt{\cdot}}(c)}$ repeatedly pulls arm $i$ that
 maximizes $b_i$:
 \begin{equation}
-b_i=\overline X_i+\sqrt {\frac {\alpha \sqrt n} {n_i}}
+b_i=\overline X_i+\sqrt {\frac {c \sqrt n} {n_i}}
 \end{equation}
@@
 \begin{thm}  For any $0<\eta<1$  and $\gamma>1$ there exists $N$ such that for
-  any number of samples $n>N$ the simple regret of the  UCB$_{\sqrt{\cdot}}$($\alpha$)
+  any number of samples $n>N$ the simple regret of the  UCB$_{\sqrt{\cdot}}$($c$)
 sampling scheme is  bounded from above as
@@
 \IE r_{ucb\sqrt{\cdot}} \le
-2\gamma\sum_{i=1}^K\Delta_i\exp\left(-\frac {\alpha\sqrt{n}} 2\right)
+2\gamma\sum_{i=1}^K\Delta_i\exp\left(-\frac {c\sqrt{n}} 2\right)
 \end{equation}
@@
 \end{eqnarray}
-Observe that, in probability, $n_i\to \frac {\alpha \sqrt n}
+Observe that, in probability, $n_i\to \frac {c \sqrt n}
 {\Delta_i^2}$, $n_i\le n_*$ as $n\to\infty$. Conclude that for
@@
 \begin{equation}
-P_i \le 2\gamma \exp\left(-\frac {\alpha \sqrt n} 2\right)
+P_i \le 2\gamma \exp\left(-\frac {c \sqrt n} 2\right)
 \label{eqn:usbsqrt-probbound}
@@
 \begin{equation}
-\IE r_{ucb\sqrt{\cdot}} \le 2\gamma\sum_{i=1}^K\Delta_i\exp\left(-\frac {\alpha\sqrt{n}} 2\right)
+\IE r_{ucb\sqrt{\cdot}} \le 2\gamma\sum_{i=1}^K\Delta_i\exp\left(-\frac {c\sqrt{n}} 2\right)
 \end{equation}
@@
 UCT \cite{Kocsis.uct} is a generalization of UCB for MCTS.  UCT
-applies UCB at each step of a rollout.  
-At the root node, the goal of sampling in MCTS is usually finding the
-first action to perform. Search is re-started, either from scratch or using some
-previously collected information, after observing the actual outcome (in MDPs) or
-the opponent's move (in adversarial games). As in many domains neither can be
-predicted with any degree of certainty, there is usually little utility in performing additional
-sampling once one action is shown to be the best choice with high confidence.
-As the choice of move is unlikely to change, the value of information of additional samples
-is low. Therefore, one should be able to do better than UCT by optimizing {\em simple regret},
-rather than {\em cumulative regret},
-at the root node.
+applies UCB($c$) at each step of a rollout.  At the root node,
+the goal of sampling in MCTS is
+usually finding the first action to perform. Search is re-started,
+either from scratch or using some previously collected information,
+after observing the actual outcome (in MDPs) or the opponent's move
+(in adversarial games). As in many domains neither can be predicted
+with any degree of certainty, there is usually little utility in
+performing additional sampling once one action is shown to be the best
+choice with high confidence.  As the choice of move is unlikely to
+change, the value of information of additional samples is
+low. Therefore, one should be able to do better than UCT by optimizing
+{\em simple regret}, rather than {\em cumulative regret}, at the root
+node.
 
@@
 significantly less sensitive to the tuning of the exploration factor
-$\alpha$ of UCT, since the conflict between the need for a larger
-value of $\alpha$ on the first step (simple regret) and a smaller
+$c$ of UCB($c$), since the contradiction between the need for a larger
+value of $c$ on the first step (simple regret) and a smaller
 value for the rest of the rollout (cumulative regret)
 \cite{Bubeck.pure} is resolved. In fact, a sampling scheme that uses
-UCB at all steps but a larger value of $\alpha$ for the first step
+UCB($c$) at all steps but a larger value of $c$ for the first step
 than for the rest of the steps, outperforms UCT. The pseudocode of the
@@
     \centering
-    \includegraphics[scale=0.9]{onelevel-tree.pdf}\\
+    \includegraphics[scale=0.8]{onelevel-tree.pdf}\\
     a. search tree
@@
 vs. the number of samples, computed for a range of values of
-$\alpha$. Figure~\ref{fig:sailing-cost-vs-nsamples}.a shows the median
+$c$. Figure~\ref{fig:sailing-cost-vs-nsamples}.a shows the median
 cost, and Figure~\ref{fig:sailing-cost-vs-nsamples}.b --- the minimum
 costs. UCT is always worse than either $\frac 1 2$-greedy+UCT or
-UCB$_{\sqrt{\cdot}}$+UCT, and is sensitive to the value of $\alpha$: the median cost is
+UCB$_{\sqrt{\cdot}}$+UCT, and is sensitive to the value of $c$: the median cost is
 much higher than the minimum cost for UCT. For both $\frac 1 2$-greedy+UCT
@@
 UCT, and $\frac 1 2$-greedy+UCT is better than UCT expect for a small range of
-values of the exploration factor $\alpha$.
+values of the exploration factor.
 

### 2012-01-24 paper/srcr/paper.tex
corrections

@@
 
-We then observe that optimizing the sampling process is itself a
-meta-reasoning problem, a solution of which can use value of
+Optimizing the sampling process is itself a
+metareasoning problem, a solution of which can use value of
 information (VOI) techniques.  Although the theory of VOI for search
@@
 we argue that it is inappropriate for search. The problem is not that
-UCT does not work; rather, we argue that a simple reconsideration from basic
+UCT does not work; rather, a simple reconsideration from basic
 principles can result in schemes that outperform UCT.
@@
 
-We begin with introduction of several sampling schemes with better
-bounds for the simple regret on sets. We then extend the results to
-sampling in trees by combining the proposed sampling schemes on the
-first step of a rollout with UCT during the rest of the
-rollout. Finally, we empirically evaluate the performance of the
-proposed sampling schemes on sets of Bernoulli arms, in randomly
-generated trees, and on the sailing domain.
+We begin with background definitions and related work. 
+Some sampling schemes are introduced, and shown to have better than UCB
+bounds for the simple regret on sets, the first contribution of this paper. 
+The results are applied to sampling in trees by combining the proposed sampling schemes on the
+first step of a rollout with UCT for the rest of the
+rollout. An additional sampling scheme based on metareasoning 
+principles is also suggested, another contribution
+of this paper.
+Finally, the performance of the
+proposed sampling schemes is evaluated on sets of Bernoulli arms, in randomly
+generated 2-level trees, and on the sailing domain, showing where the proposed schemes show improved
+performance, the third contribution of this paper.
 
@@
 
-Efficient algorithms for Multi-Armed Bandits based on
-distribution-independent bounds, and, in particular, UCB (which is
-close to being optimal in minimizing the cumulative regret), were
-introduced in \cite{Auer.ucb}. The UCT algorithm, an extension of UCB
-to Monte-Carlo Tree Search is described in \cite{Kocsis.uct}, and
-shown to outperform many state of the art search algorithms in both
-MDP and adversarial search \cite{Eyerich.ctp}. Pure exploration in
-Multi-armed bandits is examined in \cite{Bubeck.pure}. On the one
-hand, the \cite{Bubeck.pure} prove certain upper and lower bounds for
-UCB and uniform sampling, showing that an upper bound on the simple
-regret is exponential in the number of samples for uniform sampling,
-while only polynomial for UCB. On the other hand, empirical
-performance of UCB appears to be better than that of uniform sampling.
-
-A completely different direction of controlling sampling can use the
-principles of bounded rationality \cite{Horvitz.reasoningabout}
-and meta-reasoning -  \cite{Russell.right} provided a formal
-description of rational metareasoning and case studies of applications
-in several problem domains. Using the principles of meta-reasoning,
-one can ideally find an ``optimal'' sampling scheme, to be used
-for selecting what to sample, both at the root node \cite{HayRussell.MCTS} and elsewhere.
-However, this task is daunting for the following reasons:
-\begin{itemize}
-\item Defining the cost of a sample is not easy, and even just
-  using time-cost as an approximation, we get an intractable
-  meta-reasoning problem.
-\item To handle the complexity, one must use simplifying assumptions.
-      Applying the standard meta-reasoning myopic assumption, according to 
-    which samples would be selected as though at most one sample
-   can be taken before an action is chosen, we run into serious problems. Even the basic
-  selection problem \cite{TolpinShimony.blinkered} exhibits a
-  non-concave utility function and results in premature stopping of the
-  standard myopic algorithms. This is due to the fact that the value of information of
-  a single measurement (analogous to a sample in MCTS) is frequently
-  less than its time-cost, even though this is not true for multiple
-  measurements.  When applying the selection problem to MCTS, the
-  situation is exacerbated.  The utility of an action is usually
-  bounded, and thus in many cases a single sample may be insufficient
-  to change the current best action, \emph{regardless} of its
-  outcome. As a result, we frequently get a \emph{zero} ``myopic''
-  value of information for a single sample.
-\item Rational meta-reasoning requires a known distribution model, which may be
-  difficult to obtain.
-\end{itemize}
-
-As the above ultimate goal is extremely difficult
-to achieve, and even harder to analyze, we introduce
-simple schemes more amenable to analysis, loosely based on the meta-reasoning
-concept of value of information, and compare
-them to UCB (on sets) and UCT (in trees).
-
-\section{Sampling Based on Simple Regret}
-\label{sec:results}
+\subsection{Multi-armed bandits and UCT}
 
 In the Multi-armed Bandit problem \cite{Vermorel.bandits} we have a set
-of $K$ arms. Each arm can be pulled multiple times (sometimes a cost is
+of $K$ arms (see Figure~\ref{fig:mab-simple-regret}.a). Each arm can
+be pulled multiple times (sometimes a cost is
 associated with each pulling action). When the $i$th arm
@@
 
-Strategies that minimize the simple regret are called pure exploration
-strategies \cite{Bubeck.pure}.
+%Strategies that minimize the simple regret are called pure exploration
+%strategies \cite{Bubeck.pure}.
 
-\subsection{Sampling on Sets}
-\label{sec:sampling-on-sets}
+Much of research literature on Multi-armed bandits centered
+on minimizing the cumulative regret. The
+UCB scheme (see below) was shown to be near-optimal in minimizing the cumulative regret \cite{Auer.ucb}.
 
@@
 
-We introduce here two sampling schemes with super-polynomially
+The UCT algorithm, an extension of UCB
+to Monte-Carlo Tree Search is described in \cite{Kocsis.uct}, and
+shown to outperform many state of the art search algorithms in both
+MDP and adversarial search \cite{Eyerich.ctp}. Pure exploration in
+Multi-armed bandits is examined in \cite{Bubeck.pure}. On the one
+hand, the \cite{Bubeck.pure} prove certain upper and lower bounds for
+UCB and uniform sampling, showing that an upper bound on the simple
+regret is exponential in the number of samples for uniform sampling,
+while only polynomial for UCB. On the other hand, empirical
+performance of UCB appears to be better than that of uniform sampling.
+
+\subsection{Metareasoning}
+
+A completely different direction of controlling sampling can use the
+principles of bounded rationality \cite{Horvitz.reasoningabout}
+and metareasoning ---  \cite{Russell.right} provided a formal
+description of rational metareasoning and case studies of applications
+in several problem domains. In search, under myopic and sub-tree independence assumptions,
+one maintains a current best move  $\alpha $ at the root, 
+and finds the expected gain from finding another move 
+$\beta $ to be better than the current best \cite{Russell.aima}. The ``cost'' of search
+actions can also be factored in.
+Ideally, an ``optimal'' sampling scheme, to be used
+for selecting what to sample, both at the root node \cite{HayRussell.MCTS} and elsewhere,
+can be developed using metareasoning.
+However, this task is daunting for the following reasons:
+\begin{itemize}
+\item Defining the cost of a sample is not trivial.
+\item The method is in general intractable, necessitating simplifying assumptions.
+ However, using the standard metareasoning myopic assumption,
+    where samples would be selected as though at most one sample
+   can be taken before an action is chosen, we run into serious problems. Even the basic
+  selection problem \cite{TolpinShimony.blinkered} exhibits a
+  non-concave utility function and results in premature stopping of the
+  standard myopic algorithms. This is due to the fact that the value of information of
+  a single measurement (analogous to a sample in MCTS) is frequently
+  less than its time-cost, even though this is not true for multiple
+  measurements. 
+% When applying the selection problem to MCTS, the
+%  situation is exacerbated.  The utility of an action is usually
+%  bounded, and thus in many cases a single sample may be insufficient
+%  to change the current best action, \emph{regardless} of its
+%  outcome. As a result, we frequently get a \emph{zero} ``myopic''
+%  value of information for a single sample.
+\item Rational metareasoning requires a known distribution model, which may be
+  difficult to obtain.
+\end{itemize}
+
+As the above ultimate goal is extremely difficult
+to achieve, we introduce
+simple schemes more amenable to analysis, loosely based on the metareasoning
+concept of value of information, and compare
+them to UCB (on sets) and UCT (in trees).
+
+\section{Sampling Based on Simple Regret}
+\label{sec:results}
+
+\subsection{Analysis of Sampling on Sets}
+\label{sec:sampling-on-sets}
+
+We examine two sampling schemes with super-polynomially
 decreasing upper bounds on the simple regret. The bounds
@@
 after observing the actual outcome (in MDPs) or the opponent's move
-(in adversarial games). As in many domains neither can be predicted
-with any degree of certainty, there is usually little utility in
-performing additional sampling once one action is shown to be the best
-choice with high confidence.  As the choice of move is unlikely to
-change, the value of information of additional samples is
-low. Therefore, one should be able to do better than UCT by optimizing
-{\em simple regret}, rather than {\em cumulative regret}, at the root
-node.
+(in adversarial games). Once one move is shown to be the best
+choice with high confidence, the value of information of additional
+samples of the best move is low. Therefore, one should be able to do
+better than UCT by optimizing {\em simple regret}, rather than {\em
+  cumulative regret}, at the root node.
 
@@
 tree nodes is beneficial. For these internal nodes optimizing simple regret is not the answer, and 
-cumulative regret optimization seems more reasonable. Lacking a complete meta-reasoning for sampling,
+cumulative regret optimization seems more reasonable. Lacking a complete metareasoning for sampling,
 which would indicate the optimal way to sample both root nodes and internal nodes,
@@
 
-Such two-stage sampling scheme outperforms UCT, and, additionally, is
+We will denote realizations of SR+CR as \emph{Alg}+UCT, where
+\emph{Alg} is the sampling scheme employed at the first step of a
+rollout (e.g. $\frac 1 2$-greedy+UCT). Such two-stage sampling scheme
+outperforms UCT, and, additionally, is
 significantly less sensitive to the tuning of the exploration factor
@@
 than for the rest of the steps, outperforms UCT. The pseudocode of the
-two-stage rollout is in Algorithm~\ref{alg:two-stage-mcts}.
+two-stage rollout is in
+Algorithm~\ref{alg:two-stage-mcts}: \textsc{FirstAction} selects the first
+step of a layout (line~\ref{alg:srcr-first-action}), and
+\textsc{NextAction} (line~\ref{alg:srcr-next-action}) selects steps during
+the rest of the layout (usually using UCB).
 
@@
   \Else
-    \If {depth=1}
-      \State action $\gets$ \Call{FirstAction}{node}
-    \Else
-      \State action $\gets$ \Call{NextAction}{node}
-    \EndIf
+    \State \textbf{if} depth=1 \textbf{then} action $\gets$ \Call{FirstAction}{node} \label{alg:srcr-first-action}   
+    \State \textbf{else} action $\gets$ \Call{NextAction}{node} \label{alg:srcr-next-action}
     \State next-node $\gets$ \Call{NextState}{node, action}
@@
 value of information (VOI) of the rollouts and choosing rollouts that
-maximize the VOI. This is done by maintaining a current best move at the root,
-and finding the expected gain from finding another move to be better than the
-current best \cite{Russell.aima}.
+maximize the VOI. 
 However, as indicated above, actually computing the VOI is infeasible.
@@
 \item An estimate of the probability that one or more rollouts will make another action
-appear better than the current best.
+appear better than the current best $\alpha $.
 \item An estimate of the gain that may be incurred if such a change occurs.
@@
 in order to handle the initial case of no previous samples) in the denominator.
+
+These VOI estimates are used in the ``VOI-aware'' sampling scheme as follows: sample the
+action that has maximum estimated VOI. We judged these estimates to be too crude
+to be used as ``stopping criteria'' that can be used to cut off sampling,
+leaving this issue for future research.
 Early experiments with this approach demonstrate a significantly lower simple regret.
@@
 Figure~\ref{fig:mab-simple-regret}.b shows the regret
-vs. the number of samples, averaged over $10^4$ experiments for
+vs. the number of samples, averaged over $10000$ experiments for
 randomly generated instances of 32 arms. For smaller numbers of
@@
 number of samples for trees with root degree 16
-(Figure~\ref{fig:mcts-regret}.b) and 64 (Figure~\ref{fig:mcts-regret}.c). The
-dependencies look differently from Multi-armed bandit instances, but
-the algorithms exhibit a similar relative performance: either $\frac 1
+(Figure~\ref{fig:mcts-regret}.b) and 64 (Figure~\ref{fig:mcts-regret}.c).
+The algorithms exhibit a similar relative performance: either $\frac 1
 2$-greedy+UCT or UCB$_{\sqrt{\cdot}}$+UCT
@@
   \includegraphics[scale=0.45]{tree-identity-k=32-uqb=8+voi.pdf}\\
-  \caption{MCTS in random trees, VOI+UCT vs. other schemes.}
+  \caption{MCTS in random trees, including VOI+UCT.}
   \label{fig:mcts-regret-voi}
@@
 numerous search applications. We argue that although UCT is already very efficient,
-one can do better if the sampling scheme is considered from a meta-reasoning
+one can do better if the sampling scheme is considered from a metareasoning
 perspective of value of information (VOI).
@@
 
-\section*{Acknowledgments}
-
-Blinded.
+%\section*{Acknowledgments}
 

### 2012-01-27 paper/srcr/paper.tex
Bayesian MCTS

--- 2012-01-23 paper/srcr/paper.tex
@@
 principles of bounded rationality \cite{Horvitz.reasoningabout}
-and meta-reasoning -  \cite{Russell.right} provided a formal
+and meta-reasoning ---  \cite{Russell.right} provided a formal
 description of rational metareasoning and case studies of applications

### 2012-01-27 paper/ecai2012/ecai2012.tex

\documentclass{ecai2012}
\usepackage{times}
\usepackage{graphicx}
\usepackage{latexsym}

%%\ecaisubmission   % inserts page numbers. Use only for submission of paper.
                  % Do NOT use for camera-ready version of paper.

\begin{document}

\title{Guidelines for Preparing a Paper for the \\
European Conference on Artificial Intelligence}

\author{Name1 Surname1 \and Name2 Surname2 \and Name3 Surname3\institute{University of Leipzig,
Germany, email: somename@informatik.uni-leipzig.de} }

\maketitle
\bibliographystyle{ecai2012}

\begin{abstract}
  The purpose of this paper is to show a contributor the required
  style for a paper for ECAI-2012 and PAIS-2012. The specifications
  for layout are described so that non-\LaTeX\ users can create their
  own style sheet to achieve the same layout. The source for the
  sample file is available for \LaTeX\ users. The PostScript and the
  PDF file is available for all. A Microsoft Word template is also available.
 The layout is identical to ECAI'10
  papers. The publisher (IOS Press) will insert a footer for each
  page.
\end{abstract}

\section{PAGE LIMIT}
The page limit for ECAI-2012 and PAIS-2012 long papers is \textbf{6
  (six)} pages in the required format. The page limit for short
submissions is 2 pages. This is a strict limit. Overlength papers will
not be reviewed.

\section{GENERAL SPECIFICATIONS}
The following details should allow contributors to set up the general
page description for their paper:

\begin{enumerate}
\item The paper is set in two columns each 20.5 picas (86 mm) wide
  with a column separator of 1.5 picas (6 mm).

\item The typeface is Times Modern Roman.

\item The body text size is 9 point (3.15 mm) on a body of 11
point (3.85 mm) (i.e., 61 lines of text).

\item The effective text height for each page is 56 picas (237 mm).
The first page has less text height. It requires an additional footer
space of 3.5 picas (14.8 mm) for the copyright inserted by the publisher
and 1.5 picas (6 mm) of space before the title.
The effective text height of the first page is 51 picas (216 mm).

\item There are no running feet for the final camera-ready version of the
paper. The submission paper should have page numbers in the running feet.

\end {enumerate}

\section{TITLE, AUTHOR, AFFILIATION, COPYRIGHT
AND RUNNING FEET}
\subsection{Title}
The title is set in 20 point (7 mm) bold with leading of 22 point (7.7
mm), centered over the full text measure, with 1.5 picas (6 mm) of
space before and after.

\subsection{Author}
The author's name is set in 11 point (3.85 mm) bold with leading of 12
point (4.2 mm), centered over full text measure, with 1.5 picas (6 mm)
of space below. A footnote indicator is set in 11 point (3.85 mm)
medium and positioned as a superscript character.

\subsection{Affiliation}
The affiliation is set as a footnote to the first column. This is set
in 8 point (2.8 mm) medium with leading of 8.6 point (3.1 mm), with a
1 point (0.35 mm) footnote rule to column width.

\subsection{Copyright}
The copyright details will be inserted by the publisher.

\subsection{Running feet}
The running feet are inserted by the publisher. For submission you may
insert page numbers in the middle of the running feet. Do not,
however, insert page numbers for the camera-ready version of the
paper.

\section{ABSTRACT}
The abstract for the paper is set in 9 point (3.15 mm) medium, on a
body of 10 point (3.5 mm). The word Abstract is set in bold, followed
by a full point and a 0.5 pica space.

\section{HEADINGS}\label{heads}
Three heading levels have been specified:

\begin{enumerate}
\item{A level headings}

\begin{itemize}
\item The first level of heading is set is 11 point (3.85 mm) bold, on
  a body of 12 point (4.2 mm), 1.5 lines of space above and 0.5 lines
  of space below.

\item The heading is numbered to one digit with a 1 pica space
  separating it from the text.

\item The text is keyed in capitals and is unjustified.
\end{itemize}

\item{B level headings}
\begin{itemize}
\item The second level of heading is set is 11 point (3.85 mm) bold,
  on a body of 12 point (4.2 mm), 1.5 lines of space above and 0.5
  lines of space below.

\item The heading is numbered to two digits separated with a full
  point, with a 1 pica space separating it from the text.

\item The text is keyed in upper and lower case with an initial
  capital for first word only, and is unjustified.
\end{itemize}

\item{C level headings}
\begin{itemize}
\item The third level of heading is set is 10 point (3.5 mm) italic,
  on a body of 11 point (3.85 mm), 1.5 lines of space above and 0.5
  lines of space below.

\item The heading is numbered to three digits separated with a full
  point, with a 1 pica space separating it from the text.

\item The text is keyed in upper and lower case with an initial
  capital for first word only, and is unjustified.
\end{itemize}

\item{Acknowledgements}

  This heading is the same style as an A level heading but is not
  numbered.
\end{enumerate}

\section{TEXT}
The first paragraph of text following any heading is set to the
complete measure (i.e., do not indent the first line).

Subsequent paragraphs are set with the first line indented
by 1 pica (3.85 mm).

There isn't any inter-paragraph spacing.

\section{LISTS}
The list identifier may be an arabic number, a bullet, an em
rule or a roman numeral.

The items in a list are set in text size and indented by 1
pica (4.2 mm) from the left margin. Half a line of space is
set above and below the list to separate it from surrounding
text.

See layout of Section \ref{heads} on headings to see the results of the list macros.

\section{TABLES}
Tables are set in 8 point (2.8 mm) on a body of 10 point (3.5 mm).
The table caption is set centered at the start of the table, with
the word Table and the number in bold. The caption is set in medium
with a 1 pica (4.2 mm) space separating it from the table number.

A one line space separates the table from surrounding text.

\begin{table}
\begin{center}
{\caption{The table caption is centered on the table measure. If it
extends to two lines each is centered.}\label{table1}}
\begin{tabular}{lccccccc}
\hline
\rule{0pt}{12pt}
&\multicolumn{7}{c}{Processors}\\
&1&\multicolumn{3}{c}{2}&\multicolumn{3}{c}{4}\\
\cline{2-8}
\rule{0pt}{12pt}
Window&$\Diamond$&$\Diamond$&$\Box$&$\bigtriangleup$&$\Diamond$&$\Box$&$\bigtriangleup$
\\
\hline
\\[-6pt]
\quad1&1273&110&21.79&89\%&6717&22.42&61\%\\
\quad2&2145&116&10.99&50\%&5386&10.77&19\%\\
\quad3&3014&117&41.77&89\%&7783&42.31&58\%\\
\quad4&4753&151&71.55&77\%&7477&61.97&49\%\\
\quad5&5576&148&61.60&80\%&7551&91.80&45\%
\\
\hline
\\[-6pt]
\multicolumn{8}{l}{$\Diamond$ execution time in ticks\ \
$\Box$ speed-up values\ \
$\bigtriangleup$ efficiency values}
\end{tabular}
\end{center}
\end{table}

\section{FIGURES}
A figure caption is set centered in 8 point (2.8 mm) medium on a
leading of 10 point (3.5 mm).  It is set under the figure, with the
word Figure and the number in bold and with a 1 pica (4.2 mm) space
separating the caption text from the figure number.

One line of space separates the figure from the caption. A one line
space separates the figure from surrounding text.

\begin{figure}
\centerline{\includegraphics[height=3.5in]{ecaif01.ps}}
\caption{Network of transputers and the structure of individual
processes } \label{procstructfig}
\end{figure}

\section{EQUATIONS}
A display equation is numbered, using arabic numbers in parentheses.
It is centered and set with one line of space above and below to
separate it from surrounding text. The following example is a simple
single line equation:

\begin{equation}
Ax=b
\label{thesystem}
\end{equation}

The next example is a multi-line equation:

\begin{eqnarray}
(x+y)(x-y) & = & x^2-xy+xy-y^2\\
(x+y)^2    & = & x^2+2xy+y^2
\end{eqnarray}

The equal signs are aligned in a multi-line equation.

\section{PROGRAM LISTINGS}
Program listings are set in 9 point (3.15 mm) Courier on a
leading of 11 point (3.85 mm). That is to say, a non-proportional
font is used to ensure the correct alignment.

A one line space separates the program listing from surrounding text.

\begin{verbatim}
void inc(x)
int* x;
{
    *x++;
}

int y = 1;
inc(&y);
printf("%d\n",y);
\end{verbatim}

\section{THEOREMS}
The text of a theorem is set in 9 point (3.15 mm) italic on a
leading of 11 point (3.85 mm). The word Theorem and its number
are set in 9 point (3.15 mm) bold.

A one line space separates the theorem from surrounding text.

\begin{theorem}
Let us assume this is a valid theorem. In reality it is a piece
of text set in the theorem environment.
\end{theorem}

\section{FOOTNOTES}
Footnotes are set in 8 point (2.8 mm) medium with leading of 8.6 point (3.1
mm), with a 1 point (0.35 mm) footnote rule to column
width\footnote{This is an example of a footnote that occurs in
the text. If the text runs to two lines the second line aligns
with the start of text in the first line.} .

\section{REFERENCES}
The reference identifier in the text is set as a sequential number in
square brackets. The reference entry itself is set in 8 point
(2.8 mm) with a leading of 10 point (3.5 mm), and appears
in the sequence in which it is cited in the paper.

\section{SAMPLE CODING}
The remainder of this paper contains examples of the specifications
detailed above and can be used for reference if required.

\section{PROGRAMMING MODEL}
\label{par}
Our algorithms were implemented using the \emph{single program,
  multiple data} model (SPMD). SPMD involves writing a single code
that will run on all the processors co-operating on a task. The data
are partitioned among the processors which know what portions of the
data they will work on \cite{kn:Golub89}.

\subsection{Structure of processes and processors}
\label{procstruct}
The grid has $P=P_{\rm{r}}\times P_{\rm{c}}$ processors, where
$P_{\rm{r}}$ is the number of rows of processors and $P_{\rm{c}}$ is
the number of columns of processors.

\subsubsection{Routing information on the grid}
\label{routing} A message may be either \emph{broadcast} or {\em
  specific}. A broadcast message originates on a processor and is
relayed through the network until it reaches all other processors. A
specific message is one that is directed to a particular target
processor.

Broadcast messages originate from a processor called \emph{central}
which is situated in the `middle' of the grid. This processor has
co-ordinates $(\lfloor P_{\rm{r}}/2 \rfloor,\lfloor P_{\rm{c}}/2
\rfloor)$.  Messages are broadcast using the \emph{row--column
  broadcast} algorithm (RCB), which uses the following strategy.  The
number of steps required to complete the RCB algorithm (i.e., until
all processors have received the broadcast value) is given by $\lfloor
P_{\rm{r}}/2\rfloor+\lfloor P_{\rm{c}}/2\rfloor$.

A specific message is routed through the processors using the
\emph{find-row--find-column} algorithm (FRFC) detailed in
\cite{kn:deCarlini91}.  The message is sent from the \emph{originator}
processor vertically until it reaches a processor sitting in the same
row as the \emph{target} processor. The message is then moved
horizontally across the processors in that row until it reaches the
target processor. An accumulation based on the recursive doubling
technique \cite[pp. 56--61]{kn:Modi88}, would require the same number
of steps as the RCB requires. If either the row or column of the
originator and target processors are the same then the message will
travel only in a horizontal or vertical direction, respectively (see
\cite{kn:Smith85}).

\section{DATA PARTITIONING}
\label{datapart}
We use \emph{data partitioning by contiguity}, defined in the
following way. To partition the data (i.e., vectors and matrices)
among the processors, we divide the set of variables
\mbox{$V=\{\,i\,\}_{i=1}^{N}$} into $P$ subsets
$\{\,W_{p}\,\}_{p=1}^{P}$ of $s=N/P$ elements each. We assume without
loss of generality that $N$ is an integer multiple of $P$. We define
each subset as $W_{p}=\{(p-1)s+j\}_{j=1}^{s}$ (see
\cite{kn:Schofield89}, \cite{kn:daCunha92a} and \cite{kn:Atkin} for
details).

Each processor $p$ is responsible for performing the computations over
the variables contained in $W_{p}$. In the case of vector operations,
each processor will hold segments of $s$ variables. The data
partitioning for operations involving matrices is discussed in Section
\ref{mv}.
\section{LINEAR ALGEBRA OPERATIONS}
\label{blas}
\subsection{Saxpy}
\label{saxpy}
The saxpy $w=u+\alpha v$ operation, where $u$, $v$ and $w$ are vectors
and $\alpha$ is a scalar value, has the characteristic that its
computation is \emph{disjoint elementwise} with respect to $u, v$ and
$w$. This means that we can compute a saxpy without any communication
between processors; the resulting vector $w$ does not need to be
distributed among the processors. Parallelism is exploited in the
saxpy by the fact that $P$ processors will compute the same operation
with a substantially smaller amount of data. The saxpy is computed as
\begin{equation}
w_{i}=u_{i}+\alpha v_{i},\quad \forall i\in\{W_{p}\}_{p=1}^{P}
\label{distsaxpy}
\end{equation}
\subsection{Inner-product and vector 2-norm}
\label{utv}
The inner-product $\alpha=u^{T}v=\sum_{i=1}^{N}{u_{i}v_{i}}$ is an
operation that involves accumulation of data, implying a high level of
communication between all processors. The mesh topology and the
processes architecture used allowed a more efficient use of the
processors than, for instance, a ring topology, reducing the time that
processors are idle waiting for the computed inner-product value to
arrive, but the problem still remains. The use of the SPMD paradigm
also implies the global broadcast of the final computed value to all
processors.

The inner-product is computed in three distinct phases. Phase 1 is the
computation of partial sums of the form
\begin{equation}
\alpha_{p}=\sum_{\forall i\in\{W_{p}\}}{u_{i}\times v_{i}},\quad p=1,\ldots,P
\label{distutv}
\end{equation}

The accumulation phase of the inner-product using the RCA algorithm is
completed in the same number of steps as the RCB algorithm (Section
\ref{routing}). This is because of the need to relay partial values
between processors without any accumulation taking place, owing to the
connectivity of the grid topology.

The vector 2-norm $\alpha=||\,u\,||_{2}=\sqrt{u^{T}u}$ is computed
using the inner-product algorithm described above. Once the
inner-product value is received by a processor during the final
broadcast phase, it computes the square root of that value giving the
required 2-norm value.

\subsection{Matrix--vector product}
\label{mv}
For the matrix--vector product $v=Au$, we use a \emph{column
  partitioning} of $A$.  Each processor holds a set $W_{p}$ (see
Section \ref{datapart}) of $s$ columns each of $N$ elements of $A$ and
$s$ elements of $u$.  The $s$ elements of $u$ stored locally have a
one-to-one correspondence to the $s$ columns of $A$ (e.g. a processor
holding element $u_{j}$ also holds the $j$-th column of $A$). Note
that whereas we have $A$ partitioned by columns among the processors,
the matrix--vector product is to be computed by \emph{rows}.

The algorithm for computing the matrix--vector product using column
partitioning is a generalization of the inner-product algorithm
described in Section \ref{utv} (without the need for a final broadcast
phase). At a given time during the execution of the algorithm, each
one of $P-1$ processors is computing a vector $w$ of $s$ elements
containing partial sums required for the segment of the vector $v$ in
the remaining `target' processor.  After this computation is complete,
each of the $P$ processors stores a vector $w$.  The resulting segment
of the matrix--vector product vector which is to be stored in the
target processor is obtained by summing together the $P$ vectors $w$,
as described below.

Each processor other than the target processor sends its $w$ vector to
one of its neighboring processors. A processor decides whether to send
the vector in either the row or column direction to reach the target
processor based on the FRFC algorithm (see Section \ref{routing}). If
a vector passes through further processors in its route to the target
processor the $w$ vectors are accumulated. Thus the target processor
will receive at most four $w$ vectors which, when summed to its own
$w$ vector, yield the desired set of $s$ elements of $v$.

\subsection{Matrix--vector product---finite-difference approximation}
We now consider a preconditioned version of the conjugate-gradients
method \cite{kn:Golub89}. Note that we do not need to form $A$
explicitly.  This implies a very low degree of information exchange
between the processors which can be effectively exploited with
transputers, since the required values of $u$ can be exchanged
independently through each link.

The preconditioning used in our implementations is the polynomial
preconditioning (see \cite{kn:Saad85}, \cite{kn:Eisenstat81},
\cite{kn:Adams85} and \cite{kn:Johnson83}), which can be implemented
very efficiently in a parallel architecture since it is expressed as a
sequence of saxpys and matrix--vector products.

We have $l$ rows and columns in the discretization grid, which we want
to partition among a $P_{\rm{r}}\times P_{\rm{c}}$ mesh of
processors. Each processor will then carry out the computations
associated with a block of $\lfloor
l/P_{\rm{r}}\rfloor+\hbox{sign}\left( l\bmod P_{\rm{r}}\right)$ rows
and $\lfloor l/P_{\rm{c}}\rfloor+\hbox{sign}\left( l\bmod
  P_{\rm{c}}\right)$ columns of the interior points of the grid.

The matrix--vector product using the column partitioning is highly
parallel. Since there is no broadcast operation involved, as soon as a
processor on the boundary of the grid (either rows or columns) has
computed and sent a $w_{p}$ vector destined to a processor `A', it can
compute and (possibly) send a $w_{p}$ vector to processor `B', at
which time its neighboring processors may also have started computing
and sending their own $w$ vectors to processor `B'.

At a given point in the matrix--vector product computation, the
processors are computing $w$ vectors destined to processor A. When
these vectors have been accumulated in the row of that processor (step
1), the processors in the top and bottom rows compute and send the $w$
vectors for processor B, while the processors on the left and right
columns of the row of processor A send the accumulated $r$ vectors to
processor A (step 2). Processor A now stores its set of the resulting
$v$ vector (which is the accumulation of the $w$ vectors). In step 3,
the processors in the bottom row compute and send the $w$ vectors for
processor C while the processor at the left-hand end of the row of
processor B sends the accumulated $w$ vectors of that column towards
processor B. The next steps are similar to the above.

In our implementation, we exploit the geometry associated with the
regular grid of points used to approximate the PDE. A geometric
partitioning is used to match the topology and connectivity present in
the grid of transputers (Section \ref{procstruct}).

The discretization of the PDE is obtained by specifying a grid size
$l$ defining an associated grid of $N=l^2$ interior points (note that
this is the order of the linear system to be solved). With each
interior point, we associate a set of values, namely the coefficients
$C, N, S, E\,$ and $W$.

\section{CONCLUSION}
We have shown that an iterative method such as the preconditioned
conjugate-gradients method may be successfully parallelized by using
highly efficient parallel implementations of the linear algebra
operations involved. We have used the same approach to parallelize
other iterative methods with similar degrees of efficiency (see
\cite{kn:daCunha92a} and \cite {kn:daCunha92b}).

\ack We would like to thank the referees for their comments, which
helped improve this paper considerably

\bibliography{ecai2012}
\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2012-01-28 paper/srcr/paper.tex
added papers

--- 2012-01-24 paper/srcr/paper.tex
@@
-%File: formatting-instruction.tex
 \documentclass[letterpaper]{article}
@@
 proposed sampling schemes is evaluated on sets of Bernoulli arms, in randomly
-generated 2-level trees, and on the sailing domain, showing where the proposed schemes show improved
+generated 2-level trees, and on the sailing domain, showing where the proposed schemes have improved
 performance, the third contribution of this paper.
@@
 (MDP).  An MDP is defined by the set of states $S$, the set of actions
-$A$, the transition table $T(s, a, s')$, the reward table $R(s, a,
-s')$, the initial state $s$ and the goal state $t$: $(S, A, T, R, s,
+$A$ (also called {\em moves} in this paper),
+the transition distribution $T(s, a, s')$, the reward function $R(s, a,
+s')$, the initial state $s$ and an optional goal state $t$: $(S, A, T, R, s,
 t)$ \cite{Russell.aima}.  Several MCTS schemes explore an MDP by
@@
 of $K$ arms (see Figure~\ref{fig:mab-simple-regret}.a). Each arm can
-be pulled multiple times (sometimes a cost is
-associated with each pulling action). When the $i$th arm
+be pulled multiple times. Sometimes a cost is
+associated with each pulling action. When the $i$th arm
 is pulled, a random reward $X_i$ from an unknown stationary
-distribution is encountered.  The reward is bounded between 0 and 1.
-In the cumulative setting, all encountered rewards are collected
-by the agent. In the simple (selection) setting, the agent gets to
+distribution is encountered.  The reward is usually bounded between 0 and 1.
+In the cumulative setting (the focus of much of the research literature on Multi-armed bandits),
+all encountered rewards are collected by the agent. 
+The UCB scheme was shown to be near-optimal in this respect \cite{Auer.ucb}:
+
+\begin{dfn} Scheme $\mathbf{UCB(c)}$ pulls arm $i$ that maximizes 
+upper confidence bound $b_i$ on the reward:
+\begin{equation}
+b_i=\overline X_i+\sqrt {\frac {c \log (n)} {n_i}}
+\label{eqn:ucb}
+\end{equation}
+where $\overline X_i$ is the average reward obtained from arm $i$,
+$n_i$ is the number of times arm $i$ was pulled, and $n$ is the total
+number of pulls so far. \end{dfn}
+
+The UCT algorithm, an extension of UCB
+to Monte-Carlo Tree Search is described in \cite{Kocsis.uct}, and
+shown to outperform many state of the art search algorithms in both
+MDP and adversarial search \cite{Eyerich.ctp}.
+
+In the simple regret (selection) setting, the agent gets to
 collect only the reward of the last pull.
@@
 
-%Strategies that minimize the simple regret are called pure exploration
-%strategies \cite{Bubeck.pure}.
-
-Much of research literature on Multi-armed bandits centered
-on minimizing the cumulative regret. The
-UCB scheme (see below) was shown to be near-optimal in minimizing the cumulative regret \cite{Auer.ucb}.
-
-\begin{dfn} Scheme $\mathbf{UCB(c)}$ pulls arm $i$ that maximizes 
-upper confidence bound $b_i$ on the reward:
-\begin{equation}
-b_i=\overline X_i+\sqrt {\frac {c \ln n} {n_i}}
-\label{eqn:ucb}
-\end{equation}
-where $\overline X_i$ is the average reward obtained from arm $i$,
-$n_i$ is the number of times arm $i$ was pulled, and $n$ is the total
-number of pulls so far. \end{dfn} The best known upper bound on the simple
-regret of UCB($c$) is polynomially decreasing in the number of samples
-(see \cite{Bubeck.pure}, Theorems~2,3).
-
-An upper bound on the simple regret of uniform sampling is
+Strategies that minimize the simple regret are called pure exploration
+strategies \cite{Bubeck.pure}. An upper bound on the simple regret of uniform sampling is
 exponentially decreasing in the number of samples (see
-\cite{Bubeck.pure}, Proposition~1). However, empirically
-UCB($c$) yields a lower simple regret than uniform
+\cite{Bubeck.pure}, Proposition~1). For UCB($c$) the best known respective upper bound on the simple
+regret of UCB($c$) is only polynomially decreasing in the number of samples
+(see \cite{Bubeck.pure}, Theorems~2,3). However, empirically
+UCB($c$) appears to yield a lower simple regret than uniform
 sampling.
 
-The UCT algorithm, an extension of UCB
-to Monte-Carlo Tree Search is described in \cite{Kocsis.uct}, and
-shown to outperform many state of the art search algorithms in both
-MDP and adversarial search \cite{Eyerich.ctp}. Pure exploration in
-Multi-armed bandits is examined in \cite{Bubeck.pure}. On the one
-hand, the \cite{Bubeck.pure} prove certain upper and lower bounds for
-UCB and uniform sampling, showing that an upper bound on the simple
-regret is exponential in the number of samples for uniform sampling,
-while only polynomial for UCB. On the other hand, empirical
-performance of UCB appears to be better than that of uniform sampling.
+%Pure exploration in
+%Multi-armed bandits is examined in \cite{Bubeck.pure}. On the one
+%hand, the \cite{Bubeck.pure} prove certain upper and lower bounds for
+%UCB and uniform sampling, showing that an upper bound on the simple
+%regret is exponential in the number of samples for uniform sampling,
+%while only polynomial for UCB. On the other hand, empirical
+%performance of UCB appears to be better than that of uniform sampling.
 
@@
 
-A completely different direction of controlling sampling can use the
+A completely different scheme for control of sampling can use the
 principles of bounded rationality \cite{Horvitz.reasoningabout}
@@
 and finds the expected gain from finding another move 
-$\beta $ to be better than the current best \cite{Russell.aima}. The ``cost'' of search
+$\beta $ to be better than the current best \cite{Russell.right}. The ``cost'' of search
 actions can also be factored in.
@@
 \begin{itemize}
-\item Defining the cost of a sample is not trivial.
 \item The method is in general intractable, necessitating simplifying assumptions.
@@
   difficult to obtain.
+\item Defining the time-cost of a sample is not trivial.
 \end{itemize}
@@
 This sampling scheme exhibits an exponentially decreasing simple regret:
-\begin{thm} For any $0<\eta<1$  and $\gamma>1$ there exists $N$ such that for
+\begin{thm} For every $0<\eta<1$  and $\gamma>1$ there exists $N$ such that for
   any number of samples $n>N$ the simple regret of the  $\varepsilon$-greedy
@@
 example, $\sqrt\cdot$.
-\begin{dfn} Scheme $\mathbf{UCB_{\sqrt{\cdot}}(c)}$ repeatedly pulls arm $i$ that
-maximizes $b_i$:
+\begin{dfn} Scheme $\mathbf{UCB_{\sqrt{\cdot}}(c)}$ pulls arm $i$ that
+maximizes $b_i$, where:
 \begin{equation}
@@
 This scheme also exhibits a super-polynomially decreasing simple regret:
-\begin{thm}  For any $0<\eta<1$  and $\gamma>1$ there exists $N$ such that for
+\begin{thm}  For every $0<\eta<1$  and $\gamma>1$ there exists $N$ such that for
   any number of samples $n>N$ the simple regret of the  UCB$_{\sqrt{\cdot}}$($c$)
@@
 
-UCT \cite{Kocsis.uct} is a generalization of UCB for MCTS.  UCT
+As mentioned above,
+UCT \cite{Kocsis.uct} is an extension of UCB for MCTS, that
 applies UCB($c$) at each step of a rollout.  At the root node,
-the goal of sampling in MCTS is
-usually finding the first action to perform. Search is re-started,
+the sampling in MCTS is usually aimed at finding the first move to perform. Search is re-started,
 either from scratch or using some previously collected information,
@@
 
-For nodes deeper in the search tree the situation is not as clear, but here the task appears different.
+Nodes deeper in the search tree are a different matter.
 In order to support an optimal move choice at the root, a precise {\em value} of the search
-tree nodes is beneficial. For these internal nodes optimizing simple regret is not the answer, and 
+tree nodes is beneficial. For these internal nodes, optimizing simple regret is not the answer, and 
 cumulative regret optimization seems more reasonable. Lacking a complete metareasoning for sampling,
@@
 
-We will denote realizations of SR+CR as \emph{Alg}+UCT, where
-\emph{Alg} is the sampling scheme employed at the first step of a
-rollout (e.g. $\frac 1 2$-greedy+UCT). Such two-stage sampling scheme
-outperforms UCT, and, additionally, is
-significantly less sensitive to the tuning of the exploration factor
-$c$ of UCB($c$), since the contradiction between the need for a larger
-value of $c$ on the first step (simple regret) and a smaller
-value for the rest of the rollout (cumulative regret)
-\cite{Bubeck.pure} is resolved. In fact, a sampling scheme that uses
-UCB($c$) at all steps but a larger value of $c$ for the first step
-than for the rest of the steps, outperforms UCT. The pseudocode of the
-two-stage rollout is in
+The pseudocode of this two-stage rollout is in
 Algorithm~\ref{alg:two-stage-mcts}: \textsc{FirstAction} selects the first
@@
 the rest of the layout (usually using UCB).
+We denote such two-step realizations of SR+CR as \emph{Alg}+UCT, where
+\emph{Alg} is the sampling scheme employed at the first step of a
+rollout (e.g. $\frac 1 2$-greedy+UCT).
 
@@
 \end{algorithm}
+
+We expect such two-stage sampling schemes to
+outperform UCT and be
+significantly less sensitive to the tuning of the exploration factor
+$c$ of UCB($c$). That is since the contradiction between the need for a larger
+value of $c$ on the first step (simple regret) and a smaller
+value for the rest of the rollout (cumulative regret)
+\cite{Bubeck.pure} is resolved. In fact, a sampling scheme that uses
+UCB($c$) at all steps but a larger value of $c$ for the first step
+than for the rest of the steps, outperforms UCT.
 
@@
 leaving this issue for future research.
-Early experiments with this approach demonstrate a significantly lower simple regret.
+Unfortunately, this scheme appears too complicated to be amenable to a formal
+analysis. Early experiments with this approach demonstrate a significantly lower simple regret.
 
@@
 The second set of experiments was performed on randomly generated
-trees crafted in such a way that uniform random sampling selects a
-direction at the root randomly. The degree of the root is a parameter of
-the tree generator. The degree of all nodes at distance 1 from the
-root is 2, and all nodes at distance 2 from the roots are leaves. The
-average reward of two children of each node at distance 1 is
-0.5. Thus, a uniform sampling scheme results in the same average reward for
-all edges at the root, and an adaptive sampling scheme, such as UCT,
-has to be used.
+2-level max-max trees crafted so as to deliberately deceive uniform sampling
+(Figure~\ref{fig:mcts-regret}.a), necessitating  an adaptive sampling scheme, 
+such as UCT. That is due to the switch nodes, each with 2 children with anti-symmetric
+values, which would cause a uniform sampling scheme to incorrectly give them all a value of 0.5.
 
-Figure~\ref{fig:mcts-regret} shows a sketch of the search tree
-(Figure~\ref{fig:mcts-regret}.a) and the dependency of the regret vs. the
-number of samples for trees with root degree 16
+Simple regret vs. the number of samples are shown for trees with root degree 16
 (Figure~\ref{fig:mcts-regret}.b) and 64 (Figure~\ref{fig:mcts-regret}.c).
@@
   \label{fig:sailing-cost-vs-factor}
+  \vspace{-1.5em}
 \end{figure}
@@
 the results for 32 arms. VOI+UCT, the scheme based on a VOI estimate,
-outperforms all other sampling schemes in this example. 
+outperforms all other sampling schemes in this example. Similar performance improvements
+(not shown) also occur for the sailing domain.
 \begin{figure}[h!]
@@
   \label{fig:mcts-regret-voi} 
+  \vspace{-1.5em}
 \end{figure}
@@
 
+Finding a better sampling scheme for non-root nodes,
+as well as the root node, should also be possible.
+Although cumulative regret does reasonably
+well there, it is far from optimal, as meta-reasoning principles imply that an optimal scheme
+for these nodes must be asymmetrical (e.g. it is not helpful to find out that the
+value of the current best action is even better than previously believed).
+
 %\section*{Acknowledgments}

### 2012-01-28 paper/vago/paper.tex
added papers

@@
   modified and based solely on the sample cost, specified as
-  a constant parameter. 
+  a constant parameter. The heuristic stopping condition for the
+  original UCT policy was left unchanged.
 \item The time-allocation mode based on the fixed number of samples
-  was modified for both the original UCT policy and the VOI-aware
-  policies such that 
+  was modified for \textit{both the original UCT policy and the VOI-aware
+  policies} such that 
   \begin{enumerate}

### 2012-02-21 paper/srcr/paper.tex
presentation narrative in progress

@@
 
-%\title{MCTS Based on Simple Regret}
-%\author {David Tolpin, Solomon Eyal Shimony \\
-%Department of Computer Science, \\
-%Ben-Gurion University of the Negev, Beer Sheva, Israel \\
-%\{tolpin,shimony\}@cs.bgu.ac.il}
+\title{MCTS Based on Simple Regret}
+\author {David Tolpin, Solomon Eyal Shimony \\
+Department of Computer Science, \\
+Ben-Gurion University of the Negev, Beer Sheva, Israel \\
+\{tolpin,shimony\}@cs.bgu.ac.il}
 
 \title{MCTS Based on Simple Regret}
-\author {Submission 198}
 
@@
 
-%\section*{Acknowledgments}
+\section*{Acknowledgments}
 
-%The research is partially supported by Israel
-%Science Foundation grant 305/09, by the Lynne and William Frankel
-%Center for Computer Sciences, and by the Paul Ivanier Center for
-%Robotics Research and Production Management.
+The research is partially supported by Israel
+Science Foundation grant 305/09, by the Lynne and William Frankel
+Center for Computer Sciences, and by the Paul Ivanier Center for
+Robotics Research and Production Management.
 

### 2012-02-22 slides/slides.tex
slides in progress

@@
 
-\title{Rational Monte Carlo Tree Search}
+\title{MCTS Based on Simple Regret}
 \author{David Tolpin, Solomon Eyal Shimony}
 \institute{Ben-Gurion University of the Negev\\Beer Sheva, Israel}
-\date{June 1, 2011}
+\date{February 22, 2012}
 
@@
 
+\begin{frame}{Hard to solve search problems}
+Search problems are often hard too solve in practice when:
+\begin{itemize}
+\item search space is extremely large;
+\item {\it and} good heuristics are unknown.
+\end{itemize}
+\textbf{Easier} to solve:
+\begin{itemize}
+\item Chess --- search space size is manageable ($10^{50}$).
+\item Timetabling --- good heuristics.
+\end{itemize}
+\textbf{Hard} to solve:
+\begin{itemize}
+\item Compute Go ($10^{180}$), Poker ($10^{70}$).
+\item Canadian Traveller Problem.
+\end{itemize}
+\end{frame}
+
+\begin{frame}{MCTS}
+
+{\bf M}onte {\bf C}arlo {\bf T}ree {\bf S}earch helps in large search spaces.
+\begin{itemize}
+\item Starts with the root only.
+\item Repeats:
+  \begin{enumerate}
+    \item \textbf{Selection:} select a branch to explore.
+    \item \textbf{Expansion:} adds children of the leaf to the stored
+      tree.
+    \item \textbf{Simulation:} continues search (using a simple
+      strategy) until a goal state is reached.
+    \item \textbf{Backpropagation:} values of each stored node 
+      are updated.
+  \end{enumerate}
+\end{itemize}
+\textbf{Adaptive} MCTS samples `good' moves more frequently,\\
+but sometimes {\bf explores} new directions.
+\end{frame}
+
+\begin{frame}{Multi-armed Bandit Problem and UCB}
+Multi-armed Bandit Problem:
+\begin{itemize}
+\item We are given a set of $K$ arms.
+\item Each arm can be pulled multiple times.
+\item The reward is drawn from an {\bf unknown} (but normally {\it
+    stationary} and {\it bounded}) distribution.
+\item The {\bf total reward} must be maximized.
+\end{itemize}
+
+{\bf UCB} is near-optimal for MAB --- solves
+{\it exploration/exploitation} tradeoff.
+\begin{itemize}
+\item pulls an arm that maximizes {\bf U}pper {\bf C}onfidence {\bf
+    B}ound: $b_i=\overline X_i+\sqrt {\frac {c \log (n)} {n_i}}$
+\item the cumulative regret is $O(\log n)$.
+\end{itemize}
+\end{frame}
+
 \begin{frame}{UCT}
-UCT (Upper Confidence Bounds on Trees) is popular for Monte
-  Carlo search in large trees:
+UCT ({\bf U}pper {\bf C}onfidence Bounds applied to {\bf T}rees) is 
+based on UCB.
 \begin{itemize}
-\item Selects an action $a_i$ that maximizes $\overline r_i + C\sqrt
-  {\frac {\log n} {n_i}}$
-\item Chooses a non-optimal action $n_i\ge\rho\log n$ times ($\rho$ is
-  some constant).
-\item Based on UCB1, that achieves logarithmic regret for multi-armed
-  bandits.
-\item But {\bf{\Large no bandits in Monte Carlo:}} {\bf no reward} is given for {\bf
-  sampling} a good action.
+\item Adaptive MCTS.
+\item Applies the UCB selection scheme at each step of the rollout.
+\item Demonstrated good performance in Computer Go (MoGo, CrazyStone, Fuego,
+  Pachi, ...) as well as in other domains.
 \end{itemize}
+However, the first step of a rollout is different:
+\begin{itemize}
+\item The purpose of MCTS is to choose an action with the greatest utility.
+\item Therefore, the {\bf simple regret} must be minimized.
+\end{itemize}
+\end{frame}
+
+\begin{frame}{SRCR}
+
+{\bf S}imple {\bf R}egret followed by {\bf C}umulative {\bf R}egret.
+\begin{itemize}
+\item Maximizes {\bf simple regret} at the {\bf first step}.
+\item Continues with UCT from the {\bf second step on}.
+\end{itemize}
+\vspace{1em}
+\begin{algorithmic}[1]
+\Procedure{Rollout}{node, depth=1}
+  \If {\Call{IsLeaf}{node, depth}}
+    \State \textbf{return} 0
+  \Else
+    \State \textbf{if} depth=1 \textbf{then} action $\gets$ \Call{FirstAction}{node} \label{alg:srcr-first-action}   
+    \State \textbf{else} action $\gets$ \Call{NextAction}{node} \label{alg:srcr-next-action}
+    \State next-node $\gets$ \Call{NextState}{node, action}
+    \State reward $\gets$ \Call{Reward}{node, action, next-node}
+     \State \hspace{4em} + \Call{Rollout}{next-node, depth+1}
+    \State \Call{UpdateStats}{node, action, reward}
+  \EndIf
+\EndProcedure
+\end{algorithmic}
+\end{frame}
+
+\begin{frame}{Sampling for Simple Regret}
+Sampling schemes for miniminizing the simple regret:
+\begin{enumerate}
+\item $\varepsilon$-greedy sampling.
+\item a modified version of UCB (worse for cumulative, better for
+  simple regret).
+\item<+-> VOI-based sampling.
+\end{enumerate}
+
+\begin{itemize}
+\item<+-> 1, 2 --- heuristic selection criterion, theoretical upper bounds
+  can be obtained.
+\item<+-> 3 --- based on principles of {\it Rational Metareasoning},
+  but harder to analyze.
+\end{itemize}
+
 \end{frame}

### 2012-02-22 slides/slides.tex
heuristic sampling schemes

@@
 \begin{itemize}
-\item Starts with the root only.
-\item Repeats:
+\item<+-> Starts with the root only.
+\item<+-> Repeats:
   \begin{enumerate}
     \item \textbf{Selection:} select a branch to explore.
-    \item \textbf{Expansion:} adds children of the leaf to the stored
+    \item<+-> \textbf{Expansion:} adds children of the leaf to the stored
       tree.
-    \item \textbf{Simulation:} continues search (using a simple
+    \item<+-> \textbf{Simulation:} continues search (using a simple
       strategy) until a goal state is reached.
-    \item \textbf{Backpropagation:} values of each stored node 
+    \item<+-> \textbf{Backpropagation:} values of each stored node 
       are updated.
@@
 \end{itemize}
+\vspace{1em}
 \textbf{Adaptive} MCTS samples `good' moves more frequently,\\
@@
 
+\begin{frame}{Heuristic sampling schemes}
+
+\onslide<+->
+$\varepsilon$-greedy:
+\begin{itemize} 
+\item Pulls the empirically best arm with probability $\varepsilon$.
+\item Any other arm with probability $\frac {1-epsilon} {K-1}$.
+\item Exhibits exponentially decreasing simple regret.
+\item Uniform sampling when $\varepsilon=\frac 1 K$, much better when
+  $\varepsilon=\frac 1 2$.
+\end{itemize}
+\vspace{1em}
+\onslide<+->
+$UCB_{\sqrt{\cdot}}$ ($\sqrt{\cdot}$ instead of $\log$): 
+\begin{itemize}
+\item Pulls arm $i$ that maximizes
+  $b_i=\overline X_i+\sqrt {\frac {c \sqrt n} {n_i}}$. 
+\item Exhibits superpolynomially decreasing simple regret.
+\end{itemize}
+\end{frame}
+
 \begin{frame}{Doing better than UCT on sets}

### 2012-02-22 slides/slides.tex
voi-aware sampling
all goes well

@@
 
+\begin{frame}{VOI-aware sampling}
+
+\begin{itemize}
+\item Chooses an action with the maximum VOI estimate.
+\item Estimates the VOI based on bounds on:
+  \begin{itemize}
+  \item the probability that one or more rollouts will make another
+    action appear better than the current best;
+  \item the gain that may be incurred if such a change occurs.
+  \end{itemize}
+\end{itemize}
+\begin{eqnarray*}
+VOI_\alpha&\approx&\frac {\overline X_\beta} {n_\alpha+1}
+\exp\left(-2(\overline X_\alpha - \overline X_\beta)^2 n_\alpha\right)\\
+VOI_i&\approx&\frac {1-\overline X_\alpha} {n_i+1}
+\exp\left(-2(\overline X_\alpha - \overline X_i)^2 n_i\right),\; i\ne\alpha\nonumber\\
+\mbox{where }&&\alpha=\arg\max_i \overline X_i,\quad
+             \beta=\arg\max_{i,\,i\ne\alpha} \overline X_i\nonumber
+\end{eqnarray*}
+\end{frame}
+
 \begin{frame}{Doing better than UCT on sets}
@@
 
+\begin{frame}{Summary}
+\begin{itemize}
+\item<+-> Improved MCTS scheme, SRCR, introduced.
+\item<+-> SRCR performs better than unmodified UCT.
+\item<+-> VOI-aware sampling for minimizing sampling regret proposed.
+\item<+-> Better sampling schemes can be developed based on principles
+  of Rational Metareasoning.
+\item<+-> UCT is not well understood. Better understanding will help
+  in developing efficient  MCTS schemes and adapting to different domains.
+\end{itemize}
+\end{frame}
+
+\begin{frame}{}
+\begin{center}
+\LARGE{Thank you!}
+\end{center}
+\end{frame}
+
 \end{document}

### 2012-02-22 slides/slides.tex
empirical evaluation in progress

@@
 
-\begin{frame}{Doing better than UCT on sets}
+\begin{frame}{Simple regret in MAB}
+
 \begin{figure}[h]
+  \begin{minipage}[c]{0.4\linewidth}
     \centering
-\includegraphics[scale=0.8]{onelevel-tree.pdf}
+    \includegraphics[scale=0.65]{onelevel-tree.pdf}\\
+    {\tiny Fig. 1: Search tree}
+  \end{minipage}
+  \begin{minipage}[c]{0.4\linewidth}
+    \centering
+    \includegraphics[scale=0.4]{flat-trilevel-k=64-uqb=8.pdf}\\
+    {\tiny Fig. 2: Regret vs number of samples}
+  \end{minipage}
+  \label{fig:mab-simple-regret}
 \end{figure}
-When an arm is selected based on the {\bf sample mean}:
 \begin{itemize}
-\item Regret of UCB decreases {\it polynomially} with $n$.
-\item Regret of $\epsilon$-greedy decreases {\it exponentially} with
-  $n$.
-\item Regret of UVB: $\max V_i$, $V_{i_{best}}=\frac {1-1/k}
-  {n_{i_{best}}},\quad V_{i_{other}}=\frac {1/k} {n_{i_{other}}}$
-  decreases exponentially with $n$, faster than $\epsilon$-greedy.
+\item For smaller numbers of
+samples, $\frac 1 2$-greedy achieves the best
+performance.
+\item For larger numbers of samples, UCB$_{\sqrt{\cdot}}$
+outperforms $\frac 1 2$-greedy. 
+\item A combination of $\frac 1
+2$-greedy  and UCB$_{\sqrt{\cdot}}$ dominates UCB over the
+whole range.
 \end{itemize}

### 2012-02-22 slides/slides.tex
empirical in progress

@@
     \includegraphics[scale=0.65]{onelevel-tree.pdf}\\
-    {\tiny Fig. 1: Search tree}
+    {\scriptsize Fig. 1: Search tree}
   \end{minipage}
@@
     \includegraphics[scale=0.4]{flat-trilevel-k=64-uqb=8.pdf}\\
-    {\tiny Fig. 2: Regret vs number of samples}
+    {\scriptsize Fig. 2: Regret vs \# of samples (32 arms)}
   \end{minipage}
@@
 whole range.
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Monte Carlo tree search}
+
+\begin{figure}[h]
+  \begin{minipage}[c]{0.4\linewidth}
+    \centering
+    \includegraphics[scale=0.45]{twolevel-tree.pdf}\\
+    {\scriptsize Fig. 1: Search tree}
+  \end{minipage}
+  \begin{minipage}[c]{0.4\linewidth}
+    \centering
+    \includegraphics[scale=0.4]{tree-identity-k=16-uqb=8.pdf}\\
+    {\scriptsize Fig. 2: Regret vs \# of samples (16 switches)}
+  \end{minipage}
+  \label{fig:mab-simple-regret}
+\end{figure}
+\begin{itemize}
+\item Either $\frac 1
+2$-greedy+UCT or UCB$_{\sqrt{\cdot}}$+UCT
+gives the lowest regret.
+\item UCB$_{\sqrt{\cdot}}$+UCT dominates UCT everywhere
+except for small numbers of instances. 
+\item The advantage of both $\frac 1
+2$-greedy+UCT and UCB$_{\sqrt{\cdot}}$+UCT grows with the number of
+arms.
 \end{itemize}

### 2012-02-22 slides/slides.tex
final changes to the slides

@@
 
-\begin{frame}{UCB vs. $\epsilon$-greedy vs UVB}
+\begin{frame}{Sailing domain}
+\begin{itemize}
+\item A square lake.
+\item A sailboat has to find the shortest path between corners.
+\item The wind changes randomly.
+\end{itemize}
+\begin{figure}[h!]
+  \begin{minipage}[b]{0.45\linewidth}
+    \centering
+    \includegraphics[scale=0.32]{costs-size=6-group=median.pdf}\\
+    {\scriptsize Fig. 1: Median cost}
+  \end{minipage}
+  \begin{minipage}[b]{0.5\linewidth}
+    \centering
+    \includegraphics[scale=0.32]{costs-size=6-group=minimum.pdf}\\
+    {\scriptsize Fig. 2: Minimum cost}
+  \end{minipage}
+  \label{fig:sailing-cost-vs-nsamples}
+\end{figure}
+\begin{itemize}
+\item UCT is always worse than $\frac 1 2$-greedy+UCT or
+UCB$_{\sqrt{\cdot}}$+UCT. 
+\item UCT is sensitive to the value of $c$: the median cost is
+much higher than the minimum cost.
+\end{itemize}
+\end{frame}
+
+\begin{frame}{VOI-aware MCTS}
 \begin{figure}[h]
   \centering
-\includegraphics[scale=0.45]{onelevel-64.pdf}
-
-64 Bernoulli arms, randomly generated
+  \includegraphics[scale=0.6]{tree-identity-k=32-uqb=8+voi.pdf}\\
 \end{figure}
-\end{frame}
-
-\begin{frame}{Doing Better Than UCT on Trees}
-Uniform sampling is useless in this tree:
-\begin{figure}[h]
-\centering
-\includegraphics[scale=0.8]{twolevel-tree.pdf}
-\end{figure}
-Rational sampling:
 \begin{itemize}
-\item first, choose an action that maximizes VOI (UVB);
-\item then, choose actions that maximize average reward (UCB).
+\item The experiments were performed on
+randomly generated trees.
+\item VOI+UCT, the scheme based on a VOI estimate,
+outperforms all other sampling schemes in this example.
 \end{itemize}
-\end{frame}{}
-
-\begin{frame}{UVT vs. VCT (UVB+UCT) vs. UCT}
-\begin{figure}[h]
-\centering
-\includegraphics[scale=0.45]{twolevel-64.pdf}
-
-64 Bernoulli arms, randomly generated
-\end{figure}
 \end{frame}

### 2012-02-22 slides/slides.tex
max->min
done/todo

@@
 \begin{itemize}
-\item Maximizes {\bf simple regret} at the {\bf first step}.
+\item Minimizes {\bf simple regret} at the {\bf first step}.
 \item Continues with UCT from the {\bf second step on}.
@@
 \item Pulls the empirically best arm with probability $\varepsilon$.
-\item Any other arm with probability $\frac {1-epsilon} {K-1}$.
+\item Any other arm with probability $\frac {1-\varepsilon} {K-1}$.
 \item Exhibits exponentially decreasing simple regret.
@@
 \begin{frame}{Summary}
+\onslide<+->
+Done:
 \begin{itemize}
-\item<+-> Improved MCTS scheme, SRCR, introduced.
-\item<+-> SRCR performs better than unmodified UCT.
-\item<+-> VOI-aware sampling for minimizing sampling regret proposed.
-\item<+-> Better sampling schemes can be developed based on principles
+\item Improved MCTS scheme, SRCR, introduced.
+\item SRCR performs better than unmodified UCT.
+\item VOI-aware sampling for minimizing sampling regret proposed.
+\end{itemize}
+\onslide<+->
+TODO:
+\begin{itemize}
+\item Better sampling schemes can be developed based on principles
   of Rational Metareasoning.
-\item<+-> UCT is not well understood. Better understanding will help
+\item UCT is not well understood. Better understanding will help
   in developing efficient  MCTS schemes and adapting to different domains.

### 2012-02-26 paper/vago.ecai/paper.tex
short paper started

--- 2012-01-28 paper/vago/paper.tex
@@
-\documentclass{article}
+\documentclass{ecai2012}
 \usepackage{enumerate}
@@
 
-\title{VOI-aware Monte-Carlo Sampling in Trees}
+\title{VOI-aware MCTS}
 \author {David Tolpin, Solomon Eyal Shimony \\
@@
 Upper bounds for the VOI are provided for pure exploration in the
-Multi-armed Bandit Problem. Sampling policies based on the upper
-bounds are suggested. Empirical evaluation of the policies and
+Multi-armed Bandit Problem. A sampling policy based on the upper
+bounds is suggested. Empirical evaluation of the policy and
 comparison to the UCB1 and UCT policies is provided
@@
 approximating assumption must be used instead.
-
-\section{Some Concentration Inequalities}
-
-Let $X_1, \ldots, X_n$ be i.i.d. random variables with values from $[0,1]$,
-$X=\frac 1 n \sum_{i=1}^n X_i$. Then 
-\begin{description}
-\item[Hoeffding inequality \rm{\cite{Hoeffding.ineq}}:] 
-\begin{equation}
-\Pr(X-\IE[X] \ge a) \le \exp ( -2a^2n)
-\label{eqn:conc-hoeffding}
-\end{equation}
-\item[Empirical Bernstein inequality \rm{(derived in Appendix~\ref{app:deriv-conc-empbernstein})}:]
-\begin{eqnarray}
-\Pr(X-\IE[X] \ge a) &\le& 2\exp \left( - \frac {a^2n} {\frac {14} {3}
-                          \frac {n} {n-1}a+2\overline\sigma_n^2}\right)\nonumber\\
-                    &\le& 2\exp \left( - \frac {a^2n} {10a+2\overline\sigma_n^2}\right)
-\label{eqn:conc-empbernstein}
-\end{eqnarray}
-where sample variance $\overline\sigma_n^2$ is
-\begin{equation}
-\overline\sigma_n^2=\frac 1 {n(n-1)} \sum_{1\le i < j\le n}(X_i-X_j)^2
-\label{eqn:sample-variance}
-\end{equation}
-\end{description}
-Bounds (\ref{eqn:conc-hoeffding}, \ref{eqn:conc-empbernstein}) are symmetrical
-around the mean. Bound~(\ref{eqn:conc-empbernstein}) is tighter than
-(\ref{eqn:conc-hoeffding}) for small $a$ and $\overline\sigma_n^2$.
 
@@
 \centering
-\includegraphics[scale=0.8]{flat.pdf}
+\includegraphics[scale=0.5]{flat.pdf}
 \caption{Random instances: regret vs. number of samples}
@@
 \centering
-\includegraphics[scale=0.8]{vct-wins.pdf}
+\includegraphics[scale=0.5]{vct-wins.pdf}
 \caption{Go: winning rate --- UCT against VCT}
@@
 \centering
-\includegraphics[scale=0.8]{ect-wins.pdf}
+\includegraphics[scale=0.5]{ect-wins.pdf}
 \caption{Go: winning rate --- UCT against ECT}
@@
 \centering
-\includegraphics[scale=0.8]{bct-wins.pdf}
+\includegraphics[scale=0.5]{bct-wins.pdf}
 \caption{Go: winning rate --- UCT against BCT}
@@
 \centering
-\includegraphics[scale=0.8]{bests-bw.pdf}
+\includegraphics[scale=0.5]{bests-bw.pdf}
 \caption{Go: best winning rate comparison}
@@
 
-\section{Empirical Bernstein Inequality}
-\label{app:deriv-conc-empbernstein}
-
-Theorem~4 in~\cite{MaurerPontil.benrstein} states that
-\begin{equation}
-\Pr\left(\IE[X]-\overline X_n
-    \ge \sqrt { \frac {2\overline\sigma_n^2 \ln 2/\delta} n } + \frac {7 \ln 2/\delta} {3(n-1)}\right)\le \delta
-\end{equation}
-Therefore
-\begin{equation}
-\Pr\left(\IE[X]-\overline X_n \ge
-              \sqrt { \left(\frac {7 \ln 2/\delta} {3(n-1)}\right)^2+\frac {2\overline\sigma_n^2 \ln 2/\delta} n }
-                      + \frac {7 \ln 2/\delta} {3(n-1)}\right)
-   \le \delta.
-\end{equation}
-$a=\sqrt { \left(\frac {7 \ln 2/\delta} {3(n-1)}\right)^2+\frac {2\overline\sigma_n^2 \ln 2/\delta} n } + \frac {7 \ln 2/\delta} {3(n-1)}$
- is a root of square equation $a^2-a\frac {14 \ln 2/\delta} {3(n-1)} -\frac {2\overline\sigma_n^2 \ln 2/\delta} n=0$
-which, solved for $\delta\triangleq\Pr(\IE[X]-\overline X_n\ge a)$,
-gives
-\begin{equation*}
-\Pr(\IE[X]-\overline X_n\ge a)\le 2\exp \left( - \frac {na^2} {\frac {14} {3} \frac {n} {n-1}a+2\overline\sigma_n^2}\right)
-\end{equation*}
-Other derivations, giving slightly different results, are possible.
-
-\section{Better Hoeffding-Based Bound on Value of Perfect Information}
-\label{app:better-hoeffding-bound}
-
-Bounds (\ref{eqn:bound-perf-hoeffding}, \ref{eqn:bound-blnk-hoeffding}) have the form
-\begin{equation}
-  \Lambda_i \le \hat\Lambda_i=\left\{
-  \begin{array}{l l}
-    (\overline X_\beta-B)\exp\left(- \gamma(\overline X_i - \overline X_\beta)^2 n_i \right) & \mbox{{\rm if} $i=\alpha$} \\
-    (A-\overline  X_\alpha)\exp\left(- \gamma(\overline X_\alpha - \overline X_i)^2 n_i \right)  &  \mbox{\rm otherwise}
-  \end{array} \right.
-\label{eqn:bound-generic-hoeffding}
-\end{equation}
-where $A$ is the upper bound on the posterior sample mean $\overline
-X_i'$ if $i\ne \alpha$, $B$ is the lower bound on $\overline X_i'$ if
-$i=\alpha$. For $i=\alpha$, the bound can be improved by selecting an
-intermediate point $y$, $B < y < \overline X_\beta$, and computing the bound
-as the sum of two parts: 
-\begin{itemize}
-\item $\overline X_\beta-y$ multiplied by the probability that
-  $\overline X_i' \le \overline X_\beta$;
-\item $\overline X_\beta-B$ multiplied by the probability that $\overline X_i'\le
-  y$.
-\end{itemize}.
-\begin{equation}
-\Lambda_{i|i=\alpha} \le \hat \Lambda_{i|i=\alpha} =
- (\overline X_\beta-y)\exp\left(-\gamma (\overline X_i-\overline X_\beta)^2n_i\right)
- +(\overline X_\beta-B)\exp\left(-\gamma (\overline X_i-y)^2n_i\right)
-\label{eqn:bound-twopart-hoeffding-alpha}
-\end{equation}
-If $y$ that minimizes $\hat\Lambda_{i|i=\alpha}$ exists, the minimum is achieved when
- $\frac {d\hat\Lambda_{i|i=\alpha}} {dy}=0$, that
-is, when $y$ is the root of the following equation:
-\begin{equation}
-  2\gamma (\overline X_\beta-B)(\overline X_i-y)n_i = \exp\left(-\gamma \left(\frac {\overline X_i-\overline X_\beta}
-    {\overline X_i-y}\right)^2n_i\right)
-\end{equation}
-The derivation for the case $i\ne \alpha$ is obtained by substitution $1-A, 1-\overline X_i, 1-\overline X_\alpha, 1-y$ instead of
-$B, \overline X_i, \overline X_\beta, y$ into (\ref{eqn:bound-twopart-hoeffding-alpha}):
-\begin{equation}
-\Lambda_{i|i\ne\alpha} \le \hat\Lambda_{i|i\ne\alpha} = (y-\overline X_\alpha)\exp\left(-\gamma(\overline X_\alpha-\overline
-  X_i)^2 n_i\right)+(A-\overline X_\alpha)\exp\left(-\gamma (y-\overline X_i)^2n_i\right)
-\label{eqn:bound-twopart-hoeffding-rest}
-\end{equation}
-A closed-form solution for $y$ cannot be obtained, but given
-$A, B, \overline X_\alpha, \overline X_\beta, n$, the value of $y$ can be efficiently
-computed.
-
-\bibliographystyle{plain}
+\bibliographystyle{ecai2012}
 \bibliography{refs}

### 2012-02-26 paper/vago.ecai/paper.tex
short paper underway

@@
 \title{VOI-aware MCTS}
-\author {David Tolpin, Solomon Eyal Shimony \\
-Department of Computer Science, \\
-Ben-Gurion University of the Negev, Beer Sheva, Israel \\
-\{tolpin,shimony\}@cs.bgu.ac.il}
+%\author {David Tolpin, Solomon Eyal Shimony \\
+%Department of Computer Science, \\
+%Ben-Gurion University of the Negev, Beer Sheva, Israel \\
+%\{tolpin,shimony\}@cs.bgu.ac.il}
 
@@
 
-\begin{dfn} The \textbf{blinkered estimate} of intrinsic VOI information of the
-  $i$th arm is the intrinsic VOI of pulling the $i$th arm for the
-  remaining budget.
-\end{dfn}
-
 \begin{thm} Denote the current number of samples of the $i$th arm by
-  $n_i$. The blinkered estimate $\Lambda_i^b$ of intrinsic value of
-  information of pulling the $i$th arm for the remaining budget of $N$
+  $n_i$. The intrinsic value of
+  information $\Lambda_i^b$ of pulling the $i$th arm for the remaining budget of $N$
   samples is bounded from above as
@@
 Appendix~\ref{app:hoeffding-prob-bounds-proof}) is based on the
-Hoeffding inequality (\ref{eqn:conc-hoeffding}):
+Hoeffding inequality \cite{Hoeffding.ineq}.
 \begin{thm} The probabilities in equations (\ref{eqn:thm-vopi},
@@
 example, based on the empirical Bernstein inequality
-(\ref{eqn:conc-empbernstein}) or through a more careful application of
-the Hoeffding inequality (Appendix~\ref{app:better-hoeffding-bound}).
+ or through a more careful application of
+the Hoeffding inequality.
 
@@
 The experiments compare the UCT algorithm \cite{Kocsis.uct} with
-modified versions of UCT in which the samples at the first step are
-selected according to their VOI estimates. Three blinkered VOI
-estimates were used for the comparison:
-\begin{description}
-\item[VCT:] based on Hoeffding inequality (\ref{eqn:conc-hoeffding});
-\item[ECT:] based on Hoeffding inequality with an intermediate point (see
-Appendix~\ref{app:better-hoeffding-bound});
-\item[BCT:] based on empirical Bernstein inequality
-(\ref{eqn:conc-empbernstein}).
-\end{description}
-For simplicity, the bounds on probability for an unlimited number
-of samples were used in the estimate formulas. The estimates where
-computed as follows:
-\begin{eqnarray}\
-  \Lambda_i^{VCT}& =&\left\{
-  \begin{array}{l l}
-    \frac {\overline X_\beta} {n_i} \cdot \exp\left(- 2(\overline X_i - \overline X_\beta)^2 n_i\right)
-      & \mbox{{\rm if} $i=\alpha$} \\
-    \frac {(1-\overline  X_\alpha)} {n_i}\cdot \exp\left(- 2(\overline X_\alpha - \overline X_i)^2 n_i\right)
-      &  \mbox{\rm otherwise}
-  \end{array}  \right.\nonumber\\
-  \Lambda_i^{ECT}& =&\min\left(\Lambda_i^{VCT}, \min_y\left\{
-  \begin{array}{l l}
-    \frac {\overline X_\beta-y} {n_i}\cdot \exp\left(-2 (\overline X_i-\overline X_\beta)^2n_i\right)
-    +\frac {\overline X_\beta} {n_i}\cdot \exp\left(-2 (\overline X_i-y)^2n_i\right)
-      & \mbox{{\rm if} $i=\alpha$} \\
-\frac {y-\overline X_\alpha} {n_i}\cdot \exp\left(-2(\overline X_\alpha-\overline X_i)^2n_i\right)
-    +\frac {1-\overline X_\alpha} {n_i}\cdot \exp\left(-2 (y-\overline X_i)^2n_i\right)
-      &  \mbox{\rm otherwise}
-  \end{array} \right.\right) \nonumber\\
-  \Lambda_i^{BCT}& =&\min\left(\Lambda_i^{VCT},\left\{
-  \begin{array}{l l}
-    \frac {\overline X_\beta} {n_i}\cdot 2\exp\left(- \frac {(\overline X_i - \overline X_\beta)^2 n_i} {\frac {14} 3 \frac {n_i} {n_i-1}(\overline X_i - \overline X_\beta) + 2 \overline X_i(1-\overline X_i)} \right)
-      & \mbox{{\rm if} $i=\alpha$} \\
-    \frac {(1-\overline  X_\alpha)} {n_i}\cdot 2\exp\left(- \frac {(\overline X_\alpha - \overline X_i)^2 n_i} {\frac {14} 3 \frac {n_i} {n_i-1}(\overline X_\alpha - \overline X_i) + 2 \overline X_i(1-\overline X_i)} \right)
-      &  \mbox{\rm otherwise}
-  \end{array} \right.\right)
-\label{eqn:emp-estimates}
-\end{eqnarray}
+a modified version of UCT in which the samples at the first step are
+selected according to their VOI estimates.
 
@@
 \end{figure}
-\begin{figure}[t]
-\centering
-\includegraphics[scale=0.5]{ect-wins.pdf}
-\caption{Go: winning rate --- UCT against ECT}
-\label{fig:uct-against-ect}
-\end{figure}
-\begin{figure}[t]
-\centering
-\includegraphics[scale=0.5]{bct-wins.pdf}
-\caption{Go: winning rate --- UCT against BCT}
-\label{fig:uct-against-bct}
-\end{figure}
+
 The engines were compared on the 9x9 board, for 5000, 7000, 1000, and
 15000 samples (game simulations) per ply, each experiment was repeated
-1000 times. Figures~\ref{fig:uct-against-vct}--\ref{fig:uct-against-bct}
-show the winning rate of UCT against VCT, ECT, BCT, correspondingly,
+1000 times. Figure~\ref{fig:uct-against-vct}
+show the winning rate of UCT against VCT
 vs. the stopping threshold (if the maximum VOI of a sample is below
@@
 compared to UCT.
-
-As the results show, without VOI-aware sample redistribution the
-VOI-aware policies are at best equal to UCT for lower numbers of
-samples per ply (5000, 7000), and the advantage of the VOI-aware policies
-for larger number of samples is rather modest (e.g., ECT wins in 55\%
-of games for 10000 samples per ply). For the stopping threshold of
-$10^{-6}$, however, the VOI-aware policies are almost always better
-than UCT (only VCT for 5000 samples wins in less than 50\% of games),
-with stronger policies (ECT, VCT) reaching the winning rate of 66\%.
 
@@
 increasing the maximum number of samples in a future state is low.
-
-\begin{figure}[t]
-\centering
-\includegraphics[scale=0.5]{bests-bw.pdf}
-\caption{Go: best winning rate comparison}
-\label{fig:best-winning-rate}
-\end{figure}
-Figure~\ref{fig:best-winning-rate} presents relative performance of
-different VOI-aware policies depending on the number of samples per
-ply by comparing the best achieved winning rates against UCT (again,
-the lower the bar, the better is the policy). Both ECT (based on the
-Hoeffding inequality with an intermediate point) and BCT (based on the empirical
-Bernstein inequality) are better than VCT. The difference between ECT
-and VCT is most prominent for smaller number of samples, and the
-winning rate of ECT increases relatively slowly with the number of
-samples. In contrast, BCT is worse than ECT and only slightly better
-than VCT for smaller number of samples, but the winning rate rapidly
-increases with the number of samples, and for 10000 and 15000 samples
-per ply BCT is the best of the policies with the winning rate reaching
-67\%.
 
@@
 Equations (\ref{eqn:probound-perf-hoeffding}) is a direct
-application of the Hoeffding inequality (\ref{eqn:conc-hoeffding}).
+application of the Hoeffding inequality \cite{Hoeffding.ineq}.
 

### 2012-02-26 paper/vago.ecai/paper.tex
2 and a half

@@
 \documentclass{ecai2012}
+\usepackage{times}
 \usepackage{enumerate}
@@
 in the regret compared to selecting an arm without pulling any arm at
-all. The \textit{myopic} VOI estimate is of limited applicability to
-Monte-Carlo sampling, since the effect of a single sample is small,
-and the myopic VOI estimate will often be zero, resulting in premature
-termination of the search. However, $\Lambda_i$ can be estimated as the intrinsic 
-value of perfect information $\Lambda_i^p$ about the mean reward of the $i$th arm. Two
-cases are possible:
+all. Two cases are possible:
 \begin{itemize}
@@
 \end{itemize}
-$\Lambda_i^p$ can be bounded from above as the probability that a
-different arm is selected, multiplied by the
-maximum possible increase in the reward:
-\begin{thm} The intrinsic value of perfect information $\Lambda_i^p$ about the $i$th arm is
-  bounded from above as
-\begin{equation}
-  \Lambda_i^p \le \left\{
-  \begin{array}{l l}
-    \overline X_\beta \Pr(\IE[X_i] \le \overline X_\beta) & \mbox{{\rm if} $i=\alpha$} \\
-    (1-\overline X_\alpha)\Pr(\IE[X_i] \ge \overline X_\alpha) & \mbox{\rm otherwise}
-  \end{array} \right.
-\label{eqn:thm-vopi}
-\end{equation}
-\label{thm:vopi}
-\end{thm}
-
-\begin{proof} For the case $i\ne \alpha$, the probability that the perfect
-  information about the $i$th arm changes the final choice is
-  $\Pr(\IE[X_i] \ge \overline X_\alpha)$. $\IE[X_i] \le 1$ by definition,
-  therefore the maximum increase in the expected reward is
-  $(1-\overline X_\alpha)$. Thus the intrinsic value of perfect
-  information is at most $\Pr(\IE[X_i] \ge \overline
-  X_\alpha)(1-\overline X_\alpha)$.
-  Proof for the case $i=\alpha$ is similar.
-\end{proof}
-
-The search time is finite, and in a simple case the \textit{search
-  budget} is specified as the maximum number of samples.
-An estimate based on the perfect intrinsic VOI does not take in
-consideration the remaining number of samples. Given two arms
-with the same intrinsic perfect VOI, the VOI
-estimate of the arm pulled fewer times so far should be higher.
-
+The \textit{myopic} VOI estimate is of limited applicability to
+Monte-Carlo sampling, since the effect of a single sample is small,
+and the myopic VOI estimate will often be zero, resulting in premature
+termination of the search. However, $\Lambda_i$ can be estimated as the intrinsic 
+value of information $\Lambda_i^b$ of pulling the $i$th arm for the
+rest of the budget of $N$ samples:
 \begin{thm} Denote the current number of samples of the $i$th arm by
-  $n_i$. The intrinsic value of
-  information $\Lambda_i^b$ of pulling the $i$th arm for the remaining budget of $N$
-  samples is bounded from above as
+  $n_i$. $\Lambda_i^b$ is bounded from above as
 \begin{equation}
@@
 \end{equation}
-where $\overline X_i'$ is the sample mean of the $i$th arm after $n_i+N$ 
-samples.
+where $\overline X_i'$ is the sample mean of the $i$th arm after $n_i+N$ samples.
 \label{thm:be}
@@
 
-\begin{proof}
-The proof is similar to the proof of Theorem~\ref{thm:vopi}, with
-$\overline X'_i$ substituted instead of $\IE[X_i]$, and, if
-$i\ne\alpha$, an upper bound on $X_i'$ is $X_\alpha+(1-X_\alpha)
-\frac N {N+n_i}$; if $i=\alpha$, a lower bound on $X_i'$ is
-$X_\beta - X_\beta \frac N {N+n_i}$.
-\end{proof}
-
-The probabilities in equations (\ref{eqn:thm-vopi}, \ref{eqn:thm-be})
-can be bounded from above using concentration inequalities. In
-particular, Theorem~\ref{thm:hoeffding-prob-bounds} (proved in 
-Appendix~\ref{app:hoeffding-prob-bounds-proof}) is based on the
+The probabilities can be bounded from above using the
 Hoeffding inequality \cite{Hoeffding.ineq}.
-\begin{thm} The probabilities in equations (\ref{eqn:thm-vopi},
-\ref{eqn:thm-be}) are bounded from above as
-\begin{eqnarray}
-\Pr(\IE[X_i] \le \overline X_\beta|i=\alpha)&
-   \le & \exp(-2 (\overline X_i - \overline X_\beta)^2 n_i)\nonumber\\
-\Pr(\IE[X_i] \ge \overline X_\alpha|i\ne\alpha)&
-   \le & \exp(-2 (\overline X_\alpha - \overline X_i)^2 n_i)
-\label{eqn:probound-perf-hoeffding}
-\end{eqnarray}
-\vspace{-1.5\baselineskip}
-\begin{eqnarray}
-\Pr(\overline X_i' \le \overline X_\beta|i=\alpha) &
-   \le & 2\exp\left(- 2\left(\frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}}(\overline X_i - \overline X_\beta)\right)^2 n_i \right)
-   \le 2\exp\left(- 1.37(\overline X_i - \overline X_\beta)^2 n_i \right) \nonumber\\
-\Pr(\overline X_i' \ge \overline X_\alpha|i\ne\alpha) &
-   \le & 2\exp\left(- 2\left(\frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}} (\overline X_\alpha - \overline X_i)\right)^2 n_i \right)
-    \le 2\exp\left(- 1.37(\overline X_\alpha - \overline X_i)^2 n_i \right)
+\begin{thm} The probabilities in equations (\ref{eqn:thm-be}) are bounded from above as
+\begin{align}
+\Pr&(\overline X_\alpha' \le \overline X_\beta)
+  \le 2\exp\left(- \varphi(n_\alpha)(\overline X_\alpha - \overline X_\beta)^2 n_\alpha
+   \right)\nonumber\\
+\Pr&(\overline X_i' \ge \overline X_\beta)
+   \le 2\exp\left(- \varphi(n_i) (\overline X_\alpha -\overline  X_i)^2 n_i \right)
 \label{eqn:probound-blnk-hoeffding}
-\end{eqnarray}
+\end{align}
+where $\varphi(n)=2(\frac {1+n/N} {1+\sqrt {n/N}})^2 \ge 1.37$.
 \label{thm:hoeffding-prob-bounds}
@@
 
-\begin{crl} An upper bound on the intrinsic value of perfect information is obtained by substituting
-(\ref{eqn:probound-perf-hoeffding}) into (\ref{eqn:thm-vopi}):
-\begin{equation}
-  \Lambda_i^p \le \hat\Lambda_i^p=\left\{
-  \begin{array}{l l}
-    \overline X_\beta\cdot \exp\left(- 2(\overline X_i - \overline X_\beta)^2 n_i \right) & \mbox{{\rm if} $i=\alpha$} \\
-    (1-\overline  X_\alpha)\cdot \exp\left(- 2(\overline X_\alpha - \overline X_i)^2 n_i \right)  &  \mbox{\rm otherwise}
-  \end{array} \right.
-\label{eqn:bound-perf-hoeffding}
-\end{equation}
-\label{crl:bound-perf-hoeffding}
-\end{crl}
 \begin{crl}
-An upper bound on the blinkered VOI estimate with ordering independent of $N$ is obtained
+An upper bound on the VOI estimate $\Lambda_i^b$ is obtained
 by substituting (\ref{eqn:probound-blnk-hoeffding}) into (\ref{eqn:thm-be}).
@@
 \end{crl}
-Bound~\ref{eqn:bound-blnk-hoeffding} is generally tighter when $N$ is
-small relative to $n_i$. Estimate $\hat\Lambda_i^b$ can be viewed as a
+Estimate $\hat\Lambda_i^b$ can be viewed as a
 relaxation of the myopic assumption such that sequences consisting
-of more than one action are taken into considertion. Better bounds can
-be obtained through tighter estimates on the probabilities, for
-example, based on the empirical Bernstein inequality
- or through a more careful application of
-the Hoeffding inequality.
+of more than one action are taken into considertion.
 
-\section{VOI-based Sampling Control}
-
-\subsection{Selection Criterion}
+\subsection{VOI-based Sample Allocation}
 
@@
 the highest VOI should be pulled at each step. The upper bounds
-established in Theorems~\ref{thm:vopi},~\ref{thm:be} can be used
-as VOI estimates. Blinkered VOI estimates~(\ref{eqn:thm-be},
- \ref{eqn:bound-blnk-hoeffding}) can be viewed
-as approximations of VOI estimates for myopic or
-receding horizon policies. As illustrated by the empirical evaluation
+established in Corollary~\ref{crl:bound-blnk-hoeffding} can be used
+as VOI estimates. As illustrated by the empirical evaluation
 (Section~\ref{sec:empirical-evaluation}), estimates based on upper
@@
 heuristic algorithms.
-
-\subsection{Termination Condition}
-\label{sec:control-termination-condition}
-
-The simplest termination condition for a sampling policy is the
-budget---a fixed number of samples performed before a decision is
-made. When a sample has a cost commensurable with the value of
-information of a measurement, an upper bound on the intrinsic VOI can
-be used to stop the sampling if the intrinsic VOI of any action
-is less than the cost of sampling $C$:
-\begin{equation}
-\mbox{stop if } \max_i \Lambda_i \le C
-\end{equation}
-Blinkered VOI estimates~(\ref{eqn:thm-be},
-\ref{eqn:bound-blnk-hoeffding}) include the remaining budget $N$ as a
-factor, but given the cost of a single sample $c$, the cost of the
-remaining samples accounted for in estimating the intrinsic VOI is
-$C=cN$. $N$ can be dropped on both sides of the inequality,
-and a viable stopping condition is
-\begin{equation}
-\begin{array}{c r r l l}
-&\frac 1 N \Lambda_\alpha^b
-\le\frac {\overline  X_\beta} {n_\alpha}\Pr(\overline X_\alpha'\le\overline X_\beta)&\le& c
-&\nonumber\\
-\mbox{and}&&&&\nonumber\\
-&\frac 1 N \max_i\Lambda_i^b
-\le \max_i\frac {(1-\overline  X_\alpha)} {n_i}\Pr(\overline X_i'\ge\overline X_\alpha)&\le& c
- & \forall i: i\ne\alpha
-\end{array}
-\label{eqn:stopping-blnk}
-\end{equation}
-The empirical evaluation (Section~\ref{sec:empirical-evaluation})
-confirms viability of this stopping condition and illustrates the
-influence of the sample cost $c$ on the performance of
-the sampling policy.
-
-\subsection{Sample Redistribution in Trees}
-\label{sec:control-redistribution}
-
-Monte-Carlo tree search \cite{Chaslot.montecarlo} selects at each
-search step an action that appears to be the best based on outcomes
-of \textit{search rollouts}. Two different criteria are employed in
-action selection during a rollout:
-\begin{itemize}
-\item at the first step, the value of partial
-information should be maximized to minimize the \textit{simple} regret
-of the selection at the current root node;
-\item as the goal of sampling in deeper tree nodes is estimating the
-value of a node, rather than selection, it makes sense to minimize
-the \textit{cumulative} regret of the rollouts from the second step
-onwards.
-\end{itemize}
-
-A natural approach would be to apply an algorithm that minimizes the
-simple regret in Multi-armed bandits at the first step, and the
-cumulative regret during the rest of the rollout.  However, this
-approach assumes that the information obtained from rollouts in the
-current state is discarded after an action is selected. In practice,
-many successful Monte-Carlo tree search algorithms re-use rollouts
-originating from an earlier search state and passing through the
-current search state; thus, the value of information of a rollout is
-determined not just by the influence on the choice of the action at
-the current state, but also by the influence on the choice at future
-search states, provided the rollout passes through the search states
-which will be visited.
-
-One way to account for this re-use would be to incorporate the
-`future' value of information into a VOI estimate. However, this 
-approach appears to be complicated. Alternatively, one can behave
-myopically in search tree depth:
-\begin{enumerate}
-\item estimate VOI as though the information is discarded after each step;
-\item stop early if the VOI is below a certain threshold
-   (see Section~\ref{sec:control-termination-condition});
-\item save the unused sample budget for search in future state, such that
-   if the nominal budget is $N$, and the unused budget in the last state
-   is $N_u$, the search budget in the next state will be $N+N_u$.
-\end{enumerate}
-In this approach, the cost of a sample in the current state is the
-VOI of increasing the budget of a future state by one sample.  It is
-unclear whether this cost can be accurately estimated, but supposedly
-a fixed value for a given problem type and algorithm implementation
-would work. Indeed, the empirical evaluation (Section~\ref{sec:emp-go})
-confirms that stopping and sample redistribution based on a learned
-fixed cost  substantially improve the performance of the VOI-based
-sampling policy in game tree search.
 
@@
 The sampling policies are first compared on random Multi-armed bandit 
-problem instances, where the Monte-Carlo Tree Search algorithm
-is reduced to the sample selection policy at the first step: UCT is 
-actually UCB1. In the problem of pure exploration in Multi-armed bandits
-\cite{Bubeck.pure}, the smaller the exploration factor, the higher
-the expected simple regret is; indeed, UCT exhibits the lowest simple
-regret in the experiments for the default exploration factor $Cp=\frac
-{\sqrt 2} 2$ \cite{Kocsis.uct}.
-
-Figure~\ref{fig:random-instances} shows experiment results for
+problem instance. Figure~\ref{fig:random-instances} shows experiment results for
 randomly-generated Multi-armed bandits with 32 Bernoulli arms, with
@@
 step of $2$. The experiment for each number of samples was repeated
-20000 times. UCT is worse than any of the VOI-aware sampling policies,
-for example, for 384 samples the simple regret of UCT is $\approx 40$
-times greater than of VCT, and $\approx 100$ times greater than of
-ECT. On average, it takes $2.5$ times as many samples for UCT to reach
-the regret of VCT. Different numbers of arms and distributions of
-means give similar results.
+20000 times. UCT is always considerably worse than the VOI-aware sampling policy.
 
@@
 
-For the experiments, a modified version of Pachi \cite{Braudis.pachi},
-a state of the art Go program, was used:
-\begin{itemize}
-\item The UCT engine of Pachi was extended with VOI-aware sampling
-  policies at the first step. 
-\item The stopping condition for the VOI-aware policies was
-  modified and based solely on the sample cost, specified as
-  a constant parameter. The heuristic stopping condition for the
-  original UCT policy was left unchanged.
-\item The time-allocation mode based on the fixed number of samples
-  was modified for \textit{both the original UCT policy and the VOI-aware
-  policies} such that 
-  \begin{enumerate}
-    \item the same number of samples is available to
-      the agent at each step, independently of the number of pre-simulated
-      games;  
-    \item if samples were unused at the current step,
-      they become available at the next step. 
-  \end{enumerate}
-\end{itemize}
-While the UCT engine is not the most powerful engine of Pachi, it is
-still a strong player. On the other hand, additional features of more
-advanced engines would obstruct the Monte-Carlo sampling phenomena
-which are the subject of the experiment.
-
+A modified version of Pachi \cite{Braudis.pachi}, a state of the art
+Go program, was used for the experiments. The UCT engine was extended
+with a VOI-aware sampling policy, and the time allocation mode
+ensuring that both the original UCT policy and the VOI-aware policy
+use the same average number of samples per node. While the UCT engine
+is not the most powerful engine of Pachi, it is still a strong
+player. On the other hand, additional features of more advanced
+engines would obstruct the Monte-Carlo sampling phenomena which are
+the subject of the experiment.
 \begin{figure}[t]
@@
 \caption{Go: winning rate --- UCT against VCT}
-\label{fig:uct-against-vct}
+\label{fig:vct-against-uct}
 \end{figure}
-
 The engines were compared on the 9x9 board, for 5000, 7000, 1000, and
 15000 samples (game simulations) per ply, each experiment was repeated
-1000 times. Figure~\ref{fig:uct-against-vct}
-show the winning rate of UCT against VCT
-vs. the stopping threshold (if the maximum VOI of a sample is below
-the threshold, the simulation is stopped, and a move is chosen). Each
-curve in the figures corresponds to a certain number of samples per
-ply. The lower the curve, the better is the corresponding policy
-compared to UCT.
-
-In agreement with intuition
-(Section~\ref{sec:control-redistribution}), VOI-based stopping and
-sample redistribution is most influential for medium numbers of
-samples per ply. When the maximum number of samples is too low, early
-stopping would result in poorly selected moves. On the other hand,
-when the maximum number of samples is sufficiently high, the VOI of
-increasing the maximum number of samples in a future state is low.
+1000 times. Figure~\ref{fig:vct-against-uct}
+shows the winning rate of VOI+UCT against UCT vs. the number of
+samples. For most numbers of samples per node, VO+UCT outperforms UCT.
 
@@
 
-\section{Proof of Theorem~\ref{thm:hoeffding-prob-bounds}}
-\label{app:hoeffding-prob-bounds-proof}
-
-\begin{proof}[]
-Equations (\ref{eqn:probound-perf-hoeffding}) is a direct
-application of the Hoeffding inequality \cite{Hoeffding.ineq}.
-
-Equations (\ref{eqn:probound-blnk-hoeffding}) follow from the
-observation that if $i\ne\alpha$, $\overline X_i'>\overline X_\alpha$
-if and only if the mean $\overline X_i^+$ of $N$ samples from $n_i+1$
-to $N$ is at least $\overline X_\alpha+(\overline X_\alpha-\overline
-X_i)\frac {n_i} N$.
-
-For any $\delta$, the probability that $\overline X_i'$ is greater
-than $\overline X_\alpha$ is less than the probability that
-$\IE[X_i]\ge\overline X_i+\delta$
-\emph{or} $\overline X_i^+\ge \IE[X_i]+\overline X_\alpha
-- \overline X_i - \delta +(\overline X_\alpha - \overline X_i)\frac {n_i} N$,
-thus, by the union bound, less than the sum of the probabilities:
-\begin{equation}
-\Pr(\overline X_i'\ge \overline X_\alpha)\le
-  \Pr(\IE[X_i]-\overline X_i \ge \delta)+
-  \Pr\left(\overline X_i^+ - \IE[X_i] \ge \overline X_\alpha
-           - \overline X_i - \delta +(\overline X_\alpha - \overline X_i)\frac {n_i} N\right)
-\end{equation}
-Bounding the probabilities on the right-hand side using the Hoeffding
-inequality, obtain:
-\begin{equation}
-\Pr(\overline X_i'\ge \overline X_\alpha)\le 
-    \exp(-2\delta^2n_i)+
-         \exp\left(-2\left((\overline X_\alpha 
-         - \overline X_i)\left(1+\frac {n_i} N\right)
-         - \delta\right)^2N\right)
-\label{eqn:app-hoeffding-le-maxexp}
-\end{equation}
-Find $\delta$ for which the two terms on the right-hand side of
-(\ref{eqn:app-hoeffding-le-maxexp}) are equal:
-\begin{equation}
-\exp(-\delta^2n) = \exp(-2\left((\overline X_\alpha - \overline X_i)(1+\frac {n_i} N) - \delta\right)^2N)\label{eqn:app-hoeffding-eq-exp}
-\end{equation}
-Solve (\ref{eqn:app-hoeffding-eq-exp}) for $\delta$: $\delta=\frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}} (\overline X_\alpha
-- \overline X_i) \ge (\sqrt 2 - 1)(\overline X_\alpha-\overline X_i)$. Substitute $\delta$ into 
-(\ref{eqn:app-hoeffding-le-maxexp}) and obtain
-\begin{eqnarray}
-\Pr(\overline X_i'\ge \overline X_\alpha) 
-& \le & 2\exp\left(-2\left( \frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}}
-                          (\overline X_\alpha - \overline X_i)\right)^2 n_i\right)\nonumber \\
-& \le & 2\exp(-2(\sqrt 2 - 1)^2(\overline X_\alpha - \overline X_i)^2n_i)
- \le 2\exp(-1.37(\overline X_\alpha - \overline X_i)^2n_i)
-\end{eqnarray}
-Derivation for the case $i=\alpha$ is similar.
-\end{proof}
-
 \bibliographystyle{ecai2012}

### 2012-02-26 paper/vago.ecai/paper.tex
2 and a half

@@
 Go program, was used for the experiments. The UCT engine was extended
-with a VOI-aware sampling policy, and the time allocation mode
-ensuring that both the original UCT policy and the VOI-aware policy
-use the same average number of samples per node. While the UCT engine
-is not the most powerful engine of Pachi, it is still a strong
+with a VOI-aware sampling policy, and a time allocation mode ensuring
+that both the original UCT policy and the VOI-aware policy use the
+same average number of samples per node was added. While the UCT
+engine is not the most powerful engine of Pachi, it is still a strong
 player. On the other hand, additional features of more advanced

### 2012-02-26 paper/vago.ecai/paper.tex
making it shorter

@@
 Upper bounds for the VOI are provided for pure exploration in the
-Multi-armed Bandit Problem. A sampling policy based on the upper
-bounds is suggested. Empirical evaluation of the policy and
-comparison to the UCB1 and UCT policies is provided
-on random problem instances as well as on the Go game.
+MCTS. A sampling policy based on the upper bounds is
+suggested. Empirical evaluation of the policy and comparison to UCB1
+and UCT is performed on random bandit instances as well as on Computer
+Go.
 \end{abstract}
@@
 
-\subsection{VOI-based Sample Allocation}
+\section{VOI-based Sample Allocation}
 
@@
 
-The experiments compare the UCT algorithm \cite{Kocsis.uct} with
-a modified version of UCT in which the samples at the first step are
-selected according to their VOI estimates.
-
 \subsection{Selecting The Best Arm}
@@
 The sampling policies are first compared on random Multi-armed bandit 
-problem instance. Figure~\ref{fig:random-instances} shows experiment results for
+problem instances. Figure~\ref{fig:random-instances} shows results for
 randomly-generated Multi-armed bandits with 32 Bernoulli arms, with
 the mean rewards of the arms distributed uniformly in the range $[0,
-  1]$, for a range of sample budgets $32..2048$, with multiplicative
+  1]$, for a range of sample budgets $32..1024$, with multiplicative
 step of $2$. The experiment for each number of samples was repeated
-20000 times. UCT is always considerably worse than the VOI-aware sampling policy.
+10000 times. UCB1 is always considerably worse than the
+VOI-aware sampling policy.
 
@@
 
-One search domain in which Monte-Carlo tree search based on the UCT
-sampling policy has been particularly successful is playing the game Go
-\cite{Gelly.mogo}. This series of experiments compares
-the winning rate of UCT against VOI-aware policies (VCT, ECT, BCT)
-in Go and investigates the effect of early stopping and sample
-redistribution.
-
-A modified version of Pachi \cite{Braudis.pachi}, a state of the art
+The policies were also compared on Computer Go, a  search domain
+in which UCT-based MCTS has been particularly successful
+\cite{Gelly.mogo}. A modified version of Pachi \cite{Braudis.pachi}, a state of the art
 Go program, was used for the experiments. The UCT engine was extended
@@
 that both the original UCT policy and the VOI-aware policy use the
-same average number of samples per node was added. While the UCT
+same average number of samples per node was added. (While the UCT
 engine is not the most powerful engine of Pachi, it is still a strong
-player. On the other hand, additional features of more advanced
-engines would obstruct the Monte-Carlo sampling phenomena which are
-the subject of the experiment.
+player; on the other hand, additional features of more advanced
+engines would obstruct the MCTS phenomena which are the subject of
+the experiment.)
 \begin{figure}[t]
@@
 The engines were compared on the 9x9 board, for 5000, 7000, 1000, and
-15000 samples (game simulations) per ply, each experiment was repeated
+15000 samples per ply, each experiment was repeated
 1000 times. Figure~\ref{fig:vct-against-uct}
 shows the winning rate of VOI+UCT against UCT vs. the number of
-samples. For most numbers of samples per node, VO+UCT outperforms UCT.
+samples. For most numbers of samples per node, VOI+UCT outperforms UCT.
 

### 2012-02-26 paper/vago.ecai/paper.tex
short vago paper in progress

@@
 \usepackage{color}
-\usepackage{geometry}
 
-\geometry{margin=1.2in}
+\ecaisubmission
 
@@
 \begin{abstract}
-Upper bounds for the VOI are provided for pure exploration in the
-MCTS. A sampling policy based on the upper bounds is
-suggested. Empirical evaluation of the policy and comparison to UCB1
-and UCT is performed on random bandit instances as well as on Computer
-Go.
+An MCTS sampling policy based on VOI estimates of rollouts is
+suggested. Empirical evaluation of the policy and comparison to
+UCB1 and UCT is performed on random Multi-armed Bandit instances
+as well as on Computer Go.
 \end{abstract}
@@
 \section{Introduction and Definitions}
+
+TODO: start with MCTS, then go to Bandits.
 
@@
 \cite{Russell.right} suggest that at each step the arm with the
-greatest value of information (VOI) estimate must be pulled, and the
-sampling must be stopped and a decision must be made when no arm has
-positive VOI estimate.
-
-In order to estimate the VOI of pulling an arm, a sampling model is
-needed.  The generic sampling model used is: 
-\begin{itemize}
-\item The actual value of an arm has a certain distribution.
-\item Pulling an arm is an i.i.d. noisy measurement of the actual
-  value of the arm.
-\end{itemize}
-In \cite{HayRussell.MCTS}, Section 4, two instances of this model are
-examined:
-\begin{enumerate}[a)]
-\item uniformly distributed prior of the value, with Bernouli
-  sampling, and 
-\item normally distributed prior of the value, with normally
-  distributed sampling.
-\end{enumerate}
-Here we examine the same general setting (i.i.d. samples), but instead
-of using a specific distribution model, use {\em concentration
-inequalities} to derive distribution-independent bounds on the
-VOI. Our bounds are applicable to any bounded distribution. Similar
-bounds can be derived for certain unbounded distributions, such as the
-normally distributed prior of the value with normally disributed
-sampling, a subject of future work.
+greatest value of information (VOI) estimate must be pulled.
 
@@
 \cite{Auer.ucb}. The UCT algorithm, an extension of UCB1 to
-Monte-Carlo Tree Search is described in \cite{Kocsis.uct}, and a successful
-application of UCT to playing the Go game is discussed in
-\cite{Gelly.mogo}.
-
-Pure exploration in Multi-armed bandits is explored in
+MCTS is described in \cite{Kocsis.uct}, and a successful
+application of UCT to playing Go is discussed in
+\cite{Gelly.mogo}. Pure exploration in Multi-armed bandits is explored in
 \cite{Bubeck.pure}. On the one hand, the paper proves certain upper
@@
 empirical performance of UCB1 appears to be better than of uniform
-sampling. \cite{Mnih.bernstop} investigate stopping criteria for
-sampling based on the empirical Bernstein inequality; however, the
-stopping criteria are based on error probabilities rather than on
-value of information measures, and do not directly address the
-objective of regret minimization.
-
-The principles of bounded rationality appear in
-\cite{Horvitz.reasoningabout}. \cite{Russell.right} provided a formal
+sampling. The principles of bounded rationality appear in
+\cite{Horvitz.reasoningabout}. \cite{Russell.right} provides a formal
 description of rational metareasoning and case studies of applications
-in several problem domains. One obstacle to straightforward
-application of the principles of rational metareasoning to Monte-Carlo
-sampling is the metagreedy assumption, according to which samples must
-be selected as though at most one sample can be taken before an action
-is chosen. In Monte-Carlo sampling, the value of information of any
-single sample in a given search state is often zero, so a different
-approximating assumption must be used instead.
+in several problem domains.
 
@@
 \begin{itemize}
-\item the arm $\alpha$ with the highest sample mean is pulled, and the 
-actual arm value is lower than the sample mean of the second-best arm $\beta$;
-\item another arm is pulled, and the actual arm value is higher
-than the sample mean of the best arm $\alpha$.
+\item the arm $\alpha$ with the highest sample mean $\overline
+  X_\alpha$ is pulled, and $\overline X_\alpha$ becomes lower than
+  $\overline X_\beta$ of the second-best arm $\beta$;
+\item another arm $i$ is pulled, and $\overline X_i$ becomes higher
+than $\overline X_\alpha$.
 \end{itemize}
@@
 Monte-Carlo sampling, since the effect of a single sample is small,
-and the myopic VOI estimate will often be zero, resulting in premature
-termination of the search. However, $\Lambda_i$ can be estimated as the intrinsic 
-value of information $\Lambda_i^b$ of pulling the $i$th arm for the
-rest of the budget of $N$ samples:
+and the myopic VOI estimate will often be zero. However, $\Lambda_i$
+can be estimated as the intrinsic value of information $\Lambda_i^b$
+of pulling the $i$th arm for the rest of the budget of $N$ samples:
 \begin{thm} Denote the current number of samples of the $i$th arm by
   $n_i$. $\Lambda_i^b$ is bounded from above as
-\begin{equation}
-  \Lambda_i^b \le \left\{
-  \begin{array}{l l}
-    \overline X_\beta\frac N {N+n_i}\Pr(\overline X_i'\le\overline X_\beta)
-    \le N\frac {\overline X_\beta} {n_i}\Pr(\overline X_i'\le\overline X_\beta) & \mbox{{\rm if} $i=\alpha$} \\
-    (1-\overline  X_\alpha)\frac N {N+n_i}\Pr(\overline X_i'\ge\overline X_\alpha)
-       \le N\frac {(1-\overline  X_\alpha)} {n_i}\Pr(\overline X_i'\ge\overline X_\alpha) & \mbox{\rm otherwise}
-  \end{array} \right.
+\begin{align}
+  \Lambda&_\alpha^b \le \frac {N \overline X_\beta} {N+n_i}\Pr(\overline X_i'\le\overline X_\beta)
+    \le \frac {N \overline X_\beta} {n_i} \Pr(\overline X_i'\le\overline X_\beta)\nonumber\\
+\Lambda&_{i|i\ne\alpha}^b \le \frac{ N(1-\overline  X_\alpha)} {N+n_i}\Pr(\overline X_i'\ge\overline X_\alpha)
+     \le \frac {N(1-\overline X_\alpha)} {n_i}\Pr(\overline   X_i'\ge\overline X_\alpha)
+\vspace{-24pt}
 \label{eqn:thm-be}
-\end{equation}
+\end{align}
 where $\overline X_i'$ is the sample mean of the $i$th arm after $n_i+N$ samples.
@@
 The probabilities can be bounded from above using the
-Hoeffding inequality \cite{Hoeffding.ineq}.
+Hoeffding inequality \cite{Hoeffding.ineq}:
 \begin{thm} The probabilities in equations (\ref{eqn:thm-be}) are bounded from above as
@@
    \right)\nonumber\\
-\Pr&(\overline X_i' \ge \overline X_\beta)
+\Pr&(\overline X_{i|i\ne\alpha}' \ge \overline X_\beta)
    \le 2\exp\left(- \varphi(n_i) (\overline X_\alpha -\overline  X_i)^2 n_i \right)
@@
 \end{align}
-where $\varphi(n)=2(\frac {1+n/N} {1+\sqrt {n/N}})^2 \ge 1.37$.
+where $\varphi(n)=2(\frac {1+n/N} {1+\sqrt {n/N}})^2 > 1.37$.
 \label{thm:hoeffding-prob-bounds}
 \end{thm}
-
+\vspace{-16pt}
 \begin{crl}
@@
 by substituting (\ref{eqn:probound-blnk-hoeffding}) into (\ref{eqn:thm-be}).
-\begin{equation}
-  \Lambda_i^b \le \hat\Lambda_i^b=\left\{
-  \begin{array}{l l}
-    N\frac {\overline X_\beta} {n_i}\cdot 2\exp\left(- 1.37(\overline X_i - \overline X_\beta)^2 n_i\right)
-      & \mbox{{\rm if} $i=\alpha$} \\
-    N\frac {(1-\overline  X_\alpha)} {n_i}\cdot 2\exp\left(- 1.37(\overline X_\alpha - \overline X_i)^2 n_i\right)
-      &  \mbox{\rm otherwise}
-  \end{array} \right.
+\begin{align}
+  \Lambda&_\alpha^b \le \hat\Lambda_\alpha^b=\frac {2N\overline X_\beta} {n_\alpha}\exp\left(- 1.37(\overline X_\alpha - \overline X_\beta)^2 n_\alpha\right)\nonumber\\
+  \Lambda&_{i|i\ne\alpha}^b\le \hat\Lambda_i^b=  \frac {2N(1-\overline  X_\alpha)} {n_i}\exp\left(- 1.37(\overline X_\alpha - \overline X_i)^2 n_i\right)
 \label{eqn:bound-blnk-hoeffding}
-\end{equation}
+\end{align}
 \label{crl:bound-blnk-hoeffding}
 \end{crl}
+\vspace{-16pt}
 Estimate $\hat\Lambda_i^b$ can be viewed as a
@@
 
-Following the principles of rational metareasoning, an arm with
+Following the principles of rational metareasoning, for pure
+exploration in Multi-armed Bandits an arm with
 the highest VOI should be pulled at each step. The upper bounds
 established in Corollary~\ref{crl:bound-blnk-hoeffding} can be used
-as VOI estimates. As illustrated by the empirical evaluation
+as VOI estimates. In MCTS, pure exploration takes place at the first
+step of a rollout, where an action with the highest utility must
+be chosen. MCTS differs from pure epxloration in Multi-armed Bandits
+in that the distributions of the rewards are not stationary. However,
+VOI estimates computed as for stationary distributions work well in
+practice. As illustrated by the empirical evaluation
 (Section~\ref{sec:empirical-evaluation}), estimates based on upper
-bounds on the VOI result in rational sampling policies, and exhibit
-performance comparable to the performance of some state-of-the-art
-heuristic algorithms.
+bounds on the VOI result in a rational sampling policy exceeding the
+performance of some state-of-the-art heuristic algorithms.
 
@@
 
-\begin{figure}[t]
+\begin{figure}[h!]
 \centering
 \includegraphics[scale=0.5]{flat.pdf}
+\vspace{-8pt}
 \caption{Random instances: regret vs. number of samples}
 \label{fig:random-instances}
+\vspace{-16pt}
 \end{figure}
@@
 randomly-generated Multi-armed bandits with 32 Bernoulli arms, with
-the mean rewards of the arms distributed uniformly in the range $[0,
-  1]$, for a range of sample budgets $32..1024$, with multiplicative
-step of $2$. The experiment for each number of samples was repeated
+the mean rewards of the arms distributed uniformly in the range~$[0,
+  1]$, for a range of sample budgets~$32..1024$, with multiplicative
+step of~$2$. The experiment for each number of samples was repeated
 10000 times. UCB1 is always considerably worse than the
@@
 the experiment.)
-\begin{figure}[t]
+\begin{figure}[h!]
 \centering
 \includegraphics[scale=0.5]{vct-wins.pdf}
+\vspace{-8pt}
 \caption{Go: winning rate --- UCT against VCT}
 \label{fig:vct-against-uct}
+\vspace{-16pt}
 \end{figure}
@@
 
-This work suggested Monte-Carlo sampling policies in which sample
-selection and termination are based on upper bounds on the value of
-information. A simplified model of accounting for future value of
-information of a sample, based on early termination and sample
-redistribution, was proposed for Monte-Carlo sampling in
-trees. Empirical evaluation showed that these policies outperform
-heuristic algorithms for simple regret minimization in Multi-armed
-bandits, as well as for tree search.
+This work suggested a Monte-Carlo sampling policy in which sample
+selection is based on upper bounds on the value of
+information. Empirical evaluation showed that this policy outperforms
+heuristic algorithms for pure exploration in Multi-armed
+bandits, as well as for MCTS.
 
-Monte-Carlo tree search still remains a largely unexplored field of
+MCTS still remains a largely unexplored field of
 application of VOI-aware algorithms. More elaborated VOI estimates,
 taking into consideration re-use of samples in future search states
-should be considered. The policies introduced in the paper differ from
+should be considered. The policy introduced in the paper differs from
 the UCT algorithm only at the first step, where the VOI-aware
@@
 metareasoning at all steps of a rollout may further improve the
-sampling policies.
+sampling.
 

### 2012-02-27 paper/vago.ecai/paper.tex
2 pages again

@@
 \title{VOI-aware MCTS}
+\author {Blind submission}
 %\author {David Tolpin, Solomon Eyal Shimony \\
@@
 \begin{abstract}
-An MCTS sampling policy based on VOI estimates of rollouts is
+UCT, a state-of-the art algorithm for Monte Carlo tree search (MCTS)
+in games and Markov decision processes, is based on UCB1, a sampling
+policy for the Multi-armed Bandit problem (MAB) that 
+minimizes the cumulative regret.  However, search differs from MAB in
+that in MCTS it is usually only the final ``arm pull'' (the actual
+move selection) that collects a reward, rather than all ``arm pulls''.
+In this paper, an MCTS sampling policy based on VOI estimates of rollouts is
 suggested. Empirical evaluation of the policy and comparison to
-UCB1 and UCT is performed on random Multi-armed Bandit instances
+UCB1 and UCT is performed on random MAB instances
 as well as on Computer Go.
@@
 
-\section{Introduction and Definitions}
+\section{Introduction}
 
-TODO: start with MCTS, then go to Bandits.
+MCTS, and especially UCT \cite{Kocsis.uct} appears in numerous search
+applications, such as \cite{Eyerich.ctp}. Although these methods are
+shown to be successful empirically, most authors appear to be using
+UCT ``because it has been shown to be successful in the past'', and
+``because it does a good job of trading off exploration and
+exploitation''. While the latter statement may be correct for the
+Multi-armed Bandit problem and for the UCB1 algorithm \cite{Auer.ucb},
+we argue that a simple reconsideration from basic principles can
+result in schemes that outperform UCT.
 
-Taking a sequence of samples in order to minimize the
-regret of a decision based on the samples is abstracted by the
-{\em Multi-armed Bandit Problem.} In the Multi-armed Bandit problem
-we have a set of $K$ arms. Each arm can be pulled multiple
-times. When the $i$th arm is pulled, a random reward $X_i$ from an
-unknown stationary distribution is returned.  The reward is bounded
-between 0 and 1.
+The core issue is that in MCTS for adversarial search and search in
+``games against nature'' the goal is typically to find the best
+first action of a good (or even optimal) policy, which is closer to
+minimizing the simple regret, rather than the cumulative regret
+minimized by UCB1.  However, the simple and the cumulative regret
+cannot be minimized simultaneously; moreover, \cite{Bubeck.pure} shows
+that in many cases the smaller the cumulative regret, the greater the
+simple regret.
 
-The simple regret of a sampling policy for the Multi-armed Bandit
-Problem is the expected difference between the best expected reward
-$\max\limits_i\IE[X_i]$ and the expected reward $\IE[X_j]$ of the arm
-with the best sample mean $\overline X_j=\max\limits_i\overline X_i$:
+We begin with background definitions and related work.  VOI estimates
+for arm pulls in MAB are presented, and a VOI-aware sampling policy is
+suggested, both for the simple regret in MAB and for MCTS.  Finally,
+the performance of the proposed sampling policy is evaluated on sets
+of Bernoulli arms and on Computer GO, showing the improved
+performance.
+
+\section{Background and Related Work}
+\label{sec:related-work}
+
+Monte-Carlo tree search was initially suggested as a scheme for
+finding approximately optimal policies for Markov Decision Processes
+(MDP).  MCTS explores an MDP by performing
+\emph{rollouts}---trajectories from the current state to a state in
+which a termination condition is satisfied (either the goal or a
+cutoff state).
+
+Taking a sequence of samples in order to minimize the regret of a
+decision based on the samples is captured by the Multi-armed Bandit
+problem (MAB) \cite{Vermorel.bandits}. In MAB, we have a set of $K$
+arms. Each arm can be pulled multiple times. When the $i$th arm is
+pulled, a random reward $X_i$ from an unknown stationary distribution
+is encountered. In the \textit{cumulative setting}, all encountered rewards are
+collected.  UCB1 \cite{Auer.ucb} was shown to be
+near-optimal in this respect. UCT, an extension of UCB1 to MCTS is
+described in \cite{Kocsis.uct}, and shown to outperform many state of
+the art search algorithms in both MDP and adversarial search
+\cite{Gelly.mogo,Eyerich.ctp}. In the \textit{simple regret setting}, the agent
+gets to collect only the reward of the last pull.
+\begin{dfn}
+The \textbf{simple regret} of a sampling policy for MAB
+is the expected difference between the best expected reward
+$\mu_*$ and the expected reward $\mu_j$ of the empirically best arm
+$\overline X_j=\max_i\overline X_i$:
 \begin{equation}
+\IE r=\sum_{j=1}^K\Delta_j\Pr(\overline X_j=\max_i\overline X_i)
 \label{eqn:simple-regret}
-\IE[R]=\sum_{j=1}^K\Delta_j\Pr(\overline X_j=\max_i\overline X_i)
+\vspace{-8pt}
 \end{equation}
-where $\Delta_j=\max\limits_i\IE[X_i]-\IE[X_j]$.  Strategies that
-minimize the simple regret are called pure exploration strategies
-\cite{Bubeck.pure}. Principles of rational metareasoning
-\cite{Russell.right} suggest that at each step the arm with the
-greatest value of information (VOI) estimate must be pulled.
+where $\Delta_j=\mu_*-\mu_j$.
+\end{dfn}
+Strategies that minimize the simple regret are called pure exploration
+strategies, explored in \cite{Bubeck.pure}.
 
-\section{Related Work}
-
-Efficient algorithms for Multi-Armed Bandits based on
-distribution-independent bounds, in particular UCB1, are introduced in
-\cite{Auer.ucb}. The UCT algorithm, an extension of UCB1 to
-MCTS is described in \cite{Kocsis.uct}, and a successful
-application of UCT to playing Go is discussed in
-\cite{Gelly.mogo}. Pure exploration in Multi-armed bandits is explored in
-\cite{Bubeck.pure}. On the one hand, the paper proves certain upper
-and lower bounds for UCB1 and uniform sampling, showing that an upper
-bound on the simple regret is exponential in the number of samples for
-uniform sampling, while only polynomial for UCB1. On the other hand,
-empirical performance of UCB1 appears to be better than of uniform
-sampling. The principles of bounded rationality appear in
-\cite{Horvitz.reasoningabout}. \cite{Russell.right} provides a formal
-description of rational metareasoning and case studies of applications
-in several problem domains.
+A different scheme for control of sampling can use the principles of
+bounded rationality \cite{Horvitz.reasoningabout} and rational
+metareasoning \cite{Russell.right}.  In search, one maintains a
+current best action $\alpha$, and finds the expected gain
+from finding another action $\beta $ to be better than the current best.
 
@@
   $n_i$. $\Lambda_i^b$ is bounded from above as
-\begin{align}
-  \Lambda&_\alpha^b \le \frac {N \overline X_\beta} {N+n_i}\Pr(\overline X_i'\le\overline X_\beta)
-    \le \frac {N \overline X_\beta} {n_i} \Pr(\overline X_i'\le\overline X_\beta)\nonumber\\
-\Lambda&_{i|i\ne\alpha}^b \le \frac{ N(1-\overline  X_\alpha)} {N+n_i}\Pr(\overline X_i'\ge\overline X_\alpha)
-     \le \frac {N(1-\overline X_\alpha)} {n_i}\Pr(\overline   X_i'\ge\overline X_\alpha)
-\vspace{-24pt}
+\begin{eqnarray}
+  &\Lambda_\alpha^b \le \frac {N \overline X_\beta} {N+n_i}\Pr(\overline X_i'\le\overline X_\beta)
+    \le \frac {N \overline X_\beta} {n_i} \Pr(\overline X_i'\le\overline X_\beta)\\
+&\Lambda_{i|i\ne\alpha}^b \le \frac{ N(1-\overline  X_\alpha)} {N+n_i}\Pr(\overline X_i'\ge\overline X_\alpha)
+     \le \frac {N(1-\overline X_\alpha)} {n_i}\Pr(\overline   X_i'\ge\overline X_\alpha)\nonumber
 \label{eqn:thm-be}
-\end{align}
+\end{eqnarray}
 where $\overline X_i'$ is the sample mean of the $i$th arm after $n_i+N$ samples.
@@
 \end{thm}
-\vspace{-16pt}
 \begin{crl}
@@
 \end{crl}
-\vspace{-16pt}
 Estimate $\hat\Lambda_i^b$ can be viewed as a
@@
 information. Empirical evaluation showed that this policy outperforms
-heuristic algorithms for pure exploration in Multi-armed
-bandits, as well as for MCTS.
+heuristic algorithms for pure exploration in MAB, as well as for MCTS.
 

### 2012-02-27 paper/vago.ecai/paper.tex
spell-checked

@@
 relaxation of the myopic assumption such that sequences consisting
-of more than one action are taken into considertion.
+of more than one action are taken into consideration.
 
@@
 step of a rollout, where an action with the highest utility must
-be chosen. MCTS differs from pure epxloration in Multi-armed Bandits
+be chosen. MCTS differs from pure exploration in Multi-armed Bandits
 in that the distributions of the rewards are not stationary. However,

### 2012-02-27 paper/vago.ecai/paper.tex
graphs

@@
 \vspace{-8pt}
-\caption{Go: winning rate --- UCT against VCT}
+\caption{Go: winning rate --- VOI against UCT}
 \label{fig:vct-against-uct}
@@
 1000 times. Figure~\ref{fig:vct-against-uct}
-shows the winning rate of VOI+UCT against UCT vs. the number of
-samples. For most numbers of samples per node, VOI+UCT outperforms UCT.
+shows the winning rate of VOI against UCT vs. the number of
+samples. For most numbers of samples per node, VOI outperforms UCT.
 

### 2012-02-27 paper/vago.ecai/data/figures.R

@@ -1,0 +1,38 @@
+draw.regrets <- function(regrets_file, log="xy", legend.position='topright', xlab=expression(N[samples]), main=NA) {
+	regrets <- read.table(regrets_file, header=F)
+
+        names <- as.matrix(regrets[1,])
+        regrets <- as.matrix(regrets[-1,])
+        regrets <- matrix(as.numeric(regrets), nrow=nrow(regrets), ncol=ncol(regrets))
+
+        xlim <- range(regrets[,1])
+	values <- unlist(regrets[,-1])
+	ylim <- range(values[values>0])
+
+	plot(x=regrets[,1], type='n', log=log, xlim=xlim, ylim=ylim, yaxs='i', xlab=xlab, ylab='Regrets', main=if(is.na(main)) unlist(strsplit(regrets_file, "\\."))[[1]] else main)
+	for(i in (2:ncol(regrets))) {
+          lines(x=regrets[,1], y=regrets[,i],
+                type="o", pch=i-1)
+	}
+	legend(x=legend.position, legend=sapply(names[-1], function(n) substring(n, 3)), pch=1:length(names[-1]))
+}
+
+
+draw.winrate <- function(regrets_file, log="xy", legend.position='topright', xlab=expression(N[samples]), main=NA) {
+	regrets <- read.table(regrets_file, header=F)
+
+        names <- as.matrix(regrets[1,])
+        regrets <- as.matrix(regrets[-1,])
+        regrets <- matrix(as.numeric(regrets), nrow=nrow(regrets), ncol=ncol(regrets))
+
+    xlim <- range(regrets[,1])
+    ylim <- range(max(0, min(regrets[,-1])*0.95), min(100, max(regrets[,-1])*1.05))
+
+	plot(x=regrets[,1], y=regrets[,2], type='o', log=log, xlim=xlim, ylim=ylim, yaxs='i', xlab=xlab, ylab='VOI wins, %', main=if(is.na(main)) unlist(strsplit(regrets_file, "\\."))[[1]] else main, axes=F)
+	box()
+	axis(1, at=regrets[,1])
+	axis(2)
+	abline(h=50.0, lty='dashed');
+}
+
+

### 2012-03-11 code/erf.py
added python implementation of erf function

@@ -1,0 +1,22 @@
+import math
+
+def erf(x):
+    # constants
+    a1 =  0.254829592
+    a2 = -0.284496736
+    a3 =  1.421413741
+    a4 = -1.453152027
+    a5 =  1.061405429
+    p  =  0.3275911
+
+    # Save the sign of x
+    sign = 1
+    if x < 0:
+        sign = -1
+    x = abs(x)
+
+    # A & S 7.1.26
+    t = 1.0/(1.0 + p*x)
+    y = 1.0 - (((((a5*t + a4)*t) + a3)*t + a2)*t + a1)*t*math.exp(-x*x)
+
+    return sign*y

### 2012-03-14 paper/vago/paper.tex

@@
 Solve (\ref{eqn:app-hoeffding-eq-exp}) for $\delta$: $\delta=\frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}} (\overline X_\alpha
-- \overline X_i) \ge (\sqrt 2 - 1)(\overline X_\alpha-\overline X_i)$. Substitute $\delta$ into 
+- \overline X_i) \ge 2(\sqrt 2 - 1)(\overline X_\alpha-\overline X_i)$. Substitute $\delta$ into 
 (\ref{eqn:app-hoeffding-le-maxexp}) and obtain
@@
                           (\overline X_\alpha - \overline X_i)\right)^2 n_i\right)\nonumber \\
-& \le & 2\exp(-2(\sqrt 2 - 1)^2(\overline X_\alpha - \overline X_i)^2n_i)
+& \le & 2\exp(-8(\sqrt 2 - 1)^2(\overline X_\alpha - \overline X_i)^2n_i)
  \le 2\exp(-1.37(\overline X_\alpha - \overline X_i)^2n_i)

### 2012-03-14 paper/srcr/paper.tex
answers to the questions
fixed illustration

@@
 \begin{dfn}
-The \textbf{simple regret} of a sampling policy for the Multi-armed Bandit
+The \textbf{simple regret} $\IE r$ of a sampling policy for the Multi-armed Bandit
 Problem is the expected difference between the best expected reward
@@
     \State \Call{UpdateStats}{node, action, reward}
+    \State \textbf{return} reward
   \EndIf

### 2012-03-20 paper/vago.uai/paper.tex
vago.uai in progress

--- 2012-02-27 paper/vago.ecai/paper.tex
@@
-\documentclass{ecai2012}
-\usepackage{times}
+\documentclass[]{article}
+\usepackage{proceed2e}
 \usepackage{enumerate}
@@
 \usepackage{color}
-
-\ecaisubmission
 
@@
 
-\title{VOI-aware MCTS}
-\author {Blind submission}
-%\author {David Tolpin, Solomon Eyal Shimony \\
-%Department of Computer Science, \\
-%Ben-Gurion University of the Negev, Beer Sheva, Israel \\
-%\{tolpin,shimony\}@cs.bgu.ac.il}
-
 \begin{document}
 
-\maketitle
-
-\begin{abstract}
-UCT, a state-of-the art algorithm for Monte Carlo tree search (MCTS)
-in games and Markov decision processes, is based on UCB1, a sampling
-policy for the Multi-armed Bandit problem (MAB) that 
-minimizes the cumulative regret.  However, search differs from MAB in
-that in MCTS it is usually only the final ``arm pull'' (the actual
-move selection) that collects a reward, rather than all ``arm pulls''.
-In this paper, an MCTS sampling policy based on VOI estimates of rollouts is
-suggested. Empirical evaluation of the policy and comparison to
-UCB1 and UCT is performed on random MAB instances
-as well as on Computer Go.
-\end{abstract}
-
-\section{Introduction}
-
-MCTS, and especially UCT \cite{Kocsis.uct} appears in numerous search
-applications, such as \cite{Eyerich.ctp}. Although these methods are
-shown to be successful empirically, most authors appear to be using
-UCT ``because it has been shown to be successful in the past'', and
-``because it does a good job of trading off exploration and
-exploitation''. While the latter statement may be correct for the
-Multi-armed Bandit problem and for the UCB1 algorithm \cite{Auer.ucb},
-we argue that a simple reconsideration from basic principles can
-result in schemes that outperform UCT.
-
-The core issue is that in MCTS for adversarial search and search in
-``games against nature'' the goal is typically to find the best
-first action of a good (or even optimal) policy, which is closer to
-minimizing the simple regret, rather than the cumulative regret
-minimized by UCB1.  However, the simple and the cumulative regret
-cannot be minimized simultaneously; moreover, \cite{Bubeck.pure} shows
-that in many cases the smaller the cumulative regret, the greater the
-simple regret.
-
-We begin with background definitions and related work.  VOI estimates
-for arm pulls in MAB are presented, and a VOI-aware sampling policy is
-suggested, both for the simple regret in MAB and for MCTS.  Finally,
-the performance of the proposed sampling policy is evaluated on sets
-of Bernoulli arms and on Computer GO, showing the improved
-performance.
-
-\section{Background and Related Work}
-\label{sec:related-work}
-
-Monte-Carlo tree search was initially suggested as a scheme for
-finding approximately optimal policies for Markov Decision Processes
-(MDP).  MCTS explores an MDP by performing
-\emph{rollouts}---trajectories from the current state to a state in
-which a termination condition is satisfied (either the goal or a
-cutoff state).
-
-Taking a sequence of samples in order to minimize the regret of a
-decision based on the samples is captured by the Multi-armed Bandit
-problem (MAB) \cite{Vermorel.bandits}. In MAB, we have a set of $K$
-arms. Each arm can be pulled multiple times. When the $i$th arm is
-pulled, a random reward $X_i$ from an unknown stationary distribution
-is encountered. In the \textit{cumulative setting}, all encountered rewards are
-collected.  UCB1 \cite{Auer.ucb} was shown to be
-near-optimal in this respect. UCT, an extension of UCB1 to MCTS is
-described in \cite{Kocsis.uct}, and shown to outperform many state of
-the art search algorithms in both MDP and adversarial search
-\cite{Gelly.mogo,Eyerich.ctp}. In the \textit{simple regret setting}, the agent
-gets to collect only the reward of the last pull.
-\begin{dfn}
-The \textbf{simple regret} of a sampling policy for MAB
-is the expected difference between the best expected reward
-$\mu_*$ and the expected reward $\mu_j$ of the empirically best arm
-$\overline X_j=\max_i\overline X_i$:
-\begin{equation}
-\IE r=\sum_{j=1}^K\Delta_j\Pr(\overline X_j=\max_i\overline X_i)
-\label{eqn:simple-regret}
-\vspace{-8pt}
-\end{equation}
-where $\Delta_j=\mu_*-\mu_j$.
-\end{dfn}
-Strategies that minimize the simple regret are called pure exploration
-strategies, explored in \cite{Bubeck.pure}.
-
-A different scheme for control of sampling can use the principles of
-bounded rationality \cite{Horvitz.reasoningabout} and rational
-metareasoning \cite{Russell.right}.  In search, one maintains a
-current best action $\alpha$, and finds the expected gain
-from finding another action $\beta $ to be better than the current best.
-
-\section{Upper Bounds on Value of Information}
+\section*{Upper Bounds on Value of Information}
 
 The intrinsic VOI $\Lambda_i$ of pulling an arm is the expected decrease
-in the regret compared to selecting an arm without pulling any arm at
+in the regret compared to selecting the best arm without pulling any arm at
 all. Two cases are possible:
@@
 Monte-Carlo sampling, since the effect of a single sample is small,
-and the myopic VOI estimate will often be zero. However, $\Lambda_i$
-can be estimated as the intrinsic value of information $\Lambda_i^b$
-of pulling the $i$th arm for the rest of the budget of $N$ samples:
-\begin{thm} Denote the current number of samples of the $i$th arm by
-  $n_i$. $\Lambda_i^b$ is bounded from above as
+and the myopic VOI estimate will often be zero. However, for the
+common case of a fixed budget of samples per node, $\Lambda_i$ can be
+estimated as the intrinsic VOI $\Lambda_i^b$ of pulling the $i$th arm
+for the rest of the budget.  Let us denote the current number of
+samples of the $i$th arm by $n_i$, and the remaining number of samples
+by $N$:
+\begin{thm} $\Lambda_i^b$ is bounded from above as
 \begin{eqnarray}
+\label{eqn:thm-be}
   &\Lambda_\alpha^b \le \frac {N \overline X_\beta} {N+n_i}\Pr(\overline X_i'\le\overline X_\beta)
@@
      \le \frac {N(1-\overline X_\alpha)} {n_i}\Pr(\overline   X_i'\ge\overline X_\alpha)\nonumber
-\label{eqn:thm-be}
 \end{eqnarray}
@@
 \begin{align}
+  \label{eqn:probound-blnk-hoeffding}
   \Pr&(\overline X_\alpha' \le \overline X_\beta)
@@
   \le 2\exp\left(- \varphi(n_i) (\overline X_\alpha -\overline  X_i)^2 n_i \right)
-\label{eqn:probound-blnk-hoeffding}
 \end{align}
@@
 \begin{align}
+  \label{eqn:bound-blnk-hoeffding}
   \Lambda&_\alpha^b \le \hat\Lambda_\alpha^b=\frac {2N\overline X_\beta} {n_\alpha}\exp\left(- 1.37(\overline X_\alpha - \overline X_\beta)^2 n_\alpha\right)\nonumber\\
   \Lambda&_{i|i\ne\alpha}^b\le \hat\Lambda_i^b=  \frac {2N(1-\overline  X_\alpha)} {n_i}\exp\left(- 1.37(\overline X_\alpha - \overline X_i)^2 n_i\right)
-\label{eqn:bound-blnk-hoeffding}
 \end{align}
@@
 \end{crl}
-Estimate $\hat\Lambda_i^b$ can be viewed as a
-relaxation of the myopic assumption such that sequences consisting
-of more than one action are taken into consideration.
 
-\section{VOI-based Sample Allocation}
+\section*{VOI-based Sample Allocation}
 
@@
 
-\section{Empirical Evaluation}
+\section*{Empirical Evaluation}
 \label{sec:empirical-evaluation}
 
-\subsection{Selecting The Best Arm}
+\subsection*{Selecting The Best Arm}
 \label{sec:emp-arm}
@@
 \centering
-\includegraphics[scale=0.5]{flat.pdf}
+\includegraphics[scale=0.6]{flat.pdf}
 \vspace{-8pt}
@@
 
-\subsection{Playing Go Against UCT}
+\subsection*{Playing Go Against UCT}
 \label{sec:emp-go}
@@
 \centering
-\includegraphics[scale=0.5]{vct-wins.pdf}
+\includegraphics[scale=0.6]{vct-wins.pdf}
 \vspace{-8pt}
@@
 \end{figure}
-The engines were compared on the 9x9 board, for 5000, 7000, 1000, and
+The engines were compared on the 9x9 board, for 5000, 7000, 10000, and
 15000 samples per ply, each experiment was repeated
@@
 
-\section{Summary and Future Work}
-
-This work suggested a Monte-Carlo sampling policy in which sample
-selection is based on upper bounds on the value of
-information. Empirical evaluation showed that this policy outperforms
-heuristic algorithms for pure exploration in MAB, as well as for MCTS.
-
-MCTS still remains a largely unexplored field of
-application of VOI-aware algorithms. More elaborated VOI estimates,
-taking into consideration re-use of samples in future search states
-should be considered. The policy introduced in the paper differs from
-the UCT algorithm only at the first step, where the VOI-aware
-decisions are made. Consistent application of principles of rational
-metareasoning at all steps of a rollout may further improve the
-sampling.
-
-\bibliographystyle{ecai2012}
+\bibliographystyle{plain}
 \bibliography{refs}

### 2012-03-20 paper/vago.uai/sample2e.tex
vago.uai in progress

% Sample LaTeX file for creating a paper in the Morgan Kaufmannn two 
% column, 8 1/2 by 11 inch proceedings format.

\documentclass[]{article}
\usepackage{proceed2e}

\title{Instructions for Authors}

\author{} % LEAVE BLANK FOR ORIGINAL SUBMISSION.
          % UAI 2010 reviewing is double-blind.

% The author names and affiliations should appear only in the accepted paper.
%
%\author{ {\bf Harry Q.~Bovik\thanks{Footnote for author to give an 
%alternate address.}} \\  
%Computer Science Dept. \\  
%Cranberry University\\ 
%Pittsburgh, PA 15213 \\ 
%\And 
%{\bf Coauthor}  \\ 
%Affiliation          \\ 
%Address \\              
%\And 
%{\bf Coauthor}   \\ 
%Affiliation \\          
%Address    \\           
%(if needed)\\ 
%}

\begin{document}

\maketitle

\begin{abstract} 
The Abstract paragraph should be indented 0.25 inch (1.5 picas) on
both left and right-hand margins.  Use 10~point type, with a vertical
spacing of 11~points.  {\bf Abstract} must be centered, bold, and in
point size 12. Two line spaces precede the Abstract. The Abstract must
be limited to one paragraph.
\end{abstract}

\section{GENERAL FORMATTING INSTRUCTIONS}

Papers are in 2 columns with the overall line width of 6.75~inches
(41~picas).  Each column is 3.25~inches wide (19.5~picas).  The space
between the columns is .25~inches wide (1.5~picas).  The left margin
is 1~inch (6~picas).  Use 10~point type with a vertical spacing of
11~points.  Times Roman is the preferred typeface throughout.

Paper title is 16~point, caps/lc, bold, centered between 2~horizontal
rules.  Top rule is 4~points thick and bottom rule is 1~point thick.
Allow 1/4~inch space above and below title to rules.

Reviewing is double-blind, so do not include author names, affiliations, or any
other identifying information in the original submission.  If you include urls
to supplementary material, make sure the urls also do not disclose your identity.

After a paper is accepted, for the camera-ready submission, Authors' names are
centered, initial caps.  The lead author's name is to be listed first
(left-most), and the Co-authors' names (if different address) are set to
follow.  If only one co-author, center both the author and co-author,
side-by-side.

One-half line space between paragraphs, with no indent.

\section{FIRST LEVEL HEADINGS}

First level headings are all caps, flush left, bold and in point size
12. One line space before the first level heading and 1/2~line space
after the first level heading.

\subsection{SECOND LEVEL HEADING}

Second level headings must be flush left, all caps, bold and in point
size 10. One line space before the second level heading and 1/2~line
space after the second level heading.

\subsubsection{Third Level Heading}

Third level headings must be flush left, initial caps, bold, and in
point size 10.  One line space before the third level heading and
1/2~line space after the third level heading.

\vskip .5pc 
Fourth Level Heading

Fourth level headings must be flush left and initial caps.
One line space before the fourth level heading and 1/2~line space
after the fourth level heading.

\subsection{CITATIONS, FIGURES, REFERENCES}

\subsubsection{Citations in Text}

Citations within the text should include the author's last name and
year, e.g., (Cheesman, 1985). Reference style should follow the style
that you are used to using, as long as the citation style is
consistent.

For the original submission, take care not to reveal the authors' identity through
the manner in which one's own previous work is cited.  For example, writing
``In (Bovik, 1970), we studied the problem of AI'' would be inappropriate, as
it reveals the author's identity.  Instead, write ``(Bovik, 1970) studied the
problem of AI.''

\subsubsection{Footnotes}

Indicate footnotes with a number\footnote{Sample of the first
footnote} in the text. Use 8 point type for footnotes.  Place the
footnotes at the bottom of the page on which they appear.  Precede the
footnote with a 0.5 point horizontal rule 1~inch (6~picas)
long.\footnote{Sample of the second footnote}

\subsubsection{Figures}

All artwork must be centered, neat, clean, and legible. Figure number
and caption always appear below the figure.  Leave 2 line spaces
between the figure and the caption. The figure caption is initial caps
and each figure numbered consecutively.

Make sure that the figure caption does not get separated from the
figure. Leave extra white space at the bottom of the page rather than
splitting the figure and figure caption.
\begin{figure}[h] 
\vspace{1in} 
\caption{Sample Figure Caption} 
\end{figure}

\subsubsection{Tables}

All tables must be centered, neat, clean, and legible. Table number
and title always appear above the table.  See
Table~\ref{sample-table}.

One line space before the table title, one line space after the table
title, and one line space after the table. The table title must be
initial caps and each table numbered consecutively.

\begin{table}[h] 
\caption{Sample Table Title} 
\label{sample-table} 
\begin{center} 
\begin{tabular}{ll} 
\multicolumn{1}{c}{\bf PART}  &\multicolumn{1}{c}{\bf DESCRIPTION} \\ 
\hline \\ 
Dendrite         &Input terminal \\ 
Axon             &Output terminal \\ 
Soma             &Cell body (contains cell nucleus) \\ 
\end{tabular} 
\end{center} 
\end{table}

\newpage

\subsubsection*{Acknowledgements}

Use unnumbered third level headings for the acknowledgements title.
All acknowledgements go at the end of the paper.

\subsubsection*{References}

References follow the acknowledgements.  Use unnumbered third level
heading for the references title.  Any choice of citation style is
acceptable as long as you are consistent.

J.~Alspector, B.~Gupta, and R.~B.~Allen  (1989). Performance of a 
stochastic learning microchip.  In D. S. Touretzky (ed.), {\it Advances 
in Neural Information Processing Systems 1}, 748-760.  San Mateo, Calif.: 
Morgan Kaufmann.

F.~Rosenblatt (1962). {\it Principles of Neurodynamics.} Washington, 
D.C.: Spartan Books.

G.~Tesauro (1989). Neurogammon wins computer Olympiad.  {\it Neural 
Computation} {\bf 1}(3):321-323.

\end{document}

### 2012-03-20 paper/vago.uai/uai12-flat.tex
vago.uai in progress

\documentclass{article}
\begin{document}

\section{Introduction}

%% Build on the intro from the Hay and Russell tech report

[[Sequential decision problem approximation by simulating possible futures]]

[[Metareasoning to select simulations]]

[[Monte-Carlo tree search]]

[[UCT, which is based on the asymptotically regret-optimal bandit control rule UCB]]

[[Bandit problems vs. selection problems]]

[[Resulting problems for UCT: biased against selecting negative options, no natural stopping criteria]]

[[explain nature of our results: basic theoretical foundations,
    initial foray into new hueristics for selection and some empirical
    results]]

\section{On optimal policies for selection}

[[Formal definition of selection problems and the metalevel MDP with cost per sample (time value); also mention budgeted learning.]]

[[Theorem: if Optimal stops in x, myopic stops in x (converse is more useful!)]]

[[Theorem: if Myopic stops in all states reachable from x, optimal stops in x]]

[[Theorem: expected number of samples is bounded; actual number bounded wp 1]]

[[Theorems: actual number of samples bounded for flat Normal, Bernoulli]]

[[Counterexample to actual boundedness in general (SPRT case)]]

\section{Context effects and non-indexability}

[[No index theorem; via context inversion counter-example]]

[[Theorem: context effect occurs only for a single convex interval of
    context value (how general can we make this?]]

\section{Regret bounds and approximate policies}

[[Regret models: simple regret, regret with cost per sampling; regret goes to zero as c does]]

[[Expected simple regret bounds for normal case?]]

[[Blinkered sampling]]

[[ESPb: Frazier's continuous time approximation]]

%% [[Selection criteria]]
%% [[Termination criteria]]

[[Concentration upper bounds on VOI]]

[[Control rules: VCT, ECT, BCT]]

\section{Application to Monte Carlo tree search}

[[summarize MCTS]]

[[Root rules vs. subtree rules; here just focus on former.]]

[[Efficiency gains from reusing samples between time steps; Shimony's model of this]]

[[Simple regret in bandit problem [VEB]CT vs. UCB]]

[[Improvement in performance in Go (or other trees) [VEB]CT vs. UCT]]

\end{document}

### 2012-03-20 paper/vago.ecai/paper.tex
entered submission number
changes
vago for uai in progress

@@
 move selection) that collects a reward, rather than all ``arm pulls''.
-In this paper, an MCTS sampling policy based on VOI estimates of rollouts is
-suggested. Empirical evaluation of the policy and comparison to
-UCB1 and UCT is performed on random MAB instances
+In this paper, an MCTS sampling policy based on Value of Information
+(VOI) estimates of rollouts is suggested. Empirical evaluation of the
+policy and comparison to UCB1 and UCT is performed on random MAB instances
 as well as on Computer Go.
@@
 Strategies that minimize the simple regret are called pure exploration
-strategies, explored in \cite{Bubeck.pure}.
+strategies \cite{Bubeck.pure}.
 
@@
 bounded rationality \cite{Horvitz.reasoningabout} and rational
-metareasoning \cite{Russell.right}.  In search, one maintains a
-current best action $\alpha$, and finds the expected gain
-from finding another action $\beta $ to be better than the current best.
+metareasoning \cite{Russell.right,HayRussell.MCTS}.  In search, one
+maintains a current best action $\alpha$, and finds the expected gain
+from finding another action $\beta $ to be better than the current
+best.
 
@@
 The intrinsic VOI $\Lambda_i$ of pulling an arm is the expected decrease
-in the regret compared to selecting an arm without pulling any arm at
+in the regret compared to selecting the best arm without pulling any arm at
 all. Two cases are possible:
@@
 Monte-Carlo sampling, since the effect of a single sample is small,
-and the myopic VOI estimate will often be zero. However, $\Lambda_i$
-can be estimated as the intrinsic value of information $\Lambda_i^b$
-of pulling the $i$th arm for the rest of the budget of $N$ samples:
-\begin{thm} Denote the current number of samples of the $i$th arm by
-  $n_i$. $\Lambda_i^b$ is bounded from above as
+and the myopic VOI estimate will often be zero. However, for the
+common case of a fixed budget of samples per node, $\Lambda_i$ can be
+estimated as the intrinsic VOI $\Lambda_i^b$ of pulling the $i$th arm
+for the rest of the budget.  Let us denote the current number of
+samples of the $i$th arm by $n_i$, and the remaining number of samples
+by $N$:
+\begin{thm} $\Lambda_i^b$ is bounded from above as
 \begin{eqnarray}
-  &\Lambda_\alpha^b \le \frac {N \overline X_\beta} {N+n_i}\Pr(\overline X_i'\le\overline X_\beta)
-    \le \frac {N \overline X_\beta} {n_i} \Pr(\overline X_i'\le\overline X_\beta)\\
+\label{eqn:thm-be}
+  &\Lambda_\alpha^b \le \frac {N \overline X_\beta} {N+n_\alpha}\Pr(\overline X_\alpha'\le\overline X_\beta)
+    \le \frac {N \overline X_\beta} {n_\alpha} \Pr(\overline X_\alpha'\le\overline X_\beta)\\
 &\Lambda_{i|i\ne\alpha}^b \le \frac{ N(1-\overline  X_\alpha)} {N+n_i}\Pr(\overline X_i'\ge\overline X_\alpha)
      \le \frac {N(1-\overline X_\alpha)} {n_i}\Pr(\overline   X_i'\ge\overline X_\alpha)\nonumber
-\label{eqn:thm-be}
 \end{eqnarray}
@@
 \end{crl}
-Estimate $\hat\Lambda_i^b$ can be viewed as a
-relaxation of the myopic assumption such that sequences consisting
-of more than one action are taken into consideration.
 
@@
 MCTS still remains a largely unexplored field of
-application of VOI-aware algorithms. More elaborated VOI estimates,
+application of VOI-aware algorithms. More elaborate VOI estimates,
 taking into consideration re-use of samples in future search states

### 2012-03-20 paper/vago.uai/paper.tex
vago for uai in progress

@@
 
-\section*{Upper Bounds on Value of Information}
+\section{Upper Bounds on Value of Information}
 
@@
 \begin{thm} $\Lambda_i^b$ is bounded from above as
-\begin{eqnarray}
+\begin{align}
 \label{eqn:thm-be}
-  &\Lambda_\alpha^b \le \frac {N \overline X_\beta} {N+n_i}\Pr(\overline X_i'\le\overline X_\beta)
-    \le \frac {N \overline X_\beta} {n_i} \Pr(\overline X_i'\le\overline X_\beta)\\
-&\Lambda_{i|i\ne\alpha}^b \le \frac{ N(1-\overline  X_\alpha)} {N+n_i}\Pr(\overline X_i'\ge\overline X_\alpha)
-     \le \frac {N(1-\overline X_\alpha)} {n_i}\Pr(\overline   X_i'\ge\overline X_\alpha)\nonumber
-\end{eqnarray}
-where $\overline X_i'$ is the sample mean of the $i$th arm after $n_i+N$ samples.
+  \Lambda_\alpha^b& \le \frac {N \overline X_\beta^n} {N+n_\alpha}\Pr(\overline X_\alpha^{n+N}\le\overline X_\beta^n)\nonumber\\
+    &\le \frac {N \overline X_\beta^n} {n_\alpha} \Pr(\overline X_\alpha^{n+N}\le\overline X_\beta^n)\nonumber\\
+\Lambda_{i|i\ne\alpha}^b& \le \frac{ N(1-\overline  X_\alpha^{n_i})} {N+n_i}\Pr(\overline X_i^{{n_i}+N}\ge\overline X_\alpha^{n_i})\nonumber\\
+     &\le \frac {N(1-\overline X_\alpha^{n_i})} {n_i}\Pr(\overline   X_i^{{n_i}+N}\ge\overline X_\alpha^{n_i})
+\end{align}
 \label{thm:be}
@@
   \label{eqn:probound-blnk-hoeffding}
-  \Pr&(\overline X_\alpha' \le \overline X_\beta)
-  \le 2\exp\left(- \varphi(n_\alpha)(\overline X_\alpha - \overline X_\beta)^2 n_\alpha
+  \Pr&(\overline X_\alpha^{{n_\alpha}+N} \le \overline X_\beta^{n_\alpha})
+  \le 2\exp\left(- \varphi(n_\alpha)(\overline X_\alpha^{n_\alpha} - \overline X_\beta^{n_\alpha})^2 n_\alpha
   \right)\nonumber\\
-  \Pr&(\overline X_{i|i\ne\alpha}' \ge \overline X_\beta)
-  \le 2\exp\left(- \varphi(n_i) (\overline X_\alpha -\overline  X_i)^2 n_i \right)
+  \Pr&(\overline X_{i|i\ne\alpha}^{n+N} \ge \overline X_\beta^n)
+  \le 2\exp\left(- \varphi(n_i) (\overline X_\alpha^n -\overline  X_i^n)^2 n_i \right)
 \end{align}
@@
   \label{eqn:bound-blnk-hoeffding}
-  \Lambda&_\alpha^b \le \hat\Lambda_\alpha^b=\frac {2N\overline X_\beta} {n_\alpha}\exp\left(- 1.37(\overline X_\alpha - \overline X_\beta)^2 n_\alpha\right)\nonumber\\
-  \Lambda&_{i|i\ne\alpha}^b\le \hat\Lambda_i^b=  \frac {2N(1-\overline  X_\alpha)} {n_i}\exp\left(- 1.37(\overline X_\alpha - \overline X_i)^2 n_i\right)
+  \Lambda&_\alpha^b \le \hat\Lambda_\alpha^b=\frac {2N\overline X_\beta^{n_\alpha}} {n_\alpha}\exp\left(- 1.37(\overline X_\alpha^{n_\alpha} - \overline X_\beta^{n_\alpha})^2 n_\alpha\right)\nonumber\\
+  \Lambda&_{i|i\ne\alpha}^b\le \hat\Lambda_i^b=  \frac {2N(1-\overline  X_\alpha^{n_i})} {n_i}\exp\left(- 1.37(\overline X_\alpha^{n_i} - \overline X_i^{n_i})^2 n_i\right)
 \end{align}
@@
 \end{crl}
+Better bounds can be obtained through tighter estimates on the
+probabilities, for example, based on the empirical Bernstein
+inequality~\cite{MaurerPontil.benrstein}, or through a more careful
+application of the Hoeffding inequality.
 
-\section*{VOI-based Sample Allocation}
+\section{VOI-based Sampling Control}
 
-Following the principles of rational metareasoning, for pure
-exploration in Multi-armed Bandits an arm with
-the highest VOI should be pulled at each step. The upper bounds
-established in Corollary~\ref{crl:bound-blnk-hoeffding} can be used
-as VOI estimates. In MCTS, pure exploration takes place at the first
-step of a rollout, where an action with the highest utility must
-be chosen. MCTS differs from pure exploration in Multi-armed Bandits
-in that the distributions of the rewards are not stationary. However,
-VOI estimates computed as for stationary distributions work well in
-practice. As illustrated by the empirical evaluation
-(Section~\ref{sec:empirical-evaluation}), estimates based on upper
-bounds on the VOI result in a rational sampling policy exceeding the
-performance of some state-of-the-art heuristic algorithms.
+\subsection{Selection Criterion}
 
-\section*{Empirical Evaluation}
+Following the principles of rational metareasoning, an arm with the
+highest VOI should be pulled at each step. The upper bounds
+established in Theorem~\ref{thm:be} can be used as VOI estimates. MCTS
+differs from pure exploration in Multi-armed Bandits in that the
+distributions of the rewards are not stationary. However, VOI
+estimates computed as for stationary distributions work well in
+practice. As illustrated by the empirical evaluation. As illustrated
+by the empirical evaluation (Section~\ref{sec:empirical-evaluation}),
+estimates based on upper bounds on the VOI result in rational sampling
+policies, and exhibit performance comparable to the performance of
+some state-of-the-art heuristic algorithms.
+
+\subsection{Termination Condition}
+\label{sec:control-termination-condition}
+
+The simplest termination condition for a sampling policy is the
+budget---a fixed number of samples performed before a decision is
+made. When a sample has a cost commensurable with the value of
+information of a measurement, an upper bound on the intrinsic VOI can
+be used to stop the sampling if the intrinsic VOI of any action
+is less than the cost of sampling $C$:
+\begin{equation}
+\mbox{stop if } \max_i \Lambda_i \le C
+\end{equation}
+Blinkered VOI estimates~(\ref{eqn:thm-be},
+\ref{eqn:bound-blnk-hoeffding}) include the remaining budget $N$ as a
+factor, but given the cost of a single sample $c$, the cost of the
+remaining samples accounted for in estimating the intrinsic VOI is
+$C=cN$. $N$ can be dropped on both sides of the inequality,
+and a viable stopping condition is
+\begin{align}
+\frac 1 N \Lambda_\alpha^b \le&\frac {\overline X_\beta^{n_\alpha}}
+  {n_\alpha}\Pr(\overline X_\alpha^{n_\alpha+N}\le\overline
+  X_\beta^{n_\alpha})\le c\nonumber\\
+\frac 1 N \max_i\Lambda_i^b\le &\max_i\frac {(1-\overline X_\alpha^{n_i})} {n_i}\Pr(\overline
+  X_i^{n_i+N}\ge\overline X_\alpha^{n_i})\le c\nonumber\\
+    &\forall i: i\ne\alpha
+\label{eqn:stopping-blnk}
+\end{align}
+The empirical evaluation (Section~\ref{sec:empirical-evaluation})
+confirms viability of this stopping condition and illustrates the
+influence of the sample cost $c$ on the performance of
+the sampling policy.
+
+\subsection{Sample Redistribution in Trees}
+\label{sec:control-redistribution}
+
+Monte-Carlo tree search \cite{Chaslot.montecarlo} selects at each
+search step an action that appears to be the best based on outcomes
+of \textit{search rollouts}. Two different criteria are employed in
+action selection during a rollout:
+\begin{itemize}
+\item at the first step, the value of partial
+information should be maximized to minimize the \textit{simple} regret
+of the selection at the current root node;
+\item as the goal of sampling in deeper tree nodes is estimating the
+value of a node, rather than selection, it makes sense to minimize
+the \textit{cumulative} regret of the rollouts from the second step
+onwards.
+\end{itemize}
+
+A natural approach would be to apply an algorithm that minimizes the
+simple regret in Multi-armed bandits at the first step, and the
+cumulative regret during the rest of the rollout.  However, this
+approach assumes that the information obtained from rollouts in the
+current state is discarded after an action is selected. In practice,
+many successful Monte-Carlo tree search algorithms re-use rollouts
+originating from an earlier search state and passing through the
+current search state; thus, the value of information of a rollout is
+determined not just by the influence on the choice of the action at
+the current state, but also by the influence on the choice at future
+search states, provided the rollout passes through the search states
+which will be visited.
+
+One way to account for this re-use would be to incorporate the
+`future' value of information into a VOI estimate. However, this 
+approach appears to be complicated. Alternatively, one can behave
+myopically in search tree depth:
+\begin{enumerate}
+\item estimate VOI as though the information is discarded after each step;
+\item stop early if the VOI is below a certain threshold
+   (see Section~\ref{sec:control-termination-condition});
+\item save the unused sample budget for search in future state, such that
+   if the nominal budget is $N$, and the unused budget in the last state
+   is $N_u$, the search budget in the next state will be $N+N_u$.
+\end{enumerate}
+In this approach, the cost of a sample in the current state is the
+VOI of increasing the budget of a future state by one sample.  It is
+unclear whether this cost can be accurately estimated, but supposedly
+a fixed value for a given problem type and algorithm implementation
+would work. Indeed, the empirical evaluation (Section~\ref{sec:emp-go})
+confirms that stopping and sample redistribution based on a learned
+fixed cost  substantially improve the performance of the VOI-based
+sampling policy in game tree search.
+
+\section{Empirical Evaluation}
 \label{sec:empirical-evaluation}
 
-\subsection*{Selecting The Best Arm}
+\subsection{Selecting The Best Arm}
 \label{sec:emp-arm}
@@
 \includegraphics[scale=0.6]{flat.pdf}
-\vspace{-8pt}
 \caption{Random instances: regret vs. number of samples}
 \label{fig:random-instances}
-\vspace{-16pt}
 \end{figure}
@@
 
-\subsection*{Playing Go Against UCT}
+\subsection{Playing Go Against UCT}
 \label{sec:emp-go}
@@
 \includegraphics[scale=0.6]{vct-wins.pdf}
-\vspace{-8pt}
 \caption{Go: winning rate --- VOI against UCT}
 \label{fig:vct-against-uct}
-\vspace{-16pt}
 \end{figure}

### 2012-03-20 paper/vago.uai/paper.tex
vago.uai our part first draft

@@
 \cite{Gelly.mogo}. A modified version of Pachi \cite{Braudis.pachi}, a state of the art
-Go program, was used for the experiments. The UCT engine was extended
-with a VOI-aware sampling policy, and a time allocation mode ensuring
-that both the original UCT policy and the VOI-aware policy use the
-same average number of samples per node was added. (While the UCT
-engine is not the most powerful engine of Pachi, it is still a strong
-player; on the other hand, additional features of more advanced
+Go program, was used for the experiments:
+\begin{itemize}
+\item The UCT engine of Pachi was extended with VOI-aware sampling
+  policies at the first step. 
+\item The stopping condition for the VOI-aware policy was
+  modified and based solely on the sample cost, specified as
+  a constant parameter. The heuristic stopping condition for the
+  original UCT policy was left unchanged.
+\item The time-allocation mode based on the fixed number of samples
+  was modified for \textit{both the original UCT policy and the VOI-aware
+  policies} such that 
+  \begin{itemize}
+    \item the same number of samples is available to
+      the agent at each step, independently of the number of pre-simulated
+      games;  
+    \item if samples were unused at the current step,
+      they become available at the next step. 
+  \end{itemize}
+\end{itemize}
+While the UCT engine is not the most powerful engine of Pachi, it is still a strong
+player. On the other hand, additional features of more advanced
 engines would obstruct the MCTS phenomena which are the subject of
-the experiment.)
+the experiment.
 \begin{figure}[h!]
 \centering
-\includegraphics[scale=0.6]{vct-wins.pdf}
-\caption{Go: winning rate --- VOI against UCT}
-\label{fig:vct-against-uct}
+\includegraphics[scale=0.6]{uctvoi.pdf}
+\caption{Go: winning rate vs. $c$}
+\label{fig:uctvoi}
 \end{figure}
-The engines were compared on the 9x9 board, for 5000, 7000, 10000, and
-15000 samples per ply, each experiment was repeated
-1000 times. Figure~\ref{fig:vct-against-uct}
+The engines were compared on the 9x9 board, for 5000, 7000, 1000, and
+15000 samples (game simulations) per ply, each experiment was repeated
+1000 times. Figure~\ref{fig:uctvoi}
+show the winning rate of UCT against the VOI-aware policy
+vs. the stopping threshold $c$ (if the maximum VOI of a sample is below
+the threshold, the simulation is stopped, and a move is chosen). Each
+curve in the figure corresponds to a certain number of samples per
+ply. For the stopping threshold of $10^{-6}$, the VOI-aware policy
+is almost always better than UCT, and  reaches the winning rate of
+64\% for 10000 samples per ply.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[scale=0.6]{voi-wins.pdf}
+\caption{Go: winning rate for $c=10^{-6}$}
+\label{fig:voi-wins}
+\end{figure}
+
+Figure~\ref{fig:voi-wins}
 shows the winning rate of VOI against UCT vs. the number of
-samples. For most numbers of samples per node, VOI outperforms UCT.
+samples for $c=10^{-6}$. In agreement with the intuition
+(Section~\ref{sec:control-redistribution}), VOI-based stopping and
+sample redistribution is most influential for medium numbers of
+samples per ply. When the maximum number of samples is too low, early
+stopping would result in poorly selected moves. On the other hand,
+when the maximum number of samples is sufficiently high, the VOI of
+increasing the maximum number of samples in a future state is low.
 

### 2012-03-20 code/figures.R

@@ -86,23 +86,33 @@ go.draw.curves <- function(agent="vct") {
     fname <- paste("uct", agent, "-time=", time, ".shomersu.dat", sep="")
     res[[time]] <- read.table(fname, header=T)
     res[[time]]$MAXVOI <- res[[time]]$MAXVOI+1E-8
-    res[[time]]$WIN_W <- res[[time]]$WIN_W*100 
+    res[[time]]$WIN_W <- 100-res[[time]]$WIN_W*100 
     if(lo > min(res[[time]]$WIN_W)) lo <- min(res[[time]]$WIN_W)
     if(hi < max(res[[time]]$WIN_W)) hi <- max(res[[time]]$WIN_W)
   }
-  lo <- min(lo, 30)
+  lo <- max(lo, 30)
   hi <- min(hi, 70)
   
 
+  par(xpd=F, mgp=c(2.5, 1, 0))
   plot(res[['5000']]$MAXVOI, res[['5000']]$WIN_W, ylim=c(lo,hi), log="x", type='n',
-       main=NA, xlab=expression(VOI[min]+10^-8), ylab='UCT wins, %')
-  abline(h=50, lty='dashed')
-  lines
+       main=NA, xlab='c', ylab='VOI wins, %', xaxs='i', xaxp=c(1E-7, 1, 1))
+  abline(h=50, lty='longdash')
+  abline(v=1.01E-6, lty='dashed')
+  par(xpd=NA)
+  segments(x0=1E-8, y0=lo, y1=lo-3)
+  text(x=1E-8, y=lo-6, "0")
+  segments(x0=1.01E-6, y0=lo, y1=lo-3)
+  text(x=1.01E-6, y=lo-6, "1e-06")
+  segments(x0=1.E-4, y0=lo, y1=lo-3)
+  text(x=1.E-4, y=lo-6, "1e-04")
+
+  par(xpd=F)
   for(time in go.times) {
     lines(res[[time]]$MAXVOI, res[[time]]$WIN_W, type='o', pch=go.pchs[[time]])
   }
 
-  legend(x='topleft', legend=go.times, pch=sapply(go.times, function(time) go.pchs[[time]]))
+  legend(x='topright', legend=go.times, pch=sapply(go.times, function(time) go.pchs[[time]]))
 }
 
 go.draw.bests <- function(bestsfile, colors=rainbow(3), densities=c(20, 20, 30)) {

### 2012-03-20 paper/vago.uai/notation.tex
in progress

\documentclass[]{article}

\begin{document}

\begin{tabular}{ll} 
    $K$ & Number of actions/arms \\ 
    $n$ & Index of the number of samples; $N$ as an upper bound/total 
number of samples \\ 
    $U_i$ & Utility of $i$th action (random variable) \\ 
    $X_{i,j}$ & $j$th sample of $i$th action (random variable) \\     
    $\overline{X}^n_{i}$ & Sample mean of $i$th action after $n$ samples 
(random variable) \\ 
    $\Lambda^p_i$ & Intrinsic value of perfect information for $i$th action \\ 
    $\Lambda^{VCT}_i$ & VCT upper bound on $\Lambda^p_i$; similarly for 
ECT, BCT \\ 
    $\mu_{i,n}$    & Posterior mean of $i$th action after $n$ samples (random 
variable) \\ 
    $\pi$ & Metalevel policy / sampling policy (pick one and use 
consistently?) \\ 
    $R^\pi_n$ & Simple regret of sampling policy $\pi$ given $n$ samples \\ 
    $R^\pi_c$ & Simple regret of sampling policy $\pi$ with cost of 
sampling $c$ \\     
    $\Delta_j$ & Regret of choosing action $i$; $\Delta_j = (\max_i U_i) - 
U_j$ \\ 
    $M$ & Metalevel MDP \\ 
    $M_{\mu_0}$ & Metalevel MDP $M$ with action of known value $\mu_0$ 
adjoined \\ 
    $S$ & Set of belief states in a given metalevel MDP \\     
    $\mu_i(s)$ & Posterior mean of $i$th action in belief state $s$ \\         
    $V^\pi_{M,c}(s)$ & Value of metalevel policy $\pi$ in metalevel MDP $M$ 
with cost of sampling $c$ \\ 
    $Q^\pi_{M,c}(s)$ & Q-value of .... \\ 
\end{tabular}

\end{

### 2012-03-21 paper/vago.uai/paper.tex
some corrections

@@
 \end{itemize}
+
+In order to estimate the VOI of pulling an arm, a sampling model is
+needed.  The generic sampling model used is: 
+\begin{itemize}
+\item The actual value of an arm has a certain distribution.
+\item Pulling an arm is an i.i.d. noisy measurement of the actual
+  value of the arm.
+\end{itemize}
+Instead of using a specific distribution model, we use {\em concentration
+inequalities} to derive distribution-independent bounds on the
+VOI. Our bounds are applicable to any bounded distribution. Similar
+bounds can be derived for certain unbounded distributions, such as the
+normally distributed prior of the value with normally disributed
+sampling, a subject of future work.
+
 The \textit{myopic} VOI estimate is of limited applicability to
@@
 \label{eqn:thm-be}
-  \Lambda_\alpha^b& \le \frac {N \overline X_\beta^n} {N+n_\alpha}\Pr(\overline X_\alpha^{n+N}\le\overline X_\beta^n)\nonumber\\
-    &\le \frac {N \overline X_\beta^n} {n_\alpha} \Pr(\overline X_\alpha^{n+N}\le\overline X_\beta^n)\nonumber\\
-\Lambda_{i|i\ne\alpha}^b& \le \frac{ N(1-\overline  X_\alpha^{n_i})} {N+n_i}\Pr(\overline X_i^{{n_i}+N}\ge\overline X_\alpha^{n_i})\nonumber\\
-     &\le \frac {N(1-\overline X_\alpha^{n_i})} {n_i}\Pr(\overline   X_i^{{n_i}+N}\ge\overline X_\alpha^{n_i})
+  \Lambda_\alpha^b&\le \frac {N \overline X_\beta^{n_\beta}} {n_\alpha} \Pr(\overline X_\alpha^{n_\alpha+N}\le\overline X_\beta^{n_\beta})\nonumber\\
+\Lambda_{i|i\ne\alpha}^b&\le \frac {N(1-\overline X_\alpha^{n_\alpha})} {n_i}\Pr(\overline   X_i^{{n_i}+N}\ge\overline X_\alpha^{n_\alpha})
 \end{align}
@@
 \end{thm}
+\vspace{-2em}
+\begin{proof} For the case $i\ne \alpha$, the probability that the
+  $i$th arm is finally chosen instead of $\alpha$ is
+  $\Pr(\overline X_i^{n_i+N} \ge \overline X_\alpha^{n_\alpha})$. $X_i \le 1$,
+  therefore $\overline X_i^{n_i+N}\le \overline
+  X_\alpha^{n_\alpha}+\frac {N(1-\overline X_\alpha^{n_\alpha})} {N+n_i}$. Hence, the intrinsic value of perfect
+  information is at most $\frac{ N(1-\overline  X_\alpha^{n_\alpha})}
+  {N+n_i}\Pr(\overline X_i^{{n_i}+N}\ge\overline X_\alpha^{n_\alpha})\le \frac {N \overline X_\beta^{n_\beta}} {n_\alpha} \Pr(\overline X_\alpha^{n_\alpha+N}\le\overline X_\beta^{n_\beta})$.
+  Proof for the case $i=\alpha$ is similar.
+\end{proof}
 
@@
   \label{eqn:probound-blnk-hoeffding}
-  \Pr&(\overline X_\alpha^{{n_\alpha}+N} \le \overline X_\beta^{n_\alpha})
-  \le 2\exp\left(- \varphi(n_\alpha)(\overline X_\alpha^{n_\alpha} - \overline X_\beta^{n_\alpha})^2 n_\alpha
+  \Pr&(\overline X_\alpha^{{n_\alpha}+N} \le \overline X_\beta^{n_\beta})
+  \le 2\exp\left(- \varphi (\overline X_\alpha^{n_\alpha} - \overline X_\beta^{n_\beta})^2 n_\alpha
   \right)\nonumber\\
-  \Pr&(\overline X_{i|i\ne\alpha}^{n+N} \ge \overline X_\beta^n)
-  \le 2\exp\left(- \varphi(n_i) (\overline X_\alpha^n -\overline  X_i^n)^2 n_i \right)
+  \Pr&(\overline X_{i|i\ne\alpha}^{n_\alpha+N} \ge \overline X_\beta^{n_\beta})
+  \le 2\exp\left(- \varphi (\overline X_\alpha^{n_\alpha} -\overline  X_i^{n_i})^2 n_i \right)
 \end{align}
-where $\varphi(n)=2(\frac {1+n/N} {1+\sqrt {n/N}})^2 > 1.37$.
+where $\varphi=\min \left(2(\frac {1+n/N} {1+\sqrt {n/N}})^2\right)=8(\sqrt 2 - 1)^2 > 1.37$.
 \label{thm:hoeffding-prob-bounds}
 \end{thm}
+\begin{proof}
+Equations (\ref{eqn:probound-blnk-hoeffding}) follow from the
+observation that if $i\ne\alpha$, $\overline X_i^{n_i+N}>\overline X_\alpha^{n_i}$
+if and only if the mean $\overline X_i^N$ of $N$ samples from $n_i+1$
+to $n_i+N$ is at least $\overline X_\alpha^{n_i}+(\overline X_\alpha^{n_i}-\overline
+X_i^{n_i})\frac {n_i} N$.
+
+For any $\delta$, the probability that $\overline X_i^{n_i+N}$ is greater
+than $\overline X_\alpha^{n_i}$ is less than the probability that
+$\IE[X_i]\ge\overline X_i^n+\delta$
+\emph{or} $\overline X_i^N\ge \IE[X_i]+\overline X_\alpha^{n_\alpha}
+- \overline X_i^{n_i} - \delta +(\overline X_\alpha^{n_\alpha} - \overline X_i^{n_i})\frac {n_i} N$,
+thus, by the union bound, less than the sum of the probabilities:
+\begin{align}
+\Pr&(\overline X_i^{n_i+N}\ge \overline X_\alpha^{n_i})\nonumber\\
+   &\le\Pr(\IE[X_i]-\overline X_i^{n_i} \ge \delta)\\
+   &\hspace{0.25em}+\Pr\left(\overline X_i^N\hspace{-0.5em} - \IE[X_i] \ge \overline X_\alpha^{n_\alpha}\hspace{-0.5em}
+           - \overline X_i^{n_i}\hspace{-0.5em} - \delta +(\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_i^{n_i})\frac {n_i} N\right)\nonumber
+\end{align}
+Bounding the probabilities on the right-hand side using the Hoeffding
+inequality, obtain:
+\begin{align}
+\Pr&(\overline X_i^{n_i+N}\ge \overline X_\alpha^{n_\alpha})\le \nonumber\\
+    &\exp(-2\delta^2n_i)+\nonumber\\
+    &\quad\exp\left(-2\left((\overline X_\alpha^{n_\alpha}\hspace{-0.5em}
+         - \overline X_i^{n_i})\left(1+\frac {n_i} N\right)
+         - \delta\right)^2N\right)
+\label{eqn:app-hoeffding-le-maxexp}
+\end{align}
+Find $\delta$ for which the two terms on the right-hand side of
+(\ref{eqn:app-hoeffding-le-maxexp}) are equal:
+\begin{equation}
+\exp(-\delta^2n) = \exp\left(-2\left((\overline X_\alpha - \overline X_i)(1+\frac {n_i} N) - \delta\right)^2N\right)\label{eqn:app-hoeffding-eq-exp}
+\end{equation}
+Solve (\ref{eqn:app-hoeffding-eq-exp}) for $\delta$: $\delta=\frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}} (\overline X_\alpha^{n_\alpha}\hspace{-0.5em}
+- \overline X_i^{n_i}) \ge 2(\sqrt 2 - 1)(\overline X_\alpha^{n_\alpha}\hspace{-0.5em}-\overline X_i^{n_i})$. Substitute $\delta$ into 
+(\ref{eqn:app-hoeffding-le-maxexp}) and obtain
+\begin{align}
+\Pr&(\overline X_i^{n_i}\ge \overline X_\alpha^{n_\alpha}) \nonumber\\
+& \le 2\exp\left(-2\left( \frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}}
+                          (\overline X_\alpha^{n_\alpha} - \overline X_i^{n_i})\right)^2 n_i\right)\nonumber \\
+& \le 2\exp(-8(\sqrt 2 - 1)^2(\overline X_\alpha^{n_\alpha} - \overline X_i^{n_i})^2n_i)\nonumber\\
+& = 2\exp(-\varphi(\overline X_\alpha^{n_\alpha} - \overline X_i^{n_i})^2n_i)
+\end{align}
+Derivation for the case $i=\alpha$ is similar.
+\end{proof}
+
 \begin{crl}
@@
   \label{eqn:bound-blnk-hoeffding}
-  \Lambda&_\alpha^b \le \hat\Lambda_\alpha^b=\frac {2N\overline X_\beta^{n_\alpha}} {n_\alpha}\exp\left(- 1.37(\overline X_\alpha^{n_\alpha} - \overline X_\beta^{n_\alpha})^2 n_\alpha\right)\nonumber\\
-  \Lambda&_{i|i\ne\alpha}^b\le \hat\Lambda_i^b=  \frac {2N(1-\overline  X_\alpha^{n_i})} {n_i}\exp\left(- 1.37(\overline X_\alpha^{n_i} - \overline X_i^{n_i})^2 n_i\right)
+  \Lambda&_\alpha^b \le \hat\Lambda_\alpha^b=\frac {2N\overline X_\beta^{n_\beta}} {n_\alpha}\exp\left(- \varphi(\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_\beta^{n_\beta})^2 n_\alpha\right)\nonumber\\
+  \Lambda&_{i|i\ne\alpha}^b\le \hat\Lambda_i^b=  \frac {2N(1-\overline  X_\alpha^{n_\alpha})} {n_i}\exp\left(- \varphi(\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_i^{n_i})^2 n_i\right)
 \end{align}
@@
 \end{crl}
+\vspace{-2em}
 Better bounds can be obtained through tighter estimates on the
@@
 estimates computed as for stationary distributions work well in
-practice. As illustrated by the empirical evaluation. As illustrated
+practice. As illustrated
 by the empirical evaluation (Section~\ref{sec:empirical-evaluation}),
@@
 
-\subsection{Termination Condition}
-\label{sec:control-termination-condition}
+\subsection{Stopping Criterion}
+\label{sec:control-stopping-criterion}
 
-The simplest termination condition for a sampling policy is the
+The simplest stopping criterion for a sampling policy is the
 budget---a fixed number of samples performed before a decision is
@@
 \end{equation}
-Blinkered VOI estimates~(\ref{eqn:thm-be},
+VOI estimates~(\ref{eqn:thm-be},
 \ref{eqn:bound-blnk-hoeffding}) include the remaining budget $N$ as a
@@
 \begin{align}
-\frac 1 N \Lambda_\alpha^b \le&\frac {\overline X_\beta^{n_\alpha}}
+\frac 1 N \Lambda_\alpha^b \le&\frac {\overline X_\beta^{n_\beta}}
   {n_\alpha}\Pr(\overline X_\alpha^{n_\alpha+N}\le\overline
   X_\beta^{n_\alpha})\le c\nonumber\\
-\frac 1 N \max_i\Lambda_i^b\le &\max_i\frac {(1-\overline X_\alpha^{n_i})} {n_i}\Pr(\overline
-  X_i^{n_i+N}\ge\overline X_\alpha^{n_i})\le c\nonumber\\
+\frac 1 N \max_i\Lambda_i^b\le &\max_i\frac {(1-\overline X_\alpha^{n_\alpha})} {n_i}\Pr(\overline
+  X_i^{n_i+N}\ge\overline X_\alpha^{n_\alpha})\le c\nonumber\\
     &\forall i: i\ne\alpha
@@
 \item stop early if the VOI is below a certain threshold
-   (see Section~\ref{sec:control-termination-condition});
+   (see Section~\ref{sec:control-stopping-criterion});
 \item save the unused sample budget for search in future state, such that
@@
   policies at the first step. 
-\item The stopping condition for the VOI-aware policy was
+\item The stopping criterion for the VOI-aware policy was
   modified and based solely on the sample cost, specified as
-  a constant parameter. The heuristic stopping condition for the
+  a constant parameter. The heuristic stopping criterion for the
   original UCT policy was left unchanged.
@@
 1000 times. Figure~\ref{fig:uctvoi}
-show the winning rate of UCT against the VOI-aware policy
+show the winning rate of the VOI-aware policy against UCT
 vs. the stopping threshold $c$ (if the maximum VOI of a sample is below
@@
 (Section~\ref{sec:control-redistribution}), VOI-based stopping and
-sample redistribution is most influential for medium numbers of
+sample redistribution is most influential for intermediate numbers of
 samples per ply. When the maximum number of samples is too low, early

### 2012-03-21 paper/vago.uai/paper.tex
corrected version

@@
 inequality~\cite{MaurerPontil.benrstein}, or through a more careful
-application of the Hoeffding inequality.
+application of the Hoeffding inequality:
+\begin{align}
+\Lambda_i^b&\le\frac {N\sqrt \pi} {n_i \sqrt {n_i}}
+  \left[\mathrm{erf}\left((1\hspace{-0.25em}-\hspace{-0.25em}\overline X_i^{n_i})\sqrt {n_i}\right)
+      \hspace{-0.25em}-\hspace{-0.25em}\mathrm{erf}\left((\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_i^{n_i})\sqrt{n_i}\right)\right]\nonumber\\
+\Lambda_\alpha^b&\le\frac {N\sqrt \pi} {n_\alpha \sqrt {n_\alpha}}
+  \left[\mathrm{erf}\left(\overline X_\alpha^{n_\alpha}\sqrt {n_\alpha}\right)
+      -\mathrm{erf}\left((\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_\beta^{n_\beta})\sqrt{n_\alpha}\right)\right]
+\label{eqn:erf-blinkered}
+\end{align}
 
@@
 problem instances. Figure~\ref{fig:random-instances} shows results for
-randomly-generated Multi-armed bandits with 32 Bernoulli arms, with
+randomly-generated Multi-armed bandits with 25 Bernoulli arms, with
 the mean rewards of the arms distributed uniformly in the range~$[0,
-  1]$, for a range of sample budgets~$32..1024$, with multiplicative
+  1]$, for a range of sample budgets~$200..2000$, with multiplicative
 step of~$2$. The experiment for each number of samples was repeated
-10000 times. UCB1 is always considerably worse than the
-VOI-aware sampling policy.
+10000 times. UCB1, as well as policies  based on bounds
+(\ref{eqn:bound-blnk-hoeffding}) (VOI) and
+(\ref{eqn:erf-blinkered}) (VOI+) are compared. UCB1 is always considerably worse than the
+VOI-aware sampling policies.
 

### 2012-03-21 paper/vago.uai/paper.tex
draft sent

@@
 
-The intrinsic VOI $\Lambda_i$ of pulling an arm is the expected decrease
-in the regret compared to selecting the best arm without pulling any arm at
-all. Two cases are possible:
+The intrinsic (myopic greedy) VOI $\Lambda_i$ of testing an arm $i$ is
+the expected reward of the best arm given the information obtained by the test,
+minus the expected reward of the best arm according to the current information state.
+We will assume here a distribution model where the arm with the highest sample mean
+in an information state is the one with the highest expected reward.
+Two cases are possible:
 \begin{itemize}
 \item the arm $\alpha$ with the highest sample mean $\overline
-  X_\alpha$ is pulled, and $\overline X_\alpha$ becomes lower than
+  X_\alpha$ is tested, and $\overline X_\alpha$ becomes lower than
   $\overline X_\beta$ of the second-best arm $\beta$;
-\item another arm $i$ is pulled, and $\overline X_i$ becomes higher
+\item another arm $i$ is tested, and $\overline X_i$ becomes higher
 than $\overline X_\alpha$.
@@
 
-In order to estimate the VOI of pulling an arm, a sampling model is
+In order to estimate the VOI of testing an arm, a sampling model is
 needed.  The generic sampling model used is: 
 \begin{itemize}
-\item The actual value of an arm has a certain distribution.
-\item Pulling an arm is an i.i.d. noisy measurement of the actual
+\item The actual value of an arm has a certain (possibly unknown) distribution.
+\item Testing an arm is an i.i.d. noisy measurement of the actual
   value of the arm.
 \end{itemize}
-Instead of using a specific distribution model, we use {\em concentration
+When the distribution model is known, one can use the methods of Section \ref{sec:normal}.
+Otherwise, one can still use {\em concentration
 inequalities} to derive distribution-independent bounds on the
@@
 bounds can be derived for certain unbounded distributions, such as the
-normally distributed prior of the value with normally disributed
-sampling, a subject of future work.
+normally distributed prior of the value with normally distributed
+sampling.
 
-The \textit{myopic} VOI estimate is of limited applicability to
+Unfortunately, the pure \textit{myopic} VOI estimate is of little use in
 Monte-Carlo sampling, since the effect of a single sample is small,
-and the myopic VOI estimate will often be zero. However, for the
-common case of a fixed budget of samples per node, $\Lambda_i$ can be
-estimated as the intrinsic VOI $\Lambda_i^b$ of pulling the $i$th arm
-for the rest of the budget.  Let us denote the current number of
-samples of the $i$th arm by $n_i$, and the remaining number of samples
-by $N$:
+and the myopic VOI estimate will often be zero. Instead, we will derive
+an estimate for the case where we decide to test an arm $N$ times,
+where $N$ can be the remaining ``budget'' of available samples or
+any other integer quantity.
+Denote by  $\Lambda_i^b$ the intrinsic VOI of testing the $i$th arm
+$N$ times, and the number of
+samples already taken from the $i$th arm by $n_i$.
 \begin{thm} $\Lambda_i^b$ is bounded from above as
@@
   X_\alpha^{n_\alpha}+\frac {N(1-\overline X_\alpha^{n_\alpha})} {N+n_i}$. Hence, the intrinsic value of perfect
-  information is at most $\frac{ N(1-\overline  X_\alpha^{n_\alpha})}
-  {N+n_i}\Pr(\overline X_i^{{n_i}+N}\ge\overline X_\alpha^{n_\alpha})\le \frac {N \overline X_\beta^{n_\beta}} {n_\alpha} \Pr(\overline X_\alpha^{n_\alpha+N}\le\overline X_\beta^{n_\beta})$.
+  information is at most: 
+\begin{equation}
+\label{eq:simplistic}
+\frac{ N(1-\overline  X_\alpha^{n_\alpha})}
+  {N+n_i}\Pr(\overline X_i^{{n_i}+N}\hspace{-0.5em}\ge\overline X_\alpha^{n_\alpha})\le \frac {N \overline X_\beta^{n_\beta}} {n_\alpha} \Pr(\overline X_\alpha^{n_\alpha+N}\hspace{-0.5em}\le\overline X_\beta^{n_\beta})
+\end{equation}
   Proof for the case $i=\alpha$ is similar.
@@
 \vspace{-2em}
-Better bounds can be obtained through tighter estimates on the
-probabilities, for example, based on the empirical Bernstein
+
+More refined bounds can be obtained through tighter estimates on the
+probabilities in Equation \ref{eq:simplistic}, for example, based on the empirical Bernstein
 inequality~\cite{MaurerPontil.benrstein}, or through a more careful
-application of the Hoeffding inequality:
+application of the Hoeffding inequality, resulting in:
 \begin{align}
@@
 
-Following the principles of rational metareasoning, an arm with the
-highest VOI should be pulled at each step. The upper bounds
-established in Theorem~\ref{thm:be} can be used as VOI estimates. MCTS
+A reasonable greedy policy is to test the arm that has the best
+current VOI estimate. When the distributions are unknown, it makes sense
+to use the upper bounds established in Theorem~\ref{thm:be} instead. Although MCTS
 differs from pure exploration in Multi-armed Bandits in that the
-distributions of the rewards are not stationary. However, VOI
+distributions of the rewards are not stationary, VOI
 estimates computed as for stationary distributions work well in
@@
 by the empirical evaluation (Section~\ref{sec:empirical-evaluation}),
-estimates based on upper bounds on the VOI result in rational sampling
-policies, and exhibit performance comparable to the performance of
+estimates based on upper bounds on the VOI result in good sampling
+policies, which exhibit performance comparable to the performance of
 some state-of-the-art heuristic algorithms.
@@
 budget---a fixed number of samples performed before a decision is
-made. When a sample has a cost commensurable with the value of
-information of a measurement, an upper bound on the intrinsic VOI can
+made. When a sample has a known cost commensurable with the value of
+information of a measurement, an upper bound on the intrinsic VOI can also
 be used to stop the sampling if the intrinsic VOI of any action
@@
 VOI estimates~(\ref{eqn:thm-be},
-\ref{eqn:bound-blnk-hoeffding}) include the remaining budget $N$ as a
+\ref{eqn:bound-blnk-hoeffding}) include the remaining sample budget $N$ as a
 factor, but given the cost of a single sample $c$, the cost of the
@@
 $C=cN$. $N$ can be dropped on both sides of the inequality,
-and a viable stopping condition is
+and a reasonable stopping criterion would be:
 \begin{align}
@@
 The empirical evaluation (Section~\ref{sec:empirical-evaluation})
-confirms viability of this stopping condition and illustrates the
+confirms the viability of this stopping criterion and illustrates the
 influence of the sample cost $c$ on the performance of
@@
 \item at the first step, the value of partial
-information should be maximized to minimize the \textit{simple} regret
-of the selection at the current root node;
-\item as the goal of sampling in deeper tree nodes is estimating the
-value of a node, rather than selection, it makes sense to minimize
-the \textit{cumulative} regret of the rollouts from the second step
-onwards.
+information should be used to select nodes to sample, as the goal is
+selection of a move at the current root node;
+\item as the goal of sampling in deeper tree nodes is usually to improve the
+node value estimate, rather than selection, a different sampling scheme
+should be used there. At these levels, the standard practice
+of using UCT-like rollouts makes sense, as this method is known to be nearly optimal
+in minimizing the \textit{cumulative} regret.
 \end{itemize}
 
-A natural approach would be to apply an algorithm that minimizes the
-simple regret in Multi-armed bandits at the first step, and the
-cumulative regret during the rest of the rollout.  However, this
-approach assumes that the information obtained from rollouts in the
+The above hybrid approach assumes
+that the information obtained from rollouts in the
 current state is discarded after an action is selected. In practice,
 many successful Monte-Carlo tree search algorithms re-use rollouts
-originating from an earlier search state and passing through the
-current search state; thus, the value of information of a rollout is
+generated at earlier search states, if the sample traverses the
+current search state during the rollout; thus, the value of information of a rollout is
 determined not just by the influence on the choice of the action at
-the current state, but also by the influence on the choice at future
-search states, provided the rollout passes through the search states
-which will be visited.
+the current state, but also by its potential influence on the choice at future
+search states.
 
@@
 `future' value of information into a VOI estimate. However, this 
-approach appears to be complicated. Alternatively, one can behave
-myopically in search tree depth:
+approach requires a non-trivial extension of the theory of meta-reasoning for search.
+Alternately, one can behave myopically w.r.t. search tree depth:
 \begin{enumerate}
@@
    (see Section~\ref{sec:control-stopping-criterion});
-\item save the unused sample budget for search in future state, such that
+\item save the unused sample budget for search in future states, such that
    if the nominal budget is $N$, and the unused budget in the last state

### 2012-03-28 paper/vago.uai/context.tex
fixed typo

%% \section{Context effects and non-indexability}

\note{No index theorem; via context inversion counter-example}

de

\begin{figure}[htb]
\centering
\includegraphics[scale=0.7, trim=90 70 400 300]{swap-counterex.pdf}
\caption{Regret vs. cost}
\label{fig:swap-counterex}
\end{figure}

\begin{comment}
set samples 1000

max(x,y) = x>y ? x : y
max3(x,y,z) = max(max(x,y),z)

q0(u) = max(1,u)
qa(u) = 0.5*max3(u, 1.5, -0.2 + 0.5*max(u,1.5) + 0.5*max(u, 1.75)) + 0.5*max3(u, 1, -0.2 + 0.5*max(u,0.25) + 0.5*max(u, 1.75)) - 0.2
qb(u) = 0.5*max3(u, 0.25, -0.2 + 0.5*max(u,1.5) + 0.5*max(u, 0.25)) + 0.5*max3(u, 1.75, -0.2 + 0.5*max(u,1.75) + 0.5*max(u, 1.75)) - 0.2

set terminal postscript enhanced linewidth 2
set size 0.5,0.5
set output "swap-counterex.ps"
set xlabel "Utility of fixed alternative (u)"
set ylabel "Utility relative to stopping"
set xtics 0.5
set ytics 0.1
set key left top
plot [u=-2:2] [-0.3:0.3] 0 title "stop" with lines lw 2, qa(u)-q0(u) title "observe A" with lines lw 2, qb(u)-q0(u) title "observe B" with lines lw 2	
\end{comment}

%% \note{Theorem: /stopping/ context effect occurs only for a single convex interval of
%%    context value (how general can we make this?)}

\begin{dfn}\label{dfn:context}
	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$, and a constant $\mu\in\R$,
	define $M_\mu = (S,s_0,A_s,T,R_\mu)$ to be $M$ with an additional action of known 
	value $\mu$, defined by:
	\begin{align*}
		R_\mu(s,E,s')      &= R(s,E,s') \\
		R_\mu(s,\bot,\bot) &= \max\{\mu, R(s,\bot,\bot)\}
	\end{align*}
\end{dfn}

\begin{thm}
	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$, 
	there exists a real interval $I(s)$ for every state $s\in I$ such that
	it is optimal to stop in state $s$ in $M_\mu$ iff $\mu\in I(s)$.
	Furthermore, $I(s)$ contains $\max_i\mu_i(s)$ whenever it is nonempty.
\end{thm}

\begin{proof}

With $m(s) = \max_i\mu_i(s)$, the utility of a policy $\pi$ starting in state $s$ of $M_\mu$ is
\[
	V^\pi_{M_\mu}(s) = \IE_{\pi}[-c\,N + \max(\mu,m(S_N)) \given S_0=s]
\]
and the utility of stopping in this state $\max(\mu,m(s_0))$.
We wish to show that the set of $\mu$ such that
\[
	\max_\pi \IE_{\pi}[-c\,N + \max(\mu,m(S_N)) - \max(\mu,m(S_0)) \given S_0=s] \le 0
\]
forms an interval.

Observe that for any random variable $X$, 
	$\IE[\max(\mu,X)]$ is monotonically increasing in $\mu$ with subderivative between zero and one.
As a result, for any $v_1$
	$\IE[\max(\mu,X)] - \max(\mu,v_1)$ 
	is monotonically increasing for $\mu<v$, 
	and monotonically decreasing thereafter.
Therefore, the set if $\mu$ such that this expression is at most $v_2$ forms an interval, containing $v_1$ if non-empty.

Applying this with $v_1 = m(s_0)$ and $v_2=\IE_{\pi}[c\,N]$, and observing that the union of intervals containing
a point is an interval containing that point, gives the result.
\end{proof}

%% \note{Example showing that choice between computations to perform 
%% can vary arbitrarily many times?  Does one exist?}

### 2012-03-28 paper/vago.uai/introduction.tex
fixed typo

The broad family of sequential decision problems includes
combinatorial search problems, game playing, robotic path planning,
model-predictive control problems, Markov decision processes (fully or
partially observable), and a huge range of applications. In almost all
realistic instances, exact solution is intractable and approximate
methods are sought.  Perhaps the most popular approach is to simulate
a limited number of possible future action sequences. For example,
a typical game-playing algorithm explores a tree or graph of action sequences
with some limit on depth, using pruning methods to avoid irrelevant
subtrees; based on what it finds in the explored portion of the state space, the algorithm
then selects a move.

Clearly, it is desirable to select the best possible move from the
least possible amount of exploration. For a given amount of
exploration, decision quality can be improved by directing exploration
towards those actions sequences whose outcomes are helpful in selecting
a good move. Thus, the {\em metalevel} decision problem is to choose
what future action sequences to explore (or, more generally, what
deliberative computations to do), while the {\em object-level}
decision problem is to choose an action to execute in the real world.

That the metalevel decision problem can itself be formulated and
solved decision-theoretically was noted by \citet{Matheson:1968},
borrowing directly from the related concept of {\em information value
  theory}~\citep{Howard:1966}. In essence, computations can be
selected according to the expected improvement in decision quality
resulting from their execution. I. J.~\citet{Good:1968} independently
proposed using this idea to control search in chess, and later defined
``Type II rationality'' to refer to agents that optimally solve the
metalevel decision problem before acting. As interest in probabilistic
and decision-theoretic approaches in AI grew during the 1980s, several
authors explored these ideas further~\citep{Dean+Boddy:1988,Doyle:1988,Fehling+Breese:1988,Horvitz:1987b}.
Work by
\citet{Russell+Wefald:1988a,Russell+Wefald:1991a,Russell+Wefald:1991b}
formulated the metalevel sequential decision problem, employing an
explicit model of the results of computational actions, and applied
this to the control of game-playing search in Othello with encouraging
results.

An independent thread of research on metalevel control began with work
by \citet{Kocsis+Szepesvari:2006} on the UCT algorithm, which operates
in the context of {\em Monte Carlo tree search} (MCTS) algorithms.
In MCTS, each computation takes the form
of a simulating a randomized sequence of actions leading from a leaf of the
current tree to a terminal state. UCT is primarily a method for
selecting a leaf from which to conduct the next simulation.
It forms the core of the successful \textsc{MoGo} algorithm for Go playing \citep{Gelly+Silver:2011}.  It is situated 
within the theory of bandit problems \citep{Berry+Fristedt:1985}, applying the bandit algorithm
UCB1 \citep{Auer+et+al:2002} recursively to select actions to perform within simulations.

It is natural to consider whether the two independent threads are
consistent; for example, are bandit algorithms such as UCB1 approximate
solutions to some particular case of the metalevel decision problem
defined by Russell and Wefald? The answer, perhaps surprisingly, is no.
The essential difference is that, in bandit problems, every trial involves
executing a real object-level action with real costs, whereas in the metareasoning
problem the trials are {\em simulations} whose cost is usually independent of the
utility of the action being simulated. 
Hence, in UCT, bandit algorithms are applied to problems that are not bandit problems.

One consequence of the mismatch is that bandit policies are
inappropriately biased away from exploring actions whose current
utility estimates are low.  Another consequence is the absence of any
notion of ``stopping'' in bandit algorithms, which are designed for
infinite sequences of trials.  A metalevel policy needs to decide
when to stop deliberating and execute a real action.

Analyzing the metalevel problem within an appropriate theoretical
framework ought to lead to more effective algorithms than those
obtained within the bandit framework.  For Monte Carlo computations,
in which samples are gathered to estimate the utilities of actions,
the metalevel decision problem is an instance of the {\em selection
problem} studied in statistics~\cite{early-selection-refs}.  Despite
recent work by \citet{Frazier+Powell:2010}, the theory of selection
problems is less well understood than that of bandit problems.
Accordingly, we present in \secrefs{optimal-section}{context-section}
a number of results concerning optimal policies for the general case
and for specific sampling distributions. In particular, we give finite
bounds on the number of samples collected by optimal policies for
Bernoulli and Gaussian distributions. We also provide a simple
counterexample to the intuitive conjecture that an optimal policy
should not spend more on deciding than the decision is worth; in fact,
it is possible for an optimal policy to compute forever. We also show
by a simple counterexample that {\em index
policies}~\cite{Gittins:1989} do not necessarily exist for selection
problems.

Motivated by this theoretical analysis, we propose in \secrefs{approx-bayesian-section}{approx-nonbayesian-section}
two families of heuristic approximations, one for the Bayesian case and one for the distribution-free setting.
We show empirically that these rules give better performance than UCB1 on a wide range of standard (non-sequential) selection problems.
\secref{mcts-section} shows similar results for the case of guiding Monte Carlo tree search in the game of Go.

### 2012-03-28 paper/vago.uai/optimalpolicies.tex
fixed typo

%% \section{On optimal policies for selection}

%% Define selection problem and metalevel MDP representation.

%% \note{Unsure about terminology here: selection problems, sampling problems, 
%% metalevel probability model, metalevel decision problem (conflicts with Markov decision process?).}

%% ============================
\subsection{Selection problems}
%% ============================

\note{Formal definition of selection problems and the metalevel MDP with cost per sample (time value); also mention budgeted learning.}

In a \term{selection problem} the decision maker is faced with the choice between
$k$ alternative actions.  To make this choice, they may gather evidence about the
utility of each of these alternative, at some cost.  What the decision maker can know
and what they can find out is formalized in:

\begin{dfn} \label{dfn:metalevel-model}
	A \term{metalevel probability model} is a tuple $M=(U_1,\dots,U_k,\Evidence)$ 
	consisting of jointly distributed random variables:
	\begin{itemize}
		\item Real random variables $U_1,\dots,U_k$, where $U_i$ is the utility of performing action~$i$, and
		\item A countable set $\Evidence$ of random variables, each variable $E\in\Evidence$ being 
		      a computation that can be performed and whose value is the result of that computation.
	\end{itemize}
\end{dfn}
%% Not actually true that we necessarily define this with U's as roots:  in sampling models they are naturally /leaves/.

A metalevel probability model, when combined with a cost $c>0$ of computation,%
\footnote{The assumption of a fixed cost of computation is a simplification; 
	precise conditions for its validity are given by \citep{Harada:1997}.} 
defines a metalevel decision problem: what is the optimal strategy with which to choose a sequence 
of computations $E\in\Evidence$ to observe in order to maximize the expected utility 
of the action finally taken less the cost of computation?  Intuitively, 
this strategy should observe the variables which give the most evidence relevant
to deciding which action to perform, stopping when the cost of computation 
outweighs the benefit gained.

\begin{example}[Bernoulli sampling]\label{example:bernoulli}
One simple metalevel probability model is used as a model of the results of simulations \cite{someone}.
Each action will either succeed or not $U_i\in\{0,1\}$, with an unknown latent frequency of success $\Theta_i$, 
and stochastic simulations of possible consequences $E_{ij}$ can be performed:
\begin{align*}
	\Theta_i &\iid {\rm Uniform}[0,1]                      &&\quad\text{for $i=1,\dots,k$}\\
	U_i \given \Theta_i &\sample {\rm Bernoulli}(\Theta_i) &&\quad\text{for $i=1,\dots,k$}\\
	E_{ij} \given \Theta_i &\iid {\rm Bernoulli}(\Theta_i) &&\quad\text{for $i=1,\dots,k$ and $j=1,\dots$}
\end{align*}
\end{example}

%% Bounedness assumption
For simplicity, in the below we'll assume the utilities $U_i$ are bounded and so, 
without loss of generality, lie in $[0,1]$.

To formalize this problem, we use the theory of Markov Decision Processes (MDPs; 
see, for example, \citet{Puterman:1994}):

\begin{dfn}
A (countable state, undiscounted) \term{Markov Decision Process} (MDP) is a tuple $M=(S,s_0,A_s,T,R)$ where:
	$S$ is a countable set of states,
	$s_0\in S$ is the fixed initial state,
	$A_s$ is a countable set of actions available in state $s\in S$,
	$T(s,a,s')$ is the transition distribution 
	giving the probability of transitioning from a state $s\in S$ to a state $s'\in S$ after performing action $a\in A_s$,
	and $R(s,a,s')$ is the expected reward received when transitioning from $s\in S$ to $s'\in S$ after performing action $a\in A_s$.
\end{dfn}

\begin{dfn}\label{dfn:metalevel-mdp}
	Given a metalevel probability model%
		\footnote{Definition~\ref{dfn:metalevel-model} made no assumption about the computational result
			variables $E_i\in\Evidence$, but for simplicity in the following we'll assume that
			each $E_i$ takes one of a countable set of values.  Without loss of generality, 
			we'll further assume the domains of the computational variables $E\in\Evidence$ are disjoint.}
	$M=(U_1,\dots,U_k,\Evidence)$ and
	a cost of computation $c>0$, a corresponding \term{metalevel decision problem}
	is the MDP $M=(S,s_0,A_s,T,R)$ with
	\begin{align*}
		S &= \{\bot\}\cup\{\langle e_1\dots, e_n \rangle : e_i\in E_i \text{ for all $i$,} \\
								& \qquad\qquad \text{for finite $n\ge0$ and distinct $E_i\in\Evidence$}\} \\
		s_0 &= \langle \rangle \\
		A_s &= \{\bot\}\cup\Evidence_s \\
	%
	\intertext{where $\bot\in S$ is the unique terminal state,
	where $\Evidence_s\subseteq\Evidence$ is a potentially state-dependent subset of allowed computations,
	and when given any $s=\langle e_1, \dots, e_n \rangle\in S$,
	computational action $E\in\Evidence$, 
	and $s'= \langle e_1, \dots, e_n, e \rangle\in S$ where $e\in E$, we have:}
	%
		T(s,E,s') &= P(E=e \given E_1=e_1,\dots,E_n=e_n) \\
		T(s,\bot,\bot) &= 1 \\
		R(s,E,s') &= -c \\
		R(s,\bot,\bot) &= \max_i \mu_i(s) % \max_i \IE[U_i \given E_1=e_1,\dots,E_n=e_n]
	\end{align*}
	where $\mu_i(s) = \IE[U_i \given E_1=e_1,\dots,E_n=e_n]$.
\end{dfn}

\begin{example}[Bernoulli sampling]\label{example:bernoulli2}
In the Bernoulli metalevel probability model (\exampleref{example:bernoulli}),
note that: 
\begin{align}
	\Theta_i \given E_{i1},\dots,E_{in_i} &\sim {\rm Beta}(s_i+1, f_i+1)  \label{eq:bernoulli1}\\
	E_{i(n_i+1)} \given E_{i1},\dots,E_{in_i} &\sim {\rm Bernoulli}\left(\frac{s_i+1}{n_i+2}\right) \label{eq:bernoulli2} \\
	\IE[U_i \given E_{i1},\dots,E_{in_i}] &= (s_i+1)/(n_i+2) \label{eq:bernoulli3}
\end{align}
by standard properties of these distributions, where $s_i=\sum_{j=1}^{n_i} E_{in_i}$
is the number of simulated successes of action $i$, and $f_i=n_i-s_i$ the failures.  By \eqref{eq:bernoulli1}, 
the state space is the set of all $k$ pairs $(s_i,f_i)$, and \eqrefs{eq:bernoulli2}{eq:bernoulli3}
suffice to give the transition probabilities and terminal rewards, respectively.
\end{example}

Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$ one defines policies and 
value functions as in all MDPs.  A (deterministic, stationary) \term{metalevel policy} 
$\pi$ is a function mapping states $s\in S$ to actions to take in that state $\pi(s)\in A_s$.
The \term{value function} for a policy $\pi$ gives the expected total reward
received under that policy starting from a given state $s\in S$: 
\begin{equation}\label{eq:value-fn}
	V^\pi_M(s) = \IE^\pi_M\left[ \sum_{i=0}^N R(S_i,\pi(S_i),S_{i+1}) \given S_0 = s\right]	
\end{equation}

where $N\in[0,\infty]$ is the random time the MDP is terminated, i.e.,
the unique time where $\pi(S_N)=\bot$.%
\footnote{Whenever $N<\infty$, $S_{N+1}=\bot$ the unique terminal state.
Instead of having a random termination time, one can equivalently make
the state $\bot$ absorbing with zero reward on all transitions.}

\begin{thm}\label{thm:value-of-computation}
	The value function of a metalevel decision process $M=(S,s_0,A_s,T,R)$ is of the form
	\[
		V^\pi_M(s) = \IE^\pi_M[ -c\,N + \max_i \mu_i(S_N) \given S_0=s]
	\]
	where the random variable $N$ denotes the total number of computations performed.
\end{thm}

\begin{proof}
	Follows immediately from \eqref{eq:value-fn} and the definition of the
	reward function $R$ in \dfnref{dfn:metalevel-mdp}.
\end{proof}

\begin{thm}
	MDPs $M=(S,s_0,A_s,T,R)$ where there exists a unique terminal state $\bot\in S$,
	a distinguished action $\bot\in A_s$ for all $s\in S$, a constant $c>0$, and bounded functions 
	$\mu_1,\dots,\mu_k\maps S\to\R$ such that for all $s\in S$, $a\in A_s\setminus\{\bot\}$, and $i=1,\dots,k$:
	\begin{align*}
		\mu_i(s) &= \sum_{s'\in S} T(s,a,s')\,\mu_i(s') \\
		T(s,\bot,\bot) &= 1 \\
		R(s,a,s') &= -c \\
		R(s,\bot,\bot) &= \max_i \mu_i(s)
	\end{align*}
	are equivalent to metalevel decision problems with bounded utilities, and conversely any 
	metalevel decision problem with bounded utilities $U_i$ is such an MDP.
\end{thm}

\begin{proof}
Given such an MDP, for simplicity will assume that $\mu_i(s)\in(0,1)$ for all $i$ and $s\in S$; the proof generalizes 
to any bounds $(\mu^-,\mu^+)$.  We assume further the MDP is acyclic, which can be ensured, e.g., by augmenting
the state space with a time variable incrementing on each transition.

Define $\{0,1\}$-valued random variables $U_1,\dots,U_k$ 
and $E_{sa}$ for all $s\in S$, $a\in A_s$ by
\begin{align*}
	P(U_i=1) &= \mu_i(s_0) \\
	P(E_{sa}=s'\given U_i=1) &= \frac{T(s,a,s')\,\mu_i(s')}{\mu_i(s)} \\
	P(E_{sa}=s'\given U_i=0) &= \frac{T(s,a,s')\,(1-\mu_i(s'))}{1- \mu_i(s)}
\end{align*}
where $U_1,\dots,U_k$ are independent, and the $E_{sa}$ are conditionally 
independent given $(U_1,\dots,U_k)$.
These together form a metalevel probability model, so we can form the 
corresponding metalevel decision problem following definition~\ref{dfn:metalevel-mdp},
to avoid conflict of notation we denote the latter by $M'=(H,h_0,B_h,T',R')$,
and where since $B_h$ is underspecified in the definition we define
\[
	B_h = \{\bot\} \cup \{E_{t(h)a} : a \in A_{t(h)a}\}
\]
where the tail function $t\maps H\to S$ has $t(\langle\rangle)=s_0$ and otherwise
\[
	t(\langle e_{s_0a_0},\dots,e_{s_n a_n}\rangle) = e_{s_n a_n} \in S.
\]
This definition of $B_h$ means that all elements of $H$ reachable from $h_0=\langle\rangle$ with positive probability are 
sequences $h=\langle e_{s_0a_0},\dots,e_{s_n a_n}\rangle$
with $e_{s_ia_i} = s_{i+1} \in S$ and $a_i\in A_{s_i}$, i.e., are possible state sequences 
in the original MDP $M$ under some sequence of actions.  Fix such a sequence, and
note that given $E_{s_n a}\in B_h$, any $h'$ with positive transition probability
is the form $h'=\langle e_{s_0a_0},\dots,e_{s_n a_n},e_{s_{n+1}a}\rangle$, and
\[
	T'(h,E_{s_n a},h') = T(t(h), a, t(h')).
\]

Finally, observe that by acyclicity and construction
\begin{align*}
	& P(U_i=1, E_{s_0a_0}=s_1,\dots,E_{s_n a_n}=s_{n+1}) \\
	&= \mu_i(s_0) \prod_{i=0}^n \frac{T(s_i,a_i,s_{i+1})\,\mu_i(s_{i+1})}{\mu_i(s_i)} \\
	&= \mu_i(s_{n+1}) \prod_{i=0}^n T(s_i,a_i,s_{i+1}),
%
\intertext{and similarly}
%
	& P(U_i=0, E_{s_0a_0}=s_1,\dots,E_{s_n a_n}=s_{n+1}) \\
	&= (1-\mu_i(s_{n+1})) \prod_{i=0}^n T(s_i,a_i,s_{i+1}).
\end{align*}
As a result
\begin{align*}
	& \IE(U_i\given E_{s_0a_0}=s_1,\dots,E_{s_n a_n}=s_{n+1})\\
	& = \mu_i(s_n) = \mu_i(t(h))\\
	& P(E_{s_n a_n}=s_{n+1} \given E_{s_0a_0}=s_1,\dots,E_{s_{n-1} a_{n-1}}=s_{n}) \\
	&=  T(s_n,a,s_{n+1}) = T(t(h),a,t(h'))
\end{align*}

Thus the metalevel decision problem $M'$ is equivalent to the original MDP $M$
under the mapping $t\maps H\to S$.

For the converse, observe that the properties hold with $c$ the cost of computation
and \[\mu_i(s) = \IE(U_i\given E_1=e_1,\dots,E_n=e_n)\] where $s=\langle e_1,\dots,e_n \rangle$.
\end{proof}

As usual, an \term{optimal policy} $\pi^*$ is one which maximises 
the value from every state $s\in S$, i.e., if we define for each $s\in S$
\[
	V^*_M(s) = \sup_\pi V^\pi_M(s),
\]
then an optimal policy $\pi^*$ satisfies $V^{\pi^*}_M(s) = V^*_M(s)$
for all $s\in S$.

\note{Something about existence of optimal solutions?  Probably do not exist in general,
but do for finite $S$ despite discounting as this is a positive bounded model.}

%% Positive bounded model.  We do have upper and lower bounds on V^*.  For finite s we can get optimality...

\begin{dfn}
	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$,
	the \term{myopic policy} $\pi^m(s)$ is defined to equal
	\begin{equation}\label{eq:myopic}
		\argmax_{E\in A_s} \IE_M[ -c + \max_i \mu_i(S_1) \given S_0 = s, A_0 = E]		
	\end{equation}
	whenever this maximum is positive, $\pi^m(s)=\bot$ otherwise.
\end{dfn}

The myopic policy takes the best action, to either stop or perform a computation,
under the assumption that at most one further computation can be performed.  This policy
is typically easy to compute, although not necessarily very good.  It does have
importance, however, through its relation to the optimal policy in \thmref{thm:optimal-myopic}
and its partial converse \thmref{thm:myopic-optimal}:

%% \note{Theorem: if Optimal stops in x, myopic stops in x (converse is more useful!)}

\begin{thm}\label{thm:optimal-myopic}
	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$
	if the optimal policy stops in a state $s\in S$
	then the myopic policy stops too, i.e., if $\pi^*(s)=\bot$ then $\pi^m(s)=\bot$.
\end{thm}

\begin{proof}
	If the optimal policy stops in a state $s\in S$ then
	\[
		V^{\pi^*}(s) \le \max_i \mu_i(s),
	\]
	but then the maximum of \eqref{eq:myopic} is non-positive, so the 
	myopic policy stops by definition.
\end{proof}

\begin{dfn}
	Given an MDP $M=(S,s_0,A_s,T,R)$ a state $s'$ is \term{reachable} from a state $s$
	if there is a sequence of states $s=s_0,s_1,\dots,s_n=s'$ and actions $a_i\in A_{s_i}$
	with positive probability under the transition distribution $T$.
\end{dfn}

%% \note{Theorem: if Myopic stops in all states reachable from x, optimal stops in x}

\begin{thm}\label{thm:myopic-optimal}
	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$
	if the myopic policy stops in all states $s'\in S$ reachable
	from a given state $s$ then the optimal policy stops too.
	then the myopic policy stops too, i.e., if $\pi^*(s)=\bot$ then $\pi^m(s)=\bot$.
\end{thm}

\begin{proof}
Defining $m(s) = \max_i\mu_i(s)$, observe the myopic stopping for all
reachable states implies that
\begin{align*}
	\IE^{\pi}[m(S_{j+1}) - c; j<N\given S_0=s] \\
	\le \IE^{\pi}[m(S_{j}); j<N\given S_0=s]
\end{align*}
holds for all $j$, and as a result:
\begin{align*}
	V^\pi(s) 
	&= \IE^{\pi}[ - c N + m(S_N) |S_0=s] \\
	&= \IE^{\pi}[m(S_0) + \sum_{j=0}^{N-1} (m(S_{j+1}) - c - m(S_j)) |S_0=s] \\	
	&\le \IE^{\pi}[m(S_0) + \sum_{j=0}^{N-1} 0 |S_0=s] \\		
	&= \max_i\mu_i(s)
\end{align*}
\end{proof}

% Similarity to an optimal stopping result
	% Search <blah> for one-step lookahead rules.
	% But HERE is generalized to a stopping problem where there are many ways to continue.

\note{Define value of perfect information here for the below?}

\begin{thm}\label{thm:bounded-expected-computations}
	The optimal policy $\pi^*$ has bounded expected number of computations
	\[
		\IE^{\pi^*}[N\given S_0=s] \le \frac{1}{c} \left(\IE[\max_i U_i\given S_0=s] - \max_i \mu_i(s)\right).
	\]
	Further, any policy $\pi$ with infinite expected number of computations 
	$\IE^\pi[N]=\infty$ has negative infinite value, so in particular the optimal
	policy stops with probability 1.
\end{thm}

\begin{proof}
	The first follows as in state $s$ the optimal policy has value at least that
	of stopping immediately $\max_i \mu_i(s)$, and as $\IE \max_i\mu_i(S_N) \le \IE \max_i U_i$ by Jensen's inequality.
	The second follows from \thmref{thm:value-of-computation} as $c>0$.
\end{proof}

Although the \emph{expected} number of computations is always bounded,
there are important cases in which the \emph{actual} number is not:

\begin{example}\label{example:sprt}
Consider the Bernoulli sampling model for two actions but with different priors:
	$\Theta_1=1/2$,
	and $\Theta_2$ is $1/3$ or $2/3$ with equal probability.
%
Simulating action 1 gains nothing, and after $(s,f)$ simulated successes and failures of action 2
the posterior odds is
\[
	\frac{P(\Theta_2=2/3\given s,f)}{P(\Theta_2=1/3\given s,f)} = 2^{s-f}.
\]
The difference $s-f$ is a sufficient statistic for $\Theta$ given the computations,
so there exists a metalevel MDP with this difference as its state.  For sufficiently
low cost, the optimal policy samples in states $-1$, $0$, and $1$.  But with probability
$1/3$ state $0$ transitions to state $0$ after two samples, giving finite probability
to arbitrarily long sequences of computations.
\end{example}

However in a number of settings, including the original Bernoulli model of \exampleref{example:bernoulli},
we can upper bound the number of computations:

\begin{thm}\label{thm:bernoulli-bound}
In the Bernoulli sampling problem of \exampleref{example:bernoulli},
at most $k/4c$ computations are performed.	
\end{thm}

\begin{proof}

\end{proof}

As an important consequence, we can truncate the MDP to

\note{Gaussian bound too?}

%% \note{Reference ranking and selection here?}

The bound in \thmref{thm:bernoulli-bound} allows us

\begin{dfn}
	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$,
	the \term{blinkered policy} $\pi^b(s)$ \term{myopic policy} $\pi^m(s)$ is defined to equal
	\begin{equation}\label{eq:blinkered}
		V^b(s) = \sup_{\pi\in\Pi_b} V^\pi_M(s),
	\end{equation}
	where $\Pi_b$ is the set of policies which sample .....

	...need comparisons between $A_s$'s to exist.
\end{dfn}

\note{Graph comparing blinkered to other policies}

\begin{figure}[htb]
\centering
\includegraphics[scale=0.7, trim=90 70 400 300]{blinkered-regret.pdf}
\caption{Regret vs. cost}
\label{fig:blinkered}
\end{figure}

\begin{comment}
set xlabel "Cost"
set ylabel "Regret"

set terminal postscript enhanced linewidth 2
set output "blinkered-regret.ps"
set size 0.5, 0.5

set logscale xy

plot [0.1:0.0005] [0.05:1] "blinkered-regret.dat" title "Blinkered" with lines lw 2, "espb-regret.dat" title "ESPb" with lines lw 2, "ucb-blinkered-regret.dat" title "UCB1-B" with lines lw 2, "ucb-blinkered-table-regret.dat" with lines lw 2 title "UCB1-b"	
\end{comment}

### 2012-03-30 paper/vago.uai/blinkered.tex
added files, probably too many

%% \section{The blinkered policy}

%% This is the key section.  Write it, and then see what prereqs are required.

The myopic policy is an extreme approximation, often stopping significantly early.
For a natural class of metalevel probability models we can define a better approximation:

\begin{dfn}\label{dfn:independent-actions}
	A metalevel probability model $\mathcal{M}=(U_1,\dots,U_k,\Evidence)$ 
	has \term{independent actions} if the computational variables can be partitioned 
	$\Evidence = \Evidence_1\cup\dots\cup\Evidence_k$ such that such that
	the sets $\{U_i\}\cup\Evidence_i$ are independent of each other for different $i$.	
\end{dfn}

\begin{dfn}\label{dfn:blinkered}
	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$ with independent actions,
	the \term{blinkered policy} $\pi^b$ is defined by $\pi^b(s) = \argmax_{a\in A_s} Q^b(s,a)$ where
	$Q^b(s,\bot) = \bot$ and for $E_i\in\Evidence_i$
	\begin{equation}\label{eq:blinkered}
		Q^b(s,E_i) = \sup_{\pi\in\Pi^b_i} Q^\pi(s,E_i)
	\end{equation}
	where $\Pi^b_i$ is the set of policies $\pi$ where $\pi(s)\in\Evidence_i$ for all $s\in S$.
\end{dfn}

This is a better approximation than the myopic policy as it takes into account more possible future strategies.
Indeed, $Q^m(s,a) \le Q^b(s,a) \le Q^*(s,a)$.  Intuitively, whereas the myopic policy can only see one 
computation ahead, the blinkered policy can see arbitrary far ahead but with blinkers on so can only see
the computations involving a single action at a time.

\begin{dfn}\label{dfn:one-action}
	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$ with independent actions,
	a \term{one-action metalevel decision problem} for $i=1,\dots,k$ is the metalevel decision
	problem $M^1_{i,m} = (S_i,s_0,A_{s0},T_i,R_i)$ defined by the metalevel probability
	model $(U_0,U_i,\Evidence_i)$ with $U_0=m$.
\end{dfn}

Note that given that state $s$ of a metalevel decision problem, we can form a state
$s_i$ by taking only the results of computations in $\Evidence_i$ (see \dfnref{dfn:metalevel-mdp}).
By action independence, $\mu_i(s)$ is a function only of $s_i$.

\begin{thm}\label{thm:blinkered}
Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$ with independent actions,
let $M^1_{i,\mu}$ be the $i$th one-action metalevel decision problem for $i=1,\dots,k$.
Then for any $s\in S$, whenever $E_i\in A_s\cap\Evidence_i$ we have:
\[
	Q^b_M(s,E_i) = Q^*_{M^1_{i,m_i}}(s_i, E_i)
\]
where $m_i = \max_{j\neq i} \mu_i(s)$.
\end{thm}

\begin{comment}
	\begin{proof}
	Fix a state $s$, a $E_i\in A_s$ and take any $\pi\in\Pi^b_i$.  Note that such policies
	are equivalent to polices $\pi'$ on $M^1_{1,m}$, and all such policies are represented.
	Consider $Q^\pi(s,E_i)$.  As $\pi(s)\in\Evidence_i$ for all $s\in S$, by action independence $\mu_j(S_n) = \mu_j(s)$.
	By this and \thmref{thm:value-of-computation}, then,
	\[
		Q^\pi(s,E_i) = \IE^\pi_M[ -c\,N + \max(\mu_i(S_N), m_i) \given S_0=s, A_0=E_i].
	\]
	Noting that $\mu_i(S_N)$ is a function only of $(S_N)_i$, and that since 
	But then this is exactly $Q^{\pi'}_{M^1_{i,m_i}}(s_i, E_i)$.  Taking the supremum
	over $\pi$ gives the result.
	\end{proof}	
\end{comment}

\thmref{thm:blinkered} shows that to compute the blinkered policy we need
only compute the optimal policies for all $i$th one-action metalevel decision problems.

For the Bernoulli problem with $k$ actions, the one-action metalevel decision problems
are all one-action Bernoulli problems (\exampleref{example:bernoulli}).  By \thmref{thm:one-action-bound}
these policies perform at most $1/4c - 3$ computations.
As a result, the blinkered policy can be numerically computed in time $O(D/c^2)$ 
independent of $k$ by backwards induction, where $D$ is the number of points $m\in[0,1]$
we compute $Q^*_{M^1_{i,m}}(s)$ for\footnote{in our experiments below, $D=129$ points are equally spaced,
using linear interpolation between points}.  This will be worth the cost in 
the same situations as mentioned at the of \secref{sec:optimal}.

\figref{fig:blinkered} compared the expected regret of the blinked policy on the 
Bernoulli sampling problem with $k=25$ and varying cost of samples to several 
alternatives, where the regret include the cost of sampling:
\[
	R = (\max_i U_i) - U_j + c\,n
\]
if the policy performs $n$ computations and selects action $j$.
Myopic is the myopic policy for Bernoulli sampling. ESPb is the 
policy recently proposed by \citet{Chick+Frazier:2011}, which approximates the 
blinkered policy for sampling normal distributions by moving to continuous time
and numerically solving a diffusion equation.  UCB1-B uses UCB$(1/\sqrt2)$ \cite{Auer+et+al:2002} to
choose which actions to sample and the blinkered policy to stop.  UCB1-b uses
UCB$(1/\sqrt2)$ to sample, is given the average number of samples blinkered
uses for the given cost.

The blinkered policy significantly outperforms all others.  The myopic policy
plateaus as it quickly reaches a position where no single computation can
change the final decision between actions.  ESPb performs quite well given
that is making a normal approximation to the Beta posterior.  
UCB1-B and UCB1-b's curve shows that even given a good stopping rule, UCB1's
choice of actions to sample can be significantly improved on.

\begin{figure}[htb]
\centering
\includegraphics[scale=0.7, trim=90 70 400 300]{blinkered-regret.pdf}
\caption{Average regret of various policies as a function of the cost in 
a 25-action Bernoulli sampling problem.  Averaged over 1000 trials.}
\label{fig:blinkered}
\end{figure}

\begin{comment}
set xlabel "Cost"
set ylabel "Regret"

set terminal postscript enhanced linewidth 2
set output "blinkered-regret.ps"
set size 0.5, 0.5

set logscale xy

set key left bottom
plot [0.1:0.0005] [0.05:1] "blinkered-regret.dat" title "Blinkered" with lines lw 2, "myopic-regret.dat" title "Myopic" with lines lw 2, "espb-regret.dat" title "ESPb" with lines lw 2, "ucb-blinkered-regret.dat" title "UCB1-B" with lines lw 2, "ucb-blinkered-table-regret.dat" with lines lw 2 title "UCB1-b"
\end{comment}

### 2012-03-30 paper/vago.uai/context.tex
added files, probably too many

@@
 
-\note{No index theorem; via context inversion counter-example}
+%% \note{No index theorem; via context inversion counter-example}
 
-de
+The Gittins index theorem \citep{Gittins:1979} is a famous structural
+result for bandit problems.  It states that in bandit problems with
+independent reward distribution for each arm and geometric discounting,
+the optimal policy is an \term{index policy}:  each arm is assigned a
+real-valued index based on its state only, such that it is optimal to
+sample the arm with greatest index.
+
+The analogous result does \emph{not} hold for metalevel decision problems,
+even when the action's values are independent (this formalized later in \dfnref{dfn:independent-actions}):
+
+\begin{example}[Non-indexability]
+Consider a metalevel probability model with three actions. 
+	$U_0=m$ has a known value (the context),
+	$U_1$ is equally likely to be -1.5 or 1.5 (low mean, high variance),
+	and $U_2$ is equally likely to be 0.25 or 1.75 (high mean, low variance).
+The two computations are to observe exactly $U_1$ and $U_2$, respectively, with a cost 0.2 of computation.
+The corresponding metalevel MDP has 9 states and can be solved exactly.
+\figref{fig:swap-counterex} plots $Q^*_m(s_0,U_i) - Q^*_m(s_0,\bot)$ as a function of the known value $m$.
+As the context $m$ varies the optimal action inverts from observing 1 to observing 2.
+Inversions like this are impossible for index policies.
+\end{example}
 
@@
 \includegraphics[scale=0.7, trim=90 70 400 300]{swap-counterex.pdf}
-\caption{Regret vs. cost}
+\caption{Optimal Q-values as a function of the utility of the fixed alternative.  Note the inversion
+where for low $m$ observing action~$1$ is strictly optimal, while for medium $u$ observing action~$2$ is strictly optimal.}
 \label{fig:swap-counterex}
@@
 set output "swap-counterex.ps"
-set xlabel "Utility of fixed alternative (u)"
+set xlabel "Utility of fixed alternative (m)"
 set ylabel "Utility relative to stopping"
@@
 set key left top
-plot [u=-2:2] [-0.3:0.3] 0 title "stop" with lines lw 2, qa(u)-q0(u) title "observe A" with lines lw 2, qb(u)-q0(u) title "observe B" with lines lw 2	
+plot [u=-2:2] [-0.3:0.3] 0 title "stop" with lines lw 2, qa(u)-q0(u) title "observe 1" with lines lw 2, qb(u)-q0(u) title "observe 2" with lines lw 2	
 \end{comment}
@@
 %%    context value (how general can we make this?)}
+
+There is, however, a restriction on what kind of influence the context can have:
 
@@
 	there exists a real interval $I(s)$ for every state $s\in I$ such that
-	it is optimal to stop in state $s$ in $M_\mu$ iff $\mu\in I(s)$.
+	it is optimal to stop in state $s$ in $M_\mu$ iff $\mu\notin I(s)$.
 	Furthermore, $I(s)$ contains $\max_i\mu_i(s)$ whenever it is nonempty.
@@
 
+\begin{comment}
 	\begin{proof}
-
 	With $m(s) = \max_i\mu_i(s)$, the utility of a policy $\pi$ starting in state $s$ of $M_\mu$ is
@@
 	\end{proof}	
+\end{comment}
 

### 2012-03-30 paper/vago.uai/dropped-parts.tex
added files, probably too many

--- 2012-03-28 paper/vago.uai/optimalpolicies.tex
@@
-%% \section{On optimal policies for selection}
+\begin{comment}
+	\begin{lmm}\label{lmm:equivalent-mdp}
+		Given MDPs $M^1=(S^1,s^1_0,A^1_s,T^1,R^1)$ and $M^2=(S^2,s^2_0,A^2_s,T^2,R^2)$,
+		if there exists $f\maps S^1\to S^2$ such that for all $s,s'\in S^1$ and $a\in A_s$,
+		we have $A^2_{f(s)} = A^1_s$ and
+		\begin{align*}
+			T^2(f(s),a,f(s')) &= T(s_2,E,s_2') \\
+			\max_i \mu_i(s_1) &= \max_i \mu_i(s_2),\\
+		%
+		\intertext{then for any policy $\pi^2$ in $M^2$, with policy $\pi^1$ in $M^1$
+		defined by $\pi^1(s) = \pi^2(f(s))$ we have}
+		%
+			V^{\pi^1}_{M^1}(s) &= V^{\pi^2}_{M^2}(f(s)).
+		\end{align*}
+	\end{lmm}		
+\end{comment}
 
-%% Define selection problem and metalevel MDP representation.
-
-%% \note{Unsure about terminology here: selection problems, sampling problems, 
-%% metalevel probability model, metalevel decision problem (conflicts with Markov decision process?).}
-
-%% ============================
-\subsection{Selection problems}
-%% ============================
-
-\note{Formal definition of selection problems and the metalevel MDP with cost per sample (time value); also mention budgeted learning.}
-
-In a \term{selection problem} the decision maker is faced with the choice between
-$k$ alternative actions.  To make this choice, they may gather evidence about the
-utility of each of these alternative, at some cost.  What the decision maker can know
-and what they can find out is formalized in:
-
-\begin{dfn} \label{dfn:metalevel-model}
-	A \term{metalevel probability model} is a tuple $M=(U_1,\dots,U_k,\Evidence)$ 
-	consisting of jointly distributed random variables:
-	\begin{itemize}
-		\item Real random variables $U_1,\dots,U_k$, where $U_i$ is the utility of performing action~$i$, and
-		\item A countable set $\Evidence$ of random variables, each variable $E\in\Evidence$ being 
-		      a computation that can be performed and whose value is the result of that computation.
-	\end{itemize}
-\end{dfn}
-%% Not actually true that we necessarily define this with U's as roots:  in sampling models they are naturally /leaves/.
-
-A metalevel probability model, when combined with a cost $c>0$ of computation,%
-\footnote{The assumption of a fixed cost of computation is a simplification; 
-	precise conditions for its validity are given by \citep{Harada:1997}.} 
-defines a metalevel decision problem: what is the optimal strategy with which to choose a sequence 
-of computations $E\in\Evidence$ to observe in order to maximize the expected utility 
-of the action finally taken less the cost of computation?  Intuitively, 
-this strategy should observe the variables which give the most evidence relevant
-to deciding which action to perform, stopping when the cost of computation 
-outweighs the benefit gained.
-
-\begin{example}[Bernoulli sampling]\label{example:bernoulli}
-One simple metalevel probability model is used as a model of the results of simulations \cite{someone}.
-Each action will either succeed or not $U_i\in\{0,1\}$, with an unknown latent frequency of success $\Theta_i$, 
-and stochastic simulations of possible consequences $E_{ij}$ can be performed:
-\begin{align*}
-	\Theta_i &\iid {\rm Uniform}[0,1]                      &&\quad\text{for $i=1,\dots,k$}\\
-	U_i \given \Theta_i &\sample {\rm Bernoulli}(\Theta_i) &&\quad\text{for $i=1,\dots,k$}\\
-	E_{ij} \given \Theta_i &\iid {\rm Bernoulli}(\Theta_i) &&\quad\text{for $i=1,\dots,k$ and $j=1,\dots$}
-\end{align*}
-\end{example}
-
-%% Bounedness assumption
-For simplicity, in the below we'll assume the utilities $U_i$ are bounded and so, 
-without loss of generality, lie in $[0,1]$.
-
-To formalize this problem, we use the theory of Markov Decision Processes (MDPs; 
-see, for example, \citet{Puterman:1994}):
-
-\begin{dfn}
-A (countable state, undiscounted) \term{Markov Decision Process} (MDP) is a tuple $M=(S,s_0,A_s,T,R)$ where:
-	$S$ is a countable set of states,
-	$s_0\in S$ is the fixed initial state,
-	$A_s$ is a countable set of actions available in state $s\in S$,
-	$T(s,a,s')$ is the transition distribution 
-	giving the probability of transitioning from a state $s\in S$ to a state $s'\in S$ after performing action $a\in A_s$,
-	and $R(s,a,s')$ is the expected reward received when transitioning from $s\in S$ to $s'\in S$ after performing action $a\in A_s$.
-\end{dfn}
-
-\begin{dfn}\label{dfn:metalevel-mdp}
-	Given a metalevel probability model%
-		\footnote{Definition~\ref{dfn:metalevel-model} made no assumption about the computational result
-			variables $E_i\in\Evidence$, but for simplicity in the following we'll assume that
-			each $E_i$ takes one of a countable set of values.  Without loss of generality, 
-			we'll further assume the domains of the computational variables $E\in\Evidence$ are disjoint.}
-	$M=(U_1,\dots,U_k,\Evidence)$ and
-	a cost of computation $c>0$, a corresponding \term{metalevel decision problem}
-	is the MDP $M=(S,s_0,A_s,T,R)$ with
-	\begin{align*}
-		S &= \{\bot\}\cup\{\langle e_1\dots, e_n \rangle : e_i\in E_i \text{ for all $i$,} \\
-								& \qquad\qquad \text{for finite $n\ge0$ and distinct $E_i\in\Evidence$}\} \\
-		s_0 &= \langle \rangle \\
-		A_s &= \{\bot\}\cup\Evidence_s \\
-	%
-	\intertext{where $\bot\in S$ is the unique terminal state,
-	where $\Evidence_s\subseteq\Evidence$ is a potentially state-dependent subset of allowed computations,
-	and when given any $s=\langle e_1, \dots, e_n \rangle\in S$,
-	computational action $E\in\Evidence$, 
-	and $s'= \langle e_1, \dots, e_n, e \rangle\in S$ where $e\in E$, we have:}
-	%
-		T(s,E,s') &= P(E=e \given E_1=e_1,\dots,E_n=e_n) \\
-		T(s,\bot,\bot) &= 1 \\
-		R(s,E,s') &= -c \\
-		R(s,\bot,\bot) &= \max_i \mu_i(s) % \max_i \IE[U_i \given E_1=e_1,\dots,E_n=e_n]
-	\end{align*}
-	where $\mu_i(s) = \IE[U_i \given E_1=e_1,\dots,E_n=e_n]$.
-\end{dfn}
-
-\begin{example}[Bernoulli sampling]\label{example:bernoulli2}
-In the Bernoulli metalevel probability model (\exampleref{example:bernoulli}),
-note that: 
-\begin{align}
-	\Theta_i \given E_{i1},\dots,E_{in_i} &\sim {\rm Beta}(s_i+1, f_i+1)  \label{eq:bernoulli1}\\
-	E_{i(n_i+1)} \given E_{i1},\dots,E_{in_i} &\sim {\rm Bernoulli}\left(\frac{s_i+1}{n_i+2}\right) \label{eq:bernoulli2} \\
-	\IE[U_i \given E_{i1},\dots,E_{in_i}] &= (s_i+1)/(n_i+2) \label{eq:bernoulli3}
-\end{align}
-by standard properties of these distributions, where $s_i=\sum_{j=1}^{n_i} E_{in_i}$
-is the number of simulated successes of action $i$, and $f_i=n_i-s_i$ the failures.  By \eqref{eq:bernoulli1}, 
-the state space is the set of all $k$ pairs $(s_i,f_i)$, and \eqrefs{eq:bernoulli2}{eq:bernoulli3}
-suffice to give the transition probabilities and terminal rewards, respectively.
-\end{example}
-
-Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$ one defines policies and 
-value functions as in all MDPs.  A (deterministic, stationary) \term{metalevel policy} 
-$\pi$ is a function mapping states $s\in S$ to actions to take in that state $\pi(s)\in A_s$.
-The \term{value function} for a policy $\pi$ gives the expected total reward
-received under that policy starting from a given state $s\in S$: 
-\begin{equation}\label{eq:value-fn}
-	V^\pi_M(s) = \IE^\pi_M\left[ \sum_{i=0}^N R(S_i,\pi(S_i),S_{i+1}) \given S_0 = s\right]	
-\end{equation}
-
-where $N\in[0,\infty]$ is the random time the MDP is terminated, i.e.,
-the unique time where $\pi(S_N)=\bot$.%
-\footnote{Whenever $N<\infty$, $S_{N+1}=\bot$ the unique terminal state.
-Instead of having a random termination time, one can equivalently make
-the state $\bot$ absorbing with zero reward on all transitions.}
-
-\begin{thm}\label{thm:value-of-computation}
-	The value function of a metalevel decision process $M=(S,s_0,A_s,T,R)$ is of the form
-	\[
-		V^\pi_M(s) = \IE^\pi_M[ -c\,N + \max_i \mu_i(S_N) \given S_0=s]
-	\]
-	where the random variable $N$ denotes the total number of computations performed.
-\end{thm}
-
-\begin{proof}
-	Follows immediately from \eqref{eq:value-fn} and the definition of the
-	reward function $R$ in \dfnref{dfn:metalevel-mdp}.
-\end{proof}
-
+\begin{comment}
 	\begin{thm}
@@
 	\end{proof}
+\end{comment}
 
-As usual, an \term{optimal policy} $\pi^*$ is one which maximises 
-the value from every state $s\in S$, i.e., if we define for each $s\in S$
-\[
-	V^*_M(s) = \sup_\pi V^\pi_M(s),
-\]
-then an optimal policy $\pi^*$ satisfies $V^{\pi^*}_M(s) = V^*_M(s)$
-for all $s\in S$.
-
-\note{Something about existence of optimal solutions?  Probably do not exist in general,
-but do for finite $S$ despite discounting as this is a positive bounded model.}
-
+%% \note{Something about existence of optimal solutions?  Probably do not exist in general,
+%% but do for finite $S$ despite discounting as this is a positive bounded model.}
 %% Positive bounded model.  We do have upper and lower bounds on V^*.  For finite s we can get optimality...
 
-\begin{dfn}
-	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$,
-	the \term{myopic policy} $\pi^m(s)$ is defined to equal
-	\begin{equation}\label{eq:myopic}
-		\argmax_{E\in A_s} \IE_M[ -c + \max_i \mu_i(S_1) \given S_0 = s, A_0 = E]		
-	\end{equation}
-	whenever this maximum is positive, $\pi^m(s)=\bot$ otherwise.
-\end{dfn}
-
-The myopic policy takes the best action, to either stop or perform a computation,
-under the assumption that at most one further computation can be performed.  This policy
-is typically easy to compute, although not necessarily very good.  It does have
-importance, however, through its relation to the optimal policy in \thmref{thm:optimal-myopic}
-and its partial converse \thmref{thm:myopic-optimal}:
-
-%% \note{Theorem: if Optimal stops in x, myopic stops in x (converse is more useful!)}
-
-\begin{thm}\label{thm:optimal-myopic}
-	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$
-	if the optimal policy stops in a state $s\in S$
-	then the myopic policy stops too, i.e., if $\pi^*(s)=\bot$ then $\pi^m(s)=\bot$.
-\end{thm}
-
-\begin{proof}
-	If the optimal policy stops in a state $s\in S$ then
-	\[
-		V^{\pi^*}(s) \le \max_i \mu_i(s),
-	\]
-	but then the maximum of \eqref{eq:myopic} is non-positive, so the 
-	myopic policy stops by definition.
-\end{proof}
+%% \note{Define value of perfect information here for the below?}
 
@@
 \begin{align*}
-	\IE^{\pi}[m(S_{j+1}) - c; j<N\given S_0=s] \\
-	\le \IE^{\pi}[m(S_{j}); j<N\given S_0=s]
+	\IE^{\pi}[(m(S_{j+1}) - c)\, 1(j<N)\given S_0=s] \\
+	\le \IE^{\pi}[m(S_{j})\, 1(j<N)\given S_0=s]
 \end{align*}
@@
 	&= \IE^{\pi}[m(S_0) + \sum_{j=0}^{N-1} (m(S_{j+1}) - c - m(S_j)) |S_0=s] \\	
-	&\le \IE^{\pi}[m(S_0) + \sum_{j=0}^{N-1} 0 |S_0=s] \\		
-	&= \max_i\mu_i(s)
+%%	&\le \IE^{\pi}[m(S_0) + \sum_{j=0}^{N-1} 0 |S_0=s] \\		
+	&\le \max_i\mu_i(s) \qedhere
 \end{align*}
@@
 
-% Similarity to an optimal stopping result
-	% Search <blah> for one-step lookahead rules.
-	% But HERE is generalized to a stopping problem where there are many ways to continue.
-
-\note{Define value of perfect information here for the below?}
-
-\begin{thm}\label{thm:bounded-expected-computations}
-	The optimal policy $\pi^*$ has bounded expected number of computations
-	\[
-		\IE^{\pi^*}[N\given S_0=s] \le \frac{1}{c} \left(\IE[\max_i U_i\given S_0=s] - \max_i \mu_i(s)\right).
-	\]
-	Further, any policy $\pi$ with infinite expected number of computations 
-	$\IE^\pi[N]=\infty$ has negative infinite value, so in particular the optimal
-	policy stops with probability 1.
+\begin{thm}\label{thm:closed}
+	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$,
+	a subset $S'\subseteq S$ of states closed under transitions,
+	and a computation $E\in\Evidence$,
+	if $Q^m(s,E) \le Q^m(s,\bot)$ for all states $s\in S'$ where $E\in A_s$,
+	then $Q^*(s,E) \le \max{E\neq a\in A_s} Q^*(s,a)$ for all states $s\in S'$ where $E\in A_s$.
 \end{thm}
 
-\begin{proof}
-	The first follows as in state $s$ the optimal policy has value at least that
-	of stopping immediately $\max_i \mu_i(s)$, and as $\IE \max_i\mu_i(S_N) \le \IE \max_i U_i$ by Jensen's inequality.
-	The second follows from \thmref{thm:value-of-computation} as $c>0$.
-\end{proof}
-
-Although the \emph{expected} number of computations is always bounded,
-there are important cases in which the \emph{actual} number is not:
-
-\begin{example}\label{example:sprt}
-Consider the Bernoulli sampling model for two actions but with different priors:
-	$\Theta_1=1/2$,
-	and $\Theta_2$ is $1/3$ or $2/3$ with equal probability.
+The state space in definition \dfnref{metalevel-mdp} is often a bit cumbersome for analysis
+of particular examples, with many redundant.  However, we can use the following to reduce 
+the MDP to something more manageable:
+\begin{lmm}\label{lmm:equivalent-mdp}
+	Given MDPs $M^1=(S^1,s^1_0,A^1_s,T^1,R^1)$ and $M^2=(S^2,s^2_0,A^2_s,T^2,R^2)$,
+	if there exists $f\maps S^1\to S^2$ such that for all $s,s'\in S^1$ and $a\in A_s$,
+	we have $A^2_{f(s)} = A^1_s$ and
+	\begin{align*}
+		T^2(f(s),a,f(s')) &= T(s_2,E,s_2') \\
+		\max_i \mu_i(s_1) &= \max_i \mu_i(s_2),\\
 	%
-Simulating action 1 gains nothing, and after $(s,f)$ simulated successes and failures of action 2
-the posterior odds is
-\[
-	\frac{P(\Theta_2=2/3\given s,f)}{P(\Theta_2=1/3\given s,f)} = 2^{s-f}.
-\]
-The difference $s-f$ is a sufficient statistic for $\Theta$ given the computations,
-so there exists a metalevel MDP with this difference as its state.  For sufficiently
-low cost, the optimal policy samples in states $-1$, $0$, and $1$.  But with probability
-$1/3$ state $0$ transitions to state $0$ after two samples, giving finite probability
-to arbitrarily long sequences of computations.
-\end{example}
-
-However in a number of settings, including the original Bernoulli model of \exampleref{example:bernoulli},
-we can upper bound the number of computations:
-
-\begin{thm}\label{thm:bernoulli-bound}
-In the Bernoulli sampling problem of \exampleref{example:bernoulli},
-at most $k/4c$ computations are performed.	
-\end{thm}
-
-\begin{proof}
-
-\end{proof}
-
-As an important consequence, we can truncate the MDP to
-
-\note{Gaussian bound too?}
-
-%% \note{Reference ranking and selection here?}
-
-The bound in \thmref{thm:bernoulli-bound} allows us
-
-\begin{dfn}
-	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$,
-	the \term{blinkered policy} $\pi^b(s)$ \term{myopic policy} $\pi^m(s)$ is defined to equal
-	\begin{equation}\label{eq:blinkered}
-		V^b(s) = \sup_{\pi\in\Pi_b} V^\pi_M(s),
-	\end{equation}
-	where $\Pi_b$ is the set of policies which sample .....
-
-	...need comparisons between $A_s$'s to exist.
-\end{dfn}
-
-\note{Graph comparing blinkered to other policies}
-
-\begin{figure}[htb]
-\centering
-\includegraphics[scale=0.7, trim=90 70 400 300]{blinkered-regret.pdf}
-\caption{Regret vs. cost}
-\label{fig:blinkered}
-\end{figure}
-
-\begin{comment}
-set xlabel "Cost"
-set ylabel "Regret"
-
-set terminal postscript enhanced linewidth 2
-set output "blinkered-regret.ps"
-set size 0.5, 0.5
-
-set logscale xy
-
-plot [0.1:0.0005] [0.05:1] "blinkered-regret.dat" title "Blinkered" with lines lw 2, "espb-regret.dat" title "ESPb" with lines lw 2, "ucb-blinkered-regret.dat" title "UCB1-B" with lines lw 2, "ucb-blinkered-table-regret.dat" with lines lw 2 title "UCB1-b"
-\end{comment}
+	\intertext{then for any policy $\pi^2$ in $M^2$, with policy $\pi^1$ in $M^1$
+	defined by $\pi^1(s) = \pi^2(f(s))$ we have}
+	%
+		V^{\pi^1}_{M^1}(s) &= V^{\pi^2}_{M^2}(f(s)).
+	\end{align*}
+\end{lmm}

### 2012-03-30 paper/vago.uai/introduction.tex
added files, probably too many

@@
 combinatorial search problems, game playing, robotic path planning,
-model-predictive control problems, Markov decision processes (fully or
-partially observable), and a huge range of applications. In almost all
+model-predictive control problems, Markov decision processes (MDP), whether fully or
+partially observable, and a huge range of applications. In almost all
 realistic instances, exact solution is intractable and approximate
 methods are sought. Perhaps the most popular approach is to simulate
-a limited number of possible future action sequences. For example,
-a typical game-playing algorithm explores a tree or graph of action sequences
+a limited number of possible future action sequences, 
+in order to find a move in the current state that is (hopefully)
+near-optimal. One of the most successful
+examples of this approach are sampling algorithms for
+playing the game of Go \citep{Gelly.mogo} based on the upper-confidence bound (UCB)
+method \citep{Kocsis+Szepesvari:2006}. This paper examines the issue of sampling for move selection from
+first principles, by defining an appropriate meta-level MDP and formally
+analyzing some of its properties. Although
+the meta-level MDP is also intractable, approximation schemes based on
+provable value of information bounds for this problem are presented. 
+These lead to practical sampling schemes that
+are promising: they out-perform the UCB-based schemes,
+as indicated by empirical results on both ``flat'' selection and
+game-playing in Go.
+
+As selection is extremely general \citep{TolpinShimony:2012}, we focus on the adversarial
+game-playing domain as an example for clarity.
+A typical game-playing algorithm explores a tree or graph of action sequences
 with some limit on depth, using pruning methods to avoid irrelevant
 subtrees; based on what it finds in the explored portion of the state space, the algorithm
-then selects a move.
+then selects a move. This exploration is repeated after observing the move
+made by the opponent (whether or not information from previous explorations
+is reused is discussed in Section \ref{sec:control-redistribution}).
 
-Clearly, it is desirable to select the best possible move from the
+Clearly, it is desirable to select the best possible move using the
 least possible amount of exploration. For a given amount of
@@
 current tree to a terminal state. UCT is primarily a method for
-selecting a leaf from which to conduct the next simulation.
-It forms the core of the successful \textsc{MoGo} algorithm for Go playing \citep{Gelly+Silver:2011}.  It is situated 
-within the theory of bandit problems \citep{Berry+Fristedt:1985}, applying the bandit algorithm
-UCB1 \citep{Auer+et+al:2002} recursively to select actions to perform within simulations.
+selecting a leaf from which to conduct the next simulation, and
+forms the core of the successful \textsc{MoGo} algorithm for Go 
+playing \citep{Gelly+Silver:2011}.  The UCT algorithm is
+based on the the theory of bandit problems \citep{Berry+Fristedt:1985} and the near-optimal
+UCB1 bandit algorithm \citep{Auer+et+al:2002}. UCT applies
+UCB1 recursively to select actions to perform within simulations.
 
@@
 the metalevel decision problem is an instance of the {\em selection
-problem} studied in statistics~\cite{early-selection-refs}.  Despite
-recent work by \citet{Frazier+Powell:2010}, the theory of selection
+problem} studied in statistics~\citep{Bechhofer:1954,Swisher+et+al:2003}.  Despite
+recent work by \citet{Frazier+Powell:2010} and \citet{TolpinShimony:2012}, the theory of selection
 problems is less well understood than that of bandit problems.
-Accordingly, we present in \secrefs{optimal-section}{context-section}
+Accordingly, we present in \secrefs{sec:optimal}{sec:context}
 a number of results concerning optimal policies for the general case
@@
 it is possible for an optimal policy to compute forever. We also show
-by a simple counterexample that {\em index
-policies}~\cite{Gittins:1989} do not necessarily exist for selection
+by a simple counterexample that optimal {\em index
+policies}~\citep{Gittins:1989} do not necessarily exist for selection
 problems.
@@
 Motivated by this theoretical analysis, we propose in \secrefs{approx-bayesian-section}{approx-nonbayesian-section}
-two families of heuristic approximations, one for the Bayesian case and one for the distribution-free setting.
-We show empirically that these rules give better performance than UCB1 on a wide range of standard (non-sequential) selection problems.
-\secref{mcts-section} shows similar results for the case of guiding Monte Carlo tree search in the game of Go.
+two families of heuristic approximations, one for the
+Bayesian case and one for the distribution-free setting.
+We show empirically that these rules give better performance than UCB1 on 
+a wide range of standard (non-sequential) selection problems.
+\secref{mcts-section} shows similar results for the case of guiding Monte Carlo tree search
+in the game of Go.

### 2012-03-30 paper/vago.uai/optimalpolicies.tex
added files, probably too many

@@
 %% ============================
-\subsection{Selection problems}
+%% \subsection{Selection problems}
 %% ============================
 
-\note{Formal definition of selection problems and the metalevel MDP with cost per sample (time value); also mention budgeted learning.}
+%% \note{Formal definition of selection problems and the metalevel MDP with cost per sample (time value); also mention budgeted learning.}
 
 In a \term{selection problem} the decision maker is faced with the choice between
-$k$ alternative actions.  To make this choice, they may gather evidence about the
-utility of each of these alternative, at some cost.  What the decision maker can know
-and what they can find out is formalized in:
+alternative actions.  To make this choice, they may gather evidence about the
+utility of each of these alternative, at some cost.  The objective is to maximize
+the expected utility of the final action selected, less the cost of computation.
+
+%% This problem has been studied in a number of related forms.  \note{Here}
+
+%% The Bayesian ranking and selection problem \citep{Swisher+et+al:2003} is an important
+%% special case where the computations performed are samples.
+
+%%What the decision maker can know and what they can find out is formalized in:
 
@@
 	\begin{itemize}
-		\item Real random variables $U_1,\dots,U_k$, where $U_i$ is the utility of performing action~$i$, and
+		\item Real random variables $U_1,\dots,U_k$, where $U_i$ is the utility of performing action~$i$ (called ``arms" in the bandit setting), and
 		\item A countable set $\Evidence$ of random variables, each variable $E\in\Evidence$ being 
@@
 \begin{example}[Bernoulli sampling]\label{example:bernoulli}
-One simple metalevel probability model is used as a model of the results of simulations \cite{someone}.
-Each action will either succeed or not $U_i\in\{0,1\}$, with an unknown latent frequency of success $\Theta_i$, 
+%% One simple metalevel probability model is used as a model of the results of simulations \citep{someone}.
+In the \term{Bernoulli metalevel probability model},
+each action will either succeed or not $U_i\in\{0,1\}$, with an unknown latent frequency of success $\Theta_i$, 
 and stochastic simulations of possible consequences $E_{ij}$ can be performed:
@@
 \end{align*}
+
+The \term{one-action Bernoulli metalevel probability model} has $k=2$,
+$\Theta_1=m\in[0,1]$ a constant, and $\Theta_2\sample{\rm Uniform}[0,1]$.
 \end{example}
@@
 	a cost of computation $c>0$, a corresponding \term{metalevel decision problem}
-	is the MDP $M=(S,s_0,A_s,T,R)$ with
-	\begin{align*}
+	is any MDP $M=(S,s_0,A_s,T,R)$ \begin{align*}
 		S &= \{\bot\}\cup\{\langle e_1\dots, e_n \rangle : e_i\in E_i \text{ for all $i$,} \\
@@
 	\end{align*}		
-	where $\mu_i(s) = \IE[U_i \given E_1=e_1,\dots,E_n=e_n]$.
+	and where $\mu_i(s) = \IE[U_i \given E_1=e_1,\dots,E_n=e_n]$.
 \end{dfn}
+
+One can optionally add an external constraint on the number of computational actions, or their total cost,
+in the form of a deadline (called a budget henceforth).  This bridges with the related area of budgeted learning \citep{Madani+et+al:2004}.
+Although this feature is not formalized in the MDP, it can be added by including either time or past total cost 
+as part of the state.
 
@@
 is the number of simulated successes of action $i$, and $f_i=n_i-s_i$ the failures.  By \eqref{eq:bernoulli1}, 
-the state space is the set of all $k$ pairs $(s_i,f_i)$, and \eqrefs{eq:bernoulli2}{eq:bernoulli3}
+the state space is the set of all $k$ pairs $(s_i,f_i)$, \eqrefs{eq:bernoulli2}{eq:bernoulli3}
 suffice to give the transition probabilities and terminal rewards, respectively.
+The one-action Bernoulli case is similar, requiring only the two dimensional 
+space $(s,f)$ giving the posterior over $\Theta_2$.
 \end{example}
@@
 $\pi$ is a function mapping states $s\in S$ to actions to take in that state $\pi(s)\in A_s$.
+
 The \term{value function} for a policy $\pi$ gives the expected total reward
-received under that policy starting from a given state $s\in S$: 
-\begin{equation}\label{eq:value-fn}
-	V^\pi_M(s) = \IE^\pi_M\left[ \sum_{i=0}^N R(S_i,\pi(S_i),S_{i+1}) \given S_0 = s\right]	
-\end{equation}
-
+received under that policy starting from a given state $s\in S$, and the \term{Q-function}
+does the state when starting in a state $s\in S$ and taking a given action $a\in A_s$: 
+\[
+	V^\pi_M(s) = \IE^\pi_M\left[ \sum_{i=0}^N R(S_i,\pi(S_i),S_{i+1}) \given S_0 = s\right]	\label{eq:value-fn}\\
+%%	Q^\pi_M(s,a) &= \IE^\pi_M\left[ \sum_{i=0}^N R(S_i,\pi(S_i),S_{i+1}) \given S_0 = s, A_0 = a\right] \label{eq:qvalue-fn}
+\]
 where $N\in[0,\infty]$ is the random time the MDP is terminated, i.e.,
-the unique time where $\pi(S_N)=\bot$.%
+the unique time where $\pi(S_N)=\bot$%
 \footnote{Whenever $N<\infty$, $S_{N+1}=\bot$ the unique terminal state.
 Instead of having a random termination time, one can equivalently make
-the state $\bot$ absorbing with zero reward on all transitions.}
+the state $\bot$ absorbing with zero reward on all transitions.}, 
+and similarly for the Q-function $Q^\pi_M(s,a)$.
+
+As usual, an \term{optimal policy} $\pi^*$, when it exists, is one which maximises 
+the value from every state $s\in S$, i.e., if we define for each $s\in S$
+\[
+	V^*_M(s)   = \sup_\pi V^\pi_M(s),
+%%	Q^*_M(s,a) &= \sup_\pi Q^*_M(s,a),
+\]
+then the optimal policy $\pi^*$ satisfies $V^{\pi^*}_M(s) = V^*_M(s)$
+for all $s\in S$, where we break ties towards stopping $\bot$.
+%% and by standard results, $\pi^*(s) \in \argmax_{a\in A_s} Q^*_M(s,a)$, where
+
+The optimal policy must balance the cost of computations with the improved decision
+quality that results.  This is made clear by giving by expressing the value function:
 
@@
 	\]
-	where the random variable $N$ denotes the total number of computations performed.
+	where the random variable $N$ denotes the total number of computations performed,
+	and similarly for $Q^\pi_M(s,a)$.
 \end{thm}
@@
 	Follows immediately from \eqref{eq:value-fn} and the definition of the
-	reward function $R$ in \dfnref{dfn:metalevel-mdp}.
+	reward function in \dfnref{dfn:metalevel-mdp}.
 \end{proof}
 
-\begin{thm}
-	MDPs $M=(S,s_0,A_s,T,R)$ where there exists a unique terminal state $\bot\in S$,
-	a distinguished action $\bot\in A_s$ for all $s\in S$, a constant $c>0$, and bounded functions 
-	$\mu_1,\dots,\mu_k\maps S\to\R$ such that for all $s\in S$, $a\in A_s\setminus\{\bot\}$, and $i=1,\dots,k$:
-	\begin{align*}
-		\mu_i(s) &= \sum_{s'\in S} T(s,a,s')\,\mu_i(s') \\
-		T(s,\bot,\bot) &= 1 \\
-		R(s,a,s') &= -c \\
-		R(s,\bot,\bot) &= \max_i \mu_i(s)
-	\end{align*}
-	are equivalent to metalevel decision problems with bounded utilities, and conversely any 
-	metalevel decision problem with bounded utilities $U_i$ is such an MDP.
-\end{thm}
+In many problems, including the Bernoulli sampling model of \exampleref{example:bernoulli2},
+the state space is infinite.  Does this preclude solving for the optimal policy?  Can 
+infinitely many computations be performed?
 
-\begin{proof}
-Given such an MDP, for simplicity will assume that $\mu_i(s)\in(0,1)$ for all $i$ and $s\in S$; the proof generalizes 
-to any bounds $(\mu^-,\mu^+)$.  We assume further the MDP is acyclic, which can be ensured, e.g., by augmenting
-the state space with a time variable incrementing on each transition.
-
-Define $\{0,1\}$-valued random variables $U_1,\dots,U_k$ 
-and $E_{sa}$ for all $s\in S$, $a\in A_s$ by
-\begin{align*}
-	P(U_i=1) &= \mu_i(s_0) \\
-	P(E_{sa}=s'\given U_i=1) &= \frac{T(s,a,s')\,\mu_i(s')}{\mu_i(s)} \\
-	P(E_{sa}=s'\given U_i=0) &= \frac{T(s,a,s')\,(1-\mu_i(s'))}{1- \mu_i(s)}
-\end{align*}
-where $U_1,\dots,U_k$ are independent, and the $E_{sa}$ are conditionally 
-independent given $(U_1,\dots,U_k)$.
-These together form a metalevel probability model, so we can form the 
-corresponding metalevel decision problem following definition~\ref{dfn:metalevel-mdp},
-to avoid conflict of notation we denote the latter by $M'=(H,h_0,B_h,T',R')$,
-and where since $B_h$ is underspecified in the definition we define
-\[
-	B_h = \{\bot\} \cup \{E_{t(h)a} : a \in A_{t(h)a}\}
-\]
-where the tail function $t\maps H\to S$ has $t(\langle\rangle)=s_0$ and otherwise
-\[
-	t(\langle e_{s_0a_0},\dots,e_{s_n a_n}\rangle) = e_{s_n a_n} \in S.
-\]
-This definition of $B_h$ means that all elements of $H$ reachable from $h_0=\langle\rangle$ with positive probability are 
-sequences $h=\langle e_{s_0a_0},\dots,e_{s_n a_n}\rangle$
-with $e_{s_ia_i} = s_{i+1} \in S$ and $a_i\in A_{s_i}$, i.e., are possible state sequences 
-in the original MDP $M$ under some sequence of actions.  Fix such a sequence, and
-note that given $E_{s_n a}\in B_h$, any $h'$ with positive transition probability
-is the form $h'=\langle e_{s_0a_0},\dots,e_{s_n a_n},e_{s_{n+1}a}\rangle$, and
-\[
-	T'(h,E_{s_n a},h') = T(t(h), a, t(h')).
-\]
-
-Finally, observe that by acyclicity and construction
-\begin{align*}
-	& P(U_i=1, E_{s_0a_0}=s_1,\dots,E_{s_n a_n}=s_{n+1}) \\
-	&= \mu_i(s_0) \prod_{i=0}^n \frac{T(s_i,a_i,s_{i+1})\,\mu_i(s_{i+1})}{\mu_i(s_i)} \\
-	&= \mu_i(s_{n+1}) \prod_{i=0}^n T(s_i,a_i,s_{i+1}),
-%
-\intertext{and similarly}
-%
-	& P(U_i=0, E_{s_0a_0}=s_1,\dots,E_{s_n a_n}=s_{n+1}) \\
-	&= (1-\mu_i(s_{n+1})) \prod_{i=0}^n T(s_i,a_i,s_{i+1}).
-\end{align*}
-As a result
-\begin{align*}
-	& \IE(U_i\given E_{s_0a_0}=s_1,\dots,E_{s_n a_n}=s_{n+1})\\
-	& = \mu_i(s_n) = \mu_i(t(h))\\
-	& P(E_{s_n a_n}=s_{n+1} \given E_{s_0a_0}=s_1,\dots,E_{s_{n-1} a_{n-1}}=s_{n}) \\
-	&=  T(s_n,a,s_{n+1}) = T(t(h),a,t(h'))
-\end{align*}
-
-Thus the metalevel decision problem $M'$ is equivalent to the original MDP $M$
-under the mapping $t\maps H\to S$.
-
-For the converse, observe that the properties hold with $c$ the cost of computation
-and \[\mu_i(s) = \IE(U_i\given E_1=e_1,\dots,E_n=e_n)\] where $s=\langle e_1,\dots,e_n \rangle$.
-\end{proof}
-
-As usual, an \term{optimal policy} $\pi^*$ is one which maximises 
-the value from every state $s\in S$, i.e., if we define for each $s\in S$
-\[
-	V^*_M(s) = \sup_\pi V^\pi_M(s),
-\]
-then an optimal policy $\pi^*$ satisfies $V^{\pi^*}_M(s) = V^*_M(s)$
-for all $s\in S$.
-
-\note{Something about existence of optimal solutions?  Probably do not exist in general,
-but do for finite $S$ despite discounting as this is a positive bounded model.}
-
-%% Positive bounded model.  We do have upper and lower bounds on V^*.  For finite s we can get optimality...
-
-\begin{dfn}
-	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$,
-	the \term{myopic policy} $\pi^m(s)$ is defined to equal
-	\begin{equation}\label{eq:myopic}
-		\argmax_{E\in A_s} \IE_M[ -c + \max_i \mu_i(S_1) \given S_0 = s, A_0 = E]		
-	\end{equation}
-	whenever this maximum is positive, $\pi^m(s)=\bot$ otherwise.
-\end{dfn}
-
-The myopic policy takes the best action, to either stop or perform a computation,
-under the assumption that at most one further computation can be performed.  This policy
-is typically easy to compute, although not necessarily very good.  It does have
-importance, however, through its relation to the optimal policy in \thmref{thm:optimal-myopic}
-and its partial converse \thmref{thm:myopic-optimal}:
-
-%% \note{Theorem: if Optimal stops in x, myopic stops in x (converse is more useful!)}
-
-\begin{thm}\label{thm:optimal-myopic}
-	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$
-	if the optimal policy stops in a state $s\in S$
-	then the myopic policy stops too, i.e., if $\pi^*(s)=\bot$ then $\pi^m(s)=\bot$.
-\end{thm}
-
-\begin{proof}
-	If the optimal policy stops in a state $s\in S$ then
-	\[
-		V^{\pi^*}(s) \le \max_i \mu_i(s),
-	\]
-	but then the maximum of \eqref{eq:myopic} is non-positive, so the 
-	myopic policy stops by definition.
-\end{proof}
-
-\begin{dfn}
-	Given an MDP $M=(S,s_0,A_s,T,R)$ a state $s'$ is \term{reachable} from a state $s$
-	if there is a sequence of states $s=s_0,s_1,\dots,s_n=s'$ and actions $a_i\in A_{s_i}$
-	with positive probability under the transition distribution $T$.
-\end{dfn}
-
-%% \note{Theorem: if Myopic stops in all states reachable from x, optimal stops in x}
-
-\begin{thm}\label{thm:myopic-optimal}
-	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$
-	if the myopic policy stops in all states $s'\in S$ reachable
-	from a given state $s$ then the optimal policy stops too.
-	then the myopic policy stops too, i.e., if $\pi^*(s)=\bot$ then $\pi^m(s)=\bot$.
-\end{thm}
-
-\begin{proof}
-Defining $m(s) = \max_i\mu_i(s)$, observe the myopic stopping for all
-reachable states implies that
-\begin{align*}
-	\IE^{\pi}[m(S_{j+1}) - c; j<N\given S_0=s] \\
-	\le \IE^{\pi}[m(S_{j}); j<N\given S_0=s]
-\end{align*}
-holds for all $j$, and as a result:
-\begin{align*}
-	V^\pi(s) 
-	&= \IE^{\pi}[ - c N + m(S_N) |S_0=s] \\
-	&= \IE^{\pi}[m(S_0) + \sum_{j=0}^{N-1} (m(S_{j+1}) - c - m(S_j)) |S_0=s] \\	
-	&\le \IE^{\pi}[m(S_0) + \sum_{j=0}^{N-1} 0 |S_0=s] \\		
-	&= \max_i\mu_i(s)
-\end{align*}
-\end{proof}
-
-% Similarity to an optimal stopping result
-	% Search <blah> for one-step lookahead rules.
-	% But HERE is generalized to a stopping problem where there are many ways to continue.
-
-\note{Define value of perfect information here for the below?}
+There is in full generality an upper bound on the \emph{expected} number of computations
+a policy performs:
 
 \begin{thm}\label{thm:bounded-expected-computations}
-	The optimal policy $\pi^*$ has bounded expected number of computations
+	The optimal policy $\pi^*$'s expected number of computations is bounded by the 
+	perfect value of information \citep{Howard:1966} times the inverse cost $1/c$:
 	\[
@@
 
-Although the \emph{expected} number of computations is always bounded,
-there are important cases in which the \emph{actual} number is not:
+However, although the \emph{expected} number of computations is always bounded,
+there are important cases in which the \emph{actual} number is not, such as
+the following example inspired by the sequential probability ratio test \citep{Wald+1945}:
 
@@
 \[
-	\frac{P(\Theta_2=2/3\given s,f)}{P(\Theta_2=1/3\given s,f)} = 2^{s-f}.
+	\frac{P(\Theta_2=2/3\given s,f)}{P(\Theta_2=1/3\given s,f)} = \frac{(2/3)^s(1/3)^f}{(1/3)^s(2/3)^f}= 2^{s-f}.
 \]
-The difference $s-f$ is a sufficient statistic for $\Theta$ given the computations,
-so there exists a metalevel MDP with this difference as its state.  For sufficiently
-low cost, the optimal policy samples in states $-1$, $0$, and $1$.  But with probability
-$1/3$ state $0$ transitions to state $0$ after two samples, giving finite probability
-to arbitrarily long sequences of computations.
+Note that these odds completely specify the posterior distribution of $\Theta_2$,
+and thus the distribution of the utility and all future computations.  Thus, whether
+it is optimal to continue is a function only of this posterior odds, and thus $s-f$.
+%% In particular,
+%% note that after equal numbers of successes and failures, $s-f=0$ and so the posterior 
+%% distribution over $\Theta_2$ is equal to the prior.
+For sufficiently low cost, the 
+optimal policy samples when $s-f$ equals $-1$, $0$, and $1$.  But with probability
+$1/3$, a state with $s-f=0$ transitions to another state $s-f=0$ after two samples, 
+giving finite, although exponentially decreasing, probability to arbitrarily long 
+sequences of computations.
 \end{example}
 
-However in a number of settings, including the original Bernoulli model of \exampleref{example:bernoulli},
-we can upper bound the number of computations:
+Fortunately, in a number of settings, including the original Bernoulli model of \exampleref{example:bernoulli},
+we can prove an upper bound the number of computations.  For reasons of space,
+and for its later use in \secref{sec:blinkered}, we prove here the bound for the one-action Bernoulli model.
 
-\begin{thm}\label{thm:bernoulli-bound}
-In the Bernoulli sampling problem of \exampleref{example:bernoulli},
-at most $k/4c$ computations are performed.	
+Before we can do this, we need to get an analytical handle on the optimal policy.
+The key is through a natural approximate policy:
+
+\begin{dfn}\label{dfn:myopic}
+	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$,
+	the \term{myopic policy} $\pi^m(s)$ is defined to equal $\argmax_{a\in A_s} Q^m(s,a)$ 
+	where $Q^m(s,\bot) = \max_i \mu_i(s)$ and
+	\begin{equation}\label{eq:myopic}
+		 Q^m(s,E) = \IE_M[ -c + \max_i \mu_i(S_1) \given S_0 = s, A_0 = E].		
+	\end{equation}
+\end{dfn}
+
+The myopic policy takes the best action, to either stop or perform a computation,
+under the assumption that at most one further computation can be performed.
+%% This policy is typically easy to compute, although not necessarily very good.
+
+It does have importance, however, in how it constrains the optimal policy in \thmref{thm:optimal-myopic}
+and its partial converse \thmref{thm:myopic-optimal}:
+
+%% \note{Theorem: if Optimal stops in x, myopic stops in x (converse is more useful!)}
+
+\begin{thm}\label{thm:optimal-myopic}
+	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$
+	if the myopic policy performs some computation in state $s\in S$,
+	then the optimal policy does too, i.e., if $\pi^m(s)\neq\bot$ then $\pi^*(s)\neq\bot$.
 \end{thm}
 
+\begin{comment}
 	\begin{proof}
+		Observe that the myopic Q-function \eqref{dfn:myopic} is equivalently given by
+		\[
+			Q^m(s,a) = Q^\bot(s,a)
+		\]
+		where $\bot$ is the policy which immediately stops $\bot(s)=\bot$.
+		Thus $Q^m(s,a) \le Q^*(s,a)$.  If the optimal policy stops in a state $s\in S$ then
+		\[
+			Q^{\pi^*}(s,a) \le \max_i \mu_i(s),
+		\]
+		and so the same holds for $Q^m$, showing the myopic stops.
+	\end{proof}
+\end{comment}
 
-\end{proof}
+% Similarity to an optimal stopping result
+	% Search <blah> for one-step lookahead rules.
+	% But HERE is generalized to a stopping problem where there are many ways to continue.
 
-As an important consequence, we can truncate the MDP to
+This shows that the myopic policy is conservative in stopping, so stopping too early
+is one of its approximation errors.  However myopic stopping can be used to control optimal stopping:
 
-\note{Gaussian bound too?}
-
-%% \note{Reference ranking and selection here?}
-
-The bound in \thmref{thm:bernoulli-bound} allows us
-
-\begin{dfn}
+\begin{dfn}\label{dfn:closed}
 	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$,
-	the \term{blinkered policy} $\pi^b(s)$ \term{myopic policy} $\pi^m(s)$ is defined to equal
-	\begin{equation}\label{eq:blinkered}
-		V^b(s) = \sup_{\pi\in\Pi_b} V^\pi_M(s),
-	\end{equation}
-	where $\Pi_b$ is the set of policies which sample .....
-
-	...need comparisons between $A_s$'s to exist.
+	a subset $S'\subseteq S$ of states is \term{closed under transitions}
+	if whenever $s'\in S'$, $a\in A_{s'}$, $s''\in S$, and $T(s',a,s'')>0$,
+	we have $s''\in S'$.
 \end{dfn}
 
-\note{Graph comparing blinkered to other policies}
-
-\begin{figure}[htb]
-\centering
-\includegraphics[scale=0.7, trim=90 70 400 300]{blinkered-regret.pdf}
-\caption{Regret vs. cost}
-\label{fig:blinkered}
-\end{figure}
+\begin{thm}\label{thm:myopic-optimal}
+	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$
+	and a subset $S'\subseteq S$ of states closed under transitions,	
+	if the myopic policy stops in all states $s\in S$
+	then the optimal policy does too.	
+\end{thm}
 
 \begin{comment}
-set xlabel "Cost"
-set ylabel "Regret"
+	\begin{proof}
+	Take any $s^*\in S'$, and note that all states the chain can transition
+	to from $s^*$ are also in $S'$, by transition closure.  Defining $m(s) = \max_i\mu_i(s)$, 
+	observe the myopic stopping for all such states implies that
+	\begin{align*}
+		\IE^{\pi}[(m(S_{j+1}) - c)\, 1(j<N)\given S_0=s^*] \\
+		\le \IE^{\pi}[m(S_{j})\, 1(j<N)\given S_0=s^*]
+	\end{align*}
+	holds for all $j$, and as a result:
+	\begin{align*}
+		V^\pi(s) 
+		&= \IE^{\pi}[ - c N + m(S_N) \given S_0=s^*] \\
+		&= \IE^{\pi}[m(S_0) + \sum_{j=0}^{N-1} (m(S_{j+1}) - c - m(S_j)) \given S_0=s^*] \\
+	%%	&\le \IE^{\pi}[m(S_0) + \sum_{j=0}^{N-1} 0 |S_0=s] \\		
+		&\le \max_i\mu_i(s^*) \qedhere
+	\end{align*}
+	\end{proof}	
+\end{comment}
 
-set terminal postscript enhanced linewidth 2
-set output "blinkered-regret.ps"
-set size 0.5, 0.5
+Using this, we can prove our bound:
 
-set logscale xy
+\begin{thm}\label{thm:one-action-bound}
+	The one-action Bernoulli decision process with constant arm $m\in[0,1]$ 
+	performs at most $m(1-m)/c-3 \le 1/4c-3$ computations.
+\end{thm}
 
-plot [0.1:0.0005] [0.05:1] "blinkered-regret.dat" title "Blinkered" with lines lw 2, "espb-regret.dat" title "ESPb" with lines lw 2, "ucb-blinkered-regret.dat" title "UCB1-B" with lines lw 2, "ucb-blinkered-table-regret.dat" with lines lw 2 title "UCB1-b"
+\begin{comment}
+	\begin{proof}
+		By \dfnref{dfn:myopic} and \exampleref{example:bernoulli2}, the myopic policy stops in
+		a state $(s,f)$ when
+		\begin{align}
+			c \ge &\mu\max(\mu^+,m) + (1-\mu_i)\max(\mu^-, m) - \max(\mu,m) \label{eq:stopping}
+		\end{align}
+		where $\mu=(s+1)/(n+2)$  is the mean utility for action~2, where $n=s+f$,
+		$\mu^- = \mu - \mu/(n+3)$, and $\mu^+ = \mu + (1-\mu)/(n+3)$
+		are the posterior means of action 2 after simulating a failure and a success,
+		%% and $\mu^-_i=(s_i+1)/(s_i+f_i+3) $ and $\mu^+_i = (s_i+2)/(s_i+f_i+3)$
+		%% are the posterior means of action $i$ after simulating a failure and a success,
+		respectively.  Whenever \eqref{eq:stopping} holds, stopping is preferred to sampling.
+
+		Fixing $n$ and maximizing over $\mu$, we get sufficient condition for stopping
+		\begin{equation}
+			c \ge \frac{m(1-m)}{(n_i+3)} \qquad\qquad   n_i\ge \frac{m(1-m)}{c} - 3  \label{eq:myopic-stopping}
+		\end{equation}
+		Since the set of states satisfying \eqref{eq:myopic-stopping} is closed under
+		transitions ($n$ only increases), by \thmref{thm:optimal-myopic}.  Finally, note $\max_{m\in[0,1]} m(1-m)=1/4$.
+	\end{proof}	
 \end{comment}
+
+A key implication is that the \emph{optimal} policy can be computed
+in time quadratic in the inverse cost $O(1/c^2)$.  This is particularly appropriate when the cost of 
+computation is relatively high, such as in simulation experiments \citep{Swisher+et+al:2003},
+or when the decision to be made is critical.

### 2012-03-30 paper/vago.uai/paper.tex
added files, probably too many

@@
 \usepackage{color}
+\usepackage[round]{natbib}
+\usepackage{comment}
+
+\def\citeapos#1{\citeauthor{#1}'s (\citeyear{#1})}
 
@@
 
-\newtheorem{dfn}{Definition}
+\newcommand {\term}[1] {\textbf{#1}}
+\newcommand {\note}[1] {\textcolor{red}{[[#1]]}}
+\newcommand {\Evidence} {\mathcal{E}}
+\newcommand {\R} {\mathbb{R}}
+\newcommand {\maps} {\colon}
+\newcommand {\given} {\mathrel{|}}
+\newcommand {\sample} {\sim}
+\newcommand {\iid} {\stackrel{\rm iid}{\sample}}
+\DeclareMathOperator*{\argmax}{argmax}
+
 \newtheorem{thm}{Theorem}
-\newtheorem{lmm}{Lemma}
-\newtheorem{crl}{Corollary}
+\newtheorem{dfn}[thm]{Definition}
+\newtheorem{lmm}[thm]{Lemma}
+\newtheorem{crl}[thm]{Corollary}
+\newtheorem{example}{Example}
+
+\newcommand{\secref}[1]{Section~\ref{#1}}
+\newcommand{\secrefs}[2]{Sections~\ref{#1} and~\ref{#2}}
+\newcommand{\figref}[1]{Figure~\ref{#1}}
+\renewcommand{\eqref}[1]{Equation~(\ref{#1})}
+\newcommand{\eqrefs}[2]{Equations~(\ref{#1}) and~(\ref{#2})}
+\newcommand{\thmref}[1]{Theorem~\ref{#1}}
+\newcommand{\dfnref}[1]{Definition~\ref{#1}}
+\newcommand{\exampleref}[1]{Example~\ref{#1}}
+
+\title{Selecting Computations: Theory and Applications}
+
+\author{}
+%\author{ {\bf Nicholas Hay} and {\bf Stuart Russell} \\   
+%Computer Science Division \\  
+%University of California\\ 
+%Berkeley, CA 94720 \\ 
+%\And 
+%{\bf Solomon Eyal Shimony} and {\bf David Tolpin}  \\ 
+%Department of Computer Science \\ 
+%Ben-Gurion University of the Negev\\
+%Beer Sheva, Israel
+%}
 
@@
 
-\section{Upper Bounds on Value of Information}
+\maketitle
 
-The intrinsic (myopic greedy) VOI $\Lambda_i$ of testing an arm $i$ is
+\begin{abstract}
+Sequential decision problems are often approximately solvable by
+simulating possible future action sequences.  {\em Metalevel} decision
+procedures have been developed for selecting {\em which} action
+sequences to simulate, based on some estimate of the expected
+improvement in decision quality resulting from any particular
+simulation; an example is the recent work on using selection rules
+derived for bandit problems to control Monte Carlo tree search in the
+game of Go.  In this paper we develop a theoretical basis for
+metalevel decisions in the statistical framework of {\em selection problems}, 
+arguing that this is more appropriate than the bandit
+framework.  We derive a number of basic results applicable to simple
+Monte Carlo selection problems, including the first finite sampling
+bounds for optimal policies in certain cases; we also provide a simple
+counterexample to the intuitive conjecture that an optimal policy will
+necessarily reach a decision in all cases.  We then derive heuristic
+approximations in both Bayesian and distribution-free settings and
+demonstrate their superiority to bandit-based heuristics in one-shot
+decision problems and in Go.
+\end{abstract}
+
+\section{Introduction}
+
+\input{introduction}
+
+%% \input{related}
+
+\section{On optimal policies for selection}\label{sec:optimal}
+
+\input{optimalpolicies}
+
+\section{Context effects and non-indexability}\label{sec:context}
+
+\input{context}
+
+\section{The blinkered policy}\label{sec:blinkered}\label{approx-bayesian-section}
+
+\input{blinkered}
+
+%% \section{Regret bounds and approximate policies}
+
+%% [[Regret models: simple regret, regret with cost per sampling; regret goes to zero as c does]]
+
+%% [[Expected simple regret bounds for normal case?]]
+
+%% [[Blinkered sampling]]
+
+%% [[ESPb: Frazier's continuous time approximation]]
+
+\section{Upper bounds on value of information}\label{approx-nonbayesian-section}
+
+%% Solomon: Text presumably should be placed after the relevant theorems.
+
+Theorem \ref{thm:optimal-myopic} provides necessary conditions
+for a stopping condition for the optimal policy in terms of a myopic
+policy, but that does not preclude premature stopping of a greedy policy.
+Conversely, Theorem \ref{thm:myopic-optimal} provides sufficient conditions, but these 
+conditions are hard to evaluate in practice.
+
+The myopic selection policy uses
+the intrinsic value of information (VOI) $\Lambda_i$ of testing an arm $i$, which is
 the expected reward of the best arm given the information obtained by the test,
 minus the expected reward of the best arm according to the current information state.
-We will assume here a distribution model where the arm with the highest sample mean
-in an information state is the one with the highest expected reward.
-Two cases are possible:
+Unfortunately, the pure \textit{myopic} VOI estimate is of little use in
+Monte-Carlo sampling, since the effect of a single sample is small,
+and the myopic VOI estimate will often be zero.
+
+Instead, we generalize myopic policies to include ``semi-myopic"
+up to $N$ samples of the same arm; these are called ``blinkered" policies.
+Such policies allow potentially unlimited lookahead, but only in a single ``direction'' (one specific arm),
+as if we ``had our blinkers on''.
+
+%% Solomon: end text to be (possibly moved).
+
+In many practical applications of the selection problem, such as search in
+the game of Go, prior distributions are unavailable\footnote{The analysis is also applicable to
+some Bayesian settings, using ``fake" samples to simulate prior distributions.}
+In such cases, one can still bound
+the value of information w.r.t. myopic policies using {\em concentration
+inequalities} to derive distribution-independent bounds on the
+VOI. We obtain such bounds under the
+following assumptions:
+\begin{enumerate}
+\item Samples are iid given the value of the arms (variables), as in the Bayesian schemes such as Bernoulli
+sampling.
+\item The expectation of a selection in a belief state is equal to the sample mean (and therefore,
+   after sampling terminates, the arm with the greatest sample mean will be selected).
+\end{enumerate}
+
+When considering possible samples in the ``blinkered'' semi-myopic setting,
+two cases are possible:
 \begin{itemize}
@@
 
-In order to estimate the VOI of testing an arm, a sampling model is
-needed.  The generic sampling model used is: 
-\begin{itemize}
-\item The actual value of an arm has a certain (possibly unknown) distribution.
-\item Testing an arm is an i.i.d. noisy measurement of the actual
-  value of the arm.
-\end{itemize}
-When the distribution model is known, one can use the methods of Section \ref{sec:normal}.
-Otherwise, one can still use {\em concentration
-inequalities} to derive distribution-independent bounds on the
-VOI. Our bounds are applicable to any bounded distribution. Similar
+Our bounds below are applicable to any bounded distribution (with rewards being
+bounded to $[0,1]$). Similar
 bounds can be derived for certain unbounded distributions, such as the
@@
 sampling.
-
-Unfortunately, the pure \textit{myopic} VOI estimate is of little use in
-Monte-Carlo sampling, since the effect of a single sample is small,
-and the myopic VOI estimate will often be zero. Instead, we will derive
-an estimate for the case where we decide to test an arm $N$ times,
+We derive a VOI bound for testing an arm $N$ times,
 where $N$ can be the remaining ``budget'' of available samples or
@@
   therefore $\overline X_i^{n_i+N}\le \overline
-  X_\alpha^{n_\alpha}+\frac {N(1-\overline X_\alpha^{n_\alpha})} {N+n_i}$. Hence, the intrinsic value of perfect
+  X_\alpha^{n_\alpha}+\frac {N(1-\overline X_\alpha^{n_\alpha})} {N+n_i}$. Hence, the intrinsic value of blinkered
   information is at most: 
-\begin{equation}
+\begin{align}
 \label{eq:simplistic}
 \frac{ N(1-\overline  X_\alpha^{n_\alpha})}
-  {N+n_i}\Pr(\overline X_i^{{n_i}+N}\hspace{-0.5em}\ge\overline X_\alpha^{n_\alpha})\le \frac {N \overline X_\beta^{n_\beta}} {n_\alpha} \Pr(\overline X_\alpha^{n_\alpha+N}\hspace{-0.5em}\le\overline X_\beta^{n_\beta})
-\end{equation}
+  {N+n_i}&\Pr(\overline X_i^{{n_i}+N}\hspace{-0.5em}\ge\overline X_\alpha^{n_\alpha})\nonumber \\
+&\le\frac{ N(1-\overline  X_\alpha^{n_\alpha})}
+{n_i}\Pr(\overline X_i^{{n_i}+N}\hspace{-0.5em}\ge\overline X_\alpha^{n_\alpha})
+\end{align}
   Proof for the case $i=\alpha$ is similar.
@@
 The probabilities can be bounded from above using the
-Hoeffding inequality \cite{Hoeffding.ineq}:
+Hoeffding inequality \citep{Hoeffding.ineq}:
 \begin{thm} The probabilities in equations (\ref{eqn:thm-be}) are bounded from above as
@@
 \end{thm}
+
+\begin{comment}
 	\begin{proof}
@@
 probabilities in Equation \ref{eq:simplistic}, for example, based on the empirical Bernstein
-inequality~\cite{MaurerPontil.benrstein}, or through a more careful
+inequality~\citep{MaurerPontil.benrstein}, or through a more careful
 application of the Hoeffding inequality, resulting in:
@@
 
-\section{VOI-based Sampling Control}
+\section{Practical sampling control: empirical evaluation}
+\label{sec:empirical-evaluation}\label{mcts-section}
 
-\subsection{Selection Criterion}
+As the selection MDP is intractable, we use myopic or semi-myopic policies instead.
+The myopic policies usually separate out the decision of {\em whether} to
+sample or to stop (called the stopping policy), and {\em what} to sample.
+We first examine the latter issue.
 
-A reasonable greedy policy is to test the arm that has the best
-current VOI estimate. When the distributions are unknown, it makes sense
-to use the upper bounds established in Theorem~\ref{thm:be} instead. Although MCTS
-differs from pure exploration in Multi-armed Bandits in that the
-distributions of the rewards are not stationary, VOI
-estimates computed as for stationary distributions work well in
+\subsection{Selecting The Best Arm}
+\label{sec:emp-arm}
+
+Assuming that the sample costs are constant,
+a semi-myopic policy will decide to test the arm that has the best
+current VOI estimate. 
+When the distributions are unknown, it makes sense
+to use the upper bounds established in Theorem~\ref{thm:be} instead,
+which is the method adopted here.
+This evaluation assumes a fixed budget of samples, which is
+completely used up by each of the candidate schemes, making a stopping
+criterion irrelevant.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[scale=0.6]{flat.pdf}
+\caption{Random instances: regret vs. number of samples}
+\label{fig:random-instances}
+\end{figure}
+
+The sampling policies are compared on random Multi-armed bandit 
+selection problem instances. Figure~\ref{fig:random-instances} shows results for
+randomly-generated Multi-armed bandits with 25 Bernoulli arms, with
+the mean rewards of the arms distributed uniformly in the range~$[0,
+  1]$, for a range of sample budgets~$200..2000$, with multiplicative
+step of~$2$. The experiment for each number of samples was repeated
+10000 times. UCB1, as well as policies  based on bounds
+(\ref{eqn:bound-blnk-hoeffding}) (VOI) and
+(\ref{eqn:erf-blinkered}) (VOI+) are compared. UCB1 is always considerably worse than the
+VOI-aware sampling policies.
+
+\subsection{Sampling in trees}
+
+Sampling in trees is considerably more complicated.
+The goal of Monte-Carlo tree search \citep{Chaslot.montecarlo} at the root node 
+is usually to select an action that appears to be the best based on outcomes
+of \textit{search rollouts}.
+But the goal of rollouts at non-root nodes
+is different than at the root: here it is important to better approximate the ``true'' 
+value of the node, so that selection at the root can be more informed. The exact analysis
+of sampling at internal nodes is outside the scope of this paper. At present we 
+have no better proposal for internal nodes than to use UCT there.
+
+We thus propose the following hybrid sampling scheme:
+\begin{itemize}
+\item at the root node, sample based on the VOI estimate;
+\item at non-root nodes, sample using UCT.
+\end{itemize}
+
+Strictly speaking, even at the root node the assumptions underlying our belief-state
+MDP for selection do not hold exactly. UCT is an adaptive scheme, and therefore the values
+generated by sampling at non-root nodes will typically cause valued observed at
+children of the root node to be non-stationary. 
+Nevertheless, sampling based on VOI estimates
+computed as for stationary distributions work well in
 practice. As illustrated
@@
 
-\subsection{Stopping Criterion}
+\subsubsection{Stopping criterion}
 \label{sec:control-stopping-criterion}
 
-The simplest stopping criterion for a sampling policy is the
-budget---a fixed number of samples performed before a decision is
-made. When a sample has a known cost commensurable with the value of
+When a sample has a known cost commensurable with the value of
 information of a measurement, an upper bound on the intrinsic VOI can also
@@
 influence of the sample cost $c$ on the performance of
-the sampling policy.
+the sampling policy. When the sample cost $c$ is unknown, one can perform initial calibration experiments
+to determine a reasonable value, as done in this paper.
 
-\subsection{Sample Redistribution in Trees}
+\subsubsection{Sample redistribution in trees}
 \label{sec:control-redistribution}
-
-Monte-Carlo tree search \cite{Chaslot.montecarlo} selects at each
-search step an action that appears to be the best based on outcomes
-of \textit{search rollouts}. Two different criteria are employed in
-action selection during a rollout:
-\begin{itemize}
-\item at the first step, the value of partial
-information should be used to select nodes to sample, as the goal is
-selection of a move at the current root node;
-\item as the goal of sampling in deeper tree nodes is usually to improve the
-node value estimate, rather than selection, a different sampling scheme
-should be used there. At these levels, the standard practice
-of using UCT-like rollouts makes sense, as this method is known to be nearly optimal
-in minimizing the \textit{cumulative} regret.
-\end{itemize}
 
@@
 that the information obtained from rollouts in the
-current state is discarded after an action is selected. In practice,
+current state is discarded after an real-world action is selected. In practice,
 many successful Monte-Carlo tree search algorithms re-use rollouts
@@
 
-\section{Empirical Evaluation}
-\label{sec:empirical-evaluation}
-
-\subsection{Selecting The Best Arm}
-\label{sec:emp-arm}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=0.6]{flat.pdf}
-\caption{Random instances: regret vs. number of samples}
-\label{fig:random-instances}
-\end{figure}
-
-The sampling policies are first compared on random Multi-armed bandit 
-problem instances. Figure~\ref{fig:random-instances} shows results for
-randomly-generated Multi-armed bandits with 25 Bernoulli arms, with
-the mean rewards of the arms distributed uniformly in the range~$[0,
-  1]$, for a range of sample budgets~$200..2000$, with multiplicative
-step of~$2$. The experiment for each number of samples was repeated
-10000 times. UCB1, as well as policies  based on bounds
-(\ref{eqn:bound-blnk-hoeffding}) (VOI) and
-(\ref{eqn:erf-blinkered}) (VOI+) are compared. UCB1 is always considerably worse than the
-VOI-aware sampling policies.
-
-\subsection{Playing Go Against UCT}
+\subsubsection{Playing Go against UCT}
 \label{sec:emp-go}
 
-The policies were also compared on Computer Go, a  search domain
+The hybrid policies were compared on Computer Go, a search domain
 in which UCT-based MCTS has been particularly successful
-\cite{Gelly.mogo}. A modified version of Pachi \cite{Braudis.pachi}, a state of the art
+\citep{Gelly.mogo}. A modified version of Pachi \citep{Braudis.pachi}, a state of the art
 Go program, was used for the experiments:
@@
 15000 samples (game simulations) per ply, each experiment was repeated
-1000 times. Figure~\ref{fig:uctvoi}
-show the winning rate of the VOI-aware policy against UCT
+1000 times. Figure~\ref{fig:uctvoi} depicts a calibration experiment,
+showing the winning rate of the VOI-aware policy against UCT
 vs. the stopping threshold $c$ (if the maximum VOI of a sample is below
@@
 
-\bibliographystyle{plain}
+Note that if we disallowed re-use of samples in both Pachi and
+in our VOI-based scheme, the VOI based-scheme
+win rate is even higher than shown in Figure \ref{fig:voi-wins}. This is as expected,
+as this setting (which is somewhat unfair to Pachi) is closer to
+meeting the assumptions underlying the selection MDP.
+
+\bibliographystyle{abbrvnat}
 \bibliography{refs}

### 2012-04-03 paper/vago.uai/abstract.tex
as submitted

\begin{abstract}
Sequential decision problems are often approximately solvable by
simulating possible future action sequences.  {\em Metalevel} decision
procedures have been developed for selecting {\em which} action
sequences to simulate, based on estimating the expected
improvement in decision quality that would result from any particular
simulation; an example is the recent work on using bandit algorithms
to control Monte Carlo tree search in the
game of Go.  In this paper we develop a theoretical basis for
metalevel decisions in the statistical framework of {\em selection problems}, 
arguing (as others have done) that this is more appropriate than the bandit
framework.  We derive a number of basic results applicable to 
Monte Carlo selection problems, including the first finite sampling
bounds for optimal policies in certain cases; we also provide a simple
counterexample to the intuitive conjecture that an optimal policy will
necessarily reach a decision in all cases.  We then derive heuristic
approximations in both Bayesian and distribution-free settings and
demonstrate their superiority to bandit-based heuristics in one-shot
decision problems and in Go.
\end{abstract}

### 2012-04-03 paper/vago.uai/blinkered.tex
as submitted

@@
+%% \section{The blinkered policy}
+
 %% \section{The blinkered policy}
@@
 
-The myopic policy is an extreme approximation, often stopping significantly early.
-For a natural class of metalevel probability models we can define a better approximation:
+The myopic policy is an extreme approximation, often stopping far too early.
+A better approximation can be obtained, at least for the case where
+each computation can only affect the value of one action.
+The technical definition (closely related to {\em subtree independence} in Russell and Wefald's work) is
+as follows:
 
 \begin{dfn}\label{dfn:independent-actions}
-	A metalevel probability model $\mathcal{M}=(U_1,\dots,U_k,\Evidence)$ 
+	A metalevel probability model $(U_1,\dots,U_k,\Evidence)$ 
 	has \term{independent actions} if the computational variables can be partitioned 
@@
 \end{dfn}
+
+With independent actions, we can talk about metalevel policies that focus on
+computations affecting a single action. These policies are not myopic---they can consider arbitrarily many computations---but they are {\em blinkered} because they can look
+in only a single direction at a time:
 
@@
 
-This is a better approximation than the myopic policy as it takes into account more possible future strategies.
-Indeed, $Q^m(s,a) \le Q^b(s,a) \le Q^*(s,a)$.  Intuitively, whereas the myopic policy can only see one 
-computation ahead, the blinkered policy can see arbitrary far ahead but with blinkers on so can only see
-the computations involving a single action at a time.
+Clearly, blinkered policies are better than myopic: $Q^m(s,a) \le Q^b(s,a) \le Q^*(s,a)$.
+Moreover, the blinkered policy can be computed in time proportional to the number of arms, by breaking the
+decision problem into separate subproblems:
 
@@
 	a \term{one-action metalevel decision problem} for $i=1,\dots,k$ is the metalevel decision
-	problem $M^1_{i,m} = (S_i,s_0,A_{s0},T_i,R_i)$ defined by the metalevel probability
-	model $(U_0,U_i,\Evidence_i)$ with $U_0=m$.
+	problem $M^1_{i,\lambda} = (S_i,s_0,A_{s0},T_i,R_i)$ defined by the metalevel probability
+	model $(U_0,U_i,\Evidence_i)$ with $U_0=\lambda$.
 \end{dfn}
 
-Note that given that state $s$ of a metalevel decision problem, we can form a state
+Note that given a state $s$ of a metalevel decision problem, we can form a state
 $s_i$ by taking only the results of computations in $\Evidence_i$ (see \dfnref{dfn:metalevel-mdp}).
@@
 Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$ with independent actions,
-let $M^1_{i,\mu}$ be the $i$th one-action metalevel decision problem for $i=1,\dots,k$.
+let $M^1_{i,\lambda_i}$ be the $i$th one-action metalevel decision problem for $i=1,\dots,k$.
 Then for any $s\in S$, whenever $E_i\in A_s\cap\Evidence_i$ we have:
 \[
-	Q^b_M(s,E_i) = Q^*_{M^1_{i,m_i}}(s_i, E_i)
+	Q^b_M(s,E_i) = Q^*_{M^1_{i,\mu^*_{-i}}}(s_i, E_i)
 \]
-where $m_i = \max_{j\neq i} \mu_i(s)$.
+where $\mu^*_{-i} = \max_{j\neq i} \mu_j(s)$.
 \end{thm}
 
-\begin{comment}
+\begin{hiddenproof}
 	\begin{proof}
@@
 	\[
-		Q^\pi(s,E_i) = \IE^\pi_M[ -c\,N + \max(\mu_i(S_N), m_i) \given S_0=s, A_0=E_i].
+		Q^\pi_M(s,E_i) = \IE^\pi_M[ -c\,N + \max(\mu_i(S_N), m_i) \given S_0=s, A_0=E_i].
 	\]
 	Noting that $\mu_i(S_N)$ is a function only of $(S_N)_i$, and that since 
-	But then this is exactly $Q^{\pi'}_{M^1_{i,m_i}}(s_i, E_i)$.  Taking the supremum
+	But then this is exactly $Q^*_{M^1_{i,\mu^*_{-i}}}(s_i, E_i)$.  Taking the supremum
 	over $\pi$ gives the result.
 	\end{proof}	
-\end{comment}
+\end{hiddenproof}
 
 \thmref{thm:blinkered} shows that to compute the blinkered policy we need
-only compute the optimal policies for all $i$th one-action metalevel decision problems.
+only compute the optimal policies for $k$ separate one-action problems.
 
@@
 As a result, the blinkered policy can be numerically computed in time $O(D/c^2)$ 
-independent of $k$ by backwards induction, where $D$ is the number of points $m\in[0,1]$
-we compute $Q^*_{M^1_{i,m}}(s)$ for\footnote{in our experiments below, $D=129$ points are equally spaced,
-using linear interpolation between points}.  This will be worth the cost in 
-the same situations as mentioned at the of \secref{sec:optimal}.
+independent of $k$ by backwards induction, where $D$ is the number of points $\lambda\in[0,1]$
+for which we compute $Q^*_{M^1_{i,m}}(s)$.\footnote{in our experiments below, $D=129$ points are equally spaced,
+using linear interpolation between points.}  This will be worth the cost in 
+the same situations as mentioned at the end of \secref{sec:optimal}.
 
-\figref{fig:blinkered} compared the expected regret of the blinked policy on the 
-Bernoulli sampling problem with $k=25$ and varying cost of samples to several 
-alternatives, where the regret include the cost of sampling:
-\[
-	R = (\max_i U_i) - U_j + c\,n
-\]
-if the policy performs $n$ computations and selects action $j$.
-Myopic is the myopic policy for Bernoulli sampling. ESPb is the 
-policy recently proposed by \citet{Chick+Frazier:2011}, which approximates the 
-blinkered policy for sampling normal distributions by moving to continuous time
-and numerically solving a diffusion equation.  UCB1-B uses UCB$(1/\sqrt2)$ \cite{Auer+et+al:2002} to
-choose which actions to sample and the blinkered policy to stop.  UCB1-b uses
-UCB$(1/\sqrt2)$ to sample, is given the average number of samples blinkered
-uses for the given cost.
-
+\figref{fig:blinkered} compares the blinkered policy 
+to several other policies from the literature, using a
+Bernoulli sampling problem with $k=25$ and a wide range of values for the step cost $c$.
+Performance is measured by expected {\em regret}, where the regret includes the cost of sampling:
+$R = (\max_i U_i) - U_j + c\,n$
+where $n$ is the number of computations and $j$ is the action actually selected.
 The blinkered policy significantly outperforms all others.  The myopic policy
 plateaus as it quickly reaches a position where no single computation can
-change the final decision between actions.  ESPb performs quite well given
+change the final action choice.  ESPb performs quite well given
 that is making a normal approximation to the Beta posterior.  
-UCB1-B and UCB1-b's curve shows that even given a good stopping rule, UCB1's
-choice of actions to sample can be significantly improved on.
+The curves for UCB1-B and UCB1-b show that even given a good stopping rule, UCB1's
+choice of actions to sample is not ideal.
 
@@
 \includegraphics[scale=0.7, trim=90 70 400 300]{blinkered-regret.pdf}
-\caption{Average regret of various policies as a function of the cost in 
-a 25-action Bernoulli sampling problem.  Averaged over 1000 trials.}
+\caption{Average regret of various policies as a function of the cost
+in a 25-action Bernoulli sampling problem, over 1000 trials. Error bars
+omitted as they are negligible (the relative error is at most 0.03).}
 \label{fig:blinkered}
@@
 \begin{comment}
+# N: In ~/research/src/mlrl
+
+set terminal postscript enhanced linewidth 2
+
+set size 0.5, 0.5
+set logscale xy
+
+set xlabel "Cost"
+set ylabel "Samples"
+set key left top
+
+set output "blinkered-steps.ps"
+
+plot [0.1:0.00047] "blinkered-steps.dat" title "Blinkered" with lines lw 2, "myopic-steps.dat" title "Myopic" with lines lw 2, "espb-steps.dat" title "ESPb" with lines lw 2, "ucb-blinkered-steps.dat" title "UCB1-B" with lines lw 2, "ucb-blinkered-table-steps.dat" with lines lw 2 title "UCB1-b"
+
 set xlabel "Cost"
 set ylabel "Regret"
+set key left bottom
 
-set terminal postscript enhanced linewidth 2
 set output "blinkered-regret.ps"
-set size 0.5, 0.5
 
-set logscale xy
+plot [0.1:0.00047] [0.05:1] "blinkered-regret.dat" title "Blinkered" with lines lw 2, "myopic-regret.dat" title "Myopic" with lines lw 2, "espb-regret.dat" title "ESPb" with lines lw 2, "ucb-blinkered-regret.dat" title "UCB1-B" with lines lw 2, "ucb-blinkered-table-regret.dat" with lines lw 2 title "UCB1-b"
 
-set key left bottom
-plot [0.1:0.0005] [0.05:1] "blinkered-regret.dat" title "Blinkered" with lines lw 2, "myopic-regret.dat" title "Myopic" with lines lw 2, "espb-regret.dat" title "ESPb" with lines lw 2, "ucb-blinkered-regret.dat" title "UCB1-B" with lines lw 2, "ucb-blinkered-table-regret.dat" with lines lw 2 title "UCB1-b"
+set output "blinkered-regret-linespoints.ps"
+
+plot [0.1:0.00047] [0.05:1] "blinkered-regret.dat" title "Blinkered" with linespoints lw 2, "myopic-regret.dat" title "Myopic" with linespoints lw 2, "espb-regret.dat" title "ESPb" with linespoints lw 2, "ucb-blinkered-regret.dat" title "UCB1-B" with linespoints lw 2, "ucb-blinkered-table-regret.dat" with linespoints lw 2 title "UCB1-b"
+
+set output "blinkered-regret-errorlines.ps"
+
+plot [0.1:0.00047] [0.05:1] "blinkered-regret2.dat" title "Blinkered" with errorlines lw 2, "myopic-regret2.dat" title "Myopic" with errorlines lw 2, "espb-regret2.dat" title "ESPb" with errorlines lw 2, "ucb-blinkered-regret2.dat" title "UCB1-B" with errorlines lw 2, "ucb-blinkered-table-regret2.dat" with errorlines lw 2 title "UCB1-b"
+
+set output "blinkered-regret-errorlines-nopoints.ps"
+
+plot [0.1:0.00047] [0.05:1] "blinkered-regret.dat" notitle with lines lw 2, "myopic-regret.dat" notitle with lines lw 2, "espb-regret.dat" notitle with lines lw 2, "ucb-blinkered-regret.dat" notitle with lines lw 2, "ucb-blinkered-table-regret.dat" with lines lw 2 notitle, "blinkered-regret2.dat" title "Blinkered" with errorbars lw 2, "myopic-regret2.dat" title "Myopic" with errorbars lw 2, "espb-regret2.dat" title "ESPb" with errorbars lw 2, "ucb-blinkered-regret2.dat" title "UCB1-B" with errorbars lw 2, "ucb-blinkered-table-regret2.dat" with errorlines lw 2 title "UCB1-b"
 \end{comment}

### 2012-04-03 paper/vago.uai/conclusion.tex
as submitted

\section{Conclusion}

The selection problem has numerous applications. This paper formalized the
problem as a belief-state MDP and proved some important properties of the 
resulting formalism. An application of the selection problem to control of
sampling was examined, and the insights provided by properties of the MDP 
led to approximate solutions that improve the state of the art. This was
shown in empirical evaluation both in ``flat" selection and when extending
the methods to game-tree search for the game of Go.

The methods proposed in the paper open up several new research
directions. The first is a better approximate solution of the MDP,
that should lead to even better flat sampling algorithms for
selection. A more ambitious goal is extending the formalism to
trees---in particular, achieving better sampling at non-root nodes,
for which the purpose of sampling differs from that at the root.

### 2012-04-03 paper/vago.uai/context.tex
as submitted

@@
 Consider a metalevel probability model with three actions. 
-	$U_0=m$ has a known value (the context),
-	$U_1$ is equally likely to be -1.5 or 1.5 (low mean, high variance),
-	and $U_2$ is equally likely to be 0.25 or 1.75 (high mean, low variance).
-The two computations are to observe exactly $U_1$ and $U_2$, respectively, with a cost 0.2 of computation.
+	$U_1$ is equally likely to be $-1.5$ or $1.5$ (low mean, high variance),
+	$U_2$ is equally likely to be $0.25$ or $1.75$ (high mean, low variance),
+	and $U_3=\lambda$ has a known value (the context).
+The two computations are to observe exactly $U_1$ and $U_2$, respectively, each with cost $0.2$.
 The corresponding metalevel MDP has 9 states and can be solved exactly.
-\figref{fig:swap-counterex} plots $Q^*_m(s_0,U_i) - Q^*_m(s_0,\bot)$ as a function of the known value $m$.
-As the context $m$ varies the optimal action inverts from observing 1 to observing 2.
+\figref{fig:swap-counterex} plots $Q^*_\lambda(s_0,U_i) - Q^*_\lambda(s_0,\bot)$ as a function of the known value $\lambda$.
+As the context $\lambda$ varies the optimal action inverts from observing 1 to observing 2.
 Inversions like this are impossible for index policies.
@@
 \includegraphics[scale=0.7, trim=90 70 400 300]{swap-counterex.pdf}
-\caption{Optimal Q-values as a function of the utility of the fixed alternative.  Note the inversion
-where for low $m$ observing action~$1$ is strictly optimal, while for medium $u$ observing action~$2$ is strictly optimal.}
+\caption{Optimal Q-values of computing relative to stopping as a function of the utility of the fixed alternative.  Note the inversion
+where for low $\lambda$ observing action~$1$ is strictly optimal, while for medium $\lambda$ observing action~$2$ is strictly optimal.}
 \label{fig:swap-counterex}
@@
 	set output "swap-counterex.ps"
-set xlabel "Utility of fixed alternative (m)"
+	set xlabel "Utility of fixed alternative ({/Symbol l})"
 	set ylabel "Utility relative to stopping"
@@
 \begin{dfn}\label{dfn:context}
-	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$, and a constant $\mu\in\R$,
-	define $M_\mu = (S,s_0,A_s,T,R_\mu)$ to be $M$ with an additional action of known 
-	value $\mu$, defined by:
+	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$, and a constant $\lambda\in\R$,
+	define $M_\lambda = (S,s_0,A_s,T,R_\lambda)$ to be $M$ with an additional action of known 
+	value $\lambda$, defined by:
 	\begin{align*}
-		R_\mu(s,E,s')      &= R(s,E,s') \\
-		R_\mu(s,\bot,\bot) &= \max\{\mu, R(s,\bot,\bot)\}
+		R_\lambda(s,E,s')      &= R(s,E,s') \\
+		R_\lambda(s,\bot,\bot) &= \max\{\lambda, R(s,\bot,\bot)\}
 	\end{align*}
@@
 
+Note this is equivalent to adding an extra arm with constant value $U_{k+1}=\lambda$.
+
 \begin{thm}
 	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$, 
-	there exists a real interval $I(s)$ for every state $s\in I$ such that
+	there exists a real interval $I(s)$ for every state $s\in S$ such that
 	it is optimal to stop in state $s$ in $M_\mu$ iff $\mu\notin I(s)$.
@@
 
-\begin{comment}
+\begin{hiddenproof}
 	\begin{proof}
@@
 	\end{proof}	
-\end{comment}
+\end{hiddenproof}
 

### 2012-04-03 paper/vago.uai/introduction.tex
as submitted

@@
 in order to find a move in the current state that is (hopefully)
-near-optimal. One of the most successful
-examples of this approach are sampling algorithms for
-playing the game of Go \citep{Gelly.mogo} based on the upper-confidence bound (UCB)
-method \citep{Kocsis+Szepesvari:2006}. This paper examines the issue of sampling for move selection from
-first principles, by defining an appropriate meta-level MDP and formally
-analyzing some of its properties. Although
-the meta-level MDP is also intractable, approximation schemes based on
-provable value of information bounds for this problem are presented. 
-These lead to practical sampling schemes that
-are promising: they out-perform the UCB-based schemes,
-as indicated by empirical results on both ``flat'' selection and
-game-playing in Go.
+near-optimal. In this paper, we develop a theoretical framework to examine the problem of selecting
+{\em which} future sequences to simulate.
+We derive a number of new results concerning optimal policies for this selection problem
+as well as new heuristic policies for controlling Monte Carlo simulations.
+As described below, these policies outperform previously published methods for
+``flat'' selection and game-playing in Go.
 
-As selection is extremely general \citep{TolpinShimony:2012}, we focus on the adversarial
-game-playing domain as an example for clarity.
-A typical game-playing algorithm explores a tree or graph of action sequences
-with some limit on depth, using pruning methods to avoid irrelevant
-subtrees; based on what it finds in the explored portion of the state space, the algorithm
-then selects a move. This exploration is repeated after observing the move
-made by the opponent (whether or not information from previous explorations
-is reused is discussed in Section \ref{sec:control-redistribution}).
-
-Clearly, it is desirable to select the best possible move using the
-least possible amount of exploration. For a given amount of
-exploration, decision quality can be improved by directing exploration
-towards those actions sequences whose outcomes are helpful in selecting
-a good move. Thus, the {\em metalevel} decision problem is to choose
-what future action sequences to explore (or, more generally, what
-deliberative computations to do), while the {\em object-level}
-decision problem is to choose an action to execute in the real world.
+The basic ideas behind our approach are best explained in a familiar
+context such as game playing.  A typical game-playing algorithm
+chooses a move by first exploring a tree or graph of move sequences
+and then selecting the most promising move based on this exploration.
+Classical algorithms typically explore in a fixed order, imposing a
+limit on exploration depth and using pruning methods to avoid
+irrelevant subtrees; they may also reuse some previous computations
+(see Section \ref{sec:control-redistribution}).  Exploring unpromising
+or highly predictable paths to great depth is often wasteful; for a
+given amount of exploration, decision quality can be improved by
+directing exploration towards those actions sequences whose outcomes
+are helpful in selecting a good move. Thus, the {\em metalevel}
+decision problem is to choose what future action sequences to explore
+(or, more generally, what deliberative computations to do), while the
+{\em object-level} decision problem is to choose an action to execute
+in the real world.
 
@@
 solved decision-theoretically was noted by \citet{Matheson:1968},
-borrowing directly from the related concept of {\em information value
+borrowing from the related concept of {\em information value
   theory}~\citep{Howard:1966}. In essence, computations can be
@@
 In MCTS, each computation takes the form
-of a simulating a randomized sequence of actions leading from a leaf of the
+of a simulation of a randomized sequence of actions leading from a leaf of the
 current tree to a terminal state. UCT is primarily a method for
@@
 playing \citep{Gelly+Silver:2011}.  The UCT algorithm is
-based on the the theory of bandit problems \citep{Berry+Fristedt:1985} and the near-optimal
+based on the the theory of bandit problems \citep{Berry+Fristedt:1985} and the asymptotically near-optimal
 UCB1 bandit algorithm \citep{Auer+et+al:2002}. UCT applies
@@
 It is natural to consider whether the two independent threads are
-consistent; for example, are bandit algorithms such as UCB1 approximate
-solutions to some particular case of the metalevel decision problem
-defined by Russell and Wefald? The answer, perhaps surprisingly, is no.
-The essential difference is that, in bandit problems, every trial involves
-executing a real object-level action with real costs, whereas in the metareasoning
-problem the trials are {\em simulations} whose cost is usually independent of the
-utility of the action being simulated. 
-Hence, in UCT, bandit algorithms are applied to problems that are not bandit problems.
+consistent; for example, are bandit algorithms such as UCB1
+approximate solutions to some particular case of the metalevel
+decision problem defined by Russell and Wefald? The answer, perhaps
+surprisingly, is no.  The essential difference is that, in bandit
+problems, every trial involves executing a real object-level action
+with real costs, whereas in the metareasoning problem the trials are
+{\em simulations} whose cost is usually independent of the utility of
+the action being simulated.  Hence, as \citet{Audibert+al:2010}
+and \citet{Bubeck+al:2011} have also noted, UCT applies bandit
+algorithms to problems that are not bandit problems.
 
@@
 problem} studied in statistics~\citep{Bechhofer:1954,Swisher+et+al:2003}.  Despite
-recent work by \citet{Frazier+Powell:2010} and \citet{TolpinShimony:2012}, the theory of selection
+some recent work \citep{Frazier+Powell:2010,TolpinShimony:2012}, the theory of selection
 problems is less well understood than that of bandit problems.
+Most work has focused on the probability of selection error rather than
+optimal policies~\citep{Bubeck+al:2011}.
 Accordingly, we present in \secrefs{sec:optimal}{sec:context}
 a number of results concerning optimal policies for the general case
-and for specific sampling distributions. In particular, we give finite
+as well as specific finite
 bounds on the number of samples collected by optimal policies for
-Bernoulli and Gaussian distributions. We also provide a simple
+Bernoulli arms with beta priors. We also provide a simple
 counterexample to the intuitive conjecture that an optimal policy
@@
 it is possible for an optimal policy to compute forever. We also show
-by a simple counterexample that optimal {\em index
-policies}~\citep{Gittins:1989} do not necessarily exist for selection
+by counterexample that optimal {\em index
+policies}~\citep{Gittins:1989} may not exist for selection
 problems.
@@
 in the game of Go.
+
+(Proofs omitted due to space constraints can be found in the supplementary material.)

### 2012-04-03 paper/vago.uai/optimalpolicies.tex
as submitted

@@
 
-In a \term{selection problem} the decision maker is faced with the choice between
-alternative actions.  To make this choice, they may gather evidence about the
-utility of each of these alternative, at some cost.  The objective is to maximize
-the expected utility of the final action selected, less the cost of computation.
-
-%% This problem has been studied in a number of related forms.  \note{Here}
-
-%% The Bayesian ranking and selection problem \citep{Swisher+et+al:2003} is an important
-%% special case where the computations performed are samples.
-
-%%What the decision maker can know and what they can find out is formalized in:
+In a selection problem the decision maker is faced with a choice among
+alternative arms\footnote{Alternative actions are known as \emph{arms} in the bandit setting;
+we borrow this terminology for uniformity.}.  To make this choice, they may gather evidence about the
+utility of each of these alternatives, at some cost.  The objective is to maximize
+the {\em net utility}, i.e., the expected utility of the final arm selected, less the cost of gathering the evidence. 
+In the classical case~\citep{Bechhofer:1954}, evidence might consist of physical samples
+from a product batch; in a metalevel problem with Monte Carlo simulations,
+the evidence consists of outcomes of sampling computations:
 
 \begin{dfn} \label{dfn:metalevel-model}
-	A \term{metalevel probability model} is a tuple $M=(U_1,\dots,U_k,\Evidence)$ 
+	A \term{metalevel probability model} is a tuple $(U_1,\dots,U_k,\Evidence)$ 
 	consisting of jointly distributed random variables:
 	\begin{itemize}
-		\item Real random variables $U_1,\dots,U_k$, where $U_i$ is the utility of performing action~$i$ (called ``arms" in the bandit setting), and
+		\item Real random variables $U_1,\dots,U_k$, where $U_i$ is the utility of arm~$i$, and
 		\item A countable set $\Evidence$ of random variables, each variable $E\in\Evidence$ being 
@@
 %% Not actually true that we necessarily define this with U's as roots:  in sampling models they are naturally /leaves/.
-
-A metalevel probability model, when combined with a cost $c>0$ of computation,%
-\footnote{The assumption of a fixed cost of computation is a simplification; 
-	precise conditions for its validity are given by \citep{Harada:1997}.} 
-defines a metalevel decision problem: what is the optimal strategy with which to choose a sequence 
-of computations $E\in\Evidence$ to observe in order to maximize the expected utility 
-of the action finally taken less the cost of computation?  Intuitively, 
-this strategy should observe the variables which give the most evidence relevant
-to deciding which action to perform, stopping when the cost of computation 
-outweighs the benefit gained.
+%% Bounedness assumption
+For simplicity, in the below we'll assume the utilities $U_i$ are bounded, 
+without loss of generality in $[0,1]$.
 
@@
 In the \term{Bernoulli metalevel probability model},
-each action will either succeed or not $U_i\in\{0,1\}$, with an unknown latent frequency of success $\Theta_i$, 
+each arm will either succeed or not $U_i\in\{0,1\}$, with an unknown latent frequency of success $\Theta_i$, 
 and stochastic simulations of possible consequences $E_{ij}$ can be performed:
 \begin{align*}
-	\Theta_i &\iid {\rm Uniform}[0,1]                      &&\quad\text{for $i=1,\dots,k$}\\
-	U_i \given \Theta_i &\sample {\rm Bernoulli}(\Theta_i) &&\quad\text{for $i=1,\dots,k$}\\
-	E_{ij} \given \Theta_i &\iid {\rm Bernoulli}(\Theta_i) &&\quad\text{for $i=1,\dots,k$ and $j=1,\dots$}
+	\Theta_i &\iid {\rm Uniform}[0,1]                      &&\quad\text{for $i\in\{1,\dots,k\}$}\\
+	U_i \given \Theta_i &\sample {\rm Bernoulli}(\Theta_i) &&\quad\text{for $i\in\{1,\dots,k\}$}\\
+	E_{ij} \given \Theta_i &\iid {\rm Bernoulli}(\Theta_i) &&\quad\text{for $i\in\{1,\dots,k\}$, $j\in\Nat$}
 \end{align*}
 
-The \term{one-action Bernoulli metalevel probability model} has $k=2$,
-$\Theta_1=m\in[0,1]$ a constant, and $\Theta_2\sample{\rm Uniform}[0,1]$.
+The \term{one-armed Bernoulli metalevel probability model} has $k=2$,
+$\Theta_1=\lambda\in[0,1]$ a constant, and $\Theta_2\sample{\rm Uniform}[0,1]$.
 \end{example}
 
-%% Bounedness assumption
-For simplicity, in the below we'll assume the utilities $U_i$ are bounded and so, 
-without loss of generality, lie in $[0,1]$.
-
-To formalize this problem, we use the theory of Markov Decision Processes (MDPs; 
-see, for example, \citet{Puterman:1994}):
+A metalevel probability model, when combined with a cost of computation $c>0$,%
+\footnote{The assumption of a fixed cost of computation is a simplification; 
+	precise conditions for its validity are given by \citet{Harada:1997}.} 
+defines a metalevel decision problem: what is the optimal strategy with which to choose a sequence 
+of computations $E\in\Evidence$ in order to maximize the agent's net utility?
+Intuitively, this strategy should choose the computations that give the most evidence relevant
+to deciding which arm to use, stopping when the cost of computation 
+outweighs the benefit gained. We formalize the selection problem as a Markov Decision Process (MDP; 
+see, e.g., \citet{Puterman:1994}):
 
@@
 	$A_s$ is a countable set of actions available in state $s\in S$,
-	$T(s,a,s')$ is the transition distribution 
-	giving the probability of transitioning from a state $s\in S$ to a state $s'\in S$ after performing action $a\in A_s$,
-	and $R(s,a,s')$ is the expected reward received when transitioning from $s\in S$ to $s'\in S$ after performing action $a\in A_s$.
+	$T(s,a,s')$ is the transition probability from $s\in S$ to $s'\in S$ after performing action $a\in A_s$,
+	and $R(s,a,s')$ is the expected reward received on such a transition.
 \end{dfn}
+
+To formulate the metalevel decision problem as an MDP, we define the states as sequences of
+computation outcomes and allow for a terminal state when the agent chooses to stop computing and act:
 
@@
 			we'll further assume the domains of the computational variables $E\in\Evidence$ are disjoint.}
-	$M=(U_1,\dots,U_k,\Evidence)$ and
+	$(U_1,\dots,U_k,\Evidence)$ and
 	a cost of computation $c>0$, a corresponding \term{metalevel decision problem}
-	is any MDP $M=(S,s_0,A_s,T,R)$ \begin{align*}
+	is any MDP $M=(S,s_0,A_s,T,R)$ such that
+	\begin{align*}
 		S &= \{\bot\}\cup\{\langle e_1\dots, e_n \rangle : e_i\in E_i \text{ for all $i$,} \\
@@
 	\intertext{where $\bot\in S$ is the unique terminal state,
-	where $\Evidence_s\subseteq\Evidence$ is a potentially state-dependent subset of allowed computations,
+	where $\Evidence_s\subseteq\Evidence$ is a state-dependent subset of allowed computations,
 	and when given any $s=\langle e_1, \dots, e_n \rangle\in S$,
@@
 	\end{align*}		
-	and where $\mu_i(s) = \IE[U_i \given E_1=e_1,\dots,E_n=e_n]$.
+	where $\mu_i(s) = \IE[U_i \given E_1=e_1,\dots,E_n=e_n]$.
 \end{dfn}
@@
 One can optionally add an external constraint on the number of computational actions, or their total cost,
-in the form of a deadline (called a budget henceforth).  This bridges with the related area of budgeted learning \citep{Madani+et+al:2004}.
+in the form of a deadline or {\em budget}.  This bridges with the related area of budgeted learning \citep{Madani+et+al:2004}.
 Although this feature is not formalized in the MDP, it can be added by including either time or past total cost 
@@
 by standard properties of these distributions, where $s_i=\sum_{j=1}^{n_i} E_{in_i}$
-is the number of simulated successes of action $i$, and $f_i=n_i-s_i$ the failures.  By \eqref{eq:bernoulli1}, 
-the state space is the set of all $k$ pairs $(s_i,f_i)$, \eqrefs{eq:bernoulli2}{eq:bernoulli3}
+is the number of simulated successes of arm~$i$, and $f_i=n_i-s_i$ the failures.  By \eqref{eq:bernoulli1}, 
+the state space is the set of all $k$ pairs $(s_i,f_i)$; \eqrefs{eq:bernoulli2}{eq:bernoulli3}
 suffice to give the transition probabilities and terminal rewards, respectively.
-The one-action Bernoulli case is similar, requiring only the two dimensional 
-space $(s,f)$ giving the posterior over $\Theta_2$.
+The one-armed Bernoulli case is similar, requiring as state just $(s,f)$ defining the posterior over $\Theta_2$.
 \end{example}
@@
 Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$ one defines policies and 
-value functions as in all MDPs.  A (deterministic, stationary) \term{metalevel policy} 
+value functions as in any MDP.  A (deterministic, stationary) \term{metalevel policy} 
 $\pi$ is a function mapping states $s\in S$ to actions to take in that state $\pi(s)\in A_s$.
@@
 received under that policy starting from a given state $s\in S$, and the \term{Q-function}
-does the state when starting in a state $s\in S$ and taking a given action $a\in A_s$: 
-\[
+does the same when starting in a state $s\in S$ and taking a given action $a\in A_s$: 
+\begin{equation}
 		V^\pi_M(s) = \IE^\pi_M\left[ \sum_{i=0}^N R(S_i,\pi(S_i),S_{i+1}) \given S_0 = s\right]	\label{eq:value-fn}\\
 	%%	Q^\pi_M(s,a) &= \IE^\pi_M\left[ \sum_{i=0}^N R(S_i,\pi(S_i),S_{i+1}) \given S_0 = s, A_0 = a\right] \label{eq:qvalue-fn}
-\]
+\end{equation}
 where $N\in[0,\infty]$ is the random time the MDP is terminated, i.e.,
-the unique time where $\pi(S_N)=\bot$%
-\footnote{Whenever $N<\infty$, $S_{N+1}=\bot$ the unique terminal state.
-Instead of having a random termination time, one can equivalently make
-the state $\bot$ absorbing with zero reward on all transitions.}, 
+the unique time where $\pi(S_N)=\bot$,
+%% \footnote{Whenever $N<\infty$, $S_{N+1}=\bot$ the unique terminal state.
+%% Instead of having a random termination time, one can equivalently make
+%% the state $\bot$ absorbing with zero reward on all transitions.}
 and similarly for the Q-function $Q^\pi_M(s,a)$.
 
-As usual, an \term{optimal policy} $\pi^*$, when it exists, is one which maximises 
+As usual, an \term{optimal policy} $\pi^*$, when it exists, is one that maximizes 
 the value from every state $s\in S$, i.e., if we define for each $s\in S$
@@
 \]
-then the optimal policy $\pi^*$ satisfies $V^{\pi^*}_M(s) = V^*_M(s)$
-for all $s\in S$, where we break ties towards stopping $\bot$.
+then an optimal policy $\pi^*$ satisfies $V^{\pi^*}_M(s) = V^*_M(s)$
+for all $s\in S$, where we break ties in favor of stopping.
 %% and by standard results, $\pi^*(s) \in \argmax_{a\in A_s} Q^*_M(s,a)$, where
@@
 The optimal policy must balance the cost of computations with the improved decision
-quality that results.  This is made clear by giving by expressing the value function:
+quality that results.  This tradeoff is made clear in the value function:
 
@@
 	\]
-	where the random variable $N$ denotes the total number of computations performed,
-	and similarly for $Q^\pi_M(s,a)$.
+	where $N$ denotes the (random) total number of computations performed;
+	similarly for $Q^\pi_M(s,a)$.
 \end{thm}
@@
 \begin{thm}\label{thm:bounded-expected-computations}
-	The optimal policy $\pi^*$'s expected number of computations is bounded by the 
-	perfect value of information \citep{Howard:1966} times the inverse cost $1/c$:
+	The optimal policy's expected number of computations is bounded by the 
+	value of perfect information \citep{Howard:1966} times the inverse cost $1/c$:
 	\[
@@
 	Further, any policy $\pi$ with infinite expected number of computations 
-	$\IE^\pi[N]=\infty$ has negative infinite value, so in particular the optimal
-	policy stops with probability 1.
+	%% $\IE^\pi[N]=\infty$
+	has negative infinite value, hence the optimal
+	policy stops with probability one.
 \end{thm}
@@
 	The first follows as in state $s$ the optimal policy has value at least that
-	of stopping immediately $\max_i \mu_i(s)$, and as $\IE \max_i\mu_i(S_N) \le \IE \max_i U_i$ by Jensen's inequality.
-	The second follows from \thmref{thm:value-of-computation} as $c>0$.
+	of stopping immediately ($\max_i \mu_i(s)$), and as $\IE \max_i\mu_i(S_N) \le \IE \max_i U_i$ by Jensen's inequality.
+	The second from \thmref{thm:value-of-computation}.
 \end{proof}
 
-However, although the \emph{expected} number of computations is always bounded,
+Although the \emph{expected} number of computations is always bounded,
 there are important cases in which the \emph{actual} number is not, such as
-the following example inspired by the sequential probability ratio test \citep{Wald+1945}:
+the following inspired by the sequential probability ratio test \citep{Wald+1945}:
 
 \begin{example}\label{example:sprt}
-Consider the Bernoulli sampling model for two actions but with different priors:
+Consider the Bernoulli sampling model for two arms but with a different prior:
 	$\Theta_1=1/2$,
@@
 %
-Simulating action 1 gains nothing, and after $(s,f)$ simulated successes and failures of action 2
-the posterior odds is
+Simulating arm~1 gains nothing, and after $(s,f)$ simulated successes and failures of arm~2
+the posterior odds ratio is
 \[
@@
 \]
-Note that these odds completely specify the posterior distribution of $\Theta_2$,
-and thus the distribution of the utility and all future computations.  Thus, whether
-it is optimal to continue is a function only of this posterior odds, and thus $s-f$.
+Note that this ratio completely specifies the posterior distribution of $\Theta_2$,
+and hence the distribution of the utilities and all future computations.  Thus, whether
+it is optimal to continue is a function only of this ratio, and thus of $s-f$.
 %% In particular,
@@
 For sufficiently low cost, the 
-optimal policy samples when $s-f$ equals $-1$, $0$, and $1$.  But with probability
+optimal policy samples when $s-f$ equals $-1$, $0$, or $1$.  But with probability
 $1/3$, a state with $s-f=0$ transitions to another state $s-f=0$ after two samples, 
@@
 
-Fortunately, in a number of settings, including the original Bernoulli model of \exampleref{example:bernoulli},
-we can prove an upper bound the number of computations.  For reasons of space,
-and for its later use in \secref{sec:blinkered}, we prove here the bound for the one-action Bernoulli model.
+However, in a number of settings, including the original Bernoulli model of \exampleref{example:bernoulli},
+we can prove an upper bound on the number of computations.  For reasons of space,
+and for its later use in \secref{sec:blinkered}, we prove here the bound for the one-armed Bernoulli model.
 
@@
 	where $Q^m(s,\bot) = \max_i \mu_i(s)$ and
-	\begin{equation}\label{eq:myopic}
+	\begin{equation*}%% \label{eq:myopic}
 		 Q^m(s,E) = \IE_M[ -c + \max_i \mu_i(S_1) \given S_0 = s, A_0 = E].		
-	\end{equation}
+	\end{equation*}
 \end{dfn}
@@
 under the assumption that at most one further computation can be performed.
-%% This policy is typically easy to compute, although not necessarily very good.
-
-It does have importance, however, in how it constrains the optimal policy in \thmref{thm:optimal-myopic}
-and its partial converse \thmref{thm:myopic-optimal}:
+It has a tendency to stop too early, because changing one's mind about which real action to take often takes more than one computation.
+In fact, we have:
 
@@
 	\end{proof}
-\end{comment}
+\end{hiddenproof}
 
@@
 
-This shows that the myopic policy is conservative in stopping, so stopping too early
-is one of its approximation errors.  However myopic stopping can be used to control optimal stopping:
+Despite this property, the stopping behavior of the myopic policy does
+have a close connection to that of the optimal policy:
 
@@
 	and a subset $S'\subseteq S$ of states closed under transitions,	
-	if the myopic policy stops in all states $s\in S$
+	if the myopic policy stops in all states $s'\in S'$
 	then the optimal policy does too.	
@@
 \begin{thm}\label{thm:one-action-bound}
-	The one-action Bernoulli decision process with constant arm $m\in[0,1]$ 
-	performs at most $m(1-m)/c-3 \le 1/4c-3$ computations.
+	The one-armed Bernoulli decision process with constant arm $\lambda\in[0,1]$ 
+	performs at most $\lambda(1-\lambda)/c-3 \le 1/4c-3$ computations.
 \end{thm}
+\begin{proof}
+Using \dfnref{dfn:myopic} and \exampleref{example:bernoulli2}, we determine
+which states the myopic policy stops in by bounding $Q^m(s,E)$.  For a state $(s,f)$,
+let $\mu=(s+1)/(n+2)$ be the mean utility for arm~2, where $n=s+f$.
+Fixing $n$ and maximizing over $\mu$, we get sufficient condition for stopping
+\begin{equation}
+	c \ge \frac{\lambda(1-\lambda)}{(n+3)} \qquad\qquad   n\ge \frac{\lambda(1-\lambda)}{c} - 3  \label{eq:myopic-stopping}
+\end{equation}
+Since the set of states satisfying \eqref{eq:myopic-stopping} is closed under
+transitions ($n$ only increases), by \thmref{thm:optimal-myopic}.  Finally, note $\max_{\lambda\in[0,1]} \lambda(1-\lambda)=1/4$.
+\end{proof}
 
-\begin{comment}
+\begin{hiddenproof}
 	\begin{proof}
@@
 		\end{align}
-		where $\mu=(s+1)/(n+2)$  is the mean utility for action~2, where $n=s+f$,
+		where $\mu=(s+1)/(n+2)$  is the mean utility for arm~2, where $n=s+f$,
 		$\mu^- = \mu - \mu/(n+3)$, and $\mu^+ = \mu + (1-\mu)/(n+3)$
-		are the posterior means of action 2 after simulating a failure and a success,
+		are the posterior means of arm~2 after simulating a failure and a success,
 		%% and $\mu^-_i=(s_i+1)/(s_i+f_i+3) $ and $\mu^+_i = (s_i+2)/(s_i+f_i+3)$
@@
 		\begin{equation}
-			c \ge \frac{m(1-m)}{(n_i+3)} \qquad\qquad   n_i\ge \frac{m(1-m)}{c} - 3  \label{eq:myopic-stopping}
+			c \ge \frac{\lambda(1-\lambda)}{(n+3)} \qquad\qquad   n\ge \frac{\lambda(1-\lambda)}{c} - 3  \label{eq:myopic-stopping}
 		\end{equation}
 		Since the set of states satisfying \eqref{eq:myopic-stopping} is closed under
-		transitions ($n$ only increases), by \thmref{thm:optimal-myopic}.  Finally, note $\max_{m\in[0,1]} m(1-m)=1/4$.
+		transitions ($n$ only increases), by \thmref{thm:optimal-myopic}.  Finally, note $\max_{\lambda\in[0,1]} \lambda(1-\lambda)=1/4$.
 	\end{proof}	
-\end{comment}
+\end{hiddenproof}
 
 A key implication is that the \emph{optimal} policy can be computed
-in time quadratic in the inverse cost $O(1/c^2)$.  This is particularly appropriate when the cost of 
+in time $O(1/c^2)$, i.e., quadratic in the inverse cost.  This is particularly appropriate when the cost of 
 computation is relatively high, such as in simulation experiments \citep{Swisher+et+al:2003},
-or when the decision to be made is critical.
+or when the decision to be made is critical. Finite bounds for the $k$-armed problem
+can also be derived and will be included in the full paper.

### 2012-04-03 paper/srcr/paper.tex
as submitted
trees fixed

@@
 
-Monte-Carlo tree sampling, and especially a version based on the
+Monte-Carlo tree search, and especially a version based on the
 UCT formula \cite{Kocsis.uct} appears in numerous search applications,
@@
 uncertainty, the goal is typically to either find a good (or optimal)
-strategy, or even to find the best first action of such a policy. Once
+strategy, or even just to find the best first action of such a policy. Once
 such an action is discovered, it is usually not beneficial to further sample
@@
 \end{equation}
-where $\overline X_i$ is the average reward obtained from arm $i$,
+where $\overline X_i$ is the average sample reward obtained from arm $i$,
 $n_i$ is the number of times arm $i$ was pulled, and $n$ is the total
@@
 The \textbf{simple regret} $\IE r$ of a sampling policy for the Multi-armed Bandit
-Problem is the expected difference between the best expected reward
-$\mu_*$ and the expected reward $\mu_j$ of the empirically best arm
-$\overline X_j=\max_i\overline X_i$:
+Problem is the expected difference between the best true expected reward
+$\mu_*$ and the true expected reward $\mu_j$ of the empirically best arm
+$j =\arg \max_i\overline X_i$:
 \begin{equation}
-\IE r=\sum_{j=1}^K\Delta_j\Pr(\overline X_j=\max_i\overline X_i)
+\IE r=\sum_{j=1}^K\Delta_j\Pr(j=\arg\max_i\overline X_i)
 \label{eqn:simple-regret}
@@
 UCB($c$) at all steps but a larger value of $c$ for the first step
-than for the rest of the steps, outperforms UCT.
+than for the rest of the steps, should outperform UCT.
 
@@
 vs. the number of samples, averaged over $10000$ experiments for
-randomly generated instances of 32 arms. For smaller numbers of
-samples, $\frac 1 2$-greedy achieves the best
-performance; for larger numbers of samples, UCB$_{\sqrt{\cdot}}$
-outperforms $\frac 1 2$-greedy. A combination of $\frac 1
-2$-greedy  and UCB$_{\sqrt{\cdot}}$ dominates UCB over the
-whole range.
+randomly generated instances of 32 arms. Either $\frac 1
+2$-greedy  or UCB$_{\sqrt{\cdot}}$ dominate UCB over the
+whole range. For larger number of samples, the advantage
+of UCB$_{\sqrt{\cdot}}$ over $\frac 1 2$-greedy becomes more significant.
 
@@
     \includegraphics[scale=0.8]{onelevel-tree.pdf}\\
-    a. search tree
+    a. search tree, the best arm shaded
     \vspace{1em}
@@
 Simple regret vs. the number of samples are shown for trees with root degree 16
-(Figure~\ref{fig:mcts-regret}.b) and 64 (Figure~\ref{fig:mcts-regret}.c).
+(Figure~\ref{fig:mcts-regret}.b) and 64
+(Figure~\ref{fig:mcts-regret}.c). The exploration factor $c$ is set to
+$2$, the default value for rewards in the range $[0, 1]$.
 The algorithms exhibit a similar relative performance: either $\frac 1
@@
     \includegraphics[scale=0.7]{twolevel-tree.pdf}\\
-    a. search tree
+    a. search tree, the path to the best arm shaded
     \vspace{0.5em}

### 2012-04-03 paper/srcr/paper.tex
fixed for submissions

@@
 UCT formula \cite{Kocsis.uct} appears in numerous search applications,
-such as \cite{Eyerich.ctp}. Although these methods are shown to be successful empirically,
+such as \cite{GellyWang.mogo,Eyerich.ctp}. Although these methods are shown to be successful empirically,
 most authors appear to be using the UCT formula ``because it has been shown
@@
   measurements.
-% When applying the selection problem to MCTS, the
-%  situation is exacerbated.  The utility of an action is usually
-%  bounded, and thus in many cases a single sample may be insufficient
-%  to change the current best action, \emph{regardless} of its
-%  outcome. As a result, we frequently get a \emph{zero} ``myopic''
-%  value of information for a single sample.
+
+When applying the selection problem to MCTS, the
+situation is exacerbated.  The utility of an action is usually
+bounded, and thus in many cases a single sample may be insufficient
+to change the current best action, \emph{regardless} of its
+outcome. As a result, we frequently get a \emph{zero} ``myopic''
+value of information for a single sample.
 \item Rational metareasoning requires a known distribution model, which may be
@@
 
-The pseudocode of this two-stage rollout is in
+The pseudocode of this two-stage rollout for an undiscounted MDP is in
 Algorithm~\ref{alg:two-stage-mcts}: \textsc{FirstAction} selects the first
@@
 \textsc{NextAction} (line~\ref{alg:srcr-next-action}) selects steps during
-the rest of the layout (usually using UCB).
+the rest of the layout (usually using UCB). The reward statistic for
+the selected action is updated (line~\ref{alg:srcr-update-stat}), and
+the sample reward is back-propagated (line~\ref{alg:return-reward}) 
+towards the current root.
+
 We denote such two-step realizations of SR+CR as \emph{Alg}+UCT, where
@@
      \State \hspace{4em} + \Call{Rollout}{next-node, depth+1}
-    \State \Call{UpdateStats}{node, action, reward}
-    \State \textbf{return} reward
+    \State \Call{UpdateStats}{node, action, reward} \label{alg:srcr-update-stat}
+    \State \textbf{return} reward \label{alg:return-reward}
   \EndIf
@@
     a. search tree, the path to the best arm shaded
-    \vspace{0.5em}
+    \vspace{1em}
   \end{minipage}
@@
     \centering
-    \includegraphics[scale=0.425]{costs-size=6-group=median.pdf}\\
+    \includegraphics[scale=0.45]{costs-size=6-group=median.pdf}\\
     a. median cost
-    \vspace{0.5em}
+    \vspace{1em}
   \end{minipage}
@@
     \centering
-    \includegraphics[scale=0.425]{costs-size=6-group=minimum.pdf}\\
+    \includegraphics[scale=0.45]{costs-size=6-group=minimum.pdf}\\
     b. minimum cost
@@
   \centering
-  \includegraphics[scale=0.425]{costs-size=6-nsamples=397.pdf}\\
+  \includegraphics[scale=0.45]{costs-size=6-nsamples=397.pdf}\\
   a. 397 rollouts\\
-  \vspace{0.5em}
-  \includegraphics[scale=0.425]{costs-size=6-nsamples=1585.pdf}\\
-  c. 1585 rollouts
+  \vspace{1em}
+  \includegraphics[scale=0.45]{costs-size=6-nsamples=1585.pdf}\\
+  b. 1585 rollouts
   \caption{The sailing domain, $6\times 6$ lake, cost vs. factor}
   \label{fig:sailing-cost-vs-factor}
-  \vspace{-1.5em}
 \end{figure}
@@
    \centering
-   \includegraphics[scale=0.425]{costs-size=6-nsamples=397.pdf}\\
-   b. $6\times 6$ lake \\
-   \vspace{0.5em}
-   \includegraphics[scale=0.425]{costs-size=10-nsamples=397.pdf}\\
-   c. $10\times 10$ lake
+   \includegraphics[scale=0.45]{costs-size=6-nsamples=397.pdf}\\
+   a. $6\times 6$ lake \\
+   \vspace{1em}
+   \includegraphics[scale=0.45]{costs-size=10-nsamples=397.pdf}\\
+   b. $10\times 10$ lake
   \caption{The sailing domain, 397 samples, cost vs. factor}
@@
   \label{fig:mcts-regret-voi} 
-  \vspace{-1.5em}
 \end{figure}
@@
 better general performance of the proposed scheme.
+
+MCTS and UCT are particularly successful in Computer
+Go \cite{GellyWang.mogo,Braudis.pachi,Enzenberger.Fuego}
+and the proposed scheme should be evaluated on this domain, a
+non-trivial task as Go programs typically use ``non-pure''
+versions of UCT, extended with domain-specific knowledge.
 

### 2012-04-04 paper/srcr/paper.tex
current state

@@
 Algorithm~\ref{alg:two-stage-mcts}: \textsc{FirstAction} selects the first
-step of a layout (line~\ref{alg:srcr-first-action}), and
+step of a rollout (line~\ref{alg:srcr-first-action}), and
 \textsc{NextAction} (line~\ref{alg:srcr-next-action}) selects steps during
-the rest of the layout (usually using UCB). The reward statistic for
+the rest of the rollout (usually using UCB). The reward statistic for
 the selected action is updated (line~\ref{alg:srcr-update-stat}), and

### 2012-04-24 paper/srcr/paper.tex

@@
 We begin with background definitions and related work. 
-Some sampling schemes are introduced, and shown to have better than UCB
-bounds for the simple regret on sets, the first contribution of this paper. 
+Some sampling schemes are introduced, and shown to have better
+bounds for the simple regret on sets than UCB, the first contribution of this paper. 
 The results are applied to sampling in trees by combining the proposed sampling schemes on the
@@
 generated 2-level trees, and on the sailing domain, showing where the proposed schemes have improved
-performance, the third contribution of this paper.
+performance.
 
@@
 shown to outperform many state of the art search algorithms in both
-MDP and adversarial search \cite{Eyerich.ctp}.
+MDP and adversarial games \cite{Eyerich.ctp,GellyWang.mogo}.
 
@@
 Problem is the expected difference between the best true expected reward
-$\mu_*$ and the true expected reward $\mu_j$ of the empirically best arm
+$\mu_*$ and the true expected reward $\mu_j$ of the arm with the greatest sample mean,
 $j =\arg \max_i\overline X_i$:
@@
 As the above ultimate goal is extremely difficult
-to achieve, we introduce
+to achieve, we introduce in this paper
 simple schemes more amenable to analysis, loosely based on the metareasoning
@@
 \begin{dfn} The \textbf{$\mathbf{\varepsilon}$-greedy} sampling scheme
-pulls the empirically best arm with probability
-$0<\varepsilon<1$ and any other
+pulls the arm that currently has the greatst sample mean, with probability
+$0<\varepsilon<1$, and any other
 arm with probability $\frac {1-\varepsilon} {K-1}$. 
@@
 Bound the probability $P_i$ that a non-optimal arm $i$ is selected. Split the interval
- $[\mu_i, \mu_*]$ at $\mu_i+\delta_i$. By Chernoff-Hoeffding bound:
+ $[\mu_i, \mu_*]$ at $\mu_i+\delta_i$. Apply the Chernoff-Hoeffding bound to get:
 \begin{eqnarray}
@@
 {n(1-\varepsilon)} {K-1}$ as $n\rightarrow \infty$. Conclude that for
-any $0<\eta<1$, $\gamma>1$ there exists $N$ such that for any $n>N$ and
+every $0<\eta<1$, $\gamma>1$ there exists $N$ such that for every $n>N$ and
 all non-optimal arms $i$:
@@
 
-$\varepsilon$-greedy is based solely on sampling the empirically best
-arm with a higher probability then the rest of the arms, and ignores
-information about empirical means of other arms. On the other hand,
-UCB distributes samples in accordance with empirical means, but, in order to
-minimize cumulative regret, chooses the empirically best arm too often.
+$\varepsilon$-greedy is based solely on sampling the
+arm that has the greatest sample mean (henceforth called the ``current best'' arm)
+with a higher probability then the rest of the arms, and ignores
+information about sample means of other arms. On the other hand,
+UCB distributes samples in accordance with sample means, but, in order to
+minimize cumulative regret, chooses the current best arm too often.
 Intuitively, a better scheme for simple regret minimization would
-distribute samples in a way similar to UCB, but would sample the best arm
-less. This can be achieved by replacing $\log(\cdot)$ in
+distribute samples in a way similar to UCB, but would sample the current best arm
+less often. This can be achieved by replacing $\log(\cdot)$ in
 Equation~\ref{eqn:ucb} with a faster growing sublinear function, for
@@
   non-optimal arm $i$ is chosen. Split the interval $[\mu_i, \mu_*]$
-  at $\mu_i+\frac {\Delta_i} 2$. By Chernoff-Hoeffding bound:
+  at $\mu_i+\frac {\Delta_i} 2$. Apply the Chernoff-Hoeffding bound to get:
 \begin{eqnarray}
@@
 {\Delta_i^2}$, $n_i\le n_*$ as $n\to\infty$. Conclude that for
-any $0<\eta<1$, $\gamma>1$ there exists $N$ such that for any $n>N$ and
+every $0<\eta<1$, $\gamma>1$ there exists $N$ such that for every $n>N$ and
 all non-optimal arms $i$:
@@
 choice with high confidence, the value of information of additional
-samples of the best move is low. Therefore, one should be able to do
-better than UCT by optimizing {\em simple regret}, rather than {\em
-  cumulative regret}, at the root node.
+samples of the best move (or, in fact, of any other samples) is low. Therefore, one should be able to do
+better than UCT by optimizing {\em simple regret}, rather than {\em cumulative regret}, at the root node.
 
 Nodes deeper in the search tree are a different matter.
-In order to support an optimal move choice at the root, a precise {\em value} of the search
-tree nodes is beneficial. For these internal nodes, optimizing simple regret is not the answer, and 
-cumulative regret optimization seems more reasonable. Lacking a complete metareasoning for sampling,
+In order to support an optimal move choice at the root, it is beneficial in many cases
+to find a more precise estimate of the {\em value} of the state in these search
+tree nodes. For these internal nodes, optimizing simple regret is not the answer, and 
+cumulative regret optimization is not so far off the mark. Lacking a complete metareasoning for sampling,
 which would indicate the optimal way to sample both root nodes and internal nodes,
@@
 current root node according to a scheme suitable for minimizing 
-the simple regret (\textbf{SR}), such as $\frac 1 2$-greedy or UCB$_{\sqrt{\cdot}}$, and
-then selects actions according to UCB, minimizing the cumulative regret (\textbf{CR}).
+the simple regret (\textbf{SR}), such as $\frac 1 2$-greedy or UCB$_{\sqrt{\cdot}}$, and (at non-root nodes)
+then selects actions according to UCB, which approximately minimizes the cumulative regret (\textbf{CR}).
 \end{dfn}
@@
 UCB($c$) at all steps but a larger value of $c$ for the first step
-than for the rest of the steps, should outperform UCT.
+than for the rest of the steps, should also outperform UCT.
 
@@
 \end{eqnarray}
-with $VOI_{\alpha}$ being (approximate) value for sampling the current best action,
+with $VOI_{\alpha}$ being the (approximate) value for sampling the current best action,
 and $VOI_i$ is the (approximate) value for sampling some other action $i$.
@@
 leaving this issue for future research.
-Unfortunately, this scheme appears too complicated to be amenable to a formal
-analysis. Early experiments with this approach demonstrate a significantly lower simple regret.
+Although this scheme appears too complicated to be amenable to a formal
+analysis, early experiments (Section \ref{sec:emp})
+with this approach demonstrate a significantly lower simple regret.
 
@@
     \includegraphics[scale=0.8]{onelevel-tree.pdf}\\
-    a. search tree, the best arm shaded
+    a. search tree, with best arm shaded
     \vspace{1em}
@@
 2$-greedy+UCT or UCB$_{\sqrt{\cdot}}$+UCT
-gives the lowest regret, UCB$_{\sqrt{\cdot}}$+UCT dominates UCT everywhere
-except for small numbers of instances. The advantage of both $\frac 1
+result in the lowest regret, UCB$_{\sqrt{\cdot}}$+UCT dominates UCT everywhere
+except when the number of samples is small. The advantage of both $\frac 1
 2$-greedy+UCT and UCB$_{\sqrt{\cdot}}$+UCT grows with the number of arms.
@@
     \includegraphics[scale=0.7]{twolevel-tree.pdf}\\
-    a. search tree, the path to the best arm shaded
+    a. search tree, with path to the best arm shaded
     \vspace{1em}
@@
    b. $10\times 10$ lake
-  \caption{The sailing domain, 397 samples, cost vs. factor}
+  \caption{The sailing domain, 397 rollouts, cost vs. factor}
   \label{fig:sailing-lake-size}
@@
 
-\section{Summary and Future Work}
+\section{Conclusion and Future Work}
 \label{sec:summary}
@@
 The MCTS SR+CR scheme presented in the paper differs
-from UCT mainly in the first step of the rollout, when the `simple' selection
-regret is minimized instead of the cumulative regret. Both the
+from UCT mainly in the first step of the rollout, when we attempt to minimize
+the `simple' selection regret rather than the cumulative regret. Both the
 theoretical analysis and the empirical evaluation provide evidence for
 better general performance of the proposed scheme.
-
-MCTS and UCT are particularly successful in Computer
-Go \cite{GellyWang.mogo,Braudis.pachi,Enzenberger.Fuego}
-and the proposed scheme should be evaluated on this domain, a
-non-trivial task as Go programs typically use ``non-pure''
-versions of UCT, extended with domain-specific knowledge.
 
@@
 
+Finally, applying VOI methods in complex deployed applications that already use
+MCTS is a challenge that should be addressed in future work.
+In particular, UCT is extremely successful in Computer
+Go \cite{GellyWang.mogo,Braudis.pachi,Enzenberger.Fuego},
+and the proposed scheme should be evaluated on this domain. This is non-trivial, since
+Go programs typically use ``non-pure'' versions of UCT, extended with
+domain-specific knowledge. For example, Pachi \cite{Braudis.pachi}
+typically re-uses information from rollouts generated for earlier
+moves, thereby violating our underlying assumption that information is
+only used for selecting the current move.  In early experiments not
+shown here (disallowing re-use of samples, admittedly not really a
+fair comparison) the VOI-aware scheme apears to dominate UCT.
+Nevertheless, it should also be possible to adapt the VOI-aware
+schemes to take into account expected re-use of samples, another topic
+for future research.
+
 \section*{Acknowledgments}

### 2012-05-29 paper/srcr.socs/paper.tex
trimming in progress

--- 2012-04-24 paper/srcr/paper.tex
@@
 \begin{abstract}
-UCT, a state-of-the art algorithm for Monte Carlo tree search (MCTS)
-in games and Markov decision processes, is based on UCB, a sampling
-policy for the Multi-armed Bandit problem (MAB) that 
+UCT, a state-of-the art algorithm for Monte Carlo tree search (MCTS),
+is based on UCB, a policy for the Multi-armed Bandit problem (MAB) that 
 minimizes the cumulative regret.  However, search differs from MAB in
-that in MCTS it is usually only the final ``arm pull'' (the actual
-move selection) that collects a reward, rather than all ``arm pulls''.
-Therefore, it makes more sense to minimize the simple regret, as
-opposed to the cumulative regret. We begin by introducing policies for
-multi-armed bandits with lower finite-time and asymptotic simple
-regret than UCB, using it to develop a two-stage scheme (SR+CR) for MCTS
-which outperforms UCT empirically.
-
-Optimizing the sampling process is itself a
-metareasoning problem, a solution of which can use value of
-information (VOI) techniques.  Although the theory of VOI for search
-exists, applying it to MCTS is non-trivial, as typical myopic
-assumptions fail. Lacking a complete working VOI theory for MCTS, we
-nevertheless propose a sampling scheme that is ``aware'' of VOI,
-achieving an algorithm that in empirical evaluation outperforms 
+that in MCTS it is usually only the final ``arm pull''
+that collects a reward, rather than all ``arm pulls''.
+Therefore, it makes more sense to minimize the simple, rather than
+cumulative, regret. We  introduce policies for
+multi-armed bandits with lower simple
+regret than UCB and develop a two-stage scheme (SR+CR) for MCTS
+which outperforms UCT empirically. Optimizing the sampling process is a
+metareasoning problem. However, applying the theory of VOI for search
+to MCTS is non-trivial. We nevertheless propose a sampling scheme
+that is ``aware'' of VOI, achieving an algorithm that outperforms 
 both UCT and the other proposed algorithms.
@@
 
-We begin with background definitions and related work. 
-Some sampling schemes are introduced, and shown to have better
-bounds for the simple regret on sets than UCB, the first contribution of this paper. 
-The results are applied to sampling in trees by combining the proposed sampling schemes on the
-first step of a rollout with UCT for the rest of the
-rollout. An additional sampling scheme based on metareasoning 
-principles is also suggested, another contribution
-of this paper.
-Finally, the performance of the
-proposed sampling schemes is evaluated on sets of Bernoulli arms, in randomly
-generated 2-level trees, and on the sailing domain, showing where the proposed schemes have improved
-performance.
-
 \section{Background and Related Work}
@@
 
-Monte-Carlo tree search was initially suggested as a scheme for
-finding approximately optimal policies for Markov Decision Processes
-(MDP).  An MDP is defined by the set of states $S$, the set of actions
-$A$ (also called {\em moves} in this paper),
-the transition distribution $T(s, a, s')$, the reward function $R(s, a,
-s')$, the initial state $s$ and an optional goal state $t$: $(S, A, T, R, s,
-t)$ \cite{Russell.aima}.  Several MCTS schemes explore an MDP by
-performing \emph{rollouts}---trajectories from the current state to a
-state in which a termination condition is satisfied (either the goal
-state, or a cutoff state for which the reward is evaluated
-approximately).
-
-\subsection{Multi-armed bandits and UCT}
-
-In the Multi-armed Bandit problem \cite{Vermorel.bandits} we have a set
+In the {\bf Multi-armed Bandit problem} \cite{Vermorel.bandits} we have a set
 of $K$ arms (see Figure~\ref{fig:mab-simple-regret}.a). Each arm can
@@
 \begin{equation}
-b_i=\overline X_i+\sqrt {\frac {c \log (n)} {n_i}}
+b_i=\overline X_i+\sqrt {{c \log (n)} / {n_i}}
 \label{eqn:ucb}
@@
 
-The UCT algorithm, an extension of UCB
+The {\bf UCT algorithm,} an extension of UCB
 to Monte-Carlo Tree Search is described in \cite{Kocsis.uct}, and
@@
 
-In the simple regret (selection) setting, the agent gets to
-collect only the reward of the last pull.
-
-\begin{dfn}
-The \textbf{simple regret} $\IE r$ of a sampling policy for the Multi-armed Bandit
-Problem is the expected difference between the best true expected reward
-$\mu_*$ and the true expected reward $\mu_j$ of the arm with the greatest sample mean,
-$j =\arg \max_i\overline X_i$:
-\begin{equation}
-\IE r=\sum_{j=1}^K\Delta_j\Pr(j=\arg\max_i\overline X_i)
-\label{eqn:simple-regret}
-\end{equation}
-where $\Delta_j=\mu_*-\mu_j$.
-\end{dfn}
-
-Strategies that minimize the simple regret are called pure exploration
+In the {\bf simple regret (selection) setting}, the agent gets to
+collect only the reward of the last pull. Strategies that minimize the
+simple regret are called pure exploration
 strategies \cite{Bubeck.pure}. An upper bound on the simple regret of uniform sampling is
@@
 
-%Pure exploration in
-%Multi-armed bandits is examined in \cite{Bubeck.pure}. On the one
-%hand, the \cite{Bubeck.pure} prove certain upper and lower bounds for
-%UCB and uniform sampling, showing that an upper bound on the simple
-%regret is exponential in the number of samples for uniform sampling,
-%while only polynomial for UCB. On the other hand, empirical
-%performance of UCB appears to be better than that of uniform sampling.
-
-\subsection{Metareasoning}
-
 A completely different scheme for control of sampling can use the
-principles of bounded rationality \cite{Horvitz.reasoningabout}
-and metareasoning ---  \cite{Russell.right} provided a formal
-description of rational metareasoning and case studies of applications
-in several problem domains. In search, under myopic and sub-tree independence assumptions,
-one maintains a current best move  $\alpha $ at the root, 
-and finds the expected gain from finding another move 
-$\beta $ to be better than the current best \cite{Russell.right}. The ``cost'' of search
-actions can also be factored in.
-Ideally, an ``optimal'' sampling scheme, to be used
-for selecting what to sample, both at the root node \cite{HayRussell.MCTS} and elsewhere,
-can be developed using metareasoning.
-However, this task is daunting for the following reasons:
-\begin{itemize}
-\item The method is in general intractable, necessitating simplifying assumptions.
- However, using the standard metareasoning myopic assumption,
-    where samples would be selected as though at most one sample
-   can be taken before an action is chosen, we run into serious problems. Even the basic
-  selection problem \cite{TolpinShimony.blinkered} exhibits a
-  non-concave utility function and results in premature stopping of the
-  standard myopic algorithms. This is due to the fact that the value of information of
-  a single measurement (analogous to a sample in MCTS) is frequently
-  less than its time-cost, even though this is not true for multiple
-  measurements.
-
-When applying the selection problem to MCTS, the
-situation is exacerbated.  The utility of an action is usually
-bounded, and thus in many cases a single sample may be insufficient
-to change the current best action, \emph{regardless} of its
-outcome. As a result, we frequently get a \emph{zero} ``myopic''
-value of information for a single sample.
-\item Rational metareasoning requires a known distribution model, which may be
-  difficult to obtain.
-\item Defining the time-cost of a sample is not trivial.
-\end{itemize}
-
-As the above ultimate goal is extremely difficult
-to achieve, we introduce in this paper
-simple schemes more amenable to analysis, loosely based on the metareasoning
-concept of value of information, and compare
-them to UCB (on sets) and UCT (in trees).
+principles of {\bf bounded rationality} \cite{Horvitz.reasoningabout} and
+{\bf rational metareasoning} \cite{Russell.right}. In search, under myopic
+and sub-tree independence assumptions, one maintains a current best
+move $\alpha $ at the root, and finds the expected gain from finding
+another move $\beta $ to be better than the current best
+\cite{Russell.right}. The ``cost'' of search actions can also be
+factored in.  Ideally, an ``optimal'' sampling scheme, to be used for
+selecting what to sample, both at the root node \cite{HayRussell.MCTS}
+and elsewhere, can be developed using metareasoning.  However, this
+ultimate goal is extremely difficult to achieve. We introduce in this
+paper simple schemes more amenable to analysis, loosely based on the
+metareasoning concept of value of information, and compare them to UCB
+(on sets) and UCT (in trees).
 
@@
 
-We examine two sampling schemes with super-polynomially
+We examine some sampling schemes with super-polynomially
 decreasing upper bounds on the simple regret. The bounds
@@
 
-\begin{proof}[Proof outline:]
-
-Bound the probability $P_i$ that a non-optimal arm $i$ is selected. Split the interval
- $[\mu_i, \mu_*]$ at $\mu_i+\delta_i$. Apply the Chernoff-Hoeffding bound to get:
-\begin{eqnarray}
-P_i&\le&\Pr[\overline X_i>\mu_i+\delta_i]+\Pr[\overline X_*<\mu_*-(\Delta_i-\delta_i)]\nonumber\\
-   &\le&\exp\left(-2\delta_i^2n_i\right)+\exp\left(-2(\Delta_i-\delta_i)^2n_*\right)
-\end{eqnarray}
-Observe that, in probability, $\overline X_i \rightarrow \mu_i$ as $n\rightarrow\infty$, 
-therefore $n_*\rightarrow n\varepsilon$, $n_i\rightarrow\frac
-{n(1-\varepsilon)} {K-1}$ as $n\rightarrow \infty$. Conclude that for
-every $0<\eta<1$, $\gamma>1$ there exists $N$ such that for every $n>N$ and
-all non-optimal arms $i$:
-\begin{equation}
-P_i \le \gamma \left(\exp\left(\frac {-2\delta_i^2n(1-\varepsilon)}{K-1}\right)
-+\exp\left(-2(\Delta_i-\delta_i)^2n\varepsilon\right)\right)
-\label{eqn:epsgreedy-probbound}
-\end{equation}
-Require
-\begin{eqnarray}
-\exp\left(-\frac {2\delta_i^2n(1-\varepsilon)}{K-1}\right)
-&=&\exp\left(-2(\Delta_i-\delta_i)^2n\varepsilon\right)\nonumber\\
-\frac {\delta_i} {\Delta_i-\delta_i}&=&\sqrt{\frac {(K-1)\varepsilon} {1-\varepsilon}}
-\label{eqn:epsgreedy-constant-varepsilon}
-\end{eqnarray}
-Substitute (\ref{eqn:epsgreedy-probbound}) together with
-(\ref{eqn:epsgreedy-constant-varepsilon}) into
-(\ref{eqn:simple-regret}) and obtain
-\begin{equation}
-\IE r_{\varepsilon\mbox{-}greedy}\le 2\gamma \sum_{i=1}^K\Delta_i\exp\left(\frac {-2\Delta_i^2n\varepsilon}
-  {\left(1+\sqrt{\frac {(K-1)\varepsilon}
-        {1-\varepsilon}}\right)^2}\right)
-\end{equation}
-\end{proof}
-
 In particular, as the number of arms $K$ grows, the bound for $\frac 1
@@
 \IE r_{\frac 1 2\mbox{-}greedy}&\le& 2\gamma \sum_{i=1}^K\Delta_i\exp\left(\frac {-2\Delta_i^2n}
-  {\left(1+\sqrt{K-1}\right)^2}\right)\\
-  &\approx& 2\gamma \sum_{i=1}^K\Delta_i\exp\left(\frac
-    {-2\Delta_i^2n} {K}\right)\mbox{ for }K\gg 1\nonumber
+  {\left(1+\sqrt{K-1}\right)^2}\right)
 \end{eqnarray}
@@
 \begin{equation}
-b_i=\overline X_i+\sqrt {\frac {c \sqrt n} {n_i}}
+b_i=\overline X_i+\sqrt {{c \sqrt n} / {n_i}}
 \end{equation}
@@
 \IE r_{ucb\sqrt{\cdot}} \le
-2\gamma\sum_{i=1}^K\Delta_i\exp\left(-\frac {c\sqrt{n}} 2\right)
+2\gamma\sum_{i=1}^K\Delta_i\exp\left(-{c\sqrt{n}}/2\right)
 \end{equation}
@@
 
-\begin{proof}[Proof outline:] Bound the probability $P_i$ that a
-  non-optimal arm $i$ is chosen. Split the interval $[\mu_i, \mu_*]$
-  at $\mu_i+\frac {\Delta_i} 2$. Apply the Chernoff-Hoeffding bound to get:
-\begin{eqnarray}
-P_i&\le&Pr\left[\overline X_i>\mu_i+\frac {\Delta_i} 2\right]+\Pr\left[\overline X_*<\mu_*-\frac {\Delta_i}
-  2\right]\nonumber \\
-   &\le&\exp\left(-\frac {\Delta_i^2n_i} 2\right)+\exp\left(-\frac {\Delta_i^2n_*} 2\right)
-\end{eqnarray}
-Observe that, in probability, $n_i\to \frac {c \sqrt n}
-{\Delta_i^2}$, $n_i\le n_*$ as $n\to\infty$. Conclude that for
-every $0<\eta<1$, $\gamma>1$ there exists $N$ such that for every $n>N$ and
-all non-optimal arms $i$:
-\begin{equation}
-P_i \le 2\gamma \exp\left(-\frac {c \sqrt n} 2\right)
-\label{eqn:usbsqrt-probbound}
-\end{equation}
-Substitute (\ref{eqn:usbsqrt-probbound}) into
-(\ref{eqn:simple-regret}) and obtain
-\begin{equation}
-\IE r_{ucb\sqrt{\cdot}} \le 2\gamma\sum_{i=1}^K\Delta_i\exp\left(-\frac {c\sqrt{n}} 2\right)
-\end{equation}
-\end{proof}
-
 \subsection{Sampling in Trees}
@@
 
-As mentioned above,
-UCT \cite{Kocsis.uct} is an extension of UCB for MCTS, that
-applies UCB($c$) at each step of a rollout.  At the root node,
-the sampling in MCTS is usually aimed at finding the first move to perform. Search is re-started,
-either from scratch or using some previously collected information,
-after observing the actual outcome (in MDPs) or the opponent's move
-(in adversarial games). Once one move is shown to be the best
-choice with high confidence, the value of information of additional
-samples of the best move (or, in fact, of any other samples) is low. Therefore, one should be able to do
-better than UCT by optimizing {\em simple regret}, rather than {\em cumulative regret}, at the root node.
-
-Nodes deeper in the search tree are a different matter.
-In order to support an optimal move choice at the root, it is beneficial in many cases
-to find a more precise estimate of the {\em value} of the state in these search
-tree nodes. For these internal nodes, optimizing simple regret is not the answer, and 
-cumulative regret optimization is not so far off the mark. Lacking a complete metareasoning for sampling,
-which would indicate the optimal way to sample both root nodes and internal nodes,
-our suggested improvement to UCT thus combines different sampling schemes on the first step and
-during the rest of each rollout:
+As mentioned above, UCT \cite{Kocsis.uct} is an extension of UCB for
+MCTS, that applies UCB($c$) at each step of a rollout.  At the root
+node, the sampling in MCTS is usually aimed at finding the first move
+to perform. Search is re-started, either from scratch or using some
+previously collected information, after observing the actual outcome
+(in MDPs) or the opponent's move (in adversarial games). Once one move
+is shown to be the best choice with high confidence, the value of
+information of additional samples of the best move (or, in fact, of
+any other samples) is low. Therefore, one should be able to do better
+than UCT by optimizing {\em simple regret}, rather than {\em
+cumulative regret}, at the root node.
+However, for thee internal nodes optimizing simple regret is not the
+answer, and cumulative regret optimization is not so far off the
+mark. Our suggested improvement to UCT thus combines different sampling
+schemes on the first step and during the rest of each rollout:
 \begin{dfn}
@@
 \end{dfn}
-
-The pseudocode of this two-stage rollout for an undiscounted MDP is in
-Algorithm~\ref{alg:two-stage-mcts}: \textsc{FirstAction} selects the first
-step of a rollout (line~\ref{alg:srcr-first-action}), and
-\textsc{NextAction} (line~\ref{alg:srcr-next-action}) selects steps during
-the rest of the rollout (usually using UCB). The reward statistic for
-the selected action is updated (line~\ref{alg:srcr-update-stat}), and
-the sample reward is back-propagated (line~\ref{alg:return-reward}) 
-towards the current root.
-
 We denote such two-step realizations of SR+CR as \emph{Alg}+UCT, where
 \emph{Alg} is the sampling scheme employed at the first step of a
-rollout (e.g. $\frac 1 2$-greedy+UCT).
-
-\begin{algorithm}[h!]
-\caption{Two-stage Monte-Carlo tree search sampling}
-\label{alg:two-stage-mcts}
-\begin{algorithmic}[1]
-\Procedure{Rollout}{node, depth=1}
-  \If {\Call{IsLeaf}{node, depth}}
-    \State \textbf{return} 0
-  \Else
-    \State \textbf{if} depth=1 \textbf{then} action $\gets$ \Call{FirstAction}{node} \label{alg:srcr-first-action}   
-    \State \textbf{else} action $\gets$ \Call{NextAction}{node} \label{alg:srcr-next-action}
-    \State next-node $\gets$ \Call{NextState}{node, action}
-    \State reward $\gets$ \Call{Reward}{node, action, next-node}
-     \State \hspace{4em} + \Call{Rollout}{next-node, depth+1}
-    \State \Call{UpdateStats}{node, action, reward} \label{alg:srcr-update-stat}
-    \State \textbf{return} reward \label{alg:return-reward}
-  \EndIf
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
-
-We expect such two-stage sampling schemes to
+rollout (e.g. $\frac 1 2$-greedy+UCT). We expect such two-stage sampling schemes to
 outperform UCT and be
@@
 and $VOI_i$ is the (approximate) value for sampling some other action $i$.
-
-These equations were derived as follows. The gain from switching from the current best action $\alpha$ to another
-action can be bounded:  by  the current expectation of the value the current second-best action
-for the case where we sample only $\alpha$, and by 1 (the maximum reward) minus the current expectation
-of $\alpha $ when sampling any other action. 
-The probability that another action be found
-best can be bounded by an exponential function of the difference in expectations when the true value
-of the actions becomes known. But the effect of each individual sample on the sample mean
-is inversely proportional to the current number of samples, hence the current number of samples (plus one
-in order to handle the initial case of no previous samples) in the denominator.
 

### 2012-05-29 paper/srcr.socs/paper.tex
3 pages

@@
 \author {David Tolpin, Solomon Eyal Shimony \\
-Department of Computer Science, \\
-Ben-Gurion University of the Negev, Beer Sheva, Israel \\
+Department of Computer Science, Ben-Gurion University of the Negev, Israel \\
 \{tolpin,shimony\}@cs.bgu.ac.il}
@@
 regret than UCB and develop a two-stage scheme (SR+CR) for MCTS
-which outperforms UCT empirically. Optimizing the sampling process is a
-metareasoning problem. However, applying the theory of VOI for search
-to MCTS is non-trivial. We nevertheless propose a sampling scheme
-that is ``aware'' of VOI, achieving an algorithm that outperforms 
-both UCT and the other proposed algorithms.
+which outperforms UCT empirically. In addition, we propose a sampling
+scheme based on value of information (VOI), achieving an algorithm
+that empirically outperforms other proposed algorithms.
 \end{abstract}
@@
 uncertainty, the goal is typically to either find a good (or optimal)
-strategy, or even just to find the best first action of such a policy. Once
-such an action is discovered, it is usually not beneficial to further sample
-that action, ``exploitation'' is thus meaningless for search
-problems. Finding a good first action is closer to the pure
+strategy, or even just to find the best first action of such a
+policy. Finding a good first action is closer to the pure
 exploration variant, as seen in the selection problem
@@
 In the {\bf Multi-armed Bandit problem} \cite{Vermorel.bandits} we have a set
-of $K$ arms (see Figure~\ref{fig:mab-simple-regret}.a). Each arm can
+of $K$ arms. Each arm can
 be pulled multiple times. Sometimes a cost is
@@
 is pulled, a random reward $X_i$ from an unknown stationary
-distribution is encountered.  The reward is usually bounded between 0 and 1.
+distribution is encountered.
 In the cumulative setting (the focus of much of the research literature on Multi-armed bandits),
@@
 In the {\bf simple regret (selection) setting}, the agent gets to
-collect only the reward of the last pull. Strategies that minimize the
-simple regret are called pure exploration
-strategies \cite{Bubeck.pure}. An upper bound on the simple regret of uniform sampling is
-exponentially decreasing in the number of samples (see
-\cite{Bubeck.pure}, Proposition~1). For UCB($c$) the best known respective upper bound on the simple
-regret of UCB($c$) is only polynomially decreasing in the number of samples
-(see \cite{Bubeck.pure}, Theorems~2,3). However, empirically
-UCB($c$) appears to yield a lower simple regret than uniform
-sampling.
+collect only the reward of the last pull. An upper bound on the simple
+regret of uniform sampling is exponentially decreasing in the number
+of samples (see \cite{Bubeck.pure}, Proposition~1). For UCB($c$) the
+best known respective upper bound on the simple regret of UCB($c$) is
+only polynomially decreasing in the number of samples (see
+\cite{Bubeck.pure}, Theorems~2,3). However, empirically UCB($c$)
+appears to yield a lower simple regret than uniform sampling.
 
 A completely different scheme for control of sampling can use the
-principles of {\bf bounded rationality} \cite{Horvitz.reasoningabout} and
-{\bf rational metareasoning} \cite{Russell.right}. In search, under myopic
-and sub-tree independence assumptions, one maintains a current best
-move $\alpha $ at the root, and finds the expected gain from finding
-another move $\beta $ to be better than the current best
-\cite{Russell.right}. The ``cost'' of search actions can also be
-factored in.  Ideally, an ``optimal'' sampling scheme, to be used for
-selecting what to sample, both at the root node \cite{HayRussell.MCTS}
-and elsewhere, can be developed using metareasoning.  However, this
-ultimate goal is extremely difficult to achieve. We introduce in this
-paper simple schemes more amenable to analysis, loosely based on the
-metareasoning concept of value of information, and compare them to UCB
-(on sets) and UCT (in trees).
+principles of {\bf bounded rationality} \cite{Horvitz.reasoningabout}
+and {\bf rational metareasoning} \cite{Russell.right}. In search, one
+maintains a current best move $\alpha $ at the root, and finds the
+expected gain from finding another move $\beta $ to be better than the
+current best. We introduce in this paper simple
+schemes loosely based on the metareasoning concept of value of
+information, and compare them to UCB (on sets) and UCT (in trees).
 
@@
 than uniform sampling; indeed, this is confirmed
-by experiments.
-
-We first consider $\varepsilon$-greedy sampling as a straightforward
+by experiments. First, we consider $\varepsilon$-greedy sampling as a straightforward
 generalization of uniform sampling:
@@
 \end{dfn}
-This sampling scheme exhibits an exponentially decreasing simple regret:
-\begin{thm} For every $0<\eta<1$  and $\gamma>1$ there exists $N$ such that for
-  any number of samples $n>N$ the simple regret of the  $\varepsilon$-greedy
-sampling scheme is  bounded from above as
-\begin{equation}
-\IE r_{\varepsilon\mbox{-}greedy}\le 2\gamma \sum_{i=1}^K\Delta_i\exp\left(\frac {-2\Delta_j^2n\varepsilon}
-  {\left(1+\sqrt{\frac {(K-1)\varepsilon}
-        {1-\varepsilon}}\right)^2}\right)
-\end{equation}
-with probability at least $1-\eta$.
-\end{thm}
-
+This sampling scheme exhibits an exponentially decreasing simple
+regret.
 In particular, as the number of arms $K$ grows, the bound for $\frac 1
@@
 random sampling ($\varepsilon=\frac 1 K$):
-\begin{crl}
+\begin{thm}
 For uniform random sampling, 
@@
 \end{eqnarray}
-\end{crl}
+\end{thm}
 
-$\varepsilon$-greedy is based solely on sampling the
-arm that has the greatest sample mean (henceforth called the ``current best'' arm)
-with a higher probability then the rest of the arms, and ignores
-information about sample means of other arms. On the other hand,
-UCB distributes samples in accordance with sample means, but, in order to
-minimize cumulative regret, chooses the current best arm too often.
-Intuitively, a better scheme for simple regret minimization would
-distribute samples in a way similar to UCB, but would sample the current best arm
+Another sampling scheme for simple regret minimization can be obtained
+by distributing samples in a way similar to UCB, but sampling the current best arm
 less often. This can be achieved by replacing $\log(\cdot)$ in
-Equation~\ref{eqn:ucb} with a faster growing sublinear function, for
-example, $\sqrt\cdot$.
+Equation~\ref{eqn:ucb} with a faster growing function, e.g. $\sqrt\cdot$.
 \begin{dfn} Scheme $\mathbf{UCB_{\sqrt{\cdot}}(c)}$ pulls arm $i$ that
-maximizes $b_i$, where:
-\begin{equation}
-b_i=\overline X_i+\sqrt {{c \sqrt n} / {n_i}}
-\end{equation}
-where, as before, $\overline X_i$ is the average reward obtained from arm $i$,
-$n_i$ is the number of times arm $i$ was pulled, and $n$ is the total
-number of pulls so far. \end{dfn}
-This scheme also exhibits a super-polynomially decreasing simple regret:
-\begin{thm}  For every $0<\eta<1$  and $\gamma>1$ there exists $N$ such that for
-  any number of samples $n>N$ the simple regret of the  UCB$_{\sqrt{\cdot}}$($c$)
-sampling scheme is  bounded from above as
-\begin{equation}
-\IE r_{ucb\sqrt{\cdot}} \le
-2\gamma\sum_{i=1}^K\Delta_i\exp\left(-{c\sqrt{n}}/2\right)
-\end{equation}
-with probability at least $1-\eta$.
-\end{thm}
+maximizes $b_i$, where $b_i=\overline X_i+\sqrt {{c \sqrt n} / {n_i}}$.
+\end{dfn}
+This scheme also exhibits a super-polynomially decreasing simple regret.
 
@@
 cumulative regret}, at the root node.
+
 However, for thee internal nodes optimizing simple regret is not the
 answer, and cumulative regret optimization is not so far off the
-mark. Our suggested improvement to UCT thus combines different sampling
-schemes on the first step and during the rest of each rollout:
-\begin{dfn}
-The \textbf{SR+CR MCTS sampling scheme} selects an action at the
-current root node according to a scheme suitable for minimizing 
-the simple regret (\textbf{SR}), such as $\frac 1 2$-greedy or UCB$_{\sqrt{\cdot}}$, and (at non-root nodes)
-then selects actions according to UCB, which approximately minimizes the cumulative regret (\textbf{CR}).
-\end{dfn}
-We denote such two-step realizations of SR+CR as \emph{Alg}+UCT, where
-\emph{Alg} is the sampling scheme employed at the first step of a
-rollout (e.g. $\frac 1 2$-greedy+UCT). We expect such two-stage sampling schemes to
-outperform UCT and be
-significantly less sensitive to the tuning of the exploration factor
-$c$ of UCB($c$). That is since the contradiction between the need for a larger
-value of $c$ on the first step (simple regret) and a smaller
-value for the rest of the rollout (cumulative regret)
-\cite{Bubeck.pure} is resolved. In fact, a sampling scheme that uses
-UCB($c$) at all steps but a larger value of $c$ for the first step
-than for the rest of the steps, should also outperform UCT.
+mark. Our suggested improvement to UCT is the \textbf{SR+CR MCTS
+sampling scheme} that combines different sampling schemes on the first
+step and during the rest of each rollout.  We expect such two-stage
+sampling schemes to outperform UCT and be significantly less sensitive
+to the tuning of the exploration factor $c$ of UCB($c$).
 
@@
 value of information (VOI) of the rollouts and choosing rollouts that
-maximize the VOI. 
-However, as indicated above, actually computing the VOI is infeasible.
-Instead we suggest the following scheme based on the following
-features of value of information:
-\begin{enumerate}
-\item An estimate of the probability that one or more rollouts will make another action
-appear better than the current best $\alpha $.
-\item An estimate of the gain that may be incurred if such a change occurs.
-\end{enumerate}
-
-If the distribution of results generated by the rollouts were known, the
-above features could be easily computed. However, this is not the case for
-most MCTS applications. Therefore, we estimate bounds on the feature values from
-the current set of samples, based on the myopic assumption that the algorithm will only
-sample one of the actions, and use these bounds as the feature values, to get:
+maximize the VOI. We estimate the VOI from
+the current set of samples:
 \begin{eqnarray}
@@
 \end{eqnarray}
-with $VOI_{\alpha}$ being the (approximate) value for sampling the current best action,
-and $VOI_i$ is the (approximate) value for sampling some other action $i$.
-
-These VOI estimates are used in the ``VOI-aware'' sampling scheme as follows: sample the
-action that has maximum estimated VOI. We judged these estimates to be too crude
-to be used as ``stopping criteria'' that can be used to cut off sampling,
-leaving this issue for future research.
-Although this scheme appears too complicated to be amenable to a formal
-analysis, early experiments (Section \ref{sec:emp})
-with this approach demonstrate a significantly lower simple regret.
+The ``VOI-aware'' scheme samples the action that has the maximum
+estimated VOI.  Although this scheme appears too complicated for
+a formal analysis, early experiments demonstrate a significantly lower
+simple regret.
 
@@
 
-\subsection{Simple regret in multi-armed bandits}
-\label{sec:emp-mab}
-
-Figure~\ref{fig:mab-simple-regret} presents a comparison of MCTS sampling
-schemes on Multi-armed bandits. Figure~\ref{fig:mab-simple-regret}.a shows the search tree
-corresponding to a problem instance. Each arm returns a random reward
-drawn from a Bernoulli distribution. The search selects an arm
-and compares the expected reward, unknown to the algorithm during the
-sampling, to the expected reward of the best arm.
-
-Figure~\ref{fig:mab-simple-regret}.b shows the regret
-vs. the number of samples, averaged over $10000$ experiments for
-randomly generated instances of 32 arms. Either $\frac 1
-2$-greedy  or UCB$_{\sqrt{\cdot}}$ dominate UCB over the
-whole range. For larger number of samples, the advantage
-of UCB$_{\sqrt{\cdot}}$ over $\frac 1 2$-greedy becomes more significant.
-
-\begin{figure}[h!]
-  \begin{minipage}[c]{1.0\linewidth}
-    \centering
-    \includegraphics[scale=0.8]{onelevel-tree.pdf}\\
-    a. search tree, with best arm shaded
-    \vspace{1em}
-  \end{minipage}
-  \begin{minipage}[c]{1.0\linewidth}
-    \centering
-    \includegraphics[scale=0.45]{flat-trilevel-k=64-uqb=8.pdf}\\
-    b. regret vs. number of samples
-  \end{minipage}
-  \caption{Simple regret in MAB}
-  \label{fig:mab-simple-regret}
-\end{figure}
-
-\subsection{Monte Carlo tree search}
-\label{sec:emp-mcts}
-
-The second set of experiments was performed on randomly generated
-2-level max-max trees crafted so as to deliberately deceive uniform sampling
-(Figure~\ref{fig:mcts-regret}.a), necessitating  an adaptive sampling scheme, 
-such as UCT. That is due to the switch nodes, each with 2 children with anti-symmetric
-values, which would cause a uniform sampling scheme to incorrectly give them all a value of 0.5.
-
-Simple regret vs. the number of samples are shown for trees with root degree 16
-(Figure~\ref{fig:mcts-regret}.b) and 64
-(Figure~\ref{fig:mcts-regret}.c). The exploration factor $c$ is set to
-$2$, the default value for rewards in the range $[0, 1]$.
-The algorithms exhibit a similar relative performance: either $\frac 1
+Figure~\ref{fig:mcts-regret} shows
+results for experiments on randomly generated trees. VOI+UCT, the scheme based on a VOI estimate,
+outperforms all other sampling schemes in this example; either $\frac 1
 2$-greedy+UCT or UCB$_{\sqrt{\cdot}}$+UCT
-result in the lowest regret, UCB$_{\sqrt{\cdot}}$+UCT dominates UCT everywhere
+result in the second lowest regret, UCB$_{\sqrt{\cdot}}$+UCT dominates UCT everywhere
 except when the number of samples is small. The advantage of both $\frac 1
 2$-greedy+UCT and UCB$_{\sqrt{\cdot}}$+UCT grows with the number of arms.
-
 \begin{figure}[h!]
-  \begin{minipage}[c]{1.0\linewidth}
   \centering
-    \includegraphics[scale=0.7]{twolevel-tree.pdf}\\
-    a. search tree, with path to the best arm shaded
-    \vspace{1em}
-  \end{minipage}
-  \begin{minipage}[c]{1.0\linewidth}
-    \centering
-    \includegraphics[scale=0.45]{tree-identity-k=16-uqb=8.pdf}\\ 
-    b. 16 arms
-    \vspace{0.5em}
-  \end{minipage}
-  \begin{minipage}[c]{1.0\linewidth}
-    \centering
-    \includegraphics[scale=0.45]{tree-identity-k=64-uqb=8.pdf} \\
-    c. 64 arms
-  \end{minipage}
-  \caption{MCTS in random trees}
+  \includegraphics[scale=0.5]{tree-identity-k=32-uqb=8+voi.pdf}\\
+  \caption{MCTS in random trees.}
   \label{fig:mcts-regret} 
@@
 
-\subsection{The sailing domain}
-\label{sec:emp-sailing}
-
-Figures~\ref{fig:sailing-cost-vs-nsamples}--\ref{fig:sailing-lake-size}
-show results of experiments on the sailing
-domain. Figure~\ref{fig:sailing-cost-vs-nsamples} shows the regret
-vs. the number of samples, computed for a range of values of
-$c$. Figure~\ref{fig:sailing-cost-vs-nsamples}.a shows the median
-cost, and Figure~\ref{fig:sailing-cost-vs-nsamples}.b --- the minimum
-costs. UCT is always worse than either $\frac 1 2$-greedy+UCT or
-UCB$_{\sqrt{\cdot}}$+UCT, and is sensitive to the value of $c$: the median cost is
-much higher than the minimum cost for UCT. For both $\frac 1 2$-greedy+UCT
-and UCB$_{\sqrt{\cdot}}$+UCT, the difference is significantly less prominent.
-
-\begin{figure}[h!]
-  \begin{minipage}[b]{1.0\linewidth}
-    \centering
-    \includegraphics[scale=0.45]{costs-size=6-group=median.pdf}\\
-    a. median cost
-    \vspace{1em}
-  \end{minipage}
-  \begin{minipage}[b]{1.0\linewidth}
-    \centering
-    \includegraphics[scale=0.45]{costs-size=6-group=minimum.pdf}\\
-    b. minimum cost
-  \end{minipage}
-  \caption{The sailing domain, $6\times 6$ lake, cost vs. samples}
-  \label{fig:sailing-cost-vs-nsamples}
-\end{figure}
-
-Figure~\ref{fig:sailing-cost-vs-factor} shows the regret vs. the
-exploration factor for different numbers of samples. UCB$_{\sqrt{\cdot}}$+UCT is always better than
-UCT, and $\frac 1 2$-greedy+UCT is better than UCT expect for a small range of
-values of the exploration factor.
+Figure~\ref{fig:sailing-cost-vs-nsamples}
+shows results of experiments on the sailing
+domain. UCT is always worse than either $\frac 1 2$-greedy+UCT or
+UCB$_{\sqrt{\cdot}}$+UCT.
 
@@
   \centering
-  \includegraphics[scale=0.45]{costs-size=6-nsamples=397.pdf}\\
-  a. 397 rollouts\\
-  \vspace{1em}
-  \includegraphics[scale=0.45]{costs-size=6-nsamples=1585.pdf}\\
-  b. 1585 rollouts
-  \caption{The sailing domain, $6\times 6$ lake, cost vs. factor}
-  \label{fig:sailing-cost-vs-factor}
-\end{figure}
-
-Figure~\ref{fig:sailing-lake-size} shows the cost vs. the exploration
-factor for lakes of different sizes. The relative difference between
-the sampling schemes becomes more prominent when the lake size
-increases.
-\begin{figure}[h!]
-   \centering
-   \includegraphics[scale=0.45]{costs-size=6-nsamples=397.pdf}\\
-   a. $6\times 6$ lake \\
-   \vspace{1em}
-   \includegraphics[scale=0.45]{costs-size=10-nsamples=397.pdf}\\
-   b. $10\times 10$ lake
-  \caption{The sailing domain, 397 rollouts, cost vs. factor}
-  \label{fig:sailing-lake-size}
-\end{figure}
-
-\subsection{VOI-aware MCTS}
-
-Finally, the VOI-aware sampling scheme was empirically compared to
-other sampling schemes (UCT, $\frac 1 2$-greedy+UCT,
-UCT$_{\sqrt{\cdot}}$+UCT). Again, the experiments were performed on
-randomly generated trees with structure shown in
-Figure~\ref{fig:mcts-regret}.a. Figure~\ref{fig:mcts-regret-voi} shows
-the results for 32 arms. VOI+UCT, the scheme based on a VOI estimate,
-outperforms all other sampling schemes in this example. Similar performance improvements
-(not shown) also occur for the sailing domain.
-\begin{figure}[h!]
-  \centering
-  \includegraphics[scale=0.45]{tree-identity-k=32-uqb=8+voi.pdf}\\
-  \caption{MCTS in random trees, including VOI+UCT.}
-  \label{fig:mcts-regret-voi} 
+  \includegraphics[scale=0.45]{costs-size=6-group=median.pdf}\\
+  \caption{The sailing domain, $6\times 6$ lake, cost vs. samples}
+  \label{fig:sailing-cost-vs-nsamples}
 \end{figure}

### 2012-05-29 paper/srcr.socs/paper.tex
squeezed into 2 pages

@@
 regret than UCB and develop a two-stage scheme (SR+CR) for MCTS
-which outperforms UCT empirically. In addition, we propose a sampling
+which outperforms UCT empirically. We also propose a sampling
 scheme based on value of information (VOI), achieving an algorithm
@@
 
-Monte-Carlo tree search, and especially a version based on the
-UCT formula \cite{Kocsis.uct} appears in numerous search applications,
-such as \cite{GellyWang.mogo,Eyerich.ctp}. Although these methods are shown to be successful empirically,
-most authors appear to be using the UCT formula ``because it has been shown
-to be successful in the past'', and ``because it does a good job of
-trading off exploration and exploitation''. While the latter statement may be
-correct for the multi-armed bandit and for the UCB method \cite{Auer.ucb},
-we argue that it is inappropriate for search. The problem is not that
-UCT does not work; rather, a simple reconsideration from basic
-principles can result in schemes that outperform UCT.
+Monte-Carlo tree search, and especially a version based on the UCT
+formula \cite{Kocsis.uct} appears in numerous search applications,
+such as \cite{GellyWang.mogo,Eyerich.ctp}. Although these methods are
+shown to be successful empirically, we argue that a simple
+reconsideration from basic principles can result in schemes that
+outperform UCT.
 
-The core issue is that in adversarial search
-and search in ``games against nature'' --- optimizing behavior under
-uncertainty, the goal is typically to either find a good (or optimal)
-strategy, or even just to find the best first action of such a
-policy. Finding a good first action is closer to the pure
+In adversarial search and optimization under uncertainty, the goal is
+typically to either find a good strategy, or even just to
+find the best first action of such a strategy, which is closer to the pure
 exploration variant, as seen in the selection problem
@@
 it is much better to minimize the \emph{simple} regret.  However, the
-simple and the cumulative regret cannot be minimized simultaneously;
-moreover, \cite{Bubeck.pure} shows that in many cases the smaller the
-cumulative regret, the greater the simple regret.
-
-\section{Background and Related Work}
-\label{sec:related-work}
+simple and the cumulative regret cannot be minimized
+simultaneously \cite{Bubeck.pure}.
 
@@
 of $K$ arms. Each arm can
-be pulled multiple times. Sometimes a cost is
-associated with each pulling action. When the $i$th arm
+be pulled multiple times. When the $i$th arm
 is pulled, a random reward $X_i$ from an unknown stationary
@@
 all encountered rewards are collected by the agent. 
-The UCB scheme was shown to be near-optimal in this respect \cite{Auer.ucb}:
-
-\begin{dfn} Scheme $\mathbf{UCB(c)}$ pulls arm $i$ that maximizes 
-upper confidence bound $b_i$ on the reward:
+In this respect, the $\mathbf{UCB(c)}$ scheme that pulls arm $i$ maximizing
+upper confidence bound $b_i$ on the reward 
 \begin{equation}
@@
 \end{equation}
-where $\overline X_i$ is the average sample reward obtained from arm $i$,
-$n_i$ is the number of times arm $i$ was pulled, and $n$ is the total
-number of pulls so far. \end{dfn}
+is near-optimal \cite{Auer.ucb}.
 
 The {\bf UCT algorithm,} an extension of UCB
-to Monte-Carlo Tree Search is described in \cite{Kocsis.uct}, and
+to Monte-Carlo Tree Search \cite{Kocsis.uct} is
 shown to outperform many state of the art search algorithms in both
 MDP and adversarial games \cite{Eyerich.ctp,GellyWang.mogo}. 
-
-In the {\bf simple regret (selection) setting}, the agent gets to
-collect only the reward of the last pull. An upper bound on the simple
-regret of uniform sampling is exponentially decreasing in the number
-of samples (see \cite{Bubeck.pure}, Proposition~1). For UCB($c$) the
-best known respective upper bound on the simple regret of UCB($c$) is
-only polynomially decreasing in the number of samples (see
-\cite{Bubeck.pure}, Theorems~2,3). However, empirically UCB($c$)
-appears to yield a lower simple regret than uniform sampling.
+However, the best known upper bound on the simple regret of UCB is
+only polynomially decreasing in the number of samples \cite{Bubeck.pure}.
 
@@
 principles of {\bf bounded rationality} \cite{Horvitz.reasoningabout}
-and {\bf rational metareasoning} \cite{Russell.right}. In search, one
-maintains a current best move $\alpha $ at the root, and finds the
-expected gain from finding another move $\beta $ to be better than the
-current best. We introduce in this paper simple
+and {\bf rational metareasoning} \cite{Russell.right}. We introduce in this paper simple
 schemes loosely based on the metareasoning concept of value of
-information, and compare them to UCB (on sets) and UCT (in trees).
+information (VOI), and compare them to UCB (on sets) and UCT (in trees).
 
@@
 than uniform sampling; indeed, this is confirmed
-by experiments. First, we consider $\varepsilon$-greedy sampling as a straightforward
-generalization of uniform sampling:
-\begin{dfn} The \textbf{$\mathbf{\varepsilon}$-greedy} sampling scheme
+by experiments. First, we consider $\varepsilon$-greedy sampling, which
 pulls the arm that currently has the greatst sample mean, with probability
-$0<\varepsilon<1$, and any other
-arm with probability $\frac {1-\varepsilon} {K-1}$. 
-\end{dfn}
+$0<\varepsilon<1$, and any other arm with probability $\frac {1-\varepsilon} {K-1}$. 
 This sampling scheme exhibits an exponentially decreasing simple
-regret.
-In particular, as the number of arms $K$ grows, the bound for $\frac 1
+regret. Moreover, as the number of arms $K$ grows, the bound for $\frac 1
 2$-greedy sampling ($\varepsilon=\frac 1 2$) becomes considerably tighter than for uniform
-random sampling ($\varepsilon=\frac 1 K$):
-\begin{thm}
-For uniform random sampling, 
-\begin{equation}
-\IE r_{uniform}\le 2\gamma \sum_{i=1}^K\Delta_i\exp\left(-\frac {\Delta_i^2n} {K}\right)
-\end{equation}
-For $\frac 1 2$-greedy sampling,
-\begin{eqnarray}
-\IE r_{\frac 1 2\mbox{-}greedy}&\le& 2\gamma \sum_{i=1}^K\Delta_i\exp\left(\frac {-2\Delta_i^2n}
-  {\left(1+\sqrt{K-1}\right)^2}\right)
-\end{eqnarray}
-\end{thm}
+random sampling ($\varepsilon=\frac 1 K$).
 
@@
 less often. This can be achieved by replacing $\log(\cdot)$ in
-Equation~\ref{eqn:ucb} with a faster growing function, e.g. $\sqrt\cdot$.
-\begin{dfn} Scheme $\mathbf{UCB_{\sqrt{\cdot}}(c)}$ pulls arm $i$ that
-maximizes $b_i$, where $b_i=\overline X_i+\sqrt {{c \sqrt n} / {n_i}}$.
-\end{dfn}
-This scheme also exhibits a super-polynomially decreasing simple regret.
+Equation~\ref{eqn:ucb} with $\sqrt\cdot$. This scheme also exhibits
+a super-polynomially decreasing simple regret.
 
@@
 previously collected information, after observing the actual outcome
-(in MDPs) or the opponent's move (in adversarial games). Once one move
-is shown to be the best choice with high confidence, the value of
-information of additional samples of the best move (or, in fact, of
-any other samples) is low. Therefore, one should be able to do better
-than UCT by optimizing {\em simple regret}, rather than {\em
-cumulative regret}, at the root node.
+(in MDPs) or the opponent's move (in adversarial games). Therefore,
+one should be able to do better than UCT by optimizing {\em simple}
+rather than {\em cumulative regret}, at the root node.
 
@@
 step and during the rest of each rollout.  We expect such two-stage
-sampling schemes to outperform UCT and be significantly less sensitive
-to the tuning of the exploration factor $c$ of UCB($c$).
+sampling schemes to outperform UCT.
 
@@
 Further improvement can be achieved by computing or estimating the
-value of information (VOI) of the rollouts and choosing rollouts that
+VOI of the rollouts and choosing rollouts that
 maximize the VOI. We estimate the VOI from
-the current set of samples:
-\begin{eqnarray}
-VOI_\alpha&\approx&\frac {\overline X_\beta} {n_\alpha+1}
-\exp\left(-2(\overline X_\alpha - \overline X_\beta)^2 n_\alpha\right)\\
-VOI_i&\approx&\frac {1-\overline X_\alpha} {n_i+1}
-\exp\left(-2(\overline X_\alpha - \overline X_i)^2 n_i\right),\; i\ne\alpha\nonumber\\
-\mbox{where }&&\alpha=\arg\max_i \overline X_i,\quad
-             \beta=\arg\max_{i,\,i\ne\alpha} \overline X_i\nonumber
-\end{eqnarray}
+the current set of samples using distribution-independent concentration inequalities.
+%\begin{eqnarray}
+%VOI_\alpha&\approx& \frac {\overline X_\beta} {n_\alpha+1}
+%\exp\left(-2(\overline X_\alpha - \overline X_\beta)^2 n_\alpha\right)\\
+%VOI_i&\approx& \frac {1-\overline X_\alpha} {n_i+1}
+%\exp\left(-2(\overline X_\alpha - \overline X_i)^2 n_i\right),\; i\ne\alpha\nonumber\\
+%\mbox{where }&&\alpha=\arg\max_i \overline X_i,\quad
+%             \beta=\arg\max_{i,\,i\ne\alpha} \overline X_i\nonumber
+%\end{eqnarray}
 The ``VOI-aware'' scheme samples the action that has the maximum
@@
 The results were empirically verified on Multi-armed Bandit instances,
-on search trees, and on the sailing domain, as defined in
+on search trees, and on the sailing domain
 \cite{Kocsis.uct}. In most cases, the experiments showed a lower average
@@
 
-Figure~\ref{fig:sailing-cost-vs-nsamples}
-shows results of experiments on the sailing
-domain. UCT is always worse than either $\frac 1 2$-greedy+UCT or
-UCB$_{\sqrt{\cdot}}$+UCT.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[scale=0.45]{costs-size=6-group=median.pdf}\\
-  \caption{The sailing domain, $6\times 6$ lake, cost vs. samples}
-  \label{fig:sailing-cost-vs-nsamples}
-\end{figure}
+%Figure~\ref{fig:sailing-cost-vs-nsamples}
+%shows results of experiments on the sailing
+%domain. UCT is always worse than either $\frac 1 2$-greedy+UCT or
+%UCB$_{\sqrt{\cdot}}$+UCT.
+%
+%\begin{figure}[h!]
+%  \centering
+%  \includegraphics[scale=0.45]{costs-size=6-group=median.pdf}\\
+%  \caption{The sailing domain, $6\times 6$ lake, cost vs. samples}
+%  \label{fig:sailing-cost-vs-nsamples}
+%\end{figure}
 
@@
 for finding good actions in both MDPs and adversarial games.
-Further improvement of the sampling scheme is thus of interest in
-numerous search applications. We argue that although UCT is already very efficient,
-one can do better if the sampling scheme is considered from a metareasoning
-perspective of value of information (VOI).
+Both the theoretical analysis and the empirical evaluation provide evidence for
+better general performance of the proposed SR+CR scheme.
 
-The MCTS SR+CR scheme presented in the paper differs
-from UCT mainly in the first step of the rollout, when we attempt to minimize
-the `simple' selection regret rather than the cumulative regret. Both the
-theoretical analysis and the empirical evaluation provide evidence for
-better general performance of the proposed scheme.
+Using VOI to control sampling ab-initio should provide further
+improvement. However, application of the theory of rational
+metareasoning to MCTS is an open problem, and both a solid theoretical
+model and empirically efficient VOI estimates need to be developed.
 
-Although SR+CR is inspired by the notion of VOI,
-the VOI is used there implicitly in the analysis of the algorithm,
-rather than computed or learned explicitly in order to plan the
-rollouts. Ideally, using VOI to control sampling ab-initio should do even better,
-but the theory for doing that is still not up to speed. Instead we suggest
-a  ``VOI-aware'' sampling scheme based on crude probability and value estimates,
-which despite its simplicity already shows a marked improvement in minimizing regret.
-However, application of the theory of rational metareasoning
-to Monte Carlo Tree Search is an open problem \cite{HayRussell.MCTS},
-and both a solid theoretical model and empirically efficient VOI
-estimates need to be developed.
-
-Finding a better sampling scheme for non-root nodes,
-as well as the root node, should also be possible.
-Although cumulative regret does reasonably
-well there, it is far from optimal, as meta-reasoning principles imply that an optimal scheme
-for these nodes must be asymmetrical (e.g. it is not helpful to find out that the
-value of the current best action is even better than previously believed).
-
-Finally, applying VOI methods in complex deployed applications that already use
-MCTS is a challenge that should be addressed in future work.
-In particular, UCT is extremely successful in Computer
-Go \cite{GellyWang.mogo,Braudis.pachi,Enzenberger.Fuego},
-and the proposed scheme should be evaluated on this domain. This is non-trivial, since
-Go programs typically use ``non-pure'' versions of UCT, extended with
-domain-specific knowledge. For example, Pachi \cite{Braudis.pachi}
-typically re-uses information from rollouts generated for earlier
-moves, thereby violating our underlying assumption that information is
-only used for selecting the current move.  In early experiments not
-shown here (disallowing re-use of samples, admittedly not really a
-fair comparison) the VOI-aware scheme apears to dominate UCT.
-Nevertheless, it should also be possible to adapt the VOI-aware
-schemes to take into account expected re-use of samples, another topic
-for future research.
+Applying VOI methods in complex deployed applications that
+already use MCTS should be addressed.  In particular, UCT is extremely
+successful in Computer Go
+\cite{GellyWang.mogo,Braudis.pachi,Enzenberger.Fuego}, and the
+proposed scheme should be evaluated on this domain. This is
+non-trivial, since Go programs typically use ``non-pure'' versions of
+UCT.
 

### 2012-06-05 paper/srcr.socs/paper.tex
added ucb root
socs final

@@
 
-\title{MCTS Based on Simple Regret}
+\title{MCTS Based on Simple Regret$^*$}
 
@@
 \maketitle
-
-\begin{abstract}
-UCT, a state-of-the art algorithm for Monte Carlo tree search (MCTS),
-is based on UCB, a policy for the Multi-armed Bandit problem (MAB) that 
-minimizes the cumulative regret.  However, search differs from MAB in
-that in MCTS it is usually only the final ``arm pull''
-that collects a reward, rather than all ``arm pulls''.
-Therefore, it makes more sense to minimize the simple, rather than
-cumulative, regret. We  introduce policies for
-multi-armed bandits with lower simple
-regret than UCB and develop a two-stage scheme (SR+CR) for MCTS
-which outperforms UCT empirically. We also propose a sampling
-scheme based on value of information (VOI), achieving an algorithm
-that empirically outperforms other proposed algorithms.
-\end{abstract}
+\insert\footins{\vspace{\baselineskip} \noindent\footnotesize Extended abstract of AAAI 2012 paper.}
 
@@
 typically to either find a good strategy, or even just to
-find the best first action of such a strategy, which is closer to the pure
-exploration variant, as seen in the selection problem
-\cite{Bubeck.pure,TolpinShimony.blinkered}. In the selection problem,
+find the best first action of such a strategy. This is closer to the pure
+exploration variant \cite{Bubeck.pure} as seen in the selection problem
+\cite{TolpinShimony.blinkered}. In the selection problem,
 it is much better to minimize the \emph{simple} regret.  However, the
@@
 is near-optimal \cite{Auer.ucb}.
+
+In the simple regret (selection) setting, the agent gets to collect
+only the reward of the last pull. The \textbf{simple regret} $\IE r$
+of a sampling policy for the Multi-armed Bandit Problem is the
+expected difference between the best true expected reward $\mu_*$ and
+the true expected reward $\mu_j$ of the arm with the greatest sample
+mean, $j =\arg \max_i\overline X_i$:
+\begin{equation}
+\IE r=\sum_{j=1}^K\Delta_j\Pr(j=\arg\max_i\overline X_i)
+\label{eqn:simple-regret}
+\end{equation}
+where $\Delta_j=\mu_*-\mu_j$.
 
@@
 
-We examine some sampling schemes with super-polynomially
-decreasing upper bounds on the simple regret. The bounds
-suggest that these schemes achieve a lower simple regret
-than uniform sampling; indeed, this is confirmed
-by experiments. First, we consider $\varepsilon$-greedy sampling, which
+First, we consider $\varepsilon$-greedy sampling, which
 pulls the arm that currently has the greatst sample mean, with probability
@@
 less often. This can be achieved by replacing $\log(\cdot)$ in
-Equation~\ref{eqn:ucb} with $\sqrt\cdot$. This scheme also exhibits
+Equation~\ref{eqn:ucb} with $\sqrt\cdot$. This scheme (UCB$_{\sqrt{\cdot}}$) also exhibits
 a super-polynomially decreasing simple regret.
@@
 mark. Our suggested improvement to UCT is the \textbf{SR+CR MCTS
-sampling scheme} that combines different sampling schemes on the first
-step and during the rest of each rollout.  We expect such two-stage
-sampling schemes to outperform UCT.
+  sampling scheme}. The scheme selects an action at the current
+root node according to a scheme suitable for minimizing the simple
+regret (\textbf{SR}), such as $\frac 1 2$-greedy or
+UCB$_{\sqrt{\cdot}}$, and then (at non-root nodes) selects actions
+according to UCB, which approximately minimizes the cumulative regret
+(\textbf{CR}).  We expect such two-stage sampling scheme to outperform
+UCT.
 
@@
 Further improvement can be achieved by computing or estimating the
-VOI of the rollouts and choosing rollouts that
+value of information (VOI) of the rollouts and choosing rollouts that
 maximize the VOI. We estimate the VOI from

### 2012-06-10 paper/vago.ecai/case-for-support.tex
case for support

\documentclass[11pt]{letter}
\usepackage{eurosym}

\signature{David Tolpin}

\begin{document}

\begin{letter}{}
\opening{Dear ECCAI Travel Grant Officer,}

I, David Tolpin, am a full-time PhD student at the Department of
Computer Science at Ben-Gurion University of the Negev, Israel. For the last
three years, I have been researching {\it metareasoning in
  problem-solving search}, a field of Artificial Intelligence seeking
systematic approaches to development of heuristic algorithms for hard
problems. My and my advisor's joint paper, {\bf VOI-aware MCTS},
presenting a result of the ongoing result, has been accepted to
ECAI~2012 for poster presentation.

The paper addresses an important issue of simple regret minimization
in Monte Carlo sampling in trees, and presenting the paper at the
conference and getting feedback from other researchers,
as well as attending other presentations in this and other fields of
AI is important for my ongoing studies in AI. However, travel,
logding, and registration costs are high, and without financial 
support I will have difficulty attending the conference.

The anticipated expenses are :

\begin{itemize}
\item registration --- \euro{120}
\item accomodation --- \euro{180} (student room)
\item airplane tickets --- \euro{680} (current lowest available rate
  from the current travel agent)
\end{itemize}

{\bf Total: \euro{980}}

As the expenses exceed the ECCAI travel grant, the Computer Science
department agreed to provide additional funds (see the letter from the 
department).

\closing{Sincerely Yours,}

\end{letter}
\end{document}

### 2012-06-10 paper/vago.ecai/case-for-support.tex
fixes

@@
 problems. My and my advisor's joint paper, {\bf VOI-aware MCTS},
-presenting a result of the ongoing result, has been accepted to
+presenting a result of ongoing research, has been accepted to
 ECAI~2012 for poster presentation.
@@
 \item airplane tickets --- \euro{680} (current lowest available rate
-  from the current travel agent)
+  from the local travel agent)
 \end{itemize}

### 2012-06-10 paper/vago.ecai/case-for-support.tex
fixes

@@
 as well as attending other presentations in this and other fields of
-AI is important for my ongoing studies in AI. However, travel,
-logding, and registration costs are high, and without financial 
+AI is important for my studies in AI. However, travel,
+lodging, and registration costs are high, and without financial 
 support I will have difficulty attending the conference.

### 2012-06-26 paper/vago.uai/paper.tex
current state

@@
 \newcommand {\maps} {\colon}
-\newcommand {\given} {\mathrel{|}}
+\newcommand {\given} {\mid} %{\mathrel{|}}
 \newcommand {\sample} {\sim}
 \newcommand {\iid} {\stackrel{\rm iid}{\sample}}
+\newcommand {\Nat} {\mathbb{N}}
 \DeclareMathOperator*{\argmax}{argmax}
@@
 
+\excludecomment{hiddenproof}
+
 \title{Selecting Computations: Theory and Applications}
 
-\author{}
-%\author{ {\bf Nicholas Hay} and {\bf Stuart Russell} \\   
-%Computer Science Division \\  
-%University of California\\ 
-%Berkeley, CA 94720 \\ 
-%\And 
-%{\bf Solomon Eyal Shimony} and {\bf David Tolpin}  \\ 
-%Department of Computer Science \\ 
-%Ben-Gurion University of the Negev\\
-%Beer Sheva, Israel
-%}
+%\author{}
+\author{ {\bf Nicholas Hay} and {\bf Stuart Russell} \\   
+Computer Science Division \\  
+University of California\\ 
+Berkeley, CA 94720 \\ 
+\And 
+{\bf Solomon Eyal Shimony} and {\bf David Tolpin}  \\ 
+Department of Computer Science \\ 
+Ben-Gurion University of the Negev\\
+Beer Sheva, Israel
+}
 
@@
 
-\begin{abstract}
-Sequential decision problems are often approximately solvable by
-simulating possible future action sequences.  {\em Metalevel} decision
-procedures have been developed for selecting {\em which} action
-sequences to simulate, based on some estimate of the expected
-improvement in decision quality resulting from any particular
-simulation; an example is the recent work on using selection rules
-derived for bandit problems to control Monte Carlo tree search in the
-game of Go.  In this paper we develop a theoretical basis for
-metalevel decisions in the statistical framework of {\em selection problems}, 
-arguing that this is more appropriate than the bandit
-framework.  We derive a number of basic results applicable to simple
-Monte Carlo selection problems, including the first finite sampling
-bounds for optimal policies in certain cases; we also provide a simple
-counterexample to the intuitive conjecture that an optimal policy will
-necessarily reach a decision in all cases.  We then derive heuristic
-approximations in both Bayesian and distribution-free settings and
-demonstrate their superiority to bandit-based heuristics in one-shot
-decision problems and in Go.
-\end{abstract}
+\input{abstract}
 
@@
 
-\section{Upper bounds on value of information}\label{approx-nonbayesian-section}
+\section{Upper bounds on Value of Information}\label{approx-nonbayesian-section}
 
@@
 
+\begin{comment}
 	Theorem \ref{thm:optimal-myopic} provides necessary conditions
@@
 	The myopic selection policy uses
-the intrinsic value of information (VOI) $\Lambda_i$ of testing an arm $i$, which is
+	the intrinsic value of information (VOI) $\Lambda_i$ of testing an arm~$i$, which is
 	the expected reward of the best arm given the information obtained by the test,
@@
 	as if we ``had our blinkers on''.	
+\end{comment}
 
@@
 In many practical applications of the selection problem, such as search in
-the game of Go, prior distributions are unavailable\footnote{The analysis is also applicable to
+the game of Go, prior distributions are unavailable.\footnote{The analysis is also applicable to
 some Bayesian settings, using ``fake" samples to simulate prior distributions.}
 In such cases, one can still bound
-the value of information w.r.t. myopic policies using {\em concentration
+the value of information of myopic policies using {\em concentration
 inequalities} to derive distribution-independent bounds on the
@@
 
-When considering possible samples in the ``blinkered'' semi-myopic setting,
-two cases are possible:
-\begin{itemize}
-\item the arm $\alpha$ with the highest sample mean $\overline
+When considering possible samples in the blinkered semi-myopic setting,
+two cases are possible: either
+	the arm $\alpha$ with the highest sample mean $\overline
   	X_\alpha$ is tested, and $\overline X_\alpha$ becomes lower than
  	$\overline X_\beta$ of the second-best arm $\beta$;
-\item another arm $i$ is tested, and $\overline X_i$ becomes higher
+or, 
+	another arm~$i$ is tested, and $\overline X_i$ becomes higher
     than $\overline X_\alpha$.
-\end{itemize}
 
-Our bounds below are applicable to any bounded distribution (with rewards being
-bounded to $[0,1]$). Similar
+Our bounds below are applicable to any bounded distribution (without loss of generality 
+bounded in $[0,1]$). Similar
 bounds can be derived for certain unbounded distributions, such as the
-normally distributed prior of the value with normally distributed
+normally distributed prior value with normally distributed
 sampling.
-We derive a VOI bound for testing an arm $N$ times,
-where $N$ can be the remaining ``budget'' of available samples or
+We derive a VOI bound for testing an arm a fixed $N$ times,
+where $N$ can be the remaining budget of available samples or
 any other integer quantity.
@@
 \end{thm}
+\begin{hiddenproof}
 	\vspace{-2em}
@@
 	\end{proof}		
+\end{hiddenproof}
 
@@
 Hoeffding inequality \citep{Hoeffding.ineq}:
-\begin{thm} The probabilities in equations (\ref{eqn:thm-be}) are bounded from above as
+\begin{thm} The probabilities in \eqref{eqn:thm-be} are bounded from above as
 \begin{align}
@@
 
-\begin{comment}
+\begin{hiddenproof}
 	\begin{proof}
-	Equations (\ref{eqn:probound-blnk-hoeffding}) follow from the
+	\eqref{eqn:probound-blnk-hoeffding}) follow from the
 	observation that if $i\ne\alpha$, $\overline X_i^{n_i+N}>\overline X_\alpha^{n_i}$
@@
 	Find $\delta$ for which the two terms on the right-hand side of
-	(\ref{eqn:app-hoeffding-le-maxexp}) are equal:
+	\eqref{eqn:app-hoeffding-le-maxexp} are equal:
 	\begin{equation}
@@
 	\end{equation}
-	Solve (\ref{eqn:app-hoeffding-eq-exp}) for $\delta$: $\delta=\frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}} (\overline X_\alpha^{n_\alpha}\hspace{-0.5em}
+	Solve \eqref{eqn:app-hoeffding-eq-exp} for $\delta$: $\delta=\frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}} (\overline X_\alpha^{n_\alpha}\hspace{-0.5em}
 	- \overline X_i^{n_i}) \ge 2(\sqrt 2 - 1)(\overline X_\alpha^{n_\alpha}\hspace{-0.5em}-\overline X_i^{n_i})$. Substitute $\delta$ into 
-	(\ref{eqn:app-hoeffding-le-maxexp}) and obtain
+	\eqref{eqn:app-hoeffding-le-maxexp} and obtain
 	\begin{align}
@@
 An upper bound on the VOI estimate $\Lambda_i^b$ is obtained
-by substituting (\ref{eqn:probound-blnk-hoeffding}) into (\ref{eqn:thm-be}).
+by substituting \eqref{eqn:probound-blnk-hoeffding} into \eqref{eqn:thm-be}.
 \begin{align}
@@
 More refined bounds can be obtained through tighter estimates on the
-probabilities in Equation \ref{eq:simplistic}, for example, based on the empirical Bernstein
+probabilities in \eqref{eqn:thm-be}, for example, based on the empirical Bernstein
 inequality~\citep{MaurerPontil.benrstein}, or through a more careful
@@
 
-\section{Practical sampling control: empirical evaluation}
-\label{sec:empirical-evaluation}\label{mcts-section}
-
-As the selection MDP is intractable, we use myopic or semi-myopic policies instead.
-The myopic policies usually separate out the decision of {\em whether} to
+Selection problems usually separate out the decision of {\em whether} to
 sample or to stop (called the stopping policy), and {\em what} to sample.
-We first examine the latter issue.
-
-\subsection{Selecting The Best Arm}
-\label{sec:emp-arm}
+We'll examine the first issue here, along with the empirical evaluation 
+of the above approximate algorithms, and the second in the following section.
 
@@
 When the distributions are unknown, it makes sense
-to use the upper bounds established in Theorem~\ref{thm:be} instead,
-which is the method adopted here.
+to use the upper bounds established in \thmref{thm:be}, as we do in the following.
 This evaluation assumes a fixed budget of samples, which is
@@
 \includegraphics[scale=0.6]{flat.pdf}
-\caption{Random instances: regret vs. number of samples}
+\caption{Average regret of various policies as a function of the fixed number 
+of samples in a 25-action Bernoulli sampling problem, over 10000 trials.}
 \label{fig:random-instances}
@@
 
-The sampling policies are compared on random Multi-armed bandit 
-selection problem instances. Figure~\ref{fig:random-instances} shows results for
-randomly-generated Multi-armed bandits with 25 Bernoulli arms, with
-the mean rewards of the arms distributed uniformly in the range~$[0,
-  1]$, for a range of sample budgets~$200..2000$, with multiplicative
-step of~$2$. The experiment for each number of samples was repeated
-10000 times. UCB1, as well as policies  based on bounds
-(\ref{eqn:bound-blnk-hoeffding}) (VOI) and
-(\ref{eqn:erf-blinkered}) (VOI+) are compared. UCB1 is always considerably worse than the
-VOI-aware sampling policies.
+The sampling policies are compared on random Bernoulli
+selection problem instances. \figref{fig:random-instances} shows results for
+randomly-generated selection problems with 25 Bernoulli arms, where
+the mean rewards of the arms are distributed uniformly in~$[0,1]$, 
+for a range of sample budgets~$200..2000$, with multiplicative
+step of~$2$, averaging over 10000 trials.  We compare UCB1 with the 
+policies based on the bounds in
+\eqref{eqn:bound-blnk-hoeffding} (VOI) and
+\eqref{eqn:erf-blinkered} (VOI+).
+UCB1 is always considerably worse than the VOI-aware sampling policies.
 
-\subsection{Sampling in trees}
+\section{Sampling in trees}
+\label{sec:empirical-evaluation}\label{mcts-section}
 
-Sampling in trees is considerably more complicated.
-The goal of Monte-Carlo tree search \citep{Chaslot.montecarlo} at the root node 
+The previous section addressed the selection problem in the flat case.
+Selection in trees is more complicated.  The goal of Monte-Carlo tree 
+search \citep{Chaslot.montecarlo} at the root node 
 is usually to select an action that appears to be the best based on outcomes
@@
 But the goal of rollouts at non-root nodes
-is different than at the root: here it is important to better approximate the ``true'' 
+is different than at the root: here it is important to better approximate the
 value of the node, so that selection at the root can be more informed. The exact analysis
@@
 We thus propose the following hybrid sampling scheme: 
-\begin{itemize}
-\item at the root node, sample based on the VOI estimate;
-\item at non-root nodes, sample using UCT.
-\end{itemize}
+	at the \emph{root node}, sample based on the VOI estimate;
+	at \emph{non-root nodes}, sample using UCT.
 
-Strictly speaking, even at the root node the assumptions underlying our belief-state
+Strictly speaking, even at the root node the stationarity assumptions\footnote{This is not a restriction,
+however, of the general formalism in \secref{sec:optimal}.} 
+underlying our belief-state
 MDP for selection do not hold exactly. UCT is an adaptive scheme, and therefore the values
-generated by sampling at non-root nodes will typically cause valued observed at
+generated by sampling at non-root nodes will typically cause values observed at
 children of the root node to be non-stationary. 
 Nevertheless, sampling based on VOI estimates
-computed as for stationary distributions work well in
+computed as for stationary distributions works well in
 practice. As illustrated
-by the empirical evaluation (Section~\ref{sec:empirical-evaluation}),
+by the empirical evaluation (\secref{sec:empirical-evaluation}),
 estimates based on upper bounds on the VOI result in good sampling
@@
 
-\subsubsection{Stopping criterion}
+\subsection{Stopping criterion}
 \label{sec:control-stopping-criterion}
@@
 information of a measurement, an upper bound on the intrinsic VOI can also
-be used to stop the sampling if the intrinsic VOI of any action
-is less than the cost of sampling $C$:
+be used to stop the sampling if the intrinsic VOI of any arm
+is less than the total cost of sampling $C$:
 \begin{equation}
@@
 \end{equation}
-VOI estimates~(\ref{eqn:thm-be},
-\ref{eqn:bound-blnk-hoeffding}) include the remaining sample budget $N$ as a
+The VOI estimates of \eqrefs{eqn:thm-be}{eqn:bound-blnk-hoeffding} 
+include the remaining sample budget $N$ as a
 factor, but given the cost of a single sample $c$, the cost of the
@@
 $C=cN$. $N$ can be dropped on both sides of the inequality,
-and a reasonable stopping criterion would be:
+giving a reasonable stopping criterion:
 \begin{align}
@@
 \end{align}
-The empirical evaluation (Section~\ref{sec:empirical-evaluation})
+The empirical evaluation (\secref{sec:empirical-evaluation})
 confirms the viability of this stopping criterion and illustrates the
@@
 the sampling policy. When the sample cost $c$ is unknown, one can perform initial calibration experiments
-to determine a reasonable value, as done in this paper.
+to determine a reasonable value, as done in the following.
 
-\subsubsection{Sample redistribution in trees}
+\subsection{Sample redistribution in trees}
 \label{sec:control-redistribution}
@@
 current state is discarded after an real-world action is selected. In practice,
-many successful Monte-Carlo tree search algorithms re-use rollouts
+many successful Monte-Carlo tree search algorithms reuse rollouts
 generated at earlier search states, if the sample traverses the
@@
 
-One way to account for this re-use would be to incorporate the
+One way to account for this reuse would be to incorporate the
 `future' value of information into a VOI estimate. However, this 
-approach requires a non-trivial extension of the theory of meta-reasoning for search.
-Alternately, one can behave myopically w.r.t. search tree depth:
+approach requires a nontrivial extension of the theory of metareasoning for search.
+Alternately, one can behave myopically with respect to the search tree depth:
 \begin{enumerate}
-\item estimate VOI as though the information is discarded after each step;
-\item stop early if the VOI is below a certain threshold
-   (see Section~\ref{sec:control-stopping-criterion});
-\item save the unused sample budget for search in future states, such that
+\item Estimate VOI as though the information is discarded after each step,
+\item Stop early if the VOI is below a certain threshold
+   (see \secref{sec:control-stopping-criterion}), and
+\item Save the unused sample budget for search in future states, such that
    if the nominal budget is $N$, and the unused budget in the last state
@@
 VOI of increasing the budget of a future state by one sample.  It is
-unclear whether this cost can be accurately estimated, but supposedly
+unclear whether this cost can be accurately estimated, but supposing
 a fixed value for a given problem type and algorithm implementation
-would work. Indeed, the empirical evaluation (Section~\ref{sec:emp-go})
+would work. Indeed, the empirical evaluation (\secref{sec:emp-go})
 confirms that stopping and sample redistribution based on a learned
@@
 
-\subsubsection{Playing Go against UCT}
+\subsection{Playing Go against UCT}
 \label{sec:emp-go}
 
-The hybrid policies were compared on Computer Go, a search domain
+The hybrid policies were compared on the game Go, a search domain
 in which UCT-based MCTS has been particularly successful
@@
   original UCT policy was left unchanged.
-\item The time-allocation mode based on the fixed number of samples
+\item The time-allocation model based on the fixed number of samples
   was modified for \textit{both the original UCT policy and the VOI-aware
@@
   \begin{itemize}
-    \item the same number of samples is available to
+    \item The same number of samples is available to
       the agent at each step, independently of the number of pre-simulated
       games;  
-    \item if samples were unused at the current step,
+    \item If samples were unused at the current step,
       they become available at the next step. 
@@
 \includegraphics[scale=0.6]{uctvoi.pdf}
-\caption{Go: winning rate vs. $c$}
+\caption{Winning rate of the VOI-aware policy in Go as a function of the cost $c$, for varying numbers of samples per ply.}
 \label{fig:uctvoi}
@@
 The engines were compared on the 9x9 board, for 5000, 7000, 1000, and
-15000 samples (game simulations) per ply, each experiment was repeated
-1000 times. Figure~\ref{fig:uctvoi} depicts a calibration experiment,
-showing the winning rate of the VOI-aware policy against UCT
-vs. the stopping threshold $c$ (if the maximum VOI of a sample is below
+15000 samples (game simulations) per ply, each experiment repeated
+1000 times. \figref{fig:uctvoi} depicts a calibration experiment,
+showing the winning rate of the VOI-aware policy against UCT as a function of
+the stopping threshold $c$ (if the maximum VOI of a sample is below
 the threshold, the simulation is stopped, and a move is chosen). Each
@@
 \includegraphics[scale=0.6]{voi-wins.pdf}
-\caption{Go: winning rate for $c=10^{-6}$}
+\caption{Winning rate of the VOI-aware policy in Go as a function of the number of samples, fixing cost $c=10^{-6}$.}
 \label{fig:voi-wins}
@@
 
-Figure~\ref{fig:voi-wins}
-shows the winning rate of VOI against UCT vs. the number of
-samples for $c=10^{-6}$. In agreement with the intuition
-(Section~\ref{sec:control-redistribution}), VOI-based stopping and
+\figref{fig:voi-wins}
+shows the winning rate of VOI against UCT $c=10^{-6}$. In agreement with the intuition
+(\figref{sec:control-redistribution}), VOI-based stopping and
 sample redistribution is most influential for intermediate numbers of
@@
 
-Note that if we disallowed re-use of samples in both Pachi and
+Note that if we disallowed reuse of samples in both Pachi and
 in our VOI-based scheme, the VOI based-scheme
-win rate is even higher than shown in Figure \ref{fig:voi-wins}. This is as expected,
+win rate is even higher than shown in \figref{fig:voi-wins}. This is as expected,
 as this setting (which is somewhat unfair to Pachi) is closer to
 meeting the assumptions underlying the selection MDP.
+
+\input{conclusion}
 

### 2012-06-26 paper/vago.uai/proof-supplement.tex
current state

\documentclass[]{article}
\usepackage{proceed2e}
\usepackage{enumerate}
\usepackage{algpseudocode}
\usepackage[ruled]{algorithm}
\usepackage{url}
\usepackage{framed}
\usepackage{amsfonts,amsmath,amsthm,amssymb}
\usepackage{graphicx}
\usepackage{url}
\usepackage{color}
\usepackage[round]{natbib}
\usepackage{comment}

\def\citeapos#1{\citeauthor{#1}'s (\citeyear{#1})}

\newcommand {\mean} {\ensuremath {\mathop{\mathrm{mean}}}}
\newcommand {\median} {\ensuremath {\mathop{\mathrm{median}}}}
\newcommand {\N} {\ensuremath {\mathcal{N}}}
\newcommand {\IE} {\ensuremath {\mathbb{E}}}
\newcommand {\cov} {\ensuremath {\mathop{\mathrm{cov}}}}
\newcommand {\BEL} {\ensuremath {\mathop{\mathrm{BEL}}}}

\newcommand {\term}[1] {\textbf{#1}}
\newcommand {\note}[1] {\textcolor{red}{[[#1]]}}
\newcommand {\Evidence} {\mathcal{E}}
\newcommand {\R} {\mathbb{R}}
\newcommand {\maps} {\colon}
\newcommand {\given} {\mid} %{\mathrel{|}}
\newcommand {\sample} {\sim}
\newcommand {\iid} {\stackrel{\rm iid}{\sample}}
\newcommand {\Nat} {\mathbb{N}}
\DeclareMathOperator*{\argmax}{argmax}

\newtheorem{thm}{Theorem}
\newtheorem{dfn}[thm]{Definition}
\newtheorem{lmm}[thm]{Lemma}
\newtheorem{crl}[thm]{Corollary}
\newtheorem{example}{Example}

\newcommand{\secref}[1]{Section~\ref{#1}}
\newcommand{\secrefs}[2]{Sections~\ref{#1} and~\ref{#2}}
\newcommand{\figref}[1]{Figure~\ref{#1}}
\renewcommand{\eqref}[1]{Equation~(\ref{#1})}
\newcommand{\eqrefs}[2]{Equations~(\ref{#1}) and~(\ref{#2})}
\newcommand{\thmref}[1]{Theorem~\ref{#1}}
\newcommand{\dfnref}[1]{Definition~\ref{#1}}
\newcommand{\exampleref}[1]{Example~\ref{#1}}

\includecomment{hiddenproof}

\title{Selecting Computations: Supplemental proofs}

%\author{}
\author{ {\bf Nicholas Hay} and {\bf Stuart Russell} \\   
Computer Science Division \\  
University of California\\ 
Berkeley, CA 94720 \\ 
\And 
{\bf Solomon Eyal Shimony} and {\bf David Tolpin}  \\ 
Department of Computer Science \\ 
Ben-Gurion University of the Negev\\
Beer Sheva, Israel
}

\begin{document}

\maketitle

Included here are all the proofs of the results in the paper.
Definitions are included to maintain theorem numbering.

	\begin{dfn} \label{dfn:metalevel-model}
		A \term{metalevel probability model} is a tuple $(U_1,\dots,U_k,\Evidence)$ 
		consisting of jointly distributed random variables:
		\begin{itemize}
			\item Real random variables $U_1,\dots,U_k$, where $U_i$ is the utility of arm~$i$, and
			\item A countable set $\Evidence$ of random variables, each variable $E\in\Evidence$ being 
			      a computation that can be performed and whose value is the result of that computation.
		\end{itemize}
	\end{dfn}

	\begin{dfn}
	A (countable state, undiscounted) \term{Markov Decision Process} (MDP) is a tuple $M=(S,s_0,A_s,T,R)$ where:
		$S$ is a countable set of states,
		$s_0\in S$ is the fixed initial state,
		$A_s$ is a countable set of actions available in state $s\in S$,
		$T(s,a,s')$ is the transition probability from $s\in S$ to $s'\in S$ after performing action $a\in A_s$,
		and $R(s,a,s')$ is the expected reward received on such a transition.
	\end{dfn}

	\begin{dfn}\label{dfn:metalevel-mdp}
		Given a metalevel probability model%
			\footnote{Definition~\ref{dfn:metalevel-model} made no assumption about the computational result
				variables $E_i\in\Evidence$, but for simplicity in the following we'll assume that
				each $E_i$ takes one of a countable set of values.  Without loss of generality, 
				we'll further assume the domains of the computational variables $E\in\Evidence$ are disjoint.}
		$(U_1,\dots,U_k,\Evidence)$ and
		a cost of computation $c>0$, a corresponding \term{metalevel decision problem}
		is any MDP $M=(S,s_0,A_s,T,R)$ such that
		\begin{align*}
			S &= \{\bot\}\cup\{\langle e_1\dots, e_n \rangle : e_i\in E_i \text{ for all $i$,} \\
									& \qquad\qquad \text{for finite $n\ge0$ and distinct $E_i\in\Evidence$}\} \\
			s_0 &= \langle \rangle \\
			A_s &= \{\bot\}\cup\Evidence_s \\
		%
		\intertext{where $\bot\in S$ is the unique terminal state,
		where $\Evidence_s\subseteq\Evidence$ is a state-dependent subset of allowed computations,
		and when given any $s=\langle e_1, \dots, e_n \rangle\in S$,
		computational action $E\in\Evidence$, 
		and $s'= \langle e_1, \dots, e_n, e \rangle\in S$ where $e\in E$, we have:}
		%
			T(s,E,s') &= P(E=e \given E_1=e_1,\dots,E_n=e_n) \\
			T(s,\bot,\bot) &= 1 \\
			R(s,E,s') &= -c \\
			R(s,\bot,\bot) &= \max_i \mu_i(s) % \max_i \IE[U_i \given E_1=e_1,\dots,E_n=e_n]
		\end{align*}		
		where $\mu_i(s) = \IE[U_i \given E_1=e_1,\dots,E_n=e_n]$.
	\end{dfn}

	\begin{equation}
		V^\pi_M(s) = \IE^\pi_M\left[ \sum_{i=0}^N R(S_i,\pi(S_i),S_{i+1}) \given S_0 = s\right]	\label{eq:value-fn}\\		
	\end{equation}

	\begin{thm}\label{thm:value-of-computation}
		The value function of a metalevel decision process $M=(S,s_0,A_s,T,R)$ is of the form
		\[
			V^\pi_M(s) = \IE^\pi_M[ -c\,N + \max_i \mu_i(S_N) \given S_0=s]
		\]
		where $N$ denotes the (random) total number of computations performed;
		similarly for $Q^\pi_M(s,a)$.
	\end{thm}

	\begin{proof}
		Follows immediately from \eqref{eq:value-fn} and the definition of the
		reward function in \dfnref{dfn:metalevel-mdp}.
	\end{proof}

	\begin{thm}\label{thm:bounded-expected-computations}
		The optimal policy's expected number of computations is bounded by the 
		value of perfect information  times the inverse cost $1/c$:
		\[
			\IE^{\pi^*}[N\given S_0=s] \le \frac{1}{c} \left(\IE[\max_i U_i\given S_0=s] - \max_i \mu_i(s)\right).
		\]
		Further, any policy $\pi$ with infinite expected number of computations 
		%% $\IE^\pi[N]=\infty$
		has negative infinite value, hence the optimal
		policy stops with probability one.
	\end{thm}

	\begin{proof}
		The first follows as in state $s$ the optimal policy has value at least that
		of stopping immediately ($\max_i \mu_i(s)$), and as $\IE \max_i\mu_i(S_N) \le \IE \max_i U_i$ by Jensen's inequality.
		The second from \thmref{thm:value-of-computation}.
	\end{proof}

	\begin{dfn}\label{dfn:myopic}
		Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$,
		the \term{myopic policy} $\pi^m(s)$ is defined to equal $\argmax_{a\in A_s} Q^m(s,a)$ 
		where $Q^m(s,\bot) = \max_i \mu_i(s)$ and
		\begin{equation*}%% \label{eq:myopic}
			 Q^m(s,E) = \IE_M[ -c + \max_i \mu_i(S_1) \given S_0 = s, A_0 = E].		
		\end{equation*}
	\end{dfn}

	\begin{thm}\label{thm:optimal-myopic}
		Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$
		if the myopic policy performs some computation in state $s\in S$,
		then the optimal policy does too, i.e., if $\pi^m(s)\neq\bot$ then $\pi^*(s)\neq\bot$.
	\end{thm}

	\begin{hiddenproof}
		\begin{proof}
			Observe that the myopic Q-function \eqref{dfn:myopic} is equivalently given by
			\[
				Q^m(s,a) = Q^\bot(s,a)
			\]
			where $\bot$ is the policy which immediately stops $\bot(s)=\bot$.
			Thus $Q^m(s,a) \le Q^*(s,a)$.  If the optimal policy stops in a state $s\in S$ then
			\[
				Q^{\pi^*}(s,a) \le \max_i \mu_i(s),
			\]
			and so the same holds for $Q^m$, showing the myopic stops.
		\end{proof}
	\end{hiddenproof}

	\begin{dfn}\label{dfn:closed}
		Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$,
		a subset $S'\subseteq S$ of states is \term{closed under transitions}
		if whenever $s'\in S'$, $a\in A_{s'}$, $s''\in S$, and $T(s',a,s'')>0$,
		we have $s''\in S'$.
	\end{dfn}

	\begin{thm}\label{thm:myopic-optimal}
		Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$
		and a subset $S'\subseteq S$ of states closed under transitions,	
		if the myopic policy stops in all states $s'\in S'$
		then the optimal policy does too.	
	\end{thm}

	\begin{hiddenproof}
		\begin{proof}
		Take any $s^*\in S'$, and note that all states the chain can transition
		to from $s^*$ are also in $S'$, by transition closure.  Defining $m(s) = \max_i\mu_i(s)$, 
		observe the myopic stopping for all such states implies that
		\begin{align*}
			\IE^{\pi}[(m(S_{j+1}) - c)\, 1(j<N)\given S_0=s^*] \\
			\le \IE^{\pi}[m(S_{j})\, 1(j<N)\given S_0=s^*]
		\end{align*}
		holds for all $j$, and as a result:
		\begin{align*}
			V^\pi(s) 
			&= \IE^{\pi}[ - c N + m(S_N) \given S_0=s^*] \\
			&= \IE^{\pi}[m(S_0) + \sum_{j=0}^{N-1} (m(S_{j+1}) - c - m(S_j)) \given S_0=s^*] \\
		%%	&\le \IE^{\pi}[m(S_0) + \sum_{j=0}^{N-1} 0 |S_0=s] \\		
			&\le \max_i\mu_i(s^*) \qedhere
		\end{align*}
		\end{proof}	
	\end{hiddenproof}

	\begin{thm}\label{thm:one-action-bound}
		The one-armed Bernoulli decision process with constant arm $\lambda\in[0,1]$ 
		performs at most $\lambda(1-\lambda)/c-3 \le 1/4c-3$ computations.
	\end{thm}

	\begin{hiddenproof}
		\begin{proof}
			By \dfnref{dfn:myopic} and \exampleref{example:bernoulli2}, the myopic policy stops in
			a state $(s,f)$ when
			\begin{align}
				c \ge &\mu\max(\mu^+,m) + (1-\mu_i)\max(\mu^-, m) - \max(\mu,m) \label{eq:stopping}
			\end{align}
			where $\mu=(s+1)/(n+2)$  is the mean utility for arm~2, where $n=s+f$,
			$\mu^- = \mu - \mu/(n+3)$, and $\mu^+ = \mu + (1-\mu)/(n+3)$
			are the posterior means of arm~2 after simulating a failure and a success,
			%% and $\mu^-_i=(s_i+1)/(s_i+f_i+3) $ and $\mu^+_i = (s_i+2)/(s_i+f_i+3)$
			%% are the posterior means of action $i$ after simulating a failure and a success,
			respectively.  Whenever \eqref{eq:stopping} holds, stopping is preferred to sampling.

			Fixing $n$ and maximizing over $\mu$, we get sufficient condition for stopping
			\begin{equation}
				c \ge \frac{\lambda(1-\lambda)}{(n+3)} \qquad\qquad   n\ge \frac{\lambda(1-\lambda)}{c} - 3  \label{eq:myopic-stopping}
			\end{equation}
			Since the set of states satisfying \eqref{eq:myopic-stopping} is closed under
			transitions ($n$ only increases), by \thmref{thm:optimal-myopic}.  Finally, note $\max_{\lambda\in[0,1]} \lambda(1-\lambda)=1/4$.
		\end{proof}	
	\end{hiddenproof}

	\begin{dfn}\label{dfn:context}
		Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$, and a constant $\lambda\in\R$,
		define $M_\lambda = (S,s_0,A_s,T,R_\lambda)$ to be $M$ with an additional action of known 
		value $\lambda$, defined by:
		\begin{align*}
			R_\lambda(s,E,s')      &= R(s,E,s') \\
			R_\lambda(s,\bot,\bot) &= \max\{\lambda, R(s,\bot,\bot)\}
		\end{align*}
	\end{dfn}

	\begin{thm}
		Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$, 
		there exists a real interval $I(s)$ for every state $s\in S$ such that
		it is optimal to stop in state $s$ in $M_\mu$ iff $\mu\notin I(s)$.
		Furthermore, $I(s)$ contains $\max_i\mu_i(s)$ whenever it is nonempty.
	\end{thm}

	\begin{hiddenproof}
		\begin{proof}
		With $m(s) = \max_i\mu_i(s)$, the utility of a policy $\pi$ starting in state $s$ of $M_\mu$ is
		\[
			V^\pi_{M_\mu}(s) = \IE_{\pi}[-c\,N + \max(\mu,m(S_N)) \given S_0=s]
		\]
		and the utility of stopping in this state $\max(\mu,m(s_0))$.
		We wish to show that the set of $\mu$ such that
		\[
			\max_\pi \IE_{\pi}[-c\,N + \max(\mu,m(S_N)) - \max(\mu,m(S_0)) \given S_0=s] \le 0
		\]
		forms an interval.

		Observe that for any random variable $X$, 
			$\IE[\max(\mu,X)]$ is monotonically increasing in $\mu$ with subderivative between zero and one.
		As a result, for any $v_1$
			$\IE[\max(\mu,X)] - \max(\mu,v_1)$ 
			is monotonically increasing for $\mu<v$, 
			and monotonically decreasing thereafter.
		Therefore, the set if $\mu$ such that this expression is at most $v_2$ forms an interval, containing $v_1$ if non-empty.

		Applying this with $v_1 = m(s_0)$ and $v_2=\IE_{\pi}[c\,N]$, and observing that the union of intervals containing
		a point is an interval containing that point, gives the result.
		\end{proof}	
	\end{hiddenproof}

	\begin{dfn}\label{dfn:independent-actions}
		A metalevel probability model $\mathcal{M}=(U_1,\dots,U_k,\Evidence)$ 
		has \term{independent actions} if the computational variables can be partitioned 
		$\Evidence = \Evidence_1\cup\dots\cup\Evidence_k$ such that
		the sets $\{U_i\}\cup\Evidence_i$ are independent of each other for different $i$.	
	\end{dfn}

	\begin{dfn}\label{dfn:blinkered}
		Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$ with independent actions,
		the \term{blinkered policy} $\pi^b$ is defined by $\pi^b(s) = \argmax_{a\in A_s} Q^b(s,a)$ where
		$Q^b(s,\bot) = \bot$ and
		\begin{equation}\label{eq:blinkered}
			Q^b(s,E_i) = \sup_{\pi\in\Pi^b_i} Q^\pi(s,E_i)
		\end{equation}
		for $E_i\in\Evidence_i$, where $\Pi^b_i$ is the set of policies $\pi$ where $\pi(s)\in\Evidence_i$ for all $s\in S$.
	\end{dfn}

	\begin{dfn}\label{dfn:one-action}
		Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$ with independent actions,
		a \term{one-action metalevel decision problem} for $i=1,\dots,k$ is the metalevel decision
		problem $M^1_{i,m} = (S_i,s_0,A_{s0},T_i,R_i)$ defined by the metalevel probability
		model $(U_0,U_i,\Evidence_i)$ with $U_0=m$.
	\end{dfn}

	\begin{thm}\label{thm:blinkered}
	Given a metalevel decision problem $M=(S,s_0,A_s,T,R)$ with independent actions,
	let $M^1_{i,\lambda_i}$ be the $i$th one-action metalevel decision problem for $i=1,\dots,k$.
	Then for any $s\in S$, whenever $E_i\in A_s\cap\Evidence_i$ we have:
	\[
		Q^b_M(s,E_i) = Q^*_{M^1_{i,\mu^*_{-i}}}(s_i, E_i)
	\]
	where $\mu^*_{-i} = \max_{j\neq i} \mu_j(s)$.
	\end{thm}

	\begin{hiddenproof}
		\begin{proof}
		Fix a state $s$, a $E_i\in A_s$ and take any $\pi\in\Pi^b_i$.  Note that such policies
		are equivalent to polices $\pi'$ on $M^1_{1,m}$, and all such policies are represented.
		Consider $Q^\pi(s,E_i)$.  As $\pi(s)\in\Evidence_i$ for all $s\in S$, by action independence $\mu_j(S_n) = \mu_j(s)$.
		By this and \thmref{thm:value-of-computation}, then,
		\[
			Q^\pi_M(s,E_i) = \IE^\pi_M[ -c\,N + \max(\mu_i(S_N), m_i) \given S_0=s, A_0=E_i].
		\]
		Noting that $\mu_i(S_N)$ is a function only of $(S_N)_i$, and that since 
		But then this is exactly $Q^*_{M^1_{i,\mu^*_{-i}}}(s_i, E_i)$.  Taking the supremum
		over $\pi$ gives the result.
		\end{proof}	
	\end{hiddenproof}

\begin{thm} $\Lambda_i^b$ is bounded from above as
\begin{align}
\label{eqn:thm-be}
  \Lambda_\alpha^b&\le \frac {N \overline X_\beta^{n_\beta}} {n_\alpha} \Pr(\overline X_\alpha^{n_\alpha+N}\le\overline X_\beta^{n_\beta})\nonumber\\
\Lambda_{i|i\ne\alpha}^b&\le \frac {N(1-\overline X_\alpha^{n_\alpha})} {n_i}\Pr(\overline   X_i^{{n_i}+N}\ge\overline X_\alpha^{n_\alpha})
\end{align}
\label{thm:be}
\end{thm}
	\vspace{-2em}
	\begin{proof} For the case $i\ne \alpha$, the probability that the
	  $i$th arm is finally chosen instead of $\alpha$ is
	  $\Pr(\overline X_i^{n_i+N} \ge \overline X_\alpha^{n_\alpha})$. $X_i \le 1$,
	  therefore $\overline X_i^{n_i+N}\le \overline
	  X_\alpha^{n_\alpha}+\frac {N(1-\overline X_\alpha^{n_\alpha})} {N+n_i}$. Hence, the intrinsic value of blinkered
	  information is at most: 
	\begin{align}
	\label{eq:simplistic}
	\frac{ N(1-\overline  X_\alpha^{n_\alpha})}
	  {N+n_i}&\Pr(\overline X_i^{{n_i}+N}\hspace{-0.5em}\ge\overline X_\alpha^{n_\alpha})\nonumber \\
	&\le\frac{ N(1-\overline  X_\alpha^{n_\alpha})}
	{n_i}\Pr(\overline X_i^{{n_i}+N}\hspace{-0.5em}\ge\overline X_\alpha^{n_\alpha})
	\end{align}
	  Proof for the case $i=\alpha$ is similar.
	\end{proof}

\begin{thm} The probabilities in \eqref{eqn:thm-be} are bounded from above as
\begin{align}
  \label{eqn:probound-blnk-hoeffding}
  \Pr&(\overline X_\alpha^{{n_\alpha}+N} \le \overline X_\beta^{n_\beta})
  \le 2\exp\left(- \varphi (\overline X_\alpha^{n_\alpha} - \overline X_\beta^{n_\beta})^2 n_\alpha
  \right)\nonumber\\
  \Pr&(\overline X_{i|i\ne\alpha}^{n_\alpha+N} \ge \overline X_\beta^{n_\beta})
  \le 2\exp\left(- \varphi (\overline X_\alpha^{n_\alpha} -\overline  X_i^{n_i})^2 n_i \right)
\end{align}
where $\varphi=\min \left(2(\frac {1+n/N} {1+\sqrt {n/N}})^2\right)=8(\sqrt 2 - 1)^2 > 1.37$.
\label{thm:hoeffding-prob-bounds}
\end{thm}

\begin{hiddenproof}
	\begin{proof}
	\eqref{eqn:probound-blnk-hoeffding}) follow from the
	observation that if $i\ne\alpha$, $\overline X_i^{n_i+N}>\overline X_\alpha^{n_i}$
	if and only if the mean $\overline X_i^N$ of $N$ samples from $n_i+1$
	to $n_i+N$ is at least $\overline X_\alpha^{n_i}+(\overline X_\alpha^{n_i}-\overline
	X_i^{n_i})\frac {n_i} N$.

	For any $\delta$, the probability that $\overline X_i^{n_i+N}$ is greater
	than $\overline X_\alpha^{n_i}$ is less than the probability that
	$\IE[X_i]\ge\overline X_i^n+\delta$
	\emph{or} $\overline X_i^N\ge \IE[X_i]+\overline X_\alpha^{n_\alpha}
	- \overline X_i^{n_i} - \delta +(\overline X_\alpha^{n_\alpha} - \overline X_i^{n_i})\frac {n_i} N$,
	thus, by the union bound, less than the sum of the probabilities:
	\begin{align}
	\Pr&(\overline X_i^{n_i+N}\ge \overline X_\alpha^{n_i})\nonumber\\
	   &\le\Pr(\IE[X_i]-\overline X_i^{n_i} \ge \delta)\\
	   &\hspace{0.25em}+\Pr\left(\overline X_i^N\hspace{-0.5em} - \IE[X_i] \ge \overline X_\alpha^{n_\alpha}\hspace{-0.5em}
	           - \overline X_i^{n_i}\hspace{-0.5em} - \delta +(\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_i^{n_i})\frac {n_i} N\right)\nonumber
	\end{align}
	Bounding the probabilities on the right-hand side using the Hoeffding
	inequality, obtain:
	\begin{align}
	\Pr&(\overline X_i^{n_i+N}\ge \overline X_\alpha^{n_\alpha})\le \nonumber\\
	    &\exp(-2\delta^2n_i)+\nonumber\\
	    &\quad\exp\left(-2\left((\overline X_\alpha^{n_\alpha}\hspace{-0.5em}
	         - \overline X_i^{n_i})\left(1+\frac {n_i} N\right)
	         - \delta\right)^2N\right)
	\label{eqn:app-hoeffding-le-maxexp}
	\end{align}
	Find $\delta$ for which the two terms on the right-hand side of
	\eqref{eqn:app-hoeffding-le-maxexp} are equal:
	\begin{equation}
	\exp(-\delta^2n) = \exp\left(-2\left((\overline X_\alpha - \overline X_i)(1+\frac {n_i} N) - \delta\right)^2N\right)\label{eqn:app-hoeffding-eq-exp}
	\end{equation}
	Solve \eqref{eqn:app-hoeffding-eq-exp} for $\delta$: $\delta=\frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}} (\overline X_\alpha^{n_\alpha}\hspace{-0.5em}
	- \overline X_i^{n_i}) \ge 2(\sqrt 2 - 1)(\overline X_\alpha^{n_\alpha}\hspace{-0.5em}-\overline X_i^{n_i})$. Substitute $\delta$ into 
	\eqref{eqn:app-hoeffding-le-maxexp} and obtain
	\begin{align}
	\Pr&(\overline X_i^{n_i}\ge \overline X_\alpha^{n_\alpha}) \nonumber\\
	& \le 2\exp\left(-2\left( \frac {1+\frac {n_i} N} {1+\sqrt {\frac {n_i} N}}
	                          (\overline X_\alpha^{n_\alpha} - \overline X_i^{n_i})\right)^2 n_i\right)\nonumber \\
	& \le 2\exp(-8(\sqrt 2 - 1)^2(\overline X_\alpha^{n_\alpha} - \overline X_i^{n_i})^2n_i)\nonumber\\
	& = 2\exp(-\varphi(\overline X_\alpha^{n_\alpha} - \overline X_i^{n_i})^2n_i)
	\end{align}
	Derivation for the case $i=\alpha$ is similar.
	\end{proof}	
\end{hiddenproof}

\begin{crl}
An upper bound on the VOI estimate $\Lambda_i^b$ is obtained
by substituting \eqref{eqn:probound-blnk-hoeffding} into \eqref{eqn:thm-be}.
\begin{align}
  \label{eqn:bound-blnk-hoeffding}
  \Lambda&_\alpha^b \le \hat\Lambda_\alpha^b=\frac {2N\overline X_\beta^{n_\beta}} {n_\alpha}\exp\left(- \varphi(\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_\beta^{n_\beta})^2 n_\alpha\right)\nonumber\\
  \Lambda&_{i|i\ne\alpha}^b\le \hat\Lambda_i^b=  \frac {2N(1-\overline  X_\alpha^{n_\alpha})} {n_i}\exp\left(- \varphi(\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_i^{n_i})^2 n_i\right)
\end{align}
\label{crl:bound-blnk-hoeffding}
\end{crl}
\vspace{-2em}

\end{document}

### 2012-06-26 paper/vago.uai/paper.tex
as send by nick
with our changes

@@
 \title{Selecting Computations: Theory and Applications} 
-
-%\author{}
 \author{ {\bf Nicholas Hay} and {\bf Stuart Russell} \\   
@@
 \And 
-{\bf Solomon Eyal Shimony} and {\bf David Tolpin}  \\ 
+{\bf David Tolpin} and {\bf Solomon Eyal Shimony} \\ 
 Department of Computer Science \\ 
 Ben-Gurion University of the Negev\\
-Beer Sheva, Israel
+Beer Sheva 84105, Israel
 }
@@
 
-We thus propose the following hybrid sampling scheme: 
+We thus propose the following hybrid sampling scheme \citep{TolpinShimony.mcts}: 
 	at the \emph{root node}, sample based on the VOI estimate;
@@
 \end{enumerate}
-In this approach, the cost of a sample in the current state is the
+In this approach, the cost $c$ of a sample in the current state is the
 VOI of increasing the budget of a future state by one sample.  It is
@@
   \begin{itemize}
-    \item The same number of samples is available to
+    \item Initially, the same number of samples is available to
       the agent at each step, independently of the number of pre-simulated
@@
 
+\input{acknowledgements}
+
 \bibliographystyle{abbrvnat}

### 2012-06-27 paper/vago.uai/abstract.tex
9 pages

@@
 game of Go.  In this paper we develop a theoretical basis for
-metalevel decisions in the statistical framework of {\em selection problems}, 
+metalevel decisions in the statistical framework of Bayesian {\em selection problems}, 
 arguing (as others have done) that this is more appropriate than the bandit

### 2012-06-27 paper/vago.uai/introduction.tex
9 pages

@@
 Most work has focused on the probability of selection error rather than
-optimal policies~\citep{Bubeck+al:2011}.
+optimal policies in the Bayesian setting~\citep{Bubeck+al:2011}.
 Accordingly, we present in \secrefs{sec:optimal}{sec:context}
@@
 
-(Proofs omitted due to space constraints can be found in the supplementary material.)
+%% (Proofs omitted due to space constraints can be found in the supplementary material.)

### 2012-06-27 paper/vago.uai/optimalpolicies.tex
9 pages

@@
 For simplicity, in the below we'll assume the utilities $U_i$ are bounded, 
-without loss of generality in $[0,1]$.
+without loss of generality in $[0,1]$.  We will abuse notation and denote by $e\in E$ that $e$ is
+a potential value of the computation $E$.
 
@@
 each arm will either succeed or not $U_i\in\{0,1\}$, with an unknown latent frequency of success $\Theta_i$, 
-and stochastic simulations of possible consequences $E_{ij}$ can be performed:
+and a set of stochastic simulations of possible consequences
+$\mathcal{E} = \{E_{ij} | 1\le i \le k, j\in \mathbb{N}\}$ that can be performed:
 \begin{align*}
@@
 to deciding which arm to use, stopping when the cost of computation 
-outweighs the benefit gained. We formalize the selection problem as a Markov Decision Process (MDP; 
-see, e.g., \citet{Puterman:1994}):
+outweighs the benefit gained. We formalize the selection problem as a Markov Decision Process
+(see, e.g., \citet{Puterman:1994}):
 
@@
 \end{dfn}
+
+Note that when stopping in state $s$, the expected utility of 
+action $i$ is by definition $\mu_i(s)$, so the optimal action to take is $i^*\in\argmax_i \mu_i(s)$
+which has expected utility $\mu_{i*}(s) = \max_i\mu_i(s)$.
 
@@
 
-\begin{proof}
+\begin{hiddenproof}
 	Follows immediately from \eqref{eq:value-fn} and the definition of the
 	reward function in \dfnref{dfn:metalevel-mdp}.
-\end{proof}
+\end{hiddenproof}
 
@@
 infinitely many computations be performed?
+
+%%This immediately raises the question of whether
+%%one can solve for an optimal policy, and whether it is sometimes optima
+%%to perform infinitely many computations.
 
@@
 
-\begin{proof}
+\begin{hiddenproof}
 	The first follows as in state $s$ the optimal policy has value at least that
@@
 	The second from \thmref{thm:value-of-computation}.
-\end{proof}
+\end{hiddenproof}
 
@@
 
-The myopic policy takes the best action, to either stop or perform a computation,
+The myopic policy (known the metalevel greedy approximation with single-step
+assumption in \cite{Russell+Wefald:1991a}) takes the best action, to either stop or perform a computation,
 under the assumption that at most one further computation can be performed.
@@
 
-Using this, we can prove our bound:
+Using these results connecting the behavior of the optimal and myopic policies, we can prove our bound:
 
@@
 Fixing $n$ and maximizing over $\mu$, we get sufficient condition for stopping
+Since the set of states satisfying \eqref{eq:myopic-stopping} is closed under
 \begin{equation}
@@
 computation is relatively high, such as in simulation experiments \citep{Swisher+et+al:2003},
-or when the decision to be made is critical. Finite bounds for the $k$-armed problem
-can also be derived and will be included in the full paper.
+or when the decision to be made is critical.
+
+%% Finite bounds for the $k$-armed problem
+%% can also be derived and will be included in the full paper.

### 2012-06-27 paper/vago.uai/paper.tex
9 pages

@@
 University of California\\ 
-Berkeley, CA 94720 \\ 
+Berkeley, CA 94720\\
+\texttt{\{nickjhay,russell\}@cs.berkeley.edu}
 \And 
@@
 Ben-Gurion University of the Negev\\
-Beer Sheva 84105, Israel
+Beer Sheva 84105, Israel\\
+\texttt{\{tolpin,shimony\}@cs.bgu.ac.il}
 }
@@
 \end{align}
+\vspace{-24pt}
 \label{thm:be}
@@
   \label{eqn:probound-blnk-hoeffding}
-  \Pr&(\overline X_\alpha^{{n_\alpha}+N} \le \overline X_\beta^{n_\beta})
+  \Pr&(\overline X_\alpha^{{n_\alpha}+N}\hspace{-6pt} \le \overline X_\beta^{n_\beta})
   \le 2\exp\left(- \varphi (\overline X_\alpha^{n_\alpha} - \overline X_\beta^{n_\beta})^2 n_\alpha
   \right)\nonumber\\
-  \Pr&(\overline X_{i|i\ne\alpha}^{n_\alpha+N} \ge \overline X_\beta^{n_\beta})
-  \le 2\exp\left(- \varphi (\overline X_\alpha^{n_\alpha} -\overline  X_i^{n_i})^2 n_i \right)
+  \Pr&(\overline X_{i|i\ne\alpha}^{n_\alpha+N}\hspace{-6pt} \ge \overline X_\beta^{n_\beta})
+  \le 2\exp\left(- \varphi (\overline X_\alpha^{n_\alpha} -\overline  X_i^{n_i})^2 n_i \right)\hspace{-6pt}
 \end{align}
@@
 An upper bound on the VOI estimate $\Lambda_i^b$ is obtained
-by substituting \eqref{eqn:probound-blnk-hoeffding} into \eqref{eqn:thm-be}.
+by substituting \eqref{eqn:probound-blnk-hoeffding} into (\ref{eqn:thm-be}).
 \begin{align}
   \label{eqn:bound-blnk-hoeffding}
-  \Lambda&_\alpha^b \le \hat\Lambda_\alpha^b=\frac {2N\overline X_\beta^{n_\beta}} {n_\alpha}\exp\left(- \varphi(\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_\beta^{n_\beta})^2 n_\alpha\right)\nonumber\\
-  \Lambda&_{i|i\ne\alpha}^b\le \hat\Lambda_i^b=  \frac {2N(1-\overline  X_\alpha^{n_\alpha})} {n_i}\exp\left(- \varphi(\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_i^{n_i})^2 n_i\right)
+  \Lambda&_\alpha^b \le \hat\Lambda_\alpha^b=\frac {2N\overline X_\beta^{n_\beta}} {n_\alpha}\exp\left(- \varphi(\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_\beta^{n_\beta})^2 n_\alpha\right)\\
+  \Lambda&_{i|i\ne\alpha}^b\le \hat\Lambda_i^b=  \frac {2N(1-\overline  X_\alpha^{n_\alpha})} {n_i}\exp\left(- \varphi(\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_i^{n_i})^2 n_i\right)\nonumber
 \end{align}
@@
 \centering
-\includegraphics[scale=0.6]{flat.pdf}
+\includegraphics[scale=0.55]{flat.pdf}
 \caption{Average regret of various policies as a function of the fixed number 
@@
 be used to stop the sampling if the intrinsic VOI of any arm
-is less than the total cost of sampling $C$:
-\begin{equation}
-\mbox{stop if } \max_i \Lambda_i \le C
-\end{equation}
+is less than the total cost of sampling $C$: $\max_i \Lambda_i \le C$.
+
 The VOI estimates of \eqrefs{eqn:thm-be}{eqn:bound-blnk-hoeffding} 
@@
 \centering
-\includegraphics[scale=0.6]{uctvoi.pdf}
+\includegraphics[scale=0.55]{uctvoi.pdf}
 \caption{Winning rate of the VOI-aware policy in Go as a function of the cost $c$, for varying numbers of samples per ply.}
@@
 \centering
-\includegraphics[scale=0.6]{voi-wins.pdf}
+\includegraphics[scale=0.55]{voi-wins.pdf}
 \caption{Winning rate of the VOI-aware policy in Go as a function of the number of samples, fixing cost $c=10^{-6}$.}

### 2012-08-12 paper/vago.ecai/data/figures.R
vct wins with error bars

@@ -25,12 +25,15 @@ draw.winrate <- function(regrets_file, log="xy", legend.position='topright', xla
         regrets <- as.matrix(regrets[-1,])
         regrets <- matrix(as.numeric(regrets), nrow=nrow(regrets), ncol=ncol(regrets))
 
-    xlim <- range(regrets[,1])
-    ylim <- range(max(0, min(regrets[,-1])*0.95), min(100, max(regrets[,-1])*1.05))
-
-	plot(x=regrets[,1], y=regrets[,2], type='o', log=log, xlim=xlim, ylim=ylim, yaxs='i', xlab=xlab, ylab='VOI wins, %', main=if(is.na(main)) unlist(strsplit(regrets_file, "\\."))[[1]] else main, axes=F)
+        x <- regrets[,1]
+        y <- regrets[,2]
+        xlim <- range(x)
+        ylim <- range(max(0, min(y)*0.95), min(100, max(y)*1.05))
+        err <- sqrt(y*(100.0-y)/1000.0)  # sigma for 1000 trials
+	plot(x=x, y=y, type='o', log=log, xlim=xlim, ylim=ylim, yaxs='i', xlab=xlab, ylab='VOI wins, %', main=if(is.na(main)) unlist(strsplit(regrets_file, "\\."))[[1]] else main, axes=F)
+        arrows(x,y+err, x, y-err, angle=90, code=3, length=0.025)
 	box()
-	axis(1, at=regrets[,1])
+	axis(1, at=x)
 	axis(2)
 	abline(h=50.0, lty='dashed');
 }

### 2012-08-13 paper/vago.uai/slides/slides.tex
slides for UAI in progress

\documentclass{beamer}
\usepackage[latin1]{inputenc}

\usepackage[noend]{algpseudocode}
\usepackage[ruled]{algorithm}
\usepackage{url}
\usepackage{framed}
\usepackage{amsfonts,amsmath,amsthm}
\usepackage{graphicx}
\usepackage{url}
\usepackage{color}

\title{}
\author{}
\institute{}
\date{}

\AtBeginSection[]
{
  \begin{frame}<beamer>
    \frametitle{Outline}
    \tableofcontents[currentsection]
  \end{frame}
}

\begin{document}

% \begin{frame}
% \titlepage
% \end{frame}

\begin{frame}{Upper Bounds on Value of Information}
\end{frame}

\begin{frame}{Sampling in Trees}
\begin{itemize}
\item Hybrid sampling scheme:
\begin{enumerate}
\item At the {\it root node}: sample based on the VOI estimate.
\item At {\it non-root nodes}: sample using UCT.
\end{enumerate}
\item Stopping Criterion
\end{itemize}
\end{frame}

\begin{frame}{Sample Redistribution}
\begin{itemize}
\item<+-> The VOI estimate assumes that the information is {\bf discarded}
  between states.
\item<+-> MCTS {\bf re-uses rollouts} generated at earlier search states.
\vspace{\baselineskip}
\item<+-> Either incorporate `future' influence into the VOI estimate
  ({\it non-trivial!}).
\item<+-> Or behave myopically w.r.t. search tree depth:
\begin{enumerate}
\item Estimate VOI as though the information is discarded.
\item Stop early if the VOI is below a certain threshold.
\item Save the unused sample budget for search in future states.
\end{enumerate}
\item<+-> The cost $c$ of a sample  is\\\hspace{1em} the VOI of increasing a
  future budget by one sample.
\end{itemize}
\end{frame}

\begin{frame}{Playing Go Against UCT:\\\hspace{2em}Tuning the Sample Cost}
\begin{figure}[h]
  \centering
  \includegraphics[scale=0.65]{uctvoi.pdf}
\end{figure}
Best results for sample cost $c\approx 10^{-6}$:\\\hspace{2em}winning rate of {\bf
  64\%} for 10000 samples per ply.
\end{frame}

\begin{frame}{Playing Go Against UCT:
    \\\hspace{1em} Winning Rate vs. Number of Samples per Ply}

Sample cost $c$ fixed at $10^{-6}$:
\begin{figure}
  \centering
  \includegraphics[scale=0.65]{voi-wins.pdf}
\end{figure}
Best results for {\it intermediate} $N_{samples}$:
\begin{itemize}
\item When $N_{samples}$ is too low, poor moves are selected.
\item When $N_{samples}$ is too high, the VOI of further sampling
  is low.
\end{itemize}
\end{frame}

\end{document}

### 2012-08-13 paper/vago.uai/slides/slides.tex
slides in progress, mostly done

@@
 \begin{frame}{Upper Bounds on Value of Information}
+Assuming that:
+\begin{enumerate}
+\item Samples are i.i.d. given the value of the arm.
+\item The expectation of a selection in a belief state is equal to the sample mean.
+\end{enumerate}
+Upper bounds on intrinsic VOI $\Lambda^b_i$ of testing the $i$th arm N
+times are (based on Hoeffding inequality):
+\begin{align*}
+  \Lambda&_\alpha^b < \frac {2N\overline X_\beta^{n_\beta}} {n_\alpha+1}\exp\left(- 1.37(\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_\beta^{n_\beta})^2 n_\alpha\right)\\
+  \Lambda&_{i|i\ne\alpha}^b <  \frac {2N(1-\overline  X_\alpha^{n_\alpha})} {n_i+1}\exp\left(- 1.37(\overline X_\alpha^{n_\alpha}\hspace{-0.5em} - \overline X_i^{n_i})^2 n_i\right)
+\end{align*}
+Tighter bounds can be obtained (see the paper).
+
 \end{frame}
@@
 \end{enumerate}
-\item Stopping Criterion
+\item Stopping criterion: Assuming sample cost $c$ is known,\\
+\hspace{1em}stop sampling when intrinsic VOI is less than $C=cN$:
+\begin{align*}
+\frac 1 N \Lambda_\alpha^b \le&\frac {\overline X_\beta^{n_\beta}}
+  {n_\alpha+1}\Pr(\overline X_\alpha^{n_\alpha+N}\le\overline
+  X_\beta^{n_\alpha})\le c\\
+\frac 1 N \max_i\Lambda_i^b\le &\max_i\frac {(1-\overline X_\alpha^{n_\alpha})} {n_i+1}\Pr(\overline
+  X_i^{n_i+N}\ge\overline X_\alpha^{n_\alpha})\le c\\
+    &\forall i: i\ne\alpha
+\end{align*}
 \end{itemize}
@@
 \begin{itemize}
-\item<+-> The VOI estimate assumes that the information is {\bf discarded}
+\item The VOI estimate assumes that the information is {\bf discarded}
   between states.
-\item<+-> MCTS {\bf re-uses rollouts} generated at earlier search states.
+\item MCTS {\bf re-uses rollouts} generated at earlier search states.
 \vspace{\baselineskip}
-\item<+-> Either incorporate `future' influence into the VOI estimate
+\item Either incorporate `future' influence into the VOI estimate
   ({\it non-trivial!}).
-\item<+-> Or behave myopically w.r.t. search tree depth:
+\item Or behave myopically w.r.t. search tree depth:
 \begin{enumerate}
@@
 \end{enumerate}
-\item<+-> The cost $c$ of a sample  is\\\hspace{1em} the VOI of increasing a
+\item The cost $c$ of a sample  is\\\hspace{1em} the VOI of increasing a
   future budget by one sample.
